### 2024-05-10 analysis/eda.Rmd
Start OA eda to inform firther data gathering

---
title: "eda"
output: github_document
---

```{r, knitr_setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r r_setup}
library(bigrquery)
library(tidyverse)
#' Connect to DB 
#' 
#' BigQuery HOAD (Crossref, OpenAlex and open friends)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  project = "subugoe-collaborative",
  billing = "subugoe-collaborative"
)
```

## Web of Science

### Open access

```{sql add_oa_info, connection = bq_con, output.var = "wos_oa"}
WITH oa_raw AS (SELECT DISTINCT item_id, doi, issn_l, CASE WHEN online_year IS NULL THEN pubyear ELSE online_year END AS earliest_year, is_oa
FROM (
SELECT
  wos.item_id, 
  wos.doi, 
  wos.issn_l,
  wos.pubyear,
  CAST(REGEXP_EXTRACT(wos_pubdate_online, r'^[0-9]{4}') AS NUMERIC) AS online_year,
  CASE
    WHEN cc IS NOT NULL AND vor = 1 THEN TRUE
  ELSE
  FALSE
END
  AS is_oa
FROM
  `hoa-article.hoa_comparision.wos_matching_20240410` AS wos
LEFT JOIN
  `subugoe-collaborative.hoaddata.cc_md` AS cc
ON
  wos.doi = cc.doi ) )
SELECT COUNT(DISTINCT item_id) as wos_articles, issn_l, earliest_year, is_oa
FROM oa_raw
GROUP BY issn_l, earliest_year, is_oa
ORDER BY issn_l, earliest_year
```
By year

```{r}
wos_all <- wos_oa |>
  group_by(earliest_year) |>
  summarise(articles = sum(wos_articles))

wos_oa |>
  filter(is_oa == TRUE) |>
  group_by(earliest_year) |>
  summarise(oa = sum(wos_articles)) |>
  right_join(wos_all) |>
  mutate(prop = oa / articles)
```

### 2024-05-14 data-raw/001_data_gathering.R
Add more info to wos items

@@ -89,7 +89,7 @@ where
 #' backup
 write_csv(ror_wos_matching, here::here("data-raw", "ror_wos_matching.csv"))
 
-#### Prepare journal data
+#' ## Prepare journal data
 jct_issn <- hoaddata::jct_hybrid_jns |>
   distinct(issn, issn_l)
 dbWriteTable(kb_con, "jct_hybrid_jns_issn", jct_issn, overwrite = TRUE)
@@ -97,13 +97,26 @@ dbWriteTable(kb_con, "jct_hybrid_jns_issn", jct_issn, overwrite = TRUE)
 #### Retrieve articles from hybrid journals
 dbExecute(kb_con, "DROP TABLE wos_jct_items")
 
-dbExecute(kb_con, "CREATE table wos_jct_items AS select
+dbExecute(kb_con, "CREATE table wos_jct_items AS 
+select distinct 
+    issn_l,
+	item_id,
+	doi,
+	pubyear,
+	wos_pubdate_online,
+	oa_status,
+	item_type
+from (
+select
 	distinct 
-	jhji.issn_l ,
+	jhji.issn_l,
 	i.item_id,
 	doi,
 	pubyear,
-	wos_pubdate_online
+	wos_pubdate_online,
+	unnest(wos_ci) as ci,
+	oa_status,
+	item_type
 from
 	jct_hybrid_jns_issn jhji
 left join wos_b_202310.v_issn_isbn vii on
@@ -111,7 +124,16 @@ left join wos_b_202310.v_issn_isbn vii on
 left join wos_b_202310.v_items i on
 	vii.item_id = i.item_id
 where
-	pubyear > 2017")
+	pubyear > 2017 ) as tmp
+where ci in ('SCI', 'SSCI', 'AHCI')")
+
+wos_jct_items <- DBI::dbReadTable(kb_con, "wos_jct_items")
+#' backup
+wos_jct_items_df <- wos_jct_items |> 
+  as_tibble() |> 
+  mutate(item_type = as.character(gsub('\\{|\\}|"', '', item_type))) |>
+  mutate(oa_status = as.character(gsub('\\{|\\}|"', '', oa_status)))
+write_csv(wos_jct_items_df, "data-raw/wos_jct_items_df.csv")
 
 #### Add WOS affiliation data for first data and corresponding authors
 dbExecute(kb_con, "DROP TABLE wos_jct_affiliations")

### 2024-05-14 analysis/eda.Rmd
OA analysis Jan 2024

@@
 ```
+
+## Compare to HOAD
+
+```{r}
+wos_jn_all <- wos_oa |>
+  group_by(earliest_year) |>
+  summarise(journals = n_distinct(issn_l),
+            articles = sum(wos_articles))
+
+wos_oa_per_year <-  wos_oa |>
+  filter(is_oa == TRUE) |>
+  group_by(earliest_year) |>
+  summarise(oa_journals = n_distinct(issn_l),
+            oa_articles = sum(wos_articles))
+
+wos_per_year <- inner_join(wos_jn_all, wos_oa_per_year)
+
+cr_all_per_year <- hoaddata::jn_ind |> 
+  distinct(issn_l, cr_year, jn_all) |>
+  group_by(cr_year) |>
+  summarise(cr_journals = n_distinct(issn_l),
+            cr_articles = sum(jn_all))
+
+cr_oa <- hoaddata::jn_ind |>
+  filter(!is.na(cc)) |> group_by(cr_year) |>
+  summarise(cr_oa = sum(cc_total))
+
+cr_per_year <- inner_join(cr_all_per_year, cr_oa) |>
+  mutate(cr_year = as.numeric(as.character(cr_year)))
+
+my_df_year <- inner_join(cr_per_year, wos_per_year, by = c("cr_year" = "earliest_year")) |>
+  mutate(cr_oa_prop = cr_oa / cr_articles,
+         wos_oa_prop = oa_articles / articles)
+
+### journal level
+
+wos_jn_all <- wos_oa |>
+  group_by(earliest_year, issn_l) |>
+  summarise(articles = sum(wos_articles))
+
+wos_oa_per_year <-  wos_oa |>
+  filter(is_oa == TRUE) |>
+  group_by(earliest_year, issn_l) |>
+  summarise(oa_articles = sum(wos_articles))
+
+wos_per_year <- inner_join(wos_jn_all, wos_oa_per_year)
+
+cr_all_per_year <- hoaddata::jn_ind |> 
+  distinct(issn_l, cr_year, jn_all) |>
+  group_by(cr_year, issn_l) |>
+  summarise(cr_articles = sum(jn_all))
+
+cr_oa <- hoaddata::jn_ind |>
+  filter(!is.na(cc)) |> group_by(cr_year, issn_l) |>
+  summarise(cr_oa = sum(cc_total))
+
+cr_per_year <- inner_join(cr_all_per_year, cr_oa) |>
+  mutate(cr_year = as.numeric(as.character(cr_year)))
+
+my_df <- inner_join(cr_per_year, wos_per_year, by = c("cr_year" = "earliest_year", "issn_l")) |>
+  mutate(cr_oa_prop = cr_oa / cr_articles,
+         wos_oa_prop = oa_articles / articles)
+```
+
+### wos_jct_items_approach
+
+use items from kb, distinguishing between doc types and provides oa information
+
+```{r}
+wos_jct_items_raw <- readr::read_csv(here::here("data-raw/", "wos_jct_items_df.csv")) 
+wos_jct_items <- wos_jct_items_raw|>
+  mutate(core_pub = ifelse(grepl("Article|Review", item_type), 1, 0),
+         hybrid_oa = ifelse(grepl("hybrid", oa_status), 1, 0),
+         cr_oa = ifelse(tolower(doi) %in% hoaddata::cc_articles$doi, 1, 0),
+         online_year = lubridate::year(wos_pubdate_online)) |>
+  mutate(earlies_year = ifelse(!is.na(online_year), online_year, pubyear)) |>
+  distinct()
+```
+
+## stats
+
+```{r}
+core_hybrid_oa <- wos_jct_items |>
+  filter(core_pub == 1) |>
+  group_by(earlies_year) |>
+  summarise(core_hybrid_oa = sum(hybrid_oa),
+            core_cr_oa = sum(cr_oa))
+
+wos_oct_23_stat <- wos_jct_items |>
+  group_by(earlies_year) |>
+  summarise(all = n_distinct(item_id),
+            core = sum(core_pub),
+            hybrid_oa = sum(hybrid_oa),
+            cr_oa = sum(cr_oa)) |>
+  inner_join(core_hybrid_oa) |>
+  mutate(oa_prop = hybrid_oa / all,
+         core_oa_prop = core_hybrid_oa / core,
+         cr_prop = cr_oa / all,
+         core_cr_prop = core_cr_oa / core)
+
+write_csv(wos_oct_23_stat, here::here("data", "wos_oct_23_stat.csv"))
+```

### 2024-05-14 data-raw/001_data_gathering.R

@@ -119,9 +119,9 @@ select
 	item_type
 from
 	jct_hybrid_jns_issn jhji
-left join wos_b_202310.v_issn_isbn vii on
+left join wos_b_202401.v_issn_isbn vii on
 	jhji.issn = vii.sn
-left join wos_b_202310.v_items i on
+left join wos_b_202401.v_items i on
 	vii.item_id = i.item_id
 where
 	pubyear > 2017 ) as tmp

### 2024-05-15 analysis/eda.Rmd
Next, map it by esac publisher

@@
 
-wos_oct_23_stat <- wos_jct_items |>
+wos_24_stat <- wos_jct_items |>
   group_by(earlies_year) |>
@@
 
-write_csv(wos_oct_23_stat, here::here("data", "wos_oct_23_stat.csv"))
+#write_csv(wos_oct_23_stat, here::here("data", "wos_oct_23_stat.csv"))
 ```
+
+## by journal
+
+```{r}
+
+## Wos data
+
+jn_core_hybrid_oa <- wos_jct_items |>
+  filter(core_pub == 1) |>
+  group_by(earlies_year, issn_l) |>
+  summarise(core_hybrid_oa = sum(hybrid_oa),
+            core_cr_oa = sum(cr_oa))
+
+jn_wos_24_stat <- wos_jct_items |>
+  group_by(earlies_year, issn_l) |>
+  summarise(all = n_distinct(item_id),
+            core = sum(core_pub),
+            hybrid_oa = sum(hybrid_oa),
+            cr_oa = sum(cr_oa)) |>
+  inner_join(jn_core_hybrid_oa) |>
+  mutate(oa_prop = hybrid_oa / all,
+         core_oa_prop = core_hybrid_oa / core,
+         cr_prop = cr_oa / all,
+         core_cr_prop = core_cr_oa / core)
+
+## hoaddata
+
+cr_all_per_year <- hoaddata::jn_ind |> 
+  distinct(issn_l, cr_year, jn_all) |>
+  group_by(cr_year, issn_l) |>
+  summarise(hoad_articles = sum(jn_all))
+
+cr_oa <- hoaddata::jn_ind |>
+  filter(!is.na(cc)) |> 
+  group_by(cr_year, issn_l) |>
+  summarise(hoad_oa = sum(cc_total))
+
+cr_per_year <- inner_join(cr_all_per_year, cr_oa) |>
+  mutate(cr_year = as.numeric(as.character(cr_year)))
+
+jn_wos_stats <- inner_join(jn_wos_24_stat, cr_per_year, by = c("earlies_year" = "cr_year", "issn_l"))
+```
+
+### Exploration
+
+```{r}
+ggplot(jn_wos_stats, aes(hoad_oa / hoad_articles, core_cr_prop)) +
+  geom_point(alpha = 0.05) +
+  coord_equal() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~ earlies_year)
+```

### 2024-07-18 data-raw/001_data_gathering.R
Improve institution matching by a) using the vendor org id instead of organisation string and b) use only the most frequent vendor_org_id / ror combination. Case 10.1002/anie.202009335

@@ -38,7 +38,7 @@ kb_con <- dbConnect(RPostgres::Postgres(),
 #' combination is used.
 #' 
 #' Article-level publication metadata from institutions with TA
-ta_oa_inst <- readr::read_csv("https://github.com/subugoe/hoaddata/releases/download/v0.2.95/ta_oa_inst.csv.gz")
+ta_oa_inst <- readr::read_csv("https://github.com/subugoe/hoaddata/releases/download/v0.2.98/ta_oa_inst.csv.gz")
 ta_ror_pubs <- ta_oa_inst |>
   distinct(doi, ror_matching)
 #' Make sure RORs are aligned 
@@ -51,23 +51,28 @@ dbExecute(kb_con, "DROP TABLE ta_wos")
 dbExecute(kb_con, "CREATE table ta_wos AS 
 select ta.*, vi.item_id
 from ta_ror_pubs ta
-left join wos_b_202310.v_items vi on ta.doi = LOWER(vi.doi)")
+left join wos_b_202404.v_items vi on ta.doi = LOWER(vi.doi)")
 #' 
 #' Get article-level affiliation metadata
 dbExecute(kb_con, "DROP TABLE ta_wos_aff")
 dbExecute(kb_con, "CREATE TABLE ta_wos_aff AS select distinct tw.*, aa.author_seq_nr, aa.organization, aa.vendor_org_id 
 from ta_wos tw 
-inner join wos_b_202310.v_authors_affiliations aa on tw.item_id = aa.item_id 
+inner join wos_b_202404.v_authors_affiliations aa on tw.item_id = aa.item_id 
 where aa.author_seq_nr = 1 ")
 #' Create matching table where the most frequent ror / org combination is chosen to
 #' account for multiple authorships
+#' 
+#' I use the Unique ID of an organization as defined by Clarivate. WoS Organization Enhanced ("preferred name"). 
+#' Path: /REC/static_data/fullrecord_metadata/addresses/address_name/address_spec/organizations/organization@pref.
+#' 
+#' 
 ror_wos_matching <- dbGetQuery(kb_con, "select
 	*
 from
 	(
 	select
 		*,
-		row_number() over(partition by ror_matching
+		row_number() over(partition by vendor_org_id
 	order by
 		matches desc) as row_n
 	from
@@ -75,12 +80,12 @@ from
 		select
[…]

### 2024-07-18 data-raw/002_wos_ta_matching.R

@@ -10,17 +10,17 @@
 #' 
 #' To Do: Keep this data point in KB.
 library(tidyverse)
-wos_df <- readr::read_csv("data-raw/wos_jct_affiliations_20240410.csv") 
+wos_df <- readr::read_csv("data-raw/wos_jct_affiliations_20240717.csv") 
 # Matching
 wos_ror <- readr::read_csv("data-raw/ror_wos_matching.csv") |>
-  distinct(ror_matching, organization)
+  distinct(ror_matching, vendor_org_id)
 
 wos_matching <- wos_df |> 
-  left_join(wos_ror, by = "organization")
+  left_join(wos_ror, by = "vendor_org_id")
 #' The results is downloaded as csv and uploaded to Google BiGQuery using gcloud 
 #' command line tools (because file is large)
-#' `gcloud storage cp ~/Documents/thesis/hoa_validation/data-raw/wos_matching_20240410.csv gs://bigschol`
-write_csv(wos_matching, "data-raw/wos_matching_20240410.csv")
+#' `gcloud storage cp ~/Documents/thesis/hoa_validation/data-raw/wos_matching_20240717.csv gs://bigschol`
+write_csv(wos_matching, "data-raw/wos_matching_20240717.csv")
 #'
 #' 
 #' ## Matching
@@ -30,7 +30,7 @@ write_csv(wos_matching, "data-raw/wos_matching_20240410.csv")
 #' enriched. The resulting temporary table is `esac_data`
 #' 
 #' Then, we create a subset of publications from eligible institutions by merging
-#' `esac_data` with `wos_matching_20240410`. To improve the performance of the query
+#' `esac_data` with `wos_matching_20240717`. To improve the performance of the query
 #' not all fields will be returned. Also, records from institutions without TA were 
 #' excluded before the matching as indicated by NA. 
 library(bigrquery)
@@ -63,7 +63,7 @@ my_sql <- "WITH
     *,
     CAST(REGEXP_EXTRACT(wos_pubdate_online, r'^[0-9]{4}') AS NUMERIC) AS online_year
   FROM
-    `hoa-article.hoa_comparision.wos_matching_20240410`
+    `hoa-article.hoa_comparision.wos_matching_20240717`
   WHERE
     ror_matching != 'NA')
 SELECT
@@ -75,7 +75,7 @@ SELECT
   CASE WHEN online_year IS NULL THEN pubyear ELSE online_year END AS earliest_year,
[…]

### 2024-07-18 analysis/intersections.Rmd
Investigate intersections between different sets

---
title: "intersections"
output: github_document
---

```{r, knitr_setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r r_setup}
library(bigrquery)
library(tidyverse)
#' Connect to DB 
#' 
#' BigQuery HOAD (Crossref, OpenAlex and open friends)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  billing = "subugoe-collaborative",
  project = "subugoe-collaborative"
)
```

### Overlap between sets

Set 1: HOAD articles under TA as measured by first authorship, ROR and year match

Set 2: Replication data using WOS:TA as measured by first authorship, ROR and year match

Set 3: WOS correspondance author under TA as measured by Correspondance authorship, ROR and year match

```{r}
hoad_ta_articles <- readr::read_csv(here::here("data-raw", "ta_oa_inst.csv.gz"))

ta_hoad <- hoad_ta_articles |>
  filter(cr_year %in% 2018:2023, ta_active == TRUE) |>
  distinct(doi)

wos_replication <- tbl(bq_con,  I("hoa-article.hoa_comparision.wos_jct_ta"))

wos_rep_doi <- wos_replication |>
  filter(pubyear %in% 2018:2023 & author_seq_nr == 1 & ta == TRUE) |>
  distinct(doi) |>
  collect()

wos_cor_doi <- wos_replication |>
  filter(pubyear %in% 2018:2023 & corresponding == TRUE & ta == TRUE) |>
  distinct(doi) |>
  collect()

my_venn <- list(ta_hoad = unlist(ta_hoad), wos_rep_doi = unlist(wos_rep_doi), wos_cor_doi = unlist(wos_cor_doi))

library(UpSetR)
upset(fromList(my_venn), order.by = "freq")
```

### 2024-08-07 analysis/article_volume.Rmd
EDA tmp

--- 2024-05-10 analysis/eda.Rmd
@@
 ---
-title: "eda"
+title: "Article volume"
 output: github_document
@@
 
-## Web of Science
+```{r data_setup}
 
-### Open access
-
-```{sql add_oa_info, connection = bq_con, output.var = "wos_oa"}
-WITH oa_raw AS (SELECT DISTINCT item_id, doi, issn_l, CASE WHEN online_year IS NULL THEN pubyear ELSE online_year END AS earliest_year, is_oa
-FROM (
-SELECT
-  wos.item_id, 
-  wos.doi, 
-  wos.issn_l,
-  wos.pubyear,
-  CAST(REGEXP_EXTRACT(wos_pubdate_online, r'^[0-9]{4}') AS NUMERIC) AS online_year,
-  CASE
-    WHEN cc IS NOT NULL AND vor = 1 THEN TRUE
-  ELSE
-  FALSE
-END
-  AS is_oa
-FROM
-  `hoa-article.hoa_comparision.wos_matching_20240410` AS wos
-LEFT JOIN
-  `subugoe-collaborative.hoaddata.cc_md` AS cc
-ON
-  wos.doi = cc.doi ) )
-SELECT COUNT(DISTINCT item_id) as wos_articles, issn_l, earliest_year, is_oa
-FROM oa_raw
-GROUP BY issn_l, earliest_year, is_oa
-ORDER BY issn_l, earliest_year
 ```
-By year
-
-```{r}
-wos_all <- wos_oa |>
-  group_by(earliest_year) |>
-  summarise(articles = sum(wos_articles))
-
-wos_oa |>
-  filter(is_oa == TRUE) |>
-  group_by(earliest_year) |>
-  summarise(oa = sum(wos_articles)) |>
-  right_join(wos_all) |>
-  mutate(prop = oa / articles)
-```

### 2024-08-07 analysis/eda.Rmd
EDA tmp

@@
 ```
+
 By year
@@
 
-cr_per_year <- inner_join(cr_all_per_year, cr_oa) |>
+cr_per_year <- left_join(cr_all_per_year, cr_oa) |>
   mutate(cr_year = as.numeric(as.character(cr_year)))
 
-jn_wos_stats <- inner_join(jn_wos_24_stat, cr_per_year, by = c("earlies_year" = "cr_year", "issn_l"))
+jn_wos_stats <- left_join(cr_per_year, jn_wos_24_stat, by = c("cr_year" = "earlies_year" , "issn_l"))
+
+## add esac portfolios
+
+### in the anayltics, we need to decide whether we anaylse the intersection or the scenario all journals
+
+publisher_intersection_df <- hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) |>
+  inner_join(jn_wos_stats, by = "issn_l") |>
+  filter(cr_year %in% 2018:2023) |>
+  # only intersection
+  filter(!is.na(all)) |>
+  group_by(esac_publisher) |>
+  summarise(all = sum(all),
+            core = sum(core),
+            hybrid_oa = sum(hybrid_oa),
+            core_hybrid_oa = sum(core_hybrid_oa),
+            cr_oa = sum(cr_oa),
+            core_cr_oa = sum(core_cr_oa),
+            hoad_articles = sum(hoad_articles, na.rm = TRUE),
+            hoad_oa = sum(hoad_oa, na.rm = TRUE))
+
+publisher_intersection_by_year_df <- hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) |>
+  inner_join(jn_wos_stats, by = "issn_l") |>
+  filter(cr_year %in% 2018:2023) |>
+  # only intersection
+  filter(!is.na(all)) |>
+  group_by(esac_publisher, cr_year) |>
+  summarise(all = sum(all),
+            core = sum(core),
+            hybrid_oa = sum(hybrid_oa),
+            core_hybrid_oa = sum(core_hybrid_oa),
+            cr_oa = sum(cr_oa),
+            core_cr_oa = sum(core_cr_oa),
+            hoad_articles = sum(hoad_articles, na.rm = TRUE),
+            hoad_oa = sum(hoad_oa, na.rm = TRUE))
+
+p <- ggplot(publisher_intersection_by_year_df, aes(core_hybrid_oa, core_cr_oa, text = esac_publisher)) +
+  geom_point(alpha = 0.4) +
+  #scale_x_log10(limits = c(1,100000)) +
+  #scale_y_log10(limits = c(1,100000)) +
+  coord_equal() +
+  geom_abline(slope = 1, intercept = 0)  +
+  # geom_smooth(method = "lm") +
+  facet_wrap(~ cr_year) 
+plotly::ggplotly(p)
 ```
@@
   geom_point(alpha = 0.05) +
+  scale_x_log10(limits = c(0.001,1)) +
+  scale_y_log10(limits = c(0.001,1)) +
   coord_equal() +
-  geom_smooth(method = "lm") +
-  facet_wrap(~ earlies_year)
+  geom_abline(slope = 1, intercept = 0)  +
+  facet_wrap(~ cr_year)
 ```

### 2024-08-07 analysis/results.Rmd
EDA tmp

---
title: "Results"
output: github_document
---

```{r, knitr_setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  echo = TRUE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "70%",
  fig.align = "center",
  dpi = 300
)
options(scipen = 999, digits = 4)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r r_setup}
library(bigrquery)
library(tidyverse)
library(hoaddata)
#' Connect to DB 
#' 
#' BigQuery HOAD (Crossref, OpenAlex and open friends)
bq_con <- dbConnect(
  bigrquery::bigquery(),
  billing = "subugoe-collaborative",
  project = "subugoe-collaborative"
)
```

## Research questions

Are open data sources suitable to measure transformative agreements? More specifically, do analyses based on {hoaddata} yield comparable results compared to WoS and Scopus?

## Journal and article volume

- Hybrid Journal coverage
- Article coverage (unsing document types)
- classification of publisher (top 3, other)

### Journal level

- JCT Journal data: `hoaddata::jct_hybrid_jns` (excluding PNAS)
- Crossref: Journal-level data on article volume including OA: `hoaddata::jn_ind`
- Web of Science: `data-raw/wos_jct_items.csv`

Levels:

- All
- Active: Have published at least one article between 2019 - 2023
- Hybrid OA: Have published at least one OA article between 2019 - 2023

#### Overview

HOAD:

```{r hoad_journal_level_prep}
hoad_jns <- hoaddata::jct_hybrid_jns |>
  filter(issn_l != "0027-8424")

hoad_oa_excluded <- bq_table_download("subugoe-collaborative.hoaddata.cc_oa_prop") |>
  distinct(issn_l)

active_jns <- hoad_jns |>
 # filter(!issn_l %in% hoad_oa_excluded$issn_l) |>
  inner_join(hoaddata::jn_ind, by = "issn_l") |>
  filter(!cr_year %in% c("2017", "2018", "2024"))

active_jns_with_oa <- active_jns |> 
  filter(!is.na(cc)) 
```

- Number of hybrid journals: `r length(unique(hoad_jns$issn_l))`
- Number of active hybrid journals between 2019 and 2023: `r length(unique(active_jns$issn_l))`
- Number of active hybrid journals with at least one OA between 2019 and 2023: `r length(unique(active_jns_with_oa$issn_l))`

Web of Science:

```{r wos_journal_level_prep}
wos_jct_items_df <- readr::read_csv(here::here("data-raw", "wos_jct_items_df.csv"))

wos_hybrid_jns <- wos_jct_items_df |>
  # No PNAS
  filter(issn_l != "0027-8424") |>
  # No Full OA detected by OA prop
  filter(!issn_l %in% hoad_oa_excluded$issn_l)

wos_active_jns <- wos_hybrid_jns |>
  filter(pubyear %in% 2019:2023)

wos_active_jns_with_oa <- wos_active_jns |> 
  filter(oa_status == "hybrid")

# wos core 
wos_active_core <- wos_active_jns |>
  filter(item_type %in% c("Article", "Review"))
```

- Number of hybrid journals: `r length(unique(wos_hybrid_jns$issn_l))`
- Number of active hybrid journals between 2019 and 2023: `r length(unique(wos_active_jns$issn_l))`
- Number of active hybrid journals with at least one OA between 2019 and 2023: `r length(unique(wos_active_jns_with_oa$issn_l))`

Scopus

```{r}
scp_jct_items_df <- readr::read_csv(here::here("data-raw", "scp_jct_items_df.csv"))

scp_hybrid_jns <- scp_jct_items_df |>
  # No PNAS
  filter(issn_l != "0027-8424") |>
  # No Full OA detected by OA prop
  filter(!issn_l %in% hoad_oa_excluded$issn_l)

scp_active_jns <- scp_hybrid_jns |>
  filter(pubyear %in% 2019:2023)

scp_active_jns_with_oa <- scp_active_jns |> 
  filter(grepl("hybrid", oa_status))

# core
scp_active_core <- scp_active_jns |>
   filter(grepl("^Article|^Review", item_type))

scp_active_core_oa <- scp_active_core |>
    filter(grepl("hybrid", oa_status))
```

- Number of hybrid journals: `r length(unique(scp_hybrid_jns$issn_l))`
- Number of active hybrid journals between 2019 and 2023: `r length(unique(scp_active_jns$issn_l))`
- Number of active hybrid journals with at least one OA between 2019 and 2023: `r length(unique(scp_active_jns_with_oa$issn_l))`

Core

- Active Journals: `r length(unique(scp_active_core$issn_l))`
- Active journals with OA: `r length(unique(scp_active_core_oa$issn_l))`

Journals without Original Articles and Reviews (mostly Proceedings)

```{r}
scp_active_jns |> 
  filter(!issn_l %in% scp_active_core$issn_l) |> 
  group_by(issn_l) |>
  summarise(items = n_distinct(item_id)) |>
  arrange(desc(items)) |>
  janitor::adorn_totals()
```

### Journal analytics

Journals with items published between 2019 - 2023

```{r upset_jns}
library(UpSetR)

jn_list <- list(
  crossref = unique(active_jns$issn_l),
  wos = unique(wos_active_jns$issn_l),
  scp = unique(scp_active_jns$issn_l)
)

# Plot Journals Five Years Period
upset(fromList(jn_list), order.by = "freq")

```

Print out journals, which are not in Crossref, but in WoS or Scopus

WoS:

```{r}
wos_active_jns |>
  filter(issn_l %in% setdiff(jn_list$wos, jn_list$crossref)) |>
  group_by(issn_l) |>
  summarise(items = n_distinct(item_id)) |>
  arrange(desc(items)) |>
  janitor::adorn_totals()
```
(Chinese Physics B seems to have the Chinese Physics ISSN-L in WOS)

Scopus:

```{r}
scp_active_jns |>
  filter(issn_l %in% setdiff(jn_list$scp, jn_list$crossref)) |>
  group_by(issn_l) |>
  summarise(items = n_distinct(item_id)) |>
  arrange(desc(items)) |>
  janitor::adorn_totals()
```

- `0094-243X` represents API Proceedings, all content tagged as proceedings article in Crossref.
- `0893-7400` changed to `1547-1896` and is present in both datasets
- `0095-2338` changed to `1549-9596`

### Item level view

```{r}
cr_df <- readr::read_csv(here::here("data-raw/hoad_dois_all_19_23.csv"))

item_view_df <-  list(
  crossref = unique(cr_df$doi),
  wos = unique(unlist(wos_active_core[!is.na(wos_active_core$doi),"doi"])),
  scp = unique(unlist(scp_active_core[!is.na(scp_active_core$doi),"doi"]))
)

# Plot Journals Five Years Period
upset(fromList(item_view_df), order.by = "freq")
```

### OA View

Journals with at least one OA item according to database

```{r}
jn_list_with_oa <- list(
  crossref = unique(active_jns_with_oa$issn_l),
  wos = unique(wos_active_core$issn_l),
  scp = unique(scp_active_core$issn_l)
)

upset(fromList(jn_list_with_oa), order.by = "freq")
```
### Item level

### 2024-08-07 data-raw/001_data_gathering.R

@@ -293,3 +293,18 @@ scp_jct_affiliations <- tt |>
   as_tibble()
 write_csv(scp_jct_affiliations, "data-raw/scp_jct_affiliations.csv")
 
+### BigQuery raw
+
+hoad_dois_all <- DBI::dbGetQuery(bq_con, "SELECT
+  DISTINCT doi,
+  issn_l,
+  cr_year
+FROM
+  `subugoe-collaborative.hoaddata.cc_md`
+WHERE
+  cr_year BETWEEN 2019
+  AND 2023")
+
+# backup
+
+write_csv(hoad_dois_all, here::here("data-raw", "hoad_dois_all_19_23.csv"))

### 2024-08-08 analysis/results.Rmd
tmp save

@@
 wos_active_jns <- wos_hybrid_jns |>
-  filter(pubyear %in% 2019:2023)
+  # use online first, if available
+  mutate(online_year = lubridate::year(wos_pubdate_online)) |>
+  mutate(pub_year = ifelse(!is.na(online_year), online_year, pubyear)) |>
+  filter(pub_year %in% 2019:2023)
 
@@
 wos_active_core <- wos_active_jns |>
-  filter(item_type %in% c("Article", "Review"))
+  filter(grepl("Article|^Review$", item_type))
 ```
@@
 scp_active_jns <- scp_hybrid_jns |>
-  filter(pubyear %in% 2019:2023)
+   # use online first, if available
+  mutate(pub_year = ifelse(!is.na(first_pubyear), first_pubyear, pubyear)) |>
+  filter(pub_year %in% 2019:2023)
 
@@
 # Plot Journals Five Years Period
-upset(fromList(item_view_df), order.by = "freq")
+UpSetR::upset(UpSetR::fromList(item_view_df), order.by = "freq")
 ```
@@
 ```
-### Item level
+
+#### Compose a complex upset
+
+```{r upset_data}
+cr_active_upset <- cr_df |>
+  filter(issn_l %in% active_jns$issn_l) |>
+  group_by(issn_l) |>
+  summarise(n_cr = n_distinct(doi)) |>
+  mutate(crossref = TRUE)
+wos_active_upset <- wos_active_core |>
+  group_by(issn_l) |>
+  summarise(n_wos = n_distinct(item_id)) |>
+  mutate(wos = TRUE)
+scp_active_upset <- scp_active_core |>
+  group_by(issn_l) |>
+  summarise(n_scp = n_distinct(item_id)) |>
+  mutate(scopus = TRUE)
+
+jn_upset_ <- dplyr::full_join(cr_active_upset, wos_active_upset, by = "issn_l") |>
+  dplyr::full_join(scp_active_upset, by = "issn_l") |>
+  mutate(across(c(crossref, wos, scopus), ~replace_na(.x, FALSE))) |>
+  # Journal 5-years article volume
+  mutate(n = case_when(!is.na(n_cr) ~ n_cr, !is.na(n_scp) ~n_scp, TRUE ~ n_wos))
+
+# add year
+jn_age <- jn_ind |> 
+  mutate(cr_year = as.numeric(as.character(cr_year))) |> 
+  filter(cr_year %in% 2019:2023) |>
+  group_by(issn_l) |>
+  summarise(min_year = min(cr_year),
+            max_year = max(cr_year)) |>
+  mutate(jn_age = max_year - min_year + 1)
+
+jn_upset <- jn_upset_ |>
+  inner_join(jn_age, by = "issn_l")
+library(ComplexUpset)
+
+ComplexUpset::upset(
+  jn_upset,
+  c("crossref", "wos", "scopus"),
+  base_annotations = list(
+    `Journals` = (
+      intersection_size() +
+        scale_y_continuous(labels =  scales::number_format(big.mark = ","))
+    )
+  ),
+  annotations = list(
+    'Article Volume' = (
+      ggplot(jn_upset, aes(x = intersection, y = n)) +
+        geom_boxplot(outliers = FALSE) +
+        scale_y_log10(labels =  scales::number_format(big.mark = ",")) 
+    ),
+    # journal age, not very useful
+    'Journal age' = (
+      ggplot(jn_upset, aes(x = intersection, y = jn_age)) +
+         geom_violin())
+    )
+  )
+```
+
+What is the age of the journals?
+
+```{r journal_age}
+jn_age <- jn_ind |> 
+  mutate(cr_year = as.numeric(as.character(cr_year))) |> 
+  filter(cr_year %in% 2019:2023) |>
+  group_by(issn_l) |>
+  summarise(min_year = min(cr_year),
+            max_year = max(cr_year)) |>
+  mutate(jn_age = max_year - min_year + 1) 
+```

### 2024-08-15 analysis/results.Rmd
Journal database coverage: Upset by publisher

@@
 
-```{r upset_data}
+```{r upset_data, fig.asp=0.8}
 cr_active_upset <- cr_df |>
@@
   summarise(n_cr = n_distinct(doi)) |>
-  mutate(crossref = TRUE)
+  mutate(Crossref = TRUE)
 wos_active_upset <- wos_active_core |>
@@
   summarise(n_wos = n_distinct(item_id)) |>
-  mutate(wos = TRUE)
+  mutate(WoS = TRUE)
 scp_active_upset <- scp_active_core |>
@@
   summarise(n_scp = n_distinct(item_id)) |>
-  mutate(scopus = TRUE)
+  mutate(Scopus = TRUE)
 
@@
   dplyr::full_join(scp_active_upset, by = "issn_l") |>
-  mutate(across(c(crossref, wos, scopus), ~replace_na(.x, FALSE))) |>
+  mutate(across(c(Crossref, WoS, Scopus), ~replace_na(.x, FALSE))) |>
   # Journal 5-years article volume
@@
 
-jn_upset <- jn_upset_ |>
-  inner_join(jn_age, by = "issn_l")
+# Add Esac publisher
+
+esac <- hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) |>
+  mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           )) |>
+  mutate(publisher = forcats::fct_rev(publisher)) |>
+  distinct(issn_l, publisher)
+
 library(ComplexUpset)
+jn_upset <- left_join(jn_upset_, esac, by = "issn_l")
+
+size = get_size_mode('exclusive_intersection')
 
@@
   jn_upset,
-  c("crossref", "wos", "scopus"),
+  c("Crossref", "WoS", "Scopus"),
+  # Only journals in Crossref
+  min_size = 30,
+  width_ratio = 0.4,
+  stripes = "transparent",
+  name = "Data Source Coverage",
   base_annotations = list(
-    `Journals` = (
-      intersection_size() +
+    "Journals" = (
+      ComplexUpset::intersection_size(
+        counts = TRUE,
+        text = list(size = 9 / .pt),
+        text_mapping = aes(
+          label = paste0(!!upset_text_percentage(), '\n(', !!size, ')'),
+          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
+          y = ifelse(!!size > 5000, !!size - 4000, !!size)
+        )
+      ) +
         scale_y_continuous(labels = scales::number_format(big.mark = ","))
@@
   ),
-  annotations = list(
-    'Article Volume' = (
-      ggplot(jn_upset, aes(x = intersection, y = n)) +
+  annotations = list("Article Volume" = (
+    ggplot(mapping = aes(x = intersection, y = n, fill = fct_rev(publisher))) +
       geom_boxplot(outliers = FALSE) +
-        scale_y_log10(labels =  scales::number_format(big.mark = ",")) 
-    ),
-    # journal age, not very useful
-    'Journal age' = (
-      ggplot(jn_upset, aes(x = intersection, y = jn_age)) +
-         geom_violin())
-    )
-  )
+              scale_fill_manual(
+        "",
+        values = c(
+          "Elsevier" = "#e9711c",
+          "Springer Nature" = "#486a7e",
+          "Wiley" = "#068853",
+          "Other" = "#b3b3b3a0"
+        ), guide = "none") +
+      scale_y_log10(labels =  scales::number_format(big.mark = ","))),
+    "Publisher" = (
+      ggplot(mapping = aes(fill = publisher)) +
+        geom_bar(stat = "count", position ="fill") + 
+        scale_y_continuous("Publisher Share", labels=scales::percent_format()) +
+        scale_fill_manual(
+        "",
+        values = c(
+          "Elsevier" = "#e9711c",
+          "Springer Nature" = "#486a7e",
+          "Wiley" = "#068853",
+          "Other" = "#b3b3b3a0"
+        ), guide = "none"))), 
+  #,
+  # 'Total article' = (
+  #   ggplot(mapping = aes(x = intersection, y = n)) +
+  #     geom_bar(stat = "identity", color = "grey35", fill = "grey35")) +
+  #  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) ),
+  set_sizes=(
+        upset_set_size(
+          geom = geom_bar(aes(fill = publisher), width = 0.8), 
+          position='right')) +
+          scale_fill_manual("", values = c(
+          "Elsevier" = "#e9711c",
+          "Springer Nature" = "#486a7e",
+          "Wiley" = "#068853",
+          "Other" = "#b3b3b3a0")) +
+    scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
+    guides(fill = guide_legend(reverse = TRUE)) +
+    ylab("Journals"),
+      guides = "over",
+  matrix = (
+        intersection_matrix(
+            geom = geom_point(
+                shape = "square",
+                size=3.5
+            )))
+        ) +
+  patchwork::plot_layout(heights=c(2, 1.5, 2, 1))
 ```

### 2024-08-16 analysis/results.Rmd
More upsets :-)

@@
 library(hoaddata)
+library(patchwork)
+
 #' Connect to DB 
@@
   mutate(Crossref = TRUE)
-wos_active_upset <- wos_active_core |>
+wos_active_upset <- wos_active_jns |>
   group_by(issn_l) |>
@@
   mutate(WoS = TRUE)
-scp_active_upset <- scp_active_core |>
+scp_active_upset <- scp_active_jns |>
   group_by(issn_l) |>
@@
 library(ComplexUpset)
-jn_upset <- left_join(jn_upset_, esac, by = "issn_l")
+jn_upset <- inner_join(jn_upset_, esac, by = "issn_l") |>
+  distinct(issn_l, .keep_all = TRUE)
 
@@
 ```
+
+### DOI coverage
+
+```{r doi_coverage_check}
+jn_overlap <- jn_upset |> 
+  filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE)
+
+miss_doi <- cr_df |>
+  filter(issn_l %in% jn_overlap$issn_l) |> 
+  filter(!doi %in% wos_active_core$doi)
+
+# Check if it is in WOS (full)
+wos_type_diff <- wos_jct_items_df |>
+  filter(doi %in% miss_doi$doi)
+
+miss_doi |>
+  filter(!doi %in% wos_type_diff$doi) |>
+  sample_n(10)
+```
+
+### Add Article Coverage
+
+- hoaddata: Crossref exclunding paratext and proceedings as indicated by pattern macthing. Patterns mostly from Unpaywall / OpenAlex
+- WoS: All DOIs / DOIS only article and reviews
+- Scopus: All DOIs / DOIS only article and reviews
+
+```{r fig.asp=0.8}
+ doi_overlap_df <- cr_df |>
+  filter(issn_l %in% active_jns$issn_l) |>
+  mutate(in_wos_and_scopus = ifelse(doi %in% wos_active_jns$doi & doi %in% scp_active_jns$doi, 1, 0),
+         in_scopus = ifelse(doi %in% scp_active_jns$doi, 1, 0),
+         in_wos =  ifelse(doi %in% wos_active_jns$doi, 1, 0))
+doi_overlap <- doi_overlap_df |>
+  group_by(issn_l) |>
+  summarise(
+    doi_in_wos_and_scopus = sum(in_wos_and_scopus),
+    doi_in_scopus = sum(in_scopus),
+    doi_in_wos = sum(in_wos))
+
+jn_upset_articles <- jn_upset |> 
+  left_join(doi_overlap, by = "issn_l")
+
+### Plot
+size = get_size_mode('exclusive_intersection')
+
+upset_overview <- ComplexUpset::upset(
+  jn_upset_articles,
+  c("Crossref", "WoS", "Scopus"),
+  # Only journals in Crossref
+  min_size = 30,
+  width_ratio = 0.4,
+  set_sizes=FALSE,
+  stripes = "transparent",
+  wrap = TRUE,
+  name = "Data Source Coverage",
+  base_annotations = list(
+    "Journals" = (
+      ComplexUpset::intersection_size(
+        counts = TRUE,
+        text = list(size = 9 / .pt),
+        text_mapping = aes(
+          label = paste0(!!upset_text_percentage(), '\n(', !!size, ')'),
+          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
+          y = ifelse(!!size > 5000, !!size - 4000, !!size)
+        )
+      ) +
+        scale_y_continuous(labels = scales::number_format(big.mark = ","))
+    )
+  ),
+  annotations = list("Journal size\n(log scale)" = (
+    ggplot(mapping = aes(x = intersection, y = n)) +
+      geom_boxplot(outliers = FALSE) +
+      scale_y_log10(labels =  scales::number_format(big.mark = ","))),
+    "Articles" = (ggplot(mapping = aes(y = n_cr)) +
+                        geom_bar(stat = "identity", color = "grey80", aes(fill = "in Crossref only")) +
+                        geom_bar(aes(y = doi_in_wos_and_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
+                        geom_bar(aes(y = doi_in_wos, fill = "Shared"), stat = "identity", color = "#0093c7") +
+                        geom_bar(aes(y = doi_in_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
+                        scale_fill_manual(values = c(`in Crossref only` = "grey80", `Shared` = "#0093c7"), name = "") +
+                        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+                        theme(
+    legend.position = "top",
+    legend.justification = "right"
+  ) +
+    guides(fill = "none")
+    )),
+   matrix = (
+        intersection_matrix(
+            geom = geom_point(
+                shape = "square",
+                size = 2
+            )))
+        ) 
+upset_overview
+```
+
+```{r dois_over_the_years}
+doi_coverage_over_years <- doi_overlap_df |>
+  group_by(cr_year) |>
+  summarise(dois = n_distinct(doi),
+            in_wos_and_scopus = sum(in_wos_and_scopus, na.rm = TRUE),
+            in_scopus = sum(in_scopus, na.rm = TRUE),
+            in_wos = sum(in_wos)) |>
+  pivot_longer(cols = c(dois, in_wos_and_scopus, in_scopus, in_wos)) |>
+  mutate(cr_year = gsub("^20", "'", as.character(cr_year)))
+
+# dois all
+dois_all_by_year <- doi_coverage_over_years |>
+  filter(name == "dois") |>
+  select(-name)
+
+coverage_over_year_plot <- doi_coverage_over_years |>
+  filter(name != "dois") |>
+  # Fix order of panel
+  mutate(name = factor(name, levels = c("in_wos_and_scopus", "in_scopus", "in_wos"))) |>
+  ggplot(aes(x = cr_year, y = value)) +
+  geom_bar(data = dois_all_by_year, aes(fill = "in Crossref only"), color = "transparent", stat = "identity") +
+  geom_bar(aes(fill = "Shared"), color = "transparent", stat = "identity") +
+  facet_wrap( ~ name, ncol = 1) +
+  scale_fill_manual(values = c("grey80", "#0093c7"), name = "") +
+  scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+  labs(x = "Year published", y = "Articles with DOI") + 
+  theme_minimal() +
+  theme(legend.position = "top",
+        legend.justification = "right") 
+coverage_over_year_plot
+```
+
+```{r coverage_upset_patch, fig.asp=0.85}
+design = "AB"
+wrap_plots(
+  A = upset_overview,
+  B = coverage_over_year_plot,
+  design = design,
+  widths = c(3.5,1)
+) & plot_annotation(tag_levels = 'A') 
+```

### 2024-08-19 analysis/results.Rmd
Update coverage overview

--- 2024-08-08 analysis/results.Rmd
@@
 ---
-title: "Results"
+title: "Journal and article coverage"
 output: github_document
@@
 library(hoaddata)
+library(patchwork)
+library(ComplexUpset)
+
 #' Connect to DB 
@@
 
-- Hybrid Journal coverage
-- Article coverage (unsing document types)
-- classification of publisher (top 3, other)
-
-### Journal level
-
 - JCT Journal data: `hoaddata::jct_hybrid_jns` (excluding PNAS)
@@
 - Web of Science: `data-raw/wos_jct_items.csv`
+- Scopus: `data-raw/scp_jct_items.csv`
 
@@
 
-#### Overview
+### Create analytical datasets
 
-HOAD:
+#### {hoaddata}
 
@@
 
+# Full OA journals detected through OA proportion > .95
 hoad_oa_excluded <- bq_table_download("subugoe-collaborative.hoaddata.cc_oa_prop") |>
@@
 active_jns <- hoad_jns |>
- # filter(!issn_l %in% hoad_oa_excluded$issn_l) |>
+  # Only active hybrid journals from JCT
   inner_join(hoaddata::jn_ind, by = "issn_l") |>
@@
 
-Web of Science:
+#### Web of Science
 
@@
 wos_active_jns <- wos_hybrid_jns |>
-  # use online first, if available
+  # Use online first, if available
   mutate(online_year = lubridate::year(wos_pubdate_online)) |>
@@
   filter(grepl("Article|^Review$", item_type))
+
+wos_active_core_oa <- wos_active_jns |>
+   filter(grepl("hybrid", oa_status))
 ```
@@
 
-Scopus
+Core:
+
+- Active Journals: `r length(unique(wos_active_core$issn_l))`
+- Active journals with OA: `r length(unique(wos_active_core_oa$issn_l))`
+
+#### Scopus
 
@@
 scp_active_jns <- scp_hybrid_jns |>
-   # use online first, if available
+   # Use online first, if available
   mutate(pub_year = ifelse(!is.na(first_pubyear), first_pubyear, pubyear)) |>
@@
 
-### Journal analytics
+## Coverage analysis: Journal and article volume
+
+### Overview
 
@@
 
-```{r upset_jns}
-library(UpSetR)
+```{r journal_coverage_dataset}
+# DOI set too large to track it with Git, see 001_data_gathering how the data
+# was obtained from Google BQ
+cr_df <- readr::read_csv(here::here("data-raw/hoad_dois_all_19_23.csv"))
 
-jn_list <- list(
-  crossref = unique(active_jns$issn_l),
-  wos = unique(wos_active_jns$issn_l),
-  scp = unique(scp_active_jns$issn_l)
-)
+cr_active_upset <- cr_df |>
+  filter(issn_l %in% active_jns$issn_l) |>
+  group_by(issn_l) |>
+  summarise(n_cr = n_distinct(doi)) |>
+  mutate(Crossref = TRUE)
+wos_active_upset <- wos_active_jns |>
+  group_by(issn_l) |>
+  summarise(n_wos = n_distinct(item_id)) |>
+  mutate(WoS = TRUE)
+scp_active_upset <- scp_active_jns |>
+  group_by(issn_l) |>
+  summarise(n_scp = n_distinct(item_id)) |>
+  mutate(Scopus = TRUE)
 
-# Plot Journals Five Years Period
-upset(fromList(jn_list), order.by = "freq")
+jn_upset_ <- dplyr::full_join(cr_active_upset, wos_active_upset, by = "issn_l") |>
+  dplyr::full_join(scp_active_upset, by = "issn_l") |>
+  mutate(across(c(Crossref, WoS, Scopus), ~ replace_na(.x, FALSE))) |>
+  # Journal 5-years article volume
+  mutate(n = case_when(!is.na(n_cr) ~ n_cr, !is.na(n_scp) ~ n_scp, TRUE ~ n_wos))
 
+## DOI overlap
+
+doi_overlap_df <- cr_df |>
+  filter(issn_l %in% active_jns$issn_l) |>
+  mutate(in_wos_and_scopus = ifelse(doi %in% wos_active_jns$doi & doi %in% scp_active_jns$doi, 1, 0),
+         in_scopus = ifelse(doi %in% scp_active_jns$doi, 1, 0),
+         in_wos =  ifelse(doi %in% wos_active_jns$doi, 1, 0))
+doi_overlap <- doi_overlap_df |>
+  group_by(issn_l) |>
+  summarise(
+    doi_in_wos_and_scopus = sum(in_wos_and_scopus),
+    doi_in_scopus = sum(in_scopus),
+    doi_in_wos = sum(in_wos))
+
+jn_upset_articles <- jn_upset_ |> 
+  left_join(doi_overlap, by = "issn_l")
+
+# Backup
+write_csv(jn_upset_articles, here::here("data/jn_upset_articles.csv"))
 ```
 
-Print out journals, which are not in Crossref, but in WoS or Scopus
+```{r journal_article_overlap_upset}
+size <- get_size_mode("exclusive_intersection")
+
+upset_overview <- ComplexUpset::upset(
+  jn_upset_articles,
+  c("Crossref", "WoS", "Scopus"),
+  # Only journals in Crossref
+  min_size = 30,
+  width_ratio = 0.4,
+  set_sizes = FALSE,
+  stripes = "transparent",
+ # wrap = TRUE,
+  name = "Data Source Coverage",
+  base_annotations = list(
+    "Journals" = (
+      ComplexUpset::intersection_size(
+        counts = TRUE,
+        text = list(size = 9 / .pt),
+        text_mapping = aes(
+          label = paste0(!!upset_text_percentage(), '\n(', !!size, ')'),
+          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
+          y = ifelse(!!size > 5000, !!size - 4000, !!size)
+        )
+      ) +
+        scale_y_continuous(labels = scales::number_format(big.mark = ","))
+    )
+  ),
+  annotations = list("Journal size\n(log scale)" = (
+    ggplot(mapping = aes(x = intersection, y = n)) +
+      geom_boxplot(outliers = FALSE) +
+      scale_y_log10(labels =  scales::number_format(big.mark = ","))),
+    "Articles" = (ggplot(mapping = aes(y = n_cr)) +
+                        geom_bar(stat = "identity", color = "grey80", aes(fill = "in Crossref only")) +
+                        geom_bar(aes(y = doi_in_wos_and_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
+                        geom_bar(aes(y = doi_in_wos, fill = "Shared"), stat = "identity", color = "#0093c7") +
+                        geom_bar(aes(y = doi_in_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
+                        scale_fill_manual(values = c(`in Crossref only` = "grey80", `Shared` = "#0093c7"), name = "") +
+                        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+                        theme(
+    legend.position = "top",
+    legend.justification = "right"
+  ) +
+    guides(fill = "none")
+    )),
+   matrix = (
+        intersection_matrix(
+            geom = geom_point(
+                shape = "square",
+                size = 2
+            )))
+        ) 
+upset_overview
+```
+The plot is limited to journals, which are indexed in Crossref. However, there
+are `r jn_upset_articles |> filter(Crossref == FALSE) |> distinct(issn_l) |> nrow()` journals not incldued in Crossref, but in WoS or Scopus:
+
+```{r}
+jn_upset_articles |> 
+  filter(Crossref == FALSE) |> 
+  distinct(issn_l, n) |>
+  arrange(desc(n)) |>
+  janitor::adorn_totals()
+```
 
@@
 ```{r}
-wos_active_jns |>
-  filter(issn_l %in% setdiff(jn_list$wos, jn_list$crossref)) |>
-  group_by(issn_l) |>
-  summarise(items = n_distinct(item_id)) |>
-  arrange(desc(items)) |>
+jn_upset_articles |>
+  filter(Crossref == FALSE, Scopus == FALSE, WoS == TRUE) |>
+  distinct(issn_l, n) |>
+  arrange(desc(n)) |>
   janitor::adorn_totals()
@@
 ```{r}
-scp_active_jns |>
-  filter(issn_l %in% setdiff(jn_list$scp, jn_list$crossref)) |>
-  group_by(issn_l) |>
-  summarise(items = n_distinct(item_id)) |>
-  arrange(desc(items)) |>
+jn_upset_articles |>
+  filter(Crossref == FALSE, Scopus == TRUE, WoS == FALSE) |>
+  distinct(issn_l, n) |>
+  arrange(desc(n)) |>
   janitor::adorn_totals()
@@
 - `0095-2338` changed to `1549-9596`
+- `1542-5983` for `1542-5991` used in Scopus for different journal (Hearing Journal, 0745-7472). `1542-5983` journal changed to `2771-1897` since 2009
 
-### Item level view
+### By publisher
 
-```{r}
-cr_df <- readr::read_csv(here::here("data-raw/hoad_dois_all_19_23.csv"))
+```{r upset_publisher_prepare_data}
+# Add ESAC portfolio names
+esac <- hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) |>
+  mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           )) |>
+  mutate(publisher = forcats::fct_rev(publisher)) |>
+  distinct(issn_l, publisher)
 
-item_view_df <-  list(
-  crossref = unique(cr_df$doi),
-  wos = unique(unlist(wos_active_core[!is.na(wos_active_core$doi),"doi"])),
-  scp = unique(unlist(scp_active_core[!is.na(scp_active_core$doi),"doi"]))
+jn_upset_publisher <- inner_join(jn_upset_articles, esac, by = "issn_l") |>
+  # Because of Journal transfer, restrict choice to one portfolio
+  distinct(issn_l, .keep_all = TRUE)
+```
+
+```{r upset_publisher_plot}
+size = get_size_mode('exclusive_intersection')
+
+upset_publisher_plot <- ComplexUpset::upset(
+  jn_upset_publisher,
+  c("Crossref", "WoS", "Scopus"),
+  # Only journals in Crossref
+  min_size = 30,
+  width_ratio = 0.4,
+  stripes = "transparent",
+  name = "Data Source Coverage",
+  base_annotations = NULL,
+  annotations = list(
+    "Journal size\n(log-scale)" = (
+      ggplot(mapping = aes(x = intersection, y = n, fill = fct_rev(publisher))) +
+        geom_boxplot(outliers = FALSE) +
+        scale_fill_manual(
+          "",
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3"
+          ), guide = "none") +
+        scale_y_log10(labels = scales::number_format(big.mark = ",")) 
+    ),
+    "Articles" = (
+      ggplot(mapping = aes(y = n, fill = publisher, color = publisher)) +
+        geom_bar(stat = "identity") +
+        scale_fill_manual(
+          "",
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3a0"
+          ), guide = "none") +
+        scale_color_manual(
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3"
+          ), guide = "none"
+        ) + scale_y_continuous(labels = scales::number_format(big.mark = ",")) 
+    ),
+    "Publisher" = (
+      ggplot(mapping = aes(fill = publisher)) +
+        geom_bar(stat = "count", position = "fill") + 
+        scale_y_continuous("Journals", labels = scales::percent_format()) +
+        scale_fill_manual(
+          "",
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3"
+          ), guide = "none")
+    )
+  ),
+  set_sizes = (
+    upset_set_size(
+      geom = geom_bar(aes(fill = publisher), width = 0.8), 
+      position = "right"
+    ) +
+    scale_fill_manual(
+      "",
+      values = c(
+        "Elsevier" = "#e9711c",
+        "Springer Nature" = "#486a7e",
+        "Wiley" = "#068853",
+        "Other" = "#b3b3b3"
+      )
+    ) +
+    scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
+    guides(fill = guide_legend(reverse = TRUE)) +
+    ylab("Journals")
+  ),
+  guides = "over",
+  matrix = (
+    intersection_matrix(
+      geom = geom_point(
+        shape = "square",
+        size = 2
+      )
+    )
+  )
 )
 
-# Plot Journals Five Years Period
-UpSetR::upset(UpSetR::fromList(item_view_df), order.by = "freq")
-```
-
-### OA View
-
-Journals with at least one OA item according to database
-
-```{r}
-jn_list_with_oa <- list(
-  crossref = unique(active_jns_with_oa$issn_l),
-  wos = unique(wos_active_core$issn_l),
-  scp = unique(scp_active_core$issn_l)
-)
-
-upset(fromList(jn_list_with_oa), order.by = "freq")
-```
-
-#### Compose a complex upset
-
-```{r upset_data}
-cr_active_upset <- cr_df |>
-  filter(issn_l %in% active_jns$issn_l) |>
-  group_by(issn_l) |>
-  summarise(n_cr = n_distinct(doi)) |>
-  mutate(crossref = TRUE)
-wos_active_upset <- wos_active_core |>
-  group_by(issn_l) |>
-  summarise(n_wos = n_distinct(item_id)) |>
-  mutate(wos = TRUE)
-scp_active_upset <- scp_active_core |>
-  group_by(issn_l) |>
-  summarise(n_scp = n_distinct(item_id)) |>
-  mutate(scopus = TRUE)
-
-jn_upset_ <- dplyr::full_join(cr_active_upset, wos_active_upset, by = "issn_l") |>
-  dplyr::full_join(scp_active_upset, by = "issn_l") |>
-  mutate(across(c(crossref, wos, scopus), ~replace_na(.x, FALSE))) |>
-  # Journal 5-years article volume
-  mutate(n = case_when(!is.na(n_cr) ~ n_cr, !is.na(n_scp) ~n_scp, TRUE ~ n_wos))
-
-# add year
-jn_age <- jn_ind |> 
-  mutate(cr_year = as.numeric(as.character(cr_year))) |> 
-  filter(cr_year %in% 2019:2023) |>
-  group_by(issn_l) |>
-  summarise(min_year = min(cr_year),
-            max_year = max(cr_year)) |>
-  mutate(jn_age = max_year - min_year + 1)
-
-jn_upset <- jn_upset_ |>
-  inner_join(jn_age, by = "issn_l")
-library(ComplexUpset)
-
-ComplexUpset::upset(
-  jn_upset,
-  c("crossref", "wos", "scopus"),
-  base_annotations = list(
-    `Journals` = (
-      intersection_size() +
-        scale_y_continuous(labels =  scales::number_format(big.mark = ","))
-    )
-  ),
-  annotations = list(
-    'Article Volume' = (
-      ggplot(jn_upset, aes(x = intersection, y = n)) +
-        geom_boxplot(outliers = FALSE) +
-        scale_y_log10(labels =  scales::number_format(big.mark = ",")) 
-    ),
-    # journal age, not very useful
-    'Journal age' = (
-      ggplot(jn_upset, aes(x = intersection, y = jn_age)) +
-         geom_violin())
-    )
-  )
+upset_publisher_plot 
 ```
@@
 ```
+
+DOI coverage check
+
+```{r doi_coverage_check, eval = FALSE}
+jn_overlap <- jn_upset |> 
+  filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE)
+
+miss_doi <- cr_df |>
+  filter(issn_l %in% jn_overlap$issn_l) |> 
+  filter(!doi %in% wos_active_core$doi)
+
+# Check if it is in WOS (full)
+wos_type_diff <- wos_jct_items_df |>
+  filter(doi %in% miss_doi$doi)
+
+miss_doi |>
+  filter(!doi %in% wos_type_diff$doi) |>
+  sample_n(10)
+```
+
+10.1016/j.apmr.2022.08.623 -> research poster abstract
+10.1016/s1530-891x(20)39611-7 -> research abstract
+10.1007/s42835-023-01664-z -> publsihed 2024
+10.1097/yic.0000000000000534 -> published 2024
+10.1038/s41415-022-5051-7 -> others
+10.1016/j.matpr.2020.03.412 -> proceedings paper
+10.3233/jifs-219108 -> doi does not work
+10.1136/vr.m1027 -> paper review vet record!
+10.1002/jmv.29204 -> indexing date 2024, pubyear 2023 
+10.1159/000516826 -> subject index

### 2024-08-20 analysis/results.Rmd
Style upset plots

@@
 )
+
+# Set themes for plots
+my_base_size = 12
+
+my_upset_theme <-
+  theme(panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank(),
+        plot.title.position="plot",
+        plot.title = element_text(hjust = 0, vjust = 1, face = "bold"))
+
 ```
@@
 # Backup
+jn_upset_articles
 write_csv(jn_upset_articles, here::here("data/jn_upset_articles.csv"))
@@
 
-```{r journal_article_overlap_upset}
+```{r journal_article_overlap_upset, fig.asp=0.75}
+# Helper function to make text labels the same size as axis labels as seen in
+# <https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/>
+# Todo: Learn how to invoke it
+ggtext_size <- function(base_size, ratio = 0.8) {
+  ratio * base_size / ggplot2::.pt
+}
+
+bar_width <- 0.7
 size <- get_size_mode("exclusive_intersection")
@@
   min_size = 30,
-  width_ratio = 0.4,
   set_sizes = FALSE,
@@
         counts = TRUE,
-        text = list(size = 9 / .pt),
+        text = list(size = 8 / .pt),
         text_mapping = aes(
@@
           y = ifelse(!!size > 5000, !!size - 4000, !!size)
-        )
+        ), width = bar_width
       ) +
-        scale_y_continuous(labels = scales::number_format(big.mark = ","))
+        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
+        ggtitle("A")
     )
   ),
-  annotations = list("Journal size\n(log scale)" = (
+  annotations = list(
+    "Journal size\n(log scale)" = (
       ggplot(mapping = aes(x = intersection, y = n)) +
-      geom_boxplot(outliers = FALSE) +
-      scale_y_log10(labels =  scales::number_format(big.mark = ","))),
-    "Articles" = (ggplot(mapping = aes(y = n_cr)) +
-                        geom_bar(stat = "identity", color = "grey80", aes(fill = "in Crossref only")) +
-                        geom_bar(aes(y = doi_in_wos_and_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
-                        geom_bar(aes(y = doi_in_wos, fill = "Shared"), stat = "identity", color = "#0093c7") +
-                        geom_bar(aes(y = doi_in_scopus, fill = "Shared"), stat = "identity", color = "#0093c7") +
-                        scale_fill_manual(values = c(`in Crossref only` = "grey80", `Shared` = "#0093c7"), name = "") +
+        geom_boxplot(outliers = FALSE, width = bar_width) +
+        scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
+        ggtitle("C")
+    ),
+    "Articles" = (
+      ggplot(mapping = aes(y = n_cr)) +
+        geom_bar(
+          stat = "identity",
+          color = "grey80",
+          width = bar_width,
+          aes(fill = "in Crossref only")
+        ) +
+        geom_bar(
+          aes(y = doi_in_wos_and_scopus, fill = "Shared"),
+          stat = "identity",
+          width = bar_width,
+          color = "#0093c7"
+        ) +
+        geom_bar(
+          aes(y = doi_in_wos, fill = "Shared"),
+          stat = "identity",
+          width = bar_width,
+          color = "#0093c7"
+        ) +
+        geom_bar(
+          aes(y = doi_in_scopus, fill = "Shared"),
+          stat = "identity",
+          width = bar_width,
+          color = "#0093c7"
+        ) +
+        scale_fill_manual(
+          values = c(
+            `in Crossref only` = "grey80",
+            `Shared` = "#0093c7"
+          ),
+          name = "DOI"
+        ) +
         scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
-                        theme(
-    legend.position = "top",
-    legend.justification = "right"
-  ) +
-    guides(fill = "none")
-    )),
-   matrix = (
-        intersection_matrix(
-            geom = geom_point(
-                shape = "square",
-                size = 2
+        ggtitle("B")
+    )
+  ),
+  matrix = (intersection_matrix(geom = geom_point(
+    shape = "square", size = 2
   )))
 )
-upset_overview
+upset_overview & my_upset_theme
 ```
+
 The plot is limited to journals, which are indexed in Crossref. However, there
@@
 
-```{r upset_publisher_plot}
+```{r upset_publisher_plot, fig.asp=0.75}
 size = get_size_mode('exclusive_intersection')
@@
       ggplot(mapping = aes(x = intersection, y = n, fill = fct_rev(publisher))) +
-        geom_boxplot(outliers = FALSE) +
+        geom_boxplot(outliers = FALSE, width = bar_width) +
         scale_fill_manual(
@@
           ), guide = "none") +
-        scale_y_log10(labels = scales::number_format(big.mark = ",")) 
+        scale_y_log10(labels = scales::number_format(big.mark = ",")) +
+        ggtitle("C")
     ),
@@
       ggplot(mapping = aes(y = n, fill = publisher, color = publisher)) +
-        geom_bar(stat = "identity") +
+        geom_bar(stat = "identity", width = bar_width) +
         scale_fill_manual(
@@
           ), guide = "none"
-        ) + scale_y_continuous(labels = scales::number_format(big.mark = ",")) 
+        ) + scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
+        ggtitle("B") 
     ),
@@
       ggplot(mapping = aes(fill = publisher)) +
-        geom_bar(stat = "count", position = "fill") + 
+        geom_bar(stat = "count", position = "fill", width = bar_width) + 
         scale_y_continuous("Journals", labels = scales::percent_format()) +
@@
             "Other" = "#b3b3b3"
-          ), guide = "none")
+          ), guide = "none") +
+        ggtitle("A")
     )
@@
 
-upset_publisher_plot 
+upset_publisher_plot & my_upset_theme 
 ```

### 2024-10-17 analysis/results.Rmd
big mark upset, introduces new issue

@@
         text_mapping = aes(
-          label = paste0(!!upset_text_percentage(), '\n(', !!size, ')'),
+          label = paste0(!!upset_text_percentage(), '\n(', format(!!size, big.mark = ","), ')'),
           colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),

### 2024-11-07 data-raw/indicators.R
Re-calulate indicators after making sure that they are distinctively counted

@@ -156,52 +156,53 @@ wos_ta_by_year <- dbGetQuery(bq_con, "WITH
     DISTINCT item_id,
     earliest_year,
     CASE
-      WHEN author_seq_nr = 1 THEN 1
-      ELSE 0
+      WHEN author_seq_nr = 1 THEN item_id
+      ELSE NULL
   END
     AS is_first_author,
     CASE
-      WHEN corresponding = TRUE THEN 1
-      ELSE 0
+      WHEN corresponding = TRUE THEN item_id
+      ELSE NULL
   END
     AS is_corresponding_author,
-    REGEXP_CONTAINS(oa_status, 'hybrid') AS is_oa,
+    CASE
+      WHEN REGEXP_CONTAINS(oa_status, 'hybrid') THEN item_id
+      ELSE NULL
+  END
+    AS is_oa,
     core
   FROM
-    `hoa-article.hoa_comparision.wos_jct_match` ),
+    `hoa-article.hoa_comparision.wos_jct_match`),
   article_stats AS (
   SELECT
     earliest_year,
     'all' AS pubtype,
     COUNT(DISTINCT item_id) AS ta_articles,
-    SUM(CASE
-        WHEN is_oa = TRUE THEN 1
-        ELSE 0
-    END
-      ) AS ta_oa_articles,
-    -- Count for first author articles (all and OA only)
-    SUM(is_first_author) AS ta_first_author_articles,
-    SUM(CASE
-        WHEN is_oa = TRUE THEN is_first_author
-        ELSE 0
+    COUNT(DISTINCT is_oa) AS ta_oa_articles,
+    COUNT(DISTINCT is_first_author) AS ta_first_author_articles,
+    COUNT(DISTINCT
+      CASE
+        WHEN is_oa IS NOT NULL THEN is_first_author
+        ELSE NULL
     END
       ) AS ta_oa_first_author_articles,
-    -- Count for corresponding author articles (all and OA only)
-    SUM(is_corresponding_author) AS ta_corresponding_author_articles,
-    SUM(CASE
-        WHEN is_oa = TRUE THEN is_corresponding_author
-        ELSE 0
+    COUNT(DISTINCT is_corresponding_author) AS ta_corresponding_author_articles,
+    COUNT(DISTINCT
+      CASE
+        WHEN is_oa IS NOT NULL THEN is_corresponding_author
+        ELSE NULL
     END
       ) AS ta_oa_corresponding_author_articles,
[…]

### 2024-11-07 data-raw/001_data_gathering.R
Use KB dump

@@ -430,19 +430,40 @@ WHERE
 
 write_csv(hoad_dois_all, here::here("data-raw", "hoad_dois_all_19_23.csv"))
 
-# Backup OpenAlex funder information and corresponding author data
+# Backup OpenAlex funder information and corresponding author data using KB
 
-dbExecute(bq_con, "DROP TABLE hoa-article.hoaddata_sep24.oalex_cr_raw_cor_funder")
+#dbExecute(bq_con, "DROP TABLE hoa-article.hoaddata_sep24.oalex_cr_raw_cor_funder")
 
-dbExecute(bq_con, "-- Safeguard corresponding author and funding data
-CREATE table hoa-article.hoaddata_sep24.oalex_cr_raw_cor_funder AS SELECT DISTINCT cr.doi, inst.ror,  au.is_corresponding, author_position, ARRAY_LENGTH(oalex.grants) as has_funder
-  FROM `hoa-article.hoaddata_sep24.cr_raw` AS cr
-  LEFT JOIN `subugoe-collaborative.openalex.works` AS oalex 
-    ON oalex.doi = cr.doi,
-  UNNEST(authorships) AS au,
-  UNNEST(au.institutions) AS inst
-  WHERE au.is_corresponding = TRUE
-")
+dbExecute(kb_con, "CREATE TABLE oalex_cr_raw_cor_funder AS WITH filtered_works AS (
+  SELECT 
+    id,
+    REPLACE(doi, 'https://doi.org/', '') as short_doi,
+    corresponding_author_ids, 
+    corresponding_institution_ids,
+    json_array_length(corresponding_author_ids) AS num_corresponding_authors,
+    json_array_length(corresponding_institution_ids) AS num_corresponding_institutions
+  FROM 
+    fiz_openalex_rep_20240831_openbib.works
+)
+SELECT 
+    hoad.doi, 
+    w.corresponding_author_ids, 
+    w.corresponding_institution_ids,
+    w.num_corresponding_authors,
+    w.num_corresponding_institutions,
+    f.funder_id
+FROM 
+    hoad_dois_all_19_23 hoad
+LEFT JOIN 
+    filtered_works w
+ON
+    hoad.doi = w.short_doi
+LEFT JOIN
+    fiz_openalex_rep_20240831_openbib.works_funders f
+ON 
+    w.id = f.work_id")
+
[…]

### 2024-11-07 analysis/article_volume.Rmd
Start exploring article counts

@@
 library(tidyverse)
+library(patchwork)
 #' Connect to DB 
@@
 ```{r data_setup}
+# Web of Science
+wos_by_year <- readr::read_csv(here::here("data", "wos_by_year.csv"))
+wos_ta_by_year <- readr::read_csv(here::here("data", "wos_ta_by_year.csv"))
+
+wos_df <- inner_join(wos_by_year, wos_ta_by_year, by = c("earliest_year", "pub_type" = "pubtype"))
+
+# Scopus
+scp_by_year <- readr::read_csv(here::here("data", "scp_by_year.csv"))
+scp_ta_by_year <- readr::read_csv(here::here("data", "scp_ta_by_year.csv"))
+
+scp_df <- inner_join(scp_by_year, scp_ta_by_year, by = c("earliest_year" = "first_pubyear", "pub_type" = "pubtype"))
+
+#scp_jn_by_year <- readr::read_csv(here::here("data", sc))
+```
+
+```{r ta_oa_by_year}
+# core
+wos_oa_by_year <- wos_df |> 
+  filter(pub_type == "core") |> 
+  select(earliest_year, n_articles, n_oa_articles, ta_oa_articles,
+         ta_oa_first_author_articles, ta_oa_corresponding_author_articles) |> 
+  pivot_longer(c(ta_oa_articles, ta_oa_first_author_articles, ta_oa_corresponding_author_articles)) |> 
+  mutate(src = "Web of Science")
+
+scp_oa_by_year <- scp_df |> 
+  filter(pub_type == "core") |> 
+  select(earliest_year, n_articles, n_oa_articles, ta_oa_articles,
+         ta_oa_first_author_articles, ta_oa_corresponding_author_articles) |> 
+  pivot_longer(c(ta_oa_articles, ta_oa_first_author_articles, ta_oa_corresponding_author_articles)) |> 
+  mutate(src = "Scopus")
+
+wos_oa_by_year_plot <- ggplot() +
+  geom_bar(
+    data = distinct(wos_oa_by_year, earliest_year, n_oa_articles, n_articles),
+    aes(
+      x = earliest_year,
+      y = n_oa_articles,
+      fill = "All"
+    ),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  geom_bar(
+    data = wos_oa_by_year,
+    aes(
+      x = earliest_year,
+      y = value,
+      fill = "TA OA"
+    ),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  facet_grid(src ~ name) +
+  scale_fill_manual(
+    values = c("#b3b3b3a0", "#56B4E9"),
+    name = "",
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  # scale_y_continuous(
+  #   expand = expansion(mult = c(0, 0.05)),
+  #   labels = scales::percent_format(),
+  #   limits = c(0, .20),
+  #   breaks = c(0, .05, .1, .15)
+  # ) +
+  theme_minimal() +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left",
+    axis.text.x = element_blank()
+  )
+
+scp_oa_by_year_plot <- ggplot() +
+  geom_bar(
+    data = distinct(scp_oa_by_year, earliest_year, n_oa_articles, n_articles),
+    aes(
+      x = earliest_year,
+      y = n_oa_articles,
+      fill = "All"
+    ),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  geom_bar(
+    data = scp_oa_by_year,
+    aes(
+      x = earliest_year,
+      y = value,
+      fill = "TA OA"
+    ),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  facet_grid(src ~ name) +
+  scale_fill_manual(
+    values = c("#b3b3b3a0", "#56B4E9"),
+    name = "",
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  # scale_y_continuous(
+  #   expand = expansion(mult = c(0, 0.05)),
+  #   labels = scales::percent_format(),
+  #   limits = c(0, .20),
+  #   breaks = c(0, .05, .1, .15, 0.2)
+ # ) +
+  theme_minimal() +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left",
+    axis.text.x = element_blank()
+  )
+
+wos_oa_by_year_plot / scp_oa_by_year_plot + plot_layout(axes = "collect", axis_titles = "collect")
+```
+
+## Check Overlap
+
+```{r}
+# Overlap data
+jn_overlap <- readr::read_csv(here::here("data", "jn_upset_articles.csv"))
+
+# Get TA DOIs and upset
+hoad_ta_oa_dois <- readr::read_csv(here::here("data-raw", "ta_oa_inst.csv")) |>
+  filter(cr_year %in% 2018:2023, ta_active == TRUE, !is.na(cc)) |> 
+  distinct(doi) |>
+  mutate(src = "hoad")
+
+wos_scp_ta_oa_dois <- DBI::dbGetQuery(bq_con, "SELECT DISTINCT doi, issn_l, author_seq_nr, corresponding, 'Scopus' AS src
+FROM `hoa-article.hoa_comparision.scp_jct_match` 
+WHERE ta_flag = TRUE AND core = TRUE AND oa_status LIKE '%hybrid%' AND first_pubyear BETWEEN 2018 AND 2023
+UNION ALL
+SELECT DISTINCT doi, issn_l, author_seq_nr, corresponding, 'Web of Science' AS src
+FROM `hoa-article.hoa_comparision.wos_jct_match` 
+WHERE ta_flag = TRUE AND core = TRUE AND oa_status LIKE '%hybrid%' AND earliest_year BETWEEN 2018 AND 2023")
+
+my_df <- wos_scp_ta_oa_dois |>
+  filter(corresponding == TRUE) |>
+  bind_rows(hoad_ta_oa_dois) |>
+  distinct(doi, src, issn_l)
+
+# Create binary columns for each source
+data_wide <- my_df |>
+  mutate(value = 1) |>
+  # Split sources if they contain multiple values
+#  separate_rows(src, sep = " and ") %>%
+  # Create wide format with binary indicators
+  pivot_wider(
+    id_cols = c(doi, issn_l),
+    names_from = src,
+    values_from = value,
+    values_fill = 0
+  )
+
+# Create the UpSet plot
+upset_plot <- upset(
+  data_wide,
+  intersect = colnames(data_wide)[-1],  # Exclude the DOI column
+  name = "Source Intersections",
+  width_ratio = 0.1,
+  themes = upset_modify_themes(
+    list(
+      "intersections_matrix" = theme(
+        text = element_text(size = 10),
+        plot.margin = margin(1, 1, 1, 1, "cm")
+      ),
+      "Intersection size" = theme(
+        axis.text.y = element_text(size = 10),
+        axis.title.y = element_text(size = 12)
+      ),
+      "Overall set size" = theme(
+        axis.text.x = element_text(size = 10, angle = 45, hjust = 1),
+        axis.title.x = element_text(size = 12)
+      )
+    )
+  ),
+  set_sizes = (
+    upset_set_size() + 
+    theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  ),
+  matrix = (
+    intersection_matrix(
+      geom = geom_point(size = 3)
+    )
+  ),
+  base_annotations = list(
+    "Intersection size" = intersection_size(
+      mapping = aes(fill = "black")
+    )
+  )
+)
+```
+
+```{r}
+tt <- wos_scp_ta_oa_dois |>
+  mutate(cat = ifelse(author_seq_nr == 1, paste0("first_", src), paste0("corresponding_", src))) |> bind_rows(hoad_ta_oa_dois) |>
+  distinct(doi, cat) |>
+  mutate(cat = ifelse(is.na(cat), "first_hoad", cat))
+
+tt_wide <- tt |>
+ # select(-issn_l) |>
+  mutate(value = 1) |>
+  pivot_wider(
+    id_cols = c(doi),
+    names_from = cat,
+    values_from = value,
+    values_fill = 0
+  )
+
+upset_plot <- upset(
+  tt_wide,
+  intersect = colnames(tt_wide)[-1],
+  name = "Source Intersections",
+  group_by = "sets")
 

### 2024-11-08 data-raw/ta_matching.R
Add scopus affiliation id

@@ -206,6 +206,7 @@ scp_jct_match_sql <- "WITH
     scp_items.ror,
     scp_items.countrycode,
     scp_items.author_seq_nr,
+    scp_items.vendor_org_id,
     scp_items.corresponding,
     scp_items.oa_status,
     scp_items.core,

### 2024-11-12 analysis/results.Rmd
Allow scp ta matching when there is more than one affiliation ID available

@@
 10.1159/000516826 -> subject index
+
+## Replication
+
+Here, I want to check whether we find consistent results across the three databases relative to open access uptake in hybrid journals in general, and relative to open access through agreements in particular.
+A particular focus will be whether findings change substantially when counting first or corresponding authors
+
+Working definition replication: In scientific research, a replication is commonly defined as a study that is conducted using the same or similar methods as the original investigation, in order to evaluate whether consistent results can be obtained [1].
+https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/
+
+```{r}
+# Web of Science
+wos_jn_by_year <- readr::read_csv(here::here("data", "wos_jn_by_year.csv"))
+
+wos_jn_ta_by_year <- readr::read_csv(here::here("data", "wos_jn_ta_by_year.csv"))
+
+wos_ta_df <- wos_jn_by_year |>
+  # only hybrid journals with at least one oa between 2018 - 2023
+  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
+  left_join(wos_jn_ta_by_year, by = c("issn_l", "earliest_year", "pub_type" = "pubtype")) |>
+  # Focus on core
+  filter(pub_type == "core") |>
+  mutate(src = "Web of Science") |>
+  select(-pub_type, -p)
+
+# Scopus
+scp_jn_by_year <-  readr::read_csv(here::here("data", "scp_jn_by_year.csv"))
+
+scp_jn_ta_by_year <- readr::read_csv(here::here("data", "scp_ta_jn_by_year.csv"))
+
+scp_ta_df <-  scp_jn_by_year |>
+  # only hybrid journals with at least one oa between 2018 - 2023
+  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
+  left_join(scp_jn_ta_by_year, by = c("issn_l", "earliest_year" = "first_pubyear", "pub_type" = "pubtype")) |>
+  # Focus on core
+  filter(pub_type == "core") |>
+  mutate(src = "Scopus") |>
+  select(-pub_type, -p)
+
+# HOAD
+hoad_jn_by_year_ <- active_jns_with_oa |>
+  mutate(cr_year = as.numeric(as.character(cr_year))) |>
+  distinct(issn_l, cr_year, n_articles = jn_all)
+hoad_oa_jn_by_year <- hoaddata::cc_articles |>
+  group_by(issn_l, cr_year) |>
+  summarise(n_oa_articles = n_distinct(doi))
+
+hoad_jn_by_year <- inner_join(hoad_jn_by_year_, hoad_oa_jn_by_year, by = c("issn_l", "cr_year"))
+
+hoad_jn_ta_by_year <- readr::read_csv(here::here("data-raw", "ta_oa_inst.csv"))
+
+hoad_ta_by_year <- hoad_jn_ta_by_year |>
+  filter(issn_l %in% active_jns_with_oa$issn_l, ta_active == TRUE) |>
+  group_by(issn_l, cr_year) |>
+  summarise(ta_first_author_articles = n_distinct(doi))
+
+hoad_ta_oa_by_year <- hoad_jn_ta_by_year |>
+  filter(issn_l %in% active_jns_with_oa$issn_l, ta_active == TRUE, !is.na(cc)) |>
+  group_by(issn_l, cr_year) |>
+  summarise(ta_oa_first_author_articles = n_distinct(doi))
+
+hoad_jn_ta_by_year_ <- left_join(hoad_ta_by_year, hoad_ta_oa_by_year)
+
+hoad_ta_df <- hoad_jn_by_year |>
+  left_join(hoad_jn_ta_by_year_) |>
+  # https://stackoverflow.com/a/73621978
+  {\(.) {replace(.,is.na(.),0)}}() |>
+  mutate(src = "HOAD") |>
+  distinct(issn_l, earliest_year = cr_year, n_oa_articles, n_articles, ta_articles = ta_first_author_articles, ta_first_author_articles, ta_oa_articles = ta_oa_first_author_articles, ta_oa_first_author_articles, src)
+
+uptake_df <- bind_rows(wos_ta_df, scp_ta_df, hoad_ta_df) |>
+  mutate_if(is.numeric, ~replace(., is.na(.), 0))
+```
+
+```{r plot_uptake}
+# Data prep
+uptake_by_year <- uptake_df |>
+  group_by(earliest_year, src) |>
+  summarize(
+    n_articles = sum(n_articles),
+    n_oa_articles = sum(n_oa_articles),
+    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
+    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
+  )
+
+uptake_by_year_df <- uptake_by_year |>
+  # Remove HOAD because we only have first author data available
+  filter(src != "HOAD") |>
+  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
+
+hoad_panel <- uptake_by_year |>
+  filter(src == "HOAD") |>
+  ggplot(aes(x = earliest_year)) +
+  geom_bar(
+    aes(y = n_oa_articles / n_articles, fill = "All"),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  geom_bar(
+    aes(y = ta_oa_first_author_articles / n_articles, fill = "TA OA"),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  geom_text(
+    aes(
+      label = paste0(
+        round(ta_oa_first_author_articles / n_oa_articles * 100, 1),
+        "%\n(",
+        format(ta_oa_first_author_articles, big.mark = ","),
+        ")"
+      ),
+      y = ta_oa_first_author_articles / n_articles),
+    position = position_dodge(width = 1),
+    vjust = -0.2,
+    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
+    size = 0.8 * my_base_size / .pt
+  ) +
+  scale_fill_manual(
+    values = c("#b3b3b3a0", "#56B4E9"),
+    name = "",
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    limits = c(0, .20),
+    breaks = c(0, .05, .1, .15, .20)
+  ) +
+  theme_minimal(base_size = my_base_size) +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left",
+    axis.text.x = element_blank()
+  )
+
+wos_scp_uptake_panel <- uptake_by_year_df |>
+  distinct(earliest_year, n_oa_articles, n_articles, src, name, value) |>
+  # Modify factor levels to control order
+  mutate(
+    src = factor(src, levels = c("Scopus", "Web of Science")),
+    name = factor(name, 
+                 levels = c("ta_oa_first_author_articles", "ta_oa_corresponding_author_articles"))
+  ) |>
+  ggplot() +
+  geom_bar(
+    aes(
+      x = earliest_year,
+      y = n_oa_articles / n_articles,
+      fill = "All"
+    ),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  geom_bar(
+    aes(
+      x = earliest_year,
+      y = value / n_articles,
+      fill = "TA OA"
+    ),
+    stat = "identity",
+    color = "#000000"
+  ) +
+   geom_text(
+    aes(
+      label = paste0(
+        round(value / n_oa_articles * 100, 1),
+        "%\n(",
+        format(value, big.mark = ","),
+        ")"
+      ),
+      x = earliest_year,
+      y = value / n_articles),
+    position = position_dodge(width = 1),
+    vjust = -0.2,
+    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
+    size = 0.8 * my_base_size / .pt
+  ) +
+  facet_grid(fct_rev(src) ~ name) +
+  scale_fill_manual(
+    values = c("#b3b3b3a0", "#56B4E9"),
+    name = "",
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    limits = c(0, .20),
+    breaks = c(0, .05, .1, .15, .20)
+  ) +
+  theme_minimal(base_size = my_base_size) +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left",
+    axis.text.x = element_blank())
+```

### 2024-11-12 data-raw/ta_matching.R

@@ -174,17 +174,21 @@ scp_jct_match_sql <- "WITH
     scp_jct.author_seq_nr,
     scp_jct.corresponding,
     scp_jct.vendor_org_id,
+    vendor_org_id_part,
     scp_jct.countrycode,
     matching.ror
   FROM
     `hoa-article.hoa_comparision.scp_jct_affiliations` AS scp_jct
+  # Be aware of multiple ids
+  CROSS JOIN
+    UNNEST(SPLIT(scp_jct.vendor_org_id, ',')) AS vendor_org_id_part
   LEFT JOIN
     `hoa-article.hoa_comparision.scp_ror_matching` AS matching
   ON
-    scp_jct.vendor_org_id = CAST(matching.vendor_org_id AS STRING)
+    vendor_org_id_part = CAST(matching.vendor_org_id AS STRING)
   WHERE
     matching.ror != 'NA'),
-  -- Filter and add additional metadata from Web of Science items
+  -- Filter and add additional metadata from Scopus
   scp_items AS (
   SELECT
     DISTINCT scp_ror.*,
@@ -206,7 +210,7 @@ scp_jct_match_sql <- "WITH
     scp_items.ror,
     scp_items.countrycode,
     scp_items.author_seq_nr,
-    scp_items.vendor_org_id,
+    scp_items.vendor_org_id_part as vendor_org_id,
     scp_items.corresponding,
     scp_items.oa_status,
     scp_items.core,

### 2024-11-12 analysis/results.Rmd
Compare uptake

@@
   fig.asp = 0.618,
-  out.width = "70%",
+  out.width = "99%",
   fig.align = "center",
@@
 # Set themes for plots
-my_base_size = 12
+my_base_size = 10
 
@@
 
-```{r plot_uptake}
+```{r plot_uptake, fig.width=8}
 # Data prep
 uptake_by_year <- uptake_df |>
+  mutate(earliest_year = as.character(earliest_year)) |>
   group_by(earliest_year, src) |>
@@
   # Remove HOAD because we only have first author data available
-  filter(src != "HOAD") |>
+  filter(src != "HOAD", earliest_year != "2018") |>
   pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
@@
     legend.position = "top",
-    legend.justification = "left",
-    axis.text.x = element_blank()
+    legend.justification = "left"
+  )
+
+# Prepare facet labels
+my_facet_labels <- c(ta_oa_first_author_articles = "First Author",
+                  ta_oa_corresponding_author_articles = "Corresponding Author",
+                  `Web of Science` = "Web of Science",
+                  `Scopus` = "Scopus"
                 )
@@
   mutate(
-    src = factor(src, levels = c("Scopus", "Web of Science")),
+    src = factor(src, levels = c("Web of Science", "Scopus")),
     name = factor(name, 
@@
   ) +
-  facet_grid(fct_rev(src) ~ name) +
+  facet_grid(src ~ name, labeller = as_labeller(my_facet_labels)) +
   scale_fill_manual(
@@
     legend.position = "top",
-    legend.justification = "left",
-    axis.text.x = element_blank())
+    legend.justification = "left")
+
+## Combine to patchwork
+layout <- "
+AC
+BB
+"
+wrap_plots(
+  A = hoad_panel,
+  B = wos_scp_uptake_panel,
+  C = guide_area(),
+  design = layout,
+  guides = "collect",
+  axes = "collect",
+  heights = c(0.5, 1)
+) &
+  theme(
+    legend.position = "top",
+    legend.justification = "right",
+    legend.direction = "horizontal",
+ #   panel.grid.minor = element_blank(),
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title=element_text(size = 10)) &
+    plot_annotation(tag_levels = 'A')
+
 ```

### 2024-11-12 data-raw/001_data_gathering.R

@@ -379,7 +379,7 @@ scp_jct_items_df <- scp_jct_items |>
   mutate(oa_status = as.character(gsub('\\{|\\}|"', '', oa_status)))
 write_csv(scp_jct_items_df, "data-raw/scp_jct_items_df.csv")
 
-#### Add WOS affiliation data for first data and corresponding authors
+#### Add Scopus affiliation data for first data and corresponding authors
 dbExecute(kb_con, "DROP TABLE scp_jct_affiliations")
 
 dbExecute(kb_con, "CREATE table scp_jct_affiliations AS select distinct jct.item_id,

### 2024-11-20 analysis/results.Rmd
First draft oa by publisher

@@
     plot_annotation(tag_levels = 'A')
+```
+
+### Scatterplot
+
+```{r}
+uptake_jn_five_jns <- uptake_df |>
+  filter(earliest_year != "2018") |>
+  group_by(issn_l, src) |>
+  summarize(
+    n_articles = sum(n_articles),
+    n_oa_articles = sum(n_oa_articles),
+    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
+    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
+  )
+
+uptake_scatter_df <- uptake_jn_five_jns |>
+  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles)) |>
+  select(issn_l, src, n_articles, n_oa_articles, author_position_oa = name, value) |>
+  mutate(cat = tolower(paste(src, author_position_oa, sep = "_")))
+```
+
+```{r}
+plot_uptake_jn_five_jns <- uptake_jn_five_jns |>
+  mutate(oa_ta_first = ta_oa_first_author_articles,
+         oa_ta_corr =  ta_oa_corresponding_author_articles) |> 
+  distinct(issn_l, src, oa_ta_first, oa_ta_corr) |> 
+  pivot_wider(names_from = src,
+              values_from = c("oa_ta_first", "oa_ta_corr")) |> 
+  janitor::clean_names() |>
+  pivot_longer(
+    cols = c(oa_ta_first_scopus, oa_ta_first_web_of_science, oa_ta_corr_scopus, oa_ta_corr_web_of_science)
+  ) |>
+  distinct(issn_l, oa_ta_first_hoad, name, value)
+
+ggplot(plot_uptake_jn_five_jns, aes(oa_ta_first_hoad, value, group = name))  +
+  geom_point(alpha = 0.05) +
+  scale_x_log10() +
+  scale_y_log10() +
+  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray50") +
+  coord_fixed() +
+  facet_grid(~name)
+```
+
+### By publisher
+
+```{r}
+uptake_publisher <- uptake_df |>
+  inner_join(esac, by = "issn_l") |>
+  filter(earliest_year != "2018")
+
+uptake_publisher_by_year <- uptake_publisher |>
+  mutate(earliest_year = as.character(earliest_year)) |>
+  group_by(earliest_year, src, publisher) |>
+  summarize(
+    n_articles = sum(n_articles),
+    n_oa_articles = sum(n_oa_articles),
+    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
+    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
+  )
+
+uptake_by_year_publisher_df <- uptake_publisher_by_year |>
+  # Remove HOAD because we only have first author data available
+  filter(src != "HOAD", earliest_year != "2018") |>
+  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
+
+hoad_publisher_panel <- uptake_publisher_by_year |>
+  filter(src == "HOAD") |>
+  ggplot(aes(x = earliest_year)) +
+  geom_bar(
+    aes(y = n_oa_articles / n_articles, fill = "All"),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  geom_bar(
+    aes(y = ta_oa_first_author_articles / n_articles, fill = "TA OA"),
+    stat = "identity",
+    color = "#000000"
+  ) +
+  geom_text(
+    aes(
+      label = paste0(
+        round(ta_oa_first_author_articles / n_oa_articles * 100, 1),
+        "%\n(",
+        format(ta_oa_first_author_articles, big.mark = ","),
+        ")"
+      ),
+      y = ta_oa_first_author_articles / n_articles),
+    position = position_dodge(width = 1),
+    vjust = -0.1,
+    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
+    size = 0.8 * my_base_size / .pt
+  ) + 
+  facet_grid(~fct_rev(publisher)) +
+  scale_fill_manual(
+    values = c("#b3b3b3a0", "#56B4E9"),
+    name = "",
+    guide = guide_legend(reverse = TRUE)
+  ) +
+   scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    limits = c(0, .4),
+    breaks = c(0, .1, .20, .3, .4)
+  ) +
+  theme_minimal(base_size = my_base_size) +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left"
+  )
+
+## label test
+
+first_au_publisher_panel <- uptake_by_year_publisher_df |>
+  distinct(earliest_year, publisher, n_oa_articles, n_articles, src, name, value) |>
+  # Modify factor levels to control order
+  mutate(
+    src = factor(src, levels = c("Web of Science", "Scopus")),
+    name = factor(name, 
+                 levels = c("ta_oa_first_author_articles", "ta_oa_corresponding_author_articles"))
+  ) |>
+  filter(name == "ta_oa_first_author_articles") |>
+  ggplot() +
+   geom_bar(
+    aes(
+      x = earliest_year,
+      y = n_oa_articles / n_articles,
+      fill = "All"
+    ),
+    stat = "identity",
+    color = "#000000a0",
+    size = .2
+  ) +
+  geom_bar(
+    aes(
+      x = earliest_year,
+      y = value / n_articles,
+      fill = "TA OA"
+    ),
+    stat = "identity",
+    color = "#000000a0",
+    size = .2
+  ) +
+   geom_text(
+    aes(
+      label = paste0(
+        round(value / n_oa_articles * 100, 1),
+        "%\n(",
+        format(value, big.mark = ","),
+        ")"
+      ),
+      x = earliest_year,
+      y = value / n_articles),
+    position = position_dodge(width = 1),
+    vjust = -0.2,
+    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
+    size = 0.7 * my_base_size / .pt
+  ) +
+  facet_grid(src ~ fct_rev(publisher)) +
+  scale_fill_manual(
+    values = c("#b3b3b3a0", "#56B4E9"),
+    name = "",
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    limits = c(0, .4),
+    breaks = c(0, .1, .20, .3, .4)
+  ) +
+  theme_minimal(base_size = my_base_size) +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left")
+
+cor_au_publisher_panel <- uptake_by_year_publisher_df |>
+  distinct(earliest_year, publisher, n_oa_articles, n_articles, src, name, value) |>
+  # Modify factor levels to control order
+  mutate(
+    src = factor(src, levels = c("Web of Science", "Scopus")),
+    name = factor(name, 
+                 levels = c("ta_oa_first_author_articles", "ta_oa_corresponding_author_articles"))
+  ) |>
+  filter(name == "ta_oa_corresponding_author_articles") |>
+  ggplot() +
+  geom_bar(
+    aes(
+      x = earliest_year,
+      y = n_oa_articles / n_articles,
+      fill = "All"
+    ),
+    stat = "identity",
+    color = "#000000",
+    size = .2
+  ) +
+  geom_bar(
+    aes(
+      x = earliest_year,
+      y = value / n_articles,
+      fill = "TA OA"
+    ),
+    stat = "identity",
+    color = "#000000",
+    size = .2
+  ) +
+   geom_text(
+    aes(
+      label = paste0(
+        round(value / n_oa_articles * 100, 1),
+        "%\n(",
+        format(value, big.mark = ","),
+        ")"
+      ),
+      x = earliest_year,
+      y = n_oa_articles / n_articles),
+    position = position_dodge(width = 1),
+    vjust = -0.1,
+    # https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/
+    size = 0.8 * my_base_size / .pt
+  ) +
+  facet_grid(src ~ fct_rev(publisher)) +
+  scale_fill_manual(
+    values = c("#b3b3b3a0", "#56B4E9"),
+    name = "",
+    guide = guide_legend(reverse = TRUE)
+  ) +
+  scale_y_continuous(
+    expand = expansion(mult = c(0, 0.05)),
+    labels = scales::percent_format(),
+    limits = c(0, .4),
+    breaks = c(0, .1, .20, .3, .4)
+  ) +
+  theme_minimal(base_size = my_base_size) +
+  labs(y = "Hybrid OA Uptake", x = NULL) +
+  theme(
+    legend.position = "top",
+    legend.justification = "left")
+```
+
+```{r, fig.asp=0.7}
+## Combine to patchwork
+layout <- "
+A
+B
+C
+"
+wrap_plots(
+  A = hoad_publisher_panel,
+  B = first_au_publisher_panel,
+  C = cor_au_publisher_panel,
+  D = guide_area(),
+  design = layout,
+  guides = "collect",
+  axes = "collect",
+  heights = c(1, 2, 2)
+) &
+  theme(
+    legend.position = "top",
+    legend.justification = "right",
+    legend.direction = "horizontal",
+ #   panel.grid.minor = element_blank(),
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title=element_text(size = 10)) &
+    plot_annotation(tag_levels = 'A')
+```
+
+- Elsevier gap suggests different oa information, let's make a scatterplot and contrast it hoad info using a shared corpus. due to publisher-specifc license
+
+ta by country
 

### 2024-11-28 analysis/results.Rmd
Safeguard

@@
 # Full OA journals detected through OA proportion > .95
-hoad_oa_excluded <- bq_table_download("subugoe-collaborative.hoaddata.cc_oa_prop") |>
+hoad_oa_excluded <- bq_table_download("hoa-article.hoaddata_sep24.cc_oa_prop") |>
   distinct(issn_l)
@@
 wos_active_jns_with_oa <- wos_active_jns |> 
-  filter(oa_status == "hybrid")
+  filter(grepl("hybrid", oa_status))
 
@@
 
-```{r, fig.asp=0.7}
+```{r, fig.asp=1, width = 12}
 ## Combine to patchwork
@@
 
-```
+### shared
+
+```{r}
+# Create a shared corpus via DOI
+wos_dois_active_jns_with_oa_core_df <- wos_active_jns |> 
+  filter(issn_l %in% wos_active_core_oa$issn_l) |>
+  filter(grepl("Article|^Review$", item_type)) |>
+  mutate(is_hybrid_wos = grepl("hybrid", oa_status)) |>
+  distinct(doi, is_hybrid_wos) |>
+  mutate(wos = TRUE)
+
+scp_dois_active_jns_with_oa_core_df <- scp_active_jns |> 
+  filter(issn_l %in% scp_active_core_oa$issn_l) |>
+  filter(grepl("^Article|^Review", item_type)) |>
+  mutate(is_hybrid_scp = grepl("hybrid", oa_status)) |>
+  distinct(doi, is_hybrid_scp) |>
+  mutate(scopus = TRUE)
+
+hoad_dois_active_jns_with_oa <- cr_df |> 
+  filter(issn_l %in% active_jns_with_oa$issn_l, between(cr_year, 2019, 2023)) |>
+  mutate(is_hybrid_hoad = doi %in% hoaddata::cc_articles$doi) |>
+  distinct(doi, issn_l, is_hybrid_hoad) |>
+  mutate(hoad = TRUE)
+
+## Shared corpus
+all_dois <- dplyr::full_join(hoad_dois_active_jns_with_oa, scp_dois_active_jns_with_oa_core_df, by = "doi") |>
+  dplyr::full_join(wos_dois_active_jns_with_oa_core_df, by = "doi") |>
+  mutate(across(where(is.logical), ~ replace_na(.x, FALSE))) |>
+  filter(!is.na(doi))
+
+shared_dois <- all_dois |>
+  filter(wos == TRUE, scopus == TRUE, hoad == TRUE)
+
+## Merge Publisher info
+
+shared_dois_df <- hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) |>
+  inner_join(shared_dois, by = "issn_l")
+
+esac_five_years <- shared_dois_df |>
+  group_by(esac_publisher) |>
+  summarise(hoad = sum(hoad == TRUE),
+            hoad_oa = sum(is_hybrid_hoad == TRUE),
+            wos = sum(wos == TRUE),
+            wos_oa = sum(is_hybrid_wos == TRUE),
+            scp = sum(scopus == TRUE),
+            scp_oa = sum(is_hybrid_scp == TRUE)) |>
+  mutate(hoad_oa_prop = hoad_oa / hoad,
+         wos_oa_prop = wos_oa / wos,
+         scp_oa_prop = scp_oa / scp) |>
+   mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           ))
+
+oa_plot_df <- esac_five_years |> 
+    pivot_longer(cols = contains(c("wos", "scp")))
+
+oa_absolut_plot <- oa_plot_df |>
+  filter(name %in% c("wos_oa", "scp_oa")) |>
+  mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )
+  ) |>
+  mutate(publisher =
+           forcats::fct_relevel(publisher, c(
+             "Elsevier", "Springer Nature", "Wiley", "Other"
+           ))) |>
+  ggplot(aes(x = hoad_oa, y = value, size = hoad)) +
+  geom_point(aes(fill = publisher), pch = 21) +
+  scale_x_continuous(limits = c(0, max(
+    c(
+      esac_five_years$scp_oa,
+      esac_five_years$hoad_oa,
+      esac_five_years$wos_oa
+    )
+  )),
+  labels = scales::number_format(big.mark = ",")) +
+  scale_y_continuous(limits = c(0, max(
+    c(
+      esac_five_years$scp_oa,
+      esac_five_years$hoad_oa,
+      esac_five_years$wos_oa
+    )
+  )),
+  labels = scales::number_format(big.mark = ",")) +
+  scale_fill_manual(
+    "",
+    values = c(
+      "Elsevier" = "#e9711c",
+      "Springer Nature" = "#486a7e",
+      "Wiley" = "#068853",
+      "Other" = "#b3b3b3"
+    )
+  ) +
+  facet_grid( ~ name) +
+  geom_abline(slope = 1,
+              intercept = 0,
+              color = "blue") +
+  coord_fixed() +
+  theme_minimal(my_base_size)
+
+oa_share_plot <- oa_plot_df |>
+  filter(name %in% c("wos_oa", "scp_oa")) |>
+   mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           )) |>
+  ggplot(aes(x = hoad_oa / hoad, y = value / hoad, size = hoad)) +
+  geom_point(aes(fill = publisher),  pch = 21) +
+  scale_x_continuous(limits = c(0, 1), labels = scales::percent_format()) +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent_format())+
+  scale_fill_manual(
+      "",
+      values = c(
+        "Elsevier" = "#e9711c",
+        "Springer Nature" = "#486a7e",
+        "Wiley" = "#068853",
+        "Other" = "#b3b3b3"
+      )
+    ) +
+  facet_grid(~name) +
+  geom_abline(slope = 1, intercept = 0, color = "blue") +
+  coord_fixed() +
+  theme_minimal(my_base_size)
+```
+
+```{r fig_oa_compare_scatter, fig.asp=1}
+layout <- "
+C
+A
+B
+"
+wrap_plots(
+  A = oa_absolut_plot,
+  B = oa_share_plot,
+#  C = guide_area(),
+  design = layout,
+  guides = "collect",
+  axes = "collect",
+  heights = c(0.5, 1, 1)
+) &
+  theme(
+   # legend.position = "top",
+  #  legend.justification = "right",
+ #   legend.direction = "horizontal",
+ #   panel.grid.minor = element_blank(),
+    panel.border = element_rect(color = "grey50", fill = NA),
+ #   axis.title=element_text(size = 10)
+ ) &
+    plot_annotation(tag_levels = 'A')
+```
+
+###
+
+crossref / unpaywall
+
+### TA countries
+
+replication database vs database
+validation shared corpus based on dois

### 2024-12-08 data-raw/indicators.R
Revised approach (align counting methods including varliable naming and filtering for publication years); re-generating indicators scp and wos

@@ -75,70 +75,69 @@ wos_by_year |>
   filter(between(earliest_year, 2018, 2023)) |>
   write_csv(here::here("data", "wos_by_year.csv"))
 #' ### by journal
-wos_jn_by_year <- dbGetQuery(bq_con, "-- Step 1: Extract relevant data with distinct item_id
-WITH items_by_year AS (
+wos_jn_by_year <- dbGetQuery(bq_con, "WITH
+  tt AS (
   SELECT
     DISTINCT wos.item_id,
-    COALESCE(CAST(REGEXP_EXTRACT(wos_pubdate_online, r'^[0-9]{4}') AS NUMERIC), pubyear) AS earliest_year,
-    REGEXP_CONTAINS(wos.oa_status, 'hybrid') AS is_oa,
-    REGEXP_CONTAINS(item_type, 'Article|^Review$') AS core,
+    COALESCE(CAST(REGEXP_EXTRACT(wos.wos_pubdate_online, r'^[0-9]{4}') AS NUMERIC), wos.pubyear) AS earliest_year,
+    CASE
+      WHEN REGEXP_CONTAINS(oa_status, 'hybrid') THEN wos.item_id
+      ELSE NULL
+  END
+    AS is_oa,
+    CASE
+      WHEN REGEXP_CONTAINS(item_type, 'Article|^Review$') THEN TRUE
+      ELSE FALSE
+  END
+    AS is_core,
     wos.issn_l
   FROM
-    `hoa-article.hoa_comparision.wos_jct_items` AS wos
-),
-
--- Step 2: Aggregate article counts by year and issn_l
-article_counts AS (
+    `hoa-article.hoa_comparision.wos_jct_items` AS wos),
+  -- Just core
+  core_pubs AS (
   SELECT
-    issn_l,
-    earliest_year,
-    COUNT(DISTINCT item_id) AS n_articles,
-    SUM(CASE WHEN core = TRUE THEN 1 ELSE 0 END) AS n_core_articles,
-    SUM(CASE WHEN is_oa = TRUE THEN 1 ELSE 0 END) AS n_oa_articles,
-    SUM(CASE WHEN is_oa = TRUE AND core = TRUE THEN 1 ELSE 0 END) AS n_core_oa_articles
-  FROM
-    items_by_year
-  GROUP BY
-    issn_l, earliest_year
-),
-
--- Step 3: Calculate OA percentages for all articles
-all_oa AS (
-  SELECT
[…]

### 2024-12-08 data-raw/indicators.R
Minor fix

@@ -59,7 +59,8 @@ wos_by_year <- dbGetQuery(bq_con, "WITH
     n_core_oa_articles / n_core_articles AS p
   FROM
     article_counts )
-SELECT
+SELECT *
+FROM (SELECT
   *
 FROM
   all_oa
@@ -69,11 +70,11 @@ SELECT
 FROM
   core_oa
 ORDER BY
-  earliest_year DESC")
+  earliest_year DESC )
+WHERE earliest_year BETWEEN 2019 AND 2023")
 #' Save results
-wos_by_year |>
-  filter(between(earliest_year, 2018, 2023)) |>
-  write_csv(here::here("data", "wos_by_year.csv"))
+write_csv(wos_by_year, here::here("data", "wos_by_year.csv"))
+
 #' ### by journal
 wos_jn_by_year <- dbGetQuery(bq_con, "WITH
   tt AS (
@@ -461,7 +462,7 @@ FROM
 ORDER BY
   earliest_year DESC")
 #' Save results
-write_csv(here::here(scp_by_year, "data", "scp_by_year.csv"))
+write_csv(scp_by_year, here::here("data", "scp_by_year.csv"))
 
 scp_jn_by_year <- dbGetQuery(bq_con, "WITH
   tt AS (
@@ -641,14 +642,12 @@ SELECT
   ta_oa_first_corresponding_author_articles
 FROM
   article_stats
-WHERE first_pubyear BEWTEEN 2019 AND 2023
+WHERE first_pubyear BETWEEN 2019 AND 2023
 ORDER BY
   first_pubyear DESC")
 
 #' Save results
-scp_ta_by_year |>
-  filter(between(first_pubyear, 2018, 2023)) |>
-  write_csv(here::here("data", "scp_ta_by_year.csv"))
+write_csv(scp_ta_by_year, here::here("data", "scp_ta_by_year.csv"))
 
 scp_ta_jn_by_year <- dbGetQuery(bq_con, "WITH
   jct_short AS (
@@ -771,7 +770,7 @@ ORDER BY
   first_pubyear DESC,
   issn_l")
 
-write_csv(first_pubyear, here::here("data", "scp_ta_jn_by_year.csv"))
+write_csv(scp_ta_jn_by_year, here::here("data", "scp_ta_jn_by_year.csv"))
 
 ### Country Affiliations
 

### 2024-12-08 data/README.Rmd
A data overview

---
output: github_document
---

```{r}
#| include: false
#| message: false
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```

## About

This folder contains analytical datasets providing aggregated counts on the uptake of open access in hybrid journals included in transformative agreements by journal and country information between 2019 and 2023. A particular focus was on whether publications were from eligible authors who could use such agreements. For this purpose, the affiliations of  first and corresponding authors were analysed.

The following datasets were used:

- Scopus (`scp`): In-house Scopus database provided by the Kompetenznetzwerk Bibliometrie (April 2024 snapshot)
- Web of Science (`wos`): In-house Web of Science database provided by the Kompetenznetzwerk Bibliometrie (April 2024 snapshot)

See `data-raw/` how the subset of publications associated with hybrid journals and transformative agreements was retrieved. 
The indicator were determined using SQL. `indicators.R` contains the database queries.

## Data files

### Overview

Articles in hybrid journals by year

|Variable | Description |
|-------- |-----------  |
|`earliest_year`|Year first published |
|`pub_type`|`core`: Article or Review; `all`: All articles regardless of item type |
|`n_articles` | Number of items |
|`n_oa_articles` | Number of items in the database with a hybrid flag, indicating open access availability under an open license |
|`p`| Open access uptake |

#### `wos_by_year`

Web of Science

```{r}
#| message: false
readr::read_csv(here("data", "wos_by_year.csv"))
```

#### `scp_by_year`

Scopus

```{r}
#| message: false
readr::read_csv(here("data", "scp_by_year.csv"))
```

### By hybrid journal

Articles in hybrid journals by year and hybrid journal included in a transformative agreement

|Variable | Description |
|-------- |-----------  |
|`issn_l` | Linking ISSN, a unique identifier for different journal ISSN according to the ISSN agency |
|`earliest_year`|Year first published |
|`pub_type`|`core`: Article or Review; `all`: All articles regardless of item type |
|`n` | Number of items |
|`oa_articles` | Number of items in the database with a hybrid flag, indicating open access availability under an open license |

#### `wos_jn_by_year`

Web of Science

```{r}
#| message: false
readr::read_csv(here("data", "wos_jn_by_year.csv"))
```

#### `scp_jn_by_year`

Scopus

```{r}
#| message: false
readr::read_csv(here("data", "scp_jn_by_year.csv"))
```

### Country affiliations

Indicators are provided for both first and corresponding authors. Because of multiple country affiliations, full counting was applied. Note that the bibliometric database did not always include the corresponding affiliation.

|Variable | Description |
|-------- |-----------  |
|`issn_l` | Linking ISSN, a unique identifier for different journal ISSN according to the ISSN agency |
|`earliest_year`|Year first published |
|`country_code`|ISO three letter country codes |
|`pub_type`|`core`: Article or Review; `all`: All articles regardless of item type |
|`n` | Number of items |
|`oa_articles` | Number of items in the database with a hybrid flag, indicating open access availability under an open license |
|`first_author_articles`|Number of items from first authors|
|`oa_first_author_articles`|Number of open access items (tagged as hybrid) from first authors|
|`corresponding_author_articles`|Number of items from corresponding authors|
|`oa_first_author_articles`|Number of open access items (tagged as hybrid) from corresponding authors|
|`first_corresponding_author_articles`|Number of items from first authors, which are also corresponding authors|
|`oa_first_corresponding_author_articles`|Number of open access items (tagged as hybrid) from first authors, which are also corresponding authors|

#### `wos_country_aff.csv`

Web of Science

```{r}
#| message: false
readr::read_csv(here("data", "wos_country_aff.csv"))
```

#### `scp_country_aff.csv`

Scopus

```{r}
#| message: false
readr::read_csv(here("data", "scp_country_aff.csv"))
```

### Transformative agreement data: Overview

The following data files contain indicators for journals where a link were established between articles and eligible institutions participating in transformative agreements. This could be achieved by linking author affiliations with agreement data.

Note that the earliest snapshot available with information about transformative agreemnt were from July 2021, while the analysis started in 2019.

|Variable | Description |
|-------- |-----------  |
|`earliest_year`|Year first published |
|`pub_type`|`core`: Article or Review; `all`: All articles regardless of item type |
|`ta_articles` | Number of items under an transformative agreement (first OR corresponding author)|
|`ta_oa_articles` | Number of items under an transformative agreement in the database with a hybrid flag, indicating open access availability under an open license (first OR corresponding author)|
|`ta_first_author_articles`|Number items from first authors under an transformative agreement |
|`ta_oa_first_author_articles`|Number of open access items (tagged as hybrid) from first authors under an transformative agreement|
|`ta_corresponding_author_articles`|Number of items from corresponding authors under an transformative agreement|
|`ta_oa_first_author_articles`|Number of open access items (tagged as hybrid) from corresponding authors under an transformative agreement|
|`ta_first_corresponding_author_articles`|Number of items from first authors under an transformative agreement, which are also corresponding authors|
|`ta_oa_first_corresponding_author_articles`|Number of open access items (tagged as hybrid) from first authors under an transformative agreement, which are also corresponding authors|

#### `wos_ta_by_year`

Web of Science

```{r}
#| message: false
readr::read_csv(here("data", "wos_ta_by_year.csv"))
```

#### `scp_ta_by_year`

Scopus

```{r}
#| message: false
readr::read_csv(here("data", "scp_ta_by_year.csv"))
```

### Transformative agreement data by hybrid journal

Articles in hybrid journals under transformative agreements by journal.

|Variable | Description |
|-------- |-----------  |
|`issn_l` | Linking ISSN, a unique identifier for different journal ISSN according to the ISSN agency |
|`earliest_year`|Year first published |
|`pub_type`|`core`: Article or Review; `all`: All articles regardless of item type |
|`ta_articles` | Number of items under an transformative agreement (first OR corresponding author)|
|`ta_oa_articles` | Number of items under an transformative agreement in the database with a hybrid flag, indicating open access availability under an open license (first OR corresponding author)|
|`ta_first_author_articles`|Number items from first authors under an transformative agreement |
|`ta_oa_first_author_articles`|Number of open access items (tagged as hybrid) from first authors under an transformative agreement|
|`ta_corresponding_author_articles`|Number of items from corresponding authors under an transformative agreement|
|`ta_oa_first_author_articles`|Number of open access items (tagged as hybrid) from corresponding authors under an transformative agreement|
|`ta_first_corresponding_author_articles`|Number of items from first authors under an transformative agreement, which are also corresponding authors|
|`ta_oa_first_corresponding_author_articles`|Number of open access items (tagged as hybrid) from first authors under an transformative agreement, which are also corresponding authors|

#### `wos_jn_ta_by_year`

Web of Science

```{r}
#| message: false
readr::read_csv(here("data", "wos_jn_ta_by_year.csv"))
```

#### `scp_ta_jn_by_year`

Scopus

```{r}
#| message: false
readr::read_csv(here("data", "scp_ta_jn_by_year.csv"))
```

### Transformative agreement data by country affiliations

Indicators are provided for both first and corresponding authors. Because of multiple country affiliations, full counting was applied. Note that the bibliometric database did not always include the corresponding affiliation.

|Variable | Description |
|-------- |-----------  |
|`issn_l` | Linking ISSN, a unique identifier for different journal ISSN according to the ISSN agency |
|`country_code`|ISO three letter country codes |
|`earliest_year`|Year first published |
|`pub_type`|`core`: Article or Review; `all`: All articles regardless of item type |
|`ta_articles` | Number of items under an transformative agreement (first OR corresponding author)|
|`ta_oa_articles` | Number of items under an transformative agreement in the database with a hybrid flag, indicating open access availability under an open license (first OR corresponding author)|
|`ta_first_author_articles`|Number items from first authors under an transformative agreement |
|`ta_oa_first_author_articles`|Number of open access items (tagged as hybrid) from first authors under an transformative agreement|
|`ta_corresponding_author_articles`|Number of items from corresponding authors under an transformative agreement|
|`ta_oa_first_author_articles`|Number of open access items (tagged as hybrid) from corresponding authors under an transformative agreement|
|`ta_first_corresponding_author_articles`|Number of items from first authors under an transformative agreement, which are also corresponding authors|
|`ta_oa_first_corresponding_author_articles`|Number of open access items (tagged as hybrid) from first authors under an transformative agreement, which are also corresponding authors|

#### `wos_jn_country_ta_by_year`

Web of Science

```{r}
#| message: false
readr::read_csv(here("data", "wos_jn_country_ta_by_year.csv"))
```

#### `wos_jn_country_ta_by_year`

Scopus

```{r}
#| message: false
readr::read_csv(here("data", "scp_jn_country_ta_by_year.csv"))
```

### 2024-12-09 data-raw/indicators.R
Add hoaddata indicators

@@ -1316,3 +1316,94 @@ ORDER BY
 
 write_csv(scp_jn_country_ta_by_year, here::here("data", "scp_jn_country_ta_by_year.csv"))
 
+#' hoadata
+
+## hoad jn by year
+
+hoad_jn_by_year <- hoaddata::jn_ind |>
+  mutate(cr_year = as.numeric(as.character(cr_year))) |>
+  filter(between(cr_year, 2019, 2023)) |>
+  distinct(issn_l, cr_year, jn_all)  |>
+  arrange(issn_l, cr_year)
+
+# hoad_country_aff 
+hoad_country_all <- hoaddata::jn_aff |>
+  filter(between(cr_year, 2019, 2023)) |>
+  distinct(issn_l, cr_year, country_code, articles_total) |>
+  rename(articles = articles_total)
+
+hoad_country_oa <- hoaddata::jn_aff |>
+  filter(between(cr_year, 2019, 2023), !is.na(cc)) |>
+  distinct(issn_l, cr_year, country_code, cc, articles_under_cc_variant) |>
+  group_by(issn_l, cr_year, country_code) |>
+  summarise(oa_articles = sum(articles_under_cc_variant))
+
+hoad_country_aff <- left_join(hoad_country_all, hoad_country_oa, by = c("issn_l", "cr_year", "country_code")) |>
+  mutate(across(where(is.numeric), ~replace_na(.x, 0))) |>
+  arrange(issn_l, cr_year, country_code)
+write_csv(hoad_country_aff, here::here("data", "hoad_country_aff.csv"))
+# hoad ta
+# Use BigQuery
+library(bigrquery)
+library(DBI)
+bq_con <- dbConnect(
+  bigrquery::bigquery(),
+  project = "subugoe-collaborative",
+  billing = "subugoe-collaborative"
+)
+hoad_ta_country_df <- dbGetQuery(bq_con, "SELECT
+  DISTINCT main.doi,
+  main.cr_year,
+  main.issn_l,
+  country_code,
+  CASE
+    WHEN CC IS NULL THEN FALSE
+    ELSE TRUE
+END
+  is_oa
+FROM
+  `hoa-article.hoaddata_sep24.ta_oa_inst` AS main
+LEFT JOIN
+  `subugoe-collaborative.openalex.institutions` AS oalex
+ON
+  main.ror = oalex.ror
+WHERE
+  ta_active = TRUE
+  AND cr_year BETWEEN 2019 AND 2023")
+
[…]

### 2024-12-09 data-raw/indicators.R
Add hoad jn indicators

@@ -1325,6 +1325,8 @@ hoad_jn_by_year <- hoaddata::jn_ind |>
   filter(between(cr_year, 2019, 2023)) |>
   distinct(issn_l, cr_year, jn_all)  |>
   arrange(issn_l, cr_year)
+write_csv(hoad_jn_by_year, here::here("data", "hoad_jn_by_year.csv"))
+
 
 # hoad_country_aff 
 hoad_country_all <- hoaddata::jn_aff |>

### 2025-01-03 data-raw/ta_matching.R
Don't TA match through country information in ESAC; not needed and issue with multinational consortia like CAUL

@@ -36,23 +36,6 @@ bigrquery::bq_dataset_query("hoa-article.hoa_comparision",
                                    destination_table = "hoa-article.hoa_comparision.jct_full",
                                    billing = "subugoe-collaborative")
 
-#' Upload ESAC data
-data_url <- "https://keeper.mpdl.mpg.de/f/7fbb5edd24ab4c5ca157/?dl=1"
-tmp <- tempfile()
-download.file(data_url, tmp)
-esac <- readxl::read_xlsx(tmp, skip = 2) |>
-  janitor::clean_names()
-esac_countries <- esac |>
-  select(id, country) |>
-  separate_rows(country, sep = ",") |>
-  mutate(country = trimws(country)) |>
-  mutate(country_code = countrycode::countrycode(country, origin = "country.name", dest = "iso3c")) |>
-  mutate(country_code = ifelse(country == "Kosovo", "XKK", country_code))
-#' Upload to BQ
-if (bigrquery::bq_table_exists("hoa-article.hoa_comparision.esac_countries")) 
-  bigrquery::bq_table_delete("hoa-article.hoa_comparision.esac_countries") 
-bigrquery::bq_table_upload("hoa-article.hoa_comparision.esac_countries", esac_countries)
-
 ## Web of Science
 
 #' Upload Web of Science article-level affilaition data and the matching table to Cloud Storage
@@ -132,12 +115,7 @@ wos_jct_match_sql <- "WITH
   END
     AS ta_flag
   FROM
-    wos_ror_jct
-  INNER JOIN
-    `hoa-article.hoa_comparision.esac_countries` AS esac
-  ON
-    wos_ror_jct.countrycode = esac.country_code
-    AND wos_ror_jct.esac_id = esac.id )
+    wos_ror_jct)
   -- Select only the articles that are identified as published under a transformative agreement
 SELECT
   DISTINCT *
@@ -156,7 +134,7 @@ bigrquery::bq_dataset_query("hoa-article.hoa_comparision",
 
 ## Scopus
 
-#' Upload Web of Science article-level affilaition data and the matching table to Cloud Storage
+#' Upload Web of Science article-level affiliation data and the matching table to Cloud Storage
 #'  and import to BigQuery
 #' `gcloud storage cp ~/Documents/thesis/hoa_validation/data-raw/scp_jct_affiliations.csv gs://bigschol`
[…]

### 2025-01-03 analysis/results.Rmd
Redo analysis,

@@
   mutate(src = "Web of Science") |>
-  select(-pub_type, -p)
+  select(-pub_type)
 
@@
   filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-  left_join(scp_jn_ta_by_year, by = c("issn_l", "earliest_year" = "first_pubyear", "pub_type" = "pubtype")) |>
+  left_join(scp_jn_ta_by_year, by = c("issn_l", "earliest_year", "pub_type" = "pubtype")) |>
   # Focus on core
@@
   mutate(src = "Scopus") |>
-  select(-pub_type, -p)
+  select(-pub_type)
 

### 2025-01-17 analysis/results.Rmd
Fix stat calculation

@@
 ```{r doi_coverage_check, eval = FALSE}
-jn_overlap <- jn_upset |> 
+jn_overlap <- jn_upset_articles |> 
   filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE)
@@
   mutate(cr_year = as.numeric(as.character(cr_year))) |>
-  distinct(issn_l, cr_year, n_articles = jn_all)
+  distinct(issn_l, cr_year, n = jn_all)
 hoad_oa_jn_by_year <- hoaddata::cc_articles |>
   group_by(issn_l, cr_year) |>
-  summarise(n_oa_articles = n_distinct(doi))
+  summarise(oa_articles = n_distinct(doi))
 
@@
   mutate(src = "HOAD") |>
-  distinct(issn_l, earliest_year = cr_year, n_oa_articles, n_articles, ta_articles = ta_first_author_articles, ta_first_author_articles, ta_oa_articles = ta_oa_first_author_articles, ta_oa_first_author_articles, src)
+  distinct(issn_l, earliest_year = cr_year, n, oa_articles, ta_articles = ta_first_author_articles, ta_first_author_articles, ta_oa_articles = ta_oa_first_author_articles, ta_oa_first_author_articles, src)
 
@@
   summarize(
-    n_articles = sum(n_articles),
-    n_oa_articles = sum(n_oa_articles),
+    n_articles = sum(n),
+    n_oa_articles = sum(oa_articles),
     ta_oa_first_author_articles = sum(ta_oa_first_author_articles),

### 2025-01-31 analysis/manuscript.Rmd
Overview table

---
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    template: apa6.sty
    extra_dependencies: ["flafter"]
csl: apa.csl
bibliography: references.bib
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  cache.lazy = FALSE,
  echo = FALSE,
  fig.width = 6,
  fig.asp = 0.618,
  out.width = "99%",
  fig.align = "center",
  dpi = 300,
  dev = c("png", "cairo_ps", "pdf"),
  fig.path = "fig/",
  fig.pos = "ht!", out.extra = ""
)
options(scipen = 999, digits = 2)
knitr::knit_hooks$set(
  inline = function(x) {
    if (is.numeric(x)) {
      return(prettyNum(x, big.mark = ","))
    } else{
      return(x)
    }
  }
)
```

```{r dependencies}
library(tidyverse, warn.conflicts = FALSE)
library(bigrquery)
library(kableExtra)
library(patchwork)
```

# Introduction

This study aims to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
To achieve this, a replication study was conducted by comparing results from hoaddata, an openly available and continuously updated dataset on hybrid open access uptake based on Crossref, OpenAlex, and the cOAlition S Journal Checker Tool, with the established bibliometric databases Web of Science and Scopus.

This study focuses on the coverage of hybrid journal portfolios included in transformative agreements between 2019 and 2023.
Special attention is given to potential differences in open access uptake by country when comparing first-author affiliation data to corresponding authorships.
This is crucial because the lack of publicly available invoicing data corresponding to authorships plays an essential role in determining whether an open-access article is supported through transformative agreements.
Data on corresponding authorships have been available on the Web of Science and Scopus for much longer than in open databases such as OpenAlex, where this information is still being roled out at the time of writing.
Because of this weakness, open approaches such as hoaddata and related research use first-authorship data instead.

By conducting a large-scale comparative analysis, this study aims to

1. Determine the strengths and weaknesses of using open data sources in monitoring the impact of transformative agreements on hybrid open access publishing.
2. Assess the coverage and accuracy of open data sources compared with established bibliometric databases.
3. Evaluate the reliability of first author affiliation data as a proxy for corresponding authorship in the context of open access uptake analysis.

# Background -- Evidence base to measure the effects of transformative agreements

## Anforderungen an das Monitoring

- esac guidelines 
- gemeinsamkeiten und unterschiede zu apc (listenpreise, tatsächliche zahlungen, zentrales invoicing, rabatte, waivers) 
- insitutionen covern cas, jedoch kann es zu unterschiedlichen verrechnugnsformen führen (antielig mit förderer, splitting innerhaklbd er einrichtung)

## Bibliometrische Evidenzen

- allgmeeiner uptake 
- wachstum apcs 
- wachstum verträge (konsortien, forschung) 
- konsequenzen

# Data and methods

The aim of this study is to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
To achieve this, a replication study is conducted by comparing results  from hoaddata, an openly available dataset on hybrid open access uptake, with the established bibliometric databases Web of Science and Scopus.
Replication studies, which can be found in the Life, Natural and Social Sciences, typically employ the same or comparable methods as the original study to determine whether consistent results can be obtained (https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/).

This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network for Bibliometrics (Kompetenznetzwerk Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using author roles (first and corresponding) and affiliation data are described.
Overall, xxxx hybrid journals from xxx agreements formed the basis of this study. This study covered journal articles published between 2019 and 2023.

## Dataset characteristics

### {hoaddata}

### Web of Science

### Scopus

## Data processing steps

### journal matching

### authorship records

###  Identifying eligible articles under transformative agreements

## Data records

```{r dataset_filtering, cache=TRUE}
## HOAD
hoad_jns <- hoaddata::jct_hybrid_jns |>
  filter(issn_l != "0027-8424")

# Full OA journals detected through OA proportion > .95
hoad_oa_excluded <- bq_table_download("hoa-article.hoaddata_sep24.cc_oa_prop") |>
  distinct(issn_l)

active_jns <- hoad_jns |>
  # Only active hybrid journals from JCT
  inner_join(hoaddata::jn_ind, by = "issn_l") |>
  filter(!cr_year %in% c("2017", "2018", "2024"))

active_jns_with_oa <- active_jns |> 
  filter(!is.na(cc))

# Get indicators for all articles without applying paratext adn supplement detection
hoad_all <- read_csv(here::here("data", "cr_all_ind.csv"))

## Web of Science
wos_jct_items_df <- readr::read_csv(here::here("data-raw", "wos_jct_items_df.csv"))

wos_hybrid_jns <- wos_jct_items_df |>
  # No PNAS
  filter(issn_l != "0027-8424") |>
  # No Full OA detected by OA prop
  filter(!issn_l %in% hoad_oa_excluded$issn_l)

wos_active_jns <- wos_hybrid_jns |>
  # Use online first, if available
  mutate(online_year = lubridate::year(wos_pubdate_online)) |>
  mutate(pub_year = ifelse(!is.na(online_year), online_year, pubyear)) |>
  filter(pub_year %in% 2019:2023)

wos_active_jns_with_oa <- wos_active_jns |> 
  filter(grepl("hybrid", oa_status))

# wos core 
wos_active_core <- wos_active_jns |>
  filter(grepl("Article|^Review$", item_type))

wos_active_core_oa <- wos_active_jns |>
   filter(grepl("hybrid", oa_status))

## Scopus

scp_jct_items_df <- readr::read_csv(here::here("data-raw", "scp_jct_items_df.csv"))

scp_hybrid_jns <- scp_jct_items_df |>
  # No PNAS
  filter(issn_l != "0027-8424") |>
  # No Full OA detected by OA prop
  filter(!issn_l %in% hoad_oa_excluded$issn_l)

scp_active_jns <- scp_hybrid_jns |>
   # Use online first, if available
  mutate(pub_year = ifelse(!is.na(first_pubyear), first_pubyear, pubyear)) |>
  filter(pub_year %in% 2019:2023)

scp_active_jns_with_oa <- scp_active_jns |> 
  filter(grepl("hybrid", oa_status))

# core
scp_active_core <- scp_active_jns |>
   filter(grepl("^Article|^Review", item_type))

scp_active_core_oa <- scp_active_core |>
    filter(grepl("hybrid", oa_status))

```

```{r journals_overview_table, cache=TRUE}
### Journals
ind_active_jns <- c(`hoad` = hoad_all$active_journals, 
                    `wos` = length(unique(wos_active_jns$issn_l)),
                    `scp` = length(unique(scp_active_jns$issn_l)))
ind_active_jns_core <- c(`hoad` = length(unique(active_jns$issn_l)), 
                    `wos` = length(unique(wos_active_core$issn_l)),
                    `scp` = length(unique(scp_active_core$issn_l)))
ind_active_jns_core_with_oa <-  c(`hoad` = length(unique(active_jns_with_oa$issn_l)), 
                    `wos` = length(unique(wos_active_core_oa$issn_l)),
                    `scp` = length(unique(scp_active_core_oa$issn_l)))

jns_ind <- bind_rows(ind_active_jns = ind_active_jns,
          ind_active_jns_core = ind_active_jns_core,
          ind_active_jns_core_with_oa = ind_active_jns_core_with_oa, .id = "indicator")
```

```{r article_overview_table, cache=TRUE}
### Article Volume

## hoad

# DOI set too large to track it with Git, see 001_data_gathering how the data
# was obtained from Google BQ
cr_df <- readr::read_csv(here::here("data-raw/hoad_dois_all_19_23.csv"))

ind_articles_hoad <- cr_df |>
  filter(issn_l %in% active_jns_with_oa$issn_l) |>
  distinct(doi) |>
  nrow()

## WoS
ind_articles_wos <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  distinct(item_id) |>
  nrow()

# core
ind_core_articles_wos <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(item_id) |>
  nrow()
# doi
ind_articles_wos_doi <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  distinct(doi) |>
  nrow()

ind_core_articles_wos_doi <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(doi) |>
  nrow()
## scopus
ind_articles_scp <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  distinct(item_id) |>
  nrow()
# core
ind_core_articles_scp <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("^Article|^Review", item_type)) |>
  distinct(item_id) |>
  nrow()
# doi
ind_articles_scp_doi <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  distinct(doi) |>
  nrow()

ind_core_articles_scp_doi <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(doi) |>
  nrow()

# Bringing it all together
ind_articles_total <- c(`hoad` = hoad_all$active_journals_all,
                        `wos` = ind_articles_wos,
                        `scp` = ind_articles_scp)
ind_articles_core <- c(`hoad` = ind_articles_hoad,
                        `wos` = ind_core_articles_wos,
                        `scp` = ind_core_articles_scp)
ind_articles_doi <- c(`hoad` = hoad_all$active_journals_all,
                        `wos` = ind_articles_wos_doi,
                        `scp` = ind_articles_scp_doi)
ind_articles_core_doi <- c(`hoad` = ind_articles_hoad,
                        `wos` = ind_core_articles_wos_doi,
                        `scp` = ind_core_articles_scp_doi)

articles_ind <- bind_rows(ind_articles_total = ind_articles_total,
                          ind_articles_core = ind_articles_core,
                          ind_articles_doi = ind_articles_doi,
                          ind_articles_core_doi = ind_articles_core_doi, .id = "indicator" )

## oa just core with doi

### hoad
ind_oa_articles_hoad <- hoaddata::cc_articles |>
  filter(between(cr_year, 2019, 2023)) |>
  filter(issn_l %in% active_jns$issn_l) |>
  distinct(doi) |>
  nrow()

### wos

ind_oa_articles_wos_all <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()

ind_oa_articles_wos_core <- wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()

### scp
ind_oa_articles_scp_all <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
   filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()
ind_oa_articles_scp_core <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  filter(grepl("hybrid", oa_status)) |>
  distinct(item_id) |>
  nrow()

ind_oa_articles_all <- c(`hoad` = hoad_all$ind_oa_articles_all,
                                             `wos` = ind_oa_articles_wos_all,
                                             `scp` = ind_oa_articles_scp_all
                     )
ind_oa_articles_core <-  c(`hoad` = ind_oa_articles_hoad,
                                             `wos` = ind_oa_articles_wos_core,
                                             `scp` = ind_oa_articles_scp_core
                     )

oa_articles_ind <- bind_rows(ind_oa_articles_all = ind_oa_articles_all,
                          ind_oa_articles_core = ind_oa_articles_core, .id = "indicator")
```

```{r ind_affiliations}

### affiliations

### WOS
wos_active_core_with_oa_item_ids <-  wos_active_jns |>
  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
  filter(grepl("Article|^Review$", item_type)) |>
  distinct(item_id, oa_status)

wos_aff <- readr::read_csv(here::here("data-raw", "wos_jct_affiliations.csv"))

wos_aff_df <- wos_aff |>
  inner_join(wos_active_core_with_oa_item_ids, by = "item_id")

wos_first_au_with_affiliation <- wos_aff_df |>
  filter(author_seq_nr == 1, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

wos_cor_au_with_affiliation <- wos_aff_df |>
  filter(corresponding == TRUE, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

# Scopus
scp_active_core_with_oa_item_ids <- scp_active_jns |>
  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
  filter(grepl("^Article|^Review", item_type)) |>
  distinct(item_id, oa_status)

scp_aff <-  readr::read_csv(here::here("data-raw", "scp_jct_affiliations.csv"))

scp_aff_df <- scp_aff |>
  inner_join(scp_active_core_with_oa_item_ids, by = "item_id")

scp_first_au_with_affiliation <- scp_aff_df |>
  filter(author_seq_nr == 1, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

scp_cor_au_with_affiliation <- scp_aff_df |>
  filter(corresponding == TRUE, !is.na(vendor_org_id)) |>
  distinct(item_id) |>
  nrow()

# hoad
aff_oalex_stats <- readr::read_csv(here::here("data", "aff_oalex_stats.csv"))

aff_ind <- bind_rows(
  first_aff = c(hoad = unlist(aff_oalex_stats$first_author_with_ror), 
                wos = wos_first_au_with_affiliation,
                scp = scp_first_au_with_affiliation),
  cor_aff = c(hoad = unlist(aff_oalex_stats$corresponding_authors_inst),
              wos = wos_cor_au_with_affiliation,
              scp = scp_cor_au_with_affiliation), .id = "indicator"
)

```

!<-- -- internal  -- external  - matching table institutions - journal-level data - country affiliation data  -->

```{r methods_overview_table, cache=TRUE}
methods_overview_table_df <- bind_rows(
  # Journals
  jns_ind,
  # Articles
  articles_ind,
  # OA
  oa_articles_ind,
  # Affiliation and Auhtor roles
  aff_ind
) |>
  # Human-readable indicator labels
  mutate(
    indicator = case_when(
      indicator == "ind_active_jns" ~ "Active journals",
      indicator == "ind_active_jns_core" ~ "Active journals (core)",
      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (core) with OA",
      indicator == "ind_articles_total" ~ "Total published articles ",
      indicator == "ind_articles_core" ~ "Core articles",
      indicator == "ind_articles_doi" ~ "Articles with DOI",
      indicator == "ind_articles_core_doi" ~ "Core articles with DOI",
      indicator == "ind_oa_articles_all" ~ "OA articles",
      indicator == "ind_oa_articles_core" ~ "Core OA articles",
      indicator == "first_aff" ~ "First author articles",
      indicator == "cor_aff" ~ "Corresponding author articles",
    )
  )

kableExtra::kbl(
  methods_overview_table_df,
  booktabs = TRUE, 
  format.args = list(big.mark = ","), 
  digits = 1, 
  escape = TRUE, 
  position = "H",
  caption ="Coverage of hybrid journals in transformative agreements 2019-23.",
  col.names = c("", "HOAD", "Web of Science", "Scopus")
) |>
  kable_styling(position = "center") |>
  row_spec(0, bold = TRUE) |>
  pack_rows("Hybrid journal metrics", 1, 3) |>
  pack_rows("Publication metrics", 4, 5) |>
  pack_rows("Digital Object Identifier (DOI) coverage", 6, 7) |>
  pack_rows("Open Access (OA) metrics", 8, 9) |>
  pack_rows("Core articles with affiliation data", 10, 11) 
```

# results

# discussion

### 2025-01-31 analysis/manuscript.tex
Overview table

\documentclass[a4paper,man,floatsintext,longtable,noextraspace,12pt]{apa6}

\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorinlistoftodos]{todonotes}
\usepackage{hyperref}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
% make captions italic

% number lines
% \usepackage{lineno}
% \linenumbers

% bibliography

% tightlist
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\shorttitle{}

\begin{document}
\thispagestyle{otherpage}

%\maketitle

% Change \title, \date, remove \author:
\begin{large}
\textbf{Replication: Hybrid Open Access in Transformative Agreements}
\end{large}

% Add below \maketitle:
\newcommand{\orcid}{%
  \begingroup\normalfont
  \includegraphics[height=6px]{orcid_logo.png}%
  \endgroup
}
Najko Jahn\textsuperscript{1}\textsuperscript{*} (\orcid{} \href{https://orcid.org/0000-0001-5105-1463}{\color{black}{0000-0001-5105-1463}}) \\

\textsuperscript{1} Göttingen State and University Library, University of Göttingen, Germany. \\

\textsuperscript{*} Correspondence: \href{mailto:najko.jahn@sub.uni-goettingen.de}{\color{black}{najko.jahn@sub.uni-goettingen.de}}

\section*{Abstract}
{}
{\textbf{Keywords}: hybrid open access, transformative agreements, scholarly publishing, big deals, bibliometrics}

\newpage

% QSS wants numbered sections
\setcounter{secnumdepth}{2}

\section{Introduction}\label{introduction}

This study aims to demonstrate the suitability of open scholarly data
sources for assessing the impact of transformative agreements on hybrid
open access. To achieve this, a replication study was conducted by
comparing results from hoaddata, an openly available and continuously
updated dataset on hybrid open access uptake based on Crossref,
OpenAlex, and the cOAlition S Journal Checker Tool, with the established
bibliometric databases Web of Science and Scopus.

This study focuses on the coverage of hybrid journal portfolios included
in transformative agreements between 2019 and 2023. Special attention is
given to potential differences in open access uptake by country when
comparing first-author affiliation data to corresponding authorships.
This is crucial because the lack of publicly available invoicing data
corresponding to authorships plays an essential role in determining
whether an open-access article is supported through transformative
agreements. Data on corresponding authorships have been available on the
Web of Science and Scopus for much longer than in open databases such as
OpenAlex, where this information is still being roled out at the time of
writing. Because of this weakness, open approaches such as hoaddata and
related research use first-authorship data instead.

By conducting a large-scale comparative analysis, this study aims to

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Determine the strengths and weaknesses of using open data sources in
  monitoring the impact of transformative agreements on hybrid open
  access publishing.
\item
  Assess the coverage and accuracy of open data sources compared with
  established bibliometric databases.
\item
  Evaluate the reliability of first author affiliation data as a proxy
  for corresponding authorship in the context of open access uptake
  analysis.
\end{enumerate}

\section{Background -- Evidence base to measure the effects of
transformative
agreements}\label{background-evidence-base-to-measure-the-effects-of-transformative-agreements}

\subsection{Anforderungen an das
Monitoring}\label{anforderungen-an-das-monitoring}

\begin{itemize}
\tightlist
\item
  esac guidelines
\item
  gemeinsamkeiten und unterschiede zu apc (listenpreise, tatsächliche
  zahlungen, zentrales invoicing, rabatte, waivers)
\item
  insitutionen covern cas, jedoch kann es zu unterschiedlichen
  verrechnugnsformen führen (antielig mit förderer, splitting
  innerhaklbd er einrichtung)
\end{itemize}

\subsection{Bibliometrische Evidenzen}\label{bibliometrische-evidenzen}

\begin{itemize}
\tightlist
\item
  allgmeeiner uptake
\item
  wachstum apcs
\item
  wachstum verträge (konsortien, forschung)
\item
  konsequenzen
\end{itemize}

\section{Data and methods}\label{data-and-methods}

The aim of this study is to demonstrate the suitability of open
scholarly data sources for assessing the impact of transformative
agreements on hybrid open access. To achieve this, a replication study
is conducted by comparing results from hoaddata, an openly available
dataset on hybrid open access uptake, with the established bibliometric
databases Web of Science and Scopus. Replication studies, which can be
found in the Life, Natural and Social Sciences, typically employ the
same or comparable methods as the original study to determine whether
consistent results can be obtained
(\url{https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/}).

This study compares results drawn from the openly available and
regularly updated hoaddata, an R package leveraging the open scholarly
data sources Crossref and OpenAlex, with the in-house database of the
German Competence Network for Bibliometrics (Kompetenznetzwerk
Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
After describing the initial data sources used, the necessary
pre-processing steps to obtain eligible articles from transformative
agreements using author roles (first and corresponding) and affiliation
data are described. Overall, xxxx hybrid journals from xxx agreements
formed the basis of this study. This study covered journal articles
published between 2019 and 2023.

\subsection{Dataset characteristics}\label{dataset-characteristics}

\subsubsection{\{hoaddata\}}\label{hoaddata}

\subsubsection{Web of Science}\label{web-of-science}

\subsubsection{Scopus}\label{scopus}

\subsection{Data processing steps}\label{data-processing-steps}

\subsubsection{journal matching}\label{journal-matching}

\subsubsection{authorship records}\label{authorship-records}

\subsubsection{Identifying eligible articles under transformative
agreements}\label{identifying-eligible-articles-under-transformative-agreements}

\subsection{Data records}\label{data-records}

!\textless-- -- internal -- external - matching table institutions -
journal-level data - country affiliation data --\textgreater{}

\begin{table}[H]
\centering
\caption{\label{tab:methods_overview_table}Coverage of hybrid journals in transformative agreements 2019-23.}
\centering
\begin{tabular}[t]{lrrr}
\toprule
\textbf{} & \textbf{HOAD} & \textbf{Web of Science} & \textbf{Scopus}\\
\midrule
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Hybrid journal metrics}}\\
\hspace{1em}Active journals & 12,890 & 8,655 & 11,888\\
\hspace{1em}Active journals (core) & 12,888 & 8,655 & 11,878\\
\hspace{1em}Active journals (core) with OA & 11,348 & 8,392 & 11,313\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Publication metrics}}\\
\hspace{1em}Total published articles & 9,740,015 & 8,616,053 & 8,117,644\\
\hspace{1em}Core articles & 8,158,425 & 6,708,083 & 7,317,703\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Digital Object Identifier (DOI) coverage}}\\
\hspace{1em}Articles with DOI & 9,740,015 & 7,713,796 & 8,105,112\\
\hspace{1em}Core articles with DOI & 8,158,425 & 6,695,661 & 7,314,327\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Open Access (OA) metrics}}\\
\hspace{1em}OA articles & 998,699 & 1,112,758 & 974,099\\
\hspace{1em}Core OA articles & 969,817 & 1,019,784 & 922,578\\
\addlinespace[0.3em]
\multicolumn{4}{l}{\textbf{Core articles with affiliation data}}\\
\hspace{1em}First author articles & 7,242,542 & 6,294,855 & 7,232,017\\
\hspace{1em}Corresponding author articles & 5,534,207 & 6,291,441 & 6,898,487\\
\bottomrule
\end{tabular}
\end{table}

\section{results}\label{results}

\section{discussion}\label{discussion}

\end{document}

### 2025-01-31 data-raw/affiliation_data_openalex.R

@@ -1,0 +1,251 @@
+# Check affiliations and corresponding authors in OpenAlex
+library(DBI)
+library(bigrquery)
+library(RPostgres)
+library(tidyverse)
+library(here)
+#' Connect to DB 
+
+#' KB Postgres (Web of Science, Scopus)
+kb_con <- dbConnect(RPostgres::Postgres(),
+                    host = "biblio-p-db03.fiz-karlsruhe.de",
+                    port = 6432,
+                    dbname = "kbprod",
+                    user =  Sys.getenv("kb_user"),
+                    password = Sys.getenv("kb_pwd"),
+                    bigint = "numeric"
+)
+cr_dois <- readr::read_csv(here("data-raw", "hoad_dois_all_19_23.csv"))
+
+#' HOAD active journals with OA
+hoad_jns <- hoaddata::jct_hybrid_jns |>
+  filter(issn_l != "0027-8424")
+
+active_jns <- hoad_jns |>
+  # Only active hybrid journals from JCT
+  inner_join(hoaddata::jn_ind, by = "issn_l") |>
+  filter(!cr_year %in% c("2017", "2018", "2024"))
+
+active_jns_with_oa <- active_jns |> 
+  filter(!is.na(cc))
+
+dois_active_jns_with_oa <- cr_dois |>
+  filter(issn_l %in% active_jns_with_oa$issn_l)
+
+#' Send to FIZ
+DBI::dbWriteTable(kb_con, "dois_active_jns_with_oa", dois_active_jns_with_oa)
+
+#' Create table with author roles, affiliations and funding stats
+dbExecute(kb_con, "CREATE TABLE oalex_cr_cor_funder AS WITH filtered_works AS (
+  SELECT 
+    id,
+    REPLACE(doi, 'https://doi.org/', '') as short_doi,
+    corresponding_author_ids, 
+    corresponding_institution_ids,
+    json_array_length(corresponding_author_ids) AS num_corresponding_authors,
+    json_array_length(corresponding_institution_ids) AS num_corresponding_institutions
+  FROM 
+    fiz_openalex_rep_20240831_openbib_fdw.works
+)
+SELECT 
+    hoad.doi, 
+    w.corresponding_author_ids, 
+    w.corresponding_institution_ids,
[…]

### 2025-02-05 analysis/manuscript.Rmd
Describe data sources used

@@
 The aim of this study is to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
-To achieve this, a replication study is conducted by comparing results  from hoaddata, an openly available dataset on hybrid open access uptake, with the established bibliometric databases Web of Science and Scopus.
-Replication studies, which can be found in the Life, Natural and Social Sciences, typically employ the same or comparable methods as the original study to determine whether consistent results can be obtained (https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/).
+To achieve this, results from hoaddata, an openly available collection of open research information regarding hybrid open access, was compared with the established bibliometric databases Web of Science and Scopus.
+After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using open access evidence, author roles (first and corresponding) and affiliation data are presented.
+Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study.
+
+<!-- Replication studies, which can be found in the Life, Natural and Social Sciences, typically employ the same or comparable methods as the original study to determine whether consistent results can be obtained (https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/).
 
 This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network for Bibliometrics (Kompetenznetzwerk Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
-After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using author roles (first and corresponding) and affiliation data are described.
-Overall, xxxx hybrid journals from xxx agreements formed the basis of this study. This study covered journal articles published between 2019 and 2023.
+After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using author roles (first and corresponding) and affiliation data are presented.
+Overall, xxxx hybrid journals from xxx agreements formed the basis of this study. This study covered journal articles published between 2019 and 2023.   -->
 
-## Dataset characteristics
+## Data sources
 
-### {hoaddata}
+### hoaddata
+
+hoaddata, developed and maintained by the author, is an R data package comprising information about the uptake of hybrid open access since 2017 from several openly available data sources.
+It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.
+
+More specifically, hoaddata uses Crossref, a DOI registration agency, for obtaining journal publication volume and open access status through
+Creative Commons licence information relative to the published version ("version of record").
+Because of limited affiliation metadata in Crossref (https://doi.org/10.31222/osf.io/smxe5), hoaddata sources  first-author affiliations from OpenAlex.
+While the country in which a first-author was located was used to aggregate country-level statistics, Research Organization Registry (ROR) identifiers (ROR-ID) were used in conjunction with ESAC registry information on the duration of an agreement was used to estimate whether an article was published under active transformative agreements.
+This matching benefited from the availability of the ROR-IDs in both sources.
+To improve the matching, JCT data were enriched to include associated institutions, such as unviersity hospitals.
+
+hoaddata follows good practices for computational reproducibility using R.
+The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
+To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow interfaces with the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
+The package has been regularly updated since 2022 and the version including the computation log is available on GitHub.
+
+hoaddata is used as a data basis of the Hybrid Open Access Dashboard, a data analytics services for  library consortia and publishers to track the uptake of hybrid open access through transformative agreements.
+It was also used in bibliometric research (Jahn 2025).
 
@@
 
+Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020]. 
+The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).
+
+The WoS provides important data points for studying spending on open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as Openalex.
+However, it lacks information about journals and articles under transformative agreements.
+
+For programmatic access to article-level data, the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany is used to access bibliometric data.
+The KB processes raw XML data provided by Clarivate Analytics, which is provided as an in-house PostgreSQL database under a uniform schema.
+To support reproducibility, KB maintains annual snapshots of the database. 
+Accordingly, this study used the annual snapshot from April 2024, which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].
+
 ### Scopus
+
+Elsevier's Scopus, launched in 2004, is another widely used proprietary bibliometric database for measuring research [@Baas_2020].
+Similar to the Web of Science, Scopus is selective with regard to the journals it indexes.
+However, its coverage is substantially more extensive than that of the Web of Science Core collection [@Singh_2021; @Visser_2021].
+With detailed metadata about article types, open access status information derived from Unpaywall, author roles, and disambiguated affiliations, Scopus also contains important data to assess open access uptake, although no direct information regarding transformative agreements was available at the time of the study.
+
+This study used the Scopus annual snapshot of April 2024 as provided by the KB. The same KB curation effort was applied to the  Scopus raw data as for the Web of Science [@schmidt_2024_13935407].
 
@@
 
-```{r dataset_filtering, cache=TRUE}
-## HOAD
-hoad_jns <- hoaddata::jct_hybrid_jns |>
-  filter(issn_l != "0027-8424")
-
-# Full OA journals detected through OA proportion > .95
-hoad_oa_excluded <- bq_table_download("hoa-article.hoaddata_sep24.cc_oa_prop") |>
-  distinct(issn_l)
-
-active_jns <- hoad_jns |>
-  # Only active hybrid journals from JCT
-  inner_join(hoaddata::jn_ind, by = "issn_l") |>
-  filter(!cr_year %in% c("2017", "2018", "2024"))
-
-active_jns_with_oa <- active_jns |> 
-  filter(!is.na(cc))
-
-# Get indicators for all articles without applying paratext adn supplement detection
-hoad_all <- read_csv(here::here("data", "cr_all_ind.csv"))
-
-## Web of Science
-wos_jct_items_df <- readr::read_csv(here::here("data-raw", "wos_jct_items_df.csv"))
-
-wos_hybrid_jns <- wos_jct_items_df |>
-  # No PNAS
-  filter(issn_l != "0027-8424") |>
-  # No Full OA detected by OA prop
-  filter(!issn_l %in% hoad_oa_excluded$issn_l)
-
-wos_active_jns <- wos_hybrid_jns |>
-  # Use online first, if available
-  mutate(online_year = lubridate::year(wos_pubdate_online)) |>
-  mutate(pub_year = ifelse(!is.na(online_year), online_year, pubyear)) |>
-  filter(pub_year %in% 2019:2023)
-
-wos_active_jns_with_oa <- wos_active_jns |> 
-  filter(grepl("hybrid", oa_status))
-
-# wos core 
-wos_active_core <- wos_active_jns |>
-  filter(grepl("Article|^Review$", item_type))
-
-wos_active_core_oa <- wos_active_jns |>
-   filter(grepl("hybrid", oa_status))
-
-## Scopus
-
-scp_jct_items_df <- readr::read_csv(here::here("data-raw", "scp_jct_items_df.csv"))
-
-scp_hybrid_jns <- scp_jct_items_df |>
-  # No PNAS
-  filter(issn_l != "0027-8424") |>
-  # No Full OA detected by OA prop
-  filter(!issn_l %in% hoad_oa_excluded$issn_l)
-
-scp_active_jns <- scp_hybrid_jns |>
-   # Use online first, if available
-  mutate(pub_year = ifelse(!is.na(first_pubyear), first_pubyear, pubyear)) |>
-  filter(pub_year %in% 2019:2023)
-
-scp_active_jns_with_oa <- scp_active_jns |> 
-  filter(grepl("hybrid", oa_status))
-
-# core
-scp_active_core <- scp_active_jns |>
-   filter(grepl("^Article|^Review", item_type))
-
-scp_active_core_oa <- scp_active_core |>
-    filter(grepl("hybrid", oa_status))
-
-```
-
-```{r journals_overview_table, cache=TRUE}
-### Journals
-ind_active_jns <- c(`hoad` = hoad_all$active_journals, 
-                    `wos` = length(unique(wos_active_jns$issn_l)),
-                    `scp` = length(unique(scp_active_jns$issn_l)))
-ind_active_jns_core <- c(`hoad` = length(unique(active_jns$issn_l)), 
-                    `wos` = length(unique(wos_active_core$issn_l)),
-                    `scp` = length(unique(scp_active_core$issn_l)))
-ind_active_jns_core_with_oa <-  c(`hoad` = length(unique(active_jns_with_oa$issn_l)), 
-                    `wos` = length(unique(wos_active_core_oa$issn_l)),
-                    `scp` = length(unique(scp_active_core_oa$issn_l)))
-
-jns_ind <- bind_rows(ind_active_jns = ind_active_jns,
-          ind_active_jns_core = ind_active_jns_core,
-          ind_active_jns_core_with_oa = ind_active_jns_core_with_oa, .id = "indicator")
-```
-
-```{r article_overview_table, cache=TRUE}
-### Article Volume
-
-## hoad
-
-# DOI set too large to track it with Git, see 001_data_gathering how the data
-# was obtained from Google BQ
-cr_df <- readr::read_csv(here::here("data-raw/hoad_dois_all_19_23.csv"))
-
-ind_articles_hoad <- cr_df |>
-  filter(issn_l %in% active_jns_with_oa$issn_l) |>
-  distinct(doi) |>
-  nrow()
-
-## WoS
-ind_articles_wos <- wos_active_jns |>
-  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
-  distinct(item_id) |>
-  nrow()
-
-# core
-ind_core_articles_wos <- wos_active_jns |>
-  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
-  filter(grepl("Article|^Review$", item_type)) |>
-  distinct(item_id) |>
-  nrow()
-# doi
-ind_articles_wos_doi <- wos_active_jns |>
-  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
-  distinct(doi) |>
-  nrow()
-
-ind_core_articles_wos_doi <- wos_active_jns |>
-  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
-  filter(grepl("Article|^Review$", item_type)) |>
-  distinct(doi) |>
-  nrow()
-## scopus
-ind_articles_scp <- scp_active_jns |>
-  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-  distinct(item_id) |>
-  nrow()
-# core
-ind_core_articles_scp <- scp_active_jns |>
-  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-  filter(grepl("^Article|^Review", item_type)) |>
-  distinct(item_id) |>
-  nrow()
-# doi
-ind_articles_scp_doi <- scp_active_jns |>
-  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-  distinct(doi) |>
-  nrow()
-
-ind_core_articles_scp_doi <- scp_active_jns |>
-  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-  filter(grepl("Article|^Review$", item_type)) |>
-  distinct(doi) |>
-  nrow()
-
-# Bringing it all together
-ind_articles_total <- c(`hoad` = hoad_all$active_journals_all,
-                        `wos` = ind_articles_wos,
-                        `scp` = ind_articles_scp)
-ind_articles_core <- c(`hoad` = ind_articles_hoad,
-                        `wos` = ind_core_articles_wos,
-                        `scp` = ind_core_articles_scp)
-ind_articles_doi <- c(`hoad` = hoad_all$active_journals_all,
-                        `wos` = ind_articles_wos_doi,
-                        `scp` = ind_articles_scp_doi)
-ind_articles_core_doi <- c(`hoad` = ind_articles_hoad,
-                        `wos` = ind_core_articles_wos_doi,
-                        `scp` = ind_core_articles_scp_doi)
-
-articles_ind <- bind_rows(ind_articles_total = ind_articles_total,
-                          ind_articles_core = ind_articles_core,
-                          ind_articles_doi = ind_articles_doi,
-                          ind_articles_core_doi = ind_articles_core_doi, .id = "indicator" )
-
-## oa just core with doi
-
-### hoad
-ind_oa_articles_hoad <- hoaddata::cc_articles |>
-  filter(between(cr_year, 2019, 2023)) |>
-  filter(issn_l %in% active_jns$issn_l) |>
-  distinct(doi) |>
-  nrow()
-
-### wos
-
-ind_oa_articles_wos_all <- wos_active_jns |>
-  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
-  filter(grepl("hybrid", oa_status)) |>
-  distinct(item_id) |>
-  nrow()
-
-ind_oa_articles_wos_core <- wos_active_jns |>
-  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
-  filter(grepl("Article|^Review$", item_type)) |>
-  filter(grepl("hybrid", oa_status)) |>
-  distinct(item_id) |>
-  nrow()
-
-### scp
-ind_oa_articles_scp_all <- scp_active_jns |>
-  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-   filter(grepl("hybrid", oa_status)) |>
-  distinct(item_id) |>
-  nrow()
-ind_oa_articles_scp_core <- scp_active_jns |>
-  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-  filter(grepl("Article|^Review$", item_type)) |>
-  filter(grepl("hybrid", oa_status)) |>
-  distinct(item_id) |>
-  nrow()
-
-ind_oa_articles_all <- c(`hoad` = hoad_all$ind_oa_articles_all,
-                                             `wos` = ind_oa_articles_wos_all,
-                                             `scp` = ind_oa_articles_scp_all
-                     )
-ind_oa_articles_core <-  c(`hoad` = ind_oa_articles_hoad,
-                                             `wos` = ind_oa_articles_wos_core,
-                                             `scp` = ind_oa_articles_scp_core
-                     )
-
-oa_articles_ind <- bind_rows(ind_oa_articles_all = ind_oa_articles_all,
-                          ind_oa_articles_core = ind_oa_articles_core, .id = "indicator")
-```
-
-```{r ind_affiliations}
-
-### affiliations
-
-### WOS
-wos_active_core_with_oa_item_ids <-  wos_active_jns |>
-  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
-  filter(grepl("Article|^Review$", item_type)) |>
-  distinct(item_id, oa_status)
-
-wos_aff <- readr::read_csv(here::here("data-raw", "wos_jct_affiliations.csv"))
-
-wos_aff_df <- wos_aff |>
-  inner_join(wos_active_core_with_oa_item_ids, by = "item_id")
-
-wos_first_au_with_affiliation <- wos_aff_df |>
-  filter(author_seq_nr == 1, !is.na(vendor_org_id)) |>
-  distinct(item_id) |>
-  nrow()
-
-wos_cor_au_with_affiliation <- wos_aff_df |>
-  filter(corresponding == TRUE, !is.na(vendor_org_id)) |>
-  distinct(item_id) |>
-  nrow()
-
-# Scopus
-scp_active_core_with_oa_item_ids <- scp_active_jns |>
-  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
-  filter(grepl("^Article|^Review", item_type)) |>
-  distinct(item_id, oa_status)
-
-scp_aff <-  readr::read_csv(here::here("data-raw", "scp_jct_affiliations.csv"))
-
-scp_aff_df <- scp_aff |>
-  inner_join(scp_active_core_with_oa_item_ids, by = "item_id")
-
-scp_first_au_with_affiliation <- scp_aff_df |>
-  filter(author_seq_nr == 1, !is.na(vendor_org_id)) |>
-  distinct(item_id) |>
-  nrow()
-
-scp_cor_au_with_affiliation <- scp_aff_df |>
-  filter(corresponding == TRUE, !is.na(vendor_org_id)) |>
-  distinct(item_id) |>
-  nrow()
-
-# hoad
-aff_oalex_stats <- readr::read_csv(here::here("data", "aff_oalex_stats.csv"))
-
-aff_ind <- bind_rows(
-  first_aff = c(hoad = unlist(aff_oalex_stats$first_author_with_ror), 
-                wos = wos_first_au_with_affiliation,
-                scp = scp_first_au_with_affiliation),
-  cor_aff = c(hoad = unlist(aff_oalex_stats$corresponding_authors_inst),
-              wos = wos_cor_au_with_affiliation,
-              scp = scp_cor_au_with_affiliation), .id = "indicator"
-)
-
-```
-
-!<-- -- internal  -- external  - matching table institutions - journal-level data - country affiliation data  -->
-
 ```{r methods_overview_table, cache=TRUE}
+load(here::here("data", "manuscript_data.Rda"))
 methods_overview_table_df <- bind_rows(

### 2025-02-05 analysis/manuscript.tex
Describe data sources used

@@
 % bibliography
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
+\makeatletter
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+  {\begin{list}{}{%
+   \setlength{\itemindent}{0pt}
+   \setlength{\leftmargin}{0pt}
+   \setlength{\parsep}{0pt}
+   % turn on hanging indent if param 1 is 1
+   \ifodd #1
+    \setlength{\leftmargin}{\cslhangindent}
+    \setlength{\itemindent}{-1\cslhangindent}
+   \fi
+   % set entry spacing
+   \setlength{\itemsep}{#2\baselineskip}}}
+  {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 scholarly data sources for assessing the impact of transformative
-agreements on hybrid open access. To achieve this, a replication study
-is conducted by comparing results from hoaddata, an openly available
-dataset on hybrid open access uptake, with the established bibliometric
-databases Web of Science and Scopus. Replication studies, which can be
-found in the Life, Natural and Social Sciences, typically employ the
-same or comparable methods as the original study to determine whether
-consistent results can be obtained
-(\url{https://pmc.ncbi.nlm.nih.gov/articles/PMC10019630/}).
+agreements on hybrid open access. To achieve this, results from
+hoaddata, an openly available collection of open research information
+regarding hybrid open access, was compared with the established
+bibliometric databases Web of Science and Scopus. After describing the
+initial data sources used, the necessary pre-processing steps to obtain
+eligible articles from transformative agreements using open access
+evidence, author roles (first and corresponding) and affiliation data
+are presented. Overall, xxxx hybrid journals from xxx agreements that
+published at least one open access article between 2019 and 2023. formed
+the basis of this study.
 
-This study compares results drawn from the openly available and
-regularly updated hoaddata, an R package leveraging the open scholarly
-data sources Crossref and OpenAlex, with the in-house database of the
-German Competence Network for Bibliometrics (Kompetenznetzwerk
-Bibliometrie, KB), which comprises Web of Science and Scopus snapshots.
-After describing the initial data sources used, the necessary
-pre-processing steps to obtain eligible articles from transformative
-agreements using author roles (first and corresponding) and affiliation
-data are described. Overall, xxxx hybrid journals from xxx agreements
-formed the basis of this study. This study covered journal articles
-published between 2019 and 2023.
+\subsection{Data sources}\label{data-sources}
 
-\subsection{Dataset characteristics}\label{dataset-characteristics}
+\subsubsection{hoaddata}\label{hoaddata}
 
-\subsubsection{\{hoaddata\}}\label{hoaddata}
+hoaddata, developed and maintained by the author, is an R data package
+comprising information about the uptake of hybrid open access since 2017
+from several openly available data sources. It combines article-level
+metadata from Crossref and OpenAlex with transformative agreement
+information from the cOAlition S Journal Checker Tool (JCT), which links
+journal and institutional data to agreements in the ESAC registry.
+
+More specifically, hoaddata uses Crossref, a DOI registration agency,
+for obtaining journal publication volume and open access status through
+Creative Commons licence information relative to the published version
+(``version of record''). Because of limited affiliation metadata in
+Crossref (\url{https://doi.org/10.31222/osf.io/smxe5}), hoaddata sources
+first-author affiliations from OpenAlex. While the country in which a
+first-author was located was used to aggregate country-level statistics,
+Research Organization Registry (ROR) identifiers (ROR-ID) were used in
+conjunction with ESAC registry information on the duration of an
+agreement was used to estimate whether an article was published under
+active transformative agreements. This matching benefited from the
+availability of the ROR-IDs in both sources. To improve the matching,
+JCT data were enriched to include associated institutions, such as
+unviersity hospitals.
+
+hoaddata follows good practices for computational reproducibility using
+R. The package, which includes data, code, a test suite and
+documentation, is openly available on GitHub. To ensure computational
+reproducibility while aggregating the data, a GitHub Actions continuous
+integration and delivery (CI/CD) workflow interfaces with the SUB
+Göttingen's open scholarly data warehouse based on Google BigQuery,
+which provides high-performant programmatic access to monthly snapshots
+of Crossref and OpenAlex. The package has been regularly updated since
+2022 and the version including the computation log is available on
+GitHub.
+
+hoaddata is used as a data basis of the Hybrid Open Access Dashboard, a
+data analytics services for library consortia and publishers to track
+the uptake of hybrid open access through transformative agreements. It
+was also used in bibliometric research (Jahn 2025).
 
@@
 
+Clarivate Analytics' Web of Science (WoS) is a well-established
+proprietary bibliometric database consisting of several collections
+(Birkle et al., 2020). The collections considered in this study were the
+Science Citation Index Expanded (SCIE), the Social Sciences Citation
+Index (SSCI) and the Arts \& Humanities Citation Index (AHCI).
+
+The WoS provides important data points for studying spending on open
+access: author affiliations and roles, differentiation of journal
+articles into document types representing different types of journal
+contributions, such as original articles or reviews, and open access
+status information derived from OurResearch's Unpaywall, the same
+provider as Openalex. However, it lacks information about journals and
+articles under transformative agreements.
+
+For programmatic access to article-level data, the database of the
+Kompetenznetzwerk Bibliometrie (KB) in Germany is used to access
+bibliometric data. The KB processes raw XML data provided by Clarivate
+Analytics, which is provided as an in-house PostgreSQL database under a
+uniform schema. To support reproducibility, KB maintains annual
+snapshots of the database. Accordingly, this study used the annual
+snapshot from April 2024, which is considered to cover almost the entire
+previous publication year (Schmidt et al., 2024).
+
 \subsubsection{Scopus}\label{scopus}
+
+Elsevier's Scopus, launched in 2004, is another widely used proprietary
+bibliometric database for measuring research (Baas et al., 2020).
+Similar to the Web of Science, Scopus is selective with regard to the
+journals it indexes. However, its coverage is substantially more
+extensive than that of the Web of Science Core collection (Singh et al.,
+2021; Visser et al., 2021). With detailed metadata about article types,
+open access status information derived from Unpaywall, author roles, and
+disambiguated affiliations, Scopus also contains important data to
+assess open access uptake, although no direct information regarding
+transformative agreements was available at the time of the study.
+
+This study used the Scopus annual snapshot of April 2024 as provided by
+the KB. The same KB curation effort was applied to the Scopus raw data
+as for the Web of Science (Schmidt et al., 2024).
 
@@
 \subsection{Data records}\label{data-records}
-
-!\textless-- -- internal -- external - matching table institutions -
-journal-level data - country affiliation data --\textgreater{}
 
@@
 
-\section{discussion}\label{discussion}
+\section*{discussion}\label{discussion}
+\addcontentsline{toc}{section}{discussion}
+
+\phantomsection\label{refs}
+\begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-Baas_2020}
+Baas, J., Schotten, M., Plume, A., Côté, G., \& Karimi, R. (2020).
+Scopus as a curated, high-quality bibliometric data source for academic
+research in quantitative science studies. \emph{Quantitative Science
+Studies}, \emph{1}(1), 377--386.
+\url{https://doi.org/10.1162/qss_a_00019}
+
+\bibitem[\citeproctext]{ref-Birkle_2020}
+Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). Web of
+science as a data source for research on scientific and scholarly
+activity. \emph{Quantitative Science Studies}, \emph{1}(1), 363--376.
+\url{https://doi.org/10.1162/qss_a_00018}
+
+\bibitem[\citeproctext]{ref-schmidt_2024_13935407}
+Schmidt, M., Rimmert, C., Stephen, D., Lenke, C., Donner, P., Gärtner,
+S., Taubert, N., Bausenwein, T., \& Stahlschmidt, S. (2024). \emph{The
+data infrastructure of the {German Kompetenznetzwerk Bibliometrie}: An
+enabling intermediary between raw data and analysis}. Zenodo.
+\url{https://doi.org/10.5281/zenodo.13935407}
+
+\bibitem[\citeproctext]{ref-Singh_2021}
+Singh, V. K., Singh, P., Karmakar, M., Leta, J., \& Mayr, P. (2021). The
+journal coverage of web of science, scopus and dimensions: A comparative
+analysis. \emph{Scientometrics}, \emph{126}(6), 5113--5142.
+\url{https://doi.org/10.1007/s11192-021-03948-5}
+
+\bibitem[\citeproctext]{ref-Visser_2021}
+Visser, M., Eck, N. J. van, \& Waltman, L. (2021). Large-scale
+comparison of bibliographic data sources: {Scopus, Web of Science,
+Dimensions, Crossref, and Microsoft Academic}. \emph{Quantitative
+Science Studies}, \emph{2}(1), 20--41.
+\url{https://doi.org/10.1162/qss_a_00112}
+
+\end{CSLReferences}
 

### 2025-02-05 analysis/manuscript_preprocessing_data.R

@@ -1,0 +1,290 @@
+### Data preprocessing for the manuscript
+
+## Why? It just takes too long to render the manuscript.
+
+library(tidyverse, warn.conflicts = FALSE)
+library(bigrquery)
+library(here)
+
+## HOAD
+hoad_jns <- hoaddata::jct_hybrid_jns |>
+  filter(issn_l != "0027-8424")
+
+# Full OA journals detected through OA proportion > .95
+hoad_oa_excluded <- bq_table_download("hoa-article.hoaddata_sep24.cc_oa_prop") |>
+  distinct(issn_l)
+
+active_jns <- hoad_jns |>
+  # Only active hybrid journals from JCT
+  inner_join(hoaddata::jn_ind, by = "issn_l") |>
+  filter(!cr_year %in% c("2017", "2018", "2024"))
+
+active_jns_with_oa <- active_jns |> 
+  filter(!is.na(cc))
+
+# Get indicators for all articles without applying paratext adn supplement detection
+hoad_all <- read_csv(here::here("data", "cr_all_ind.csv"))
+
+## Web of Science
+wos_jct_items_df <- readr::read_csv(here::here("data-raw", "wos_jct_items_df.csv"))
+
+wos_hybrid_jns <- wos_jct_items_df |>
+  # No PNAS
+  filter(issn_l != "0027-8424") |>
+  # No Full OA detected by OA prop
+  filter(!issn_l %in% hoad_oa_excluded$issn_l)
+
+wos_active_jns <- wos_hybrid_jns |>
+  # Use online first, if available
+  mutate(online_year = lubridate::year(wos_pubdate_online)) |>
+  mutate(pub_year = ifelse(!is.na(online_year), online_year, pubyear)) |>
+  filter(pub_year %in% 2019:2023)
+
+wos_active_jns_with_oa <- wos_active_jns |> 
+  filter(grepl("hybrid", oa_status))
+
+# wos core 
+wos_active_core <- wos_active_jns |>
+  filter(grepl("Article|^Review$", item_type))
+
+wos_active_core_oa <- wos_active_jns |>
+  filter(grepl("hybrid", oa_status))
+
+## Scopus
+
+scp_jct_items_df <- readr::read_csv(here::here("data-raw", "scp_jct_items_df.csv"))
+
+scp_hybrid_jns <- scp_jct_items_df |>
+  # No PNAS
+  filter(issn_l != "0027-8424") |>
+  # No Full OA detected by OA prop
[…]

### 2025-02-10 analysis/manuscript.Rmd
Start to describe the matching procedures

@@
 
-### journal matching
+This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
+The journal data source used was a unified dataset covering various snapshots of the Transformative Agreements Public Data from the cOAlition S Journal Checker Tool (JCT), a well-established source to help authors identify suitable open access publishing venues.
+These snapshots, archived weekly by the author using a cron job via a publicly available GitHub repository, cover agreements active from July 2021 to July 2024. 
+After enriching the JCT journal data with the linking ISSN (ISSN-L) according to the ISSN Registry Agency, a comprehensive exclusion of fully open access journals, including flipped journals covering multiple journal lists (DOAJ, OpenAlex, Bielefeld) and article-level investigations, was applied as described elsewhere.
+In total, xxxx hybrid journals could be linked to xxx agreements listed in the ESAC Transformative Agreements Registry, which formed to basis for obtaining article-level data including author affiliations, open access status ans document types.
 
-### authorship records
+For each of the databases examined, article metadata such as publication identifier, document types (where available), author roles and institutional affiliations, publication date and open access information were retrieved for each journal, using all ISSN variants linked to an ISSN-L according to the ISSN Registry Agency.
+The publication period was determined using the earliest year of publication. 
+In the case of Crossref, the article data source for hoaddata, this was the publication date, although variations were observed (check ref).
+For Web of Science and Scopus, the earliest publication date was used where available. In the latter case, the KB determined the earliest known publication date by tracking versions of the raw data.
+
+<!-- literature to check https://direct.mit.edu/qss/article/3/1/99/109079/The-prevalence-and-impact-of-university-->
+
+Affiliation data, including author roles, are essential for linking articles to transformative agreements.
+Therefore, author affiliations were retrieved for the first and corresponding author at the article level for each of the databases examined.
+In order to account for different address variants in bibliometric databases, the database-specific IDs for affiliations were used. 
+For hoaddata these were the ROR IDs from OpenAlex, for Web of Science the affiliation enhanced names and for Scopus the Scopus Affiliation Identifier, an eight-digit ID number.
+In addition, the country of the author's address, represented by ISO country codes, was retrieved.
+
+At the time of our study, there was no bibliometric database that included information on whether an article was published under a transformative agreement. 
+Here, eligible articles were identified by matching hybrid journals and participating institutions from the Transformative Agreement Data Dump with the databases examined per active agreement.
+This matching process also took into account the duration of the agreements according to the ESAC registry.
+While in the case of hoaddata, which uses OpenAlex as an affiliated data source, matching was straightforward due to the availability of the ROR ID in both sources, neither Web of Science nor Scopus supported ROR at the time of the study.
+Instead, xxxx articles identified under a transformative agreement in hoaddata were processed to map the ROR ID associated with the first author to the proprietary affiliation identifier in Web of Science and Scopus representing the affiliation of the first author.
+Due to multiple affiliations, different organisational hierarchies encoded in the different organisational ID systems, and to deal with conflicting matches, an algorithm then sorted the ROR-ID and proprietary affiliation identifier pairs in descending order and selected the most frequent combination per ROR-ID as a match.
+
+On the basis of the matching tables thus compiled, articles eligible under transformative agreements could also be retrieved from the Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT.
 
 ###  Identifying eligible articles under transformative agreements
+
+### adn which of them were os
 

### 2025-02-10 analysis/manuscript.tex
Start to describe the matching procedures

@@
 
-\subsubsection{journal matching}\label{journal-matching}
+This study examined the adoption of open access for hybrid journals
+included in transformative agreements, which published at least one
+journal article in the five-year period 2019-2023. The journal data
+source used was a unified dataset covering various snapshots of the
+Transformative Agreements Public Data from the cOAlition S Journal
+Checker Tool (JCT), a well-established source to help authors identify
+suitable open access publishing venues. These snapshots, archived weekly
+by the author using a cron job via a publicly available GitHub
+repository, cover agreements active from July 2021 to July 2024. After
+enriching the JCT journal data with the linking ISSN (ISSN-L) according
+to the ISSN Registry Agency, a comprehensive exclusion of fully open
+access journals, including flipped journals covering multiple journal
+lists (DOAJ, OpenAlex, Bielefeld) and article-level investigations, was
+applied as described elsewhere. In total, xxxx hybrid journals could be
+linked to xxx agreements listed in the ESAC Transformative Agreements
+Registry, which formed to basis for obtaining article-level data
+including author affiliations, open access status ans document types.
 
-\subsubsection{authorship records}\label{authorship-records}
+For each of the databases examined, article metadata such as publication
+identifier, document types (where available), author roles and
+institutional affiliations, publication date and open access information
+were retrieved for each journal, using all ISSN variants linked to an
+ISSN-L according to the ISSN Registry Agency. The publication period was
+determined using the earliest year of publication. In the case of
+Crossref, the article data source for hoaddata, this was the publication
+date, although variations were observed (check ref). For Web of Science
+and Scopus, the earliest publication date was used where available. In
+the latter case, the KB determined the earliest known publication date
+by tracking versions of the raw data.
+
+Affiliation data, including author roles, are essential for linking
+articles to transformative agreements. Therefore, author affiliations
+were retrieved for the first and corresponding author at the article
+level for each of the databases examined. In order to account for
+different address variants in bibliometric databases, the
+database-specific IDs for affiliations were used. For hoaddata these
+were the ROR IDs from OpenAlex, for Web of Science the affiliation
+enhanced names and for Scopus the Scopus Affiliation Identifier, an
+eight-digit ID number. In addition, the country of the author's address,
+represented by ISO country codes, was retrieved.
+
+At the time of our study, there was no bibliometric database that
+included information on whether an article was published under a
+transformative agreement. Here, eligible articles were identified by
+matching hybrid journals and participating institutions from the
+Transformative Agreement Data Dump with the databases examined per
+active agreement. This matching process also took into account the
+duration of the agreements according to the ESAC registry. While in the
+case of hoaddata, which uses OpenAlex as an affiliated data source,
+matching was straightforward due to the availability of the ROR ID in
+both sources, neither Web of Science nor Scopus supported ROR at the
+time of the study. Instead, xxxx articles identified under a
+transformative agreement in hoaddata were processed to map the ROR ID
+associated with the first author to the proprietary affiliation
+identifier in Web of Science and Scopus representing the affiliation of
+the first author. Due to multiple affiliations, different organisational
+hierarchies encoded in the different organisational ID systems, and to
+deal with conflicting matches, an algorithm then sorted the ROR-ID and
+proprietary affiliation identifier pairs in descending order and
+selected the most frequent pair as a match.
+
+On the basis of the matching tables thus compiled, articles eligible
+under transformative agreements could also be retrieved from the Web of
+Science and Scopus, although they did not contain the ROR IDs used by
+the JCT.
 
@@
 agreements}\label{identifying-eligible-articles-under-transformative-agreements}
+
+\subsubsection{adn which of them were
+os}\label{adn-which-of-them-were-os}
 

### 2025-02-12 analysis/manuscript.Rmd
Tmp save methods

--- 2025-02-05 analysis/manuscript.Rmd
@@
 After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using open access evidence, author roles (first and corresponding) and affiliation data are presented.
-Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study.
+<!-- Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study. -->
 
@@
 
-hoaddata, developed and maintained by the author, is an R data package comprising information about the uptake of hybrid open access since 2017 from several openly available data sources.
+hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package comprising information about the uptake of hybrid open access since 2017 from several openly available data sources.
 It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.
@@
 Creative Commons licence information relative to the published version ("version of record").
-Because of limited affiliation metadata in Crossref (https://doi.org/10.31222/osf.io/smxe5), hoaddata sources  first-author affiliations from OpenAlex.
-While the country in which a first-author was located was used to aggregate country-level statistics, Research Organization Registry (ROR) identifiers (ROR-ID) were used in conjunction with ESAC registry information on the duration of an agreement was used to estimate whether an article was published under active transformative agreements.
+Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources first-author affiliations from OpenAlex.
+
+<!--While the country in which a first-author was located was used to aggregate country-level statistics, Research Organization Registry (ROR) identifiers (ROR-ID) were used in conjunction with ESAC registry information on the duration of an agreement was used to estimate whether an article was published under active transformative agreements.
 This matching benefited from the availability of the ROR-IDs in both sources.
-To improve the matching, JCT data were enriched to include associated institutions, such as unviersity hospitals.
+To improve the matching, JCT data were enriched to include associated institutions, such as unviersity hospitals.-->
 
@@
 To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow interfaces with the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
-The package has been regularly updated since 2022 and the version including the computation log is available on GitHub.
-
-hoaddata is used as a data basis of the Hybrid Open Access Dashboard, a data analytics services for  library consortia and publishers to track the uptake of hybrid open access through transformative agreements.
-It was also used in bibliometric research (Jahn 2025).
+The workflow has run regularly to fetch updates from these data sources since 2022.
+The package version used in this study is 0.3, containing data from the Crossref 2024-08 dump provided to Crossref Metadata Plus subscribers and the OpenAlex 2024-08-29 monthly dump.
+This version including the computation log is available on GitHub (https://github.com/subugoe/hoaddata/releases/tag/v.0.3).
 
@@
 Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020]. 
-The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).
+The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI), collectively referred to as Web of Science Primary (WoS Primary).
 
-The WoS provides important data points for studying spending on open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as Openalex.
+The WoS Primary provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as Openalex.
 However, it lacks information about journals and articles under transformative agreements.
 
-For programmatic access to article-level data, the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany is used to access bibliometric data.
+For programmatic access to article-level data, this study used the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany.
 The KB processes raw XML data provided by Clarivate Analytics, which is provided as an in-house PostgreSQL database under a uniform schema.
 To support reproducibility, KB maintains annual snapshots of the database. 
-Accordingly, this study used the annual snapshot from April 2024, which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].
+Accordingly, this study used the annual snapshot from April 2024, wos_b_202404, which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].
 
@@
 Similar to the Web of Science, Scopus is selective with regard to the journals it indexes.
-However, its coverage is substantially more extensive than that of the Web of Science Core collection [@Singh_2021; @Visser_2021].
+However, its coverage is substantially more extensive than that of the Web of Science collections considered in this study [@Singh_2021; @Visser_2021].
 With detailed metadata about article types, open access status information derived from Unpaywall, author roles, and disambiguated affiliations, Scopus also contains important data to assess open access uptake, although no direct information regarding transformative agreements was available at the time of the study.
 
-This study used the Scopus annual snapshot of April 2024 as provided by the KB. The same KB curation effort was applied to the  Scopus raw data as for the Web of Science [@schmidt_2024_13935407].
+This study used the Scopus annual snapshot of April 2024 as provided by the KB (scp_b_202404).
+The same KB curation effort was applied to the  Scopus raw data as for the Web of Science [@schmidt_2024_13935407].
 
@@
 
-### journal matching
+### Determining hybrid journal publication volume
 
-### authorship records
+This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
+The initial data source was the Transformative Agreements Public Data from the cOAlition S Journal Checker Tool (JCT), which helps authors identify suitable open access publishing venues. 
+Weekly snapshots of this data were archived using an automated GitHub Actions workflow (https://github.com/njahn82/jct_data) and into hoaddata.
+The resulting JCT data cover agreements active from July 2021 to July 2024.
 
-###  Identifying eligible articles under transformative agreements
+The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
+A comprehensive exclusion of fully open access journals was performed, following the methodology described in @jahn_2025.
+This involved checking multiple journal lists (DOAJ, OpenAlex, Bielefeld) and article-level investigations.
+The JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
+<!-- In total, xxxx hybrid journals could be linked to xxx agreements listed in the ESAC Transformative Agreements Registry, which formed to basis for obtaining article-level data including author affiliations, open access status ans document types.
+-->
+
+For each database, article metadata was retrieved using all ISSN variants linked to an ISSN-L.
+The metadata included DOIs, document types (where available), author roles, institutional affiliations, publication dates and open access information.
+Publication years were determined using the earliest known date of publication in a journal.
+In the case of Crossref, the article data source for hoaddata, this was the issued date, although variations were observed (check ref).
+For Web of Science and Scopus, the earliest publication date was used where available.
+In the latter case, the KB determined the earliest known publication date by tracking versions of the raw data.
+
+<!-- literature to check https://direct.mit.edu/qss/article/3/1/99/109079/The-prevalence-and-impact-of-university-->
+
+### Open Access
+
+Only published versions with Creative Commons licenses available on publisher platforms were considered as hybrid open access. 
+hoaddata derived this information from Crossref's license metadata, while Web of Science and Scopus  used data from Unpaywall. 
+In additiaon to Crossref, Unpaywall also parses publisher websites directly, because some publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018], although transformative agreement workflows requiring this information during DOI registration [@Geschuhn_2017].
+
+### Author affiliations
+
+```{r cache=TRUE}
+# Article level data TA
+ta_oa_inst <- readr::read_csv(here::here("data-raw", "ta_oa_inst.csv"))
+```
+Author affiliations at the article level were retrieved for both first and, if available, corresponding authors across all databases. 
+To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
+Additionally, ISO country codes were retrieved for each author's address to compile country-level statistics. 
+These country codes for Web of Science and Scopus were provided by the KB.
+
+Because neither Web of Science nor Scopus support ROR IDs—the institution identifier used by the JCT—a two-step matching process was implemented to identify articles covered by transformative agreements.
+First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
+Then, an algorithm selected the most frequent ROR ID and proprietary identifier pairs to handle multiple affiliations and organizational hierarchy differences.
+
+```{r aff_evaluation}
+scp_ror <- readr::read_csv(here::here("data-raw", "scp_wos_matching.csv")) 
+wos_ror <- readr::read_csv(here::here("data-raw", "ror_wos_matching.csv")) 
+```
+
+This process linked `r length(unique(scp_ror$ror))` ROR IDs to  `r length(unique(scp_ror$vendor_org_id))` Scopus Affiliation IDs, and  `r length(unique(wos_ror$ror))` ROR IDs to `r length(unique(scp_ror$vendor_org_id))` enhanced affiliation strings in the Web of Science.
+Quality evaluation through random sampling of 50 pairs revealed an error rate of 22% for Web of Science (11 mismatches) and 6% for Scopus (3 mismatches). 
+Upon inspection, these mismatches primarily occurred with less-represented institutions having only a few publications, introduced through multiple affiliations of single authors. 
+The difference between databases suggests that Scopus's affiliation control aligns more closely with ROR than that of the Web of Science.
+
+On the basis of these matching tables, articles eligible under transformative agreements could also be retrieved from the Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT.
+
+### Estimating open access in hybrid journals covered by transformative agreements
 
 ## Data records
+
+As a result of the above-described comprehensive data processing, data-sets were prepared for each database on open access in hybrid journals included in transformative agreements at the country and journal level by year.
+Table \ref{tab:methods_overview_table} presents an general overview of the coverage of hybrid journals in transformative agreements 2019-23 per dataset. 
+It shows that the majority of hybrid journals published at least one original article or review (core) in the five-years period.
+These journals constituted the basis for the subsequent aggregation of publication metrics. 
+While hoad only covers articles with DOI, the aggreation of Scopus and Web of Scienc ewas carried out isng the databse identifier. 
+A subsequent compatision of DOU coverage illustrates that non.core articles in Web of Scienc e lack DOIs. These are in aprticular meeting abstracts, a pecularity of this database.
+Open access was aggregated using DOI, as all alos Unpaywall only colelctzs opena cces ststaus information for articles with DOI. 
+A closer view on core articles with affiliation data reveal a lack of data withr egard to corresponding auhtors in the case of OpenAlex.
+Only aroud two third of investiaged artilces had at affilation data linked to corresponding authoirs.
+In case of first authors, the share was higher.
+In contrast, web of scienc eand scopus recorded a higher coverage.
+Accordingly, only first authro evidence were presented for haoddata in the follwoing
 

### 2025-02-12 analysis/manuscript.tex
Tmp save methods

--- 2025-02-05 analysis/manuscript.tex
@@
 evidence, author roles (first and corresponding) and affiliation data
-are presented. Overall, xxxx hybrid journals from xxx agreements that
-published at least one open access article between 2019 and 2023. formed
-the basis of this study.
+are presented.
 
@@
 
-hoaddata, developed and maintained by the author, is an R data package
-comprising information about the uptake of hybrid open access since 2017
-from several openly available data sources. It combines article-level
+hoaddata, developed and maintained by the author to support open access
+monitoring and research (Jahn, 2025), is an R data package comprising
+information about the uptake of hybrid open access since 2017 from
+several openly available data sources. It combines article-level
 metadata from Crossref and OpenAlex with transformative agreement
@@
 (``version of record''). Because of limited affiliation metadata in
-Crossref (\url{https://doi.org/10.31222/osf.io/smxe5}), hoaddata sources
-first-author affiliations from OpenAlex. While the country in which a
-first-author was located was used to aggregate country-level statistics,
-Research Organization Registry (ROR) identifiers (ROR-ID) were used in
-conjunction with ESAC registry information on the duration of an
-agreement was used to estimate whether an article was published under
-active transformative agreements. This matching benefited from the
-availability of the ROR-IDs in both sources. To improve the matching,
-JCT data were enriched to include associated institutions, such as
-unviersity hospitals.
+Crossref (Eck \& Waltman, 2022), hoaddata sources first-author
+affiliations from OpenAlex.
 
@@
 which provides high-performant programmatic access to monthly snapshots
-of Crossref and OpenAlex. The package has been regularly updated since
-2022 and the version including the computation log is available on
-GitHub.
-
-hoaddata is used as a data basis of the Hybrid Open Access Dashboard, a
-data analytics services for library consortia and publishers to track
-the uptake of hybrid open access through transformative agreements. It
-was also used in bibliometric research (Jahn 2025).
+of Crossref and OpenAlex. The workflow has run regularly to fetch
+updates from these data sources since 2022. The package version used in
+this study is 0.3, containing data from the Crossref 2024-08 dump
+provided to Crossref Metadata Plus subscribers and the OpenAlex
+2024-08-29 monthly dump. This version including the computation log is
+available on GitHub
+(\url{https://github.com/subugoe/hoaddata/releases/tag/v.0.3}).
 
@@
 Science Citation Index Expanded (SCIE), the Social Sciences Citation
-Index (SSCI) and the Arts \& Humanities Citation Index (AHCI).
+Index (SSCI) and the Arts \& Humanities Citation Index (AHCI),
+collectively referred to as Web of Science Primary (WoS Primary).
 
-The WoS provides important data points for studying spending on open
+The WoS Primary provides important data points for analysing open
 access: author affiliations and roles, differentiation of journal
@@
 
-For programmatic access to article-level data, the database of the
-Kompetenznetzwerk Bibliometrie (KB) in Germany is used to access
-bibliometric data. The KB processes raw XML data provided by Clarivate
-Analytics, which is provided as an in-house PostgreSQL database under a
-uniform schema. To support reproducibility, KB maintains annual
-snapshots of the database. Accordingly, this study used the annual
-snapshot from April 2024, which is considered to cover almost the entire
-previous publication year (Schmidt et al., 2024).
+For programmatic access to article-level data, this study used the
+database of the Kompetenznetzwerk Bibliometrie (KB) in Germany. The KB
+processes raw XML data provided by Clarivate Analytics, which is
+provided as an in-house PostgreSQL database under a uniform schema. To
+support reproducibility, KB maintains annual snapshots of the database.
+Accordingly, this study used the annual snapshot from April 2024,
+wos\_b\_202404, which is considered to cover almost the entire previous
+publication year (Schmidt et al., 2024).
 
@@
 journals it indexes. However, its coverage is substantially more
-extensive than that of the Web of Science Core collection (Singh et al.,
-2021; Visser et al., 2021). With detailed metadata about article types,
-open access status information derived from Unpaywall, author roles, and
-disambiguated affiliations, Scopus also contains important data to
-assess open access uptake, although no direct information regarding
-transformative agreements was available at the time of the study.
+extensive than that of the Web of Science collections considered in this
+study (Singh et al., 2021; Visser et al., 2021). With detailed metadata
+about article types, open access status information derived from
+Unpaywall, author roles, and disambiguated affiliations, Scopus also
+contains important data to assess open access uptake, although no direct
+information regarding transformative agreements was available at the
+time of the study.
 
 This study used the Scopus annual snapshot of April 2024 as provided by
-the KB. The same KB curation effort was applied to the Scopus raw data
-as for the Web of Science (Schmidt et al., 2024).
+the KB (scp\_b\_202404). The same KB curation effort was applied to the
+Scopus raw data as for the Web of Science (Schmidt et al., 2024).
 
@@
 
-\subsubsection{journal matching}\label{journal-matching}
+\subsubsection{Determining hybrid journal publication
+volume}\label{determining-hybrid-journal-publication-volume}
 
-\subsubsection{authorship records}\label{authorship-records}
+This study examined the adoption of open access for hybrid journals
+included in transformative agreements, which published at least one
+journal article in the five-year period 2019-2023. The initial data
+source was the Transformative Agreements Public Data from the cOAlition
+S Journal Checker Tool (JCT), which helps authors identify suitable open
+access publishing venues. Weekly snapshots of this data were archived
+using an automated GitHub Actions workflow
+(\url{https://github.com/njahn82/jct_data}) and into hoaddata. The
+resulting JCT data cover agreements active from July 2021 to July 2024.
 
-\subsubsection{Identifying eligible articles under transformative
-agreements}\label{identifying-eligible-articles-under-transformative-agreements}
+The JCT journal data was enriched with the linking ISSN (ISSN-L)
+according to the ISSN Registry Agency. A comprehensive exclusion of
+fully open access journals was performed, following the methodology
+described in (\textbf{jahn\_2025?}). This involved checking multiple
+journal lists (DOAJ, OpenAlex, Bielefeld) and article-level
+investigations. The JCT institution data was enriched with ROR-IDs from
+associated institutions, such as university hospitals or institutes of
+large research organisations such as the Max Planck Society, according
+to OpenAlex' institution entity.
+
+For each database, article metadata was retrieved using all ISSN
+variants linked to an ISSN-L. The metadata included DOIs, document types
+(where available), author roles, institutional affiliations, publication
+dates and open access information. Publication years were determined
+using the earliest known date of publication in a journal. In the case
+of Crossref, the article data source for hoaddata, this was the issued
+date, although variations were observed (check ref). For Web of Science
+and Scopus, the earliest publication date was used where available. In
+the latter case, the KB determined the earliest known publication date
+by tracking versions of the raw data.
+
+\subsubsection{Open Access}\label{open-access}
+
+Only published versions with Creative Commons licenses available on
+publisher platforms were considered as hybrid open access. hoaddata
+derived this information from Crossref's license metadata, while Web of
+Science and Scopus used data from Unpaywall. In additiaon to Crossref,
+Unpaywall also parses publisher websites directly, because some
+publishers do not provide machine-readable Creative Commons license
+information (Piwowar et al., 2018), although transformative agreement
+workflows requiring this information during DOI registration (Geschuhn
+\& Stone, 2017).
+
+\subsubsection{Author affiliations}\label{author-affiliations}
+
+Author affiliations at the article level were retrieved for both first
+and, if available, corresponding authors across all databases. To handle
+different address variants, database-specific affiliation identifiers
+were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced
+names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
+Additionally, ISO country codes were retrieved for each author's address
+to compile country-level statistics. These country codes for Web of
+Science and Scopus were provided by the KB.
+
+Because neither Web of Science nor Scopus support ROR IDs---the
+institution identifier used by the JCT---a two-step matching process was
+implemented to identify articles covered by transformative agreements.
+First, 2,782,540 articles from 6,457 institutions with ROR-IDs in the
+JCT data since 2017 (according to hoaddata) were processed to map first
+authors' ROR-IDs to corresponding proprietary affiliation identifier in
+Web of Science and Scopus using DOI matching. Then, an algorithm
+selected the most frequent ROR ID and proprietary identifier pairs to
+handle multiple affiliations and organizational hierarchy differences.
+
+This process linked 6,375 ROR IDs to 4,894 Scopus Affiliation IDs, and
+6,034 ROR IDs to 4,894 enhanced affiliation strings in the Web of
+Science. Quality evaluation through random sampling of 50 pairs revealed
+an error rate of 22\% for Web of Science (11 mismatches) and 6\% for
+Scopus (3 mismatches). Upon inspection, these mismatches primarily
+occurred with less-represented institutions having only a few
+publications, introduced through multiple affiliations of single
+authors. The difference between databases suggests that Scopus's
+affiliation control aligns more closely with ROR than that of the Web of
+Science.
+
+On the basis of these matching tables, articles eligible under
+transformative agreements could also be retrieved from the Web of
+Science and Scopus, although they did not contain the ROR IDs used by
+the JCT.
+
+\subsubsection{Estimating open access in hybrid journals covered by
+transformative
+agreements}\label{estimating-open-access-in-hybrid-journals-covered-by-transformative-agreements}
 
 \subsection{Data records}\label{data-records}
+
+As a result of the above-described comprehensive data processing,
+data-sets were prepared for each database on open access in hybrid
+journals included in transformative agreements at the country and
+journal level by year. Table \ref{tab:methods_overview_table} presents
+an general overview of the coverage of hybrid journals in transformative
+agreements 2019-23 per dataset. It shows that the majority of hybrid
+journals published at least one original article or review (core) in the
+five-years period. These journals constituted the basis for the
+subsequent aggregation of publication metrics. While hoad only covers
+articles with DOI, the aggreation of Scopus and Web of Scienc ewas
+carried out isng the databse identifier. A subsequent compatision of DOU
+coverage illustrates that non.core articles in Web of Scienc e lack
+DOIs. These are in aprticular meeting abstracts, a pecularity of this
+database. Open access was aggregated using DOI, as all alos Unpaywall
+only colelctzs opena cces ststaus information for articles with DOI. A
+closer view on core articles with affiliation data reveal a lack of data
+withr egard to corresponding auhtors in the case of OpenAlex. Only aroud
+two third of investiaged artilces had at affilation data linked to
+corresponding authoirs. In case of first authors, the share was higher.
+In contrast, web of scienc eand scopus recorded a higher coverage.
+Accordingly, only first authro evidence were presented for haoddata in
+the follwoing
 
@@
 
+\bibitem[\citeproctext]{ref-van_Eck_2022}
+Eck, N. J. van, \& Waltman, L. (2022). \emph{Crossref as a source of
+open bibliographic metadata}.
+\url{https://doi.org/10.31222/osf.io/smxe5}
+
+\bibitem[\citeproctext]{ref-Geschuhn_2017}
+Geschuhn, K., \& Stone, G. (2017). It's the workflows, stupid! What is
+required to make {``offsetting''} work for the open access transition.
+\emph{Insights the {UKSG} Journal}, \emph{30}(3), 103--114.
+\url{https://doi.org/10.1629/uksg.391}
+
+\bibitem[\citeproctext]{ref-Jahn_2025}
+Jahn, N. (2025). How open are hybrid journals included in transformative
+agreements? \emph{Quantitative Science Studies}, 1--39.
+\url{https://doi.org/10.1162/qss_a_00348}
+
+\bibitem[\citeproctext]{ref-Piwowar_2018}
+Piwowar, H., Priem, J., Larivière, V., Alperin, J. P., Matthias, L.,
+Norlander, B., Farley, A., West, J., \& Haustein, S. (2018). The state
+of {OA}: A large-scale analysis of the prevalence and impact of open
+access articles. \emph{{PeerJ}}, \emph{6}, e4375.
+\url{https://doi.org/10.7717/peerj.4375}
+
 \bibitem[\citeproctext]{ref-schmidt_2024_13935407}

### 2025-02-14 analysis/manuscript.Rmd
First draft data and methods, next results startung with coverage, followed by replication

@@
 
-The aim of this study is to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
-To achieve this, results from hoaddata, an openly available collection of open research information regarding hybrid open access, was compared with the established bibliometric databases Web of Science and Scopus.
-After describing the initial data sources used, the necessary pre-processing steps to obtain eligible articles from transformative agreements using open access evidence, author roles (first and corresponding) and affiliation data are presented.
+This study aims to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
+As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
+The subsequent section, introduces the initial data sources followed by a presentation of the necessary pre-processing steps to obtain eligible articles from transformative agreements by using author roles (first and corresponding) and harmonised affiliation data.
 <!-- Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study. -->
@@
 
+\begin{figure}[htbp]
+\centering
+\begin{tikzpicture}[node distance=2cm, thick,scale=1, every node/.style={scale=0.9}]    
+% Data Sources
+    \node (crossref) [datasource] {Crossref};
+    \node (openAlex) [datasource, right=of crossref] {OpenAlex};
+    \node (jct) [datasource, right=of openAlex] {JCT};
+    \node (wos) [datasource, below=of crossref] {Web of Science};
+    \node (scopus) [datasource, below=of jct] {Scopus};
+
+    % Processing Steps
+    \node (hoaddata) [datasource, below=1.5cm of $(crossref)!1!(openAlex)$] {hoaddata};
+    \node (kb) [process, below=1cm of $(wos)!0.5!(scopus)$] {Journal and article retrieval (including open access)};
+
+ % Matching Process - Fixed line break syntax
+    \node (matching) [process, below=1.5cm of $(hoaddata)!1!(kb)$] {Harmonising author affiliations through RoR};
+
+ % Aggregation
+    \node (transmatch) [process, below=1.5cm of $(matching)!1!(matching)$] {Estimate TA output};
+
+    % Output
+    \node (analysis) [output, below=1.5cm of $(transmatch)!1!(transmatch)$] {Aggregated analytical datasets};
+
+    % Arrows
+    \draw [arrow] (crossref) -- (hoaddata);
+    \draw [arrow] (openAlex) -- (hoaddata);
+    \draw [arrow] (jct) -- (hoaddata);
+    \draw [arrow] (hoaddata) -- (kb);
+    \draw [arrow] (wos) -- (kb);
+    \draw [arrow] (scopus) -- (kb);
+    \draw [arrow] (kb) -- (matching);
+    \draw [arrow] (matching) -- (transmatch);
+    \draw [arrow] (transmatch) -- (analysis);
+
+    % Legend
+    \node [datasource, scale=0.7] at ($(analysis)+(4,0)$) {Data Sources};
+    \node [process, scale=0.7] at ($(analysis)+(4,-1)$) {Processing Steps};
+    \node [output, scale=0.7] at ($(analysis)+(4,-2)$) {Output};
+\end{tikzpicture}
+\caption{Data processing workflow for comparing hybrid open access uptake across bibliometric data sources.
+The workflow shows how data from different sources (Crossref, OpenAlex, JCT, Web of Science, and Scopus) is processed and matched using ROR identifiers to enable comparative analysis.
+TA = Transformative Agreement}
+\label{fig:workflow}
+\end{figure}
+
 ## Data sources
@@
 
-More specifically, hoaddata uses Crossref, a DOI registration agency, for obtaining journal publication volume and open access status through
-Creative Commons licence information relative to the published version ("version of record").
+More specifically, hoaddata uses a unified dataset of historic JCT snapshots since July 2021, which were enriched ISSN variants linked to an ISSN-L and associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
+A comprehensive exclusion of fully open access journals using  multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals, as described in @Jahn_2025.
+hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
 Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources first-author affiliations from OpenAlex.
@@
 
-hoaddata follows good practices for computational reproducibility using R.
+hoaddata follows good practices for computational reproducibility using R [@Marwick_2018].
 The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
@@
 To support reproducibility, KB maintains annual snapshots of the database. 
-Accordingly, this study used the annual snapshot from April 2024, wos_b_202404, which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].
+Accordingly, this study used the annual snapshot from April 2024 (wos_b_202404), which is considered to cover almost the entire previous publication year [@schmidt_2024_13935407].
 
@@
 
-This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
-The initial data source was the Transformative Agreements Public Data from the cOAlition S Journal Checker Tool (JCT), which helps authors identify suitable open access publishing venues. 
+The hoaddata package (v.0.3) provided an enriched and unified list of hybrid journals in transformative agreements according to the JCT, covering agreements active from July 2021 to July 2024 according to the ESAC registry.
+<!-- This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
+The initial data source was the Transformative Agreements Public Data from the cOAlition S Journal Checker Tool (JCT), which helps authors identify suitable open access publishing venues, as provided by hoaddata.
  Weekly snapshots of this data were archived using an automated GitHub Actions workflow (https://github.com/njahn82/jct_data) and into hoaddata.
-The resulting JCT data cover agreements active from July 2021 to July 2024.
+The JCT data used in the examined hoaddata version, which was enriched with linking ISSN (ISSN-L) and ROR-IDs from associated institutions before inclusion, cover agreements active from July 2021 to July 2024.
+-->
 
-The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
+<!-- The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
 A comprehensive exclusion of fully open access journals was performed, following the methodology described in @jahn_2025.
@@
 -->
+Article metadata was retrieved from each databse using all ISSN variants linked to an ISSN-L.
+The metadata included DOIs, publication dates, article types, open access information as well as author roles and institutional affiliations.
 
-For each database, article metadata was retrieved using all ISSN variants linked to an ISSN-L.
-The metadata included DOIs, document types (where available), author roles, institutional affiliations, publication dates and open access information.
+The publication year for each article was determined using the earliest known publication date. 
+For articles in hoaddata, this corresponded to Crossref's issued date.
+
 Publication years were determined using the earliest known date of publication in a journal.
-In the case of Crossref, the article data source for hoaddata, this was the issued date, although variations were observed (check ref).
-For Web of Science and Scopus, the earliest publication date was used where available.
-In the latter case, the KB determined the earliest known publication date by tracking versions of the raw data.
+In the case of Crossref, the article data source for hoaddata, this was the issued date<!--, although variations were observed (check ref)-->.
+For Web of Science and Scopus, the earliest publication date was used where available, with Scopus dates specifically determined by the KB through version tracking of the raw data.
+
+Many transformative agreements typically cover only certain types of journal
+articles, in particular original research articles including reviews.
+Because of limited information on these document types in open scholarly data [@Haupka_2024], hoaddata used an extended version of Unpaywall's paratext recognition approach to exclude non-scholarly content [@Piwowar_2018].
+To exclude conference supplements, which are also often not covered by transformative agreements, only articles published in regular issues, indicated by numerical pagination, were considered.
+For Web of Science and Scopus, their established document type classifications were used to identify original research articles and reviews, referred to as core articles throughout this study.
 
@@
 
-### Open Access
+### Identifying open access articles in hybrid journals
 
-Only published versions with Creative Commons licenses available on publisher platforms were considered as hybrid open access. 
-hoaddata derived this information from Crossref's license metadata, while Web of Science and Scopus  used data from Unpaywall. 
-In additiaon to Crossref, Unpaywall also parses publisher websites directly, because some publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018], although transformative agreement workflows requiring this information during DOI registration [@Geschuhn_2017].
+Published articles were considered hybrid open access  when they were made freely available under a Creative Commons license on publishers' platforms. 
+While hoaddata sourced this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall. 
+Unpaywall supplements Crossref metadata by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018]. 
+This additional parsing remains necessary despite transformative agreement workflows requiring license information during DOI registration [@Geschuhn_2017].
+Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information.
 
-### Author affiliations
+### Harmonising author affiliations across databases
 
@@
 ```
-Author affiliations at the article level were retrieved for both first and, if available, corresponding authors across all databases. 
+Author affiliations were retrieved for both first and, if available, corresponding authors across all databases to prepare the linking between articles and institutions covered by transformative agreements.
 To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
@@
 
-Because neither Web of Science nor Scopus support ROR IDs—the institution identifier used by the JCT—a two-step matching process was implemented to identify articles covered by transformative agreements.
+Because neither Web of Science nor Scopus support ROR IDs—the institution identifier used by the JCT—a two-step matching process was implemented to harmonise affiliation data.
 First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
-Then, an algorithm selected the most frequent ROR ID and proprietary identifier pairs to handle multiple affiliations and organizational hierarchy differences.
+Then, an algorithm selected the most frequent ROR ID and proprietary identifier pairs to handle multiple affiliations and organisational hierarchy differences.
 
@@
 
-On the basis of these matching tables, articles eligible under transformative agreements could also be retrieved from the Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT.
+### Estimating open access in hybrid journals covered by transformative agreements
 
-### Estimating open access in hybrid journals covered by transformative agreements
+Based on these matching tables, articles eligible under transformative agreements could also be retrieved from Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT. 
+The estimation of eligible articles followed @Jahn_2025 and included a matching of both journals and participating institutions according to the Transformative Agreement Data dump for each of the data sources examined.
+The matching also took into account the duration of agreements according to the ESAC registry, with only those matches where an agreement was actually in place being considered for subsequent analysis.
 
@@
 
-As a result of the above-described comprehensive data processing, data-sets were prepared for each database on open access in hybrid journals included in transformative agreements at the country and journal level by year.
-Table \ref{tab:methods_overview_table} presents an general overview of the coverage of hybrid journals in transformative agreements 2019-23 per dataset. 
-It shows that the majority of hybrid journals published at least one original article or review (core) in the five-years period.
-These journals constituted the basis for the subsequent aggregation of publication metrics. 
-While hoad only covers articles with DOI, the aggreation of Scopus and Web of Scienc ewas carried out isng the databse identifier. 
-A subsequent compatision of DOU coverage illustrates that non.core articles in Web of Scienc e lack DOIs. These are in aprticular meeting abstracts, a pecularity of this database.
-Open access was aggregated using DOI, as all alos Unpaywall only colelctzs opena cces ststaus information for articles with DOI. 
-A closer view on core articles with affiliation data reveal a lack of data withr egard to corresponding auhtors in the case of OpenAlex.
-Only aroud two third of investiaged artilces had at affilation data linked to corresponding authoirs.
-In case of first authors, the share was higher.
-In contrast, web of scienc eand scopus recorded a higher coverage.
-Accordingly, only first authro evidence were presented for haoddata in the follwoing
+As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were produced for each database at country and journal level by year.
+Table \ref{tab:methods_overview_table} provides a general overview of the coverage between 2019 and 2023 per database.
+It shows that the majority of hybrid journals published at least one original article or review marked as core during the five-year period.
+These journals formed the basis for the subsequent aggregation of publication metrics.
 
-```{r methods_overview_table, cache=TRUE}
+```{r methods_overview_table}
 load(here::here("data", "manuscript_data.Rda"))
@@
 
-# results
+While hoaddata only covers articles with a DOI, the Scopus and Web of Science figures were also aggregated using the database identifier.
+A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
+This was particularly the case for meeting abstracts, a particular feature of this database.
+Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
+A closer look at the core articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
+Only about two-thirds of the articles examined provided corresponding author affiliations.
+For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(hoad) * 100`%.
+Accordingly, only first author data for hoaddata are considered in the following analysis.
+
+# Results
 

### 2025-02-14 analysis/manuscript.tex
First draft data and methods, next results startung with coverage, followed by replication

@@
 \shorttitle{}
+
+% flow chart
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows, positioning, calc, shapes.multipart}
+
+% Custom styles for the flowchart
+\tikzstyle{datasource} = [rectangle, rounded corners, minimum width=2.5cm, minimum height=1cm,text centered, draw=blue!60, fill=blue!5]
+\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=green!60, fill=green!5]
+\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=red!60, fill=red!5]
+\tikzstyle{arrow} = [thick,->,>=stealth]
+\tikzstyle{output} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm, text centered, draw=purple!60, fill=purple!5]
 
@@
 
-The aim of this study is to demonstrate the suitability of open
-scholarly data sources for assessing the impact of transformative
-agreements on hybrid open access. To achieve this, results from
-hoaddata, an openly available collection of open research information
-regarding hybrid open access, was compared with the established
-bibliometric databases Web of Science and Scopus. After describing the
-initial data sources used, the necessary pre-processing steps to obtain
-eligible articles from transformative agreements using open access
-evidence, author roles (first and corresponding) and affiliation data
-are presented.
+This study aims to demonstrate the suitability of open scholarly data
+sources for assessing the impact of transformative agreements on hybrid
+open access. As shown in Figure \ref{fig:workflow}, the methodology
+involved comparing hoaddata, an openly available collection of open
+research information on hybrid open access, with the bibliometric
+databases Web of Science and Scopus. The subsequent section, introduces
+the initial data sources followed by a presentation of the necessary
+pre-processing steps to obtain eligible articles from transformative
+agreements by using author roles (first and corresponding) and
+harmonised affiliation data.
+
+\begin{figure}[htbp]
+\centering
+\begin{tikzpicture}[node distance=2cm, thick,scale=1, every node/.style={scale=0.9}]    
+% Data Sources
+    \node (crossref) [datasource] {Crossref};
+    \node (openAlex) [datasource, right=of crossref] {OpenAlex};
+    \node (jct) [datasource, right=of openAlex] {JCT};
+    \node (wos) [datasource, below=of crossref] {Web of Science};
+    \node (scopus) [datasource, below=of jct] {Scopus};
+
+    % Processing Steps
+    \node (hoaddata) [datasource, below=1.5cm of $(crossref)!1!(openAlex)$] {hoaddata};
+    \node (kb) [process, below=1cm of $(wos)!0.5!(scopus)$] {Journal and article retrieval (including open access)};
+
+ % Matching Process - Fixed line break syntax
+    \node (matching) [process, below=1.5cm of $(hoaddata)!1!(kb)$] {Harmonising author affiliations through RoR};
+
+ % Aggregation
+    \node (transmatch) [process, below=1.5cm of $(matching)!1!(matching)$] {Estimate TA output};
+
+    % Output
+    \node (analysis) [output, below=1.5cm of $(transmatch)!1!(transmatch)$] {Aggregated analytical datasets};
+
+    % Arrows
+    \draw [arrow] (crossref) -- (hoaddata);
+    \draw [arrow] (openAlex) -- (hoaddata);
+    \draw [arrow] (jct) -- (hoaddata);
+    \draw [arrow] (hoaddata) -- (kb);
+    \draw [arrow] (wos) -- (kb);
+    \draw [arrow] (scopus) -- (kb);
+    \draw [arrow] (kb) -- (matching);
+    \draw [arrow] (matching) -- (transmatch);
+    \draw [arrow] (transmatch) -- (analysis);
+
+    % Legend
+    \node [datasource, scale=0.7] at ($(analysis)+(4,0)$) {Data Sources};
+    \node [process, scale=0.7] at ($(analysis)+(4,-1)$) {Processing Steps};
+    \node [output, scale=0.7] at ($(analysis)+(4,-2)$) {Output};
+\end{tikzpicture}
+\caption{Data processing workflow for comparing hybrid open access uptake across bibliometric data sources.
+The workflow shows how data from different sources (Crossref, OpenAlex, JCT, Web of Science, and Scopus) is processed and matched using ROR identifiers to enable comparative analysis.
+TA = Transformative Agreement}
+\label{fig:workflow}
+\end{figure}
 
@@
 
-More specifically, hoaddata uses Crossref, a DOI registration agency,
-for obtaining journal publication volume and open access status through
-Creative Commons licence information relative to the published version
-(``version of record''). Because of limited affiliation metadata in
-Crossref (Eck \& Waltman, 2022), hoaddata sources first-author
-affiliations from OpenAlex.
+More specifically, hoaddata uses a unified dataset of historic JCT
+snapshots since July 2021, which were enriched ISSN variants linked to
+an ISSN-L and associated ROR-IDs representing university hospitals or
+institutes of large research organisations, according to OpenAlex'
+institution entity. A comprehensive exclusion of fully open access
+journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld) was
+performed to identify hybrid journals, as described in Jahn (2025).
+hoaddata relies on Crossref for obtaining journal publication volume and
+open access status through Creative Commons licence information relative
+to the published version (``version of record''). Because of limited
+affiliation metadata in Crossref (Eck \& Waltman, 2022), hoaddata
+sources first-author affiliations from OpenAlex.
 
 hoaddata follows good practices for computational reproducibility using
-R. The package, which includes data, code, a test suite and
-documentation, is openly available on GitHub. To ensure computational
-reproducibility while aggregating the data, a GitHub Actions continuous
-integration and delivery (CI/CD) workflow interfaces with the SUB
-Göttingen's open scholarly data warehouse based on Google BigQuery,
-which provides high-performant programmatic access to monthly snapshots
-of Crossref and OpenAlex. The workflow has run regularly to fetch
-updates from these data sources since 2022. The package version used in
-this study is 0.3, containing data from the Crossref 2024-08 dump
-provided to Crossref Metadata Plus subscribers and the OpenAlex
+R (Marwick et al., 2018). The package, which includes data, code, a test
+suite and documentation, is openly available on GitHub. To ensure
+computational reproducibility while aggregating the data, a GitHub
+Actions continuous integration and delivery (CI/CD) workflow interfaces
+with the SUB Göttingen's open scholarly data warehouse based on Google
+BigQuery, which provides high-performant programmatic access to monthly
+snapshots of Crossref and OpenAlex. The workflow has run regularly to
+fetch updates from these data sources since 2022. The package version
+used in this study is 0.3, containing data from the Crossref 2024-08
+dump provided to Crossref Metadata Plus subscribers and the OpenAlex
 2024-08-29 monthly dump. This version including the computation log is
@@
 support reproducibility, KB maintains annual snapshots of the database.
-Accordingly, this study used the annual snapshot from April 2024,
-wos\_b\_202404, which is considered to cover almost the entire previous
-publication year (Schmidt et al., 2024).
+Accordingly, this study used the annual snapshot from April 2024
+(wos\_b\_202404), which is considered to cover almost the entire
+previous publication year (Schmidt et al., 2024).
 
@@
 
-This study examined the adoption of open access for hybrid journals
-included in transformative agreements, which published at least one
-journal article in the five-year period 2019-2023. The initial data
-source was the Transformative Agreements Public Data from the cOAlition
-S Journal Checker Tool (JCT), which helps authors identify suitable open
-access publishing venues. Weekly snapshots of this data were archived
-using an automated GitHub Actions workflow
-(\url{https://github.com/njahn82/jct_data}) and into hoaddata. The
-resulting JCT data cover agreements active from July 2021 to July 2024.
+The hoaddata package (v.0.3) provided an enriched and unified list of
+hybrid journals in transformative agreements according to the JCT,
+covering agreements active from July 2021 to July 2024 according to the
+ESAC registry.
 
-The JCT journal data was enriched with the linking ISSN (ISSN-L)
-according to the ISSN Registry Agency. A comprehensive exclusion of
-fully open access journals was performed, following the methodology
-described in (\textbf{jahn\_2025?}). This involved checking multiple
-journal lists (DOAJ, OpenAlex, Bielefeld) and article-level
-investigations. The JCT institution data was enriched with ROR-IDs from
-associated institutions, such as university hospitals or institutes of
-large research organisations such as the Max Planck Society, according
-to OpenAlex' institution entity.
+Article metadata was retrieved from each databse using all ISSN variants
+linked to an ISSN-L. The metadata included DOIs, publication dates,
+article types, open access information as well as author roles and
+institutional affiliations.
 
-For each database, article metadata was retrieved using all ISSN
-variants linked to an ISSN-L. The metadata included DOIs, document types
-(where available), author roles, institutional affiliations, publication
-dates and open access information. Publication years were determined
-using the earliest known date of publication in a journal. In the case
-of Crossref, the article data source for hoaddata, this was the issued
-date, although variations were observed (check ref). For Web of Science
-and Scopus, the earliest publication date was used where available. In
-the latter case, the KB determined the earliest known publication date
-by tracking versions of the raw data.
+The publication year for each article was determined using the earliest
+known publication date. For articles in hoaddata, this corresponded to
+Crossref's issued date.
 
-\subsubsection{Open Access}\label{open-access}
+Publication years were determined using the earliest known date of
+publication in a journal. In the case of Crossref, the article data
+source for hoaddata, this was the issued date. For Web of Science and
+Scopus, the earliest publication date was used where available, with
+Scopus dates specifically determined by the KB through version tracking
+of the raw data.
 
-Only published versions with Creative Commons licenses available on
-publisher platforms were considered as hybrid open access. hoaddata
-derived this information from Crossref's license metadata, while Web of
-Science and Scopus used data from Unpaywall. In additiaon to Crossref,
-Unpaywall also parses publisher websites directly, because some
-publishers do not provide machine-readable Creative Commons license
-information (Piwowar et al., 2018), although transformative agreement
-workflows requiring this information during DOI registration (Geschuhn
-\& Stone, 2017).
+Many transformative agreements typically cover only certain types of
+journal articles, in particular original research articles including
+reviews. Because of limited information on these document types in open
+scholarly data (Haupka et al., 2024), hoaddata used an extended version
+of Unpaywall's paratext recognition approach to exclude non-scholarly
+content (Piwowar et al., 2018). To exclude conference supplements, which
+are also often not covered by transformative agreements, only articles
+published in regular issues, indicated by numerical pagination, were
+considered. For Web of Science and Scopus, their established document
+type classifications were used to identify original research articles
+and reviews, referred to as core articles throughout this study.
 
-\subsubsection{Author affiliations}\label{author-affiliations}
+\subsubsection{Identifying open access articles in hybrid
+journals}\label{identifying-open-access-articles-in-hybrid-journals}
 
-Author affiliations at the article level were retrieved for both first
-and, if available, corresponding authors across all databases. To handle
-different address variants, database-specific affiliation identifiers
-were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced
-names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
-Additionally, ISO country codes were retrieved for each author's address
-to compile country-level statistics. These country codes for Web of
-Science and Scopus were provided by the KB.
+Published articles were considered hybrid open access when they were
+made freely available under a Creative Commons license on publishers'
+platforms. While hoaddata sourced this information from Crossref license
+metadata, Web of Science and Scopus relied on Unpaywall. Unpaywall
+supplements Crossref metadata by parsing publisher websites directly,
+addressing cases where publishers do not provide machine-readable
+Creative Commons license information (Piwowar et al., 2018). This
+additional parsing remains necessary despite transformative agreement
+workflows requiring license information during DOI registration
+(Geschuhn \& Stone, 2017). Both Web of Science and Scopus defined hybrid
+open access consistently as content available under Creative Commons
+licenses on publisher platforms, distinguishing it from bronze open
+access that lack such explicit license information.
+
+\subsubsection{Harmonising author affiliations across
+databases}\label{harmonising-author-affiliations-across-databases}
+
+Author affiliations were retrieved for both first and, if available,
+corresponding authors across all databases to prepare the linking
+between articles and institutions covered by transformative agreements.
+To handle different address variants, database-specific affiliation
+identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation
+enhanced names for Web of Science, and Scopus Affiliation Identifiers
+for Scopus. Additionally, ISO country codes were retrieved for each
+author's address to compile country-level statistics. These country
+codes for Web of Science and Scopus were provided by the KB.
 
@@
 institution identifier used by the JCT---a two-step matching process was
-implemented to identify articles covered by transformative agreements.
-First, 2,782,540 articles from 6,457 institutions with ROR-IDs in the
-JCT data since 2017 (according to hoaddata) were processed to map first
-authors' ROR-IDs to corresponding proprietary affiliation identifier in
-Web of Science and Scopus using DOI matching. Then, an algorithm
-selected the most frequent ROR ID and proprietary identifier pairs to
-handle multiple affiliations and organizational hierarchy differences.
+implemented to harmonise affiliation data. First, 2,782,540 articles
+from 6,457 institutions with ROR-IDs in the JCT data since 2017
+(according to hoaddata) were processed to map first authors' ROR-IDs to
+corresponding proprietary affiliation identifier in Web of Science and
+Scopus using DOI matching. Then, an algorithm selected the most frequent
+ROR ID and proprietary identifier pairs to handle multiple affiliations
+and organisational hierarchy differences.
 
@@
 
-On the basis of these matching tables, articles eligible under
-transformative agreements could also be retrieved from the Web of
-Science and Scopus, although they did not contain the ROR IDs used by
-the JCT.
-
 \subsubsection{Estimating open access in hybrid journals covered by
@@
 
+Based on these matching tables, articles eligible under transformative
+agreements could also be retrieved from Web of Science and Scopus,
+although they did not contain the ROR IDs used by the JCT. The
+estimation of eligible articles followed Jahn (2025) and included a
+matching of both journals and participating institutions according to
+the Transformative Agreement Data dump for each of the data sources
+examined. The matching also took into account the duration of agreements
+according to the ESAC registry, with only those matches where an
+agreement was actually in place being considered for subsequent
+analysis.
+
 \subsection{Data records}\label{data-records}
 
-As a result of the above-described comprehensive data processing,
-data-sets were prepared for each database on open access in hybrid
-journals included in transformative agreements at the country and
-journal level by year. Table \ref{tab:methods_overview_table} presents
-an general overview of the coverage of hybrid journals in transformative
-agreements 2019-23 per dataset. It shows that the majority of hybrid
-journals published at least one original article or review (core) in the
-five-years period. These journals constituted the basis for the
-subsequent aggregation of publication metrics. While hoad only covers
-articles with DOI, the aggreation of Scopus and Web of Scienc ewas
-carried out isng the databse identifier. A subsequent compatision of DOU
-coverage illustrates that non.core articles in Web of Scienc e lack
-DOIs. These are in aprticular meeting abstracts, a pecularity of this
-database. Open access was aggregated using DOI, as all alos Unpaywall
-only colelctzs opena cces ststaus information for articles with DOI. A
-closer view on core articles with affiliation data reveal a lack of data
-withr egard to corresponding auhtors in the case of OpenAlex. Only aroud
-two third of investiaged artilces had at affilation data linked to
-corresponding authoirs. In case of first authors, the share was higher.
-In contrast, web of scienc eand scopus recorded a higher coverage.
-Accordingly, only first authro evidence were presented for haoddata in
-the follwoing
+As a result of the comprehensive data processing described above,
+datasets on open access in hybrid journals included in transformative
+agreements were produced for each database at country and journal level
+by year. Table \ref{tab:methods_overview_table} provides a general
+overview of the coverage between 2019 and 2023 per database. It shows
+that the majority of hybrid journals published at least one original
+article or review marked as core during the five-year period. These
+journals formed the basis for the subsequent aggregation of publication
+metrics.
 
@@
 
-\section{results}\label{results}
+While hoaddata only covers articles with a DOI, the Scopus and Web of
+Science figures were also aggregated using the database identifier. A
+subsequent comparison of DOI coverage shows that non-core articles in
+Web of Science often lacked a DOI. This was particularly the case for
+meeting abstracts, a particular feature of this database. Open access
+indicators were aggregated by DOI, as Unpaywall only collects
+information on open access status for articles with a DOI. A closer look
+at the core articles with affiliation data reveals a lack of data on
+corresponding authors in the case of OpenAlex, the affiliation data
+source used by hoaddata, compared to Web of Science and Scopus. Only
+about two-thirds of the articles examined provided corresponding author
+affiliations. For first authors, the proportion was 89\%. Accordingly,
+only first author data for hoaddata are considered in the following
+analysis.
+
+\section{Results}\label{results}
 
@@
 
+\bibitem[\citeproctext]{ref-Haupka_2024}
+Haupka, N., Culbert, J. H., Schniedermann, A., Jahn, N., \& Mayr, P.
+(2024). \emph{Analysis of the publication and document types in
+OpenAlex, web of science, scopus, pubmed and semantic scholar}. arXiv.
+\url{https://doi.org/10.48550/ARXIV.2406.15154}
+
 \bibitem[\citeproctext]{ref-Jahn_2025}
@@
 \url{https://doi.org/10.1162/qss_a_00348}
+
+\bibitem[\citeproctext]{ref-Marwick_2018}
+Marwick, B., Boettiger, C., \& Mullen, L. (2018). Packaging data
+analytical work reproducibly using {R} (and friends). \emph{The American
+Statistician}, \emph{72}(1), 80--88.
+\url{https://doi.org/10.1080/00031305.2017.1375986}
 

### 2025-02-17 analysis/manuscript.qmd
Migrate to Quarto

--- 2025-02-14 analysis/manuscript.Rmd
@@
 ---
-output:
-  pdf_document:
+format:
+  pdf:
+    keep-tex: true
     fig_caption: yes
-    keep_tex: yes
     template: apa6.sty
@@
   message = FALSE,
-  cache = TRUE,
-  cache.lazy = FALSE,
   echo = FALSE,
@@
 As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were produced for each database at country and journal level by year.
-Table \ref{tab:methods_overview_table} provides a general overview of the coverage between 2019 and 2023 per database.
+@tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023 per database.
 It shows that the majority of hybrid journals published at least one original article or review marked as core during the five-year period.
@@
 
-```{r methods_overview_table}
+```{r methods_overview_tab}
+#| label: tbl-methods_overview_tab
+#| tbl-cap: "Coverage of hybrid journals in transformative agreements 2019-23."
 load(here::here("data", "manuscript_data.Rda"))
@@
   position = "H",
-  caption ="Coverage of hybrid journals in transformative agreements 2019-23.",
   col.names = c("", "HOAD", "Web of Science", "Scopus")

### 2025-02-17 analysis/manuscript.tex
Migrate to Quarto

@@
 
-% bibliography
 % definitions for citeproc citations
@@
   \begingroup\normalfont
-  \includegraphics[height=6px]{orcid_logo.png}%
+  \includegraphics[height=0.9em]{orcid_logo}% removed .png extension
   \endgroup
 }
-Najko Jahn\textsuperscript{1}\textsuperscript{*} (\orcid{} \href{https://orcid.org/0000-0001-5105-1463}{\color{black}{0000-0001-5105-1463}}) \\
+
+Najko Jahn\textsuperscript{1}\textsuperscript{*} 
+(\orcid{} \href{https://orcid.org/0000-0001-5105-1463}{\color{black}{0000-0001-5105-1463}})
 
@@
 available on GitHub
-(\url{https://github.com/subugoe/hoaddata/releases/tag/v.0.3}).
+(https://github.com/subugoe/hoaddata/releases/tag/v.0.3).
 
@@
 agreements were produced for each database at country and journal level
-by year. Table \ref{tab:methods_overview_table} provides a general
+by year. Table~\ref{tbl-methods_overview_tab} provides a general
 overview of the coverage between 2019 and 2023 per database. It shows
@@
 
-\begin{table}[H]
-\centering
-\caption{\label{tab:methods_overview_table}Coverage of hybrid journals in transformative agreements 2019-23.}
+\begin{table}
+
+\caption{\label{tbl-methods_overview_tab}Coverage of hybrid journals in
+transformative agreements 2019-23.}
+
+\centering{
+
 \centering
@@
 \end{tabular}
-\end{table}
+
+}
+
+\end{table}%
 

### 2025-02-17 analysis/manuscript.tex
Align with APA manual

@@
-\documentclass[a4paper,man,floatsintext,longtable,noextraspace,12pt]{apa6}
+\documentclass[a4paper,man,floatsintext,longtable,noextraspace,10pt]{apa6}
 
@@
 \tikzstyle{output} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm, text centered, draw=purple!60, fill=purple!5]
+
+\setlength\parindent{1.27cm}
 
@@
 agreements were produced for each database at country and journal level
-by year. Table~\ref{tbl-methods_overview_tab} provides a general
+by year. Table Table~\ref{tbl-methods_overview_tab} provides a general
 overview of the coverage between 2019 and 2023 per database. It shows

### 2025-02-17 analysis/manuscript.qmd
Align with APA manual
Starting results with coverage analysis

@@
 library(patchwork)
+library(ComplexUpset)
+
+# Themes
+my_upset_theme <-
+  theme(panel.grid.minor = element_blank(),
+        panel.grid.major.x = element_blank(),
+        plot.title.position="plot",
+        plot.title = element_text(hjust = 0, vjust = 1, face = "bold"))
 ```
@@
 As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were produced for each database at country and journal level by year.
-@tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023 per database.
+Table @tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023 per database.
 It shows that the majority of hybrid journals published at least one original article or review marked as core during the five-year period.
@@
 
+First, I will present a coverage analysis. 
+Then, I investigate hybrid open access uptake using the same methods for each database and the compare results against each other in order to assess the suitablility of the biblioemtric databases to examine transofrmative agreements.
+
+## Coverage comparision
+
+Overall, % (n = ) of hybrid journals included in transformative agreements were indexed in all three databases (Fig A).
+These were followed by the combination between hoaddata and Scopus, which indexes more journals than the investiaged Web of Science collection SCIE, SSCI and the Arts & Humanities Citation Index AHCI.
+Around % of journals were only indexed by hoaddata. 
+Journals
+
+Note that there were 32 journals were indexed in WoS or Scopus.
+Upon isnepction, these were due to diasmbiguity of ISSNs matching (errors ISSN).
+
+```{r coverage_results, fig.asp=0.75}
+jn_upset_articles <- readr::read_csv(here::here("data", "jn_upset_articles.csv"))
+# Helper function to make text labels the same size as axis labels as seen in
+# <https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/>
+# Todo: Learn how to invoke it
+jn_upset_articles |> group_by(Crossref, WoS, Scopus) |> summarize(across(.cols = where(is.numeric),.fns =sum, na.rm = TRUE))
+
+ggtext_size <- function(base_size, ratio = 0.8) {
+  ratio * base_size / ggplot2::.pt
+}
+
+bar_width <- 0.7
+size <- get_size_mode("exclusive_intersection")
+
+upset_overview <- ComplexUpset::upset(
+  jn_upset_articles,
+  c("Crossref", "WoS", "Scopus"),
+  # Only journals in Crossref
+  min_size = 30,
+  set_sizes = FALSE,
+  stripes = "transparent",
+  # wrap = TRUE,
+  name = "Data Source Coverage",
+  base_annotations = list(
+    "Journals" = (
+      ComplexUpset::intersection_size(
+        counts = TRUE,
+        text = list(size = 8 / .pt),
+        text_mapping = aes(
+          label = paste0(!!upset_text_percentage(), '\n(', format(!!size, big.mark = ","), ')'),
+          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
+          y = ifelse(!!size > 5000, !!size - 4000, !!size)
+        ), width = bar_width
+      ) +
+        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
+        ggtitle("A")
+    )
+  ),
+  annotations = list(
+    "Journal size\n(log scale)" = (
+      ggplot(mapping = aes(x = intersection, y = n)) +
+        geom_boxplot(outliers = FALSE, width = bar_width) +
+        scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
+        ggtitle("C")
+    ),
+    "Articles" = (
+      ggplot(mapping = aes(y = n_cr)) +
+        geom_bar(
+          stat = "identity",
+          color = "grey80",
+          width = bar_width,
+          aes(fill = "in Crossref only")
+        ) +
+        geom_bar(
+          aes(y = doi_in_wos_and_scopus, fill = "Shared"),
+          stat = "identity",
+          width = bar_width,
+          color = "#0093c7"
+        ) +
+        scale_fill_manual(
+          values = c(
+            `in Crossref only` = "grey80",
+            `Shared` = "#0093c7"
+          ),
+          name = "DOI"
+        ) +
+        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+        ggtitle("B")
+    )
+  ),
+  matrix = (intersection_matrix(geom = geom_point(
+    shape = "square", size = 2
+  )))
+)
+upset_overview & my_upset_theme
+```
+
+This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network
+Overall, xxxx hybrid journals included in xxx agreements that published at least one open access article between 2019 and 2023.
+
 # discussion

### 2025-02-17 analysis/manuscript.tex
Starting results with coverage analysis

@@
 
+First, I will present a coverage analysis. Then, I investigate hybrid
+open access uptake using the same methods for each database and the
+compare results against each other in order to assess the suitablility
+of the biblioemtric databases to examine transofrmative agreements.
+
+\subsection{Coverage comparision}\label{coverage-comparision}
+
+Overall, \% (n = ) of hybrid journals included in transformative
+agreements were indexed in all three databases (Fig A). These were
+followed by the combination between hoaddata and Scopus, which indexes
+more journals than the investiaged Web of Science collection SCIE, SSCI
+and the Arts \& Humanities Citation Index AHCI. Around \% of journals
+were only indexed by hoaddata. Journals
+
+Note that there were 32 journals were indexed in WoS or Scopus. Upon
+isnepction, these were due to diasmbiguity of ISSNs matching (errors
+ISSN).
+
+\begin{verbatim}
+#> # A tibble: 7 x 10
+#> # Groups:   Crossref, WoS [4]
+#>   Crossref WoS   Scopus    n_cr   n_wos   n_scp       n doi_in_wos_and_scopus
+#>   <lgl>    <lgl> <lgl>    <dbl>   <dbl>   <dbl>   <dbl>                 <dbl>
+#> 1 FALSE    FALSE TRUE         0       0   99354   99354                 93230
+#> 2 FALSE    TRUE  FALSE        0    5495       0    5495                     9
+#> 3 FALSE    TRUE  TRUE         0  228954  231554  231554                225943
+#> 4 TRUE     FALSE FALSE   177110       0       0  177110                   559
+#> 5 TRUE     FALSE TRUE    657697       0  529707  657697                494884
+#> 6 TRUE     TRUE  FALSE    10049    8919       0   10049                  1191
+#> 7 TRUE     TRUE  TRUE   7313569 6477606 6469446 7313569               6161590
+#> # i 2 more variables: doi_in_scopus <dbl>, doi_in_wos <dbl>
+\end{verbatim}
+
+\begin{center}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/coverage_results-1.png}
+\end{center}
+
+This study compares results drawn from the openly available and
+regularly updated hoaddata, an R package leveraging the open scholarly
+data sources Crossref and OpenAlex, with the in-house database of the
+German Competence Network Overall, xxxx hybrid journals included in xxx
+agreements that published at least one open access article between 2019
+and 2023.
+
 \section*{discussion}\label{discussion}

### 2025-02-17 analysis/manuscript_preprocessing_data.R

@@ -1,10 +1,12 @@
 ### Data preprocessing for the manuscript
 
-## Why? It just takes too long to render the manuscript.
+## Why? It just takes too long to render everything in the manuscript.
 
 library(tidyverse, warn.conflicts = FALSE)
 library(bigrquery)
 library(here)
+
+# METHOD -------
 
 ## HOAD
 hoad_jns <- hoaddata::jct_hybrid_jns |>
@@ -288,3 +290,60 @@ save(# Journals
   file = here::here("data", "manuscript_data.Rda")
   )
 
+# Results --------
+
+## Overlap -------
+
+# DOI set too large to track it with Git, see 001_data_gathering how the data
+# was obtained from Google BQ
+cr_df <- readr::read_csv(here::here("data-raw/hoad_dois_all_19_23.csv"))
+
+cr_active_upset <- cr_df |>
+  filter(issn_l %in% active_jns_with_oa$issn_l) |>
+  group_by(issn_l) |>
+  summarise(n_cr = n_distinct(doi)) |>
+  mutate(Crossref = TRUE)
+wos_active_upset <- wos_active_core |>
+  filter(issn_l %in% wos_active_core_oa$issn_l) |>
+  group_by(issn_l) |>
+  summarise(n_wos = n_distinct(item_id)) |>
+  mutate(WoS = TRUE)
+scp_active_upset <- scp_active_core |>
+  filter(issn_l %in% scp_active_core_oa$issn_l) |>
+  group_by(issn_l) |>
+  summarise(n_scp = n_distinct(item_id)) |>
+  mutate(Scopus = TRUE)
+
+jn_upset_ <- dplyr::full_join(cr_active_upset, wos_active_upset, by = "issn_l") |>
+  dplyr::full_join(scp_active_upset, by = "issn_l") |>
+  mutate(across(c(Crossref, WoS, Scopus), ~ replace_na(.x, FALSE))) |>
+  # Journal 5-years article volume
+  mutate(n = case_when(!is.na(n_cr) ~ n_cr, !is.na(n_scp) ~ n_scp, TRUE ~ n_wos))
+
+## DOI overlap
+scp_dois <- scp_active_core |>
+  filter(issn_l %in% scp_active_core_oa$issn_l) |> 
+  pull(doi)
+wos_dois <- scp_active_core |>
[…]

### 2025-02-18 analysis/manuscript.qmd
Continue coverage analysis

--- 2025-02-17 analysis/manuscript.qmd
@@
 As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
-The subsequent section, introduces the initial data sources followed by a presentation of the necessary pre-processing steps to obtain eligible articles from transformative agreements by using author roles (first and corresponding) and harmonised affiliation data.
+This section introduces the initial data sources followed by a presentation of the necessary pre-processing steps to obtain eligible articles from transformative agreements by using author roles (first and corresponding) and harmonised affiliation data.
 <!-- Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study. -->
@@
 
-hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package comprising information about the uptake of hybrid open access since 2017 from several openly available data sources.
+hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package comprising information about the uptake of hybrid open access since 2017 collected from several openly available data sources.
 It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.
 
-More specifically, hoaddata uses a unified dataset of historic JCT snapshots since July 2021, which were enriched ISSN variants linked to an ISSN-L and associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
-A comprehensive exclusion of fully open access journals using  multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals, as described in @Jahn_2025.
+As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of historic JCT snapshots since July 2021, which were enriched with ISSN variants linked to an ISSN-L and associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
+A comprehensive exclusion of fully open access journals using  multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals.
 hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
@@
 The metadata included DOIs, publication dates, article types, open access information as well as author roles and institutional affiliations.
-
-The publication year for each article was determined using the earliest known publication date. 
-For articles in hoaddata, this corresponded to Crossref's issued date.
-
 Publication years were determined using the earliest known date of publication in a journal.
-In the case of Crossref, the article data source for hoaddata, this was the issued date<!--, although variations were observed (check ref)-->.
+articles in hoaddata, this corresponded to Crossref's issued date.<!--, although variations were observed (check ref)-->
 For Web of Science and Scopus, the earliest publication date was used where available, with Scopus dates specifically determined by the KB through version tracking of the raw data.
@@
 
-As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were produced for each database at country and journal level by year.
+As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were aggregated for each database at country and journal level by year.
 @tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023 per database.
-It shows that the majority of hybrid journals published at least one original article or review marked as core during the five-year period.
+It shows that the majority of hybrid journals published at least one original research article or review marked as core during the five-year period.
 These journals formed the basis for the subsequent aggregation of publication metrics.
@@
 
-While hoaddata only covers articles with a DOI, the Scopus and Web of Science figures were also aggregated using the database identifier.
+While hoaddata only covers articles with a DOI, Scopus and Web of Science publication metrics were aggregated using the database identifier.
 A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
-This was particularly the case for meeting abstracts, a particular feature of this database.
+This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025].
 Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
@@
 For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(hoad) * 100`%.
-Accordingly, only first author data for hoaddata are considered in the following analysis.
+Therefore, only first author data for hoaddata were considered in the following analysis.
 
@@
 
+First, I will present a coverage analysis. 
+Then, I investigate hybrid open access uptake using the same methods for each database and the compare results against each other in order to assess the suitablility of the biblioemtric databases to examine transofrmative agreements.
+
+## Coverage comparision
+
+```{r}
+jn_upset_articles <- readr::read_csv(here::here("data", "jn_upset_articles.csv"))
+
+all_jns <- jn_upset_articles |> 
+  distinct(issn_l)
+
+union_jns <- jn_upset_articles |>
+  filter(WoS == TRUE, Crossref == TRUE, Scopus == TRUE) |>
+  distinct(issn_l)
+
+upset_data <- jn_upset_articles |> 
+   mutate(overlap = case_when(Crossref == TRUE & WoS == TRUE & Scopus == TRUE ~ doi_in_wos_or_scopus,
+                             Crossref == TRUE & WoS == FALSE & Scopus == TRUE ~ doi_in_scopus,
+                             Crossref == TRUE & WoS == TRUE & Scopus == FALSE ~ doi_in_wos,
+                             Crossref == TRUE & WoS == FALSE & Scopus == FALSE ~ n_cr,
+                             Crossref == FALSE & WoS == FALSE & Scopus == TRUE ~ doi_in_scopus,
+                             Crossref == FALSE & WoS == TRUE & Scopus == FALSE ~ doi_in_wos,
+                             Crossref == FALSE & WoS == TRUE & Scopus == TRUE ~ doi_in_wos_or_scopus))
+
+```
+
+Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid journals included in transformative agreements, visualising the intersections of journals and articles across the examined databases as an UpSet graph [@Lex_2014; @ComplexUpset].
+The analysis included hybrid journals that published at least one open access article between 2019 and 2023, based on open access status information from each database.
+Only original research articles and reviews were considered in the analysis.
+
+```{r upset_coverage_results, fig.asp=0.75}
+#| fig-cap: "Upset graph"
+#| label: fig-upset_coverage_results
+# Helper function to make text labels the same size as axis labels as seen in
+# <https://wjschne.github.io/posts/making-text-labels-the-same-size-as-axis-labels-in-ggplot2/>
+# Todo: Learn how to invoke it
+
+ggtext_size <- function(base_size, ratio = 0.8) {
+  ratio * base_size / ggplot2::.pt
+}
+
+bar_width <- 0.7
+size <- get_size_mode("exclusive_intersection")
+
+upset_overview <- ComplexUpset::upset(
+  upset_data,
+  c("Crossref", "WoS", "Scopus"),
+  # Only journals in Crossref
+  min_size = 30,
+  set_sizes = FALSE,
+  stripes = "transparent",
+  # wrap = TRUE,
+  name = "Data Source Coverage",
+  base_annotations = list(
+    "Journals" = (
+      ComplexUpset::intersection_size(
+        counts = FALSE,
+        text = list(size = 8 / .pt),
+        text_mapping = aes(
+          label = paste0(!!upset_text_percentage(), '\n(', format(!!size, big.mark = ","), ')'),
+          colour = ifelse(!!size > 5000, 'on_bar', 'on_background'),
+          y = ifelse(!!size > 5000, !!size - 4000, !!size)
+        ), width = bar_width
+      ) +
+        scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
+        ggtitle("A")
+    )
+  ),
+  annotations = list(
+    "Journal size\n(log scale)" = (
+      ggplot(mapping = aes(x = intersection, y = n)) +
+        geom_boxplot(outliers = FALSE, width = bar_width) +
+        scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
+        ggtitle("C")
+    ),
+    "Articles" = (
+      ggplot(mapping = aes(y = n_cr)) +
+        geom_bar(
+          stat = "identity",
+          color = "grey80",
+          width = bar_width,
+          aes(fill = "All")
+        ) +
+        geom_bar(
+          aes(y = overlap, fill = "Shared"),
+          stat = "identity",
+          width = bar_width,
+          color = "#0091c8"
+        ) +
+        scale_fill_manual(
+          values = c(
+            `All` = "grey80",
+            `Shared` = "#0091c8"
+          ),
+          name = "DOI"
+        ) +
+        scale_y_continuous(labels =  scales::number_format(big.mark = ","),
+                           breaks = c(0, 2500000, 5000000, 7500000)) +
+        ggtitle("B")
+    )
+  ),
+  matrix = (intersection_matrix(geom = geom_point(
+    shape = "square", size = 2
+  )))
+)
+upset_overview & my_upset_theme
+```
+Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases ( \ref{fig-upset_coverage_results}A).
+The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
+Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while `r round( jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory databases Scopus and Web of Science.
+Upon insepction, this latter group mainly represents hybrid journals for which no open access evidence could be retrieved from Crossref, which serves as the open access evidence source for hoaddata.
+
+In terms of article coverage, Fig B shows the DOI overlap, distinguihsing between a shared corpus and just those where a DOI was only available through hoaddata.
+Upon inspection, reasons for lackign DOI coverage in proprietory database were different publication dates and differenc ein document type definitions. 
+As such, demosntrates certain uncertainies using the open approach.
+Overall, The majority of articles can be found
+
+Using DOIs also the size of journals were calculated (Fig C).
+It highlights thate large spread in the intersection between all three datasett.
+Journals are, on average, smaller than those just covered by one or two sources.In particular journals exlcusively covered by hoaddata are significant smaller than those covered by all three sources.
+
+This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network
+Overall, xxxx hybrid journals included in xxx agreements that published at least one open access article between 2019 and 2023.
+
 # discussion

### 2025-02-18 analysis/manuscript.tex
Continue coverage analysis

--- 2025-02-17 analysis/manuscript.tex
@@
 research information on hybrid open access, with the bibliometric
-databases Web of Science and Scopus. The subsequent section, introduces
-the initial data sources followed by a presentation of the necessary
-pre-processing steps to obtain eligible articles from transformative
-agreements by using author roles (first and corresponding) and
-harmonised affiliation data.
+databases Web of Science and Scopus. This section introduces the initial
+data sources followed by a presentation of the necessary pre-processing
+steps to obtain eligible articles from transformative agreements by
+using author roles (first and corresponding) and harmonised affiliation
+data.
 
@@
 monitoring and research (Jahn, 2025), is an R data package comprising
-information about the uptake of hybrid open access since 2017 from
-several openly available data sources. It combines article-level
+information about the uptake of hybrid open access since 2017 collected
+from several openly available data sources. It combines article-level
 metadata from Crossref and OpenAlex with transformative agreement
@@
 
-More specifically, hoaddata uses a unified dataset of historic JCT
-snapshots since July 2021, which were enriched ISSN variants linked to
-an ISSN-L and associated ROR-IDs representing university hospitals or
-institutes of large research organisations, according to OpenAlex'
-institution entity. A comprehensive exclusion of fully open access
-journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld) was
-performed to identify hybrid journals, as described in Jahn (2025).
-hoaddata relies on Crossref for obtaining journal publication volume and
-open access status through Creative Commons licence information relative
-to the published version (``version of record''). Because of limited
-affiliation metadata in Crossref (Eck \& Waltman, 2022), hoaddata
-sources first-author affiliations from OpenAlex.
+As discussed elsewhere (Jahn, 2025), hoaddata uses a unified dataset of
+historic JCT snapshots since July 2021, which were enriched with ISSN
+variants linked to an ISSN-L and associated ROR-IDs representing
+university hospitals or institutes of large research organisations,
+according to OpenAlex' institution entity. A comprehensive exclusion of
+fully open access journals using multiple journal lists (DOAJ, OpenAlex,
+Bielefeld) was performed to identify hybrid journals. hoaddata relies on
+Crossref for obtaining journal publication volume and open access status
+through Creative Commons licence information relative to the published
+version (``version of record''). Because of limited affiliation metadata
+in Crossref (Eck \& Waltman, 2022), hoaddata sources first-author
+affiliations from OpenAlex.
 
@@
 article types, open access information as well as author roles and
-institutional affiliations.
-
-The publication year for each article was determined using the earliest
-known publication date. For articles in hoaddata, this corresponded to
-Crossref's issued date.
-
-Publication years were determined using the earliest known date of
-publication in a journal. In the case of Crossref, the article data
-source for hoaddata, this was the issued date. For Web of Science and
+institutional affiliations. Publication years were determined using the
+earliest known date of publication in a journal. articles in hoaddata,
+this corresponded to Crossref's issued date. For Web of Science and
 Scopus, the earliest publication date was used where available, with
@@
 datasets on open access in hybrid journals included in transformative
-agreements were produced for each database at country and journal level
-by year. Table~\ref{tbl-methods_overview_tab} provides a general
+agreements were aggregated for each database at country and journal
+level by year. Table~\ref{tbl-methods_overview_tab} provides a general
 overview of the coverage between 2019 and 2023 per database. It shows
 that the majority of hybrid journals published at least one original
-article or review marked as core during the five-year period. These
-journals formed the basis for the subsequent aggregation of publication
-metrics.
+research article or review marked as core during the five-year period.
+These journals formed the basis for the subsequent aggregation of
+publication metrics.
 
@@
 
-While hoaddata only covers articles with a DOI, the Scopus and Web of
-Science figures were also aggregated using the database identifier. A
-subsequent comparison of DOI coverage shows that non-core articles in
-Web of Science often lacked a DOI. This was particularly the case for
-meeting abstracts, a particular feature of this database. Open access
-indicators were aggregated by DOI, as Unpaywall only collects
-information on open access status for articles with a DOI. A closer look
-at the core articles with affiliation data reveals a lack of data on
-corresponding authors in the case of OpenAlex, the affiliation data
-source used by hoaddata, compared to Web of Science and Scopus. Only
-about two-thirds of the articles examined provided corresponding author
-affiliations. For first authors, the proportion was 89\%. Accordingly,
-only first author data for hoaddata are considered in the following
-analysis.
+While hoaddata only covers articles with a DOI, Scopus and Web of
+Science publication metrics were aggregated using the database
+identifier. A subsequent comparison of DOI coverage shows that non-core
+articles in Web of Science often lacked a DOI. This was particularly the
+case for meeting abstracts, which are notably prevalent in Health
+Sciences journals (Melero-Fuentes et al., 2025). Open access indicators
+were aggregated by DOI, as Unpaywall only collects information on open
+access status for articles with a DOI. A closer look at the core
+articles with affiliation data reveals a lack of data on corresponding
+authors in the case of OpenAlex, the affiliation data source used by
+hoaddata, compared to Web of Science and Scopus. Only about two-thirds
+of the articles examined provided corresponding author affiliations. For
+first authors, the proportion was 89\%. Therefore, only first author
+data for hoaddata were considered in the following analysis.
 
 \section{Results}\label{results}
+
+First, I will present a coverage analysis. Then, I investigate hybrid
+open access uptake using the same methods for each database and the
+compare results against each other in order to assess the suitablility
+of the biblioemtric databases to examine transofrmative agreements.
+
+\subsection{Coverage comparision}\label{coverage-comparision}
+
+Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid
+journals included in transformative agreements, visualising the
+intersections of journals and articles across the examined databases as
+an UpSet graph (Krassowski, 2020; Lex et al., 2014). The analysis
+included hybrid journals that published at least one open access article
+between 2019 and 2023, based on open access status information from each
+database. Only original research articles and reviews were considered in
+the analysis.
+
+\begin{figure}[ht!]
+
+\centering{
+
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results-1.png}
+
+}
+
+\caption{\label{fig-upset_coverage_results}Upset graph}
+
+\end{figure}%
+
+Journal coverage analysis revealed that 66\% (n = 7,970) of hybrid
+journals included in transformative agreements were indexed in all three
+databases ( \ref{fig-upset_coverage_results}A). The second-largest set
+consisted of journals indexed in both hoaddata and Scopus, comprising
+21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739) of journals
+were exclusively contained in hoaadata, while 3\% (n = 354) were only
+found in the proprietory databases Scopus and Web of Science. Upon
+insepction, this latter group represents hybrid journals for which no
+open access evidence could be retrieved from Crossref, which serves as
+the open access evidence source for hoaddata.
+
+In terms of article coverage, Fig B shows the DOI overlap,
+distinguihsing between a shared corpus and just those where a DOI was
+only available through hoaddata. Upon inspection, reasons for lackign
+DOI coverage in proprietory database were different publication dates
+and differenc ein document type definitions. As such, demosntrates
+certain uncertainies using the open approach. Overall, The majority of
+articles can be found
+
+Using DOIs also the size of journals were calculated (Fig C). It
+highlights thate large spread in the intersection between all three
+datasett. Journals are, on average, smaller than those just covered by
+one or two sources.In particular journals exlcusively covered by
+hoaddata are significant smaller than those covered by all three
+sources.
+
+This study compares results drawn from the openly available and
+regularly updated hoaddata, an R package leveraging the open scholarly
+data sources Crossref and OpenAlex, with the in-house database of the
+German Competence Network Overall, xxxx hybrid journals included in xxx
+agreements that published at least one open access article between 2019
+and 2023.
 
@@
 
+\bibitem[\citeproctext]{ref-ComplexUpset}
+Krassowski, M. (2020). \emph{ComplexUpset}.
+\url{https://doi.org/10.5281/zenodo.3700590}
+
+\bibitem[\citeproctext]{ref-Lex_2014}
+Lex, A., Gehlenborg, N., Strobelt, H., Vuillemot, R., \& Pfister, H.
+(2014). UpSet: Visualization of intersecting sets. \emph{IEEE
+Transactions on Visualization and Computer Graphics}, \emph{20}(12),
+1983--1992. \url{https://doi.org/10.1109/tvcg.2014.2346248}
+
 \bibitem[\citeproctext]{ref-Marwick_2018}
@@
 \url{https://doi.org/10.1080/00031305.2017.1375986}
+
+\bibitem[\citeproctext]{ref-Melero_Fuentes_2025}
+Melero-Fuentes, D., Aguilar-Moya, R., Valderrama-Zurián, J.-C., \&
+Gorraiz, J. (2025). Evolution and effect of meeting abstracts in JCR
+journals. \emph{Journal of Informetrics}, \emph{19}(1), 101631.
+\url{https://doi.org/10.1016/j.joi.2024.101631}
 

### 2025-02-18 analysis/manuscript_preprocessing_data.R

@@ -324,26 +324,48 @@ jn_upset_ <- dplyr::full_join(cr_active_upset, wos_active_upset, by = "issn_l") 
 scp_dois <- scp_active_core |>
   filter(issn_l %in% scp_active_core_oa$issn_l) |> 
   pull(doi)
-wos_dois <- scp_active_core |>
-  filter(issn_l %in% scp_active_core_oa$issn_l) |>
+wos_dois <- wos_active_core |>
+  filter(issn_l %in% wos_active_core_oa$issn_l) |>
+  pull(doi)
+cr_dois <- cr_df |>
+  filter(issn_l %in% active_jns_with_oa$issn_l) |>
   pull(doi)
 
 doi_overlap_df <- cr_df |>
-#  filter(issn_l %in% active_jns_with_oa$issn_l) |>
+ # filter(issn_l %in% active_jns_with_oa$issn_l) |>
   mutate(
-    in_wos_and_scopus = ifelse(doi %in% scp_dois & doi %in% wos_dois, 1, 0),
+    in_wos_or_scopus = ifelse(doi %in% c(scp_dois, wos_dois), 1, 0),
+    in_cr = ifelse(doi %in% cr_dois, 1, 0),
     in_scopus = ifelse(doi %in% scp_dois, 1, 0),
     in_wos =  ifelse(doi %in% wos_dois, 1, 0)
   )
 doi_overlap <- doi_overlap_df |>
   group_by(issn_l) |>
   summarise(
-    doi_in_wos_and_scopus = sum(in_wos_and_scopus),
+    doi_in_wos_or_scopus = sum(in_wos_or_scopus),
     doi_in_scopus = sum(in_scopus),
-    doi_in_wos = sum(in_wos)
+    doi_in_wos = sum(in_wos),
+    doi_cr = sum(in_cr)
   )
 
 jn_upset_articles <- jn_upset_ |>
   left_join(doi_overlap, by = "issn_l")
 
 write_csv(jn_upset_articles, here::here("data/jn_upset_articles.csv"))
+
+# Check articles exclusively in crossref
+intersection_jns <- jn_upset_articles |>
+  filter(WoS == TRUE, Scopus == TRUE, Crossref == TRUE) |>
+  distinct(issn_l)
+  
+cr_df |> 
+  filter(issn_l %in% intersection_jns$issn_l) |>
+  filter(doi %in% c(wos_dois, scp_dois)) |>
+  distinct(doi)
+
[…]

### 2025-02-19 analysis/manuscript.qmd
Check missing dois

@@
 
-hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package comprising information about the uptake of hybrid open access since 2017 collected from several openly available data sources.
+hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package collecting information on hybrid open access uptake from multiple openly available data sources.
 It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.
@@
 -->
-Article metadata was retrieved from each databse using all ISSN variants linked to an ISSN-L.
+Article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L.
 The metadata included DOIs, publication dates, article types, open access information as well as author roles and institutional affiliations.
@@
 A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
-This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025].
+This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
 Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
@@
     "Journal size\n(log scale)" = (
-      ggplot(mapping = aes(x = intersection, y = n)) +
+      ggplot(mapping = aes(x = intersection, y = overlap)) +
         geom_boxplot(outliers = FALSE, width = bar_width) +
@@
           width = bar_width,
-          aes(fill = "All")
+          aes(fill = "Surplus")
         ) +
@@
           values = c(
-            `All` = "grey80",
+            `Surplus` = "grey80",
             `Shared` = "#0091c8"
@@
 ```
-Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases ( \ref{fig-upset_coverage_results}A).
+Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases (Figure \ref{fig-upset_coverage_results}A).
 The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
-Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while `r round( jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory databases Scopus and Web of Science.
-Upon insepction, this latter group mainly represents hybrid journals for which no open access evidence could be retrieved from Crossref, which serves as the open access evidence source for hoaddata.
+Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus.
+Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in the Web of Science.
+Upon inspection, this group of hybrid journals included only in the proprietary data sources mainly represents hybrid journals for which no open access evidence could be retrieved from Crossref, which served as the open access evidence source for hoaddata.
 
-In terms of article coverage, Fig B shows the DOI overlap, distinguihsing between a shared corpus and just those where a DOI was only available through hoaddata.
-Upon inspection, reasons for lackign DOI coverage in proprietory database were different publication dates and differenc ein document type definitions. 
-As such, demosntrates certain uncertainies using the open approach.
-Overall, The majority of articles can be found
+In terms of article coverage, Figure \ref{fig-upset_coverage_results}B shows the total publication volume per combination in terms of DOI availability.
+The largest set of hybrid journals, which includes all three data sources, also contains the largest number of articles. 
+In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` overlapping articles, represented by the blue bar.
+They represent `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(wos) * 100)`% of core articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(scp) * 100)`% in Scopus.
+Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to both Scopus and hoaddata.
+Exclusively in hoaddata were `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == FALSE) |> pull(n_cr) |> sum()` articles, and exclusively in the proprietary databases were `r upset_data |> filter(Crossref == FALSE, Scopus == TRUE) |> pull(overlap) |> sum(na.rm = TRUE)` articles.
+
+Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
+After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews in hoaddata.
+Particulalry, Letters and Editorials could not be fully detected.
+Paratext recognition also failed in some cases to identify non-scientific content such as journal covers or reviewer lists, which are generally not covered by Scopus and Web of Science. 
+To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.
 
@@
 
-This study compares results drawn from the openly available and regularly updated hoaddata, an R package leveraging the open scholarly data sources Crossref and OpenAlex, with the in-house database of the German Competence Network
-Overall, xxxx hybrid journals included in xxx agreements that published at least one open access article between 2019 and 2023.
-
 # discussion

### 2025-02-19 analysis/manuscript.tex
Check missing dois

@@
 hoaddata, developed and maintained by the author to support open access
-monitoring and research (Jahn, 2025), is an R data package comprising
-information about the uptake of hybrid open access since 2017 collected
-from several openly available data sources. It combines article-level
-metadata from Crossref and OpenAlex with transformative agreement
-information from the cOAlition S Journal Checker Tool (JCT), which links
-journal and institutional data to agreements in the ESAC registry.
+monitoring and research (Jahn, 2025), is an R data package collecting
+information on hybrid open access uptake from multiple openly available
+data sources. It combines article-level metadata from Crossref and
+OpenAlex with transformative agreement information from the cOAlition S
+Journal Checker Tool (JCT), which links journal and institutional data
+to agreements in the ESAC registry.
 
@@
 
-Article metadata was retrieved from each databse using all ISSN variants
-linked to an ISSN-L. The metadata included DOIs, publication dates,
-article types, open access information as well as author roles and
-institutional affiliations. Publication years were determined using the
-earliest known date of publication in a journal. articles in hoaddata,
-this corresponded to Crossref's issued date. For Web of Science and
-Scopus, the earliest publication date was used where available, with
-Scopus dates specifically determined by the KB through version tracking
-of the raw data.
+Article metadata were retrieved from each database using all ISSN
+variants linked to an ISSN-L. The metadata included DOIs, publication
+dates, article types, open access information as well as author roles
+and institutional affiliations. Publication years were determined using
+the earliest known date of publication in a journal. articles in
+hoaddata, this corresponded to Crossref's issued date. For Web of
+Science and Scopus, the earliest publication date was used where
+available, with Scopus dates specifically determined by the KB through
+version tracking of the raw data.
 
@@
 case for meeting abstracts, which are notably prevalent in Health
-Sciences journals (Melero-Fuentes et al., 2025). Open access indicators
-were aggregated by DOI, as Unpaywall only collects information on open
-access status for articles with a DOI. A closer look at the core
-articles with affiliation data reveals a lack of data on corresponding
-authors in the case of OpenAlex, the affiliation data source used by
-hoaddata, compared to Web of Science and Scopus. Only about two-thirds
-of the articles examined provided corresponding author affiliations. For
-first authors, the proportion was 89\%. Therefore, only first author
-data for hoaddata were considered in the following analysis.
+Sciences journals (Melero-Fuentes et al., 2025) and are not indexed by
+Scopus (Donner, 2017). Open access indicators were aggregated by DOI, as
+Unpaywall only collects information on open access status for articles
+with a DOI. A closer look at the core articles with affiliation data
+reveals a lack of data on corresponding authors in the case of OpenAlex,
+the affiliation data source used by hoaddata, compared to Web of Science
+and Scopus. Only about two-thirds of the articles examined provided
+corresponding author affiliations. For first authors, the proportion was
+89\%. Therefore, only first author data for hoaddata were considered in
+the following analysis.
 
@@
 journals included in transformative agreements were indexed in all three
-databases ( \ref{fig-upset_coverage_results}A). The second-largest set
-consisted of journals indexed in both hoaddata and Scopus, comprising
-21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739) of journals
-were exclusively contained in hoaadata, while 3\% (n = 354) were only
-found in the proprietory databases Scopus and Web of Science. Upon
-insepction, this latter group represents hybrid journals for which no
-open access evidence could be retrieved from Crossref, which serves as
-the open access evidence source for hoaddata.
+databases (Figure \ref{fig-upset_coverage_results}A). The second-largest
+set consisted of journals indexed in both hoaddata and Scopus,
+comprising 21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739)
+of journals were exclusively contained in hoaadata, while 6\% (n = 748)
+were only found in the proprietory database Scopus. Of these, 354 were
+also available in the Web of Science. Upon inspection, this group of
+hybrid journals included only in the proprietary data sources mainly
+represents hybrid journals for which no open access evidence could be
+retrieved from Crossref, which served as the open access evidence source
+for hoaddata.
 
-In terms of article coverage, Fig B shows the DOI overlap,
-distinguihsing between a shared corpus and just those where a DOI was
-only available through hoaddata. Upon inspection, reasons for lackign
-DOI coverage in proprietory database were different publication dates
-and differenc ein document type definitions. As such, demosntrates
-certain uncertainies using the open approach. Overall, The majority of
-articles can be found
+In terms of article coverage, Figure \ref{fig-upset_coverage_results}B
+shows the total publication volume per combination in terms of DOI
+availability. The largest set of hybrid journals, which includes all
+three data sources, also contains the largest number of articles. In
+total, these journals recorded 6,289,687 overlapping articles,
+represented by the blue bar. They represent 94\% of core articles with
+DOI indexed in the Web of Science sample, and 86\% in Scopus. Another
+657,697 articles are exclusive to both Scopus and hoaddata. Exclusively
+in hoaddata were 177,110 articles, and exclusively in the proprietary
+databases were 325,194 articles.
+
+Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only
+available via hoaddata (grey area). After validation at the DOI level
+using the KB databases and manual inspection, the main reasons for
+missing DOI coverage in the proprietary database were insufficient
+classification of journal content as original research articles and
+reviews in hoaddata. Paratext recognition also failed in some cases to
+identify non-scientific content such as journal covers or reviewer
+lists, which are generally not covered by Scopus and Web of Science. To
+a lesser extent, differences in publication and indexing dates were a
+reason for non-overlapping DOIs.
 
@@
 sources.
-
-This study compares results drawn from the openly available and
-regularly updated hoaddata, an R package leveraging the open scholarly
-data sources Crossref and OpenAlex, with the in-house database of the
-German Competence Network Overall, xxxx hybrid journals included in xxx
-agreements that published at least one open access article between 2019
-and 2023.
 
@@
 \url{https://doi.org/10.1162/qss_a_00018}
+
+\bibitem[\citeproctext]{ref-Donner_2017}
+Donner, P. (2017). Document type assignment accuracy in the journal
+citation index data of {Web of Science}. \emph{Scientometrics},
+\emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}
 

### 2025-02-19 data-raw/missing_dois_check.R

@@ -1,0 +1,43 @@
+## Check missing dois in KB
+library(DBI)
+library(RPostgres)
+library(tidyverse)
+#' Connect to DB 
+#' 
+#' KB Postgres (Web of Science, Scopus)
+kb_con <- dbConnect(RPostgres::Postgres(),
+                    host = "biblio-p-db03.fiz-karlsruhe.de",
+                    port = 6432,
+                    dbname = "kbprod",
+                    user =  Sys.getenv("kb_user"),
+                    password = Sys.getenv("kb_pwd"),
+                    bigint = "numeric"
+)
+miss_dois <- readr::read_csv(here::here("data-raw", "missing_dois_in_active_jns.csv"))
+
+dbWriteTable(kb_con, "miss_dois", miss_dois, overwrite = TRUE)
+
+miss_dois_counts <- dbGetQuery(kb_con, "select
+	count(distinct md.doi) as n_doi,
+	i.item_type as scp_item_type,
+	i2.item_type as wos_item_type,
+	i.first_pubyear,
+	i.pubyear,
+	i2.pubyear as scp_year
+from
+	miss_dois md
+left join scp_b_202404.items i on
+	md.doi = i.doi
+left join wos_b_202304.v_items i2 on
+	md.doi = i2.doi
+group by
+	i.item_type,
+	i2.item_type,
+	i.first_pubyear,
+	i.pubyear,
+	i2.pubyear")
+
+miss_dois_counts |>
+  as_tibble() |>
+  arrange(desc(n_doi)) |>
+  write_csv(here::here("data-raw", "miss_dois_in_active_jns_ckeck.csv"))

### 2025-02-19 analysis/manuscript.qmd
Starting publisher analysis. Focus: overrep of big 3, newly launched springer hybrid oa only in hoaddata, lacking corssref license practise in some other publishers (name and shame)

@@
 
+### Overview
+
 ```{r}
@@
 
+```{r miss_dois_check}
+miss_dois <- readr::read_csv(here::here("data-raw", "missing_dois_in_active_jns.csv"))
+miss_dois_val <- readr::read_csv(here::here("data-raw", "miss_dois_in_active_jns_ckeck.csv"))
+```
 Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
+In case of hybrid ournals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
 After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews in hoaddata.
-Particulalry, Letters and Editorials could not be fully detected.
-Paratext recognition also failed in some cases to identify non-scientific content such as journal covers or reviewer lists, which are generally not covered by Scopus and Web of Science. 
+Particulalry, letters and editorials could not be fully detected.
+Paratext recognition failed in `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scientific content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science.
 To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.
 
-Using DOIs also the size of journals were calculated (Fig C).
-It highlights thate large spread in the intersection between all three datasett.
-Journals are, on average, smaller than those just covered by one or two sources.In particular journals exlcusively covered by hoaddata are significant smaller than those covered by all three sources.
+Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
+Figure \ref{fig-upset_coverage_results}C shows the distribution for each combination. It shows that there is a large spread across the journals covered by all three data sources. 
+Furthermore, journals in this set published more on average than journals covered by only one or two data sources. 
+In particular, the journals covered exclusively by hoaddata are substantially smaller than those covered by all three sources. 
+Upon inspection, these are often newly launched journals, which explains the relatively low five-year publication volume.
+
+### Coverage by publisher portfolio
+
+```{r publisher_coverage_data_prep}
+# Add ESAC portfolio names
+esac <- hoaddata::jct_hybrid_jns |>
+  distinct(issn_l, esac_publisher) |>
+  mutate(
+    publisher = case_when(
+      esac_publisher == "Elsevier" ~ "Elsevier",
+      esac_publisher == "Springer Nature" ~ "Springer Nature",
+      esac_publisher == "Wiley" ~ "Wiley",
+      is.character(esac_publisher) ~ "Other"
+    )) |>
+  mutate(publisher =
+           forcats::fct_relevel(
+             publisher,
+             c("Elsevier", "Springer Nature", "Wiley", "Other") 
+           )) |>
+  mutate(publisher = forcats::fct_rev(publisher)) |>
+  distinct(issn_l, publisher)
+
+jn_upset_publisher <- inner_join(upset_data, esac, by = "issn_l") |>
+  # Because of Journal transfer, restrict choice to one portfolio
+  distinct(issn_l, .keep_all = TRUE)
+```
+
+Figre shows the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry.
+
+```{r upset_publisher_plot, fig.asp=0.75}
+size = get_size_mode('exclusive_intersection')
+
+upset_publisher_plot <- ComplexUpset::upset(
+  jn_upset_publisher,
+  c("Crossref", "WoS", "Scopus"),
+  # Only journals in Crossref
+  min_size = 30,
+  width_ratio = 0.4,
+  stripes = "transparent",
+  name = "Data Source Coverage",
+  base_annotations = NULL,
+  annotations = list(
+    "Journal size\n(log-scale)" = (
+      ggplot(mapping = aes(x = intersection, y = overlap, fill = fct_rev(publisher))) +
+        geom_boxplot(outliers = FALSE, width = bar_width) +
+        scale_fill_manual(
+          "",
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3"
+          ), guide = "none") +
+        scale_y_log10(labels = scales::number_format(big.mark = ",")) +
+        ggtitle("C")
+    ),
+    "Articles" = (
+      ggplot(mapping = aes(y = overlap, fill = publisher, color = publisher)) +
+        geom_bar(stat = "identity", width = bar_width) +
+        scale_fill_manual(
+          "",
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3a0"
+          ), guide = "none") +
+        scale_color_manual(
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3"
+          ), guide = "none"
+        ) + 
+        scale_y_continuous(labels =  scales::number_format(big.mark = ",")) +
+        ggtitle("B") 
+    ),
+    "Publisher" = (
+      ggplot(mapping = aes(fill = publisher)) +
+        geom_bar(stat = "count", position = "fill", width = bar_width) + 
+        scale_y_continuous("Journals", labels = scales::percent_format()) +
+        scale_fill_manual(
+          "",
+          values = c(
+            "Elsevier" = "#e9711c",
+            "Springer Nature" = "#486a7e",
+            "Wiley" = "#068853",
+            "Other" = "#b3b3b3"
+          ), guide = "none") +
+        ggtitle("A")
+    )
+  ),
+  set_sizes = (
+    upset_set_size(
+      geom = geom_bar(aes(fill = publisher), width = 0.8), 
+      position = "right"
+    ) +
+    scale_fill_manual(
+      "",
+      values = c(
+        "Elsevier" = "#e9711c",
+        "Springer Nature" = "#486a7e",
+        "Wiley" = "#068853",
+        "Other" = "#b3b3b3"
+      )
+    ) +
+    scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
+    guides(fill = guide_legend(reverse = TRUE)) +
+    ylab("Journals")
+  ),
+  guides = "over",
+  matrix = (
+    intersection_matrix(
+      geom = geom_point(
+        shape = "square",
+        size = 2
+      )
+    )
+  )
+)
+
+upset_publisher_plot & my_upset_theme 
+```
 

### 2025-02-19 analysis/manuscript.tex
Starting publisher analysis. Focus: overrep of big 3, newly launched springer hybrid oa only in hoaddata, lacking corssref license practise in some other publishers (name and shame)

@@
 
+\subsubsection{Overview}\label{overview}
+
 Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid
@@
 DOI indexed in the Web of Science sample, and 86\% in Scopus. Another
-657,697 articles are exclusive to both Scopus and hoaddata. Exclusively
+657,697 articles were exclusive to both Scopus and hoaddata. Exclusively
 in hoaddata were 177,110 articles, and exclusively in the proprietary
@@
 Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only
-available via hoaddata (grey area). After validation at the DOI level
-using the KB databases and manual inspection, the main reasons for
-missing DOI coverage in the proprietary database were insufficient
-classification of journal content as original research articles and
-reviews in hoaddata. Paratext recognition also failed in some cases to
-identify non-scientific content such as journal covers or reviewer
-lists, which are generally not covered by Scopus and Web of Science. To
-a lesser extent, differences in publication and indexing dates were a
-reason for non-overlapping DOIs.
+available via hoaddata (grey area). In case of hybrid ournals covered by
+all three data sources, 1,023,882 DOIs were only present in hoaddata.
+After validation at the DOI level using the KB databases and manual
+inspection, the main reasons for missing DOI coverage in the proprietary
+database were insufficient classification of journal content as original
+research articles and reviews in hoaddata. Particulalry, letters and
+editorials could not be fully detected. Paratext recognition failed in
+37\% of DOIs to identify non-scientific content such as front matters or
+reviewer lists, which are generally not indexed by Scopus and Web of
+Science. To a lesser extent, differences in publication and indexing
+dates were a reason for non-overlapping DOIs.
 
-Using DOIs also the size of journals were calculated (Fig C). It
-highlights thate large spread in the intersection between all three
-datasett. Journals are, on average, smaller than those just covered by
-one or two sources.In particular journals exlcusively covered by
-hoaddata are significant smaller than those covered by all three
-sources.
+Using overlapping DOIs, the publication volume between 2019 and 2023 was
+also calculated for each journal. Figure
+\ref{fig-upset_coverage_results}C shows the distribution for each
+combination. It shows that there is a large spread across the journals
+covered by all three data sources. Furthermore, journals in this set
+published more on average than journals covered by only one or two data
+sources. In particular, the journals covered exclusively by hoaddata are
+substantially smaller than those covered by all three sources. Upon
+inspection, these are often newly launched journals, which explains the
+relatively low five-year publication volume.
+
+\subsubsection{Coverage by publisher
+portfolio}\label{coverage-by-publisher-portfolio}
+
+Figre shows the coverage of hybrid journals in transformative agreements
+by publisher portfolio according to the ESAC registry.
+
+\begin{center}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/upset_publisher_plot-1.png}
+\end{center}
 

### 2025-02-19 analysis/manuscript.qmd
Use pdf figures in latex output

@@
   fig.align = "center",
-  dpi = 300,
-  dev = c("png", "cairo_ps", "pdf"),
+  dpi = 600,
+  dev = c("pdf", "cairo_ps", "png"),
   fig.path = "fig/",

### 2025-02-19 analysis/manuscript.tex
Use pdf figures in latex output

@@
 
-\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results-1.png}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results-1.pdf}
 
@@
 \begin{center}
-\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/upset_publisher_plot-1.png}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/upset_publisher_plot-1.pdf}
 \end{center}

### 2025-02-25 analysis/manuscript.qmd
tmp save

@@
         geom_boxplot(outliers = FALSE, width = bar_width) +
-        scale_y_log10(labels =  scales::number_format(big.mark = ",")) +
+        scale_y_log10(labels =  scales::number_format(big.mark = ","), limits = c(1, 10000)) +
         ggtitle("C")
@@
                            breaks = c(0, 2500000, 5000000, 7500000)) +
+        guides(fill = guide_legend(reverse = TRUE)) +
         ggtitle("B")
@@
   mutate(publisher = forcats::fct_rev(publisher)) |>
-  distinct(issn_l, publisher)
+  distinct(issn_l, publisher, esac_publisher)
 
@@
 
-Figre shows the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry.
+Figure \ref{fig-upset_coverage_results_publisher} shows the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry.
 
 ```{r upset_publisher_plot, fig.asp=0.75}
+#| fig-cap: "Upset graph publisher"
+#| label: fig-upset_coverage_results_publisher
+#| 
 size = get_size_mode('exclusive_intersection')

### 2025-02-25 analysis/manuscript.tex
tmp save

@@
 
-Figre shows the coverage of hybrid journals in transformative agreements
-by publisher portfolio according to the ESAC registry.
+Figure \ref{fig-upset_coverage_results_publisher} shows the coverage of
+hybrid journals in transformative agreements by publisher portfolio
+according to the ESAC registry.
 
-\begin{center}
-\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/upset_publisher_plot-1.pdf}
-\end{center}
+\begin{figure}[ht!]
+
+\centering{
+
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-upset_coverage_results_publisher-1.pdf}
+
+}
+
+\caption{\label{fig-upset_coverage_results_publisher}Upset graph
+publisher}
+
+\end{figure}%
 

### 2025-03-04 analysis/manuscript.qmd
Continuing results: publisher analysis

@@
         plot.title.position="plot",
-        plot.title = element_text(hjust = 0, vjust = 1, face = "bold"))
+        text = element_text(size = 10),
+        plot.title = element_text(hjust = 0, vjust = 1))
 ```
@@
 
-This study aims to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
+This study investigates the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
 As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
-This section introduces the initial data sources followed by a presentation of the necessary pre-processing steps to obtain eligible articles from transformative agreements by using author roles (first and corresponding) and harmonised affiliation data.
+This section introduces the initial data sources, followed by a presentation of the necessary data processing steps to obtain eligible articles from transformative agreements by using author roles (first and corresponding) and harmonised affiliation data.
 <!-- Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study. -->
@@
 
-As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of historic JCT snapshots since July 2021, which were enriched with ISSN variants linked to an ISSN-L and associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
+As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of historic JCT snapshots, which were safeguarded since July 2021.
+Journal data were enriched with ISSN variants linked to an ISSN-L and insitution data with associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
 A comprehensive exclusion of fully open access journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals.
 hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
-Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources first-author affiliations from OpenAlex.
+Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources  affiliation metadata from OpenAlex.
 
@@
 The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
-To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow interfaces with the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
+To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow handles data retrieval from the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
 The workflow has run regularly to fetch updates from these data sources since 2022.
@@
 
-The WoS Primary provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as Openalex.
+The WoS Primary provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as OpenAlex.
 However, it lacks information about journals and articles under transformative agreements.
@@
 Publication years were determined using the earliest known date of publication in a journal.
-articles in hoaddata, this corresponded to Crossref's issued date.<!--, although variations were observed (check ref)-->
+In hoaddata, this corresponded to Crossref's issued date.<!--, although variations were observed (check ref)-->
 For Web of Science and Scopus, the earliest publication date was used where available, with Scopus dates specifically determined by the KB through version tracking of the raw data.
@@
 Unpaywall supplements Crossref metadata by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018]. 
-This additional parsing remains necessary despite transformative agreement workflows requiring license information during DOI registration [@Geschuhn_2017].
+This additional parsing remains necessary despite transformative agreement workflows recommend the depositing license information during DOI registration [@Geschuhn_2017].
 Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information.
@@
 
-This process linked `r length(unique(scp_ror$ror))` ROR IDs to  `r length(unique(scp_ror$vendor_org_id))` Scopus Affiliation IDs, and  `r length(unique(wos_ror$ror))` ROR IDs to `r length(unique(scp_ror$vendor_org_id))` enhanced affiliation strings in the Web of Science.
+This process linked `r length(unique(scp_ror$ror))` ROR IDs to  `r length(unique(scp_ror$vendor_org_id))` Scopus Affiliation IDs, and  `r length(unique(wos_ror$ror))` ROR IDs to `r length(unique(wos_ror$vendor_org_id))` enhanced affiliation strings in the Web of Science.
 Quality evaluation through random sampling of 50 pairs revealed an error rate of 22% for Web of Science (11 mismatches) and 6% for Scopus (3 mismatches). 
@@
 It shows that the majority of hybrid journals published at least one original research article or review marked as core during the five-year period.
-These journals formed the basis for the subsequent aggregation of publication metrics.
+These journals formed the basis for the subsequent calculation of article-level indicators.
 
@@
   row_spec(0, bold = TRUE) |>
-  pack_rows("Hybrid journal metrics", 1, 3) |>
-  pack_rows("Publication metrics", 4, 5) |>
+  pack_rows("Hybrid journal coverage", 1, 3) |>
+  pack_rows("Publication volume", 4, 5) |>
   pack_rows("Digital Object Identifier (DOI) coverage", 6, 7) |>
@@
 
-While hoaddata only covers articles with a DOI, Scopus and Web of Science publication metrics were aggregated using the database identifier.
+While hoaddata only covers articles with a DOI, Scopus and Web of Science publication indicators were calculated using their database identifier.
 A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
@@
         geom_boxplot(outliers = FALSE, width = bar_width) +
-        scale_y_log10(labels =  scales::number_format(big.mark = ","), limits = c(1, 10000)) +
+        scale_y_log10(labels =  scales::number_format(big.mark = ","), limits = c(5, 10000)) +
         ggtitle("C")
@@
 ```
+
 Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
-In case of hybrid ournals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
-After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews in hoaddata.
+In case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
+After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews during the compilation of hoaddata.
 Particulalry, letters and editorials could not be fully detected.
@@
 Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
-Figure \ref{fig-upset_coverage_results}C shows the distribution for each combination. It shows that there is a large spread across the journals covered by all three data sources. 
+Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows that there is a large spread across the journals covered by all three data sources. 
 Furthermore, journals in this set published more on average than journals covered by only one or two data sources. 
@@
   distinct(issn_l, .keep_all = TRUE)
+
+jn_upset_publisher_ind <- 
+  jn_upset_publisher |> 
+  filter(Crossref == TRUE, Scopus == TRUE, WoS == TRUE) |> 
+  group_by(publisher) |> 
+  summarise(n_journals = n_distinct(issn_l), articles = sum(overlap))
 ```
 
-Figure \ref{fig-upset_coverage_results_publisher} shows the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry.
+Figure \ref{fig-upset_coverage_results_publisher} illustrates the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry between 2019 and 2023.
+Elsevier was best represented in the union set of all investigated databases, with `r jn_upset_publisher_ind |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals with at least one open access article that published `r jn_upset_publisher_ind |> filter(publisher == "Elsevier") |> pull(articles)` articles between the five-year period.
+Springer Nature followed with `r jn_upset_publisher_ind |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals publishing `r jn_upset_publisher_ind |> filter(publisher == "Springer Nature") |> pull(articles)` articles, and Wiley with `r jn_upset_publisher_ind |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals and `r jn_upset_publisher_ind |> filter(publisher == "Wiley") |> pull(articles)` articles.
+The remaining `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(esac_publisher) |> nrow()` publishers recorded  a total of `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(issn_l) |> nrow()` hybrid journals that published  `r jn_upset_publisher_ind |> filter(publisher == "Other") |> pull(articles)` articles.
+
+Notably, the three largest publishers--Elsevier, Springer Nature and Wiley--were well represented in the exclusive intersection of all three data sources examined.
+In terms of journal coverage, they accounted for `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`%) of the hybrid journals included in the intersection.
+This dominance is even more pronounced in terms of article coverage, with these publishers having a joint indexing coverage of `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()` articles (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as measured by shared DOIs.
 
@@
 
+When considering journals exlcusively covered by Scopus and Web of Science
+
+When considering journals exlcusively covered by hoaddata
+
 # discussion

### 2025-03-04 analysis/manuscript.tex
Continuing results: publisher analysis

@@
 
-This study aims to demonstrate the suitability of open scholarly data
-sources for assessing the impact of transformative agreements on hybrid
-open access. As shown in Figure \ref{fig:workflow}, the methodology
-involved comparing hoaddata, an openly available collection of open
-research information on hybrid open access, with the bibliometric
-databases Web of Science and Scopus. This section introduces the initial
-data sources followed by a presentation of the necessary pre-processing
-steps to obtain eligible articles from transformative agreements by
-using author roles (first and corresponding) and harmonised affiliation
-data.
+This study investigates the suitability of open scholarly data sources
+for assessing the impact of transformative agreements on hybrid open
+access. As shown in Figure \ref{fig:workflow}, the methodology involved
+comparing hoaddata, an openly available collection of open research
+information on hybrid open access, with the bibliometric databases Web
+of Science and Scopus. This section introduces the initial data sources,
+followed by a presentation of the necessary data processing steps to
+obtain eligible articles from transformative agreements by using author
+roles (first and corresponding) and harmonised affiliation data.
 
@@
 As discussed elsewhere (Jahn, 2025), hoaddata uses a unified dataset of
-historic JCT snapshots since July 2021, which were enriched with ISSN
-variants linked to an ISSN-L and associated ROR-IDs representing
-university hospitals or institutes of large research organisations,
-according to OpenAlex' institution entity. A comprehensive exclusion of
-fully open access journals using multiple journal lists (DOAJ, OpenAlex,
-Bielefeld) was performed to identify hybrid journals. hoaddata relies on
-Crossref for obtaining journal publication volume and open access status
-through Creative Commons licence information relative to the published
-version (``version of record''). Because of limited affiliation metadata
-in Crossref (Eck \& Waltman, 2022), hoaddata sources first-author
-affiliations from OpenAlex.
+historic JCT snapshots, which were safeguarded since July 2021. Journal
+data were enriched with ISSN variants linked to an ISSN-L and insitution
+data with associated ROR-IDs representing university hospitals or
+institutes of large research organisations, according to OpenAlex'
+institution entity. A comprehensive exclusion of fully open access
+journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld) was
+performed to identify hybrid journals. hoaddata relies on Crossref for
+obtaining journal publication volume and open access status through
+Creative Commons licence information relative to the published version
+(``version of record''). Because of limited affiliation metadata in
+Crossref (Eck \& Waltman, 2022), hoaddata sources affiliation metadata
+from OpenAlex.
 
@@
 computational reproducibility while aggregating the data, a GitHub
-Actions continuous integration and delivery (CI/CD) workflow interfaces
-with the SUB Göttingen's open scholarly data warehouse based on Google
-BigQuery, which provides high-performant programmatic access to monthly
-snapshots of Crossref and OpenAlex. The workflow has run regularly to
-fetch updates from these data sources since 2022. The package version
-used in this study is 0.3, containing data from the Crossref 2024-08
-dump provided to Crossref Metadata Plus subscribers and the OpenAlex
-2024-08-29 monthly dump. This version including the computation log is
-available on GitHub
+Actions continuous integration and delivery (CI/CD) workflow handles
+data retrieval from the SUB Göttingen's open scholarly data warehouse
+based on Google BigQuery, which provides high-performant programmatic
+access to monthly snapshots of Crossref and OpenAlex. The workflow has
+run regularly to fetch updates from these data sources since 2022. The
+package version used in this study is 0.3, containing data from the
+Crossref 2024-08 dump provided to Crossref Metadata Plus subscribers and
+the OpenAlex 2024-08-29 monthly dump. This version including the
+computation log is available on GitHub
 (https://github.com/subugoe/hoaddata/releases/tag/v.0.3).
@@
 status information derived from OurResearch's Unpaywall, the same
-provider as Openalex. However, it lacks information about journals and
+provider as OpenAlex. However, it lacks information about journals and
 articles under transformative agreements.
@@
 and institutional affiliations. Publication years were determined using
-the earliest known date of publication in a journal. articles in
-hoaddata, this corresponded to Crossref's issued date. For Web of
-Science and Scopus, the earliest publication date was used where
-available, with Scopus dates specifically determined by the KB through
-version tracking of the raw data.
+the earliest known date of publication in a journal. In hoaddata, this
+corresponded to Crossref's issued date. For Web of Science and Scopus,
+the earliest publication date was used where available, with Scopus
+dates specifically determined by the KB through version tracking of the
+raw data.
 
@@
 additional parsing remains necessary despite transformative agreement
-workflows requiring license information during DOI registration
-(Geschuhn \& Stone, 2017). Both Web of Science and Scopus defined hybrid
-open access consistently as content available under Creative Commons
-licenses on publisher platforms, distinguishing it from bronze open
-access that lack such explicit license information.
+workflows recommend the depositing license information during DOI
+registration (Geschuhn \& Stone, 2017). Both Web of Science and Scopus
+defined hybrid open access consistently as content available under
+Creative Commons licenses on publisher platforms, distinguishing it from
+bronze open access that lack such explicit license information.
 
@@
 This process linked 6,375 ROR IDs to 4,894 Scopus Affiliation IDs, and
-6,034 ROR IDs to 4,894 enhanced affiliation strings in the Web of
+6,034 ROR IDs to 2,422 enhanced affiliation strings in the Web of
 Science. Quality evaluation through random sampling of 50 pairs revealed
@@
 research article or review marked as core during the five-year period.
-These journals formed the basis for the subsequent aggregation of
-publication metrics.
+These journals formed the basis for the subsequent calculation of
+article-level indicators.
 
@@
 \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Hybrid journal metrics}}\\
+\multicolumn{4}{l}{\textbf{Hybrid journal coverage}}\\
 \hspace{1em}Active journals & 12,890 & 8,655 & 11,888\\
@@
 \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Publication metrics}}\\
+\multicolumn{4}{l}{\textbf{Publication volume}}\\
 \hspace{1em}Total published articles & 9,740,015 & 8,616,053 & 8,117,644\\
@@
 While hoaddata only covers articles with a DOI, Scopus and Web of
-Science publication metrics were aggregated using the database
+Science publication indicators were calculated using their database
 identifier. A subsequent comparison of DOI coverage shows that non-core
@@
 Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only
-available via hoaddata (grey area). In case of hybrid ournals covered by
-all three data sources, 1,023,882 DOIs were only present in hoaddata.
+available via hoaddata (grey area). In case of hybrid journals covered
+by all three data sources, 1,023,882 DOIs were only present in hoaddata.
 After validation at the DOI level using the KB databases and manual
@@
 database were insufficient classification of journal content as original
-research articles and reviews in hoaddata. Particulalry, letters and
-editorials could not be fully detected. Paratext recognition failed in
-37\% of DOIs to identify non-scientific content such as front matters or
-reviewer lists, which are generally not indexed by Scopus and Web of
-Science. To a lesser extent, differences in publication and indexing
-dates were a reason for non-overlapping DOIs.
+research articles and reviews during the compilation of hoaddata.
+Particulalry, letters and editorials could not be fully detected.
+Paratext recognition failed in 37\% of DOIs to identify non-scientific
+content such as front matters or reviewer lists, which are generally not
+indexed by Scopus and Web of Science. To a lesser extent, differences in
+publication and indexing dates were a reason for non-overlapping DOIs.
 
@@
 also calculated for each journal. Figure
-\ref{fig-upset_coverage_results}C shows the distribution for each
+\ref{fig-upset_coverage_results}C illustrates the distribution for each
 combination. It shows that there is a large spread across the journals
@@
 
-Figure \ref{fig-upset_coverage_results_publisher} shows the coverage of
-hybrid journals in transformative agreements by publisher portfolio
-according to the ESAC registry.
+Figure \ref{fig-upset_coverage_results_publisher} illustrates the
+coverage of hybrid journals in transformative agreements by publisher
+portfolio according to the ESAC registry between 2019 and 2023. Elsevier
+was best represented in the union set of all investigated databases,
+with 1,672 hybrid journals with at least one open access article that
+published 2,345,205 articles between the five-year period. Springer
+Nature followed with 1,561 hybrid journals publishing 1,058,564
+articles, and Wiley with 1,151 hybrid journals and 770,546 articles. The
+remaining 54 publishers recorded a total of 6,476 hybrid journals that
+published 2,115,372 articles.
+
+Notably, the three largest publishers--Elsevier, Springer Nature and
+Wiley--were well represented in the exclusive intersection of all three
+data sources examined. In terms of journal coverage, they accounted for
+4,384 (55\%) of the hybrid journals included in the intersection. This
+dominance is even more pronounced in terms of article coverage, with
+these publishers having a joint indexing coverage of 4,174,315 articles
+(66\%).
 
@@
 \end{figure}%
+
+When considering journals exlcusively covered by Scopus and Web of
+Science
+
+When considering journals exlcusively covered by hoaddata
 

### 2025-03-05 analysis/manuscript.qmd
Finish first draft of coverage analysis, next replication

@@
   group_by(publisher) |>
-  summarise(n_journals = n_distinct(issn_l), articles = sum(overlap))
+  summarise(n_journals = n_distinct(issn_l), articles = sum(overlap, na.rm = TRUE))
+
+jn_upset_publisher_all <- 
+  jn_upset_publisher |>
+  group_by(publisher) |>
+  summarise(n_journals = n_distinct(issn_l), articles = sum(overlap, na.rm = TRUE))
+
+jn_upset_publisher_all <- jn_upset_publisher |>
+  summarise(n_journals_all = n_distinct(issn_l), articles_all = sum(overlap, na.rm = TRUE)) |>
+  bind_cols(jn_upset_publisher_all) |>
+  mutate(prop_jn = n_journals / n_journals_all,
+         prop_art = articles / articles_all)
 ```
 
-Figure \ref{fig-upset_coverage_results_publisher} illustrates the coverage of hybrid journals in transformative agreements by publisher portfolio according to the ESAC registry between 2019 and 2023.
-Elsevier was best represented in the union set of all investigated databases, with `r jn_upset_publisher_ind |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals with at least one open access article that published `r jn_upset_publisher_ind |> filter(publisher == "Elsevier") |> pull(articles)` articles between the five-year period.
-Springer Nature followed with `r jn_upset_publisher_ind |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals publishing `r jn_upset_publisher_ind |> filter(publisher == "Springer Nature") |> pull(articles)` articles, and Wiley with `r jn_upset_publisher_ind |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals and `r jn_upset_publisher_ind |> filter(publisher == "Wiley") |> pull(articles)` articles.
-The remaining `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(esac_publisher) |> nrow()` publishers recorded  a total of `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(issn_l) |> nrow()` hybrid journals that published  `r jn_upset_publisher_ind |> filter(publisher == "Other") |> pull(articles)` articles.
-
-Notably, the three largest publishers--Elsevier, Springer Nature and Wiley--were well represented in the exclusive intersection of all three data sources examined.
-In terms of journal coverage, they accounted for `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`%) of the hybrid journals included in the intersection.
-This dominance is even more pronounced in terms of article coverage, with these publishers having a joint indexing coverage of `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()` articles (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as measured by shared DOIs.
+Figure \ref{fig-upset_coverage_results_publisher} presents the coverage of hybrid journals in transformative agreements across data sources from 2019 to 2023 with a focus on publisher portfolios.
+The analysis highlights the dominance of the three largest publishers--Elsevier, Springer Nature, and Wiley--, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of articles published during this five-year period.
+In terms of article volume, Elsevier led with `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(articles)` artricles (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_art)*100`% of the total) published across `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals(`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_jn)*100`% of the total).
+Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%), while Wiley recorded `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
+The remaining `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(esac_publisher) |> nrow()` publishers collectively accounted for `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`%).
 
@@
 
-When considering journals exlcusively covered by Scopus and Web of Science
+The three largest publishers--Elsevier, Springer Nature and Wiley--were best represented in the exclusive intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
+Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage with `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()` (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as determined through shared DOIs.
+When examining publication volume per journal, (\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals, followed by Springer Nature and Wiley.
 
-When considering journals exlcusively covered by hoaddata
+A comparative analysis of publisher portfolios across different indexing sets demonstrates that publisher were not represented uniformly. 
+Springer Nature exhibited `r jn_upset_publisher |> filter(esac_publisher == "Springer Nature", WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow()` hybrid journals exclusively indexed in both hoaddata and Scopus.
+This set included journals from the Chinese Academy of Science, German-language medical journals, and Eastern European publications including the *Journal of Mathematical Sciences*, which also publishes English-language translations of Russian-language works.
+Additionally, this subset included titles with a more broader disciplianry focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
+The set also captured ceased journals, providing insights into the dynamic nature of journal portfolios and business strategies.
+
+```{r miss_portfolios}
+miss_tas <- jn_upset_publisher |> 
+  filter(Crossref == FALSE) |> 
+  group_by(esac_publisher) |> 
+  summarise(journals = n_distinct(issn_l),
+            n = sum(overlap, na.rm = TRUE)) |> 
+  arrange(desc(n))
+
+cr_tas <- jn_upset_publisher |>
+  filter(Crossref == TRUE) |> 
+  pull(esac_publisher)
+
+miss_publishers <- miss_tas |>
+  filter(!esac_publisher %in% cr_tas)
+```
+
+Further examination of publisher portfolios not covered by hoaddata but present in Scopus or Web of Science identified several publishers with missing Creative Commons (CC) license information in Crossref.
+In particular, `r miss_tas$esac_publisher[1]` represented `r miss_tas$journals[1]` journals with `r miss_tas$n[1]` articles, `r miss_tas$esac_publisher[2]` accounted for `r miss_tas$journals[2]` journals with `r miss_tas$n[2]` articles, and `r miss_tas$esac_publisher[3]` recorded `r miss_tas$journals[3]` journals and `r miss_tas$n[3]` articles.
+In total, `r nrow(miss_publishers)` publishers were not represented in Crossref.
+
+An inspection of individual journals uncovered discrepancies in Unpaywall's open access identification for certain publishers that typically deposit CC licenses with Crossref.
+Notably, some subscription-only journals contained one or two articles erroneously tagged as hybrid open access by Unpaywall, which were subsequently reflected in Scopus and Web of Science.
+Examples of such misclassifications include Elsevier's *Journal of Bioscience and Bioengineering* and Springer Nature's *Journal of Mechanical Science and Technology*.
 

### 2025-03-05 analysis/manuscript.tex
Finish first draft of coverage analysis, next replication

@@
 
-Figure \ref{fig-upset_coverage_results_publisher} illustrates the
-coverage of hybrid journals in transformative agreements by publisher
-portfolio according to the ESAC registry between 2019 and 2023. Elsevier
-was best represented in the union set of all investigated databases,
-with 1,672 hybrid journals with at least one open access article that
-published 2,345,205 articles between the five-year period. Springer
-Nature followed with 1,561 hybrid journals publishing 1,058,564
-articles, and Wiley with 1,151 hybrid journals and 770,546 articles. The
-remaining 54 publishers recorded a total of 6,476 hybrid journals that
-published 2,115,372 articles.
-
-Notably, the three largest publishers--Elsevier, Springer Nature and
-Wiley--were well represented in the exclusive intersection of all three
-data sources examined. In terms of journal coverage, they accounted for
-4,384 (55\%) of the hybrid journals included in the intersection. This
-dominance is even more pronounced in terms of article coverage, with
-these publishers having a joint indexing coverage of 4,174,315 articles
-(66\%).
+Figure \ref{fig-upset_coverage_results_publisher} presents the coverage
+of hybrid journals in transformative agreements across data sources from
+2019 to 2023 with a focus on publisher portfolios. The analysis
+highlights the dominance of the three largest publishers--Elsevier,
+Springer Nature, and Wiley--, which collectively accounted for 47\% of
+hybrid journals and 62\% of articles published during this five-year
+period. In terms of article volume, Elsevier led with 2,441,358
+artricles (33\% of the total) published across 1,951 hybrid
+journals(16\% of the total). Springer Nature followed with 1,247,578
+articles (17\%) in 2,311 hybrid journals (19\%), while Wiley recorded
+858,939 articles (12\%) in1,382 hybrid journals (11\%). The remaining 54
+publishers collectively accounted for 2,749,847 articles (38\%) in 6,476
+hybrid journals (53\%).
 
@@
 
-When considering journals exlcusively covered by Scopus and Web of
-Science
+The three largest publishers--Elsevier, Springer Nature and Wiley--were
+best represented in the exclusive intersection of all three data sources
+(hoaddata, Scopus, and Web of Science). Together, they comprised 4,384
+hybrid journals (55\% of the intersectional set) and dominated article
+coverage with 4,174,315 (66\%), as determined through shared DOIs. When
+examining publication volume per journal,
+(\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
+average, the largest journals, followed by Springer Nature and Wiley.
 
-When considering journals exlcusively covered by hoaddata
+A comparative analysis of publisher portfolios across different indexing
+sets demonstrates that publisher were not represented uniformly.
+Springer Nature exhibited 519 hybrid journals exclusively indexed in
+both hoaddata and Scopus. This set included journals from the Chinese
+Academy of Science, German-language medical journals, and Eastern
+European publications including the \emph{Journal of Mathematical
+Sciences}, which also publishes English-language translations of
+Russian-language works. Additionally, this subset included titles with a
+more broader disciplianry focus such as \emph{SN Computer Science} and
+newly launched hybrid journals like \emph{Nature Computational Science},
+which started in 2021 and was indexed in Scopus but not yet in Web of
+Science. The set also captured ceased journals, providing insights into
+the dynamic nature of journal portfolios and business strategies.
+
+Further examination of publisher portfolios not covered by hoaddata but
+present in Scopus or Web of Science identified several publishers with
+missing Creative Commons (CC) license information in Crossref. In
+particular, Emerald represented 322 journals with 86,409 articles, AIP
+Publishing accounted for 24 journals with 64,898 articles, and World
+Scientific recorded 87 journals and 42,531 articles. In total, 9
+publishers were not represented in Crossref.
+
+An inspection of individual journals uncovered discrepancies in
+Unpaywall's open access identification for certain publishers that
+typically deposit CC licenses with Crossref. Notably, some
+subscription-only journals contained one or two articles erroneously
+tagged as hybrid open access by Unpaywall, which were subsequently
+reflected in Scopus and Web of Science. Examples of such
+misclassifications include Elsevier's \emph{Journal of Bioscience and
+Bioengineering} and Springer Nature's \emph{Journal of Mechanical
+Science and Technology}.
 

### 2025-03-06 analysis/manuscript.qmd
Starting replication

@@
 
+## Replication
+
+This section examines the open access uptake in hybrid journals, focusing on the influence of transformative agreements across hoaddata, Scopus and Web of Science.
+A key objective of this analysis is to determine whether consistent results can be derived across these three data sources, despite their differing coverage and methodologies. 
+By identifying potential variations, this study evaluates the suitability of open data approaches, such as hoaddata, as alternatives to proprietary bibliometric databases like Scopus and Web of Science.
+
+Following @Jahn_2025, the impact of transformative agreements on hybrid open access is assessed by analysing the proportion of open access articles, both overall and those specifically enabled by transformative agreements.
+The analysis covers the publication years 2019 to 2023, providing insights into temporal trends in the adoption of transformative agreements and their influence on open access in hybrid journals.
+The analysis will present breakdowns by publishers and countries.
+
+A complicating factor in the comparision is the differing coverage of author affiliation data across the datasets.
+In hoaddata, which relies on OpenAlex for affiliation information, the lack of comprehensive corresponding author data restricts the analysis to first authors when estimating the influence of transformative agreements.
+In contrast, Scopus and Web of Science provide more robust coverage of corresponding author affiliations, enabling a comparative assessment of how authorship roles--first versus corresponding authors--influence estimates of open access funding through transformative agreements.
+
 # discussion

### 2025-03-06 analysis/manuscript.tex
Starting replication

@@
 
+\subsection{Replication}\label{replication}
+
+This section examines the open access uptake in hybrid journals,
+focusing on the influence of transformative agreements across hoaddata,
+Scopus and Web of Science. A key objective of this analysis is to
+determine whether consistent results can be derived across these three
+data sources, despite their differing coverage and methodologies. By
+identifying potential variations, this study evaluates the suitability
+of open data approaches, such as hoaddata, as alternatives to
+proprietary bibliometric databases like Scopus and Web of Science.
+
+Following Jahn (2025), the impact of transformative agreements on hybrid
+open access is assessed by analysing the proportion of open access
+articles, both overall and those specifically enabled by transformative
+agreements. The analysis covers the publication years 2019 to 2023,
+providing insights into temporal trends in the adoption of
+transformative agreements and their influence on open access in hybrid
+journals. The analysis will present breakdowns by publishers and
+countries.
+
+A complicating factor in the comparision is the differing coverage of
+author affiliation data across the datasets. In hoaddata, which relies
+on OpenAlex for affiliation information, the lack of comprehensive
+corresponding author data restricts the analysis to first authors when
+estimating the influence of transformative agreements. In contrast,
+Scopus and Web of Science provide more robust coverage of corresponding
+author affiliations, enabling a comparative assessment of how authorship
+roles--first versus corresponding authors--influence estimates of open
+access funding through transformative agreements.
+
 \section*{discussion}\label{discussion}

### 2025-03-10 analysis/manuscript.qmd
Trend anaylsis replication

@@
 
+```{r, rep_data}
+uptake_df <- readr::read_csv(here::here("data", "uptake_df.csv"))
+```
+
 This section examines the open access uptake in hybrid journals, focusing on the influence of transformative agreements across hoaddata, Scopus and Web of Science.
 A key objective of this analysis is to determine whether consistent results can be derived across these three data sources, despite their differing coverage and methodologies.
-By identifying potential variations, this study evaluates the suitability of open data approaches, such as hoaddata, as alternatives to proprietary bibliometric databases like Scopus and Web of Science.
+
+```{r, plot_uptake_overview, fig.asp=.5}
+uptake_df <- readr::read_csv(here::here("data", "uptake_df.csv"))
+
+# Data prep
+uptake_by_year <- uptake_df |>
+  mutate(earliest_year = gsub("^20", "'", as.character(earliest_year))) |>
+  group_by(earliest_year, src) |>
+  summarize(
+    n_articles = sum(n),
+    n_oa_articles = sum(oa_articles),
+    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
+    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
+  )
+
+uptake_by_year_df <- uptake_by_year |>
+  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
+
+plot_oa_all <- ggplot(uptake_by_year_df, aes(x = earliest_year)) +
+  geom_line(aes(y = n_oa_articles / n_articles, group = src, color = src), linetype = "solid") +
+  geom_point(aes(y = n_oa_articles / n_articles, fill = src), alpha = 0.5, shape = 21) +
+  scale_fill_manual(NULL,
+    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c"),
+    guide = "none"
+  ) +
+  scale_color_manual(NULL,
+    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c")
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  labs(y = "Hybrid OA Uptake", x = NULL)
+
+plot_ta_oa <- ggplot(filter(uptake_by_year_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c"),
+                    guide = "none") +
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = function(x) format(x, big.mark = ",", scientific = FALSE),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  labs(y = "Hybrid OA via Agreement", x = "Publication Year")
+
+plot_ta_perc <- ggplot(filter(uptake_by_year_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value / n_oa_articles,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value / n_oa_articles, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c"),
+                    guide = "none") +  
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  labs(y = "% of Hybrid OA via Agreement", x = NULL)
+
+layout <- "ABC
+EEE"
+
+combined_plot <- wrap_plots(
+  E = guide_area(),
+  A = plot_oa_all,
+  B = plot_ta_oa,
+  C = plot_ta_perc,
+  heights = c(1, 0),
+  design = layout
+) &
+  theme_minimal() +
+    theme(
+      legend.position = "bottom",
+      panel.border = element_rect(color = "grey50", fill = NA),
+      text = element_text(size = 10)
+    ) &
+  plot_annotation(tag_levels = "A")
+
+combined_plot + plot_layout(guides = "collect")
+```
+
+Following @Jahn_2025, Figure ...
+
+<!-- By identifying potential variations, this study evaluates the suitability of open data approaches, such as hoaddata, as alternatives to proprietary bibliometric databases like Scopus and Web of Science.-->
 

### 2025-03-10 analysis/manuscript.tex
Trend anaylsis replication

@@
 determine whether consistent results can be derived across these three
-data sources, despite their differing coverage and methodologies. By
-identifying potential variations, this study evaluates the suitability
-of open data approaches, such as hoaddata, as alternatives to
-proprietary bibliometric databases like Scopus and Web of Science.
+data sources, despite their differing coverage and methodologies.
+
+\begin{center}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/plot_uptake_overview-1.pdf}
+\end{center}
+
+Following Jahn (2025), Figure \ldots{}
 

### 2025-03-10 analysis/manuscript_preprocessing_data.R

@@ -278,6 +278,7 @@ aff_ind <- bind_rows(
               scp = scp_cor_au_with_affiliation), .id = "indicator"
 )
 
+
 # Save
 save(# Journals
   jns_ind,
@@ -369,3 +370,66 @@ cr_df |>
   distinct(doi) |>
   write_csv(here::here("data-raw", "missing_dois_in_active_jns.csv"))
 
+# TA OA Indicators
+# Web of Science
+wos_jn_by_year <- readr::read_csv(here::here("data", "wos_jn_by_year.csv")) 
+
+wos_jn_ta_by_year <- readr::read_csv(here::here("data", "wos_jn_ta_by_year.csv"))
+
+wos_ta_df <- wos_jn_by_year |>
+  # only hybrid journals with at least one oa between 2018 - 2023
+  filter(issn_l %in% wos_active_jns_with_oa$issn_l) |>
+  left_join(wos_jn_ta_by_year, by = c("issn_l", "earliest_year", "pub_type" = "pubtype")) |>
+  # Focus on core
+  filter(pub_type == "core") |>
+  mutate(src = "Web of Science") |>
+  select(-pub_type)
+
+# Scopus
+scp_jn_by_year <-  readr::read_csv(here::here("data", "scp_jn_by_year.csv")) 
+
+scp_jn_ta_by_year <- readr::read_csv(here::here("data", "scp_ta_jn_by_year.csv"))
+
+scp_ta_df <-  scp_jn_by_year |>
+  # only hybrid journals with at least one oa between 2018 - 2023
+  filter(issn_l %in% scp_active_jns_with_oa$issn_l) |>
+  left_join(scp_jn_ta_by_year, by = c("issn_l", "earliest_year", "pub_type" = "pubtype")) |>
+  # Focus on core
+  filter(pub_type == "core") |>
+  mutate(src = "Scopus") |>
+  select(-pub_type)
+
+# HOAD
+hoad_jn_by_year_ <- active_jns_with_oa |>
+  mutate(cr_year = as.numeric(as.character(cr_year))) |>
+  distinct(issn_l, cr_year, n = jn_all)
+hoad_oa_jn_by_year <- hoaddata::cc_articles |>
+  group_by(issn_l, cr_year) |>
+  summarise(oa_articles = n_distinct(doi))
+
[…]

### 2025-03-10 analysis/manuscript.qmd
Trend anaylsis replication: publisher mosaic

@@
 
-```{r, rep_data}
-uptake_df <- readr::read_csv(here::here("data", "uptake_df.csv"))
-```
-
 This section examines the open access uptake in hybrid journals, focusing on the influence of transformative agreements across hoaddata, Scopus and Web of Science.
@@
 
+```{r plot_publisher_uptake_overview, fig.asp=1.1}
+uptake_publisher <- uptake_df |>
+  inner_join(esac, by = "issn_l") |>
+  filter(earliest_year != "2018")
+
+uptake_publisher_by_year <- uptake_publisher |>
+  mutate(earliest_year = gsub("^20", "'", as.character(earliest_year))) |>
+  group_by(earliest_year, src, publisher) |>
+  summarize(
+    n_articles = sum(n),
+    n_oa_articles = sum(oa_articles),
+    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
+    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
+  )
+
+uptake_by_year_publisher_df <- uptake_publisher_by_year |>
+  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
+
+publisher_plot_oa_all <- ggplot(uptake_by_year_publisher_df, aes(x = earliest_year)) +
+  geom_line(aes(y = n_oa_articles / n_articles, group = src, color = src), linetype = "solid") +
+  geom_point(aes(y = n_oa_articles / n_articles, fill = src), alpha = 0.5, shape = 21) +
+  scale_fill_manual(NULL,
+    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c"),
+    guide = "none"
+  ) +
+  scale_color_manual(NULL,
+    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c")
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  facet_grid(~fct_rev(publisher)) +
+  labs(y = "Hybrid OA Uptake", x = NULL)
+
+publisher_plot_ta_oa <- ggplot(filter(uptake_by_year_publisher_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c"),
+                    guide = "none") +
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = function(x) format(x, big.mark = ",", scientific = FALSE),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  facet_grid(~fct_rev(publisher)) +
+  labs(y = "Hybrid OA via Agreement", x = NULL)
+
+publisher_plot_ta_perc <- ggplot(filter(uptake_by_year_publisher_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value / n_oa_articles,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value / n_oa_articles, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#595959", `Scopus` = "#2066a8", `Web of Science` = "#ae282c"),
+                    guide = "none") +  
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  facet_grid(~fct_rev(publisher)) +
+  labs(y = "% of Hybrid OA via Agreement",  x = "Publication Year")
+
+layout <- "A
+B
+C
+E"
+
+publisher_combined_plot <- wrap_plots(
+  E = guide_area(),
+  A = publisher_plot_oa_all,
+  B = publisher_plot_ta_oa,
+  C = publisher_plot_ta_perc,
+  heights = c(1, 1, 1, 0),
+  design = layout
+) &
+  theme_minimal() +
+    theme(
+      legend.position = "bottom",
+      panel.border = element_rect(color = "grey50", fill = NA),
+      text = element_text(size = 10)
+    ) &
+  plot_annotation(tag_levels = "A")
+
+publisher_combined_plot + plot_layout(guides = "collect")
+
+```
+
 # discussion

### 2025-03-11 analysis/manuscript.qmd
Trend anaylsis replication: writing

--- 2025-03-06 analysis/manuscript.qmd
@@
 
-## Replication
+## Open Access Indicator Comparision
 
-This section examines the open access uptake in hybrid journals, focusing on the influence of transformative agreements across hoaddata, Scopus and Web of Science.
-A key objective of this analysis is to determine whether consistent results can be derived across these three data sources, despite their differing coverage and methodologies. 
-By identifying potential variations, this study evaluates the suitability of open data approaches, such as hoaddata, as alternatives to proprietary bibliometric databases like Scopus and Web of Science.
+This section examines the uptake of open access in hybrid journals, focussing on the influence of transformative agreements across hoaddata, Scopus and Web of Science.
+The analysis aims to determine whether consistent results can be derived from these three data sources despite their differing coverage and methodologies. 
+Following @Jahn_2025, indicators include the number and proportion of open access articles, both overall and those specifically enabled by transformative agreements. 
+The analysis compares the development of these indicators over time between 2019 and 2023 across all three data sources.
 
-Following @Jahn_2025, the impact of transformative agreements on hybrid open access is assessed by analysing the proportion of open access articles, both overall and those specifically enabled by transformative agreements.
-The analysis covers the publication years 2019 to 2023, providing insights into temporal trends in the adoption of transformative agreements and their influence on open access in hybrid journals.
-The analysis will present breakdowns by publishers and countries.
+```{r, plot_uptake_overview, fig.asp=.5}
+#| fig-cap: "Open Access indicato"
+#| label: fig-uptake_overview
+#| 
+uptake_df <- readr::read_csv(here::here("data", "uptake_df.csv"))
 
-A complicating factor in the comparision is the differing coverage of author affiliation data across the datasets.
-In hoaddata, which relies on OpenAlex for affiliation information, the lack of comprehensive corresponding author data restricts the analysis to first authors when estimating the influence of transformative agreements.
-In contrast, Scopus and Web of Science provide more robust coverage of corresponding author affiliations, enabling a comparative assessment of how authorship roles--first versus corresponding authors--influence estimates of open access funding through transformative agreements.
+# Data prep
+uptake_by_year <- uptake_df |>
+  group_by(earliest_year, src) |>
+  summarize(
+    n_articles = sum(n),
+    n_oa_articles = sum(oa_articles),
+    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
+    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
+  )
+
+uptake_by_year_short <- uptake_by_year |>
+  mutate(oa_prop = n_oa_articles / n_articles,
+         ta_oa_prop = ta_oa_first_author_articles / n_oa_articles,
+         ta_oa_cor_prop = ta_oa_corresponding_author_articles / n_oa_articles)
+
+uptake_by_year_df <- uptake_by_year |>
+  mutate(earliest_year = gsub("^20", "'", as.character(earliest_year))) |>
+  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
+
+plot_oa_all <- ggplot(uptake_by_year_df, aes(x = earliest_year)) +
+  geom_line(aes(y = n_oa_articles / n_articles, group = src, color = src), linetype = "solid") +
+  geom_point(aes(y = n_oa_articles / n_articles, fill = src), alpha = 0.5, shape = 21) +
+  scale_fill_manual(NULL,
+    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+    guide = "none"
+  ) +
+  scale_color_manual(NULL,
+   values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  labs(y = "Hybrid OA Uptake", x = NULL)
+
+plot_ta_oa <- ggplot(filter(uptake_by_year_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+                    guide = "none") +
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = function(x) format(x, big.mark = ",", scientific = FALSE),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  labs(y = "Hybrid OA via Agreement", x = "Publication Year")
+
+plot_ta_perc <- ggplot(filter(uptake_by_year_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value / n_oa_articles,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value / n_oa_articles, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+                    guide = "none") +  
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  labs(y = "% of Hybrid OA via Agreement", x = NULL)
+
+layout <- "ABC
+EEE"
+
+combined_plot <- wrap_plots(
+  E = guide_area(),
+  A = plot_oa_all,
+  B = plot_ta_oa,
+  C = plot_ta_perc,
+  heights = c(1, 0),
+  design = layout
+) &
+  theme_minimal() +
+    theme(
+      legend.position = "bottom",
+      panel.border = element_rect(color = "grey50", fill = NA),
+      text = element_text(size = 10)
+    ) &
+  plot_annotation(tag_levels = "A")
+
+combined_plot + plot_layout(guides = "collect")
+```
+
+### Overview
+
+Figure \ref{fig-uptake_overview}A shows a moderate growth of open access in hybrid journals, which is consistent across hoaddata (black line), Scopus (orange line) and Web of Science (blue line).
+In 2023, hybrid open access uptake reached  `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(oa_prop)*100`%  in hoaddata (n = `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(n_oa_articles)`) and Scopus (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`), and `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop) * 100`%  in Web of Science (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`.
+
+The number and proportion of open access articles enabled through transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
+These trends remained consistent when estimating open access via first author (dashed line) and corresponding author (dotted line) affiliation metadata. 
+Using Scopus, it was possible to attribute `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_first_author_articles) |> sum()` open access articles to transformative agreements based on first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`) and `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
+Web of Science recorded `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_first_author_articles) |> sum()` open access articles via transformative agreements using first author metadata ( growing from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`)  and  `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
+The development over the years for hoadata is shown only for first authors due to unavailable corresponding author data. 
+In total, hoaddata linked `r uptake_by_year_short |> filter(src == "HOAD") |> pull(ta_oa_first_author_articles) |> sum()` open access articles to transformative agreements, making it the most comprehensive data source for first-author analysis.
+However, hoaddata showed slower growth compared to the other two data sources, as illustrated by its fewer articles than Scopus in 2023. 
+Furthermore, Scopus and Web of Science recorded slighlty fewer articles for first author articles than those reported for corresponding authors.
+
+When considering the percentage of hybrid open access through transformative agreements relative to total open access, all data sources consistently indicate that between 2021 and 2022, the majority of open access was enabled by transformative agreements.
+This trend can be obsverved for both first and corresponding author affiliations.
+For first authors, the share of open access ranged between `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus) and  `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata).
+For corresponding authors, the share was slighly larger, with Web of Science recording `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100` % and Scopus  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`%)
+Nevertheless, substantial amounts of hybrid open access were still facilitated by means other than transformative agreements, likely through APCs invoiced to researchers and paid from research funds.
+
+Despite these similarities, the results also reveal notable differences. 
+Specifically, the Web of Science recorded substantially higher open access uptake in the early years compared to Scopus and hoaddata, before converging in 2023.
+This suggests that delayed open access, i.e., open access after an embargo period, was labeled as hybrid open access in the Web of Science.
+
+### Open access by publishers
+
+When considering open access trends by publisher, it becomes evident that in particular due to Elsevier, the largest publsiher in our sample. 
+Following @Jahn_2025, Figure ...
+
+```{r plot_publisher_uptake_overview, fig.asp=1.1}
+uptake_publisher <- uptake_df |>
+  inner_join(esac, by = "issn_l") |>
+  filter(earliest_year != "2018")
+
+uptake_publisher_by_year <- uptake_publisher |>
+  mutate(earliest_year = gsub("^20", "'", as.character(earliest_year))) |>
+  group_by(earliest_year, src, publisher) |>
+  summarize(
+    n_articles = sum(n),
+    n_oa_articles = sum(oa_articles),
+    ta_oa_first_author_articles = sum(ta_oa_first_author_articles),
+    ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
+  )
+
+uptake_by_year_publisher_df <- uptake_publisher_by_year |>
+  pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
+
+publisher_plot_oa_all <- ggplot(uptake_by_year_publisher_df, aes(x = earliest_year)) +
+  geom_line(aes(y = n_oa_articles / n_articles, group = src, color = src), linetype = "solid") +
+  geom_point(aes(y = n_oa_articles / n_articles, fill = src), alpha = 0.5, shape = 21) +
+  scale_fill_manual(NULL,
+    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+    guide = "none"
+  ) +
+  scale_color_manual(NULL,
+    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  facet_grid(~fct_rev(publisher)) +
+  labs(y = "Hybrid OA Uptake", x = NULL)
+
+publisher_plot_ta_oa <- ggplot(filter(uptake_by_year_publisher_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+                    guide = "none") +
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = function(x) format(x, big.mark = ",", scientific = FALSE),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  facet_grid(~fct_rev(publisher)) +
+  labs(y = "Hybrid OA via Agreement", x = NULL)
+
+publisher_plot_ta_perc <- ggplot(filter(uptake_by_year_publisher_df, value > 0), aes(x = earliest_year)) +
+  geom_line(aes(
+    y = value / n_oa_articles,
+    group = interaction(src, name),
+    color = src, # Color by src only
+    linetype = name
+  )) + # Use linetype for name
+  geom_point(aes(y = value / n_oa_articles, fill = src), alpha = 0.8, shape = 21) +
+  scale_color_manual(NULL, values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
+  scale_fill_manual(NULL,
+                    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+                    guide = "none") +  
+  scale_linetype_manual(NULL,
+    values = c(
+      `ta_oa_first_author_articles` = "longdash",
+      `ta_oa_corresponding_author_articles` = "dotted"
+    ),
+    labels = c(
+      `ta_oa_first_author_articles` = "First",
+      `ta_oa_corresponding_author_articles` = "Corresponding"
+    )
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  facet_grid(~fct_rev(publisher)) +
+  labs(y = "% of Hybrid OA via Agreement",  x = "Publication Year")
+
+layout <- "A
+B
+C
+E"
+
+publisher_combined_plot <- wrap_plots(
+  E = guide_area(),
+  A = publisher_plot_oa_all,
+  B = publisher_plot_ta_oa,
+  C = publisher_plot_ta_perc,
+  heights = c(1, 1, 1, 0),
+  design = layout
+) &
+  theme_minimal() +
+    theme(
+      legend.position = "bottom",
+      panel.border = element_rect(color = "grey50", fill = NA),
+      text = element_text(size = 10)
+    ) &
+  plot_annotation(tag_levels = "A")
+
+publisher_combined_plot + plot_layout(guides = "collect")
+
+```
 

### 2025-03-11 analysis/manuscript.tex
Trend anaylsis replication: publisher mosaic
Trend anaylsis replication: writing

--- 2025-03-05 analysis/manuscript.tex
@@
 
+\subsection{Open Access Indicator
+Comparision}\label{open-access-indicator-comparision}
+
+This section examines the uptake of open access in hybrid journals,
+focussing on the influence of transformative agreements across hoaddata,
+Scopus and Web of Science. The analysis aims to determine whether
+consistent results can be derived from these three data sources despite
+their differing coverage and methodologies. Following Jahn (2025),
+indicators include the number and proportion of open access articles,
+both overall and those specifically enabled by transformative
+agreements. The analysis compares the development of these indicators
+over time between 2019 and 2023 across all three data sources.
+
+\begin{figure}[ht!]
+
+\centering{
+
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_overview-1.pdf}
+
+}
+
+\caption{\label{fig-uptake_overview}Open Access indicato}
+
+\end{figure}%
+
+\subsubsection{Overview}\label{overview-1}
+
+Figure \ref{fig-uptake_overview}A shows a moderate growth of open access
+in hybrid journals, which is consistent across hoaddata (black line),
+Scopus (orange line) and Web of Science (blue line). In 2023, hybrid
+open access uptake reached 19\% in hoaddata (n = 302,358) and Scopus (n
+= 322,850), and 20\% in Web of Science (n = 255,481.
+
+The number and proportion of open access articles enabled through
+transformative agreements increased between 2019 and 2023 (Figures
+\ref{fig-uptake_overview}B and C). These trends remained consistent when
+estimating open access via first author (dashed line) and corresponding
+author (dotted line) affiliation metadata. Using Scopus, it was possible
+to attribute 479,297 open access articles to transformative agreements
+based on first author metadata (increasing from 15,341 to 206,084) and
+489,262 using corresponding author metadata (from 15,444 to 210,816).
+Web of Science recorded 493,028 open access articles via transformative
+agreements using first author metadata ( growing from 21,871 to 174,126)
+and 500,076 using corresponding author metadata (from 22,092 to
+177,030). The development over the years for hoadata is shown only for
+first authors due to unavailable corresponding author data. In total,
+hoaddata linked 501,649 open access articles to transformative
+agreements, making it the most comprehensive data source for
+first-author analysis. However, hoaddata showed slower growth compared
+to the other two data sources, as illustrated by its fewer articles than
+Scopus in 2023. Furthermore, Scopus and Web of Science recorded slighlty
+fewer articles for first author articles than those reported for
+corresponding authors.
+
+When considering the percentage of hybrid open access through
+transformative agreements relative to total open access, all data
+sources consistently indicate that between 2021 and 2022, the majority
+of open access was enabled by transformative agreements. This trend can
+be obsverved for both first and corresponding author affiliations. For
+first authors, the share of open access ranged between 68 \% (Web of
+Science), 64\% (Scopus) and 61\% (hoaddata). For corresponding authors,
+the share was slighly larger, with Web of Science recording 69 \% and
+Scopus 65\%) Nevertheless, substantial amounts of hybrid open access
+were still facilitated by means other than transformative agreements,
+likely through APCs invoiced to researchers and paid from research
+funds.
+
+Despite these similarities, the results also reveal notable differences.
+Specifically, the Web of Science recorded substantially higher open
+access uptake in the early years compared to Scopus and hoaddata, before
+converging in 2023. This suggests that delayed open access, i.e., open
+access after an embargo period, was labeled as hybrid open access in the
+Web of Science.
+
+\subsubsection{Open access by
+publishers}\label{open-access-by-publishers}
+
+When considering open access trends by publisher, it becomes evident
+that in particular due to Elsevier, the largest publsiher in our sample.
+Following Jahn (2025), Figure \ldots{}
+
+\begin{center}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/plot_publisher_uptake_overview-1.pdf}
+\end{center}
+
 \section*{discussion}\label{discussion}

### 2025-03-12 analysis/manuscript.qmd
Finish first draft oa indicators by publiher

@@
 
-This section examines the uptake of open access in hybrid journals, focussing on the influence of transformative agreements across hoaddata, Scopus and Web of Science.
-The analysis aims to determine whether consistent results can be derived from these three data sources despite their differing coverage and methodologies. 
-Following @Jahn_2025, indicators include the number and proportion of open access articles, both overall and those specifically enabled by transformative agreements. 
-The analysis compares the development of these indicators over time between 2019 and 2023 across all three data sources.
+This section examines the uptake of open access in hybrid journals, focusing on the influence of transformative agreements across hoaddata, Scopus, and Web of Science.
+The analysis assesses whether consistent results can be derived from these data sources despite differences in coverage and methodologies. 
+Following @Jahn_2025, indicators were calculated for each data source and comprise the number and proportion of open access articles, including those enabled by transformative agreements, from 2019 to 2023.
+For Web of Science and Scopus, the impact of transformative agreements was estimated using both first and corresponding authorships, while hoaddata indicator calculation was limited to first-author affiliations.
 
@@
 
-Figure \ref{fig-uptake_overview}A shows a moderate growth of open access in hybrid journals, which is consistent across hoaddata (black line), Scopus (orange line) and Web of Science (blue line).
-In 2023, hybrid open access uptake reached  `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(oa_prop)*100`%  in hoaddata (n = `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(n_oa_articles)`) and Scopus (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`), and `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop) * 100`%  in Web of Science (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`.
+Figure \ref{fig-uptake_overview}A shows a moderate growth of open access in hybrid journals, which is consistent across hoaddata (black line), Scopus (orange line), and Web of Science (blue line).
+According to hoaddata, hybrid open access uptake increased from `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
+Similarly, Scopus recorded an growth from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
+However, Web of Science recorded higher open access uptake in early years, before converging in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
+This may reflect delayed open access being labeled as hybrid in Web of Science.
 
-The number and proportion of open access articles enabled through transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
-These trends remained consistent when estimating open access via first author (dashed line) and corresponding author (dotted line) affiliation metadata. 
-Using Scopus, it was possible to attribute `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_first_author_articles) |> sum()` open access articles to transformative agreements based on first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`) and `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
-Web of Science recorded `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_first_author_articles) |> sum()` open access articles via transformative agreements using first author metadata ( growing from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`)  and  `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
-The development over the years for hoadata is shown only for first authors due to unavailable corresponding author data. 
-In total, hoaddata linked `r uptake_by_year_short |> filter(src == "HOAD") |> pull(ta_oa_first_author_articles) |> sum()` open access articles to transformative agreements, making it the most comprehensive data source for first-author analysis.
-However, hoaddata showed slower growth compared to the other two data sources, as illustrated by its fewer articles than Scopus in 2023. 
-Furthermore, Scopus and Web of Science recorded slighlty fewer articles for first author articles than those reported for corresponding authors.
+Open access by transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
+Trends were consistent for first (dashed line) and corresponding author (dotted line) affiliations. 
+According to Scopus, `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_first_author_articles) |> sum()` open access articles could be attributed to transformative agreements based on first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`) and `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
+Web of Science recorded `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_first_author_articles) |> sum()` open access articles via transformative agreements using first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`)  and  `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
+hoaddata, lacking corresponding author data, linked `r uptake_by_year_short |> filter(src == "HOAD") |> pull(ta_oa_first_author_articles) |> sum()` articles to transformative agreements but showed slower growth than Scopus and Web of Science, from `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`.
 
-When considering the percentage of hybrid open access through transformative agreements relative to total open access, all data sources consistently indicate that between 2021 and 2022, the majority of open access was enabled by transformative agreements.
-This trend can be obsverved for both first and corresponding author affiliations.
-For first authors, the share of open access ranged between `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus) and  `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata).
-For corresponding authors, the share was slighly larger, with Web of Science recording `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100` % and Scopus  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`%)
-Nevertheless, substantial amounts of hybrid open access were still facilitated by means other than transformative agreements, likely through APCs invoiced to researchers and paid from research funds.
-
-Despite these similarities, the results also reveal notable differences. 
-Specifically, the Web of Science recorded substantially higher open access uptake in the early years compared to Scopus and hoaddata, before converging in 2023.
-This suggests that delayed open access, i.e., open access after an embargo period, was labeled as hybrid open access in the Web of Science.
+From 2021 (hoadata, Scopus) resp. 2022 (Web of Science),  transformative agreements enabled the majority of hybrid open access.
+For first authors, the share ranged between `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and  `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata) in 2023.
+For corresponding authors, the shares were slightly larger, with Web of Science recording `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100` % and Scopus  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% in 2023.
+However, substantial hybrid open access was still facilitated outside transformative agreements, likely through APCs payed from discretionary research funds (ref).
 
@@
 
-When considering open access trends by publisher, it becomes evident that in particular due to Elsevier, the largest publsiher in our sample. 
-Following @Jahn_2025, Figure ...
-
-```{r plot_publisher_uptake_overview, fig.asp=1.1}
+```{r uptake_publisher} 
 uptake_publisher <- uptake_df |>
@@
 uptake_publisher_by_year <- uptake_publisher |>
-  mutate(earliest_year = gsub("^20", "'", as.character(earliest_year))) |>
   group_by(earliest_year, src, publisher) |>
@@
 
+uptake_by_publisher_short <- uptake_publisher_by_year |>
+  mutate(oa_prop = n_oa_articles / n_articles,
+         ta_oa_prop = ta_oa_first_author_articles / n_oa_articles,
+         ta_oa_cor_prop = ta_oa_corresponding_author_articles / n_oa_articles)
+
 uptake_by_year_publisher_df <- uptake_publisher_by_year |>
+  mutate(earliest_year = gsub("^20", "'", as.character(earliest_year))) |>
   pivot_longer(cols = c(ta_oa_first_author_articles, ta_oa_corresponding_author_articles))
+```
 
+```{r plot_publisher_uptake_overview, fig.asp=1.1}
+#| fig-cap: "Open Access indicato"
+#| label: fig-uptake_publisher
+#| 
 publisher_plot_oa_all <- ggplot(uptake_by_year_publisher_df, aes(x = earliest_year)) +
@@
 
+When considering open access trends by publisher (see \ref{fig-uptake_publisher}), the observed differences in early uptake rates between hoaddata and Scopus compared to Web of Science can be largely attributed to articles published in Elsevier hybrid journals, the largest publisher in our sample.
+Both hoadata and Scopus reported an steady increase in open access uptake between 2019 and 2023 (hoaddata from `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%; Scopus from `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%).
+In contrast, Elsevier's share remained relatively constant, increasing only slightly from `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`% according to the Web of Science.
+Upon inspection, this discrepancy primarily stemmed from articles provided under Elsevier's user license, used for the publisher's open archive. 
+These articles, made freely available after an embargo period, were tagged as hybrid open access in Web of Science, even though its documentation specified that only articles under a CC license variant were considered.
+Previous research [@Jahn_2021; @Haustein_APC_2024] has shown that Elsevier provided a substantial portion of its articles under this license, explaining the relatively large and stable share of open access over the years.
+
+Differences in open access evidence are also evident for Wiley. 
+Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022 compared to hoaddata. 
+For these two years, hoaddata reported `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Scopus", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more open access articles than Scopus and `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Web of Science", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more than Web of Science.
+This discrepancy is presumably because of challenges in fetching full-texts by Unpaywall, the open access evidence source for Scopus and Web of Science.
+According to Unpaywall's software version history, Wiley's publisher platform redirects prevented Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
+hoaddata, which relies solely on Crossref license information, was unaffected by these issues.
+
+Despite these differences in open access evidence, the three data sources show consistent temporal trends in hybrid open access enabled by transformative agreements (see Figure \ref{fig-uptake_publisher}B and C).
+Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across all data sources, followed by Springer Nature and Elsevier.
+By 2023, the share of open access enabled by transformative agreements appeared to stabilise for all three publishers, with Wiley recording the largest share (see Figure \ref{fig-uptake_publisher}B). 
+Elsevier, however, showed a comparable late uptake, consistent with the publisher's historical reluctance to engage in negotations with library consortia. 
+Interestingly, the differences between first and corresponding author affiliations were more pronounced at the publisher level. 
+In Scopus, for example, the share of open access via transformative agreements measured by corresponding authorship was greater for Springer Nature in 2023 than when using first authorship.
+
+### Open access by country
+
 # discussion
+
+[^unpaywallwiley]: See Unpaywall version history related to Wiley fixes: <https://github.com/search?q=repo%3Aourresearch%2Foadoi+wiley&type=commits>

### 2025-03-12 analysis/manuscript.tex
Finish first draft oa indicators by publiher

--- 2025-03-10 analysis/manuscript.tex
@@
 
-\subsection{Replication}\label{replication}
+\subsection{Open Access Indicator
+Comparision}\label{open-access-indicator-comparision}
 
-This section examines the open access uptake in hybrid journals,
+This section examines the uptake of open access in hybrid journals,
 focusing on the influence of transformative agreements across hoaddata,
-Scopus and Web of Science. A key objective of this analysis is to
-determine whether consistent results can be derived across these three
-data sources, despite their differing coverage and methodologies.
+Scopus, and Web of Science. The analysis assesses whether consistent
+results can be derived from these data sources despite differences in
+coverage and methodologies. Following Jahn (2025), indicators were
+calculated for each data source and comprise the number and proportion
+of open access articles, including those enabled by transformative
+agreements, from 2019 to 2023. For Web of Science and Scopus, the impact
+of transformative agreements was estimated using both first and
+corresponding authorships, while hoaddata indicator calculation was
+limited to first-author affiliations.
 
-\begin{center}
-\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/plot_uptake_overview-1.pdf}
-\end{center}
+\begin{figure}[ht!]
 
-Following Jahn (2025), Figure \ldots{}
+\centering{
 
-Following Jahn (2025), the impact of transformative agreements on hybrid
-open access is assessed by analysing the proportion of open access
-articles, both overall and those specifically enabled by transformative
-agreements. The analysis covers the publication years 2019 to 2023,
-providing insights into temporal trends in the adoption of
-transformative agreements and their influence on open access in hybrid
-journals. The analysis will present breakdowns by publishers and
-countries.
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_overview-1.pdf}
 
-A complicating factor in the comparision is the differing coverage of
-author affiliation data across the datasets. In hoaddata, which relies
-on OpenAlex for affiliation information, the lack of comprehensive
-corresponding author data restricts the analysis to first authors when
-estimating the influence of transformative agreements. In contrast,
-Scopus and Web of Science provide more robust coverage of corresponding
-author affiliations, enabling a comparative assessment of how authorship
-roles--first versus corresponding authors--influence estimates of open
-access funding through transformative agreements.
+}
+
+\caption{\label{fig-uptake_overview}Open Access indicato}
+
+\end{figure}%
+
+\subsubsection{Overview}\label{overview-1}
+
+Figure \ref{fig-uptake_overview}A shows a moderate growth of open access
+in hybrid journals, which is consistent across hoaddata (black line),
+Scopus (orange line), and Web of Science (blue line). According to
+hoaddata, hybrid open access uptake increased from 6.4\% (n = 85,071) in
+2019 to 19\% (n = 302,358) in 2023. Similarly, Scopus recorded an growth
+from 6.5\% (n = 84,648) in 2019 to 19\% (n = 322,850) in 2023. However,
+Web of Science recorded higher open access uptake in early years, before
+converging in 2023, from 11\% (n = 137,202) in 2019 to 20\% (n =
+255,481) in 2023. This may reflect delayed open access being labeled as
+hybrid in Web of Science.
+
+Open access by transformative agreements increased between 2019 and 2023
+(Figures \ref{fig-uptake_overview}B and C). Trends were consistent for
+first (dashed line) and corresponding author (dotted line) affiliations.
+According to Scopus, 479,297 open access articles could be attributed to
+transformative agreements based on first author metadata (increasing
+from 15,341 to 206,084) and 489,262 using corresponding author metadata
+(from 15,444 to 210,816). Web of Science recorded 493,028 open access
+articles via transformative agreements using first author metadata
+(increasing from 21,871 to 174,126) and 500,076 using corresponding
+author metadata (from 22,092 to 177,030). hoaddata, lacking
+corresponding author data, linked 501,649 articles to transformative
+agreements but showed slower growth than Scopus and Web of Science, from
+16,010 to 183,757.
+
+From 2021 (hoadata, Scopus) resp. 2022 (Web of Science), transformative
+agreements enabled the majority of hybrid open access. For first
+authors, the share ranged between 68 \% (Web of Science), 64\% (Scopus),
+and 61\% (hoaddata) in 2023. For corresponding authors, the shares were
+slightly larger, with Web of Science recording 69 \% and Scopus 65\% in
+2023. However, substantial hybrid open access was still facilitated
+outside transformative agreements, likely through APCs payed from
+discretionary research funds (ref).
+
+\subsubsection{Open access by
+publishers}\label{open-access-by-publishers}
+
+\begin{figure}[ht!]
+
+\centering{
+
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_publisher-1.pdf}
+
+}
+
+\caption{\label{fig-uptake_publisher}Open Access indicato}
+
+\end{figure}%
+
+When considering open access trends by publisher (see
+\ref{fig-uptake_publisher}), the observed differences in early uptake
+rates between hoaddata and Scopus compared to Web of Science can be
+largely attributed to articles published in Elsevier hybrid journals,
+the largest publisher in our sample. Both hoadata and Scopus reported an
+steady increase in open access uptake between 2019 and 2023 (hoaddata
+from 4\% to 13\%; Scopus from 5\% to 15\%). In contrast, Elsevier's
+share remained relatively constant, increasing only slightly from 14\%
+to 15\% according to the Web of Science. Upon inspection, this
+discrepancy primarily stemmed from articles provided under Elsevier's
+user license, used for the publisher's open archive. These articles,
+made freely available after an embargo period, were tagged as hybrid
+open access in Web of Science, even though its documentation specified
+that only articles under a CC license variant were considered. Previous
+research (Haustein et al., 2024; Jahn et al., 2022) has shown that
+Elsevier provided a substantial portion of its articles under this
+license, explaining the relatively large and stable share of open access
+over the years.
+
+Differences in open access evidence are also evident for Wiley.
+Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022
+compared to hoaddata. For these two years, hoaddata reported 35,308 more
+open access articles than Scopus and 32,491 more than Web of Science.
+This discrepancy is presumably because of challenges in fetching
+full-texts by Unpaywall, the open access evidence source for Scopus and
+Web of Science. According to Unpaywall's software version history,
+Wiley's publisher platform redirects prevented Unpaywall from parsing
+license information from full-texts.\footnote{See Unpaywall version
+  history related to Wiley fixes:
+  \url{https://github.com/search?q=repo\%3Aourresearch\%2Foadoi+wiley&type=commits}}
+hoaddata, which relies solely on Crossref license information, was
+unaffected by these issues.
+
+Despite these differences in open access evidence, the three data
+sources show consistent temporal trends in hybrid open access enabled by
+transformative agreements (see Figure \ref{fig-uptake_publisher}B and
+C). Wiley emerged as the fastest-growing publisher in terms of open
+access uptake, with more than 30\% of articles in hybrid journals
+reported as open access in 2023 across all data sources, followed by
+Springer Nature and Elsevier. By 2023, the share of open access enabled
+by transformative agreements appeared to stabilise for all three
+publishers, with Wiley recording the largest share (see Figure
+\ref{fig-uptake_publisher}B). Elsevier, however, showed a comparable
+late uptake, consistent with the publisher's historical reluctance to
+engage in negotations with library consortia. Interestingly, the
+differences between first and corresponding author affiliations were
+more pronounced at the publisher level. In Scopus, for example, the
+share of open access via transformative agreements measured by
+corresponding authorship was greater for Springer Nature in 2023 than
+when using first authorship.
+
+\subsubsection{Open access by country}\label{open-access-by-country}
 
@@
 
+\bibitem[\citeproctext]{ref-Haustein_APC_2024}
+Haustein, S., Schares, E., Alperin, J. P., Hare, M., Butler, L.-A., \&
+Schönfelder, N. (2024). \emph{Estimating global article processing
+charges paid to six publishers for open access between 2019 and 2023}.
+\url{https://arxiv.org/abs/2407.16551}
+
 \bibitem[\citeproctext]{ref-Jahn_2025}
@@
 \url{https://doi.org/10.1162/qss_a_00348}
+
+\bibitem[\citeproctext]{ref-Jahn_2021}
+Jahn, N., Matthias, L., \& Laakso, M. (2022). Toward transparency of
+hybrid open access through publisher-provided metadata: An article-level
+study of {Elsevier}. \emph{Journal of the Association for Information
+Science and Technology}, \emph{73}(1), 104--118.
+\url{https://doi.org/10.1002/asi.24549}
 

### 2025-03-14 analysis/manuscript.qmd
tmp save

@@
 
+```{r country_data_prep}
+jn_df <- readr::read_csv(here::here("data", "jn_country_stats_all.csv")) |>
+  filter(issn_l != "0027-8424")
+oecd <- readr::read_csv("https://raw.githubusercontent.com/njahn82/hoa_ta_effects/refs/heads/main/data/oecd.csv")
+
+country_df <- jn_df |>
+  select(-earliest_year) |>
+  group_by(country_code) |>
+  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) |>
+  mutate(country_group = case_when(
+    country_code %in% oecd$iso3 ~ "OECD",
+    country_code %in% c("BRA", "RUS", "IND", "CHN", "ZAF") ~ "BRICS",
+    .default = "Other"
+  ))
+
+country_article_plot_df <- country_df |>
+  select(country_code, country_group, hoad_articles, scp_first_author_articles, scp_corresponding_author_articles, wos_first_author_articles, wos_corresponding_author_articles) |>
+  pivot_longer(c(scp_first_author_articles, scp_corresponding_author_articles, wos_first_author_articles, wos_corresponding_author_articles)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "first_author_articles" ~ "First",
+    indicator == "corresponding_author_articles" ~ "Corresponding"
+  ))
+
+articles_plot <- country_article_plot_df |>
+  filter(!is.na(country_code), hoad_articles > 10000) |>
+  ggplot(aes(x = hoad_articles, y = value, fill = country_group)) +
+  geom_point(pch = 21, size = 2.5) +
+  ggrepel::geom_text_repel(aes(label = country_code), size = 2) +
+  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf")) +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  coord_fixed(ratio = 1) +
+  scale_x_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  scale_y_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  theme_minimal(base_size = 10) +
+  labs(x = "Article volume (log-scale)\nhoaddata", y = "Article volume (log-scale)") +
+  theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+    legend.position = "top",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+
+## OA Uptake
+oa_prop_plot_df <- country_df |>
+  filter(!is.na(country_code), hoad_articles > 10000) |>
+  mutate(
+    hoad_oa_prop = hoad_oa_articles / hoad_articles,
+    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
+    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
+    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
+    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
+  ) |>
+  select(country_code, country_group, contains("prop")) |>
+  pivot_longer(c(scp_first_oa_prop, scp_corresponding_oa_prop, wos_first_oa_prop, wos_corresponding_oa_prop)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "first_oa_prop" ~ "First",
+    indicator == "corresponding_oa_prop" ~ "Corresponding"
+  ))
+
+oa_prop_plot <- ggplot(oa_prop_plot_df, aes(x = hoad_oa_prop, y = value, fill = country_group)) +
+  geom_point(alpha = 1, pch = 21, size = 2.5) +
+  ggrepel::geom_text_repel(aes(label = country_code), size = 2) +
+  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf"), guide = "none") +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  coord_fixed(ratio = 1) +
+  scale_x_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  theme_minimal(base_size = 10) +
+  labs(x = "Hybrid OA Uptake\nhoadata", y = "Hybrid OA Uptake") +
+    theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+ #   strip.text.x = element_blank(),
+    legend.position = "top",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+
+## Hybrid OA via agreement
+ta_oa_country_df <- country_df |>
+  filter(!is.na(country_code), hoad_articles > 10000) |>
+  select(country_code, country_group, hoad_ta_oa_articles, scp_ta_oa_first_author_articles, scp_ta_oa_corresponding_author_articles, wos_ta_oa_first_author_articles, wos_ta_oa_corresponding_author_articles) |>
+  pivot_longer(c(scp_ta_oa_first_author_articles, scp_ta_oa_corresponding_author_articles, wos_ta_oa_first_author_articles, wos_ta_oa_corresponding_author_articles)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "ta_oa_first_author_articles" ~ "First",
+    indicator == "ta_oa_corresponding_author_articles" ~ "Corresponding"
+  ))
+
+## Hybrid OA vioa agreement share
+ta_oa_country_plot <- ta_oa_country_df |>
+  filter(hoad_ta_oa_articles > 1000) |>
+  ggplot(aes(x = hoad_ta_oa_articles, y = value, fill = country_group)) +
+  geom_point(pch = 21, size = 2.5) +
+  ggrepel::geom_text_repel(aes(label = country_code), size = 2) +
+  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf"), guide = "none") +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  coord_fixed(ratio = 1) +
+  scale_x_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  scale_y_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  theme_minimal(base_size = 10) +
+  labs(x = "Hybrid OA via Agreement (log-scale)\nhoadata", y = "Hybrid OA via Agreement (log-scale)") +
+  theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+    legend.position = "top",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+
+## Hybrid OA vio agreement share
+ta_oa_prop_plot_df <- country_df |>
+  filter(hoad_ta_oa_articles > 1000) |>
+  mutate(
+    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
+    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
+    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
+    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
+    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
+  ) |>
+  select(country_code, country_group, contains("prop")) |>
+  pivot_longer(c(scp_first_oa_prop, scp_corresponding_oa_prop, wos_first_oa_prop, wos_corresponding_oa_prop)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "first_oa_prop" ~ "First",
+    indicator == "corresponding_oa_prop" ~ "Corresponding"
+  ))
+
+ta_oa_prop_plot <- ggplot(ta_oa_prop_plot_df, aes(x = hoad_oa_prop, y = value, fill = country_group)) +
+  geom_point(alpha = 1, pch = 21, size = 2.5) +
+  ggrepel::geom_text_repel(aes(label = country_code), size = 2) +
+  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf"), guide = "none") +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  coord_fixed(ratio = 1) +
+  scale_x_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  theme_minimal(base_size = 10) +
+  labs(y = "% of Hybrid OA via Agreement", x = "% of Hybrid OA via Agreement\nhoadata") +
+  theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+#    strip.text.x = element_blank(),
+    legend.position = "bottom",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+```
+
+```{r oa_country_comparision_plot_, fig.asp=1, fig.width = 8.5}
+layout <- "EE
+AC
+BD"
+
+combined_oa_country_plot_ <- wrap_plots(
+  E = guide_area(),
+  A = articles_plot,
+  B = oa_prop_plot,
+  C = ta_oa_country_plot,
+  D = ta_oa_prop_plot,
+  design = layout,
+  heights = c(0,1,1)
+) &
+    theme(
+      legend.position = "top",
+      panel.border = element_rect(color = "grey50", fill = NA),
+      text = element_text(size = 10))
+
+combined_oa_country_plot_ + plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = 'A')
+```
+
+results are consistent.
+most represented are chinese, usa and india in temrs of articles.
+however, in terms of oa, other country lead,
+these are in particular those with transformative agremments
+
+overall, briscs and oecd, however, notable expection is south africa
+
+en though correlation exisits, a closer lokk at some temproarl trend revela (dahsboard), which countries were affected by inconsitstent oa evidnecie.
+heres is wiley, and there is france.
+
 # discussion

### 2025-03-14 analysis/manuscript.tex
tmp save

@@
 
+\begin{center}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/oa_country_comparision_plot_-1.pdf}
+\end{center}
+
+results are consistent. most represented are chinese, usa and india in
+temrs of articles. however, in terms of oa, other country lead, these
+are in particular those with transformative agremments
+
+overall, briscs and oecd, however, notable expection is south africa
+
+en though correlation exisits, a closer lokk at some temproarl trend
+revela (dahsboard), which countries were affected by inconsitstent oa
+evidnecie. heres is wiley, and there is france.
+
 \section*{discussion}\label{discussion}

### 2025-03-17 analysis/manuscript.qmd
Improve figure

--- 2025-03-12 analysis/manuscript.qmd
@@
 library(ComplexUpset)
+library(ggrepel)
+library(ggpp)
 
@@
 
+```{r country_data_prep}
+jn_df <- readr::read_csv(here::here("data", "jn_country_stats_all.csv")) |>
+  filter(issn_l != "0027-8424")
+oecd <- readr::read_csv("https://raw.githubusercontent.com/njahn82/hoa_ta_effects/refs/heads/main/data/oecd.csv")
+
+country_df <- jn_df |>
+  select(-earliest_year) |>
+  group_by(country_code) |>
+  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) |>
+  mutate(country_group = case_when(
+    country_code %in% oecd$iso3 ~ "OECD",
+    country_code %in% c("BRA", "RUS", "IND", "CHN", "ZAF") ~ "BRICS",
+    .default = "Other"
+  ))
+
+country_article_plot_df <- country_df |>
+  select(country_code, country_group, hoad_articles, scp_first_author_articles, scp_corresponding_author_articles, wos_first_author_articles, wos_corresponding_author_articles) |>
+  pivot_longer(c(scp_first_author_articles, scp_corresponding_author_articles, wos_first_author_articles, wos_corresponding_author_articles)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "first_author_articles" ~ "First",
+    indicator == "corresponding_author_articles" ~ "Corresponding"
+  ))
+
+# Countries I want to highlight
+country_focus <- c("USA", "CHN", "IND", "DEU", "GBR", "NLD", "SWE", "FRA", "HKG", "ZAF", "HUN", "FIN", "AUT", "NOR", "IRL")
+
+articles_plot <- country_article_plot_df |>
+  filter(!is.na(country_code), hoad_articles > 10000) |>
+  ggplot(aes(x = hoad_articles, y = value, fill = country_group)) +
+  geom_point(pch = 21, size = 2.5) +
+  ggrepel::geom_text_repel(data = filter(country_article_plot_df, country_code %in% country_focus),
+    aes(label = country_code), size = 2, min.segment.length = 0,
+                force = 50) +
+  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf")) +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  coord_fixed(ratio = 1) +
+  scale_x_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  scale_y_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  theme_minimal(base_size = 10) +
+  labs(x = "Article volume (log-scale)\nhoaddata", y = "Article volume (log-scale)") +
+  theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+    legend.position = "top",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+
+## OA Uptake
+oa_prop_plot_df <- country_df |>
+  filter(!is.na(country_code), hoad_articles > 10000) |>
+  mutate(
+    hoad_oa_prop = hoad_oa_articles / hoad_articles,
+    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
+    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
+    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
+    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
+  ) |>
+  select(country_code, country_group, contains("prop")) |>
+  pivot_longer(c(scp_first_oa_prop, scp_corresponding_oa_prop, wos_first_oa_prop, wos_corresponding_oa_prop)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "first_oa_prop" ~ "First",
+    indicator == "corresponding_oa_prop" ~ "Corresponding"
+  ))
+
+oa_prop_plot <- ggplot(oa_prop_plot_df, aes(x = hoad_oa_prop, y = value, fill = country_group)) +
+  geom_point(alpha = 1, pch = 21, size = 2.5) +
+ggrepel::geom_text_repel(data = filter(oa_prop_plot_df, country_code %in% country_focus),
+    aes(label = country_code), size = 2, min.segment.length = 0,
+                force = 50) +  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf"), guide = "none") +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  coord_fixed(ratio = 1) +
+  scale_x_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  theme_minimal(base_size = 10) +
+  labs(x = "Hybrid OA Uptake\nhoadata", y = "Hybrid OA Uptake") +
+    theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+ #   strip.text.x = element_blank(),
+    legend.position = "top",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+
+## Hybrid OA via agreement
+ta_oa_country_df <- country_df |>
+  filter(!is.na(country_code), hoad_articles > 10000) |>
+  select(country_code, country_group, hoad_ta_oa_articles, scp_ta_oa_first_author_articles, scp_ta_oa_corresponding_author_articles, wos_ta_oa_first_author_articles, wos_ta_oa_corresponding_author_articles) |>
+  pivot_longer(c(scp_ta_oa_first_author_articles, scp_ta_oa_corresponding_author_articles, wos_ta_oa_first_author_articles, wos_ta_oa_corresponding_author_articles)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "ta_oa_first_author_articles" ~ "First",
+    indicator == "ta_oa_corresponding_author_articles" ~ "Corresponding"
+  ))
+
+## Hybrid OA vioa agreement share
+ta_oa_country_plot <- ta_oa_country_df |>
+  filter(hoad_ta_oa_articles > 1000) |>
+  ggplot(aes(x = hoad_ta_oa_articles, y = value, fill = country_group)) +
+  geom_point(pch = 21, size = 2.5) +
+ggrepel::geom_text_repel(data = filter(ta_oa_country_df, hoad_ta_oa_articles > 1000, country_code %in% country_focus),
+    aes(label = country_code), size = 2, min.segment.length = 0,
+                force = 50) +  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf"), guide = "none") +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  coord_fixed(ratio = 1) +
+  scale_x_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  scale_y_log10(guide = "axis_logticks", labels = scales::number_format(big.mark = ",")) +
+  theme_minimal(base_size = 10) +
+  labs(x = "Hybrid OA via Agreement (log-scale)\nhoadata", y = "Hybrid OA via Agreement (log-scale)") +
+  theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+    legend.position = "top",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+
+## Hybrid OA vio agreement share
+ta_oa_prop_plot_df <- country_df |>
+  filter(hoad_ta_oa_articles > 1000) |>
+  mutate(
+    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
+    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
+    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
+    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
+    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
+  ) |>
+  select(country_code, country_group, contains("prop")) |>
+  pivot_longer(c(scp_first_oa_prop, scp_corresponding_oa_prop, wos_first_oa_prop, wos_corresponding_oa_prop)) |>
+  mutate(scr = sub("_.*", "", name)) |>
+  mutate(indicator = sub("^[^_]*_", "", name)) |>
+  mutate(scr = case_when(
+    scr == "scp" ~ "Scopus",
+    scr == "wos" ~ "Web of Science"
+  )) |>
+  mutate(indicator = case_when(
+    indicator == "first_oa_prop" ~ "First",
+    indicator == "corresponding_oa_prop" ~ "Corresponding"
+  ))
+
+ta_oa_prop_plot <- ggplot(ta_oa_prop_plot_df, aes(x = hoad_oa_prop, y = value, fill = country_group)) +
+  geom_point(alpha = 1, pch = 21, size = 2.5) +
+  scale_fill_manual(NULL, values = c("OECD" = "#91bfdb", "BRICS" = "#fc8d59", "Other" = "#ffffbf"), guide = "none") +
+  facet_grid(vars(indicator), vars(scr)) +
+  geom_abline() +
+  geom_text_repel(data = filter(ta_oa_prop_plot_df, country_code %in% country_focus),
+                aes(label = country_code),
+                min.segment.length = 0,
+                force = 50,
+                size = 2) +
+  coord_fixed(ratio = 1) +
+  scale_x_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  scale_y_continuous(
+    labels = scales::percent_format(),
+    limits = c(0, NA),
+    expand = expansion(mult = c(0, 0.05))
+  ) +
+  theme_minimal(base_size = 10) +
+  labs(y = "% of Hybrid OA via Agreement", x = "% of Hybrid OA via Agreement\nhoadata") +
+  theme(
+    axis.ticks.x = element_line(size = 0.2),
+    axis.ticks.y = element_line(size = 0.2),
+    axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
+  ) +
+  theme(
+#    strip.text.x = element_blank(),
+    legend.position = "bottom",
+    aspect.ratio = 1,
+    panel.border = element_rect(color = "grey50", fill = NA),
+    axis.title = element_text(size = 10)
+  )
+```
+
+```{r oa_country_comparision_plot_, fig.asp=1, fig.width = 8}
+layout <- "EE
+AC
+BD"
+
+combined_oa_country_plot_ <- wrap_plots(
+  E = guide_area(),
+  A = articles_plot,
+  B = oa_prop_plot,
+  C = ta_oa_country_plot,
+  D = ta_oa_prop_plot,
+  design = layout,
+  heights = c(0,1,1)
+) &
+    theme(
+      legend.position = "top",
+      panel.border = element_rect(color = "grey50", fill = NA),
+      text = element_text(size = 12))
+
+combined_oa_country_plot_ + plot_layout(guides = "collect") +
+  plot_annotation(tag_levels = 'A')
+```
+
+results are consistent.
+most represented are chinese, usa and india in temrs of articles.
+however, in terms of oa, other country lead,
+these are in particular those with transformative agremments
+
+overall, briscs and oecd, however, notable expection is south africa
+
+en though correlation exisits, a closer lokk at some temproarl trend revela (dahsboard), which countries were affected by inconsitstent oa evidnecie.
+heres is wiley, and there is france.
+
 # discussion

### 2025-03-17 analysis/manuscript.qmd
Finish first full draft of result section

@@
 
+When comparing countries, consistent patterns were observed across data sources from 2019 to 2023. 
+Figure \ref{fig-uptake_country} presents hybrid open access indicators by country, comparing hoaddata (x-axis) with Web of Science and Scopus (y-axis).
+Indicators calculated from these proprietary databases are shown for both first and corresponding authors, with full counting used to account for multiple country affiliations.
+
 ```{r country_data_prep}
+#| fig-cap: "Open Access by country"
+#| label: fig-uptake_country
+#| 
 jn_df <- readr::read_csv(here::here("data", "jn_country_stats_all.csv")) |>
@@
 
-results are consistent.
-most represented are chinese, usa and india in temrs of articles.
-however, in terms of oa, other country lead,
-these are in particular those with transformative agremments
-
-overall, briscs and oecd, however, notable expection is south africa
-
-en though correlation exisits, a closer lokk at some temproarl trend revela (dahsboard), which countries were affected by inconsitstent oa evidnecie.
-heres is wiley, and there is france.
+In terms of article output, a strong correlation was observed across the data sources.
+Between 2019 and 2023, China emerged as the most productive country, followed by the United States and, by a wide margin, India, the United Kingdom, and Germany. 
+Analysis of authorship roles revealed minimal variation, indicating that first and corresponding authors were typically from the same country.
+
+When examining the percentage of open access articles in hybrid journals, a different pattern emerged.
+Authors affiliated with institutions in medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, provided a large proportion of their articles in open access. 
+Germany and the United Kingdom also had approximately 50% of their output available as open access.
+In contrast, non-OECD countries showed notably lower adoption of hybrid open access, with South Africa being the only BRICS member well-represented in the sample. 
+The United States also demonstrated a relatively low proportion of open access articles.
+These findings were consistent across all databases.
+However, France was better represented in Web of Science, likely due to its agreement with Elsevier starting in 2019, which allowed delayed open access under the publisher’s user license [@Rabesandratana_2019].
+This license was not classified as hybrid open access in Scopus or hoaddata.
+
+Transformative agreements appeared to be a key driver of open access growth.
+OECD members accounted for the majority of open access articles enabled by transformative agreements.
+As a notable exception, South Africa also featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward [^sa].
+Results were consistent across data sources, with a strong correlation observed. 
+However, Wiley surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with Wiley that started in July 2019.
+
+A closer examination of hybrid open access enabled by transformative agreements revealed that the French Elsevier agreement on delayed open access was not included in JCT data, as it did not qualify as a transformative agreement.
+Medium-sized European countries again showed a high proportion of hybrid open access through transformative agreements, highlighting the impact of this licensing model across all three data sources.
+In contrast, the United States had a low proportion of hybrid open access enabled by agreements, suggesting that a substantial number of open access articles were likely financed through other means, such as article processing charges (APCs).
 
@@
 
+<!-- [juan study] larger proportion of hybrid, suggest journals included which might not be covered by transformative agreement. -->
+
 [^unpaywallwiley]: See Unpaywall version history related to Wiley fixes: <https://github.com/search?q=repo%3Aourresearch%2Foadoi+wiley&type=commits>
+
+[^sa]: <https://sanlic.ac.za/read-and-publish-agreements/>

### 2025-03-17 analysis/manuscript.tex
Finish first full draft of result section

--- 2025-03-12 analysis/manuscript.tex
@@
 
-\section*{discussion}\label{discussion}
-\addcontentsline{toc}{section}{discussion}
+When comparing countries, consistent patterns were observed across data
+sources from 2019 to 2023. Figure \ref{fig-uptake_country} presents
+hybrid open access indicators by country, comparing hoaddata (x-axis)
+with Web of Science and Scopus (y-axis). Indicators calculated from
+these proprietary databases are shown for both first and corresponding
+authors, with full counting used to account for multiple country
+affiliations.
+
+\begin{center}
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/oa_country_comparision_plot_-1.pdf}
+\end{center}
+
+In terms of article output, a strong correlation was observed across the
+data sources. Between 2019 and 2023, China emerged as the most
+productive country, followed by the United States and, by a wide margin,
+India, the United Kingdom, and Germany. Analysis of authorship roles
+revealed minimal variation, indicating that first and corresponding
+authors were typically from the same country.
+
+When examining the percentage of open access articles in hybrid
+journals, a different pattern emerged. Authors affiliated with
+institutions in medium-sized European countries, such as Sweden, the
+Netherlands, Finland, and Hungary, provided a large proportion of their
+articles in open access. Germany and the United Kingdom also had
+approximately 50\% of their output available as open access. In
+contrast, non-OECD countries showed notably lower adoption of hybrid
+open access, with South Africa being the only BRICS member
+well-represented in the sample. The United States also demonstrated a
+relatively low proportion of open access articles. These findings were
+consistent across all databases. However, France was better represented
+in Web of Science, likely due to its agreement with Elsevier starting in
+2019, which allowed delayed open access under the publisher's user
+license (Rabesandratana, 2019). This license was not classified as
+hybrid open access in Scopus or hoaddata.
+
+Transformative agreements appeared to be a key driver of open access
+growth. OECD members accounted for the majority of open access articles
+enabled by transformative agreements. As a notable exception, South
+Africa also featured prominently, as the South African National Library
+and Information Consortium (SANLiC) successfully negotiated
+transformative agreements with major publishers from 2022 onward
+\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}.
+Results were consistent across data sources, with a strong correlation
+observed. However, Wiley surplus in hoaddata led to better rankings for
+countries where Wiley played a substantial role, such as Germany, where
+the DEAL consortium negotiated its first transformative agreements with
+Wiley that started in July 2019.
+
+A closer examination of hybrid open access enabled by transformative
+agreements revealed that the French Elsevier agreement on delayed open
+access was not included in JCT data, as it did not qualify as a
+transformative agreement. Medium-sized European countries again showed a
+high proportion of hybrid open access through transformative agreements,
+highlighting the impact of this licensing model across all three data
+sources. In contrast, the United States had a low proportion of hybrid
+open access enabled by agreements, suggesting that a substantial number
+of open access articles were likely financed through other means, such
+as article processing charges (APCs).
+
+\section{discussion}\label{discussion}
 
@@
 
+\bibitem[\citeproctext]{ref-Rabesandratana_2019}
+Rabesandratana, T. (2019). Elsevier deal with france disappoints
+open-access advocates. \emph{Science}.
+\url{https://doi.org/10.1126/science.aba5656}
+
 \bibitem[\citeproctext]{ref-schmidt_2024_13935407}

### 2025-03-18 analysis/manuscript.qmd
Some minor language fixes

@@
 library(ggrepel)
-library(ggpp)
 
@@
 As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
-This section introduces the initial data sources, followed by a presentation of the necessary data processing steps to obtain eligible articles from transformative agreements by using author roles (first and corresponding) and harmonised affiliation data.
+This section introduces the initial data sources, followed by a presentation of the necessary data processing steps to obtain eligible articles enabled by transformative agreements using author roles (first and corresponding) and harmonised affiliation data.
 <!-- Overall, xxxx hybrid journals from xxx agreements that published at least one open access article between 2019 and 2023. formed the basis of this study. -->
@@
 
-hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package collecting information on hybrid open access uptake from multiple openly available data sources.
+hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package collecting information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
 It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.
 
-As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of historic JCT snapshots, which were safeguarded since July 2021.
-Journal data were enriched with ISSN variants linked to an ISSN-L and insitution data with associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
+As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of various safeguarded JCT snapshots.
+Journal data were enriched with ISSN variants linked to an ISSN-L and institution data with associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
 A comprehensive exclusion of fully open access journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals.
@@
 The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
-To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow handles data retrieval from the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.
+To ensure computational reproducibility while aggregating the data, a GitHub Actions continuous integration and delivery (CI/CD) workflow handles data retrieval from the SUB Göttingen's open scholarly data warehouse based on Google BigQuery, which provides high-performant programmatic access to monthly snapshots of Crossref and OpenAlex.[^subugoe]
 The workflow has run regularly to fetch updates from these data sources since 2022.
 The package version used in this study is 0.3, containing data from the Crossref 2024-08 dump provided to Crossref Metadata Plus subscribers and the OpenAlex 2024-08-29 monthly dump.
+It covers agreements collected between July 2021 to July 2024 from the JCT.
 This version including the computation log is available on GitHub (https://github.com/subugoe/hoaddata/releases/tag/v.0.3).
@@
 Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020]. 
-The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI), collectively referred to as Web of Science Primary (WoS Primary).
-
-The WoS Primary provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as OpenAlex.
-However, it lacks information about journals and articles under transformative agreements.
+The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).
+
+These collections provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as OpenAlex.
+However, the Web of Science lacks information about journals and articles under transformative agreements.
 
 For programmatic access to article-level data, this study used the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany.
-The KB processes raw XML data provided by Clarivate Analytics, which is provided as an in-house PostgreSQL database under a uniform schema.
+The KB processes raw XML data provided by Clarivate Analytics, which is ingested into an in-house PostgreSQL database under a uniform schema.
 To support reproducibility, KB maintains annual snapshots of the database. 
@@
 
-The hoaddata package (v.0.3) provided an enriched and unified list of hybrid journals in transformative agreements according to the JCT, covering agreements active from July 2021 to July 2024 according to the ESAC registry.
+Using an enriched and unified list of hybrid journals in transformative agreements according to the JCT, article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L.
 <!-- This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
@@
 -->
-Article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L.
-The metadata included DOIs, publication dates, article types, open access information as well as author roles and institutional affiliations.
+The article metadata included DOIs, publication dates, open access information as well as author roles and institutional affiliations.
 Publication years were determined using the earliest known date of publication in a journal.
@@
 To exclude conference supplements, which are also often not covered by transformative agreements, only articles published in regular issues, indicated by numerical pagination, were considered.
-For Web of Science and Scopus, their established document type classifications were used to identify original research articles and reviews, referred to as core articles throughout this study.
+For Web of Science and Scopus, their established document type classifications were used to identify original research articles and reviews, referred to as original articles throughout this study.
 
@@
 
-Published articles were considered hybrid open access  when they were made freely available under a Creative Commons license on publishers' platforms. 
-While hoaddata sourced this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall. 
+Articles in hybrid journals were considered open access when they were made freely available under a Creative Commons license on publishers' platforms. 
+While hoaddata sourced this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall as evidence source. 
 Unpaywall supplements Crossref metadata by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018]. 
 This additional parsing remains necessary despite transformative agreement workflows recommend the depositing license information during DOI registration [@Geschuhn_2017].
-Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information.
+Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses.
 
@@
 ```
-Author affiliations were retrieved for both first and, if available, corresponding authors across all databases to prepare the linking between articles and institutions covered by transformative agreements.
+Author affiliations were retrieved for both first and, if available, corresponding authors to prepare the linking between articles and institutions covered by transformative agreements.
 To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
@@
 
-Because neither Web of Science nor Scopus support ROR IDs—the institution identifier used by the JCT—a two-step matching process was implemented to harmonise affiliation data.
+Because neither Web of Science nor Scopus support ROR IDs, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise affiliation data.
 First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
@@
 As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were aggregated for each database at country and journal level by year.
-@tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023 per database.
-It shows that the majority of hybrid journals published at least one original research article or review marked as core during the five-year period.
+@tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023.
+It shows that the majority of hybrid journals published at least one original research article or review marked as original during the five-year period.
 These journals formed the basis for the subsequent calculation of article-level indicators.
@@
       indicator == "ind_articles_total" ~ "Total published articles ",
-      indicator == "ind_articles_core" ~ "Core articles",
+      indicator == "ind_articles_core" ~ "Original articles",
       indicator == "ind_articles_doi" ~ "Articles with DOI",
-      indicator == "ind_articles_core_doi" ~ "Core articles with DOI",
+      indicator == "ind_articles_core_doi" ~ "Original articles with DOI",
       indicator == "ind_oa_articles_all" ~ "OA articles",
-      indicator == "ind_oa_articles_core" ~ "Core OA articles",
+      indicator == "ind_oa_articles_core" ~ "Original OA articles",
       indicator == "first_aff" ~ "First author articles",
@@
   pack_rows("Open Access (OA) metrics", 8, 9) |>
-  pack_rows("Core articles with affiliation data", 10, 11) 
+  pack_rows("Original articles with affiliation data", 10, 11) 
 ```
@@
 While hoaddata only covers articles with a DOI, Scopus and Web of Science publication indicators were calculated using their database identifier.
-A subsequent comparison of DOI coverage shows that non-core articles in Web of Science often lacked a DOI.
+A subsequent comparison of DOI coverage shows that non-original articles in Web of Science often lacked a DOI.
 This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
 Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
-A closer look at the core articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
+A closer look at original articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
 Only about two-thirds of the articles examined provided corresponding author affiliations.
@@
 The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
-Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus.
+Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus.
 Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in the Web of Science.
-Upon inspection, this group of hybrid journals included only in the proprietary data sources mainly represents hybrid journals for which no open access evidence could be retrieved from Crossref, which served as the open access evidence source for hoaddata.
+Upon inspection, this group of hybrid journals exclusively covered in the proprietary data sources mainly represented hybrid journals for which no open access evidence could be retrieved from Crossref, the open access evidence source for hoaddata.
 
@@
 In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` overlapping articles, represented by the blue bar.
-They represent `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(wos) * 100)`% of core articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(scp) * 100)`% in Scopus.
+They represented `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(wos) * 100)`% of original articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(scp) * 100)`% in Scopus.
 Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to both Scopus and hoaddata.
@@
 
-Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only available via hoaddata (grey area).
+Figure \ref{fig-upset_coverage_results}B also shows the surplus of articles with DOI that were only available via hoaddata (grey area).
 In case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
 After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews during the compilation of hoaddata.
-Particulalry, letters and editorials could not be fully detected.
-Paratext recognition failed in `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scientific content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science.
+Particularly, letters and editorials could not be fully detected.
+Paratext recognition failed in `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scholarly content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science.
 To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.
@@
 Furthermore, journals in this set published more on average than journals covered by only one or two data sources. 
-In particular, the journals covered exclusively by hoaddata are substantially smaller than those covered by all three sources. 
-Upon inspection, these are often newly launched journals, which explains the relatively low five-year publication volume.
+In particular, the journals covered exclusively by hoaddata were substantially smaller than those covered by all three sources. 
+Upon inspection, these were often newly launched journals, which explains the relatively low five-year publication volume.
+An example is *Digital society* that published `r jn_upset_articles |> filter(issn_l == "2731-4650") |> pull(n_cr)` articles.
+This hybrid journal was launched in 2022, being covered by various Springer Nature transformative agreements since then.
 
@@
 The analysis highlights the dominance of the three largest publishers--Elsevier, Springer Nature, and Wiley--, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of articles published during this five-year period.
-In terms of article volume, Elsevier led with `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(articles)` artricles (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_art)*100`% of the total) published across `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals(`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_jn)*100`% of the total).
-Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%), while Wiley recorded `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
+In terms of article volume, Elsevier led with `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_art)*100`% of the total) published across `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_jn)*100`% of the total).
+Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)`, although recording the largest number of hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%).
+Wiley accounted for `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
 The remaining `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(esac_publisher) |> nrow()` publishers collectively accounted for `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`%).
@@
 The three largest publishers--Elsevier, Springer Nature and Wiley--were best represented in the exclusive intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
-Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage with `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()` (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as determined through shared DOIs.
-When examining publication volume per journal, (\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals, followed by Springer Nature and Wiley.
+Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage (n = `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()`; `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as determined through shared DOIs.
+When examining publication volume per journal, (Figure \ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals, followed by Springer Nature and Wiley.
 
 A comparative analysis of publisher portfolios across different indexing sets demonstrates that publisher were not represented uniformly. 
-Springer Nature exhibited `r jn_upset_publisher |> filter(esac_publisher == "Springer Nature", WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow()` hybrid journals exclusively indexed in both hoaddata and Scopus.
+Notably, Springer Nature exhibited `r jn_upset_publisher |> filter(esac_publisher == "Springer Nature", WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow()` hybrid journals exclusively indexed in both hoaddata and Scopus.
 This set included journals from the Chinese Academy of Science, German-language medical journals, and Eastern European publications including the *Journal of Mathematical Sciences*, which also publishes English-language translations of Russian-language works.
-Additionally, this subset included titles with a more broader disciplianry focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
-The set also captured ceased journals, providing insights into the dynamic nature of journal portfolios and business strategies.
+Additionally, this subset included titles with a more broader disciplinary focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
+The set also captured ceased journals, providing further insights into the dynamic nature of journal portfolios and business strategies.
 
@@
 
-Further examination of publisher portfolios not covered by hoaddata but present in Scopus or Web of Science identified several publishers with missing Creative Commons (CC) license information in Crossref.
+Examining publisher portfolios not covered by hoaddata but present in Scopus or Web of Science identified several publishers with missing CC license information in Crossref.
 In particular, `r miss_tas$esac_publisher[1]` represented `r miss_tas$journals[1]` journals with `r miss_tas$n[1]` articles, `r miss_tas$esac_publisher[2]` accounted for `r miss_tas$journals[2]` journals with `r miss_tas$n[2]` articles, and `r miss_tas$esac_publisher[3]` recorded `r miss_tas$journals[3]` journals and `r miss_tas$n[3]` articles.
@@
 
-An inspection of individual journals uncovered discrepancies in Unpaywall's open access identification for certain publishers that typically deposit CC licenses with Crossref.
+An inspection of individual journals also uncovered discrepancies in Unpaywall's open access identification for certain publishers that typically deposit CC licenses with Crossref.
 Notably, some subscription-only journals contained one or two articles erroneously tagged as hybrid open access by Unpaywall, which were subsequently reflected in Scopus and Web of Science.
@@
 This section examines the uptake of open access in hybrid journals, focusing on the influence of transformative agreements across hoaddata, Scopus, and Web of Science.
-The analysis assesses whether consistent results can be derived from these data sources despite differences in coverage and methodologies. 
+The aim was to assess whether consistent results can be derived from these data sources despite differences in coverage and methodologies. 
 Following @Jahn_2025, indicators were calculated for each data source and comprise the number and proportion of open access articles, including those enabled by transformative agreements, from 2019 to 2023.
@@
 Similarly, Scopus recorded an growth from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
-However, Web of Science recorded higher open access uptake in early years, before converging in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
-This may reflect delayed open access being labeled as hybrid in Web of Science.
-
-Open access by transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
+However, Web of Science recorded higher open access uptake in early years, before converging in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023,suggesting an different approach towards labeling hybrid open access by the Web of Science.
+
+Simialry, hybrid open access by transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
 Trends were consistent for first (dashed line) and corresponding author (dotted line) affiliations. 
@@
 From 2021 (hoadata, Scopus) resp. 2022 (Web of Science),  transformative agreements enabled the majority of hybrid open access.
-For first authors, the share ranged between `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and  `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata) in 2023.
-For corresponding authors, the shares were slightly larger, with Web of Science recording `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100` % and Scopus  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% in 2023.
+For first authors, the share ranged between `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and   `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science) in 2023.
+For corresponding authors, the shares were slightly larger, with Scopus recording  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% and Web of Science `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% in 2023.
 However, substantial hybrid open access was still facilitated outside transformative agreements, likely through APCs payed from discretionary research funds (ref).
@@
 
-When considering open access trends by publisher (see \ref{fig-uptake_publisher}), the observed differences in early uptake rates between hoaddata and Scopus compared to Web of Science can be largely attributed to articles published in Elsevier hybrid journals, the largest publisher in our sample.
+When considering open access trends by publisher (see Figure \ref{fig-uptake_publisher}), the observed differences in early uptake rates between hoaddata and Scopus compared to Web of Science can be largely attributed to articles published in Elsevier hybrid journals, the largest publisher in our sample.
 Both hoadata and Scopus reported an steady increase in open access uptake between 2019 and 2023 (hoaddata from `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%; Scopus from `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%).
@@
 According to Unpaywall's software version history, Wiley's publisher platform redirects prevented Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
-hoaddata, which relies solely on Crossref license information, was unaffected by these issues.
+hoaddata, which relies solely on Crossref metadata for open access identification, was unaffected by these issues.
 
 Despite these differences in open access evidence, the three data sources show consistent temporal trends in hybrid open access enabled by transformative agreements (see Figure \ref{fig-uptake_publisher}B and C).
-Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across all data sources, followed by Springer Nature and Elsevier.
+Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across the examined data sources, followed by Springer Nature and Elsevier.
 By 2023, the share of open access enabled by transformative agreements appeared to stabilise for all three publishers, with Wiley recording the largest share (see Figure \ref{fig-uptake_publisher}B). 
@@
 
-When comparing countries, consistent patterns were observed across data sources from 2019 to 2023. 
+When comparing countries, consistent patterns were observed across data sources for the five-year period 2019 to 2023. 
 Figure \ref{fig-uptake_country} presents hybrid open access indicators by country, comparing hoaddata (x-axis) with Web of Science and Scopus (y-axis).
@@
 ```{r country_data_prep}
-#| fig-cap: "Open Access by country"
-#| label: fig-uptake_country
-#| 
 jn_df <- readr::read_csv(here::here("data", "jn_country_stats_all.csv")) |>
@@
 ```{r oa_country_comparision_plot_, fig.asp=1, fig.width = 8}
+#| fig-cap: "Open Access by country"
+#| label: fig-uptake_country
+#| 
 layout <- "EE
@@
 
-In terms of article output, a strong correlation was observed across the data sources.
-Between 2019 and 2023, China emerged as the most productive country, followed by the United States and, by a wide margin, India, the United Kingdom, and Germany. 
+In terms of article output (see Figure \ref{fig-uptake_country}A), a strong correlation was observed across the data sources.
+Between 2019 and 2023, China was the most productive country, followed by the United States and, by a certain margin, India, the United Kingdom, and Germany. 
 Analysis of authorship roles revealed minimal variation, indicating that first and corresponding authors were typically from the same country.
 
-When examining the percentage of open access articles in hybrid journals, a different pattern emerged.
-Authors affiliated with institutions in medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, provided a large proportion of their articles in open access. 
+When examining the percentage of open access articles in hybrid journals (see Figure \ref{fig-uptake_country}B), a different pattern emerged.
+Authors affiliated with institutions from medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, provided a large proportion of their articles in open access. 
 Germany and the United Kingdom also had approximately 50% of their output available as open access.
-In contrast, non-OECD countries showed notably lower adoption of hybrid open access, with South Africa being the only BRICS member well-represented in the sample. 
+In contrast, non-OECD countries showed notably lower adoption of hybrid open access, with South Africa being the only BRICS member well-represented in the data.
 The United States also demonstrated a relatively low proportion of open access articles.
@@
 
-Transformative agreements appeared to be a key driver of open access growth.
+Transformative agreements appeared to be a key driver of open access growth (see Figure \ref{fig-uptake_country}C and D).
 OECD members accounted for the majority of open access articles enabled by transformative agreements.
-As a notable exception, South Africa also featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward [^sa].
+As a notable exception, South Africa also featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward.[^sa]
 Results were consistent across data sources, with a strong correlation observed. 
@@
 
-A closer examination of hybrid open access enabled by transformative agreements revealed that the French Elsevier agreement on delayed open access was not included in JCT data, as it did not qualify as a transformative agreement.
+A closer examination of the proportion of hybrid open access enabled by transformative agreements (see Figure \ref{fig-uptake_country}D) revealed that the French Elsevier agreement on delayed open access was not included in JCT data, as it did not qualify as a transformative agreement.
 Medium-sized European countries again showed a high proportion of hybrid open access through transformative agreements, highlighting the impact of this licensing model across all three data sources.
@@
 
-# discussion
+# Discussion
 
@@
 [^sa]: <https://sanlic.ac.za/read-and-publish-agreements/>
+
+[^subugoe]: <https://subugoe.github.io/scholcomm_analytics/data.html>

### 2025-03-18 analysis/manuscript.tex
Some minor language fixes

@@
 followed by a presentation of the necessary data processing steps to
-obtain eligible articles from transformative agreements by using author
-roles (first and corresponding) and harmonised affiliation data.
+obtain eligible articles enabled by transformative agreements using
+author roles (first and corresponding) and harmonised affiliation data.
 
@@
 information on hybrid open access uptake from multiple openly available
-data sources. It combines article-level metadata from Crossref and
-OpenAlex with transformative agreement information from the cOAlition S
-Journal Checker Tool (JCT), which links journal and institutional data
-to agreements in the ESAC registry.
+data sources (Jahn, 2024). It combines article-level metadata from
+Crossref and OpenAlex with transformative agreement information from the
+cOAlition S Journal Checker Tool (JCT), which links journal and
+institutional data to agreements in the ESAC registry.
 
 As discussed elsewhere (Jahn, 2025), hoaddata uses a unified dataset of
-historic JCT snapshots, which were safeguarded since July 2021. Journal
-data were enriched with ISSN variants linked to an ISSN-L and insitution
-data with associated ROR-IDs representing university hospitals or
-institutes of large research organisations, according to OpenAlex'
-institution entity. A comprehensive exclusion of fully open access
-journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld) was
-performed to identify hybrid journals. hoaddata relies on Crossref for
-obtaining journal publication volume and open access status through
-Creative Commons licence information relative to the published version
-(``version of record''). Because of limited affiliation metadata in
-Crossref (Eck \& Waltman, 2022), hoaddata sources affiliation metadata
-from OpenAlex.
+various safeguarded JCT snapshots. Journal data were enriched with ISSN
+variants linked to an ISSN-L and institution data with associated
+ROR-IDs representing university hospitals or institutes of large
+research organisations, according to OpenAlex' institution entity. A
+comprehensive exclusion of fully open access journals using multiple
+journal lists (DOAJ, OpenAlex, Bielefeld) was performed to identify
+hybrid journals. hoaddata relies on Crossref for obtaining journal
+publication volume and open access status through Creative Commons
+licence information relative to the published version (``version of
+record''). Because of limited affiliation metadata in Crossref (Eck \&
+Waltman, 2022), hoaddata sources affiliation metadata from OpenAlex.
 
@@
 based on Google BigQuery, which provides high-performant programmatic
-access to monthly snapshots of Crossref and OpenAlex. The workflow has
-run regularly to fetch updates from these data sources since 2022. The
-package version used in this study is 0.3, containing data from the
-Crossref 2024-08 dump provided to Crossref Metadata Plus subscribers and
-the OpenAlex 2024-08-29 monthly dump. This version including the
-computation log is available on GitHub
+access to monthly snapshots of Crossref and OpenAlex.\footnote{\url{https://subugoe.github.io/scholcomm_analytics/data.html}}
+The workflow has run regularly to fetch updates from these data sources
+since 2022. The package version used in this study is 0.3, containing
+data from the Crossref 2024-08 dump provided to Crossref Metadata Plus
+subscribers and the OpenAlex 2024-08-29 monthly dump. It covers
+agreements collected between July 2021 to July 2024 from the JCT. This
+version including the computation log is available on GitHub
 (https://github.com/subugoe/hoaddata/releases/tag/v.0.3).
@@
 Science Citation Index Expanded (SCIE), the Social Sciences Citation
-Index (SSCI) and the Arts \& Humanities Citation Index (AHCI),
-collectively referred to as Web of Science Primary (WoS Primary).
+Index (SSCI) and the Arts \& Humanities Citation Index (AHCI).
 
-The WoS Primary provides important data points for analysing open
+These collections provides important data points for analysing open
 access: author affiliations and roles, differentiation of journal
@@
 status information derived from OurResearch's Unpaywall, the same
-provider as OpenAlex. However, it lacks information about journals and
-articles under transformative agreements.
+provider as OpenAlex. However, the Web of Science lacks information
+about journals and articles under transformative agreements.
 
@@
 processes raw XML data provided by Clarivate Analytics, which is
-provided as an in-house PostgreSQL database under a uniform schema. To
+ingested into an in-house PostgreSQL database under a uniform schema. To
 support reproducibility, KB maintains annual snapshots of the database.
@@
 
-The hoaddata package (v.0.3) provided an enriched and unified list of
-hybrid journals in transformative agreements according to the JCT,
-covering agreements active from July 2021 to July 2024 according to the
-ESAC registry.
+Using an enriched and unified list of hybrid journals in transformative
+agreements according to the JCT, article metadata were retrieved from
+each database using all ISSN variants linked to an ISSN-L.
 
-Article metadata were retrieved from each database using all ISSN
-variants linked to an ISSN-L. The metadata included DOIs, publication
-dates, article types, open access information as well as author roles
-and institutional affiliations. Publication years were determined using
-the earliest known date of publication in a journal. In hoaddata, this
-corresponded to Crossref's issued date. For Web of Science and Scopus,
-the earliest publication date was used where available, with Scopus
-dates specifically determined by the KB through version tracking of the
-raw data.
+The article metadata included DOIs, publication dates, open access
+information as well as author roles and institutional affiliations.
+Publication years were determined using the earliest known date of
+publication in a journal. In hoaddata, this corresponded to Crossref's
+issued date. For Web of Science and Scopus, the earliest publication
+date was used where available, with Scopus dates specifically determined
+by the KB through version tracking of the raw data.
 
@@
 type classifications were used to identify original research articles
-and reviews, referred to as core articles throughout this study.
+and reviews, referred to as original articles throughout this study.
 
@@
 
-Published articles were considered hybrid open access when they were
+Articles in hybrid journals were considered open access when they were
 made freely available under a Creative Commons license on publishers'
 platforms. While hoaddata sourced this information from Crossref license
-metadata, Web of Science and Scopus relied on Unpaywall. Unpaywall
-supplements Crossref metadata by parsing publisher websites directly,
-addressing cases where publishers do not provide machine-readable
-Creative Commons license information (Piwowar et al., 2018). This
-additional parsing remains necessary despite transformative agreement
-workflows recommend the depositing license information during DOI
-registration (Geschuhn \& Stone, 2017). Both Web of Science and Scopus
-defined hybrid open access consistently as content available under
-Creative Commons licenses on publisher platforms, distinguishing it from
-bronze open access that lack such explicit license information.
+metadata, Web of Science and Scopus relied on Unpaywall as evidence
+source. Unpaywall supplements Crossref metadata by parsing publisher
+websites directly, addressing cases where publishers do not provide
+machine-readable Creative Commons license information (Piwowar et al.,
+2018). This additional parsing remains necessary despite transformative
+agreement workflows recommend the depositing license information during
+DOI registration (Geschuhn \& Stone, 2017). Both Web of Science and
+Scopus defined hybrid open access consistently as content available
+under Creative Commons licenses on publisher platforms, distinguishing
+it from bronze open access that lack such explicit license information,
+or use publisher-specific licenses.
 
@@
 Author affiliations were retrieved for both first and, if available,
-corresponding authors across all databases to prepare the linking
-between articles and institutions covered by transformative agreements.
-To handle different address variants, database-specific affiliation
-identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation
-enhanced names for Web of Science, and Scopus Affiliation Identifiers
-for Scopus. Additionally, ISO country codes were retrieved for each
-author's address to compile country-level statistics. These country
-codes for Web of Science and Scopus were provided by the KB.
+corresponding authors to prepare the linking between articles and
+institutions covered by transformative agreements. To handle different
+address variants, database-specific affiliation identifiers were used:
+ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web
+of Science, and Scopus Affiliation Identifiers for Scopus. Additionally,
+ISO country codes were retrieved for each author's address to compile
+country-level statistics. These country codes for Web of Science and
+Scopus were provided by the KB.
 
-Because neither Web of Science nor Scopus support ROR IDs---the
-institution identifier used by the JCT---a two-step matching process was
+Because neither Web of Science nor Scopus support ROR IDs, the
+institution identifier used by the JCT, a two-step matching process was
 implemented to harmonise affiliation data. First, 2,782,540 articles
@@
 level by year. Table~\ref{tbl-methods_overview_tab} provides a general
-overview of the coverage between 2019 and 2023 per database. It shows
-that the majority of hybrid journals published at least one original
-research article or review marked as core during the five-year period.
-These journals formed the basis for the subsequent calculation of
+overview of the coverage between 2019 and 2023. It shows that the
+majority of hybrid journals published at least one original research
+article or review marked as original during the five-year period. These
+journals formed the basis for the subsequent calculation of
 article-level indicators.
@@
 \hspace{1em}Total published articles & 9,740,015 & 8,616,053 & 8,117,644\\
-\hspace{1em}Core articles & 8,158,425 & 6,708,083 & 7,317,703\\
+\hspace{1em}Original articles & 8,158,425 & 6,708,083 & 7,317,703\\
 \addlinespace[0.3em]
@@
 \hspace{1em}Articles with DOI & 9,740,015 & 7,713,796 & 8,105,112\\
-\hspace{1em}Core articles with DOI & 8,158,425 & 6,695,661 & 7,314,327\\
+\hspace{1em}Original articles with DOI & 8,158,425 & 6,695,661 & 7,314,327\\
 \addlinespace[0.3em]
@@
 \hspace{1em}OA articles & 998,699 & 1,112,758 & 974,099\\
-\hspace{1em}Core OA articles & 969,817 & 1,019,784 & 922,578\\
+\hspace{1em}Original OA articles & 969,817 & 1,019,784 & 922,578\\
 \addlinespace[0.3em]
-\multicolumn{4}{l}{\textbf{Core articles with affiliation data}}\\
+\multicolumn{4}{l}{\textbf{Original articles with affiliation data}}\\
 \hspace{1em}First author articles & 7,242,542 & 6,294,855 & 7,232,017\\
@@
 Science publication indicators were calculated using their database
-identifier. A subsequent comparison of DOI coverage shows that non-core
-articles in Web of Science often lacked a DOI. This was particularly the
-case for meeting abstracts, which are notably prevalent in Health
-Sciences journals (Melero-Fuentes et al., 2025) and are not indexed by
-Scopus (Donner, 2017). Open access indicators were aggregated by DOI, as
-Unpaywall only collects information on open access status for articles
-with a DOI. A closer look at the core articles with affiliation data
-reveals a lack of data on corresponding authors in the case of OpenAlex,
-the affiliation data source used by hoaddata, compared to Web of Science
-and Scopus. Only about two-thirds of the articles examined provided
-corresponding author affiliations. For first authors, the proportion was
-89\%. Therefore, only first author data for hoaddata were considered in
-the following analysis.
+identifier. A subsequent comparison of DOI coverage shows that
+non-original articles in Web of Science often lacked a DOI. This was
+particularly the case for meeting abstracts, which are notably prevalent
+in Health Sciences journals (Melero-Fuentes et al., 2025) and are not
+indexed by Scopus (Donner, 2017). Open access indicators were aggregated
+by DOI, as Unpaywall only collects information on open access status for
+articles with a DOI. A closer look at original articles with affiliation
+data reveals a lack of data on corresponding authors in the case of
+OpenAlex, the affiliation data source used by hoaddata, compared to Web
+of Science and Scopus. Only about two-thirds of the articles examined
+provided corresponding author affiliations. For first authors, the
+proportion was \%. Therefore, only first author data for hoaddata were
+considered in the following analysis.
 
@@
 comprising 21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739)
-of journals were exclusively contained in hoaadata, while 6\% (n = 748)
-were only found in the proprietory database Scopus. Of these, 354 were
-also available in the Web of Science. Upon inspection, this group of
-hybrid journals included only in the proprietary data sources mainly
-represents hybrid journals for which no open access evidence could be
-retrieved from Crossref, which served as the open access evidence source
-for hoaddata.
+of journals were exclusively contained in hoaadata, while another 6\% (n
+= 748) were only found in the proprietory database Scopus. Of these, 354
+were also available in the Web of Science. Upon inspection, this group
+of hybrid journals exclusively covered in the proprietary data sources
+mainly represented hybrid journals for which no open access evidence
+could be retrieved from Crossref, the open access evidence source for
+hoaddata.
 
@@
 total, these journals recorded 6,289,687 overlapping articles,
-represented by the blue bar. They represent 94\% of core articles with
-DOI indexed in the Web of Science sample, and 86\% in Scopus. Another
+represented by the blue bar. They represented \% of original articles
+with DOI indexed in the Web of Science sample, and \% in Scopus. Another
 657,697 articles were exclusive to both Scopus and hoaddata. Exclusively
@@
 
-Figure \ref{fig-upset_coverage_results}B also shows DOIs that were only
-available via hoaddata (grey area). In case of hybrid journals covered
-by all three data sources, 1,023,882 DOIs were only present in hoaddata.
-After validation at the DOI level using the KB databases and manual
-inspection, the main reasons for missing DOI coverage in the proprietary
-database were insufficient classification of journal content as original
-research articles and reviews during the compilation of hoaddata.
-Particulalry, letters and editorials could not be fully detected.
-Paratext recognition failed in 37\% of DOIs to identify non-scientific
-content such as front matters or reviewer lists, which are generally not
-indexed by Scopus and Web of Science. To a lesser extent, differences in
-publication and indexing dates were a reason for non-overlapping DOIs.
+Figure \ref{fig-upset_coverage_results}B also shows the surplus of
+articles with DOI that were only available via hoaddata (grey area). In
+case of hybrid journals covered by all three data sources, 1,023,882
+DOIs were only present in hoaddata. After validation at the DOI level
+using the KB databases and manual inspection, the main reasons for
+missing DOI coverage in the proprietary database were insufficient
+classification of journal content as original research articles and
+reviews during the compilation of hoaddata. Particularly, letters and
+editorials could not be fully detected. Paratext recognition failed in
+37\% of DOIs to identify non-scholarly content such as front matters or
+reviewer lists, which are generally not indexed by Scopus and Web of
+Science. To a lesser extent, differences in publication and indexing
+dates were a reason for non-overlapping DOIs.
 
@@
 published more on average than journals covered by only one or two data
-sources. In particular, the journals covered exclusively by hoaddata are
-substantially smaller than those covered by all three sources. Upon
-inspection, these are often newly launched journals, which explains the
-relatively low five-year publication volume.
+sources. In particular, the journals covered exclusively by hoaddata
+were substantially smaller than those covered by all three sources. Upon
+inspection, these were often newly launched journals, which explains the
+relatively low five-year publication volume. An example is \emph{Digital
+society} that published 86 articles. This hybrid journal was launched in
+2022, being covered by various Springer Nature transformative agreements
+since then.
 
@@
 hybrid journals and 62\% of articles published during this five-year
-period. In terms of article volume, Elsevier led with 2,441,358
-artricles (33\% of the total) published across 1,951 hybrid
-journals(16\% of the total). Springer Nature followed with 1,247,578
-articles (17\%) in 2,311 hybrid journals (19\%), while Wiley recorded
-858,939 articles (12\%) in1,382 hybrid journals (11\%). The remaining 54
-publishers collectively accounted for 2,749,847 articles (38\%) in 6,476
-hybrid journals (53\%).
+period. In terms of article volume, Elsevier led with 2,441,358 articles
+(33\% of the total) published across 1,951 hybrid journals (16\% of the
+total). Springer Nature followed with 1,247,578 articles (17\%) in
+2,311, although recording the largest number of hybrid journals (19\%).
+Wiley accounted for 858,939 articles (12\%) in1,382 hybrid journals
+(11\%). The remaining 54 publishers collectively accounted for 2,749,847
+articles (38\%) in 6,476 hybrid journals (53\%).
 
@@
 hybrid journals (55\% of the intersectional set) and dominated article
-coverage with 4,174,315 (66\%), as determined through shared DOIs. When
-examining publication volume per journal,
-(\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
+coverage (n = 4,174,315; 66\%), as determined through shared DOIs. When
+examining publication volume per journal, (Figure
+\ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
 average, the largest journals, followed by Springer Nature and Wiley.
@@
 sets demonstrates that publisher were not represented uniformly.
-Springer Nature exhibited 519 hybrid journals exclusively indexed in
-both hoaddata and Scopus. This set included journals from the Chinese
-Academy of Science, German-language medical journals, and Eastern
-European publications including the \emph{Journal of Mathematical
-Sciences}, which also publishes English-language translations of
-Russian-language works. Additionally, this subset included titles with a
-more broader disciplianry focus such as \emph{SN Computer Science} and
-newly launched hybrid journals like \emph{Nature Computational Science},
-which started in 2021 and was indexed in Scopus but not yet in Web of
-Science. The set also captured ceased journals, providing insights into
-the dynamic nature of journal portfolios and business strategies.
+Notably, Springer Nature exhibited 519 hybrid journals exclusively
+indexed in both hoaddata and Scopus. This set included journals from the
+Chinese Academy of Science, German-language medical journals, and
+Eastern European publications including the \emph{Journal of
+Mathematical Sciences}, which also publishes English-language
+translations of Russian-language works. Additionally, this subset
+included titles with a more broader disciplinary focus such as \emph{SN
+Computer Science} and newly launched hybrid journals like \emph{Nature
+Computational Science}, which started in 2021 and was indexed in Scopus
+but not yet in Web of Science. The set also captured ceased journals,
+providing further insights into the dynamic nature of journal portfolios
+and business strategies.
 
-Further examination of publisher portfolios not covered by hoaddata but
-present in Scopus or Web of Science identified several publishers with
-missing Creative Commons (CC) license information in Crossref. In
-particular, Emerald represented 322 journals with 86,409 articles, AIP
-Publishing accounted for 24 journals with 64,898 articles, and World
-Scientific recorded 87 journals and 42,531 articles. In total, 9
-publishers were not represented in Crossref.
+Examining publisher portfolios not covered by hoaddata but present in
+Scopus or Web of Science identified several publishers with missing CC
+license information in Crossref. In particular, Emerald represented 322
+journals with 86,409 articles, AIP Publishing accounted for 24 journals
+with 64,898 articles, and World Scientific recorded 87 journals and
+42,531 articles. In total, 9 publishers were not represented in
+Crossref.
 
-An inspection of individual journals uncovered discrepancies in
+An inspection of individual journals also uncovered discrepancies in
 Unpaywall's open access identification for certain publishers that
@@
 focusing on the influence of transformative agreements across hoaddata,
-Scopus, and Web of Science. The analysis assesses whether consistent
+Scopus, and Web of Science. The aim was to assess whether consistent
 results can be derived from these data sources despite differences in
@@
 converging in 2023, from 11\% (n = 137,202) in 2019 to 20\% (n =
-255,481) in 2023. This may reflect delayed open access being labeled as
-hybrid in Web of Science.
+255,481) in 2023,suggesting an different approach towards labeling
+hybrid open access by the Web of Science.
 
-Open access by transformative agreements increased between 2019 and 2023
-(Figures \ref{fig-uptake_overview}B and C). Trends were consistent for
-first (dashed line) and corresponding author (dotted line) affiliations.
-According to Scopus, 479,297 open access articles could be attributed to
-transformative agreements based on first author metadata (increasing
-from 15,341 to 206,084) and 489,262 using corresponding author metadata
-(from 15,444 to 210,816). Web of Science recorded 493,028 open access
-articles via transformative agreements using first author metadata
-(increasing from 21,871 to 174,126) and 500,076 using corresponding
-author metadata (from 22,092 to 177,030). hoaddata, lacking
-corresponding author data, linked 501,649 articles to transformative
-agreements but showed slower growth than Scopus and Web of Science, from
-16,010 to 183,757.
+Simialry, hybrid open access by transformative agreements increased
+between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C). Trends
+were consistent for first (dashed line) and corresponding author (dotted
+line) affiliations. According to Scopus, 479,297 open access articles
+could be attributed to transformative agreements based on first author
+metadata (increasing from 15,341 to 206,084) and 489,262 using
+corresponding author metadata (from 15,444 to 210,816). Web of Science
+recorded 493,028 open access articles via transformative agreements
+using first author metadata (increasing from 21,871 to 174,126) and
+500,076 using corresponding author metadata (from 22,092 to 177,030).
+hoaddata, lacking corresponding author data, linked 501,649 articles to
+transformative agreements but showed slower growth than Scopus and Web
+of Science, from 16,010 to 183,757.
 
@@
 agreements enabled the majority of hybrid open access. For first
-authors, the share ranged between 68 \% (Web of Science), 64\% (Scopus),
-and 61\% (hoaddata) in 2023. For corresponding authors, the shares were
-slightly larger, with Web of Science recording 69 \% and Scopus 65\% in
+authors, the share ranged between 61\% (hoaddata), 64\% (Scopus), and 68
+\% (Web of Science) in 2023. For corresponding authors, the shares were
+slightly larger, with Scopus recording 65\% and Web of Science 69\% in
 2023. However, substantial hybrid open access was still facilitated
@@
 
-When considering open access trends by publisher (see
+When considering open access trends by publisher (see Figure
 \ref{fig-uptake_publisher}), the observed differences in early uptake
@@
   \url{https://github.com/search?q=repo\%3Aourresearch\%2Foadoi+wiley&type=commits}}
-hoaddata, which relies solely on Crossref license information, was
-unaffected by these issues.
+hoaddata, which relies solely on Crossref metadata for open access
+identification, was unaffected by these issues.
 
@@
 access uptake, with more than 30\% of articles in hybrid journals
-reported as open access in 2023 across all data sources, followed by
-Springer Nature and Elsevier. By 2023, the share of open access enabled
-by transformative agreements appeared to stabilise for all three
-publishers, with Wiley recording the largest share (see Figure
+reported as open access in 2023 across the examined data sources,
+followed by Springer Nature and Elsevier. By 2023, the share of open
+access enabled by transformative agreements appeared to stabilise for
+all three publishers, with Wiley recording the largest share (see Figure
 \ref{fig-uptake_publisher}B). Elsevier, however, showed a comparable
@@
 When comparing countries, consistent patterns were observed across data
-sources from 2019 to 2023. Figure \ref{fig-uptake_country} presents
-hybrid open access indicators by country, comparing hoaddata (x-axis)
-with Web of Science and Scopus (y-axis). Indicators calculated from
-these proprietary databases are shown for both first and corresponding
-authors, with full counting used to account for multiple country
-affiliations.
+sources for the five-year period 2019 to 2023. Figure
+\ref{fig-uptake_country} presents hybrid open access indicators by
+country, comparing hoaddata (x-axis) with Web of Science and Scopus
+(y-axis). Indicators calculated from these proprietary databases are
+shown for both first and corresponding authors, with full counting used
+to account for multiple country affiliations.
 
-\begin{center}
-\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/oa_country_comparision_plot_-1.pdf}
-\end{center}
+\begin{figure}[ht!]
 
-In terms of article output, a strong correlation was observed across the
-data sources. Between 2019 and 2023, China emerged as the most
-productive country, followed by the United States and, by a wide margin,
-India, the United Kingdom, and Germany. Analysis of authorship roles
-revealed minimal variation, indicating that first and corresponding
-authors were typically from the same country.
+\centering{
 
-When examining the percentage of open access articles in hybrid
-journals, a different pattern emerged. Authors affiliated with
-institutions in medium-sized European countries, such as Sweden, the
-Netherlands, Finland, and Hungary, provided a large proportion of their
-articles in open access. Germany and the United Kingdom also had
-approximately 50\% of their output available as open access. In
-contrast, non-OECD countries showed notably lower adoption of hybrid
-open access, with South Africa being the only BRICS member
-well-represented in the sample. The United States also demonstrated a
-relatively low proportion of open access articles. These findings were
-consistent across all databases. However, France was better represented
-in Web of Science, likely due to its agreement with Elsevier starting in
-2019, which allowed delayed open access under the publisher's user
-license (Rabesandratana, 2019). This license was not classified as
-hybrid open access in Scopus or hoaddata.
+\includegraphics[width=0.99\linewidth,height=\textheight,keepaspectratio]{fig/fig-uptake_country-1.pdf}
+
+}
+
+\caption{\label{fig-uptake_country}Open Access by country}
+
+\end{figure}%
+
+In terms of article output (see Figure \ref{fig-uptake_country}A), a
+strong correlation was observed across the data sources. Between 2019
+and 2023, China was the most productive country, followed by the United
+States and, by a certain margin, India, the United Kingdom, and Germany.
+Analysis of authorship roles revealed minimal variation, indicating that
+first and corresponding authors were typically from the same country.
+
+When examining the percentage of open access articles in hybrid journals
+(see Figure \ref{fig-uptake_country}B), a different pattern emerged.
+Authors affiliated with institutions from medium-sized European
+countries, such as Sweden, the Netherlands, Finland, and Hungary,
+provided a large proportion of their articles in open access. Germany
+and the United Kingdom also had approximately 50\% of their output
+available as open access. In contrast, non-OECD countries showed notably
+lower adoption of hybrid open access, with South Africa being the only
+BRICS member well-represented in the data. The United States also
+demonstrated a relatively low proportion of open access articles. These
+findings were consistent across all databases. However, France was
+better represented in Web of Science, likely due to its agreement with
+Elsevier starting in 2019, which allowed delayed open access under the
+publisher's user license (Rabesandratana, 2019). This license was not
+classified as hybrid open access in Scopus or hoaddata.
 
 Transformative agreements appeared to be a key driver of open access
-growth. OECD members accounted for the majority of open access articles
-enabled by transformative agreements. As a notable exception, South
-Africa also featured prominently, as the South African National Library
-and Information Consortium (SANLiC) successfully negotiated
-transformative agreements with major publishers from 2022 onward
-\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}.
+growth (see Figure \ref{fig-uptake_country}C and D). OECD members
+accounted for the majority of open access articles enabled by
+transformative agreements. As a notable exception, South Africa also
+featured prominently, as the South African National Library and
+Information Consortium (SANLiC) successfully negotiated transformative
+agreements with major publishers from 2022 onward.\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}
 Results were consistent across data sources, with a strong correlation
@@
 
-A closer examination of hybrid open access enabled by transformative
-agreements revealed that the French Elsevier agreement on delayed open
-access was not included in JCT data, as it did not qualify as a
-transformative agreement. Medium-sized European countries again showed a
-high proportion of hybrid open access through transformative agreements,
+A closer examination of the proportion of hybrid open access enabled by
+transformative agreements (see Figure \ref{fig-uptake_country}D)
+revealed that the French Elsevier agreement on delayed open access was
+not included in JCT data, as it did not qualify as a transformative
+agreement. Medium-sized European countries again showed a high
+proportion of hybrid open access through transformative agreements,
 highlighting the impact of this licensing model across all three data
@@
 
-\section{discussion}\label{discussion}
+\section{Discussion}\label{discussion}
 
@@
 \url{https://arxiv.org/abs/2407.16551}
+
+\bibitem[\citeproctext]{ref-hoaddata}
+Jahn, N. (2024). \emph{Hoaddata: Data about hybrid open access journal
+publishing}. \url{https://github.com/subugoe/hoaddata}
 

### 2025-03-19 analysis/cor_tables.qmd
Prepare correlation tests

---
title: "Correlation tests"
format: 
    gfm:
      echo: true
      message: false
      warning: false
      code-fold: true
---

## About

The aim of this document is to determine the correlation between coefficients for several hybrid open access indicators between hoaddata, Web of Science and Scopus.

## Data preparation

```{r lood_data}
library(tidyverse)
library(corrr)
library(stats)
library(Hmisc)
library(here)
library(gt)

# Load data
jn_df <- readr::read_csv(here("data", "/jn_country_stats_all.csv")) |>
  filter(issn_l != "0027-8424") |>
  mutate(earliest_year = as.character(earliest_year))

# Calculate indicators by country and year
my_df_ind <- jn_df |> 
  select(-earliest_year) |> 
  group_by(country_code) |> 
  summarise(across(where(is.numeric), ~ sum(.x, na.rm = TRUE))) 
```

Automate reporting

```{r automate_tables}
#' Create a correlation matrix with significance markers
#'
#' @param data A data frame containing the variables for correlation analysis
#' @param vars A character vector of variable names to include in the correlation matrix
#' @param group_name A string to be included in the title of the correlation matrix
#' @return A list with the correlation data frame and title
#' @importFrom dplyr select all_of
#' @importFrom Hmisc rcorr
#' @importFrom stringr str_to_title
create_correlation_table <- function(data, vars, group_name) {
  # Check if all variables exist in the data
  missing_vars <- setdiff(vars, names(data))
  if (length(missing_vars) > 0) {
    warning(paste("The following variables are not in the dataset:", 
                  paste(missing_vars, collapse = ", ")))
    vars <- intersect(vars, names(data))
  }

  # Select only the specified variables
  subset_data <- data %>% select(all_of(vars))

  # Calculate Spearman correlation
  corr_result <- rcorr(as.matrix(subset_data), type = "spearman")
  r <- corr_result$r
  p <- corr_result$P

  # Create a dataframe for the table
  corr_df <- as.data.frame(r)

  # Add significance markers
  for (i in 1:nrow(r)) {
    for (j in 1:ncol(r)) {
      if (!is.na(p[i, j])) {
        if (p[i, j] < 0.001) {
          corr_df[i, j] <- paste0(round(r[i, j], 2), "***")
        } else if (p[i, j] < 0.01) {
          corr_df[i, j] <- paste0(round(r[i, j], 2), "**")
        } else if (p[i, j] < 0.05) {
          corr_df[i, j] <- paste0(round(r[i, j], 2), "*")
        } else {
          corr_df[i, j] <- round(r[i, j], 2)
        }
      }
    }
  }

  # Set diagonal to "-"
  for (i in 1:nrow(corr_df)) {
    corr_df[i, i] <- "-"
  }

  # Create a nicer version of variable names for display
  nice_names <- gsub("_", " ", vars)
  nice_names <- stringr::str_to_title(nice_names)

  # Set row and column names
  rownames(corr_df) <- nice_names
  colnames(corr_df) <- nice_names

  return(list(df = corr_df, title = paste("Spearman Correlation Matrix -", group_name)))
}

#' Display correlation matrix using gt package with row names
#'
#' @param corr_list A list containing the correlation data frame and title
#' @return A gt table object
#' @importFrom gt gt tab_header tab_footnote cells_body everything tab_options fmt_number
create_gt_upper_triangle <- function(corr_list) {
  footnote_text <- c("* p < 0.05", "** p < 0.01", "*** p < 0.001")

  # Convert the correlation dataframe to include row names as a column
  corr_df <- corr_list$df
  corr_df <- rownames_to_column(corr_df, var = "Variable")

  # Create the gt table
  gt_table <- gt(corr_df) %>%
    # Add title as header
    tab_header(
      title = corr_list$title
    ) %>%
    # Set the Variable column as row labels
    tab_stubhead(label = "Variable") %>%
    # Add footnotes for significance levels
    tab_footnote(
      footnote = footnote_text[1],
      locations = cells_body(
        columns = everything(),
        rows = everything()
      )
    ) %>%
    tab_footnote(
      footnote = footnote_text[2],
      locations = cells_body(
        columns = everything(),
        rows = everything()
      )
    ) %>%
    tab_footnote(
      footnote = footnote_text[3],
      locations = cells_body(
        columns = everything(),
        rows = everything()
      )
    ) %>%
    # Add styling
    tab_options(
      heading.title.font.size = px(16),
      heading.subtitle.font.size = px(13),
      table.font.size = px(12),
      column_labels.font.weight = "bold",
      table.width = pct(100),
      column_labels.background.color = "#f8f8f8",
      row_group.background.color = "#f8f8f8"
    ) %>%
    # Format any numeric columns
    fmt_number(
      columns = where(is.numeric),
      decimals = 2
    )

  return(gt_table)
}
```

##  Results

### Article volume 2019-23

```{r}

cor_articles <- create_correlation_table(my_df_ind, vars = c("hoad_articles", "scp_first_author_articles", "wos_first_author_articles", 
            "scp_corresponding_author_articles", "wos_corresponding_author_articles"), "Article volume")
create_gt_upper_triangle(cor_articles)
```

### 2025-03-19 analysis/cor_tables.qmd
Try again

@@
       code-fold: true
+prefer-html: true
 ---
@@
 ```{r}
-
 cor_articles <- create_correlation_table(my_df_ind, vars = c("hoad_articles", "scp_first_author_articles", "wos_first_author_articles", 

### 2025-03-19 analysis/cor_tables.qmd
Convert to pandoc

@@
 library(here)
-library(gt)
 
@@
 
-#' Display correlation matrix using gt package with row names
+#' Display correlation matrix using Pandoc table with row names
 #'
 #' @param corr_list A list containing the correlation data frame and title
-#' @return A gt table object
-#' @importFrom gt gt tab_header tab_footnote cells_body everything tab_options fmt_number
-create_gt_upper_triangle <- function(corr_list) {
+#' @return A character string containing the Pandoc table
+create_pandoc_table <- function(corr_list) {
   footnote_text <- c("* p < 0.05", "** p < 0.01", "*** p < 0.001")
@@
 
-  # Create the gt table
-  gt_table <- gt(corr_df) %>%
-    # Add title as header
-    tab_header(
-      title = corr_list$title
-    ) %>%
-    # Set the Variable column as row labels
-    tab_stubhead(label = "Variable") %>%
-    # Add footnotes for significance levels
-    tab_footnote(
-      footnote = footnote_text[1],
-      locations = cells_body(
-        columns = everything(),
-        rows = everything()
-      )
-    ) %>%
-    tab_footnote(
-      footnote = footnote_text[2],
-      locations = cells_body(
-        columns = everything(),
-        rows = everything()
-      )
-    ) %>%
-    tab_footnote(
-      footnote = footnote_text[3],
-      locations = cells_body(
-        columns = everything(),
-        rows = everything()
-      )
-    ) %>%
-    # Add styling
-    tab_options(
-      heading.title.font.size = px(16),
-      heading.subtitle.font.size = px(13),
-      table.font.size = px(12),
-      column_labels.font.weight = "bold",
-      table.width = pct(100),
-      column_labels.background.color = "#f8f8f8",
-      row_group.background.color = "#f8f8f8"
-    ) %>%
-    # Format any numeric columns
-    fmt_number(
-      columns = where(is.numeric),
-      decimals = 2
-    )
+  # Add title
+  table_title <- paste("#", corr_list$title, "\n\n")
 
-  return(gt_table)
+  # Create the Pandoc table
+  pandoc_table <- paste(capture.output(knitr::kable(corr_df, format = "pandoc")), collapse = "\n")
+
+  # Add footnotes
+  footnotes <- paste(footnote_text, collapse = "\n")
+
+  return(paste(table_title, pandoc_table, "\n\n", footnotes, sep = "\n"))
 }
+
 ```
@@
 ```{r}
+#| results: asis
+
 cor_articles <- create_correlation_table(my_df_ind, vars = c("hoad_articles", "scp_first_author_articles", "wos_first_author_articles", 
             "scp_corresponding_author_articles", "wos_corresponding_author_articles"), "Article volume")
-create_gt_upper_triangle(cor_articles)
+cat(create_pandoc_table(cor_articles))
 ```

### 2025-03-19 analysis/cor_tables.qmd
Testing correlation

@@
       code-fold: true
-prefer-html: true
+bibliography: references.bib
 ---
@@
 
-The aim of this document is to determine the correlation between coefficients for several hybrid open access indicators between hoaddata, Web of Science and Scopus.
+The aim of this document is to determine the correlation coefficients for several hybrid open access indicators between hoaddata, Web of Science and Scopus, focusing on country affiliations between 2019 and 2023.
+Related research have compared open scholarly data sources and propietory databases at the country-level using Spearman's rank correlation coefficient [@alperin2024analysissuitabilityopenalexbibliometric] and Kendal tau correlation [@Akbaritabar_2024], showing high level of correlations.
 
@@
 
-  # Create a nicer version of variable names for display
-  nice_names <- gsub("_", " ", vars)
-  nice_names <- stringr::str_to_title(nice_names)
-
-  # Set row and column names
-  rownames(corr_df) <- nice_names
-  colnames(corr_df) <- nice_names
-
-  return(list(df = corr_df, title = paste("Spearman Correlation Matrix -", group_name)))
+  return(list(df = corr_df, title = paste0(group_name, "--Spearman Correlation Matrix")))
 }
@@
 create_pandoc_table <- function(corr_list) {
-  footnote_text <- c("* p < 0.05", "** p < 0.01", "*** p < 0.001")
+  footnote_text <- c("\\* p < 0.05", "** p < 0.01", "*** p < 0.001")
 
@@
   # Add title
-  table_title <- paste("#", corr_list$title, "\n\n")
+  table_title <- paste("###", corr_list$title, "\n\n")
 
@@
 
-### Article volume 2019-23
-
 ```{r}
@@
 ```
+
+```{r}
+#| results: asis
+
+oa_articles <- create_correlation_table(my_df_ind, vars =  c("hoad_oa_articles", "scp_oa_first_corresponding_author_articles", "scp_oa_corresponding_author_articles",
+            "wos_oa_first_author_articles", "wos_oa_corresponding_author_articles"), "Open Access volume")
+cat(create_pandoc_table(oa_articles))
+```
+
+```{r}
+#| results: asis
+oa_prop_ind <- my_df_ind |>
+  # As in Figure
+  mutate(
+    hoad_oa_prop = hoad_oa_articles / hoad_articles,
+    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
+    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
+    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
+    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
+  ) |>
+  select(country_code, contains("prop")) |>
+  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
+            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "Open Access Uptake")
+cat(create_pandoc_table(oa_prop_ind))
+```
+
+```{r}
+#| results: asis
+oa_prop_ind <- my_df_ind |>
+  # As in Figure
+  filter(!is.na(country_code), hoad_articles > 10000) |>
+  mutate(
+    hoad_oa_prop = hoad_oa_articles / hoad_articles,
+    scp_first_oa_prop = scp_oa_first_author_articles / scp_first_author_articles,
+    scp_corresponding_oa_prop = scp_oa_corresponding_author_articles / scp_corresponding_author_articles,
+    wos_first_oa_prop = wos_oa_first_author_articles / wos_first_author_articles,
+    wos_corresponding_oa_prop = wos_oa_corresponding_author_articles / wos_corresponding_author_articles
+  ) |>
+  select(country_code, contains("prop")) |>
+  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
+            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "Open Access Uptake (min 1.000 articles)")
+cat(create_pandoc_table(oa_prop_ind))
+```
+
+```{r}
+#| results: asis
+
+ta_oa_articles <- create_correlation_table(my_df_ind, vars = c("hoad_ta_oa_articles", "scp_ta_oa_first_corresponding_author_articles", "scp_ta_oa_corresponding_author_articles",
+            "wos_ta_oa_first_author_articles", "wos_ta_oa_corresponding_author_articles"), "Open Access volume via Transformative Agreements")
+cat(create_pandoc_table(ta_oa_articles))
+```
+
+```{r}
+#| results: asis
+ta_oa_articles <- my_df_ind |>
+    filter(!is.na(country_code), hoad_articles > 10000) |>
+  create_correlation_table(vars = c("hoad_ta_oa_articles", "scp_ta_oa_first_corresponding_author_articles", "scp_ta_oa_corresponding_author_articles",
+            "wos_ta_oa_first_author_articles", "wos_ta_oa_corresponding_author_articles"), "Open Access volume via Transformative Agreements (min 1,000 articles)")
+cat(create_pandoc_table(ta_oa_articles))
+```
+
+```{r}
+#| results: asis
+
+ta_oa_prop <- my_df_ind |>
+ #filter(hoad_ta_oa_articles > 1000) |>
+  mutate(
+    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
+    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
+    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
+    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
+    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
+  ) |>
+  select(country_code, contains("prop")) |>
+  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
+            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "% of Hybrid OA via Agreement")
+cat(create_pandoc_table(ta_oa_prop))
+```
+
+```{r}
+#| results: asis
+
+ta_oa_prop <- my_df_ind |>
+ filter(hoad_ta_oa_articles > 1000) |>
+  mutate(
+    hoad_oa_prop = hoad_ta_oa_articles / hoad_oa_articles,
+    scp_first_oa_prop = scp_ta_oa_first_author_articles / scp_oa_first_author_articles,
+    scp_corresponding_oa_prop = scp_ta_oa_corresponding_author_articles / scp_oa_corresponding_author_articles,
+    wos_first_oa_prop = wos_ta_oa_first_author_articles / wos_oa_first_author_articles,
+    wos_corresponding_oa_prop = wos_ta_oa_corresponding_author_articles / wos_oa_corresponding_author_articles
+  ) |>
+  select(country_code, contains("prop")) |>
+  create_correlation_table(vars =  c("hoad_oa_prop", "scp_first_oa_prop", 
+            "scp_corresponding_oa_prop", "wos_first_oa_prop", "wos_corresponding_oa_prop"), "% of Hybrid OA via Agreement (min 1.000 OA articles)")
+cat(create_pandoc_table(ta_oa_prop))
+```
+
+# References

### 2025-03-20 analysis/manuscript.qmd
Report spearman

@@
 
-In terms of article output (see Figure \ref{fig-uptake_country}A), a strong correlation was observed across the data sources.
+In terms of article output by country (see Figure \ref{fig-uptake_country}A), strong positive correlation was observed across the data sources and author roles (Spearman rank correlation $\rho > .9$, $p < 0.001$).
 Between 2019 and 2023, China was the most productive country, followed by the United States and, by a certain margin, India, the United Kingdom, and Germany. 
@@
 Authors affiliated with institutions from medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, provided a large proportion of their articles in open access.
-Germany and the United Kingdom also had approximately 50% of their output available as open access.
+Germany and the United Kingdom also had approximately 40% of their output available as open access.
 In contrast, non-OECD countries showed notably lower adoption of hybrid open access, with South Africa being the only BRICS member well-represented in the data.
@@
 This license was not classified as hybrid open access in Scopus or hoaddata.
+In all cases, Spearman rank correlations were $\rho > .9$, $p < 0.001$, showing a high level of correlations between the databases and authorship roles considered.
 
@@
 
+In all cases, strong positive correlations were observed using Spearman's rank correlation: 
+$\rho > .9$, $p < 0.001$ between data sources and authorship roles, when considering countries with an minimum of 1,000 open access articles enabled by transformative agreements between 2019 and 2023.
+When this limitation was removed, the correlation remained strong but slightly lower ($\rho > .85$, $p < 0.001$).
+This difference may signal countries where only a few institutions had transformative agreements in place, as opposed to those participating in national consortia with broader participation.
+
 # Discussion

### 2025-03-20 analysis/manuscript.tex
Report spearman

@@
 
-In terms of article output (see Figure \ref{fig-uptake_country}A), a
-strong correlation was observed across the data sources. Between 2019
-and 2023, China was the most productive country, followed by the United
-States and, by a certain margin, India, the United Kingdom, and Germany.
-Analysis of authorship roles revealed minimal variation, indicating that
-first and corresponding authors were typically from the same country.
+In terms of article output by country (see Figure
+\ref{fig-uptake_country}A), strong positive correlation was observed
+across the data sources and author roles (Spearman rank correlation
+\(\rho > .9\), \(p < 0.001\)). Between 2019 and 2023, China was the most
+productive country, followed by the United States and, by a certain
+margin, India, the United Kingdom, and Germany. Analysis of authorship
+roles revealed minimal variation, indicating that first and
+corresponding authors were typically from the same country.
 
@@
 provided a large proportion of their articles in open access. Germany
-and the United Kingdom also had approximately 50\% of their output
+and the United Kingdom also had approximately 40\% of their output
 available as open access. In contrast, non-OECD countries showed notably
@@
 publisher's user license (Rabesandratana, 2019). This license was not
-classified as hybrid open access in Scopus or hoaddata.
+classified as hybrid open access in Scopus or hoaddata. In all cases,
+Spearman rank correlations were \(\rho > .9\), \(p < 0.001\), showing a
+high level of correlations between the databases and authorship roles
+considered.
 
@@
 as article processing charges (APCs).
+
+In all cases, strong positive correlations were observed using
+Spearman's rank correlation: \(\rho > .9\), \(p < 0.001\) between data
+sources and authorship roles, when considering countries with an minimum
+of 1,000 open access articles enabled by transformative agreements
+between 2019 and 2023. When this limitation was removed, the correlation
+remained strong but slightly lower (\(\rho > .85\), \(p < 0.001\)). This
+difference may signal countries where only a few institutions had
+transformative agreements in place, as opposed to those participating in
+national consortia with broader participation.
 

### 2025-03-27 analysis/manuscript.qmd
Frit draft of discussion (limitation, furutre studies, conclusion lacking)

@@
 \caption{Data processing workflow for comparing hybrid open access uptake across bibliometric data sources.
-The workflow shows how data from different sources (Crossref, OpenAlex, JCT, Web of Science, and Scopus) is processed and matched using ROR identifiers to enable comparative analysis.
-TA = Transformative Agreement}
+The workflow shows how data from different sources (hoaddata, derived from Crossref, OpenAlex, Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT), the Web of Science, and Scopus) is processed and matched using ROR IDs to enable comparative analysis.
+TA = Transformative Agreement.}
 \label{fig:workflow}
@@
 
-hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package collecting information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
-It combines article-level metadata from Crossref and OpenAlex with transformative agreement information from the cOAlition S Journal Checker Tool (JCT), which links journal and institutional data to agreements in the ESAC registry.
-
-As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of various safeguarded JCT snapshots.
+hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
+It combines article-level metadata from Crossref [@Hendricks_2020] and OpenAlex [@priem2022openalex] with transformative agreement information from Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data to agreements in the ESAC registry.
+
+As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of various safeguarded JCT snapshots [^jct_data].
 Journal data were enriched with ISSN variants linked to an ISSN-L and institution data with associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
@@
 Only about two-thirds of the articles examined provided corresponding author affiliations.
-For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(hoad) * 100`%.
+For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(hoad) * 100`%.
 Therefore, only first author data for hoaddata were considered in the following analysis.
@@
 In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` overlapping articles, represented by the blue bar.
-They represented `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(wos) * 100)`% of original articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Core articles with DOI") |> pull(scp) * 100)`% in Scopus.
+They represented `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(wos) * 100)`% of original articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(scp) * 100)`% in Scopus.
 Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to both Scopus and hoaddata.
@@
 
-<!-- [juan study] larger proportion of hybrid, suggest journals included which might not be covered by transformative agreement. -->
+Investigating over 13,000 hybrid journals included in the cOAlition Journal Checker Tool shows a substantial increase in  open access enabled by transformative agreements between 2019 and 2023, although most articles in these journals remained paywalled during this five-year period.
+While transformative agreements accounted for the majority of open access, a considerable number of open access articles are likely to have resulted from the payment of individual publication fees.
+The results confirm that transformative agreements and hybrid open access continue to be concentrated among a few large commercial publishers.
+They also highlight that high-income European countries in particular are using transformative agreements to make most of their article output open access, although South Africa showed similar levels of open access adoption.
+These trends were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science.
+
+Comparing the data sources examined, the study results show strong correlations by country affiliation, despite differences in journal and article coverage, and in the availability of metadata on corresponding authorship and open access status.
+However, several notable discrepancies warrant discussion.
+They stem not only from known limitations of the open scholarly metadata sources Crossref and OpenAlex, but also from the proprietary databases Scopus and Web of Science.
+
+The coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage.
+Differences can be found for journals targeting practitioners or local non-English language communties, with many such titles indexed exclusively in Crossref and Scopus.
+Using the open index Crossref demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
+
+The situation of hybrid journal publishing is therefore different from that of fully open access journals.
+Comparing the coverage of OpenAlex, Scopus and Web of Science, @Simard_2024 indicate that only half of the fully open access journals listed in the DOAJ are also indexed in Scopus and Web of Science.
+Notably, journals that charge no publication fees ("diamond journals") are absent from the selective Web of Science, which reinforces existing disparities in the indexing of underrepresented research communities and regions in selective bibliometric databases (ref).
+
+A notable advantage of using Scopus and Web of Science to study the impact of transformative agreements is their comprehensive differentiation of journal articles by document type, as many agreements restrict funding to eligible article types.
+This resulted in slightly higher rates of open access uptake compared to hoaddata.
+
+A similar limitation of studying open access with open scholarly data sources is the often reported lack of corresponding authorship information. 
+However, indicators using first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, which is not surprising given disciplinary norms in scholarly publishing.
+First authors are often considered to have carried out the main research underlying a paper, while the corresponding author is often recognised as having supervised a paper. 
+It is therefore not surprising that both are strongly correlated, suggesting that first and corresponding authors share the same country affiliation.
+Moreover, in most cases the first author is indeed the same as the corresponding author.
+As such, first authorship can be used confidentially in large-scale studies and those focusing on national consortia to assess the impact of transformative agreements.
+
+However, affiliation data is sparser in OpenAlex than in Scopus and Web of Science, resulting in slightly more articles that could be attributed to transformative agreements. 
+However, in order to establish a link, a thorough mapping of the database affiliation identifier to the ROR ID, the JCT's identifier for institutional participation in transformative agreements, is required.
+In contrast, ROR IDs are the canonical organisation identifier in OpenAlex, allowing for less ambiguous matching between agreement and publication.
+
+Comparison of scholarly databases is also valuable in identifying differences in open access identification. 
+Not all publishers deposit a Creative Commons licence with Crossref, the citation authority.
+A collaboration with Scopus and Web of Science helped to identify such publishers, as their underlying unpaywall data also parses licence information from websites.
+However, there were significant differences between Scopus and Web of Science in terms of open access indexing, particularly with regard to the delayed open access provided by Elsevier.
+In the case of Wiley, however, there were also problems with the analysis of the Unpaywall.
+This suggests that hybrid open access studies should not only refer to a single source of evidence.
+Instead, open access tags can be validated by the combiantion of seevral data sources, as in our case where metadata from publishers was compared with data from open access discovery services such as Unpaywall.
+It is also particularly useful to include a curated list of journals under the hybrid model to avoid misclassifications based on a journal's business model [@Visser_2021].
+
+<!-- Limitations: agreeemnts are complex, game auhtorship, invoicing not available, mapping, not feasible, requires more insitutional and publsihjer case studies involving financial data, but also improved and streamlined metadata compliance according to the evlolving stadatrd and iniatives (ESAC, Varcelona) in negotiation with publishers.
+
+Overall, -->
+
+<!--Journal indexing is also influenced by the open access evidence source used. 
+Using the same list of hybrid journals, the results show consoderable difference in indexing, because of different approach to identify open access.
+(coverage)
+
+(document types)
+( correpsonding auhtor)
+(open access)
+(pids)
+
+<!-- Hybird journals are very well indexed in both proprietory and open sources.
+confirms the  large concentrartion of publieher there is a much a less advantage of using openalex in terms of includindg missing ones as in comparision to full opena ccess jorunals with or without APCs (also called diamond oa).
+Previouly reported hybrid obsvered surplus is likely due to miscoding of full open access jorunals, as they cannot been found in list of transofrmative jorunals.
+However, uuing open schoalrly sources ournals with local audience  can be helpful to determine newly launched hybrid jorunals, as well as journals with local audience which is surprisingly given that transformative agreemnt are intented to shift the business modell of extisting ubscripton.based journal.
+
+Combining different open access evidence sources
+
+Authorship
+
+confirm metadata gaps, in particular document types.
+hoewever, it is often unclear, which are incldued in transformative agreements.
+but metadata gaps ar enot just to be found in openalex, but also in .
+Lack oif ror ids to incldue this.
+only a workaroudn made it possible to establish a link.
+a manual curation is not feaible given the amoutn of transformative agreements possible.
+
+the observed increae provide oppoirtunites for improved meatdata.
+negotiator have leverage, funder start using  open arrpaoches, standardisatuion activities in .
+These demosntrate the added value of transformative agrrements.
+
+The increasing trend with is consistent with previous research and case studie that describe an general increase with the introduction of transformative agreements. For example, Sweden reproted
+
+In terms of metadata coverage and quality, it shows the value of combining
+
+that,a t least at the country level, resuts between first and correpsodnign auhtor are storngyl correlated. 
+In line with empricial research, showing that morst first auhtros were also correposnding auhtor. 
+This can be explained by norm of auhtorship where fir
+
+The study deomnstrates the impact of transformative agreements on hybrid open access being consistent across biliometric data 
+ [juan study] larger proportion of hybrid, suggest journals included which might not be covered by transformative agreement, oa journals not tagged as fully oa, 741268 withouht any publisher. -->
 
@@
 [^subugoe]: <https://subugoe.github.io/scholcomm_analytics/data.html>
+
+[^tadump]: <https://journalcheckertool.org/transformative-agreements/>
+
+[^jct_data]: <https://github.com/njahn82/jct_data>

### 2025-03-27 analysis/manuscript.tex
Frit draft of discussion (limitation, furutre studies, conclusion lacking)

@@
 hoaddata, developed and maintained by the author to support open access
-monitoring and research (Jahn, 2025), is an R data package collecting
-information on hybrid open access uptake from multiple openly available
-data sources (Jahn, 2024). It combines article-level metadata from
-Crossref and OpenAlex with transformative agreement information from the
-cOAlition S Journal Checker Tool (JCT), which links journal and
-institutional data to agreements in the ESAC registry.
+monitoring and research (Jahn, 2025), is an R data package that
+regularly collects information on hybrid open access uptake from
+multiple openly available data sources (Jahn, 2024). It combines
+article-level metadata from Crossref (Hendricks et al., 2020) and
+OpenAlex (Priem et al., 2022) with transformative agreement information
+from Transformative Agreement Data dump used by the cOAlition S Journal
+Checker Tool (JCT)\footnote{\url{https://journalcheckertool.org/transformative-agreements/}},
+which links journal and institutional data to agreements in the ESAC
+registry.
 
 As discussed elsewhere (Jahn, 2025), hoaddata uses a unified dataset of
-various safeguarded JCT snapshots. Journal data were enriched with ISSN
-variants linked to an ISSN-L and institution data with associated
-ROR-IDs representing university hospitals or institutes of large
-research organisations, according to OpenAlex' institution entity. A
-comprehensive exclusion of fully open access journals using multiple
-journal lists (DOAJ, OpenAlex, Bielefeld) was performed to identify
-hybrid journals. hoaddata relies on Crossref for obtaining journal
-publication volume and open access status through Creative Commons
-licence information relative to the published version (``version of
-record''). Because of limited affiliation metadata in Crossref (Eck \&
-Waltman, 2022), hoaddata sources affiliation metadata from OpenAlex.
+various safeguarded JCT snapshots \footnote{\url{https://github.com/njahn82/jct_data}}.
+Journal data were enriched with ISSN variants linked to an ISSN-L and
+institution data with associated ROR-IDs representing university
+hospitals or institutes of large research organisations, according to
+OpenAlex' institution entity. A comprehensive exclusion of fully open
+access journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld)
+was performed to identify hybrid journals. hoaddata relies on Crossref
+for obtaining journal publication volume and open access status through
+Creative Commons licence information relative to the published version
+(``version of record''). Because of limited affiliation metadata in
+Crossref (Eck \& Waltman, 2022), hoaddata sources affiliation metadata
+from OpenAlex.
 
@@
 provided corresponding author affiliations. For first authors, the
-proportion was \%. Therefore, only first author data for hoaddata were
+proportion was 89\%. Therefore, only first author data for hoaddata were
 considered in the following analysis.
@@
 total, these journals recorded 6,289,687 overlapping articles,
-represented by the blue bar. They represented \% of original articles
-with DOI indexed in the Web of Science sample, and \% in Scopus. Another
-657,697 articles were exclusive to both Scopus and hoaddata. Exclusively
-in hoaddata were 177,110 articles, and exclusively in the proprietary
-databases were 325,194 articles.
+represented by the blue bar. They represented 94\% of original articles
+with DOI indexed in the Web of Science sample, and 86\% in Scopus.
+Another 657,697 articles were exclusive to both Scopus and hoaddata.
+Exclusively in hoaddata were 177,110 articles, and exclusively in the
+proprietary databases were 325,194 articles.
 
@@
 
+Investigating over 13,000 hybrid journals included in the cOAlition
+Journal Checker Tool shows a substantial increase in open access enabled
+by transformative agreements between 2019 and 2023, although most
+articles in these journals remained paywalled during this five-year
+period. While transformative agreements accounted for the majority of
+open access, a considerable number of open access articles are likely to
+have resulted from the payment of individual publication fees. The
+results confirm that transformative agreements and hybrid open access
+continue to be concentrated among a few large commercial publishers.
+They also highlight that high-income European countries in particular
+are using transformative agreements to make most of their article output
+open access, although South Africa showed similar levels of open access
+adoption. These trends were consistent across the investigated open data
+source hoaddata, derived from Crossref and OpenAlex, and the established
+proprietary bibliometric databases Scopus and Web of Science.
+
+Comparing the data sources examined, the study results show strong
+correlations by country affiliation, despite differences in journal and
+article coverage, and in the availability of metadata on corresponding
+authorship and open access status. However, several notable
+discrepancies warrant discussion. They stem not only from known
+limitations of the open scholarly metadata sources Crossref and
+OpenAlex, but also from the proprietary databases Scopus and Web of
+Science.
+
+The coverage analysis reveals that hybrid journals are well indexed in
+all three data sources, particularly in terms of article coverage.
+Differences can be found for journals targeting practitioners or local
+non-English language communties, with many such titles indexed
+exclusively in Crossref and Scopus. Using the open index Crossref
+demonstrated particular strength in identifying newly established hybrid
+journals, a notable finding given that transformative agreements
+primarily target existing subscription-based journals.
+
+The situation of hybrid journal publishing is therefore different from
+that of fully open access journals. Comparing the coverage of OpenAlex,
+Scopus and Web of Science, Simard et al. (2024) indicate that only half
+of the fully open access journals listed in the DOAJ are also indexed in
+Scopus and Web of Science. Notably, journals that charge no publication
+fees (``diamond journals'') are absent from the selective Web of
+Science, which reinforces existing disparities in the indexing of
+underrepresented research communities and regions in selective
+bibliometric databases (ref).
+
+A notable advantage of using Scopus and Web of Science to study the
+impact of transformative agreements is their comprehensive
+differentiation of journal articles by document type, as many agreements
+restrict funding to eligible article types. This resulted in slightly
+higher rates of open access uptake compared to hoaddata.
+
+A similar limitation of studying open access with open scholarly data
+sources is the often reported lack of corresponding authorship
+information. However, indicators using first authors, which have often
+been used as a proxy for determining open access funding, and
+corresponding authors show a high level of correlation, which is not
+surprising given disciplinary norms in scholarly publishing. First
+authors are often considered to have carried out the main research
+underlying a paper, while the corresponding author is often recognised
+as having supervised a paper. It is therefore not surprising that both
+are strongly correlated, suggesting that first and corresponding authors
+share the same country affiliation. Moreover, in most cases the first
+author is indeed the same as the corresponding author. As such, first
+authorship can be used confidentially in large-scale studies and those
+focusing on national consortia to assess the impact of transformative
+agreements.
+
+However, affiliation data is sparser in OpenAlex than in Scopus and Web
+of Science, resulting in slightly more articles that could be attributed
+to transformative agreements. However, in order to establish a link, a
+thorough mapping of the database affiliation identifier to the ROR ID,
+the JCT's identifier for institutional participation in transformative
+agreements, is required. In contrast, ROR IDs are the canonical
+organisation identifier in OpenAlex, allowing for less ambiguous
+matching between agreement and publication.
+
+Comparison of scholarly databases is also valuable in identifying
+differences in open access identification. Not all publishers deposit a
+Creative Commons licence with Crossref, the citation authority. A
+collaboration with Scopus and Web of Science helped to identify such
+publishers, as their underlying unpaywall data also parses licence
+information from websites. However, there were significant differences
+between Scopus and Web of Science in terms of open access indexing,
+particularly with regard to the delayed open access provided by
+Elsevier. In the case of Wiley, however, there were also problems with
+the analysis of the Unpaywall. This suggests that hybrid open access
+studies should not only refer to a single source of evidence. Instead,
+open access tags can be validated by the combiantion of seevral data
+sources, as in our case where metadata from publishers was compared with
+data from open access discovery services such as Unpaywall. It is also
+particularly useful to include a curated list of journals under the
+hybrid model to avoid misclassifications based on a journal's business
+model (Visser et al., 2021).
+
 \phantomsection\label{refs}
@@
 
+\bibitem[\citeproctext]{ref-Hendricks_2020}
+Hendricks, G., Tkaczyk, D., Lin, J., \& Feeney, P. (2020). Crossref: The
+sustainable source of community-owned scholarly metadata.
+\emph{Quantitative Science Studies}, \emph{1}(1), 414--427.
+\url{https://doi.org/10.1162/qss_a_00022}
+
 \bibitem[\citeproctext]{ref-hoaddata}
@@
 
+\bibitem[\citeproctext]{ref-priem2022openalex}
+Priem, J., Piwowar, H., \& Orr, R. (2022). \emph{OpenAlex: A fully-open
+index of scholarly works, authors, venues, institutions, and concepts}.
+\url{https://arxiv.org/abs/2205.01833}
+
 \bibitem[\citeproctext]{ref-Rabesandratana_2019}
@@
 
+\bibitem[\citeproctext]{ref-Simard_2024}
+Simard, M.-A., Basson, I., Hare, M., Lariviere, V., \& Mongeon, P.
+(2024). \emph{The open access coverage of OpenAlex, scopus and web of
+science}. \url{https://doi.org/10.48550/ARXIV.2404.01985}
+
 \bibitem[\citeproctext]{ref-Singh_2021}

### 2025-03-27 analysis/manuscript.qmd
Proofreading

@@
 hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
-It combines article-level metadata from Crossref [@Hendricks_2020] and OpenAlex [@priem2022openalex] with transformative agreement information from Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data to agreements in the ESAC registry.
-
-As discussed elsewhere [@Jahn_2025], hoaddata uses a unified dataset of various safeguarded JCT snapshots [^jct_data].
-Journal data were enriched with ISSN variants linked to an ISSN-L and institution data with associated ROR-IDs representing university hospitals or institutes of large research organisations, according to OpenAlex' institution entity.
-A comprehensive exclusion of fully open access journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld)  was performed to identify hybrid journals.
-hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
-Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources  affiliation metadata from OpenAlex.
+It combines article-level metadata from Crossref [@Hendricks_2020] and affilation metadata from OpenAlex [@priem2022openalex] with transformative agreement information from Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data about participating research organisations to agreements in the ESAC registry.
 
@@
 Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020].
+Web of Science is a selective, base research-focused database with dedicated panels to choose journals for inclusion [@Visser_2021; @Stahlschmidt_2022].
 The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).
 
-These collections provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall, the same provider as OpenAlex.
+These collections provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall [@Piwowar_2018], the same provider as OpenAlex.
 However, the Web of Science lacks information about journals and articles under transformative agreements.
@@
 Similar to the Web of Science, Scopus is selective with regard to the journals it indexes.
-However, its coverage is substantially more extensive than that of the Web of Science collections considered in this study [@Singh_2021; @Visser_2021].
+However, its journal coverage is much broader than that of the Web of Science collections considered in this study, as it also indexes a wider range of applied research journals [@Singh_2021; @Visser_2021; @Stahlschmidt_2022].
 With detailed metadata about article types, open access status information derived from Unpaywall, author roles, and disambiguated affiliations, Scopus also contains important data to assess open access uptake, although no direct information regarding transformative agreements was available at the time of the study.
@@
 
-Using an enriched and unified list of hybrid journals in transformative agreements according to the JCT, article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L.
+Following @Jahn_2025, the starting point was a unified dataset of several safeguarded JCT snapshots[^jct_data], as provided by hoaddata.
+The JCT journal data in hoaddata were enriched with ISSN variants linked to an ISSN-L. 
+To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists (DOAJ, OpenAlex, Bielefeld).
+These data were used to determine the publication volume of hybrid journals for each database independently.
+
+hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
+<!-- Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources  affiliation metadata from OpenAlex.-->
+<!-- Using an enriched and unified list of hybrid journals in transformative agreements according to the JCT, article metadata were retrieved from each database using all ISSN variants linked to an ISSN-L.
+-->
 <!-- This study examined the adoption of open access for hybrid journals included in transformative agreements, which published at least one journal article in the five-year period 2019-2023. 
@@
 -->
-
 <!-- The JCT journal data was enriched with the linking ISSN (ISSN-L) according to the ISSN Registry Agency.
@@
 Many transformative agreements typically cover only certain types of journal
-articles, in particular original research articles including reviews.
+articles, in particular original research articles including reviews [@Borrego_2020].
 Because of limited information on these document types in open scholarly data [@Haupka_2024], hoaddata used an extended version of Unpaywall's paratext recognition approach to exclude non-scholarly content [@Piwowar_2018].
 To exclude conference supplements, which are also often not covered by transformative agreements, only articles published in regular issues, indicated by numerical pagination, were considered.
-For Web of Science and Scopus, their established document type classifications were used to identify original research articles and reviews, referred to as original articles throughout this study.
+For Web of Science and Scopus, their established, mainly accurate document type classifications [@Maisano_2025; @Donner_2017] were used to identify original research articles and reviews, referred to as original articles throughout this study.
 
@@
 While hoaddata sourced this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall as evidence source. 
-Unpaywall supplements Crossref metadata by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018]. 
+Unpaywall also uses Crossref license metadata, but supplements them by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018].
 This additional parsing remains necessary despite transformative agreement workflows recommend the depositing license information during DOI registration [@Geschuhn_2017].
-Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses.
+Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms according to their documentations[^wosoa][^scopusoa], distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses.
 
@@
 ```
+
 Author affiliations were retrieved for both first and, if available, corresponding authors to prepare the linking between articles and institutions covered by transformative agreements.
-To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifiers for Scopus.
+To improve the data retrieval, JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
+To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifier.
 Additionally, ISO country codes were retrieved for each author's address to compile country-level statistics. 
@@
 
-Based on these matching tables, articles eligible under transformative agreements could also be retrieved from Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT. 
-The estimation of eligible articles followed @Jahn_2025 and included a matching of both journals and participating institutions according to the Transformative Agreement Data dump for each of the data sources examined.
+Based on these generated matching tables, articles eligible under transformative agreements could also be retrieved from Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT.
+The estimation of eligible articles followed @Jahn_2025 and included a matching of both journals and participating institutions.
 The matching also took into account the duration of agreements according to the ESAC registry, with only those matches where an agreement was actually in place being considered for subsequent analysis.
+A related study [@de_Jonge_2025], applied to publications funded by the Dutch Research Council (NWO) and validated against internal invoicing data, confirmed that such matching can accurately identify most articles under transformative agreements.
 
@@
   escape = TRUE, 
-  position = "H",
-  col.names = c("", "HOAD", "Web of Science", "Scopus")
+  position = "ht",
+  col.names = c("", "hoaddata*", "Web of Science", "Scopus")
 ) |>
@@
   pack_rows("Open Access (OA) metrics", 8, 9) |>
-  pack_rows("Original articles with affiliation data", 10, 11) 
+  pack_rows("Original articles with affiliation data", 10, 11) |>
+  footnote(symbol = c("Journal article metadata from Crossref, except affiliations from OpenAlex"))
 ```
@@
 Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
-A closer look at original articles with affiliation data reveals a lack of data on corresponding authors in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
-Only about two-thirds of the articles examined provided corresponding author affiliations.
+A closer look at original articles with affiliation data, and in line with related research [@Zhang_2024], reveals a lack of affiliation data in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
+In particular, only about two-thirds of the articles examined provided corresponding author affiliations.
 For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(hoad) * 100`%.
+At the time of writing, OpenAlex disclosed limited coverage of corresponding authorhip data [^oacorresponding].
 Therefore, only first author data for hoaddata were considered in the following analysis.
 
+## Data analysis
+
 # Results
 
-First, I will present a coverage analysis. 
-Then, I investigate hybrid open access uptake using the same methods for each database and the compare results against each other in order to assess the suitablility of the biblioemtric databases to examine transofrmative agreements.
+This section first presents the indexing coverage of hybrid open access by comparing the data from open data sources with the proprietary bibliometric databases Scopus and Web of Science. 
+Then, using the same methods, indicators at the publisher and country level are calculated independently for each database and compared with each other to assess the suitability of the bibliometric databases for investigating transformative agreements.
 
@@
 
-upset_overview <- ComplexUpset::upset(
-  upset_data,
-  c("Crossref", "WoS", "Scopus"),
+upset_overview <- upset_data |>
+  rename(hoaddata = Crossref) |>
+  ComplexUpset::upset(
+  c("hoaddata", "WoS", "Scopus"),
   # Only journals in Crossref
@@
 The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
-Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaadata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus.
+Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaddata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus.
 Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in the Web of Science.
@@
 In particular, the journals covered exclusively by hoaddata were substantially smaller than those covered by all three sources. 
-Upon inspection, these were often newly launched journals, which explains the relatively low five-year publication volume.
+Upon inspection, these were often newly launched hybrid journals, which explains the relatively low five-year publication volume.
 An example is *Digital society* that published `r jn_upset_articles |> filter(issn_l == "2731-4650") |> pull(n_cr)` articles.
@@
 Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)`, although recording the largest number of hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%).
-Wiley accounted for `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
+Wiley accounted for `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
 The remaining `r jn_upset_publisher |> filter(publisher == "Other") |> distinct(esac_publisher) |> nrow()` publishers collectively accounted for `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`%).
@@
 
-upset_publisher_plot <- ComplexUpset::upset(
-  jn_upset_publisher,
-  c("Crossref", "WoS", "Scopus"),
+upset_publisher_plot <- jn_upset_publisher |>
+  rename(hoaddata = Crossref) |>
+  ComplexUpset::upset(
+    c("hoaddata", "WoS", "Scopus"),
   # Only journals in Crossref
@@
 
-A comparative analysis of publisher portfolios across different indexing sets demonstrates that publisher were not represented uniformly. 
+Comparing publisher portfolios across different indexing sets demonstrates that publisher were not represented uniformly. 
 Notably, Springer Nature exhibited `r jn_upset_publisher |> filter(esac_publisher == "Springer Nature", WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow()` hybrid journals exclusively indexed in both hoaddata and Scopus.
@@
 Additionally, this subset included titles with a more broader disciplinary focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
-The set also captured ceased journals, providing further insights into the dynamic nature of journal portfolios and business strategies.
+The set also captured ceased journals, providing further insights into the dynamics of journal publishing.
 
@@
 Following @Jahn_2025, indicators were calculated for each data source and comprise the number and proportion of open access articles, including those enabled by transformative agreements, from 2019 to 2023.
-For Web of Science and Scopus, the impact of transformative agreements was estimated using both first and corresponding authorships, while hoaddata indicator calculation was limited to first-author affiliations.
+For Web of Science and Scopus, the impact of transformative agreements was estimated using both first and corresponding authorships, while hoaddata indicator calculation was limited to first author affiliations.
 
@@
     ta_oa_corresponding_author_articles = sum(ta_oa_corresponding_author_articles)
-  )
+  ) |>
+  mutate(src = ifelse(src == "HOAD", "hoaddata", src))
 
@@
   scale_fill_manual(NULL,
-    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+    values = c(`hoaddata` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
     guide = "none"
@@
   scale_color_manual(NULL,
-   values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")
+   values = c(`hoaddata` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")
   ) +
@@
   geom_point(aes(y = value, fill = src), alpha = 0.8, shape = 21) +
-  scale_color_manual(NULL, values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
+  scale_color_manual(NULL, values = c(`hoaddata` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
   scale_fill_manual(NULL,
-                    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+                    values = c(`hoaddata` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
                     guide = "none") +
@@
   geom_point(aes(y = value / n_oa_articles, fill = src), alpha = 0.8, shape = 21) +
-  scale_color_manual(NULL, values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
+  scale_color_manual(NULL, values = c(`hoaddata` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9")) +
   scale_fill_manual(NULL,
-                    values = c(`HOAD` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
+                    values = c(`hoaddata` = "#000000", `Scopus` = "#E69F00", `Web of Science` = "#56B4E9"),
                     guide = "none") +  
@@
 Similarly, Scopus recorded an growth from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
-However, Web of Science recorded higher open access uptake in early years, before converging in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023,suggesting an different approach towards labeling hybrid open access by the Web of Science.
-
-Simialry, hybrid open access by transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
+However, Web of Science recorded higher open access uptake in early years, before converging in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023, suggesting an different approach towards labeling hybrid open access by the Web of Science.
+
+Similarly, hybrid open access by transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
 Trends were consistent for first (dashed line) and corresponding author (dotted line) affiliations. 
@@
 Web of Science recorded `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_first_author_articles) |> sum()` open access articles via transformative agreements using first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`)  and  `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
-hoaddata, lacking corresponding author data, linked `r uptake_by_year_short |> filter(src == "HOAD") |> pull(ta_oa_first_author_articles) |> sum()` articles to transformative agreements but showed slower growth than Scopus and Web of Science, from `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`.
+hoaddata, lacking corresponding author data, linked `r uptake_by_year_short |> filter(src == "HOAD") |> pull(ta_oa_first_author_articles) |> sum()` articles to transformative agreements but showed slower growth than Scopus and Web of Science, from `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` in 2019 to `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()` in 2023.
 
@@
 For corresponding authors, the shares were slightly larger, with Scopus recording  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% and Web of Science `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% in 2023.
-However, substantial hybrid open access was still facilitated outside transformative agreements, likely through APCs payed from discretionary research funds (ref).
+However, substantial hybrid open access was still facilitated outside transformative agreements, likely through APCs paid from discretionary research funds [@Jahn_2021; @Suber_2012].
 
@@
 In contrast, Elsevier's share remained relatively constant, increasing only slightly from `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`% according to the Web of Science.
-Upon inspection, this discrepancy primarily stemmed from articles provided under Elsevier's user license, used for the publisher's open archive. 
-These articles, made freely available after an embargo period, were tagged as hybrid open access in Web of Science, even though its documentation specified that only articles under a CC license variant were considered.
+Upon inspection, this discrepancy primarily stemmed from articles in the publisher's open archive. 
+These articles, made freely available after an embargo period under Elsevier's user license, were tagged as hybrid open access in Web of Science, even though its documentation[^wosoa] specified that only articles under a CC license variant were considered.
 Previous research [@Jahn_2021; @Haustein_APC_2024] has shown that Elsevier provided a substantial portion of its articles under this license, explaining the relatively large and stable share of open access over the years.
@@
 Despite these differences in open access evidence, the three data sources show consistent temporal trends in hybrid open access enabled by transformative agreements (see Figure \ref{fig-uptake_publisher}B and C).
-Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across the examined data sources, followed by Springer Nature and Elsevier.
-By 2023, the share of open access enabled by transformative agreements appeared to stabilise for all three publishers, with Wiley recording the largest share (see Figure \ref{fig-uptake_publisher}B). 
-Elsevier, however, showed a comparable late uptake, consistent with the publisher's historical reluctance to engage in negotations with library consortia. 
+Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across the examined data sources, followed by Springer Nature and Elsevier, recording a comparable late uptake, which is consistent with the publisher's historical reluctance to engage in negotations with library consortia.
+However, by 2023, the share of open access enabled by transformative agreements appeared to stabilise for all three publishers (see Figure \ref{fig-uptake_publisher}C). 
 Interestingly, the differences between first and corresponding author affiliations were more pronounced at the publisher level. 
@@
 However, France was better represented in Web of Science, likely due to its agreement with Elsevier starting in 2019, which allowed delayed open access under the publisher’s user license [@Rabesandratana_2019].
-This license was not classified as hybrid open access in Scopus or hoaddata.
+This licence was not classified as hybrid open access in either Scopus or hoaddata.
 In all cases, Spearman rank correlations were $\rho > .9$, $p < 0.001$, showing a high level of correlations between the databases and authorship roles considered.
 
-Transformative agreements appeared to be a key driver of open access growth (see Figure \ref{fig-uptake_country}C and D).
+Transformative agreements appeared to be a key driver of national open access growth (see Figure \ref{fig-uptake_country}C and D).
 OECD members accounted for the majority of open access articles enabled by transformative agreements.
 As a notable exception, South Africa also featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward.[^sa]
-Results were consistent across data sources, with a strong correlation observed. 
-However, Wiley surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with Wiley that started in July 2019.
+Results were consistent across data sources. 
+However, Wiley's open access surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with Wiley that started in July 2019.
 
@@
 
+# References
+
 <!-- Limitations: agreeemnts are complex, game auhtorship, invoicing not available, mapping, not feasible, requires more insitutional and publsihjer case studies involving financial data, but also improved and streamlined metadata compliance according to the evlolving stadatrd and iniatives (ESAC, Varcelona) in negotiation with publishers.
@@
 [^jct_data]: <https://github.com/njahn82/jct_data>
+
+[^wosoa]: <https://webofscience.help.clarivate.com/en-us/Content/open-access.html>
+
+[^scopusoa]: <https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option>
+
+[^oacorresponding]: <https://docs.openalex.org/api-entities/works/work-object/authorship-object#is_corresponding>

### 2025-03-27 analysis/manuscript.tex
Proofreading

@@
 article-level metadata from Crossref (Hendricks et al., 2020) and
-OpenAlex (Priem et al., 2022) with transformative agreement information
-from Transformative Agreement Data dump used by the cOAlition S Journal
-Checker Tool (JCT)\footnote{\url{https://journalcheckertool.org/transformative-agreements/}},
-which links journal and institutional data to agreements in the ESAC
-registry.
-
-As discussed elsewhere (Jahn, 2025), hoaddata uses a unified dataset of
-various safeguarded JCT snapshots \footnote{\url{https://github.com/njahn82/jct_data}}.
-Journal data were enriched with ISSN variants linked to an ISSN-L and
-institution data with associated ROR-IDs representing university
-hospitals or institutes of large research organisations, according to
-OpenAlex' institution entity. A comprehensive exclusion of fully open
-access journals using multiple journal lists (DOAJ, OpenAlex, Bielefeld)
-was performed to identify hybrid journals. hoaddata relies on Crossref
-for obtaining journal publication volume and open access status through
-Creative Commons licence information relative to the published version
-(``version of record''). Because of limited affiliation metadata in
-Crossref (Eck \& Waltman, 2022), hoaddata sources affiliation metadata
-from OpenAlex.
+affilation metadata from OpenAlex (Priem et al., 2022) with
+transformative agreement information from Transformative Agreement Data
+dump used by the cOAlition S Journal Checker Tool (JCT)\footnote{\url{https://journalcheckertool.org/transformative-agreements/}},
+which links journal and institutional data about participating research
+organisations to agreements in the ESAC registry.
 
@@
 proprietary bibliometric database consisting of several collections
-(Birkle et al., 2020). The collections considered in this study were the
-Science Citation Index Expanded (SCIE), the Social Sciences Citation
-Index (SSCI) and the Arts \& Humanities Citation Index (AHCI).
+(Birkle et al., 2020). Web of Science is a selective, base
+research-focused database with dedicated panels to choose journals for
+inclusion (Stahlschmidt \& Stephen, 2022; Visser et al., 2021). The
+collections considered in this study were the Science Citation Index
+Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts
+\& Humanities Citation Index (AHCI).
 
@@
 contributions, such as original articles or reviews, and open access
-status information derived from OurResearch's Unpaywall, the same
-provider as OpenAlex. However, the Web of Science lacks information
-about journals and articles under transformative agreements.
+status information derived from OurResearch's Unpaywall (Piwowar et al.,
+2018), the same provider as OpenAlex. However, the Web of Science lacks
+information about journals and articles under transformative agreements.
 
@@
 Similar to the Web of Science, Scopus is selective with regard to the
-journals it indexes. However, its coverage is substantially more
-extensive than that of the Web of Science collections considered in this
-study (Singh et al., 2021; Visser et al., 2021). With detailed metadata
-about article types, open access status information derived from
-Unpaywall, author roles, and disambiguated affiliations, Scopus also
-contains important data to assess open access uptake, although no direct
-information regarding transformative agreements was available at the
-time of the study.
+journals it indexes. However, its journal coverage is much broader than
+that of the Web of Science collections considered in this study, as it
+also indexes a wider range of applied research journals (Singh et al.,
+2021; Stahlschmidt \& Stephen, 2022; Visser et al., 2021). With detailed
+metadata about article types, open access status information derived
+from Unpaywall, author roles, and disambiguated affiliations, Scopus
+also contains important data to assess open access uptake, although no
+direct information regarding transformative agreements was available at
+the time of the study.
 
@@
 
-Using an enriched and unified list of hybrid journals in transformative
-agreements according to the JCT, article metadata were retrieved from
-each database using all ISSN variants linked to an ISSN-L.
-
-The article metadata included DOIs, publication dates, open access
-information as well as author roles and institutional affiliations.
-Publication years were determined using the earliest known date of
-publication in a journal. In hoaddata, this corresponded to Crossref's
-issued date. For Web of Science and Scopus, the earliest publication
-date was used where available, with Scopus dates specifically determined
-by the KB through version tracking of the raw data.
+Following Jahn (2025), the starting point was a unified dataset of
+several safeguarded JCT snapshots\footnote{\url{https://github.com/njahn82/jct_data}},
+as provided by hoaddata. The JCT journal data in hoaddata were enriched
+with ISSN variants linked to an ISSN-L. To identify hybrid journals, a
+comprehensive exclusion of fully open access journals was performed
+using multiple journal lists (DOAJ, OpenAlex, Bielefeld). These data
+were used to determine the publication volume of hybrid journals for
+each database independently.
+
+hoaddata relies on Crossref for obtaining journal publication volume and
+open access status through Creative Commons licence information relative
+to the published version (``version of record''). The article metadata
+included DOIs, publication dates, open access information as well as
+author roles and institutional affiliations. Publication years were
+determined using the earliest known date of publication in a journal. In
+hoaddata, this corresponded to Crossref's issued date. For Web of
+Science and Scopus, the earliest publication date was used where
+available, with Scopus dates specifically determined by the KB through
+version tracking of the raw data.
 
@@
 journal articles, in particular original research articles including
-reviews. Because of limited information on these document types in open
-scholarly data (Haupka et al., 2024), hoaddata used an extended version
-of Unpaywall's paratext recognition approach to exclude non-scholarly
-content (Piwowar et al., 2018). To exclude conference supplements, which
-are also often not covered by transformative agreements, only articles
-published in regular issues, indicated by numerical pagination, were
-considered. For Web of Science and Scopus, their established document
-type classifications were used to identify original research articles
-and reviews, referred to as original articles throughout this study.
+reviews (Borrego et al., 2021). Because of limited information on these
+document types in open scholarly data (Haupka et al., 2024), hoaddata
+used an extended version of Unpaywall's paratext recognition approach to
+exclude non-scholarly content (Piwowar et al., 2018). To exclude
+conference supplements, which are also often not covered by
+transformative agreements, only articles published in regular issues,
+indicated by numerical pagination, were considered. For Web of Science
+and Scopus, their established, mainly accurate document type
+classifications (Donner, 2017; Maisano et al., 2025) were used to
+identify original research articles and reviews, referred to as original
+articles throughout this study.
 
@@
 metadata, Web of Science and Scopus relied on Unpaywall as evidence
-source. Unpaywall supplements Crossref metadata by parsing publisher
-websites directly, addressing cases where publishers do not provide
-machine-readable Creative Commons license information (Piwowar et al.,
-2018). This additional parsing remains necessary despite transformative
-agreement workflows recommend the depositing license information during
-DOI registration (Geschuhn \& Stone, 2017). Both Web of Science and
-Scopus defined hybrid open access consistently as content available
-under Creative Commons licenses on publisher platforms, distinguishing
-it from bronze open access that lack such explicit license information,
-or use publisher-specific licenses.
+source. Unpaywall also uses Crossref license metadata, but supplements
+them by parsing publisher websites directly, addressing cases where
+publishers do not provide machine-readable Creative Commons license
+information (Piwowar et al., 2018). This additional parsing remains
+necessary despite transformative agreement workflows recommend the
+depositing license information during DOI registration (Geschuhn \&
+Stone, 2017). Both Web of Science and Scopus defined hybrid open access
+consistently as content available under Creative Commons licenses on
+publisher platforms according to their
+documentations\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}\footnote{\url{https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option}},
+distinguishing it from bronze open access that lack such explicit
+license information, or use publisher-specific licenses.
 
@@
 corresponding authors to prepare the linking between articles and
-institutions covered by transformative agreements. To handle different
-address variants, database-specific affiliation identifiers were used:
-ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web
-of Science, and Scopus Affiliation Identifiers for Scopus. Additionally,
-ISO country codes were retrieved for each author's address to compile
-country-level statistics. These country codes for Web of Science and
-Scopus were provided by the KB.
+institutions covered by transformative agreements. To improve the data
+retrieval, JCT institution data was enriched with ROR-IDs from
+associated institutions, such as university hospitals or institutes of
+large research organisations such as the Max Planck Society, according
+to OpenAlex' institution entity. To handle different address variants,
+database-specific affiliation identifiers were used: ROR-IDs from
+OpenAlex for hoaddata, affiliation enhanced names for Web of Science,
+and Scopus Affiliation Identifier. Additionally, ISO country codes were
+retrieved for each author's address to compile country-level statistics.
+These country codes for Web of Science and Scopus were provided by the
+KB.
 
@@
 
-Based on these matching tables, articles eligible under transformative
-agreements could also be retrieved from Web of Science and Scopus,
-although they did not contain the ROR IDs used by the JCT. The
-estimation of eligible articles followed Jahn (2025) and included a
-matching of both journals and participating institutions according to
-the Transformative Agreement Data dump for each of the data sources
-examined. The matching also took into account the duration of agreements
-according to the ESAC registry, with only those matches where an
-agreement was actually in place being considered for subsequent
-analysis.
+Based on these generated matching tables, articles eligible under
+transformative agreements could also be retrieved from Web of Science
+and Scopus, although they did not contain the ROR IDs used by the JCT.
+The estimation of eligible articles followed Jahn (2025) and included a
+matching of both journals and participating institutions. The matching
+also took into account the duration of agreements according to the ESAC
+registry, with only those matches where an agreement was actually in
+place being considered for subsequent analysis. A related study (Jonge
+et al., 2025), applied to publications funded by the Dutch Research
+Council (NWO) and validated against internal invoicing data, confirmed
+that such matching can accurately identify most articles under
+transformative agreements.
 
@@
 \toprule
-\textbf{} & \textbf{HOAD} & \textbf{Web of Science} & \textbf{Scopus}\\
+\textbf{} & \textbf{hoaddata*} & \textbf{Web of Science} & \textbf{Scopus}\\
 \midrule
@@
 \bottomrule
+\multicolumn{4}{l}{\rule{0pt}{1em}\textsuperscript{*} Journal article metadata from Crossref, except affiliations from OpenAlex}\\
 \end{tabular}
@@
 articles with a DOI. A closer look at original articles with affiliation
-data reveals a lack of data on corresponding authors in the case of
-OpenAlex, the affiliation data source used by hoaddata, compared to Web
-of Science and Scopus. Only about two-thirds of the articles examined
-provided corresponding author affiliations. For first authors, the
-proportion was 89\%. Therefore, only first author data for hoaddata were
-considered in the following analysis.
+data, and in line with related research (Zhang et al., 2024), reveals a
+lack of affiliation data in the case of OpenAlex, the affiliation data
+source used by hoaddata, compared to Web of Science and Scopus. In
+particular, only about two-thirds of the articles examined provided
+corresponding author affiliations. For first authors, the proportion was
+89\%. At the time of writing, OpenAlex disclosed limited coverage of
+corresponding authorhip data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
+Therefore, only first author data for hoaddata were considered in the
+following analysis.
+
+\subsection{Data analysis}\label{data-analysis}
 
@@
 
-First, I will present a coverage analysis. Then, I investigate hybrid
-open access uptake using the same methods for each database and the
-compare results against each other in order to assess the suitablility
-of the biblioemtric databases to examine transofrmative agreements.
+This section first presents the indexing coverage of hybrid open access
+by comparing the data from open data sources with the proprietary
+bibliometric databases Scopus and Web of Science. Then, using the same
+methods, indicators at the publisher and country level are calculated
+independently for each database and compared with each other to assess
+the suitability of the bibliometric databases for investigating
+transformative agreements.
 
@@
 comprising 21\% (n = 2,595) of hybrid journals. Notably, 6\% (n = 739)
-of journals were exclusively contained in hoaadata, while another 6\% (n
+of journals were exclusively contained in hoaddata, while another 6\% (n
 = 748) were only found in the proprietory database Scopus. Of these, 354
@@
 were substantially smaller than those covered by all three sources. Upon
-inspection, these were often newly launched journals, which explains the
-relatively low five-year publication volume. An example is \emph{Digital
-society} that published 86 articles. This hybrid journal was launched in
-2022, being covered by various Springer Nature transformative agreements
-since then.
+inspection, these were often newly launched hybrid journals, which
+explains the relatively low five-year publication volume. An example is
+\emph{Digital society} that published 86 articles. This hybrid journal
+was launched in 2022, being covered by various Springer Nature
+transformative agreements since then.
 
@@
 2,311, although recording the largest number of hybrid journals (19\%).
-Wiley accounted for 858,939 articles (12\%) in1,382 hybrid journals
+Wiley accounted for 858,939 articles (12\%) in 1,382 hybrid journals
 (11\%). The remaining 54 publishers collectively accounted for 2,749,847
@@
 
-A comparative analysis of publisher portfolios across different indexing
-sets demonstrates that publisher were not represented uniformly.
-Notably, Springer Nature exhibited 519 hybrid journals exclusively
-indexed in both hoaddata and Scopus. This set included journals from the
-Chinese Academy of Science, German-language medical journals, and
-Eastern European publications including the \emph{Journal of
-Mathematical Sciences}, which also publishes English-language
-translations of Russian-language works. Additionally, this subset
-included titles with a more broader disciplinary focus such as \emph{SN
-Computer Science} and newly launched hybrid journals like \emph{Nature
-Computational Science}, which started in 2021 and was indexed in Scopus
-but not yet in Web of Science. The set also captured ceased journals,
-providing further insights into the dynamic nature of journal portfolios
-and business strategies.
+Comparing publisher portfolios across different indexing sets
+demonstrates that publisher were not represented uniformly. Notably,
+Springer Nature exhibited 519 hybrid journals exclusively indexed in
+both hoaddata and Scopus. This set included journals from the Chinese
+Academy of Science, German-language medical journals, and Eastern
+European publications including the \emph{Journal of Mathematical
+Sciences}, which also publishes English-language translations of
+Russian-language works. Additionally, this subset included titles with a
+more broader disciplinary focus such as \emph{SN Computer Science} and
+newly launched hybrid journals like \emph{Nature Computational Science},
+which started in 2021 and was indexed in Scopus but not yet in Web of
+Science. The set also captured ceased journals, providing further
+insights into the dynamics of journal publishing.
 
@@
 corresponding authorships, while hoaddata indicator calculation was
-limited to first-author affiliations.
+limited to first author affiliations.
 
@@
 Scopus (orange line), and Web of Science (blue line). According to
-hoaddata, hybrid open access uptake increased from 6.4\% (n = 85,071) in
-2019 to 19\% (n = 302,358) in 2023. Similarly, Scopus recorded an growth
-from 6.5\% (n = 84,648) in 2019 to 19\% (n = 322,850) in 2023. However,
-Web of Science recorded higher open access uptake in early years, before
-converging in 2023, from 11\% (n = 137,202) in 2019 to 20\% (n =
-255,481) in 2023,suggesting an different approach towards labeling
-hybrid open access by the Web of Science.
-
-Simialry, hybrid open access by transformative agreements increased
+hoaddata, hybrid open access uptake increased from \% (n = ) in 2019 to
+\% (n = ) in 2023. Similarly, Scopus recorded an growth from 6.5\% (n =
+84,648) in 2019 to 19\% (n = 322,850) in 2023. However, Web of Science
+recorded higher open access uptake in early years, before converging in
+2023, from 11\% (n = 137,202) in 2019 to 20\% (n = 255,481) in 2023,
+suggesting an different approach towards labeling hybrid open access by
+the Web of Science.
+
+Similarly, hybrid open access by transformative agreements increased
 between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C). Trends
@@
 500,076 using corresponding author metadata (from 22,092 to 177,030).
-hoaddata, lacking corresponding author data, linked 501,649 articles to
+hoaddata, lacking corresponding author data, linked 0 articles to
 transformative agreements but showed slower growth than Scopus and Web
-of Science, from 16,010 to 183,757.
+of Science, from 0 in 2019 to 0 in 2023.
 
@@
 agreements enabled the majority of hybrid open access. For first
-authors, the share ranged between 61\% (hoaddata), 64\% (Scopus), and 68
+authors, the share ranged between \% (hoaddata), 64\% (Scopus), and 68
 \% (Web of Science) in 2023. For corresponding authors, the shares were
@@
 2023. However, substantial hybrid open access was still facilitated
-outside transformative agreements, likely through APCs payed from
-discretionary research funds (ref).
+outside transformative agreements, likely through APCs paid from
+discretionary research funds (Jahn et al., 2022; Suber, 2012).
 
@@
 to 15\% according to the Web of Science. Upon inspection, this
-discrepancy primarily stemmed from articles provided under Elsevier's
-user license, used for the publisher's open archive. These articles,
-made freely available after an embargo period, were tagged as hybrid
-open access in Web of Science, even though its documentation specified
-that only articles under a CC license variant were considered. Previous
-research (Haustein et al., 2024; Jahn et al., 2022) has shown that
-Elsevier provided a substantial portion of its articles under this
+discrepancy primarily stemmed from articles in the publisher's open
+archive. These articles, made freely available after an embargo period
+under Elsevier's user license, were tagged as hybrid open access in Web
+of Science, even though its documentation\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
+specified that only articles under a CC license variant were considered.
+Previous research (Haustein et al., 2024; Jahn et al., 2022) has shown
+that Elsevier provided a substantial portion of its articles under this
 license, explaining the relatively large and stable share of open access
@@
 reported as open access in 2023 across the examined data sources,
-followed by Springer Nature and Elsevier. By 2023, the share of open
-access enabled by transformative agreements appeared to stabilise for
-all three publishers, with Wiley recording the largest share (see Figure
-\ref{fig-uptake_publisher}B). Elsevier, however, showed a comparable
-late uptake, consistent with the publisher's historical reluctance to
-engage in negotations with library consortia. Interestingly, the
-differences between first and corresponding author affiliations were
-more pronounced at the publisher level. In Scopus, for example, the
-share of open access via transformative agreements measured by
-corresponding authorship was greater for Springer Nature in 2023 than
-when using first authorship.
+followed by Springer Nature and Elsevier, recording a comparable late
+uptake, which is consistent with the publisher's historical reluctance
+to engage in negotations with library consortia. However, by 2023, the
+share of open access enabled by transformative agreements appeared to
+stabilise for all three publishers (see Figure
+\ref{fig-uptake_publisher}C). Interestingly, the differences between
+first and corresponding author affiliations were more pronounced at the
+publisher level. In Scopus, for example, the share of open access via
+transformative agreements measured by corresponding authorship was
+greater for Springer Nature in 2023 than when using first authorship.
 
@@
 Elsevier starting in 2019, which allowed delayed open access under the
-publisher's user license (Rabesandratana, 2019). This license was not
-classified as hybrid open access in Scopus or hoaddata. In all cases,
-Spearman rank correlations were \(\rho > .9\), \(p < 0.001\), showing a
-high level of correlations between the databases and authorship roles
-considered.
-
-Transformative agreements appeared to be a key driver of open access
-growth (see Figure \ref{fig-uptake_country}C and D). OECD members
+publisher's user license (Rabesandratana, 2019). This licence was not
+classified as hybrid open access in either Scopus or hoaddata. In all
+cases, Spearman rank correlations were \(\rho > .9\), \(p < 0.001\),
+showing a high level of correlations between the databases and
+authorship roles considered.
+
+Transformative agreements appeared to be a key driver of national open
+access growth (see Figure \ref{fig-uptake_country}C and D). OECD members
 accounted for the majority of open access articles enabled by
@@
 agreements with major publishers from 2022 onward.\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}
-Results were consistent across data sources, with a strong correlation
-observed. However, Wiley surplus in hoaddata led to better rankings for
-countries where Wiley played a substantial role, such as Germany, where
-the DEAL consortium negotiated its first transformative agreements with
-Wiley that started in July 2019.
+Results were consistent across data sources. However, Wiley's open
+access surplus in hoaddata led to better rankings for countries where
+Wiley played a substantial role, such as Germany, where the DEAL
+consortium negotiated its first transformative agreements with Wiley
+that started in July 2019.
 
@@
 
+\section{References}\label{references}
+
 \phantomsection\label{refs}
@@
 
+\bibitem[\citeproctext]{ref-Borrego_2020}
+Borrego, Á., Anglada, L., \& Abadal, E. (2021). Transformative
+agreements: Do they pave the way to open access? \emph{Learned
+Publishing}, \emph{34}(2), 216--232.
+\url{https://doi.org/10.1002/leap.1347}
+
 \bibitem[\citeproctext]{ref-Donner_2017}
@@
 \emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}
-
-\bibitem[\citeproctext]{ref-van_Eck_2022}
-Eck, N. J. van, \& Waltman, L. (2022). \emph{Crossref as a source of
-open bibliographic metadata}.
-\url{https://doi.org/10.31222/osf.io/smxe5}
 
@@
 
+\bibitem[\citeproctext]{ref-de_Jonge_2025}
+Jonge, H. de, Kramer, B., \& Sondervan, J. (2025). \emph{Tracking
+transformative agreements through open metadata: Method and validation
+using dutch research council NWO funded papers.}
+\url{https://doi.org/10.31222/osf.io/tz6be_v1}
+
 \bibitem[\citeproctext]{ref-ComplexUpset}
@@
 1983--1992. \url{https://doi.org/10.1109/tvcg.2014.2346248}
+
+\bibitem[\citeproctext]{ref-Maisano_2025}
+Maisano, D. A., Mastrogiacomo, L., Ferrara, L., \& Franceschini, F.
+(2025). A large-scale semi-automated approach for assessing
+document-type classification errors in bibliometric databases.
+\emph{Scientometrics}. \url{https://doi.org/10.1007/s11192-025-05244-y}
 
@@
 
+\bibitem[\citeproctext]{ref-Stahlschmidt_2022}
+Stahlschmidt, S., \& Stephen, D. (2022). From indexation policies
+through citation networks to normalized citation impacts: Web of
+science, scopus, and dimensions as varying resonance chambers.
+\emph{Scientometrics}, \emph{127}(5), 2413--2431.
+\url{https://doi.org/10.1007/s11192-022-04309-6}
+
+\bibitem[\citeproctext]{ref-Suber_2012}
+Suber, P. (2012). \emph{Open access}. The MIT Press.
+\url{https://doi.org/10.7551/mitpress/9286.001.0001}
+
 \bibitem[\citeproctext]{ref-Visser_2021}
@@
 
+\bibitem[\citeproctext]{ref-Zhang_2024}
+Zhang, L., Cao, Z., Shang, Y., Sivertsen, G., \& Huang, Y. (2024).
+Missing institutions in OpenAlex: Possible reasons, implications, and
+solutions. \emph{Scientometrics}.
+\url{https://doi.org/10.1007/s11192-023-04923-y}
+
 \end{CSLReferences}

### 2025-03-28 analysis/manuscript.qmd
Finish first draft discussion, next introduction

@@
 Results were consistent across data sources. 
-However, Wiley's open access surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with Wiley that started in July 2019.
+However, Wiley's open access surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with publisher that started in July 2019.
 
@@
 
-Investigating over 13,000 hybrid journals included in the cOAlition Journal Checker Tool shows a substantial increase in  open access enabled by transformative agreements between 2019 and 2023, although most articles in these journals remained paywalled during this five-year period.
-While transformative agreements accounted for the majority of open access, a considerable number of open access articles are likely to have resulted from the payment of individual publication fees.
-The results confirm that transformative agreements and hybrid open access continue to be concentrated among a few large commercial publishers.
-They also highlight that high-income European countries in particular are using transformative agreements to make most of their article output open access, although South Africa showed similar levels of open access adoption.
+This study of over 13,000 hybrid journals included in the cOAlition Journal Checker Tool shows a rise in open access due to transformative agreements between 2019 and 2023, although most articles remained paywalled.
+While transformative agreements accounted for the majority of open access, many articles continue to become open through the payment of individual publication fees.
+Hybrid open access and transformative agreements remain concentrated among a small group of large commercial publishers, 
+with high-income European countries—alongside South Africa—showing high adoption rates.
+In contrast, the three most productive countries, China, the United States, and India, were ranked much behind other countries in the adoption of transformative agreements.
 These trends were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science.
+An open question remains whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or countries voluntarily refraining from such agreements.
 
@@
 However, several notable discrepancies warrant discussion.
-They stem not only from known limitations of the open scholarly metadata sources Crossref and OpenAlex, but also from the proprietary databases Scopus and Web of Science.
+They arise from limitations in both open and proprietary databases.
 
 The coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage.
-Differences can be found for journals targeting practitioners or local non-English language communties, with many such titles indexed exclusively in Crossref and Scopus.
-Using the open index Crossref demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
+Differences can be found for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
+Using the open index Crossref as bibliometric database in hoaddata demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
 
@@
 
-A notable advantage of using Scopus and Web of Science to study the impact of transformative agreements is their comprehensive differentiation of journal articles by document type, as many agreements restrict funding to eligible article types.
-This resulted in slightly higher rates of open access uptake compared to hoaddata.
-
-A similar limitation of studying open access with open scholarly data sources is the often reported lack of corresponding authorship information. 
-However, indicators using first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, which is not surprising given disciplinary norms in scholarly publishing.
-First authors are often considered to have carried out the main research underlying a paper, while the corresponding author is often recognised as having supervised a paper. 
+A notable advantage of using Scopus and Web of Science to study the impact of transformative agreements is their comprehensive differentiation of journal articles by document type, as many agreements restrict funding to original articles and reviews.
+This resulted in slightly higher open access rates in Scopus and Web of Science.
+
+A similar limitation of studying open access with less selective databases is the often reported lack of corresponding authorship information [@Fraser_2023; @Haucap_2021; @Shu_2023]. 
+However, indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, in line with previous research, reflecting disciplinary norms in scholarly publishing with regard to contributorship and author roles and positions [@Larivi_re_2016].
+First authors are often considered to have carried out the main research underlying a paper, while the corresponding author is often recognised as having supervised the research [@Mattsson_2010; @Fox_2018].
 It is therefore not surprising that both are strongly correlated, suggesting that first and corresponding authors share the same country affiliation.
-Moreover, in most cases the first author is indeed the same as the corresponding author.
-As such, first authorship can be used confidentially in large-scale studies and those focusing on national consortia to assess the impact of transformative agreements.
-
-However, affiliation data is sparser in OpenAlex than in Scopus and Web of Science, resulting in slightly more articles that could be attributed to transformative agreements. 
-However, in order to establish a link, a thorough mapping of the database affiliation identifier to the ROR ID, the JCT's identifier for institutional participation in transformative agreements, is required.
-In contrast, ROR IDs are the canonical organisation identifier in OpenAlex, allowing for less ambiguous matching between agreement and publication.
-
-Comparison of scholarly databases is also valuable in identifying differences in open access identification. 
-Not all publishers deposit a Creative Commons licence with Crossref, the citation authority.
-A collaboration with Scopus and Web of Science helped to identify such publishers, as their underlying unpaywall data also parses licence information from websites.
-However, there were significant differences between Scopus and Web of Science in terms of open access indexing, particularly with regard to the delayed open access provided by Elsevier.
-In the case of Wiley, however, there were also problems with the analysis of the Unpaywall.
-This suggests that hybrid open access studies should not only refer to a single source of evidence.
-Instead, open access tags can be validated by the combiantion of seevral data sources, as in our case where metadata from publishers was compared with data from open access discovery services such as Unpaywall.
-It is also particularly useful to include a curated list of journals under the hybrid model to avoid misclassifications based on a journal's business model [@Visser_2021].
+Moreover, in most cases the first author is indeed the same as the corresponding author [@Chinchilla_Rodr_guez_2024].
+
+The slight increase in the number of open access articles by corresponding authors compared to the number of first authors in the last two years observed across Springer Nature hybrid journals raises the question about institutional open access sponsorship practices and if they influence the assignment of author roles among co-authors, especially when funding opportunities are not consistently available within the author teams [@Gumpenberger_2018].
+Previous research notes effects of institutionalised bibliometrics praticses on the valuations of authorship positions [@Helgesson_2020], although, to my knowledge, no such evidence was reported for open access funding yet.
+As such, first authorship can reliably be used in large-scale studies and those focusing on national consortia to assess the impact of transformative agreements.
+
+While OpenAlex’s affiliation coverage is less complete, it shows strong correlations with Scopus [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric]. 
+In this study, this limitation likely reduced the number of articles confidently attributed to transformative agreements, despite high correlations between OpenAlex, Scopus, and Web of Science at the country level.
+Nevertheless, OpenAlex’s native ROR ID integration offers a distinct advantage: unlike Scopus and Web of Science, which require reconciliation of proprietary organisation identifiers with this open standard, it enables more reliable identification of agreement-enabled articles.
+<!-- Dimension is another scholarly data source that uses open standards to identify journals, which enables analysing transformative agreements [@Fraser_2023].-->
+
+Comparing scholarly databases helps to identify  discrepancies in open access evidence. 
+Not all publishers deposit CC licences with Crossref, but Unpaywall, integrated into Scopus and Web of Science, can make this transparent, because the open access discovery services also analyse publishers' websites.
+However, inconsistencies remain: Web of Science incorrectly identifies Elsevier's delayed open access as hybrid, while Scopus correctly reflects it.
+Unpaywall also missed a substantial number of open access articles with CC license according to Crossref, resulting in fewer open access articles in Scopus and Web of Science, probably due to problems parsing the publisher's website.
+As these differences may affect comparability between databases, research and monitoring exercises should not rely on a single source of evidence.
+Instead, it is recommended that the choice of open access evidence sources should be documented and verified by combining multiple data sources [@Huang_2020].
+It is particularly useful to include expert-curated lists of journals to avoid misclassifications based on a journal's business model [@Visser_2021].
+
+It is important to note that articles from institutions involved in transformative agreements are estimates due to a lack of access to invoice data, which is not usually shared between library consortia and publishers.
+Furthermore, it is important to acknowledge that the sources of transformative agreements and the underlying ESAC registry are a voluntary effort, crowd-sourced from various consortia. 
+However, recent validation with Dutch Research Information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements. 
+With enhanced and streamlined metadata compliance, attributable to evolving standards and initiatives to support negotiations with publishers, particularly through the communities of the ESAC initiative and Barcelona Declaration supporters, the situation is likely to improve.
+In conclusion, this cross-database analysis has highlighted the potential and challenges of establishing open scholarly data as a suitable evidence base for analysing the transition from subscription-based publishing to open access.
 
@@
 
-<!-- Limitations: agreeemnts are complex, game auhtorship, invoicing not available, mapping, not feasible, requires more insitutional and publsihjer case studies involving financial data, but also improved and streamlined metadata compliance according to the evlolving stadatrd and iniatives (ESAC, Varcelona) in negotiation with publishers.
+<!-- Limitations: agreeemnts are complex, game auhtorship, invoicing not available, mapping, not feasible, requires more insitutional and publsihjer case studies involving financial data, but also improved and streamlined metadata compliance according to the evlolving stadatrd and iniatives (ESAC, Barcelona) in negotiation with publishers.
 

### 2025-03-28 analysis/manuscript.tex
Finish first draft discussion, next introduction

@@
 Wiley played a substantial role, such as Germany, where the DEAL
-consortium negotiated its first transformative agreements with Wiley
+consortium negotiated its first transformative agreements with publisher
 that started in July 2019.
@@
 
-Investigating over 13,000 hybrid journals included in the cOAlition
-Journal Checker Tool shows a substantial increase in open access enabled
-by transformative agreements between 2019 and 2023, although most
-articles in these journals remained paywalled during this five-year
-period. While transformative agreements accounted for the majority of
-open access, a considerable number of open access articles are likely to
-have resulted from the payment of individual publication fees. The
-results confirm that transformative agreements and hybrid open access
-continue to be concentrated among a few large commercial publishers.
-They also highlight that high-income European countries in particular
-are using transformative agreements to make most of their article output
-open access, although South Africa showed similar levels of open access
-adoption. These trends were consistent across the investigated open data
-source hoaddata, derived from Crossref and OpenAlex, and the established
-proprietary bibliometric databases Scopus and Web of Science.
+This study of over 13,000 hybrid journals included in the cOAlition
+Journal Checker Tool shows a rise in open access due to transformative
+agreements between 2019 and 2023, although most articles remained
+paywalled. While transformative agreements accounted for the majority of
+open access, many articles continue to become open through the payment
+of individual publication fees. Hybrid open access and transformative
+agreements remain concentrated among a small group of large commercial
+publishers, with high-income European countries---alongside South
+Africa---showing high adoption rates. In contrast, the three most
+productive countries, China, the United States, and India, were ranked
+much behind other countries in the adoption of transformative
+agreements. These trends were consistent across the investigated open
+data source hoaddata, derived from Crossref and OpenAlex, and the
+established proprietary bibliometric databases Scopus and Web of
+Science. An open question remains whether this uneven distribution
+reflects temporary implementation gaps, inherent inequities in the
+transformative agreement model, or countries voluntarily refraining from
+such agreements.
 
@@
 authorship and open access status. However, several notable
-discrepancies warrant discussion. They stem not only from known
-limitations of the open scholarly metadata sources Crossref and
-OpenAlex, but also from the proprietary databases Scopus and Web of
-Science.
+discrepancies warrant discussion. They arise from limitations in both
+open and proprietary databases.
 
@@
 Differences can be found for journals targeting practitioners or local
-non-English language communties, with many such titles indexed
-exclusively in Crossref and Scopus. Using the open index Crossref
-demonstrated particular strength in identifying newly established hybrid
-journals, a notable finding given that transformative agreements
-primarily target existing subscription-based journals.
+non-English language communities, with many such titles indexed
+exclusively in Crossref and Scopus. Using the open index Crossref as
+bibliometric database in hoaddata demonstrated particular strength in
+identifying newly established hybrid journals, a notable finding given
+that transformative agreements primarily target existing
+subscription-based journals.
 
@@
 differentiation of journal articles by document type, as many agreements
-restrict funding to eligible article types. This resulted in slightly
-higher rates of open access uptake compared to hoaddata.
-
-A similar limitation of studying open access with open scholarly data
-sources is the often reported lack of corresponding authorship
-information. However, indicators using first authors, which have often
-been used as a proxy for determining open access funding, and
-corresponding authors show a high level of correlation, which is not
-surprising given disciplinary norms in scholarly publishing. First
-authors are often considered to have carried out the main research
-underlying a paper, while the corresponding author is often recognised
-as having supervised a paper. It is therefore not surprising that both
-are strongly correlated, suggesting that first and corresponding authors
+restrict funding to original articles and reviews. This resulted in
+slightly higher open access rates in Scopus and Web of Science.
+
+A similar limitation of studying open access with less selective
+databases is the often reported lack of corresponding authorship
+information (Fraser et al., 2023; Haucap et al., 2021; Shu \& Larivière,
+2023). However, indicators based on first authors, which have often been
+used as a proxy for determining open access funding, and corresponding
+authors show a high level of correlation, in line with previous
+research, reflecting disciplinary norms in scholarly publishing with
+regard to contributorship and author roles and positions (Larivière et
+al., 2016). First authors are often considered to have carried out the
+main research underlying a paper, while the corresponding author is
+often recognised as having supervised the research (Fox et al., 2018;
+Mattsson et al., 2010). It is therefore not surprising that both are
+strongly correlated, suggesting that first and corresponding authors
 share the same country affiliation. Moreover, in most cases the first
-author is indeed the same as the corresponding author. As such, first
-authorship can be used confidentially in large-scale studies and those
-focusing on national consortia to assess the impact of transformative
-agreements.
-
-However, affiliation data is sparser in OpenAlex than in Scopus and Web
-of Science, resulting in slightly more articles that could be attributed
-to transformative agreements. However, in order to establish a link, a
-thorough mapping of the database affiliation identifier to the ROR ID,
-the JCT's identifier for institutional participation in transformative
-agreements, is required. In contrast, ROR IDs are the canonical
-organisation identifier in OpenAlex, allowing for less ambiguous
-matching between agreement and publication.
-
-Comparison of scholarly databases is also valuable in identifying
-differences in open access identification. Not all publishers deposit a
-Creative Commons licence with Crossref, the citation authority. A
-collaboration with Scopus and Web of Science helped to identify such
-publishers, as their underlying unpaywall data also parses licence
-information from websites. However, there were significant differences
-between Scopus and Web of Science in terms of open access indexing,
-particularly with regard to the delayed open access provided by
-Elsevier. In the case of Wiley, however, there were also problems with
-the analysis of the Unpaywall. This suggests that hybrid open access
-studies should not only refer to a single source of evidence. Instead,
-open access tags can be validated by the combiantion of seevral data
-sources, as in our case where metadata from publishers was compared with
-data from open access discovery services such as Unpaywall. It is also
-particularly useful to include a curated list of journals under the
-hybrid model to avoid misclassifications based on a journal's business
-model (Visser et al., 2021).
+author is indeed the same as the corresponding author
+(Chinchilla-Rodríguez et al., 2024).
+
+The slight increase in the number of open access articles by
+corresponding authors compared to the number of first authors in the
+last two years observed across Springer Nature hybrid journals raises
+the question about institutional open access sponsorship practices and
+if they influence the assignment of author roles among co-authors,
+especially when funding opportunities are not consistently available
+within the author teams (Gumpenberger et al., 2018). Previous research
+notes effects of institutionalised bibliometrics praticses on the
+valuations of authorship positions (Helgesson, 2020), although, to my
+knowledge, no such evidence was reported for open access funding yet. As
+such, first authorship can reliably be used in large-scale studies and
+those focusing on national consortia to assess the impact of
+transformative agreements.
+
+While OpenAlex's affiliation coverage is less complete, it shows strong
+correlations with Scopus (Akbaritabar et al., 2024; Alperin et al.,
+2024). In this study, this limitation likely reduced the number of
+articles confidently attributed to transformative agreements, despite
+high correlations between OpenAlex, Scopus, and Web of Science at the
+country level. Nevertheless, OpenAlex's native ROR ID integration offers
+a distinct advantage: unlike Scopus and Web of Science, which require
+reconciliation of proprietary organisation identifiers with this open
+standard, it enables more reliable identification of agreement-enabled
+articles.
+
+Comparing scholarly databases helps to identify discrepancies in open
+access evidence. Not all publishers deposit CC licences with Crossref,
+but Unpaywall, integrated into Scopus and Web of Science, can make this
+transparent, because the open access discovery services also analyse
+publishers' websites. However, inconsistencies remain: Web of Science
+incorrectly identifies Elsevier's delayed open access as hybrid, while
+Scopus correctly reflects it. Unpaywall also missed a substantial number
+of open access articles with CC license according to Crossref, resulting
+in fewer open access articles in Scopus and Web of Science, probably due
+to problems parsing the publisher's website. As these differences may
+affect comparability between databases, research and monitoring
+exercises should not rely on a single source of evidence. Instead, it is
+recommended that the choice of open access evidence sources should be
+documented and verified by combining multiple data sources (Huang et
+al., 2020). It is particularly useful to include expert-curated lists of
+journals to avoid misclassifications based on a journal's business model
+(Visser et al., 2021).
+
+It is important to note that articles from institutions involved in
+transformative agreements are estimates due to a lack of access to
+invoice data, which is not usually shared between library consortia and
+publishers. Furthermore, it is important to acknowledge that the sources
+of transformative agreements and the underlying ESAC registry are a
+voluntary effort, crowd-sourced from various consortia. However, recent
+validation with Dutch Research Information demonstrates the reliability
+of such an open approach for the assessment of articles under
+transformative agreements. With enhanced and streamlined metadata
+compliance, attributable to evolving standards and initiatives to
+support negotiations with publishers, particularly through the
+communities of the ESAC initiative and Barcelona Declaration supporters,
+the situation is likely to improve. In conclusion, this cross-database
+analysis has highlighted the potential and challenges of establishing
+open scholarly data as a suitable evidence base for analysing the
+transition from subscription-based publishing to open access.
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-Akbaritabar_2024}
+Akbaritabar, A., Theile, T., \& Zagheni, E. (2024). Bilateral flows and
+rates of international migration of scholars for 210 countries for the
+period 1998-2020. \emph{Scientific Data}, \emph{11}(1).
+\url{https://doi.org/10.1038/s41597-024-03655-9}
+
+\bibitem[\citeproctext]{ref-alperin2024analysissuitabilityopenalexbibliometric}
+Alperin, J. P., Portenoy, J., Demes, K., Larivière, V., \& Haustein, S.
+(2024). An analysis of the suitability of OpenAlex for bibliometric
+analyses. \emph{arXiv}. \url{https://arxiv.org/abs/2404.17663}
+
 \bibitem[\citeproctext]{ref-Baas_2020}
@@
 
+\bibitem[\citeproctext]{ref-Chinchilla_Rodr_guez_2024}
+Chinchilla-Rodríguez, Z., Costas, R., Robinson-García, N., \& Larivière,
+V. (2024). Examining the quality of the corresponding authorship field
+in {Web of Science} and {Scopus}. \emph{Quantitative Science Studies},
+\emph{5}(1), 76--97. \url{https://doi.org/10.1162/qss_a_00288}
+
 \bibitem[\citeproctext]{ref-Donner_2017}
@@
 \emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}
+
+\bibitem[\citeproctext]{ref-Fox_2018}
+Fox, C. W., Ritchey, J. P., \& Paine, C. E. T. (2018). Patterns of
+authorship in ecology and evolution: First, last, and corresponding
+authorship vary with gender and geography. \emph{Ecology and Evolution},
+\emph{8}(23), 11492--11507. \url{https://doi.org/10.1002/ece3.4584}
+
+\bibitem[\citeproctext]{ref-Fraser_2023}
+Fraser, N., Hobert, A., Jahn, N., Mayr, P., \& Peters, I. (2023). No
+deal: German researchers' publishing and citing behaviors after big deal
+negotiations with {Elsevier}. \emph{Quantitative Science Studies},
+\emph{4}(2), 325--352. \url{https://doi.org/10.1162/qss_a_00255}
 
@@
 
+\bibitem[\citeproctext]{ref-Gumpenberger_2018}
+Gumpenberger, C., Hölbling, L., \& Gorraiz, J. I. (2018). On the issues
+of a {``corresponding author''} field-based monitoring approach for gold
+open access publications and derivative cost calculations.
+\emph{Frontiers in Research Metrics and Analytics}, \emph{3}.
+\url{https://doi.org/10.3389/frma.2018.00001}
+
+\bibitem[\citeproctext]{ref-Haucap_2021}
+Haucap, J., Moshgbar, N., \& Schmal, W. B. (2021). The impact of the
+{German {``DEAL''}} on competition in the academic publishing market.
+\emph{Managerial and Decision Economics}, \emph{42}(8), 2027--2049.
+\url{https://doi.org/10.1002/mde.3493}
+
 \bibitem[\citeproctext]{ref-Haupka_2024}
@@
 
+\bibitem[\citeproctext]{ref-Helgesson_2020}
+Helgesson, G. (2020). Authorship order and effects of changing
+bibliometrics practices. \emph{Research Ethics}, \emph{16}(1--2), 1--7.
+\url{https://doi.org/10.1177/1747016119898403}
+
 \bibitem[\citeproctext]{ref-Hendricks_2020}
@@
 \url{https://doi.org/10.1162/qss_a_00022}
+
+\bibitem[\citeproctext]{ref-Huang_2020}
+Huang, C.-K., Neylon, C., Hosking, R., Montgomery, L., Wilson, K.,
+Ozaygen, A., \& Brookes-Kenworthy, C. (2020). \emph{Evaluating
+institutional open access performance: Sensitivity analysis}.
+\url{https://doi.org/10.1101/2020.03.19.998542}
 
@@
 
+\bibitem[\citeproctext]{ref-Larivi_re_2016}
+Larivière, V., Desrochers, N., Macaluso, B., Mongeon, P., Paul-Hus, A.,
+\& Sugimoto, C. R. (2016). Contributorship and division of labor in
+knowledge production. \emph{Social Studies of Science}, \emph{46}(3),
+417--435. \url{https://doi.org/10.1177/0306312716650046}
+
 \bibitem[\citeproctext]{ref-Lex_2014}
@@
 \url{https://doi.org/10.1080/00031305.2017.1375986}
+
+\bibitem[\citeproctext]{ref-Mattsson_2010}
+Mattsson, P., Sundberg, C. J., \& Laget, P. (2010). Is correspondence
+reflected in the author position? A bibliometric study of the relation
+between corresponding author and byline position. \emph{Scientometrics},
+\emph{87}(1), 99--105. \url{https://doi.org/10.1007/s11192-010-0310-9}
 
@@
 
+\bibitem[\citeproctext]{ref-Shu_2023}
+Shu, F., \& Larivière, V. (2023). The oligopoly of open access
+publishing. \emph{Scientometrics}, \emph{129}(1), 519--536.
+\url{https://doi.org/10.1007/s11192-023-04876-2}
+
 \bibitem[\citeproctext]{ref-Simard_2024}

### 2025-04-03 analysis/manuscript.qmd
Proofreading

@@
  % Matching Process - Fixed line break syntax
-    \node (matching) [process, below=1.5cm of $(hoaddata)!1!(kb)$] {Harmonising author affiliations through RoR};
+    \node (matching) [process, below=1.5cm of $(hoaddata)!1!(kb)$] {Harmonising author affiliations through ROR};
 
  % Aggregation
-    \node (transmatch) [process, below=1.5cm of $(matching)!1!(matching)$] {Estimate TA output};
+    \node (transmatch) [process, below=1.5cm of $(matching)!1!(matching)$] {Estimate article output under transformative agreements};
 
@@
 \caption{Data processing workflow for comparing hybrid open access uptake across bibliometric data sources.
-The workflow shows how data from different sources (hoaddata, derived from Crossref, OpenAlex, Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT), the Web of Science, and Scopus) is processed and matched using ROR IDs to enable comparative analysis.
-TA = Transformative Agreement.}
+The workflow shows how data from different sources (hoaddata, derived from Crossref, OpenAlex, Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT), the Web of Science, and Scopus) were processed to enable comparative analysis.}
 \label{fig:workflow}
@@
 
-hoaddata, developed and maintained by the author to support open access monitoring and research [@Jahn_2025], is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
-It combines article-level metadata from Crossref [@Hendricks_2020] and affilation metadata from OpenAlex [@priem2022openalex] with transformative agreement information from Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data about participating research organisations to agreements in the ESAC registry.
+hoaddata, developed and maintained by the author to support open access monitoring and research (e.g.
+@Jahn_2025), is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
+It combines article-level metadata from Crossref [@Hendricks_2020] and affilation metadata from OpenAlex [@priem2022openalex] with transformative agreement information from the Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data about participating research organisations to agreements in the ESAC registry.
 
@@
 Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020].
-Web of Science is a selective, base research-focused database with dedicated panels to choose journals for inclusion [@Visser_2021; @Stahlschmidt_2022].
+Web of Science is a selective, base research-focused database [@Visser_2021; @Stahlschmidt_2022].
 The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).
@@
 
-Following @Jahn_2025, the starting point was a unified dataset of several safeguarded JCT snapshots[^jct_data], as provided by hoaddata.
-The JCT journal data in hoaddata were enriched with ISSN variants linked to an ISSN-L. 
+Following @Jahn_2025, the starting point was a unified dataset of several safeguarded JCT snapshots[^jct_data].
+The JCT journal data were enriched with ISSN variants linked to an ISSN-L. 
 To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists (DOAJ, OpenAlex, Bielefeld).
-These data were used to determine the publication volume of hybrid journals for each database independently.
-
-hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons licence information relative to the published version ("version of record").
+These data were made available via hoaddata and used to determine the publication volume of hybrid journals for each database independently.
+
+hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons (CC) licence information relative to the published version ("version of record").
 <!-- Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources  affiliation metadata from OpenAlex.-->
@@
 -->
-The article metadata included DOIs, publication dates, open access information as well as author roles and institutional affiliations.
+The article metadata included DOIs, publication dates, open access information as well as author roles and affiliations.
 Publication years were determined using the earliest known date of publication in a journal.
@@
 
-Articles in hybrid journals were considered open access when they were made freely available under a Creative Commons license on publishers' platforms. 
-While hoaddata sourced this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall as evidence source. 
-Unpaywall also uses Crossref license metadata, but supplements them by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable Creative Commons license information [@Piwowar_2018].
-This additional parsing remains necessary despite transformative agreement workflows recommend the depositing license information during DOI registration [@Geschuhn_2017].
-Both Web of Science and Scopus defined hybrid open access consistently as content available under Creative Commons licenses on publisher platforms according to their documentations[^wosoa][^scopusoa], distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses.
+Articles in hybrid journals were considered open access when they were made freely available under a CC license on publisher platforms. 
+While hoaddata obtained this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall as evidence source. 
+Unpaywall also uses Crossref license metadata, but supplements them by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable CC license information [@Piwowar_2018].
+This additional parsing remains necessary despite transformative agreement workflows recommend to deposit CC licenses during DOI registration [@Geschuhn_2017].
+Both Web of Science and Scopus defined hybrid open access consistently as content available under CC licenses on publisher platforms according to their documentations[^wosoa][^scopusoa], distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses [@Piwowar_2018].
 
@@
 
-Because neither Web of Science nor Scopus support ROR IDs, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise affiliation data.
+Because neither Web of Science nor Scopus support ROR at the time of the data retrieval, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise affiliation data.
 First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
@@
 Upon inspection, these mismatches primarily occurred with less-represented institutions having only a few publications, introduced through multiple affiliations of single authors. 
-The difference between databases suggests that Scopus's affiliation control aligns more closely with ROR than that of the Web of Science.
+The difference between databases suggests that Scopus' affiliation control aligns more closely with ROR than that of the Web of Science.
 
@@
 
-Based on these generated matching tables, articles eligible under transformative agreements could also be retrieved from Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT.
-The estimation of eligible articles followed @Jahn_2025 and included a matching of both journals and participating institutions.
+Based on these so-compiled matching tables, articles eligible under transformative agreements could also be obtained from Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT.
+The estimation of eligible articles followed @Jahn_2025 and included a matching of both journals and participating institutions according to the JCT.
 The matching also took into account the duration of agreements according to the ESAC registry, with only those matches where an agreement was actually in place being considered for subsequent analysis.
@@
 
-This section first presents the indexing coverage of hybrid open access by comparing the data from open data sources with the proprietary bibliometric databases Scopus and Web of Science. 
+This section first presents the indexing coverage of hybrid open access by comparing open data sources with the proprietary bibliometric databases Scopus and Web of Science. 
 Then, using the same methods, indicators at the publisher and country level are calculated independently for each database and compared with each other to assess the suitability of the bibliometric databases for investigating transformative agreements.
@@
 ```{r upset_coverage_results, fig.asp=0.75}
-#| fig-cap: "Upset graph"
+#| fig-cap: "Comparison of the hybrid open access indexing coverage of hoaddata (based on Crossref), Scopus and Web of Science, 2019-2023.
+Only hybrid journals  present in the cOAlition S Transformative Agreement Data dump with at least one open access article are considered.
+A) presents the number of journals, B) the number of articles (DOI), distinguishing between shared DOI corpus and surplus in hoaddata.
+Box plots (C) shows the five-years journal article volume (log-scale)."
 #| label: fig-upset_coverage_results
@@
 Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
-Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows that there is a large spread across the journals covered by all three data sources. 
-Furthermore, journals in this set published more on average than journals covered by only one or two data sources. 
-In particular, the journals covered exclusively by hoaddata were substantially smaller than those covered by all three sources. 
+Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows a large spread across the journals jointly covered by all three data sources. 
+These journals published more on average than journals covered by only one or two data sources. 
+In particular, the journals covered exclusively by hoaddata were substantially smaller. 
 Upon inspection, these were often newly launched hybrid journals, which explains the relatively low five-year publication volume.
@@
 Figure \ref{fig-upset_coverage_results_publisher} presents the coverage of hybrid journals in transformative agreements across data sources from 2019 to 2023 with a focus on publisher portfolios.
-The analysis highlights the dominance of the three largest publishers--Elsevier, Springer Nature, and Wiley--, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of articles published during this five-year period.
+The analysis highlights the dominance of the three largest publishers, Elsevier, Springer Nature, and Wiley, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of articles published during this five-year period.
 In terms of article volume, Elsevier led with `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_art)*100`% of the total) published across `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_jn)*100`% of the total).
@@
 ```{r upset_publisher_plot, fig.asp=0.75}
-#| fig-cap: "Upset graph publisher"
+#| fig-cap: "Comparison of the hybrid open access indexing coverage of hoaddata (based on Crossref), Scopus and Web of Science, 2019-2023 by publisher.
+Only hybrid journals  present in the cOAlition S Transformative Agreement Data dump with at least one open access article are considered.
+A) presents the percentage of journals by publisher, B) the number of articles by publisher (shared DOI).
+Box plots (C) shows the five-years journal article volume (log-scale) by publisher."
 #| label: fig-upset_coverage_results_publisher
@@
 
-The three largest publishers--Elsevier, Springer Nature and Wiley--were best represented in the exclusive intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
+The three largest publishers, Elsevier, Springer Nature, and Wiley, were best represented in the exclusive intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
 Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage (n = `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()`; `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as determined through shared DOIs.
@@
 In particular, `r miss_tas$esac_publisher[1]` represented `r miss_tas$journals[1]` journals with `r miss_tas$n[1]` articles, `r miss_tas$esac_publisher[2]` accounted for `r miss_tas$journals[2]` journals with `r miss_tas$n[2]` articles, and `r miss_tas$esac_publisher[3]` recorded `r miss_tas$journals[3]` journals and `r miss_tas$n[3]` articles.
-In total, `r nrow(miss_publishers)` publishers were not represented in Crossref.
+In total, `r nrow(miss_publishers)` publishers did not share CC licenses with Crossref and were therefore not represented in hoaddata.
 
@@
 ```{r, plot_uptake_overview, fig.asp=.5}
-#| fig-cap: "Open Access indicato"
+#| fig-cap: "Development of open access in hybrid journals included in transformative agreements, 2019 and 2023 by data source and author role.
+Figure shows the open access percentage (A), the number (B) and percentage (C) of open access articles enabled transformative agreements."
 #| label: fig-uptake_overview
@@
 
-Figure \ref{fig-uptake_overview}A shows a moderate growth of open access in hybrid journals, which is consistent across hoaddata (black line), Scopus (orange line), and Web of Science (blue line).
-According to hoaddata, hybrid open access uptake increased from `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
+Figure \ref{fig-uptake_overview}A shows a moderate growth of open access in hybrid journals, which is consistent across hoaddata (black line), Scopus (yellow line), and Web of Science (blue line).
+According to hoaddata, hybrid open access uptake increased from `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
 Similarly, Scopus recorded an growth from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
@@
 Similarly, hybrid open access by transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
-Trends were consistent for first (dashed line) and corresponding author (dotted line) affiliations. 
+Trends were consistent when measuring first (dashed line) and corresponding author (dotted line) affiliations. 
 According to Scopus, `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_first_author_articles) |> sum()` open access articles could be attributed to transformative agreements based on first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`) and `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
@@
 From 2021 (hoadata, Scopus) resp. 2022 (Web of Science),  transformative agreements enabled the majority of hybrid open access.
-For first authors, the share ranged between `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and   `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science) in 2023.
+For first authors, the share ranged between `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and   `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science) in 2023.
 For corresponding authors, the shares were slightly larger, with Scopus recording  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% and Web of Science `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% in 2023.
@@
 ```{r plot_publisher_uptake_overview, fig.asp=1.1}
-#| fig-cap: "Open Access indicato"
+#| fig-cap: "Development of open access in hybrid journals included in transformative agreements, 2019 and 2023, by data source, author role and publisher.
+Figure shows the open access percentage (A), the number (B) and percentage (C) of open access articles enabled transformative agreements."
 #| label: fig-uptake_publisher
@@
 Figure \ref{fig-uptake_country} presents hybrid open access indicators by country, comparing hoaddata (x-axis) with Web of Science and Scopus (y-axis).
-Indicators calculated from these proprietary databases are shown for both first and corresponding authors, with full counting used to account for multiple country affiliations.
+Indicators calculated from these proprietary databases are shown for both first and corresponding authors, with full counting used to account for multiple country affiliations [@Hottenrott_2021].
 
@@
 ```{r oa_country_comparision_plot_, fig.asp=1, fig.width = 8}
-#| fig-cap: "Open Access by country"
+#| fig-cap: "Comparison of hybrid open access by country, 2019-2023.
+Scatterplots distinguish between proprietary database Scopus and Web of Science, and author role.
+x axis shows hoaddata indicators.
+A) Five-years article volume, B) open access percentage in hybrid journals, (A and B limited to countries with > 10.000 articles published), C) number and D) percentage of open access articles enabled by transformative agreements (limited to countries with > 1.000 open access articles enabled by transformative agreements).
+Line represents line of equality.
+An interactive version is accessible here: <https://najkoja-hoa-replication.hf.space/>"
 #| label: fig-uptake_country
@@
 
-A closer examination of the proportion of hybrid open access enabled by transformative agreements (see Figure \ref{fig-uptake_country}D) revealed that the French Elsevier agreement on delayed open access was not included in JCT data, as it did not qualify as a transformative agreement.
-Medium-sized European countries again showed a high proportion of hybrid open access through transformative agreements, highlighting the impact of this licensing model across all three data sources.
-In contrast, the United States had a low proportion of hybrid open access enabled by agreements, suggesting that a substantial number of open access articles were likely financed through other means, such as article processing charges (APCs).
+<!-- A closer examination of the proportion of hybrid open access enabled by transformative agreements (see Figure \ref{fig-uptake_country}D) revealed that the French Elsevier agreement on delayed open access was not included in JCT data, as it did not qualify as a transformative agreement.
+-->
+Medium-sized European countries again showed a high proportion of hybrid open access through transformative agreements (see Figure \ref{fig-uptake_country}D), highlighting the impact of this licensing model across all three data sources.
+In contrast, the United States had a low proportion of hybrid open access enabled by agreements, suggesting that a substantial number of open access articles from US-based authors were likely financed through other means.
 
@@
 
-This study of over 13,000 hybrid journals included in the cOAlition Journal Checker Tool shows a rise in open access due to transformative agreements between 2019 and 2023, although most articles remained paywalled.
+This study of over 13,000 hybrid journals included in the Transformative Agreement Data underlying the cOAlition S Journal Checker Tool shows a rise in open access due to transformative agreements between 2019 and 2023, although most articles remained paywalled.
 While transformative agreements accounted for the majority of open access, many articles continue to become open through the payment of individual publication fees.
@@
 In contrast, the three most productive countries, China, the United States, and India, were ranked much behind other countries in the adoption of transformative agreements.
+Open questions remain whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or countries voluntarily refraining from such agreements.
+
 These trends were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science. 
-An open question remains whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or countries voluntarily refraining from such agreements.
-
-Comparing the data sources examined, the study results show strong correlations by country affiliation, despite differences in journal and article coverage, and in the availability of metadata on corresponding authorship and open access status.
-However, several notable discrepancies warrant discussion.
-They arise from limitations in both open and proprietary databases.
-
-The coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage.
+Notably, the results show strong correlations by country affiliation, aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric], which supports the use of open research information for large-scale analyses of hybrid open access.
+However, the observed differences in coverage of journals, articles, and metadata availability warrant discussion, affecting not only open data sources but also proprietary databases when used in isolation.
+
+To begin with, the coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage.
 Differences can be found for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
-Using the open index Crossref as bibliometric database in hoaddata demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
+Using the open index Crossref as a bibliometric database in hoaddata demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
 
@@
 
-A notable advantage of using Scopus and Web of Science to study the impact of transformative agreements is their comprehensive differentiation of journal articles by document type, as many agreements restrict funding to original articles and reviews.
-This resulted in slightly higher open access rates in Scopus and Web of Science.
-
-A similar limitation of studying open access with less selective databases is the often reported lack of corresponding authorship information [@Fraser_2023; @Haucap_2021; @Shu_2023]. 
-However, indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, in line with previous research, reflecting disciplinary norms in scholarly publishing with regard to contributorship and author roles and positions [@Larivi_re_2016].
+<!-- A notable advantage of using Scopus and Web of Science to study the impact of transformative agreements is their comprehensive differentiation of journal articles by document type, as many agreements restrict funding to original articles and reviews [@Borrego_2023].
+This resulted in slightly higher open access rates in Scopus and Web of Science. -->
+
+One often reported limitation of studying open access with less selective databases is the lack of corresponding authorship information [@Fraser_2023; @Haucap_2021; @Shu_2023]. 
+However, indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, reflecting disciplinary norms in scholarly publishing with regard to contributorship and author roles and positions [@Larivi_re_2016].
 First authors are often considered to have carried out the main research underlying a paper, while the corresponding author is often recognised as having supervised the research [@Mattsson_2010; @Fox_2018].
-It is therefore not surprising that both are strongly correlated, suggesting that first and corresponding authors share the same country affiliation.
-Moreover, in most cases the first author is indeed the same as the corresponding author [@Chinchilla_Rodr_guez_2024].
-
-The slight increase in the number of open access articles by corresponding authors compared to the number of first authors in the last two years observed across Springer Nature hybrid journals raises the question about institutional open access sponsorship practices and if they influence the assignment of author roles among co-authors, especially when funding opportunities are not consistently available within the author teams [@Gumpenberger_2018].
-Previous research notes effects of institutionalised bibliometrics praticses on the valuations of authorship positions [@Helgesson_2020], although, to my knowledge, no such evidence was reported for open access funding yet.
-As such, first authorship can reliably be used in large-scale studies and those focusing on national consortia to assess the impact of transformative agreements.
-
-While OpenAlex’s affiliation coverage is less complete, it shows strong correlations with Scopus [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric]. 
-In this study, this limitation likely reduced the number of articles confidently attributed to transformative agreements, despite high correlations between OpenAlex, Scopus, and Web of Science at the country level.
-Nevertheless, OpenAlex’s native ROR ID integration offers a distinct advantage: unlike Scopus and Web of Science, which require reconciliation of proprietary organisation identifiers with this open standard, it enables more reliable identification of agreement-enabled articles.
+Unsurprisingly, measures based on first or corresponding authorships are strongly correlated, suggesting that these authors share the same country affiliation.
+Moreover, in most cases, the first author is identical the corresponding author [@Chinchilla_Rodr_guez_2024].
+Thus, first authorship can reliably be used in large-scale studies to assess the impact of transformative agreements.
+
+Despite this correlation, the recent slight increase in open access articles by corresponding authors over first authors across Springer Nature hybrid journals prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
+Previous research has highlighted how institutionalized bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], underscoring the need to be cautious of monetary authorship influences [@Maddi_2024].
+
+Affiliation data is another important data element for estimating open access enabled by transformative agreements. 
+Although OpenAlex's affiliation coverage is less comprehensive, which likely reduced the number of articles confidently attributed to transformative agreements in this study, it still shows high correlations with Scopus and Web of Science at the country level.
+However, OpenAlex's native ROR ID integration offers a distinct advantage, enabling more reliable identification of agreement-enabled articles compared to Scopus and Web of Science, which require reconciliation with proprietary organization identifiers.
+Future studies based on the Web of Science will benefit of the recent integration of ROR-IDs, announced by the end of 2024.
 <!-- Dimension is another scholarly data source that uses open standards to identify journals, which enables analysing transformative agreements [@Fraser_2023].-->
 
+Comparing scholarly databases reveals discrepancies in open access evidence.
+Not all publishers deposit Creative Commons (CC) licenses with Crossref, but comparing it with Unpaywall (integrated into Scopus and Web of Science) exposes these gaps, as these discovery services also analyze publishers' websites.
+
 Comparing scholarly databases helps to identify  discrepancies in open access evidence. 
-Not all publishers deposit CC licences with Crossref, but Unpaywall, integrated into Scopus and Web of Science, can make this transparent, because the open access discovery services also analyse publishers' websites.
-However, inconsistencies remain: Web of Science incorrectly identifies Elsevier's delayed open access as hybrid, while Scopus correctly reflects it.
-Unpaywall also missed a substantial number of open access articles with CC license according to Crossref, resulting in fewer open access articles in Scopus and Web of Science, probably due to problems parsing the publisher's website.
-As these differences may affect comparability between databases, research and monitoring exercises should not rely on a single source of evidence.
-Instead, it is recommended that the choice of open access evidence sources should be documented and verified by combining multiple data sources [@Huang_2020].
-It is particularly useful to include expert-curated lists of journals to avoid misclassifications based on a journal's business model [@Visser_2021].
+Crucially, not all publishers deposit CC licences with Crossref, a limitation that becomes apparent when contrasting Crossref with Unpaywall’s data in Scopus and Web of Science. 
+While Unpaywall can detect such gaps by parsing publishers' websites for open access licences, it nevertheless missed a substantial number of CC-licensed open access articles from Wiley journals indexed in Crossref, likely due to parsing errors on the publisher's website.
+This resulted in fewer open access articles being recorded in Scopus and Web of Science.
+Further inconsistencies could be observed between Scopus and Web of Science, despite both relying on Unpaywall: Web of Science erroneously labels Elsevier’s delayed OA as hybrid, whereas Scopus correctly categorises it.
+ONotably, Scopus and Crossref showed greater alignment than observed in a related comparison between Scopus and OpenAlex [@alperin2024analysissuitabilityopenalexbibliometric].
+This likely reflects the use of a dedicated list to identify hybrid open access journals, rather than relying solely on article-level tags.
+Given these variations, which may compromise comparability, research and monitoring exercises should avoid reliance on a single source.
+Instead, the selection of open access evidence should be cross-verified using multiple sources [@Huang_2020].
+Incorporating expert-curated journal lists can also help prevent misclassifications based on business models [@Visser_2021].
 

### 2025-04-03 analysis/manuscript.tex
Proofreading

@@
 hoaddata, developed and maintained by the author to support open access
-monitoring and research (Jahn, 2025), is an R data package that
+monitoring and research (e.g. Jahn (2025)), is an R data package that
 regularly collects information on hybrid open access uptake from
@@
 affilation metadata from OpenAlex (Priem et al., 2022) with
-transformative agreement information from Transformative Agreement Data
-dump used by the cOAlition S Journal Checker Tool (JCT)\footnote{\url{https://journalcheckertool.org/transformative-agreements/}},
+transformative agreement information from the Transformative Agreement
+Data dump used by the cOAlition S Journal Checker Tool (JCT)\footnote{\url{https://journalcheckertool.org/transformative-agreements/}},
 which links journal and institutional data about participating research
@@
 (Birkle et al., 2020). Web of Science is a selective, base
-research-focused database with dedicated panels to choose journals for
-inclusion (Stahlschmidt \& Stephen, 2022; Visser et al., 2021). The
-collections considered in this study were the Science Citation Index
-Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts
-\& Humanities Citation Index (AHCI).
+research-focused database (Stahlschmidt \& Stephen, 2022; Visser et al.,
+2021). The collections considered in this study were the Science
+Citation Index Expanded (SCIE), the Social Sciences Citation Index
+(SSCI) and the Arts \& Humanities Citation Index (AHCI).
 
@@
 Following Jahn (2025), the starting point was a unified dataset of
-several safeguarded JCT snapshots\footnote{\url{https://github.com/njahn82/jct_data}},
-as provided by hoaddata. The JCT journal data in hoaddata were enriched
-with ISSN variants linked to an ISSN-L. To identify hybrid journals, a
-comprehensive exclusion of fully open access journals was performed
-using multiple journal lists (DOAJ, OpenAlex, Bielefeld). These data
-were used to determine the publication volume of hybrid journals for
-each database independently.
+several safeguarded JCT snapshots\footnote{\url{https://github.com/njahn82/jct_data}}.
+The JCT journal data were enriched with ISSN variants linked to an
+ISSN-L. To identify hybrid journals, a comprehensive exclusion of fully
+open access journals was performed using multiple journal lists (DOAJ,
+OpenAlex, Bielefeld). These data were made available via hoaddata and
+used to determine the publication volume of hybrid journals for each
+database independently.
 
 hoaddata relies on Crossref for obtaining journal publication volume and
-open access status through Creative Commons licence information relative
-to the published version (``version of record''). The article metadata
-included DOIs, publication dates, open access information as well as
-author roles and institutional affiliations. Publication years were
-determined using the earliest known date of publication in a journal. In
-hoaddata, this corresponded to Crossref's issued date. For Web of
-Science and Scopus, the earliest publication date was used where
-available, with Scopus dates specifically determined by the KB through
-version tracking of the raw data.
+open access status through Creative Commons (CC) licence information
+relative to the published version (``version of record''). The article
+metadata included DOIs, publication dates, open access information as
+well as author roles and affiliations. Publication years were determined
+using the earliest known date of publication in a journal. In hoaddata,
+this corresponded to Crossref's issued date. For Web of Science and
+Scopus, the earliest publication date was used where available, with
+Scopus dates specifically determined by the KB through version tracking
+of the raw data.
 
@@
 Articles in hybrid journals were considered open access when they were
-made freely available under a Creative Commons license on publishers'
-platforms. While hoaddata sourced this information from Crossref license
-metadata, Web of Science and Scopus relied on Unpaywall as evidence
-source. Unpaywall also uses Crossref license metadata, but supplements
-them by parsing publisher websites directly, addressing cases where
-publishers do not provide machine-readable Creative Commons license
-information (Piwowar et al., 2018). This additional parsing remains
-necessary despite transformative agreement workflows recommend the
-depositing license information during DOI registration (Geschuhn \&
-Stone, 2017). Both Web of Science and Scopus defined hybrid open access
-consistently as content available under Creative Commons licenses on
-publisher platforms according to their
+made freely available under a CC license on publisher platforms. While
+hoaddata obtained this information from Crossref license metadata, Web
+of Science and Scopus relied on Unpaywall as evidence source. Unpaywall
+also uses Crossref license metadata, but supplements them by parsing
+publisher websites directly, addressing cases where publishers do not
+provide machine-readable CC license information (Piwowar et al., 2018).
+This additional parsing remains necessary despite transformative
+agreement workflows recommend to deposit CC licenses during DOI
+registration (Geschuhn \& Stone, 2017). Both Web of Science and Scopus
+defined hybrid open access consistently as content available under CC
+licenses on publisher platforms according to their
 documentations\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}\footnote{\url{https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option}},
 distinguishing it from bronze open access that lack such explicit
-license information, or use publisher-specific licenses.
+license information, or use publisher-specific licenses (Piwowar et al.,
+2018).
 
@@
 
-Because neither Web of Science nor Scopus support ROR IDs, the
-institution identifier used by the JCT, a two-step matching process was
-implemented to harmonise affiliation data. First, 2,782,540 articles
-from 6,457 institutions with ROR-IDs in the JCT data since 2017
-(according to hoaddata) were processed to map first authors' ROR-IDs to
-corresponding proprietary affiliation identifier in Web of Science and
-Scopus using DOI matching. Then, an algorithm selected the most frequent
-ROR ID and proprietary identifier pairs to handle multiple affiliations
-and organisational hierarchy differences.
+Because neither Web of Science nor Scopus support ROR at the time of the
+data retrieval, the institution identifier used by the JCT, a two-step
+matching process was implemented to harmonise affiliation data. First,
+2,782,540 articles from 6,457 institutions with ROR-IDs in the JCT data
+since 2017 (according to hoaddata) were processed to map first authors'
+ROR-IDs to corresponding proprietary affiliation identifier in Web of
+Science and Scopus using DOI matching. Then, an algorithm selected the
+most frequent ROR ID and proprietary identifier pairs to handle multiple
+affiliations and organisational hierarchy differences.
 
@@
 publications, introduced through multiple affiliations of single
-authors. The difference between databases suggests that Scopus's
+authors. The difference between databases suggests that Scopus'
 affiliation control aligns more closely with ROR than that of the Web of
@@
 
-Based on these generated matching tables, articles eligible under
-transformative agreements could also be retrieved from Web of Science
-and Scopus, although they did not contain the ROR IDs used by the JCT.
-The estimation of eligible articles followed Jahn (2025) and included a
-matching of both journals and participating institutions. The matching
-also took into account the duration of agreements according to the ESAC
-registry, with only those matches where an agreement was actually in
-place being considered for subsequent analysis. A related study (Jonge
-et al., 2025), applied to publications funded by the Dutch Research
-Council (NWO) and validated against internal invoicing data, confirmed
-that such matching can accurately identify most articles under
-transformative agreements.
+Based on these so-compiled matching tables, articles eligible under
+transformative agreements could also be obtained from Web of Science and
+Scopus, although they did not contain the ROR IDs used by the JCT. The
+estimation of eligible articles followed Jahn (2025) and included a
+matching of both journals and participating institutions according to
+the JCT. The matching also took into account the duration of agreements
+according to the ESAC registry, with only those matches where an
+agreement was actually in place being considered for subsequent
+analysis. A related study (Jonge et al., 2025), applied to publications
+funded by the Dutch Research Council (NWO) and validated against
+internal invoicing data, confirmed that such matching can accurately
+identify most articles under transformative agreements.
 
@@
 This section first presents the indexing coverage of hybrid open access
-by comparing the data from open data sources with the proprietary
-bibliometric databases Scopus and Web of Science. Then, using the same
-methods, indicators at the publisher and country level are calculated
+by comparing open data sources with the proprietary bibliometric
+databases Scopus and Web of Science. Then, using the same methods,
+indicators at the publisher and country level are calculated
 independently for each database and compared with each other to assess
@@
 
-\caption{\label{fig-upset_coverage_results}Upset graph}
+\caption{\label{fig-upset_coverage_results}Comparison of the hybrid open
+access indexing coverage of hoaddata (based on Crossref), Scopus and Web
+of Science, 2019-2023. Only hybrid journals present in the cOAlition S
+Transformative Agreement Data dump with at least one open access article
+are considered. A) presents the number of journals, B) the number of
+articles (DOI), distinguishing between shared DOI corpus and surplus in
+hoaddata. Box plots (C) shows the five-years journal article volume
+(log-scale).}
 
@@
 \ref{fig-upset_coverage_results}C illustrates the distribution for each
-combination. It shows that there is a large spread across the journals
-covered by all three data sources. Furthermore, journals in this set
-published more on average than journals covered by only one or two data
-sources. In particular, the journals covered exclusively by hoaddata
-were substantially smaller than those covered by all three sources. Upon
-inspection, these were often newly launched hybrid journals, which
+combination. It shows a large spread across the journals jointly covered
+by all three data sources. These journals published more on average than
+journals covered by only one or two data sources. In particular, the
+journals covered exclusively by hoaddata were substantially smaller.
+Upon inspection, these were often newly launched hybrid journals, which
 explains the relatively low five-year publication volume. An example is
@@
 2019 to 2023 with a focus on publisher portfolios. The analysis
-highlights the dominance of the three largest publishers--Elsevier,
-Springer Nature, and Wiley--, which collectively accounted for 47\% of
+highlights the dominance of the three largest publishers, Elsevier,
+Springer Nature, and Wiley, which collectively accounted for 47\% of
 hybrid journals and 62\% of articles published during this five-year
@@
 
-\caption{\label{fig-upset_coverage_results_publisher}Upset graph
-publisher}
+\caption{\label{fig-upset_coverage_results_publisher}Comparison of the
+hybrid open access indexing coverage of hoaddata (based on Crossref),
+Scopus and Web of Science, 2019-2023 by publisher. Only hybrid journals
+present in the cOAlition S Transformative Agreement Data dump with at
+least one open access article are considered. A) presents the percentage
+of journals by publisher, B) the number of articles by publisher (shared
+DOI). Box plots (C) shows the five-years journal article volume
+(log-scale) by publisher.}
 
@@
 
-The three largest publishers--Elsevier, Springer Nature and Wiley--were
+The three largest publishers, Elsevier, Springer Nature, and Wiley, were
 best represented in the exclusive intersection of all three data sources
@@
 with 64,898 articles, and World Scientific recorded 87 journals and
-42,531 articles. In total, 9 publishers were not represented in
-Crossref.
+42,531 articles. In total, 9 publishers did not share CC licenses with
+Crossref and were therefore not represented in hoaddata.
 
@@
 
-\caption{\label{fig-uptake_overview}Open Access indicato}
+\caption{\label{fig-uptake_overview}Development of open access in hybrid
+journals included in transformative agreements, 2019 and 2023 by data
+source and author role. Figure shows the open access percentage (A), the
+number (B) and percentage (C) of open access articles enabled
+transformative agreements.}
 
@@
 in hybrid journals, which is consistent across hoaddata (black line),
-Scopus (orange line), and Web of Science (blue line). According to
-hoaddata, hybrid open access uptake increased from \% (n = ) in 2019 to
-\% (n = ) in 2023. Similarly, Scopus recorded an growth from 6.5\% (n =
-84,648) in 2019 to 19\% (n = 322,850) in 2023. However, Web of Science
-recorded higher open access uptake in early years, before converging in
-2023, from 11\% (n = 137,202) in 2019 to 20\% (n = 255,481) in 2023,
-suggesting an different approach towards labeling hybrid open access by
-the Web of Science.
+Scopus (yellow line), and Web of Science (blue line). According to
+hoaddata, hybrid open access uptake increased from 6.4\% (n = 85,071) in
+2019 to 19\% (n = 302,358) in 2023. Similarly, Scopus recorded an growth
+from 6.5\% (n = 84,648) in 2019 to 19\% (n = 322,850) in 2023. However,
+Web of Science recorded higher open access uptake in early years, before
+converging in 2023, from 11\% (n = 137,202) in 2019 to 20\% (n =
+255,481) in 2023, suggesting an different approach towards labeling
+hybrid open access by the Web of Science.
 
@@
 between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C). Trends
-were consistent for first (dashed line) and corresponding author (dotted
-line) affiliations. According to Scopus, 479,297 open access articles
-could be attributed to transformative agreements based on first author
-metadata (increasing from 15,341 to 206,084) and 489,262 using
-corresponding author metadata (from 15,444 to 210,816). Web of Science
-recorded 493,028 open access articles via transformative agreements
-using first author metadata (increasing from 21,871 to 174,126) and
-500,076 using corresponding author metadata (from 22,092 to 177,030).
-hoaddata, lacking corresponding author data, linked 0 articles to
-transformative agreements but showed slower growth than Scopus and Web
-of Science, from 0 in 2019 to 0 in 2023.
+were consistent when measuring first (dashed line) and corresponding
+author (dotted line) affiliations. According to Scopus, 479,297 open
+access articles could be attributed to transformative agreements based
+on first author metadata (increasing from 15,341 to 206,084) and 489,262
+using corresponding author metadata (from 15,444 to 210,816). Web of
+Science recorded 493,028 open access articles via transformative
+agreements using first author metadata (increasing from 21,871 to
+174,126) and 500,076 using corresponding author metadata (from 22,092 to
+177,030). hoaddata, lacking corresponding author data, linked 0 articles
+to transformative agreements but showed slower growth than Scopus and
+Web of Science, from 0 in 2019 to 0 in 2023.
 
@@
 agreements enabled the majority of hybrid open access. For first
-authors, the share ranged between \% (hoaddata), 64\% (Scopus), and 68
+authors, the share ranged between 61\% (hoaddata), 64\% (Scopus), and 68
 \% (Web of Science) in 2023. For corresponding authors, the shares were
@@
 
-\caption{\label{fig-uptake_publisher}Open Access indicato}
+\caption{\label{fig-uptake_publisher}Development of open access in
+hybrid journals included in transformative agreements, 2019 and 2023, by
+data source, author role and publisher. Figure shows the open access
+percentage (A), the number (B) and percentage (C) of open access
+articles enabled transformative agreements.}
 
@@
 shown for both first and corresponding authors, with full counting used
-to account for multiple country affiliations.
+to account for multiple country affiliations (Hottenrott et al., 2021).
 
@@
 
-\caption{\label{fig-uptake_country}Open Access by country}
+\caption{\label{fig-uptake_country}Comparison of hybrid open access by
+country, 2019-2023. Scatterplots distinguish between proprietary
+database Scopus and Web of Science, and author role. x axis shows
+hoaddata indicators. A) Five-years article volume, B) open access
+percentage in hybrid journals, (A and B limited to countries with
+\textgreater{} 10.000 articles published), C) number and D) percentage
+of open access articles enabled by transformative agreements (limited to
+countries with \textgreater{} 1.000 open access articles enabled by
+transformative agreements). Line represents line of equality. An
+interactive version is accessible here:
+\url{https://najkoja-hoa-replication.hf.space/}}
 
@@
 
-A closer examination of the proportion of hybrid open access enabled by
-transformative agreements (see Figure \ref{fig-uptake_country}D)
-revealed that the French Elsevier agreement on delayed open access was
-not included in JCT data, as it did not qualify as a transformative
-agreement. Medium-sized European countries again showed a high
-proportion of hybrid open access through transformative agreements,
-highlighting the impact of this licensing model across all three data
-sources. In contrast, the United States had a low proportion of hybrid
-open access enabled by agreements, suggesting that a substantial number
-of open access articles were likely financed through other means, such
-as article processing charges (APCs).
+Medium-sized European countries again showed a high proportion of hybrid
+open access through transformative agreements (see Figure
+\ref{fig-uptake_country}D), highlighting the impact of this licensing
+model across all three data sources. In contrast, the United States had
+a low proportion of hybrid open access enabled by agreements, suggesting
+that a substantial number of open access articles from US-based authors
+were likely financed through other means.
 
@@
 
-This study of over 13,000 hybrid journals included in the cOAlition
-Journal Checker Tool shows a rise in open access due to transformative
-agreements between 2019 and 2023, although most articles remained
-paywalled. While transformative agreements accounted for the majority of
-open access, many articles continue to become open through the payment
-of individual publication fees. Hybrid open access and transformative
-agreements remain concentrated among a small group of large commercial
-publishers, with high-income European countries---alongside South
-Africa---showing high adoption rates. In contrast, the three most
-productive countries, China, the United States, and India, were ranked
-much behind other countries in the adoption of transformative
-agreements. These trends were consistent across the investigated open
-data source hoaddata, derived from Crossref and OpenAlex, and the
-established proprietary bibliometric databases Scopus and Web of
-Science. An open question remains whether this uneven distribution
-reflects temporary implementation gaps, inherent inequities in the
-transformative agreement model, or countries voluntarily refraining from
-such agreements.
-
-Comparing the data sources examined, the study results show strong
-correlations by country affiliation, despite differences in journal and
-article coverage, and in the availability of metadata on corresponding
-authorship and open access status. However, several notable
-discrepancies warrant discussion. They arise from limitations in both
-open and proprietary databases.
-
-The coverage analysis reveals that hybrid journals are well indexed in
-all three data sources, particularly in terms of article coverage.
-Differences can be found for journals targeting practitioners or local
-non-English language communities, with many such titles indexed
-exclusively in Crossref and Scopus. Using the open index Crossref as
+This study of over 13,000 hybrid journals included in the Transformative
+Agreement Data underlying the cOAlition S Journal Checker Tool shows a
+rise in open access due to transformative agreements between 2019 and
+2023, although most articles remained paywalled. While transformative
+agreements accounted for the majority of open access, many articles
+continue to become open through the payment of individual publication
+fees. Hybrid open access and transformative agreements remain
+concentrated among a small group of large commercial publishers, with
+high-income European countries---alongside South Africa---showing high
+adoption rates. In contrast, the three most productive countries, China,
+the United States, and India, were ranked much behind other countries in
+the adoption of transformative agreements. Open questions remain whether
+this uneven distribution reflects temporary implementation gaps,
+inherent inequities in the transformative agreement model, or countries
+voluntarily refraining from such agreements.
+
+These trends were consistent across the investigated open data source
+hoaddata, derived from Crossref and OpenAlex, and the established
+proprietary bibliometric databases Scopus and Web of Science. Notably,
+the results show strong correlations by country affiliation, aligning
+with previous studies (Akbaritabar et al., 2024; Alperin et al., 2024),
+which supports the use of open research information for large-scale
+analyses of hybrid open access. However, the observed differences in
+coverage of journals, articles, and metadata availability warrant
+discussion, affecting not only open data sources but also proprietary
+databases when used in isolation.
+
+To begin with, the coverage analysis reveals that hybrid journals are
+well indexed in all three data sources, particularly in terms of article
+coverage. Differences can be found for journals targeting practitioners
+or local non-English language communities, with many such titles indexed
+exclusively in Crossref and Scopus. Using the open index Crossref as a
 bibliometric database in hoaddata demonstrated particular strength in
@@
 
-A notable advantage of using Scopus and Web of Science to study the
-impact of transformative agreements is their comprehensive
-differentiation of journal articles by document type, as many agreements
-restrict funding to original articles and reviews. This resulted in
-slightly higher open access rates in Scopus and Web of Science.
-
-A similar limitation of studying open access with less selective
-databases is the often reported lack of corresponding authorship
-information (Fraser et al., 2023; Haucap et al., 2021; Shu \& Larivière,
-2023). However, indicators based on first authors, which have often been
-used as a proxy for determining open access funding, and corresponding
-authors show a high level of correlation, in line with previous
-research, reflecting disciplinary norms in scholarly publishing with
-regard to contributorship and author roles and positions (Larivière et
-al., 2016). First authors are often considered to have carried out the
-main research underlying a paper, while the corresponding author is
-often recognised as having supervised the research (Fox et al., 2018;
-Mattsson et al., 2010). It is therefore not surprising that both are
-strongly correlated, suggesting that first and corresponding authors
-share the same country affiliation. Moreover, in most cases the first
-author is indeed the same as the corresponding author
-(Chinchilla-Rodríguez et al., 2024).
-
-The slight increase in the number of open access articles by
-corresponding authors compared to the number of first authors in the
-last two years observed across Springer Nature hybrid journals raises
-the question about institutional open access sponsorship practices and
-if they influence the assignment of author roles among co-authors,
-especially when funding opportunities are not consistently available
-within the author teams (Gumpenberger et al., 2018). Previous research
-notes effects of institutionalised bibliometrics praticses on the
-valuations of authorship positions (Helgesson, 2020), although, to my
-knowledge, no such evidence was reported for open access funding yet. As
-such, first authorship can reliably be used in large-scale studies and
-those focusing on national consortia to assess the impact of
-transformative agreements.
-
-While OpenAlex's affiliation coverage is less complete, it shows strong
-correlations with Scopus (Akbaritabar et al., 2024; Alperin et al.,
-2024). In this study, this limitation likely reduced the number of
-articles confidently attributed to transformative agreements, despite
-high correlations between OpenAlex, Scopus, and Web of Science at the
-country level. Nevertheless, OpenAlex's native ROR ID integration offers
-a distinct advantage: unlike Scopus and Web of Science, which require
-reconciliation of proprietary organisation identifiers with this open
-standard, it enables more reliable identification of agreement-enabled
-articles.
+One often reported limitation of studying open access with less
+selective databases is the lack of corresponding authorship information
+(Fraser et al., 2023; Haucap et al., 2021; Shu \& Larivière, 2023).
+However, indicators based on first authors, which have often been used
+as a proxy for determining open access funding, and corresponding
+authors show a high level of correlation, reflecting disciplinary norms
+in scholarly publishing with regard to contributorship and author roles
+and positions (Larivière et al., 2016). First authors are often
+considered to have carried out the main research underlying a paper,
+while the corresponding author is often recognised as having supervised
+the research (Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly,
+measures based on first or corresponding authorships are strongly
+correlated, suggesting that these authors share the same country
+affiliation. Moreover, in most cases, the first author is identical the
+corresponding author (Chinchilla-Rodríguez et al., 2024). Thus, first
+authorship can reliably be used in large-scale studies to assess the
+impact of transformative agreements.
+
+Despite this correlation, the recent slight increase in open access
+articles by corresponding authors over first authors across Springer
+Nature hybrid journals prompts questions about how institutional open
+access sponsorship practices influence author roles and assignments
+within co-author teams, especially as funding opportunities vary
+(Gumpenberger et al., 2018). Previous research has highlighted how
+institutionalized bibliometric practices can affect the valuation of
+authorship positions (Helgesson, 2020), underscoring the need to be
+cautious of monetary authorship influences (Maddi \& Silva, 2024).
+
+Affiliation data is another important data element for estimating open
+access enabled by transformative agreements. Although OpenAlex's
+affiliation coverage is less comprehensive, which likely reduced the
+number of articles confidently attributed to transformative agreements
+in this study, it still shows high correlations with Scopus and Web of
+Science at the country level. However, OpenAlex's native ROR ID
+integration offers a distinct advantage, enabling more reliable
+identification of agreement-enabled articles compared to Scopus and Web
+of Science, which require reconciliation with proprietary organization
+identifiers. Future studies based on the Web of Science will benefit of
+the recent integration of ROR-IDs, announced by the end of 2024.
+
+Comparing scholarly databases reveals discrepancies in open access
+evidence. Not all publishers deposit Creative Commons (CC) licenses with
+Crossref, but comparing it with Unpaywall (integrated into Scopus and
+Web of Science) exposes these gaps, as these discovery services also
+analyze publishers' websites.
 
 Comparing scholarly databases helps to identify discrepancies in open
-access evidence. Not all publishers deposit CC licences with Crossref,
-but Unpaywall, integrated into Scopus and Web of Science, can make this
-transparent, because the open access discovery services also analyse
-publishers' websites. However, inconsistencies remain: Web of Science
-incorrectly identifies Elsevier's delayed open access as hybrid, while
-Scopus correctly reflects it. Unpaywall also missed a substantial number
-of open access articles with CC license according to Crossref, resulting
-in fewer open access articles in Scopus and Web of Science, probably due
-to problems parsing the publisher's website. As these differences may
-affect comparability between databases, research and monitoring
-exercises should not rely on a single source of evidence. Instead, it is
-recommended that the choice of open access evidence sources should be
-documented and verified by combining multiple data sources (Huang et
-al., 2020). It is particularly useful to include expert-curated lists of
-journals to avoid misclassifications based on a journal's business model
-(Visser et al., 2021).
+access evidence. Crucially, not all publishers deposit CC licences with
+Crossref, a limitation that becomes apparent when contrasting Crossref
+with Unpaywall's data in Scopus and Web of Science. While Unpaywall can
+detect such gaps by parsing publishers' websites for open access
+licences, it nevertheless missed a substantial number of CC-licensed
+open access articles from Wiley journals indexed in Crossref, likely due
+to parsing errors on the publisher's website. This resulted in fewer
+open access articles being recorded in Scopus and Web of Science.
+Further inconsistencies could be observed between Scopus and Web of
+Science, despite both relying on Unpaywall: Web of Science erroneously
+labels Elsevier's delayed OA as hybrid, whereas Scopus correctly
+categorises it. ONotably, Scopus and Crossref showed greater alignment
+than observed in a related comparison between Scopus and OpenAlex
+(Alperin et al., 2024). This likely reflects the use of a dedicated list
+to identify hybrid open access journals, rather than relying solely on
+article-level tags. Given these variations, which may compromise
+comparability, research and monitoring exercises should avoid reliance
+on a single source. Instead, the selection of open access evidence
+should be cross-verified using multiple sources (Huang et al., 2020).
+Incorporating expert-curated journal lists can also help prevent
+misclassifications based on business models (Visser et al., 2021).
 
@@
 
+\bibitem[\citeproctext]{ref-Hottenrott_2021}
+Hottenrott, H., Rose, M. E., \& Lawson, C. (2021). The rise of multiple
+institutional affiliations in academia. \emph{Journal of the Association
+for Information Science and Technology}, \emph{72}(8), 1039--1058.
+\url{https://doi.org/10.1002/asi.24472}
+
 \bibitem[\citeproctext]{ref-Huang_2020}
@@
 
+\bibitem[\citeproctext]{ref-Maddi_2024}
+Maddi, A., \& Silva, J. A. T. da. (2024). Beyond authorship: Analyzing
+contributions in PLOS ONE and the challenges of appropriate attribution.
+\emph{Journal of Data and Information Science}, \emph{9}(3), 88--115.
+\url{https://doi.org/10.2478/jdis-2024-0015}
+
 \bibitem[\citeproctext]{ref-Maisano_2025}

### 2025-04-04 analysis/manuscript.qmd
Rough intro outline

@@
 
-This study aims to demonstrate the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
-To achieve this, a replication study was conducted by comparing results from hoaddata, an openly available and continuously updated dataset on hybrid open access uptake based on Crossref, OpenAlex, and the cOAlition S Journal Checker Tool, with the established bibliometric databases Web of Science and Scopus.
-
-This study focuses on the coverage of hybrid journal portfolios included in transformative agreements between 2019 and 2023.
-Special attention is given to potential differences in open access uptake by country when comparing first-author affiliation data to corresponding authorships.
-This is crucial because the lack of publicly available invoicing data corresponding to authorships plays an essential role in determining whether an open-access article is supported through transformative agreements.
-Data on corresponding authorships have been available on the Web of Science and Scopus for much longer than in open databases such as OpenAlex, where this information is still being roled out at the time of writing.
-Because of this weakness, open approaches such as hoaddata and related research use first-authorship data instead.
-
-By conducting a large-scale comparative analysis, this study aims to
-
-1. Determine the strengths and weaknesses of using open data sources in monitoring the impact of transformative agreements on hybrid open access publishing.
-2. Assess the coverage and accuracy of open data sources compared with established bibliometric databases.
-3. Evaluate the reliability of first author affiliation data as a proxy for corresponding authorship in the context of open access uptake analysis.
-
-# Background -- Evidence base to measure the effects of transformative agreements
-
-## Anforderungen an das Monitoring
-
-- esac guidelines 
-- gemeinsamkeiten und unterschiede zu apc (listenpreise, tatsächliche zahlungen, zentrales invoicing, rabatte, waivers) 
-- insitutionen covern cas, jedoch kann es zu unterschiedlichen verrechnugnsformen führen (antielig mit förderer, splitting innerhaklbd er einrichtung)
-
-## Bibliometrische Evidenzen
-
-- allgmeeiner uptake 
-- wachstum apcs 
-- wachstum verträge (konsortien, forschung) 
-- konsequenzen
+Transformative agreements have emerged as a much-discussed licensing model for the transition from subscription-based publishing to full open access in the last ten years. 
+While these agreements can vary substantially, increasing coordination exists in standardising the negotiation, administration, and evaluation of these agreements. 
+Simultaneously, efforts to improve transparency through open data about contracts and payments are growing. 
+However, this information remains only partially available through bibliometric databases, limiting large-scale studies about the impact of transformative agreements on the transition of hybrid journals to full open access through increased open access uptake.
+This study compares findings from hoaddata, an openly available,  continuously updated dataset on hybrid open access, built on open data from Crossref, OpenAlex and the cOAlition S Journal Checler Tool, with the established bibliometric databases Web of Science and Scopus.
+The purpose of this comparison is not only to evaluate the suitability of open metadata for measuring hybrid open access but also to inform bibliometric research and practice by highlighting the strengths and weaknesses of each data source.
+
+Since their introduction, transformative agreements have sparked considerable academic debate. 
+Whilst standardisation activities, particularly through the OA2020 and ESAC (Efficiency and Standards for Article Charges) initiatives, have led to improved workflows between library consortia and publishers, as well as increased open access publishing, the initial focus on a few large publishers by national consortia from high-income countries has drawn criticism.
+Critics have raised concerns about market concentration, the lack of journal flipping to full open access models, reduced competition in the publishing market, and wider gaps between well-resourced and under-resourced institutions and researchers.
+
+For instance, studying over 700 agreements, Jahn (2025) observed a strong growth in open access publishing to 15% in 2022 due to the varying implementation of transformative agreements across countries, predominantly driven by three commercial publishers: Elsevier, Springer Nature, and Wiley.
+This aligns with the rapid proliferation of transformative agreements in recent years.
+Previously, only a minimal proportion of articles were made open access in hybrid journals (Piwowar, 2018; Laakso, 2016).
+
+Evaluations conducted by library consortia confirm these trends but draw different conclusions.
+For instance, the British Joint Information Systems Committee's evaluation (Brayman, Devenney et al., 2024) noted transformative agreements' significant impact on national open access growth but limited effect on global open access transition, suggesting a need to reevaluate their strategy.
+Similar policy reconsiderations emerged in Norway and Sweden (Holden, Skoie et al., 2023; Widding, 2024).
+Meanwhile, cOAlition S recommended ending financial support for these agreements in 2024, but still considers hybrid open access compliant with funder open access policies.
+Conversely, Germany's DEAL consortium extended agreements with major publishers through 2028, while Colombia established Latin America's first transformative agreements (Muñoz-Vélez, Pallares et al., 2024).
+A recent Max Planck Digital Library study demonstrates how cost savings can be achieved through this model and how it can be further developed to achieve full open access while controlling expenditure.
+
+Open databases like OpenAlex, which build upon Crossref, particularly benefit from these developments.
+These resources have gained considerable attention in research evaluation due to their inclusivity, openness, and comprehensive coverage of publication metadata.
+Recent studies [REFERENCE NEEDED] have demonstrated that open bibliometric databases increasingly match or exceed the coverage of traditional commercial databases in many disciplines, whilst providing improved access to standardised and interlinked open research information.
+
+Against this background, this study evaluates the suitability of open metadata to measure the impact of transformative agreements on hybrid open access.
+Only a few studies have attempted this analysis, primarily due to data limitations and the need to combine multiple sources.
+Transformative agreements present unique measurement challenges because of their complex workflow requirements, which are not fully reflected in bibliometric databases.
+These challenges include varying journal portfolios across agreements, invoicing procedures specific to submitting authors, and article caps that limit open access publishing within agreements.
+
+The present study addresses these gaps by replicating and applying consistent methodologies across Web of Science, Scopus, and open data sources.
+This comparative approach enables the research to overcome the limitations of any single source whilst systematically assessing the strengths and weaknesses of different bibliometric data sources for monitoring transformative agreement outcomes.
+By triangulating findings across these platforms, the study aims to provide a more comprehensive understanding of how transformative agreements influence hybrid open access publishing patterns within the ongoing scholarly publishing transition.
 
@@
 
-This study of over 13,000 hybrid journals included in the Transformative Agreement Data underlying the cOAlition S Journal Checker Tool shows a rise in open access due to transformative agreements between 2019 and 2023, although most articles remained paywalled.
+This study of over 13,000 hybrid journals shows a rise in open access due to transformative agreements between 2019 and 2023, although most articles remained paywalled.
 While transformative agreements accounted for the majority of open access, many articles continue to become open through the payment of individual publication fees.
@@
 with high-income European countries—alongside South Africa—showing high adoption rates.
-In contrast, the three most productive countries, China, the United States, and India, were ranked much behind other countries in the adoption of transformative agreements.
-Open questions remain whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or countries voluntarily refraining from such agreements.
-
-These trends were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science. 
-Notably, the results show strong correlations by country affiliation, aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric], which supports the use of open research information for large-scale analyses of hybrid open access.
-However, the observed differences in coverage of journals, articles, and metadata availability warrant discussion, affecting not only open data sources but also proprietary databases when used in isolation.
-
-To begin with, the coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage.
-Differences can be found for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
+In contrast, the three most productive countries, China, the United States, and India, were substantially ranked lower in transformative agreement adoption.
+Open questions remain whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or countries deliberately refraining from such agreements.
+
+The findings were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science. 
+Notably, the results show strong correlations by country affiliation, aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73], which supports using open metadata for large-scale analyses of hybrid open access.
+However, the observed differences in journal coverage and metadata availability warrant discussion, affecting not only open data sources but also proprietary databases when used in isolation.
+
+The coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage.
+Differences emerge for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
 Using the open index Crossref as a bibliometric database in hoaddata demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
 
-The situation of hybrid journal publishing is therefore different from that of fully open access journals.
+The landscape of hybrid journal publishing differs markedly from that of fully open access journals.
 Comparing the coverage of OpenAlex, Scopus and Web of Science, @Simard_2024 indicate that only half of the fully open access journals listed in the DOAJ are also indexed in Scopus and Web of Science.
@@
 
-One often reported limitation of studying open access with less selective databases is the lack of corresponding authorship information [@Fraser_2023; @Haucap_2021; @Shu_2023]. 
-However, indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, reflecting disciplinary norms in scholarly publishing with regard to contributorship and author roles and positions [@Larivi_re_2016].
-First authors are often considered to have carried out the main research underlying a paper, while the corresponding author is often recognised as having supervised the research [@Mattsson_2010; @Fox_2018].
+A frequently reported limitation of studying open access with less selective databases is the lack of corresponding authorship information [@Fraser_2023; @Haucap_2021; @Shu_2023]. 
+However, this analysis demonstrates that indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, reflecting disciplinary norms in scholarly publishing with regard to contributorship and author roles and positions [@Larivi_re_2016].
+First authors typically conduct the main research underlying a paper, while the corresponding author often supervises the research [@Mattsson_2010; @Fox_2018]. 
 Unsurprisingly, measures based on first or corresponding authorships are strongly correlated, suggesting that these authors share the same country affiliation.
-Moreover, in most cases, the first author is identical the corresponding author [@Chinchilla_Rodr_guez_2024].
+Moreover, in most cases, the first author is identical to the corresponding author [@Chinchilla_Rodr_guez_2024].
 Thus, first authorship can reliably be used in large-scale studies to assess the impact of transformative agreements.
 
-Despite this correlation, the recent slight increase in open access articles by corresponding authors over first authors across Springer Nature hybrid journals prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
-Previous research has highlighted how institutionalized bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], underscoring the need to be cautious of monetary authorship influences [@Maddi_2024].
-
-Affiliation data is another important data element for estimating open access enabled by transformative agreements. 
+Despite this correlation, the study observed a recent slight increase in open access articles by corresponding authors over first authors across Springer Nature hybrid journals prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
+Previous research has highlighted how institutionalized bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], suggesting the need to monitor potential influences on authorships [@Maddi_2024].
+
+Another critical data element in the study is affiliation data, which is essential for estimating open access enabled by transformative agreements. 
 Although OpenAlex's affiliation coverage is less comprehensive, which likely reduced the number of articles confidently attributed to transformative agreements in this study, it still shows high correlations with Scopus and Web of Science at the country level.
 However, OpenAlex's native ROR ID integration offers a distinct advantage, enabling more reliable identification of agreement-enabled articles compared to Scopus and Web of Science, which require reconciliation with proprietary organization identifiers.
-Future studies based on the Web of Science will benefit of the recent integration of ROR-IDs, announced by the end of 2024.
+Future studies based on the Web of Science will benefit from the recent integration of ROR-IDs, announced by the end of 2024.
 <!-- Dimension is another scholarly data source that uses open standards to identify journals, which enables analysing transformative agreements [@Fraser_2023].-->
 
-Comparing scholarly databases reveals discrepancies in open access evidence.
-Not all publishers deposit Creative Commons (CC) licenses with Crossref, but comparing it with Unpaywall (integrated into Scopus and Web of Science) exposes these gaps, as these discovery services also analyze publishers' websites.
-
-Comparing scholarly databases helps to identify  discrepancies in open access evidence. 
+The database comparison revealed important discrepancies in open access evidence. 
 Crucially, not all publishers deposit CC licences with Crossref, a limitation that becomes apparent when contrasting Crossref with Unpaywall’s data in Scopus and Web of Science. 
@@
 This resulted in fewer open access articles being recorded in Scopus and Web of Science.
-Further inconsistencies could be observed between Scopus and Web of Science, despite both relying on Unpaywall: Web of Science erroneously labels Elsevier’s delayed OA as hybrid, whereas Scopus correctly categorises it.
-ONotably, Scopus and Crossref showed greater alignment than observed in a related comparison between Scopus and OpenAlex [@alperin2024analysissuitabilityopenalexbibliometric].
-This likely reflects the use of a dedicated list to identify hybrid open access journals, rather than relying solely on article-level tags.
+Further inconsistencies emerged between Scopus and Web of Science, despite both relying on Unpaywall: Web of Science erroneously labels Elsevier’s delayed open access as hybrid, whereas Scopus correctly categorises it.
+Notably, Scopus and Crossref showed greater alignment than observed in a related comparison between Scopus and OpenAlex [@alperin2024analysissuitabilityopenalexbibliometric].
+This likely reflects the use of a curated list to identify hybrid open access journals, rather than relying solely on article-level tags.
 Given these variations, which may compromise comparability, research and monitoring exercises should avoid reliance on a single source.
@@
 
-It is important to note that articles from institutions involved in transformative agreements are estimates due to a lack of access to invoice data, which is not usually shared between library consortia and publishers.
-Furthermore, it is important to acknowledge that the sources of transformative agreements and the underlying ESAC registry are a voluntary effort, crowd-sourced from various consortia. 
-However, recent validation with Dutch Research Information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements. 
+It is important to note that the study's estimates of articles from institutions involved in transformative agreements are approximations due to a lack of access to invoice data, which is not usually shared by library consortia and publishers.
+Furthermore, the data sources about transformative agreements including the cOAlition S Journal Checker Tool and the ESAC registry are a voluntary effort, crowd-sourced from various consortia. 
+However, recent validation with Dutch Research Information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements [@de_Jonge_2025]. 
 With enhanced and streamlined metadata compliance, attributable to evolving standards and initiatives to support negotiations with publishers, particularly through the communities of the ESAC initiative and Barcelona Declaration supporters, the situation is likely to improve.
-In conclusion, this cross-database analysis has highlighted the potential and challenges of establishing open scholarly data as a suitable evidence base for analysing the transition from subscription-based publishing to open access.
+In conclusion, this study has shown that open scholarly data are suitable for analysing the transition from subscription-based publishing to full open access.
 

### 2025-04-04 analysis/manuscript.tex
Rough intro outline

@@
 
-This study aims to demonstrate the suitability of open scholarly data
-sources for assessing the impact of transformative agreements on hybrid
-open access. To achieve this, a replication study was conducted by
-comparing results from hoaddata, an openly available and continuously
-updated dataset on hybrid open access uptake based on Crossref,
-OpenAlex, and the cOAlition S Journal Checker Tool, with the established
-bibliometric databases Web of Science and Scopus.
-
-This study focuses on the coverage of hybrid journal portfolios included
-in transformative agreements between 2019 and 2023. Special attention is
-given to potential differences in open access uptake by country when
-comparing first-author affiliation data to corresponding authorships.
-This is crucial because the lack of publicly available invoicing data
-corresponding to authorships plays an essential role in determining
-whether an open-access article is supported through transformative
-agreements. Data on corresponding authorships have been available on the
-Web of Science and Scopus for much longer than in open databases such as
-OpenAlex, where this information is still being roled out at the time of
-writing. Because of this weakness, open approaches such as hoaddata and
-related research use first-authorship data instead.
-
-By conducting a large-scale comparative analysis, this study aims to
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Determine the strengths and weaknesses of using open data sources in
-  monitoring the impact of transformative agreements on hybrid open
-  access publishing.
-\item
-  Assess the coverage and accuracy of open data sources compared with
-  established bibliometric databases.
-\item
-  Evaluate the reliability of first author affiliation data as a proxy
-  for corresponding authorship in the context of open access uptake
-  analysis.
-\end{enumerate}
-
-\section{Background -- Evidence base to measure the effects of
-transformative
-agreements}\label{background-evidence-base-to-measure-the-effects-of-transformative-agreements}
-
-\subsection{Anforderungen an das
-Monitoring}\label{anforderungen-an-das-monitoring}
-
-\begin{itemize}
-\tightlist
-\item
-  esac guidelines
-\item
-  gemeinsamkeiten und unterschiede zu apc (listenpreise, tatsächliche
-  zahlungen, zentrales invoicing, rabatte, waivers)
-\item
-  insitutionen covern cas, jedoch kann es zu unterschiedlichen
-  verrechnugnsformen führen (antielig mit förderer, splitting
-  innerhaklbd er einrichtung)
-\end{itemize}
-
-\subsection{Bibliometrische Evidenzen}\label{bibliometrische-evidenzen}
-
-\begin{itemize}
-\tightlist
-\item
-  allgmeeiner uptake
-\item
-  wachstum apcs
-\item
-  wachstum verträge (konsortien, forschung)
-\item
-  konsequenzen
-\end{itemize}
+Transformative agreements have emerged as a much-discussed licensing
+model for the transition from subscription-based publishing to full open
+access in the last ten years. While these agreements can vary
+substantially, increasing coordination exists in standardising the
+negotiation, administration, and evaluation of these agreements.
+Simultaneously, efforts to improve transparency through open data about
+contracts and payments are growing. However, this information remains
+only partially available through bibliometric databases, limiting
+large-scale studies about the impact of transformative agreements on the
+transition of hybrid journals to full open access through increased open
+access uptake. This study compares findings from hoaddata, an openly
+available, continuously updated dataset on hybrid open access, built on
+open data from Crossref, OpenAlex and the cOAlition S Journal Checler
+Tool, with the established bibliometric databases Web of Science and
+Scopus. The purpose of this comparison is not only to evaluate the
+suitability of open metadata for measuring hybrid open access but also
+to inform bibliometric research and practice by highlighting the
+strengths and weaknesses of each data source.
+
+Since their introduction, transformative agreements have sparked
+considerable academic debate. Whilst standardisation activities,
+particularly through the OA2020 and ESAC (Efficiency and Standards for
+Article Charges) initiatives, have led to improved workflows between
+library consortia and publishers, as well as increased open access
+publishing, the initial focus on a few large publishers by national
+consortia from high-income countries has drawn criticism. Critics have
+raised concerns about market concentration, the lack of journal flipping
+to full open access models, reduced competition in the publishing
+market, and wider gaps between well-resourced and under-resourced
+institutions and researchers.
+
+For instance, studying over 700 agreements, Jahn (2025) observed a
+strong growth in open access publishing to 15\% in 2022 due to the
+varying implementation of transformative agreements across countries,
+predominantly driven by three commercial publishers: Elsevier, Springer
+Nature, and Wiley. This aligns with the rapid proliferation of
+transformative agreements in recent years. Previously, only a minimal
+proportion of articles were made open access in hybrid journals
+(Piwowar, 2018; Laakso, 2016).
+
+Evaluations conducted by library consortia confirm these trends but draw
+different conclusions. For instance, the British Joint Information
+Systems Committee's evaluation (Brayman, Devenney et al., 2024) noted
+transformative agreements' significant impact on national open access
+growth but limited effect on global open access transition, suggesting a
+need to reevaluate their strategy. Similar policy reconsiderations
+emerged in Norway and Sweden (Holden, Skoie et al., 2023; Widding,
+2024). Meanwhile, cOAlition S recommended ending financial support for
+these agreements in 2024, but still considers hybrid open access
+compliant with funder open access policies. Conversely, Germany's DEAL
+consortium extended agreements with major publishers through 2028, while
+Colombia established Latin America's first transformative agreements
+(Muñoz-Vélez, Pallares et al., 2024). A recent Max Planck Digital
+Library study demonstrates how cost savings can be achieved through this
+model and how it can be further developed to achieve full open access
+while controlling expenditure.
+
+Open databases like OpenAlex, which build upon Crossref, particularly
+benefit from these developments. These resources have gained
+considerable attention in research evaluation due to their inclusivity,
+openness, and comprehensive coverage of publication metadata. Recent
+studies {[}REFERENCE NEEDED{]} have demonstrated that open bibliometric
+databases increasingly match or exceed the coverage of traditional
+commercial databases in many disciplines, whilst providing improved
+access to standardised and interlinked open research information.
+
+Against this background, this study evaluates the suitability of open
+metadata to measure the impact of transformative agreements on hybrid
+open access. Only a few studies have attempted this analysis, primarily
+due to data limitations and the need to combine multiple sources.
+Transformative agreements present unique measurement challenges because
+of their complex workflow requirements, which are not fully reflected in
+bibliometric databases. These challenges include varying journal
+portfolios across agreements, invoicing procedures specific to
+submitting authors, and article caps that limit open access publishing
+within agreements.
+
+The present study addresses these gaps by replicating and applying
+consistent methodologies across Web of Science, Scopus, and open data
+sources. This comparative approach enables the research to overcome the
+limitations of any single source whilst systematically assessing the
+strengths and weaknesses of different bibliometric data sources for
+monitoring transformative agreement outcomes. By triangulating findings
+across these platforms, the study aims to provide a more comprehensive
+understanding of how transformative agreements influence hybrid open
+access publishing patterns within the ongoing scholarly publishing
+transition.
 
@@
 
-This study of over 13,000 hybrid journals included in the Transformative
-Agreement Data underlying the cOAlition S Journal Checker Tool shows a
-rise in open access due to transformative agreements between 2019 and
-2023, although most articles remained paywalled. While transformative
-agreements accounted for the majority of open access, many articles
-continue to become open through the payment of individual publication
-fees. Hybrid open access and transformative agreements remain
-concentrated among a small group of large commercial publishers, with
-high-income European countries---alongside South Africa---showing high
-adoption rates. In contrast, the three most productive countries, China,
-the United States, and India, were ranked much behind other countries in
-the adoption of transformative agreements. Open questions remain whether
-this uneven distribution reflects temporary implementation gaps,
-inherent inequities in the transformative agreement model, or countries
-voluntarily refraining from such agreements.
-
-These trends were consistent across the investigated open data source
+This study of over 13,000 hybrid journals shows a rise in open access
+due to transformative agreements between 2019 and 2023, although most
+articles remained paywalled. While transformative agreements accounted
+for the majority of open access, many articles continue to become open
+through the payment of individual publication fees. Hybrid open access
+and transformative agreements remain concentrated among a small group of
+large commercial publishers, with high-income European
+countries---alongside South Africa---showing high adoption rates. In
+contrast, the three most productive countries, China, the United States,
+and India, were substantially ranked lower in transformative agreement
+adoption. Open questions remain whether this uneven distribution
+reflects temporary implementation gaps, inherent inequities in the
+transformative agreement model, or countries deliberately refraining
+from such agreements.
+
+The findings were consistent across the investigated open data source
 hoaddata, derived from Crossref and OpenAlex, and the established
@@
 the results show strong correlations by country affiliation, aligning
-with previous studies (Akbaritabar et al., 2024; Alperin et al., 2024),
-which supports the use of open research information for large-scale
-analyses of hybrid open access. However, the observed differences in
-coverage of journals, articles, and metadata availability warrant
+with previous studies (Akbaritabar et al., 2024; Alperin et al., 2024)
+and rankings (Eck et al., 2024), which supports using open metadata for
+large-scale analyses of hybrid open access. However, the observed
+differences in journal coverage and metadata availability warrant
 discussion, affecting not only open data sources but also proprietary
@@
 
-To begin with, the coverage analysis reveals that hybrid journals are
-well indexed in all three data sources, particularly in terms of article
-coverage. Differences can be found for journals targeting practitioners
-or local non-English language communities, with many such titles indexed
+The coverage analysis reveals that hybrid journals are well indexed in
+all three data sources, particularly in terms of article coverage.
+Differences emerge for journals targeting practitioners or local
+non-English language communities, with many such titles indexed
 exclusively in Crossref and Scopus. Using the open index Crossref as a
@@
 
-The situation of hybrid journal publishing is therefore different from
-that of fully open access journals. Comparing the coverage of OpenAlex,
-Scopus and Web of Science, Simard et al. (2024) indicate that only half
-of the fully open access journals listed in the DOAJ are also indexed in
-Scopus and Web of Science. Notably, journals that charge no publication
-fees (``diamond journals'') are absent from the selective Web of
-Science, which reinforces existing disparities in the indexing of
+The landscape of hybrid journal publishing differs markedly from that of
+fully open access journals. Comparing the coverage of OpenAlex, Scopus
+and Web of Science, Simard et al. (2024) indicate that only half of the
+fully open access journals listed in the DOAJ are also indexed in Scopus
+and Web of Science. Notably, journals that charge no publication fees
+(``diamond journals'') are absent from the selective Web of Science,
+which reinforces existing disparities in the indexing of
 underrepresented research communities and regions in selective
@@
 
-One often reported limitation of studying open access with less
+A frequently reported limitation of studying open access with less
 selective databases is the lack of corresponding authorship information
 (Fraser et al., 2023; Haucap et al., 2021; Shu \& Larivière, 2023).
-However, indicators based on first authors, which have often been used
-as a proxy for determining open access funding, and corresponding
-authors show a high level of correlation, reflecting disciplinary norms
-in scholarly publishing with regard to contributorship and author roles
-and positions (Larivière et al., 2016). First authors are often
-considered to have carried out the main research underlying a paper,
-while the corresponding author is often recognised as having supervised
-the research (Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly,
-measures based on first or corresponding authorships are strongly
-correlated, suggesting that these authors share the same country
-affiliation. Moreover, in most cases, the first author is identical the
+However, this analysis demonstrates that indicators based on first
+authors, which have often been used as a proxy for determining open
+access funding, and corresponding authors show a high level of
+correlation, reflecting disciplinary norms in scholarly publishing with
+regard to contributorship and author roles and positions (Larivière et
+al., 2016). First authors typically conduct the main research underlying
+a paper, while the corresponding author often supervises the research
+(Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly, measures
+based on first or corresponding authorships are strongly correlated,
+suggesting that these authors share the same country affiliation.
+Moreover, in most cases, the first author is identical to the
 corresponding author (Chinchilla-Rodríguez et al., 2024). Thus, first
@@
 
-Despite this correlation, the recent slight increase in open access
-articles by corresponding authors over first authors across Springer
-Nature hybrid journals prompts questions about how institutional open
-access sponsorship practices influence author roles and assignments
-within co-author teams, especially as funding opportunities vary
-(Gumpenberger et al., 2018). Previous research has highlighted how
-institutionalized bibliometric practices can affect the valuation of
-authorship positions (Helgesson, 2020), underscoring the need to be
-cautious of monetary authorship influences (Maddi \& Silva, 2024).
-
-Affiliation data is another important data element for estimating open
-access enabled by transformative agreements. Although OpenAlex's
-affiliation coverage is less comprehensive, which likely reduced the
-number of articles confidently attributed to transformative agreements
-in this study, it still shows high correlations with Scopus and Web of
-Science at the country level. However, OpenAlex's native ROR ID
-integration offers a distinct advantage, enabling more reliable
-identification of agreement-enabled articles compared to Scopus and Web
-of Science, which require reconciliation with proprietary organization
-identifiers. Future studies based on the Web of Science will benefit of
-the recent integration of ROR-IDs, announced by the end of 2024.
-
-Comparing scholarly databases reveals discrepancies in open access
-evidence. Not all publishers deposit Creative Commons (CC) licenses with
-Crossref, but comparing it with Unpaywall (integrated into Scopus and
-Web of Science) exposes these gaps, as these discovery services also
-analyze publishers' websites.
-
-Comparing scholarly databases helps to identify discrepancies in open
-access evidence. Crucially, not all publishers deposit CC licences with
+Despite this correlation, the study observed a recent slight increase in
+open access articles by corresponding authors over first authors across
+Springer Nature hybrid journals prompts questions about how
+institutional open access sponsorship practices influence author roles
+and assignments within co-author teams, especially as funding
+opportunities vary (Gumpenberger et al., 2018). Previous research has
+highlighted how institutionalized bibliometric practices can affect the
+valuation of authorship positions (Helgesson, 2020), suggesting the need
+to monitor potential influences on authorships (Maddi \& Silva, 2024).
+
+Another critical data element in the study is affiliation data, which is
+essential for estimating open access enabled by transformative
+agreements. Although OpenAlex's affiliation coverage is less
+comprehensive, which likely reduced the number of articles confidently
+attributed to transformative agreements in this study, it still shows
+high correlations with Scopus and Web of Science at the country level.
+However, OpenAlex's native ROR ID integration offers a distinct
+advantage, enabling more reliable identification of agreement-enabled
+articles compared to Scopus and Web of Science, which require
+reconciliation with proprietary organization identifiers. Future studies
+based on the Web of Science will benefit from the recent integration of
+ROR-IDs, announced by the end of 2024.
+
+The database comparison revealed important discrepancies in open access
+evidence. Crucially, not all publishers deposit CC licences with
 Crossref, a limitation that becomes apparent when contrasting Crossref
@@
 open access articles being recorded in Scopus and Web of Science.
-Further inconsistencies could be observed between Scopus and Web of
-Science, despite both relying on Unpaywall: Web of Science erroneously
-labels Elsevier's delayed OA as hybrid, whereas Scopus correctly
-categorises it. ONotably, Scopus and Crossref showed greater alignment
+Further inconsistencies emerged between Scopus and Web of Science,
+despite both relying on Unpaywall: Web of Science erroneously labels
+Elsevier's delayed open access as hybrid, whereas Scopus correctly
+categorises it. Notably, Scopus and Crossref showed greater alignment
 than observed in a related comparison between Scopus and OpenAlex
-(Alperin et al., 2024). This likely reflects the use of a dedicated list
+(Alperin et al., 2024). This likely reflects the use of a curated list
 to identify hybrid open access journals, rather than relying solely on
@@
 
-It is important to note that articles from institutions involved in
-transformative agreements are estimates due to a lack of access to
-invoice data, which is not usually shared between library consortia and
-publishers. Furthermore, it is important to acknowledge that the sources
-of transformative agreements and the underlying ESAC registry are a
-voluntary effort, crowd-sourced from various consortia. However, recent
-validation with Dutch Research Information demonstrates the reliability
-of such an open approach for the assessment of articles under
-transformative agreements. With enhanced and streamlined metadata
-compliance, attributable to evolving standards and initiatives to
-support negotiations with publishers, particularly through the
-communities of the ESAC initiative and Barcelona Declaration supporters,
-the situation is likely to improve. In conclusion, this cross-database
-analysis has highlighted the potential and challenges of establishing
-open scholarly data as a suitable evidence base for analysing the
-transition from subscription-based publishing to open access.
+It is important to note that the study's estimates of articles from
+institutions involved in transformative agreements are approximations
+due to a lack of access to invoice data, which is not usually shared by
+library consortia and publishers. Furthermore, the data sources about
+transformative agreements including the cOAlition S Journal Checker Tool
+and the ESAC registry are a voluntary effort, crowd-sourced from various
+consortia. However, recent validation with Dutch Research Information
+demonstrates the reliability of such an open approach for the assessment
+of articles under transformative agreements (Jonge et al., 2025). With
+enhanced and streamlined metadata compliance, attributable to evolving
+standards and initiatives to support negotiations with publishers,
+particularly through the communities of the ESAC initiative and
+Barcelona Declaration supporters, the situation is likely to improve. In
+conclusion, this study has shown that open scholarly data are suitable
+for analysing the transition from subscription-based publishing to full
+open access.
 
@@
 
+\bibitem[\citeproctext]{ref-eck_2024_6bx73}
+Eck, N. J. van, Waltman, L., \& Neijssel, M. (2024). \emph{{Launch of
+the CWTS Leiden Ranking Open Edition 2024}}. Leiden Madtrics.
+\url{https://doi.org/10.59350/r512t-r8h93}
+
 \bibitem[\citeproctext]{ref-Fox_2018}

### 2025-04-07 analysis/manuscript.qmd
First complete draft, to be shared

@@
 
-Transformative agreements have emerged as a much-discussed licensing model for the transition from subscription-based publishing to full open access in the last ten years. 
-While these agreements can vary substantially, increasing coordination exists in standardising the negotiation, administration, and evaluation of these agreements. 
-Simultaneously, efforts to improve transparency through open data about contracts and payments are growing. 
-However, this information remains only partially available through bibliometric databases, limiting large-scale studies about the impact of transformative agreements on the transition of hybrid journals to full open access through increased open access uptake.
-This study compares findings from hoaddata, an openly available,  continuously updated dataset on hybrid open access, built on open data from Crossref, OpenAlex and the cOAlition S Journal Checler Tool, with the established bibliometric databases Web of Science and Scopus.
+Transformative agreements are a much-discussed library licensing model to transition subscription-based journal publishing to full open access.
+While these agreements can vary substantially, they mainly target hybrid journal portfolios, enabling researchers from participating institutions to publish open access in them while also providing reading access to the entire portfolio.
+Responding to experiences with the subscription model, library consortia and funders have begun to coordinate the standardisation of negotiation, administration, and evaluation practices for these agreements.
+Simultaneously, efforts to improve transparency through open data about contracts and payments are ongoing. 
+However, this information remains only partially available through bibliometric databases, limiting large-scale studies about the impact of transformative agreements on the transition of hybrid journals to full open access.
+This study compares findings from hoaddata, an openly available, continuously updated dataset on hybrid open access, built on open data from Crossref, OpenAlex and the cOAlition S Journal Checker Tool, with the established bibliometric databases Web of Science and Scopus.
 The purpose of this comparison is not only to evaluate the suitability of open metadata for measuring hybrid open access but also to inform bibliometric research and practice by highlighting the strengths and weaknesses of each data source.
 
-Since their introduction, transformative agreements have sparked considerable academic debate. 
-Whilst standardisation activities, particularly through the OA2020 and ESAC (Efficiency and Standards for Article Charges) initiatives, have led to improved workflows between library consortia and publishers, as well as increased open access publishing, the initial focus on a few large publishers by national consortia from high-income countries has drawn criticism.
-Critics have raised concerns about market concentration, the lack of journal flipping to full open access models, reduced competition in the publishing market, and wider gaps between well-resourced and under-resourced institutions and researchers.
-
-For instance, studying over 700 agreements, Jahn (2025) observed a strong growth in open access publishing to 15% in 2022 due to the varying implementation of transformative agreements across countries, predominantly driven by three commercial publishers: Elsevier, Springer Nature, and Wiley.
-This aligns with the rapid proliferation of transformative agreements in recent years.
-Previously, only a minimal proportion of articles were made open access in hybrid journals (Piwowar, 2018; Laakso, 2016).
-
-Evaluations conducted by library consortia confirm these trends but draw different conclusions.
-For instance, the British Joint Information Systems Committee's evaluation (Brayman, Devenney et al., 2024) noted transformative agreements' significant impact on national open access growth but limited effect on global open access transition, suggesting a need to reevaluate their strategy.
-Similar policy reconsiderations emerged in Norway and Sweden (Holden, Skoie et al., 2023; Widding, 2024).
+Since the initial proposal to further develop hybrid open access as a means to cost-effectively repurpose subscription spending toward open access [@Schimmer_2015], transformative agreements have led to considerable academic debate.
+While standardisation activities, particularly through the OA2020 and ESAC (Efficiency and Standards for Article Charges) initiatives, have led to improved workflows between library consortia and publishers, as well as increased open access publishing [@ifla; @Geschuhn_2017], the initial focus on a few large publishers by national consortia from high-income countries has drawn criticism [@Borrego_2020].
+Critics have in particular raised concerns about market concentration [@Shu_2023; @Butler_2023], the failure of most hybrid journals to convert to full OA [@Momeni_2021; @Kiley_2024], reduced competition and incentives [@McCabe_2024; @Schmal_competition], and the continuation of publication fees (APCs) that widen gaps between well-resourced and under-resourced institutions and researchers, raising equity concerns [@budapest; @Ross_Hellauer_2022].
+
+In terms of bibliometric evidence, studying over 700 agreements, @Jahn_2025 observed a strong growth in open access publishing to 15% in 2022 due to the varying implementation of transformative agreements across countries, predominantly driven by European countries with national consortia that successfully negotiated these agreements and the three commercial publishers Elsevier, Springer Nature, and Wiley.
+This aligns with the rapid proliferation of transformative agreements in recent years [@McCabe_2024].
+Before, only a small proportion of articles were made open access in hybrid journals [@Laakso_2016, @Piwowar_2018], with only few European countries with dedicated hybrid open access funding policies and the high-energy physics SCOAP^3 consortium standing out [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020; @Kohls_2018].
+
+While consortia evaluations confirm the growth in open access publishing through transformative agreements, they reach mixed policy conclusions.
+For instance, the British Joint Information Systems Committee's evaluation [@Jisc_2024] noted transformative agreements' significant impact on national open access growth but limited effect on global open access transition, suggesting a need to reevaluate their strategy.
+Similar policy reconsiderations emerged in Norway and Sweden [@norway; @Widding_2024].
 Meanwhile, cOAlition S recommended ending financial support for these agreements in 2024, but still considers hybrid open access compliant with funder open access policies.
-Conversely, Germany's DEAL consortium extended agreements with major publishers through 2028, while Colombia established Latin America's first transformative agreements (Muñoz-Vélez, Pallares et al., 2024).
-A recent Max Planck Digital Library study demonstrates how cost savings can be achieved through this model and how it can be further developed to achieve full open access while controlling expenditure.
-
-Open databases like OpenAlex, which build upon Crossref, particularly benefit from these developments.
-These resources have gained considerable attention in research evaluation due to their inclusivity, openness, and comprehensive coverage of publication metadata.
-Recent studies [REFERENCE NEEDED] have demonstrated that open bibliometric databases increasingly match or exceed the coverage of traditional commercial databases in many disciplines, whilst providing improved access to standardised and interlinked open research information.
+Conversely, Germany's DEAL consortium extended agreements with major publishers through 2028, while Colombia established Latin America's first transformative agreements [@Mu_oz_V_lez_2024].
+Presenting experiences at the Max Planck Digital Library, [@D_r_2025] outlines strategies to achieve full open access while controlling costs.
+
+Open research information can benefit from transformative agreements as library consortia also negotiate for more comprehensive open metadata to avoid commodification of data analytics capabilities needed to negotiate and evaluate agreements [@Aspesi_2020; @McCabe_2024].
+For instance, ESAC recommends using Crossref [@Hendricks_2020] to share metadata supporting workflows including open access licenses and funding information [@Geschuhn_2017]. 
+Simultaneously, the bibliometric community pushes for more open metadata through this publisher-driven DOI registration platform and continuously monitors progress [@eck_crossref_2022].
+
+Numerous open scholarly data services building upon Crossref benefit from these developments.
+A prominent example is OpenAlex, which has gained considerable attention by promising inclusivity, openness, and comprehensive coverage of publication metadata [@priem2022openalex].
+In a recent evaluation of OpenAlex, [@C_spedes_2025] reviewed the current evidence base about its strengths and shortcomings.
+They highlight both its increasing use in bibliometric research and adoption across research organisations.
+Comparative studies generally perceive OpenAlex as a reliable alternative to commercial databases [@alperin2024analysissuitabilityopenalexbibliometric; @culbert2024referencecoverageanalysisopenalex], although data quality issues were observed regarding important metadata fields needed to analyse open access, including evidence about open access status itself [@jahn2023analysing; @Simard_2024], as well as authors [@culbert2024referencecoverageanalysisopenalex], affiliations [@Zhang_2024], author roles [@Jahn_2025], and document types used to distinguish between different article types [@Haupka_2024].
+However, these studies emphasise the new possibilities of combining open metadata and OpenAlex's curation efforts in conjunction with growing awareness about the importance of open metadata as demonstrated by the Barcelona Declaration on Open Research Information.
 
 Against this background, this study evaluates the suitability of open metadata to measure the impact of transformative agreements on hybrid open access.
-Only a few studies have attempted this analysis, primarily due to data limitations and the need to combine multiple sources.
-Transformative agreements present unique measurement challenges because of their complex workflow requirements, which are not fully reflected in bibliometric databases.
-These challenges include varying journal portfolios across agreements, invoicing procedures specific to submitting authors, and article caps that limit open access publishing within agreements.
-
-The present study addresses these gaps by replicating and applying consistent methodologies across Web of Science, Scopus, and open data sources.
-This comparative approach enables the research to overcome the limitations of any single source whilst systematically assessing the strengths and weaknesses of different bibliometric data sources for monitoring transformative agreement outcomes.
-By triangulating findings across these platforms, the study aims to provide a more comprehensive understanding of how transformative agreements influence hybrid open access publishing patterns within the ongoing scholarly publishing transition.
+Few studies have attempted this analysis due to data limitations and the need to combine multiple sources.
+@de_Jonge_2025 provides a notable exception by comparing transformative agreement data from the cOAlition S Journal Checker tool and OpenAlex with institutional invoicing records for publications funded by Dutch research funders through the Dutch library consortium.
+Their analysis found 89% accuracy in identifying open access articles enabled by transformative agreements, with discrepancies resulting from both open data limitations and the inherent complexity of agreement structures that require data from both library consortia and publishers—information that is rarely made publicly available.
+Recent large-scale studies [@Bakker_2024; @Haustein_APC_2024; @Kramer_2024; @Jahn_2025] confirm that these challenges extend beyond the Dutch context, highlighting monitoring difficulties related to varying journal portfolios, challenges in using author roles as a proxy for invoicing data, and article caps within agreements.
+
+The present study addresses these gaps by adapting the open approach used in @Jahn_2025, which combines cOAlition S Journal Checker tool data, Crossref and OpenAlex to measure the impact of transformative agreements on the open access uptake of hybrid journals, and applying this same methodology to Web of Science and Scopus data.
+This comparative approach enables assessment of different bibliometric data sources for monitoring transformative agreements while evaluating their respective strengths and limitations.
+Specifically, this study compares the indexing coverage of more than 13,000 hybrid journals across these sources and examines whether consistent results can be obtained regarding open access uptake between 2019 and 2023, analysed by publisher and country.
+By comparing findings across these platforms, this large-scale study aims to provide a more comprehensive understanding of how transformative agreements impact the ongoing transition from subscription-based journal publishing to full open access.
 
@@
 ## Data analysis
+
+Throughout this largely automated data collection and analysis process, Tidyverse tools [@tidyverse] for the R programming language [@r] were used. 
+Rank correlations were calculated using the Hmisc package [@Harrell_Jr_2003].
+The source code analysis, including all queries used to obtain the data, is available on GitHub (add link). 
+An interactive supplement exploring correlations between the data sources examined by country and publisher is available on HuggingFace Spaces: https://huggingface.co/spaces/najkoja/hoa_replication.
 
@@
 In contrast, the three most productive countries, China, the United States, and India, were substantially ranked lower in transformative agreement adoption.
-Open questions remain whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or countries deliberately refraining from such agreements.
+Open questions remain whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or a deliberate avoidance of such agreements.
 
 The findings were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science. 
-Notably, the results show strong correlations by country affiliation, aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73], which supports using open metadata for large-scale analyses of hybrid open access.
+Notably, and aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73], the results show strong correlations by country affiliation, which supports using open metadata for large-scale analyses of hybrid open access.
 However, the observed differences in journal coverage and metadata availability warrant discussion, affecting not only open data sources but also proprietary databases when used in isolation.
 
-The coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage.
+The coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage, reflecting the dominance of major publishers whose established journal portfolios are comprehensively indexed in proprietary databases [@vanbellen2024oligopolyacademicpublisherspersists].
 Differences emerge for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
-Using the open index Crossref as a bibliometric database in hoaddata demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
-
-The landscape of hybrid journal publishing differs markedly from that of fully open access journals.
+Using Crossref as a bibliometric database in hoaddata demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
+The landscape of hybrid journal publishing thus differs markedly from that of fully open access journals.
 Comparing the coverage of OpenAlex, Scopus and Web of Science, @Simard_2024 indicate that only half of the fully open access journals listed in the DOAJ are also indexed in Scopus and Web of Science.
-Notably, journals that charge no publication fees ("diamond journals") are absent from the selective Web of Science, which reinforces existing disparities in the indexing of underrepresented research communities and regions in selective bibliometric databases (ref).
+Notably, journals that charge no publication fees ("diamond journals") are absent from the selective Web of Science, which reinforces existing disparities in the indexing of under-represented research communities and regions in selective bibliometric databases [@Simard_2024]
 
@@
 Moreover, in most cases, the first author is identical to the corresponding author [@Chinchilla_Rodr_guez_2024].
-Thus, first authorship can reliably be used in large-scale studies to assess the impact of transformative agreements.
-
+<!-- Thus, first authorship can reliably be used in large-scale studies to assess the impact of transformative agreements. -->
 Despite this correlation, the study observed a recent slight increase in open access articles by corresponding authors over first authors across Springer Nature hybrid journals prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
-Previous research has highlighted how institutionalized bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], suggesting the need to monitor potential influences on authorships [@Maddi_2024].
+Previous research has highlighted how institutionalized bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], suggesting the need to monitor potential influences of the availability of open access funding on authorships [@Maddi_2024].
 
@@
 Given these variations, which may compromise comparability, research and monitoring exercises should avoid reliance on a single source.
-Instead, the selection of open access evidence should be cross-verified using multiple sources [@Huang_2020].
-Incorporating expert-curated journal lists can also help prevent misclassifications based on business models [@Visser_2021].
+Instead, the selection of open access evidence should be cross-verified using multiple sources and snapshots that can be used to track changes in the data over time [@Huang_2020].
+Incorporating expert-curated journal lists can also help prevent misclassification based on business models [@Visser_2021].
 
@@
 Furthermore, the data sources about transformative agreements including the cOAlition S Journal Checker Tool and the ESAC registry are a voluntary effort, crowd-sourced from various consortia. 
-However, recent validation with Dutch Research Information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements [@de_Jonge_2025]. 
-With enhanced and streamlined metadata compliance, attributable to evolving standards and initiatives to support negotiations with publishers, particularly through the communities of the ESAC initiative and Barcelona Declaration supporters, the situation is likely to improve.
-In conclusion, this study has shown that open scholarly data are suitable for analysing the transition from subscription-based publishing to full open access.
+However, recent validation with Dutch research information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements [@de_Jonge_2025]. 
+With enhanced open metadata compliance, attributable to evolving standards and initiatives to support negotiations with publishers, particularly through the communities of the ESAC initiative and Barcelona Declaration supporters, the situation is likely to improve.
+In conclusion, this study has shown that open metadata are suitable for analysing the transition from subscription-based journal publishing to full open access.
 

### 2025-04-07 analysis/manuscript.tex
First complete draft, to be shared

@@
 
-Transformative agreements have emerged as a much-discussed licensing
-model for the transition from subscription-based publishing to full open
-access in the last ten years. While these agreements can vary
-substantially, increasing coordination exists in standardising the
-negotiation, administration, and evaluation of these agreements.
-Simultaneously, efforts to improve transparency through open data about
-contracts and payments are growing. However, this information remains
-only partially available through bibliometric databases, limiting
-large-scale studies about the impact of transformative agreements on the
-transition of hybrid journals to full open access through increased open
-access uptake. This study compares findings from hoaddata, an openly
+Transformative agreements are a much-discussed library licensing model
+to transition subscription-based journal publishing to full open access.
+While these agreements can vary substantially, they mainly target hybrid
+journal portfolios, enabling researchers from participating institutions
+to publish open access in them while also providing reading access to
+the entire portfolio. Responding to experiences with the subscription
+model, library consortia and funders have begun to coordinate the
+standardisation of negotiation, administration, and evaluation practices
+for these agreements. Simultaneously, efforts to improve transparency
+through open data about contracts and payments are ongoing. However,
+this information remains only partially available through bibliometric
+databases, limiting large-scale studies about the impact of
+transformative agreements on the transition of hybrid journals to full
+open access. This study compares findings from hoaddata, an openly
 available, continuously updated dataset on hybrid open access, built on
-open data from Crossref, OpenAlex and the cOAlition S Journal Checler
+open data from Crossref, OpenAlex and the cOAlition S Journal Checker
 Tool, with the established bibliometric databases Web of Science and
@@
 
-Since their introduction, transformative agreements have sparked
-considerable academic debate. Whilst standardisation activities,
+Since the initial proposal to further develop hybrid open access as a
+means to cost-effectively repurpose subscription spending toward open
+access (Schimmer et al., 2015), transformative agreements have led to
+considerable academic debate. While standardisation activities,
 particularly through the OA2020 and ESAC (Efficiency and Standards for
@@
 library consortia and publishers, as well as increased open access
-publishing, the initial focus on a few large publishers by national
-consortia from high-income countries has drawn criticism. Critics have
-raised concerns about market concentration, the lack of journal flipping
-to full open access models, reduced competition in the publishing
-market, and wider gaps between well-resourced and under-resourced
-institutions and researchers.
-
-For instance, studying over 700 agreements, Jahn (2025) observed a
-strong growth in open access publishing to 15\% in 2022 due to the
-varying implementation of transformative agreements across countries,
-predominantly driven by three commercial publishers: Elsevier, Springer
-Nature, and Wiley. This aligns with the rapid proliferation of
-transformative agreements in recent years. Previously, only a minimal
-proportion of articles were made open access in hybrid journals
-(Piwowar, 2018; Laakso, 2016).
-
-Evaluations conducted by library consortia confirm these trends but draw
-different conclusions. For instance, the British Joint Information
-Systems Committee's evaluation (Brayman, Devenney et al., 2024) noted
-transformative agreements' significant impact on national open access
-growth but limited effect on global open access transition, suggesting a
-need to reevaluate their strategy. Similar policy reconsiderations
-emerged in Norway and Sweden (Holden, Skoie et al., 2023; Widding,
-2024). Meanwhile, cOAlition S recommended ending financial support for
-these agreements in 2024, but still considers hybrid open access
-compliant with funder open access policies. Conversely, Germany's DEAL
-consortium extended agreements with major publishers through 2028, while
-Colombia established Latin America's first transformative agreements
-(Muñoz-Vélez, Pallares et al., 2024). A recent Max Planck Digital
-Library study demonstrates how cost savings can be achieved through this
-model and how it can be further developed to achieve full open access
-while controlling expenditure.
-
-Open databases like OpenAlex, which build upon Crossref, particularly
-benefit from these developments. These resources have gained
-considerable attention in research evaluation due to their inclusivity,
-openness, and comprehensive coverage of publication metadata. Recent
-studies {[}REFERENCE NEEDED{]} have demonstrated that open bibliometric
-databases increasingly match or exceed the coverage of traditional
-commercial databases in many disciplines, whilst providing improved
-access to standardised and interlinked open research information.
+publishing (Campbell et al., 2022; Geschuhn \& Stone, 2017), the initial
+focus on a few large publishers by national consortia from high-income
+countries has drawn criticism (Borrego et al., 2021). Critics have in
+particular raised concerns about market concentration (Butler et al.,
+2023; Shu \& Larivière, 2023), the failure of most hybrid journals to
+convert to full OA (Kiley, 2024; Momeni et al., 2021), reduced
+competition and incentives (McCabe \& Mueller-Langer, 2024; Schmal,
+2024), and the continuation of publication fees (APCs) that widen gaps
+between well-resourced and under-resourced institutions and researchers,
+raising equity concerns (Babini et al., 2022; Ross-Hellauer et al.,
+2022).
+
+In terms of bibliometric evidence, studying over 700 agreements, Jahn
+(2025) observed a strong growth in open access publishing to 15\% in
+2022 due to the varying implementation of transformative agreements
+across countries, predominantly driven by European countries with
+national consortia that successfully negotiated these agreements and the
+three commercial publishers Elsevier, Springer Nature, and Wiley. This
+aligns with the rapid proliferation of transformative agreements in
+recent years (McCabe \& Mueller-Langer, 2024). Before, only a small
+proportion of articles were made open access in hybrid journals Piwowar
+et al. (2018), with only few European countries with dedicated hybrid
+open access funding policies and the high-energy physics SCOAP\^{}3
+consortium standing out (Huang et al., 2020; Kohls \& Mele, 2018;
+Pinfield, 2015; Robinson-Garcia et al., 2020).
+
+While consortia evaluations confirm the growth in open access publishing
+through transformative agreements, they reach mixed policy conclusions.
+For instance, the British Joint Information Systems Committee's
+evaluation (Brayman et al., 2024) noted transformative agreements'
+significant impact on national open access growth but limited effect on
+global open access transition, suggesting a need to reevaluate their
+strategy. Similar policy reconsiderations emerged in Norway and Sweden
+(Holden et al., 2023; Widding, 2024). Meanwhile, cOAlition S recommended
+ending financial support for these agreements in 2024, but still
+considers hybrid open access compliant with funder open access policies.
+Conversely, Germany's DEAL consortium extended agreements with major
+publishers through 2028, while Colombia established Latin America's
+first transformative agreements (Muñoz-Vélez et al., 2024). Presenting
+experiences at the Max Planck Digital Library, (Dér, 2025) outlines
+strategies to achieve full open access while controlling costs.
+
+Open research information can benefit from transformative agreements as
+library consortia also negotiate for more comprehensive open metadata to
+avoid commodification of data analytics capabilities needed to negotiate
+and evaluate agreements (Aspesi \& Brand, 2020; McCabe \&
+Mueller-Langer, 2024). For instance, ESAC recommends using Crossref
+(Hendricks et al., 2020) to share metadata supporting workflows
+including open access licenses and funding information (Geschuhn \&
+Stone, 2017). Simultaneously, the bibliometric community pushes for more
+open metadata through this publisher-driven DOI registration platform
+and continuously monitors progress (N. J. van Eck \& Waltman, 2022).
+
+Numerous open scholarly data services building upon Crossref benefit
+from these developments. A prominent example is OpenAlex, which has
+gained considerable attention by promising inclusivity, openness, and
+comprehensive coverage of publication metadata (Priem et al., 2022). In
+a recent evaluation of OpenAlex, (Céspedes et al., 2025) reviewed the
+current evidence base about its strengths and shortcomings. They
+highlight both its increasing use in bibliometric research and adoption
+across research organisations. Comparative studies generally perceive
+OpenAlex as a reliable alternative to commercial databases (Alperin et
+al., 2024; Culbert et al., 2024), although data quality issues were
+observed regarding important metadata fields needed to analyse open
+access, including evidence about open access status itself (Jahn et al.,
+2023; Simard et al., 2024), as well as authors (Culbert et al., 2024),
+affiliations (Zhang et al., 2024), author roles (Jahn, 2025), and
+document types used to distinguish between different article types
+(Haupka et al., 2024). However, these studies emphasise the new
+possibilities of combining open metadata and OpenAlex's curation efforts
+in conjunction with growing awareness about the importance of open
+metadata as demonstrated by the Barcelona Declaration on Open Research
+Information.
 
@@
 metadata to measure the impact of transformative agreements on hybrid
-open access. Only a few studies have attempted this analysis, primarily
-due to data limitations and the need to combine multiple sources.
-Transformative agreements present unique measurement challenges because
-of their complex workflow requirements, which are not fully reflected in
-bibliometric databases. These challenges include varying journal
-portfolios across agreements, invoicing procedures specific to
-submitting authors, and article caps that limit open access publishing
-within agreements.
-
-The present study addresses these gaps by replicating and applying
-consistent methodologies across Web of Science, Scopus, and open data
-sources. This comparative approach enables the research to overcome the
-limitations of any single source whilst systematically assessing the
-strengths and weaknesses of different bibliometric data sources for
-monitoring transformative agreement outcomes. By triangulating findings
-across these platforms, the study aims to provide a more comprehensive
-understanding of how transformative agreements influence hybrid open
-access publishing patterns within the ongoing scholarly publishing
-transition.
+open access. Few studies have attempted this analysis due to data
+limitations and the need to combine multiple sources. Jonge et al.
+(2025) provides a notable exception by comparing transformative
+agreement data from the cOAlition S Journal Checker tool and OpenAlex
+with institutional invoicing records for publications funded by Dutch
+research funders through the Dutch library consortium. Their analysis
+found 89\% accuracy in identifying open access articles enabled by
+transformative agreements, with discrepancies resulting from both open
+data limitations and the inherent complexity of agreement structures
+that require data from both library consortia and
+publishers---information that is rarely made publicly available. Recent
+large-scale studies (Bakker et al., 2024; Haustein et al., 2024; Jahn,
+2025; Kramer, 2024) confirm that these challenges extend beyond the
+Dutch context, highlighting monitoring difficulties related to varying
+journal portfolios, challenges in using author roles as a proxy for
+invoicing data, and article caps within agreements.
+
+The present study addresses these gaps by adapting the open approach
+used in Jahn (2025), which combines cOAlition S Journal Checker tool
+data, Crossref and OpenAlex to measure the impact of transformative
+agreements on the open access uptake of hybrid journals, and applying
+this same methodology to Web of Science and Scopus data. This
+comparative approach enables assessment of different bibliometric data
+sources for monitoring transformative agreements while evaluating their
+respective strengths and limitations. Specifically, this study compares
+the indexing coverage of more than 13,000 hybrid journals across these
+sources and examines whether consistent results can be obtained
+regarding open access uptake between 2019 and 2023, analysed by
+publisher and country. By comparing findings across these platforms,
+this large-scale study aims to provide a more comprehensive
+understanding of how transformative agreements impact the ongoing
+transition from subscription-based journal publishing to full open
+access.
 
@@
 
+Throughout this largely automated data collection and analysis process,
+Tidyverse tools (Wickham et al., 2019) for the R programming language (R
+Core Team, 2024) were used. Rank correlations were calculated using the
+Hmisc package (Harrell Jr, 2003). The source code analysis, including
+all queries used to obtain the data, is available on GitHub (add link).
+An interactive supplement exploring correlations between the data
+sources examined by country and publisher is available on HuggingFace
+Spaces: https://huggingface.co/spaces/najkoja/hoa\_replication.
+
 \section{Results}\label{results}
@@
 reflects temporary implementation gaps, inherent inequities in the
-transformative agreement model, or countries deliberately refraining
-from such agreements.
+transformative agreement model, or a deliberate avoidance of such
+agreements.
 
@@
 proprietary bibliometric databases Scopus and Web of Science. Notably,
-the results show strong correlations by country affiliation, aligning
-with previous studies (Akbaritabar et al., 2024; Alperin et al., 2024)
-and rankings (Eck et al., 2024), which supports using open metadata for
-large-scale analyses of hybrid open access. However, the observed
-differences in journal coverage and metadata availability warrant
-discussion, affecting not only open data sources but also proprietary
-databases when used in isolation.
+and aligning with previous studies (Akbaritabar et al., 2024; Alperin et
+al., 2024) and rankings (N. J. van Eck et al., 2024), the results show
+strong correlations by country affiliation, which supports using open
+metadata for large-scale analyses of hybrid open access. However, the
+observed differences in journal coverage and metadata availability
+warrant discussion, affecting not only open data sources but also
+proprietary databases when used in isolation.
 
 The coverage analysis reveals that hybrid journals are well indexed in
-all three data sources, particularly in terms of article coverage.
-Differences emerge for journals targeting practitioners or local
-non-English language communities, with many such titles indexed
-exclusively in Crossref and Scopus. Using the open index Crossref as a
-bibliometric database in hoaddata demonstrated particular strength in
-identifying newly established hybrid journals, a notable finding given
-that transformative agreements primarily target existing
-subscription-based journals.
-
-The landscape of hybrid journal publishing differs markedly from that of
-fully open access journals. Comparing the coverage of OpenAlex, Scopus
-and Web of Science, Simard et al. (2024) indicate that only half of the
-fully open access journals listed in the DOAJ are also indexed in Scopus
-and Web of Science. Notably, journals that charge no publication fees
-(``diamond journals'') are absent from the selective Web of Science,
-which reinforces existing disparities in the indexing of
-underrepresented research communities and regions in selective
-bibliometric databases (ref).
+all three data sources, particularly in terms of article coverage,
+reflecting the dominance of major publishers whose established journal
+portfolios are comprehensively indexed in proprietary databases (Bellen
+et al., 2024). Differences emerge for journals targeting practitioners
+or local non-English language communities, with many such titles indexed
+exclusively in Crossref and Scopus. Using Crossref as a bibliometric
+database in hoaddata demonstrated particular strength in identifying
+newly established hybrid journals, a notable finding given that
+transformative agreements primarily target existing subscription-based
+journals. The landscape of hybrid journal publishing thus differs
+markedly from that of fully open access journals. Comparing the coverage
+of OpenAlex, Scopus and Web of Science, Simard et al. (2024) indicate
+that only half of the fully open access journals listed in the DOAJ are
+also indexed in Scopus and Web of Science. Notably, journals that charge
+no publication fees (``diamond journals'') are absent from the selective
+Web of Science, which reinforces existing disparities in the indexing of
+under-represented research communities and regions in selective
+bibliometric databases (Simard et al., 2024)
 
@@
 Moreover, in most cases, the first author is identical to the
-corresponding author (Chinchilla-Rodríguez et al., 2024). Thus, first
-authorship can reliably be used in large-scale studies to assess the
-impact of transformative agreements.
-
-Despite this correlation, the study observed a recent slight increase in
-open access articles by corresponding authors over first authors across
-Springer Nature hybrid journals prompts questions about how
-institutional open access sponsorship practices influence author roles
-and assignments within co-author teams, especially as funding
-opportunities vary (Gumpenberger et al., 2018). Previous research has
-highlighted how institutionalized bibliometric practices can affect the
-valuation of authorship positions (Helgesson, 2020), suggesting the need
-to monitor potential influences on authorships (Maddi \& Silva, 2024).
+corresponding author (Chinchilla-Rodríguez et al., 2024). Despite this
+correlation, the study observed a recent slight increase in open access
+articles by corresponding authors over first authors across Springer
+Nature hybrid journals prompts questions about how institutional open
+access sponsorship practices influence author roles and assignments
+within co-author teams, especially as funding opportunities vary
+(Gumpenberger et al., 2018). Previous research has highlighted how
+institutionalized bibliometric practices can affect the valuation of
+authorship positions (Helgesson, 2020), suggesting the need to monitor
+potential influences of the availability of open access funding on
+authorships (Maddi \& Silva, 2024).
 
@@
 on a single source. Instead, the selection of open access evidence
-should be cross-verified using multiple sources (Huang et al., 2020).
+should be cross-verified using multiple sources and snapshots that can
+be used to track changes in the data over time (Huang et al., 2020).
 Incorporating expert-curated journal lists can also help prevent
-misclassifications based on business models (Visser et al., 2021).
+misclassification based on business models (Visser et al., 2021).
 
@@
 and the ESAC registry are a voluntary effort, crowd-sourced from various
-consortia. However, recent validation with Dutch Research Information
+consortia. However, recent validation with Dutch research information
 demonstrates the reliability of such an open approach for the assessment
 of articles under transformative agreements (Jonge et al., 2025). With
-enhanced and streamlined metadata compliance, attributable to evolving
-standards and initiatives to support negotiations with publishers,
-particularly through the communities of the ESAC initiative and
-Barcelona Declaration supporters, the situation is likely to improve. In
-conclusion, this study has shown that open scholarly data are suitable
-for analysing the transition from subscription-based publishing to full
-open access.
+enhanced open metadata compliance, attributable to evolving standards
+and initiatives to support negotiations with publishers, particularly
+through the communities of the ESAC initiative and Barcelona Declaration
+supporters, the situation is likely to improve. In conclusion, this
+study has shown that open metadata are suitable for analysing the
+transition from subscription-based journal publishing to full open
+access.
 
@@
 
+\bibitem[\citeproctext]{ref-Aspesi_2020}
+Aspesi, C., \& Brand, A. (2020). In pursuit of open science, open access
+is not enough. \emph{Science}, \emph{368}(6491), 574--577.
+\url{https://doi.org/10.1126/science.aba3763}
+
 \bibitem[\citeproctext]{ref-Baas_2020}
@@
 
+\bibitem[\citeproctext]{ref-budapest}
+Babini, D., Chan, L., Hagemann, M., Joseph, H., Kuchma, I., \& Suber, P.
+(2022). \emph{{The Budapest Open Access Initiative-20th. Anniversary
+recommendations (BOAI20)}}.
+\url{https://www.budapestopenaccessinitiative.org/boai20/}
+
+\bibitem[\citeproctext]{ref-Bakker_2024}
+Bakker, C., Langham-Putrow, A., \& Riegelman, A. (2024). Impact of
+transformative agreements on publication patterns: An analysis based on
+agreements from the ESAC registry. \emph{International Journal of
+Librarianship}, \emph{8}(4), 67--96.
+\url{https://doi.org/10.23974/ijol.2024.vol8.4.341}
+
+\bibitem[\citeproctext]{ref-vanbellen2024oligopolyacademicpublisherspersists}
+Bellen, S. van, Alperin, J. P., \& Larivière, V. (2024). \emph{The
+oligopoly of academic publishers persists in exclusive database}.
+\url{https://arxiv.org/abs/2406.17893}
+
 \bibitem[\citeproctext]{ref-Birkle_2020}
-Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). Web of
-science as a data source for research on scientific and scholarly
+Birkle, C., Pendlebury, D. A., Schnell, J., \& Adams, J. (2020). {Web of
+Science} as a data source for research on scientific and scholarly
 activity. \emph{Quantitative Science Studies}, \emph{1}(1), 363--376.
@@
 
+\bibitem[\citeproctext]{ref-Jisc_2024}
+Brayman, K., Devenney, A., Dobson, H., Marques, M., \& Vernon, A.
+(2024). \emph{A review of transitional agreements in the {UK}}. Zenodo.
+\url{https://doi.org/10.5281/zenodo.10787392}
+
+\bibitem[\citeproctext]{ref-Butler_2023}
+Butler, L.-A., Matthias, L., Simard, M.-A., Mongeon, P., \& Haustein, S.
+(2023). The oligopoly's shift to open access: How the big five academic
+publishers profit from article processing charges. \emph{Quantitative
+Science Studies}, \emph{4}(4), 778--799.
+\url{https://doi.org/10.1162/qss_a_00272}
+
+\bibitem[\citeproctext]{ref-ifla}
+Campbell, C., Dér, Á., Geschuhn, K., \& Valente, A. (2022). How are
+transformative agreements transforming libraries? \emph{87th IFLA World
+Library and Information Congress (WLIC) / 2022 in Dublin, Ireland}.
+\url{https://repository.ifla.org/handle/123456789/1973}
+
+\bibitem[\citeproctext]{ref-C_spedes_2025}
+Céspedes, L., Kozlowski, D., Pradier, C., Sainte‐Marie, M. H., Shokida,
+N. S., Benz, P., Poitras, C., Ninkov, A. B., Ebrahimy, S., Ayeni, P.,
+Filali, S., Li, B., \& Larivière, V. (2025). Evaluating the linguistic
+coverage of {OpenAlex}: An assessment of metadata accuracy and
+completeness. \emph{Journal of the Association for Information Science
+and Technology}. \url{https://doi.org/10.1002/asi.24979}
+
 \bibitem[\citeproctext]{ref-Chinchilla_Rodr_guez_2024}
@@
 
+\bibitem[\citeproctext]{ref-culbert2024referencecoverageanalysisopenalex}
+Culbert, J., Hobert, A., Jahn, N., Haupka, N., Schmidt, M., Donner, P.,
+\& Mayr, P. (2024). \emph{Reference coverage analysis of OpenAlex
+compared to web of science and scopus}.
+\url{https://arxiv.org/abs/2401.16359}
+
+\bibitem[\citeproctext]{ref-D_r_2025}
+Dér, Á. (2025). What gets missed in the discourse on transformative
+agreements. \emph{Katina Magazine}.
+\url{https://doi.org/10.1146/katina-20250212-1}
+
 \bibitem[\citeproctext]{ref-Donner_2017}
@@
 \emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}
+
+\bibitem[\citeproctext]{ref-eck_crossref_2022}
+Eck, N. J. van, \& Waltman, L. (2022). \emph{Crossref as a source of
+open bibliographic metadata}. MetaArXiv.
+\url{https://doi.org/10.31222/osf.io/smxe5}
 
@@
 
+\bibitem[\citeproctext]{ref-Harrell_Jr_2003}
+Harrell Jr, F. E. (2003). Hmisc: Harrell miscellaneous. In \emph{CRAN:
+Contributed Packages}. The R Foundation.
+\url{https://doi.org/10.32614/cran.package.hmisc}
+
 \bibitem[\citeproctext]{ref-Haucap_2021}
@@
 
+\bibitem[\citeproctext]{ref-norway}
+Holden, L., Skoie, M., Røeggen, V., Bjerde, K. W., Wenaas, L., Bakke,
+P., Løvhaug, J. W., Karlsen, E. S., \& Qvenild, M. (2023).
+\emph{Strategi for vitenskapelig publisering etter 2024}. Sikt.
+\url{https://doi.org/10.18711/2KZ1-BA97}
+
 \bibitem[\citeproctext]{ref-Hottenrott_2021}
@@
 Ozaygen, A., \& Brookes-Kenworthy, C. (2020). \emph{Evaluating
-institutional open access performance: Sensitivity analysis}.
+institutional open access performance: Sensitivity analysis}. bioRxiv.
 \url{https://doi.org/10.1101/2020.03.19.998542}
@@
 Jahn, N. (2025). How open are hybrid journals included in transformative
-agreements? \emph{Quantitative Science Studies}, 1--39.
+agreements? \emph{Quantitative Science Studies}, \emph{6}, 242--262.
 \url{https://doi.org/10.1162/qss_a_00348}
+
+\bibitem[\citeproctext]{ref-jahn2023analysing}
+Jahn, N., Haupka, N., \& Hobert, A. (2023). \emph{Analysing and
+reclassifying open access information in OpenAlex}. {Scholarly
+Communication Analytics Blog}.
+\url{https://subugoe.github.io/scholcomm_analytics/posts/oalex_oa_status/}
 
@@
 transformative agreements through open metadata: Method and validation
-using dutch research council NWO funded papers.}
+using {Dutch Research Council NWO} funded papers}. MetaArXiv.
 \url{https://doi.org/10.31222/osf.io/tz6be_v1}
+
+\bibitem[\citeproctext]{ref-Kiley_2024}
+Kiley, R. (2024). \emph{{Transformative Journals}: Analysis from the
+2023 reports}. {cOAlition S}.
+\url{https://www.coalition-s.org/blog/transformative-journals-analysis-from-the-2023-reports/}
+
+\bibitem[\citeproctext]{ref-Kohls_2018}
+Kohls, A., \& Mele, S. (2018). Converting the literature of a scientific
+field to open access through global collaboration: The experience of
+SCOAP3 in particle physics. \emph{Publications}, \emph{6}(2), 15.
+\url{https://doi.org/10.3390/publications6020015}
+
+\bibitem[\citeproctext]{ref-Kramer_2024}
+Kramer, B. (2024). \emph{Study on scientific publishing in {Europe} --
+{Development}, diversity, and transparency of costs}. Publications
+Office of the European Union. \url{https://doi.org/doi/10.2777/89349}
 
@@
 \url{https://doi.org/10.5281/zenodo.3700590}
+
+\bibitem[\citeproctext]{ref-Laakso_2016}
+Laakso, M., \& Björk, B.-C. (2016). Hybrid open access--a longitudinal
+study. \emph{Journal of Informetrics}, \emph{10}(4), 919--932.
+\url{https://doi.org/10.1016/j.joi.2016.08.002}
 
@@
 
+\bibitem[\citeproctext]{ref-McCabe_2024}
+McCabe, M. J., \& Mueller-Langer, F. (2024). Open access is shaping
+scientific communication. \emph{Science}, \emph{385}(6714), 1170--1172.
+\url{https://doi.org/10.1126/science.adp8882}
+
 \bibitem[\citeproctext]{ref-Melero_Fuentes_2025}
@@
 \url{https://doi.org/10.1016/j.joi.2024.101631}
+
+\bibitem[\citeproctext]{ref-Momeni_2021}
+Momeni, F., Mayr, P., Fraser, N., \& Peters, I. (2021). What happens
+when a journal converts to open access? A bibliometric analysis.
+\emph{Scientometrics}, \emph{126}(12), 9811--9827.
+\url{https://doi.org/10.1007/s11192-021-03972-5}
+
+\bibitem[\citeproctext]{ref-Mu_oz_V_lez_2024}
+Muñoz-Vélez, H., Pallares, C., Echavarría, A. F., Contreras, J., Pavas,
+A., Bello, D., Rendón, C., Calderón-Rojas, J., \& Garzón, F. (2024).
+Strategies for negotiating and signing transformative agreements in the
+{Global South}: The {Colombia Consortium} experience. \emph{Journal of
+Library Administration}, \emph{64}(1), 80--98.
+\url{https://doi.org/10.1080/01930826.2023.2287945}
+
+\bibitem[\citeproctext]{ref-Pinfield_2015}
+Pinfield, S. (2015). Making open access work: The {``state-of-the-art''}
+in providing open access to scholarly literature. \emph{Online
+Information Review}, \emph{39}(5), 604--636.
+\url{https://doi.org/10.1108/oir-05-2015-0167}
 
@@
 
+\bibitem[\citeproctext]{ref-r}
+R Core Team. (2024). \emph{R: A language and environment for statistical
+computing}. R Foundation for Statistical Computing.
+\url{https://www.R-project.org/}
+
 \bibitem[\citeproctext]{ref-Rabesandratana_2019}
-Rabesandratana, T. (2019). Elsevier deal with france disappoints
+Rabesandratana, T. (2019). Elsevier deal with {France} disappoints
 open-access advocates. \emph{Science}.
 \url{https://doi.org/10.1126/science.aba5656}
+
+\bibitem[\citeproctext]{ref-Robinson_Garcia_2020}
+Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
+access uptake by universities worldwide. \emph{{PeerJ}}, \emph{8},
+e9410. \url{https://doi.org/10.7717/peerj.9410}
+
+\bibitem[\citeproctext]{ref-Ross_Hellauer_2022}
+Ross-Hellauer, T., Reichmann, S., Cole, N. L., Fessl, A., Klebel, T., \&
+Pontika, N. (2022). Dynamics of cumulative advantage and threats to
+equity in open science: A scoping review. \emph{Royal Society Open
+Science}, \emph{9}(1). \url{https://doi.org/10.1098/rsos.211032}
+
+\bibitem[\citeproctext]{ref-Schimmer_2015}
+Schimmer, R., Geschuhn, K., \& Vogler, A. (2015). \emph{{Disrupting the
+subscription journals'business model for the necessary large-scale
+transformation to open access}}. Max Planck Digital Library.
+\url{https://doi.org/10.17617/1.3}
+
+\bibitem[\citeproctext]{ref-Schmal_competition}
+Schmal, W. B. (2024). \emph{La révolution dévore ses enfants: Pricing
+implications of transformative agreements}. arXiv.
+\url{https://doi.org/10.48550/ARXIV.2403.03597}
 
@@
 Simard, M.-A., Basson, I., Hare, M., Lariviere, V., \& Mongeon, P.
-(2024). \emph{The open access coverage of OpenAlex, scopus and web of
-science}. \url{https://doi.org/10.48550/ARXIV.2404.01985}
+(2024). \emph{The open access coverage of {OpenAlex}, {Scopus} and {Web
+of Science}}. \url{https://doi.org/10.48550/ARXIV.2404.01985}
 
@@
 Stahlschmidt, S., \& Stephen, D. (2022). From indexation policies
-through citation networks to normalized citation impacts: Web of
-science, scopus, and dimensions as varying resonance chambers.
+through citation networks to normalized citation impacts: {Web of
+Science}, {Scopus}, and {Dimensions} as varying resonance chambers.
 \emph{Scientometrics}, \emph{127}(5), 2413--2431.
@@
 
+\bibitem[\citeproctext]{ref-tidyverse}
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
+Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
+Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
+Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
+\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
+
+\bibitem[\citeproctext]{ref-Widding_2024}
+Widding, A. S. (2024). Beyond transformative agreements: Ways forward
+for universities. \emph{European Review}, \emph{32}(S1), S28--S38.
+\url{https://doi.org/10.1017/s1062798724000036}
+
 \bibitem[\citeproctext]{ref-Zhang_2024}

### 2025-04-15 analysis/manuscript.qmd
Incorporate feedback, in particular from Mikael concerning the framing, and Birgit

@@
 
-Transformative agreements are a much-discussed library licensing model to transition subscription-based journal publishing to full open access.
-While these agreements can vary substantially, they mainly target hybrid journal portfolios, enabling researchers from participating institutions to publish open access in them while also providing reading access to the entire portfolio.
-Responding to experiences with the subscription model, library consortia and funders have begun to coordinate the standardisation of negotiation, administration, and evaluation practices for these agreements.
-Simultaneously, efforts to improve transparency through open data about contracts and payments are ongoing. 
-However, this information remains only partially available through bibliometric databases, limiting large-scale studies about the impact of transformative agreements on the transition of hybrid journals to full open access.
-This study compares findings from hoaddata, an openly available, continuously updated dataset on hybrid open access, built on open data from Crossref, OpenAlex and the cOAlition S Journal Checker Tool, with the established bibliometric databases Web of Science and Scopus.
-The purpose of this comparison is not only to evaluate the suitability of open metadata for measuring hybrid open access but also to inform bibliometric research and practice by highlighting the strengths and weaknesses of each data source.
-
-Since the initial proposal to further develop hybrid open access as a means to cost-effectively repurpose subscription spending toward open access [@Schimmer_2015], transformative agreements have led to considerable academic debate.
-While standardisation activities, particularly through the OA2020 and ESAC (Efficiency and Standards for Article Charges) initiatives, have led to improved workflows between library consortia and publishers, as well as increased open access publishing [@ifla; @Geschuhn_2017], the initial focus on a few large publishers by national consortia from high-income countries has drawn criticism [@Borrego_2020].
-Critics have in particular raised concerns about market concentration [@Shu_2023; @Butler_2023], the failure of most hybrid journals to convert to full OA [@Momeni_2021; @Kiley_2024], reduced competition and incentives [@McCabe_2024; @Schmal_competition], and the continuation of publication fees (APCs) that widen gaps between well-resourced and under-resourced institutions and researchers, raising equity concerns [@budapest; @Ross_Hellauer_2022].
-
-In terms of bibliometric evidence, studying over 700 agreements, @Jahn_2025 observed a strong growth in open access publishing to 15% in 2022 due to the varying implementation of transformative agreements across countries, predominantly driven by European countries with national consortia that successfully negotiated these agreements and the three commercial publishers Elsevier, Springer Nature, and Wiley.
-This aligns with the rapid proliferation of transformative agreements in recent years [@McCabe_2024].
-Before, only a small proportion of articles were made open access in hybrid journals [@Laakso_2016, @Piwowar_2018], with only few European countries with dedicated hybrid open access funding policies and the high-energy physics SCOAP^3 consortium standing out [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020; @Kohls_2018].
-
-While consortia evaluations confirm the growth in open access publishing through transformative agreements, they reach mixed policy conclusions.
-For instance, the British Joint Information Systems Committee's evaluation [@Jisc_2024] noted transformative agreements' significant impact on national open access growth but limited effect on global open access transition, suggesting a need to reevaluate their strategy.
+Transformative agreements are a much-discussed library licensing model for transitioning subscription-based journal publishing to full open access.
+Although these agreements can vary considerably, they are mainly aimed at hybrid journal bundles, allowing authors from participating institutions to publish open access in these journals, while providing reading access to the entire portfolio.
+Library consortia and funders have begun coordinating how these agreements are negotiated, managed and evaluated.
+At the same time, efforts are underway to improve transparency through open data on contracts and payments. 
+However, this information is still only partially available, which limits large-scale studies of the impact of transformative agreements on the transition of hybrid journals to full open access.
+By comparing open metadata from hoaddata, an openly available dataset on hybrid open access based on metadata from Crossref, OpenAlex and the cOAlition S Journal Checker Tool, with the established bibliometric databases Scopus and Web of Science, this study aims to provide a robust estimate of the impact of transformative agreements on hybrid open access.
+This comparative approach informs bibliometric research and practice by highlighting the strengths and weaknesses of each data source individually and collectively.
+
+Since their initial proposal by the Max Planck Digital Library [@Schimmer_2015], transformative agreements have led to a rapid growth in the number of negotiated agreements and open access articles [@Jahn_2025; @Rothfritz_2024; @McCabe_2024].
+Before, only a small proportion of articles in hybrid journals were openly available [@Bj_rk_2012; @Laakso_2016; @Piwowar_2018], with publications from only a few European countries with dedicated funding policies and the high-energy physics SCOAP^3 consortium standing out [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020; @Kohls_2018].
+Related coordination activities, in particular through the OA2020 and ESAC (Efficiency and Standards for Article Charges) initiatives, have also led to improved workflows between library consortia and publishers [@ifla; @Geschuhn_2017].
+However, critics have raised concerns about these agreements, particularly regarding market concentration [@Shu_2023; @Butler_2023], the failure of most hybrid journals to convert to full open access [@Momeni_2021; @Kiley_2024], reduced competition and incentives [@McCabe_2024; @Schmal_competition], and continuation of publication fees that widen gaps between well-resourced and under-resourced institutions [@budapest; @Ross_Hellauer_2022]
+
+While consortia evaluations confirm the growth of open access through transformative agreements and acknowledge increasing coordination, they reach mixed policy conclusions.
+For instance, the British Jisc evaluation [@Jisc_2024] found that the consortium's transformative agreements had a significant impact on national open access growth but a limited effect on the global open access transition, suggesting a need to reevaluate their strategy.
 Similar policy reconsiderations emerged in Norway and Sweden [@norway; @Widding_2024].
-Meanwhile, cOAlition S recommended ending financial support for these agreements in 2024, but still considers hybrid open access compliant with funder open access policies.
-Conversely, Germany's DEAL consortium extended agreements with major publishers through 2028, while Colombia established Latin America's first transformative agreements [@Mu_oz_V_lez_2024].
-Presenting experiences at the Max Planck Digital Library, [@D_r_2025] outlines strategies to achieve full open access while controlling costs.
-
-Open research information can benefit from transformative agreements as library consortia also negotiate for more comprehensive open metadata to avoid commodification of data analytics capabilities needed to negotiate and evaluate agreements [@Aspesi_2020; @McCabe_2024].
+Meanwhile, the funder initiative cOAlition S recommended ending financial support for these agreements in 2024, but still considers hybrid open access to be compliant with funders' open access policies.
+Conversely, the German DEAL consortium extended agreements with major publishers until 2028, while also outside of Europe more and more transformative agreements are being put in place (for instance in Latin America, as described in @Mu_oz_V_lez_2024).
+Responding to criticism, the Max Planck Digital Library recently outlined strategies to evolve the original model, focusing on achieving full open access while controlling costs and addressing identified weaknesses [@D_r_2025]
+
+Despite controversies, transformative agreements can benefit the landscape of open research information as library consortia negotiate for comprehensive open metadata, helping avoid data analytics becoming a commodity controlled by publishers [@Aspesi_2020; @McCabe_2024].
 For instance, ESAC recommends using Crossref [@Hendricks_2020] to share metadata supporting workflows including open access licenses and funding information [@Geschuhn_2017]. 
@@
 
-Numerous open scholarly data services building upon Crossref benefit from these developments.
-A prominent example is OpenAlex, which has gained considerable attention by promising inclusivity, openness, and comprehensive coverage of publication metadata [@priem2022openalex].
-In a recent evaluation of OpenAlex, [@C_spedes_2025] reviewed the current evidence base about its strengths and shortcomings.
-They highlight both its increasing use in bibliometric research and adoption across research organisations.
-Comparative studies generally perceive OpenAlex as a reliable alternative to commercial databases [@alperin2024analysissuitabilityopenalexbibliometric; @culbert2024referencecoverageanalysisopenalex], although data quality issues were observed regarding important metadata fields needed to analyse open access, including evidence about open access status itself [@jahn2023analysing; @Simard_2024], as well as authors [@culbert2024referencecoverageanalysisopenalex], affiliations [@Zhang_2024], author roles [@Jahn_2025], and document types used to distinguish between different article types [@Haupka_2024].
-However, these studies emphasise the new possibilities of combining open metadata and OpenAlex's curation efforts in conjunction with growing awareness about the importance of open metadata as demonstrated by the Barcelona Declaration on Open Research Information.
-
-Against this background, this study evaluates the suitability of open metadata to measure the impact of transformative agreements on hybrid open access.
-Few studies have attempted this analysis due to data limitations and the need to combine multiple sources.
-@de_Jonge_2025 provides a notable exception by comparing transformative agreement data from the cOAlition S Journal Checker tool and OpenAlex with institutional invoicing records for publications funded by Dutch research funders through the Dutch library consortium.
-Their analysis found 89% accuracy in identifying open access articles enabled by transformative agreements, with discrepancies resulting from both open data limitations and the inherent complexity of agreement structures that require data from both library consortia and publishers—information that is rarely made publicly available.
-Recent large-scale studies [@Bakker_2024; @Haustein_APC_2024; @Kramer_2024; @Jahn_2025] confirm that these challenges extend beyond the Dutch context, highlighting monitoring difficulties related to varying journal portfolios, challenges in using author roles as a proxy for invoicing data, and article caps within agreements.
-
-The present study addresses these gaps by adapting the open approach used in @Jahn_2025, which combines cOAlition S Journal Checker tool data, Crossref and OpenAlex to measure the impact of transformative agreements on the open access uptake of hybrid journals, and applying this same methodology to Web of Science and Scopus data.
-This comparative approach enables assessment of different bibliometric data sources for monitoring transformative agreements while evaluating their respective strengths and limitations.
-Specifically, this study compares the indexing coverage of more than 13,000 hybrid journals across these sources and examines whether consistent results can be obtained regarding open access uptake between 2019 and 2023, analysed by publisher and country.
-By comparing findings across these platforms, this large-scale study aims to provide a more comprehensive understanding of how transformative agreements impact the ongoing transition from subscription-based journal publishing to full open access.
+Many open scholarly data services that build on Crossref benefit from these developments.
+A prominent example is OpenAlex, which has gained attention for its inclusivity, openness and comprehensive coverage of publication metadata [@priem2022openalex].
+In a recent evaluation, @C_spedes_2025 reviewed the current evidence base on the strengths and weaknesses of OpenAlex.
+They highlight both its increasing use in bibliometric research and its adoption by research organisations.
+Comparative studies have generally found OpenAlex to be a reliable alternative to commercial databases [@alperin2024analysissuitabilityopenalexbibliometric; @culbert2024referencecoverageanalysisopenalex], although data quality issues occurred for key metadata fields needed for open access analysis, including evidence of open access status itself [@jahn2023analysing; @Simard_2024], as well as authorships [@culbert2024referencecoverageanalysisopenalex], affiliations [@Zhang_2024], author roles [@Jahn_2025] and document types used to distinguish between different article types [@Haupka_2024].
+However, these studies highlight the rapid improvements and the new possibilities of combining open metadata and the curation efforts of OpenAlex, in conjunction with the growing awareness of the importance of open metadata, as demonstrated by the Barcelona Declaration on Open Research Information.
+
+As a consequence, there is growing interest in using open bibliometric databases such as Crossref and OpenAlex to assess the impact of transformative agreements. 
+Similar to established proprietary bibliometric databases such as Scopus and Web of Science, these data sources lack specific data points, particularly about open access payments, which are typically not shared [@Haustein_APC_2024].
+However, bibliometric data can be enriched with open metadata about the journals and institutions involved in transformative agreements [@Jahn_2025]. 
+@de_Jonge_2025 demonstrated this potential by combining data from the cOAlition S Journal Checker tool and OpenAlex, achieving 89% accuracy when validated against Dutch research information. 
+While open data approaches show promise, understanding how these results compare when the same methodology is applied to proprietary databases such as Web of Science and Scopus remains unexplored.
+
+This study compares open and proprietary data sources for evaluating transformative agreements.
+It adapts the approach from @Jahn_2025, which combines data from the cOAlition S Journal Checker tool, Crossref and OpenAlex, and applies the same methods to Web of Science and Scopus data.
+This comparative approach allows the evaluation of different bibliometric data sources for monitoring transformative agreements, while assessing their respective strengths and limitations.
+Specifically, this study compares the indexing of more than 13,000 hybrid journals across these sources, and examines whether consistent results can be obtained in terms of open access adoption between 2019 and 2023, analysed by publisher and country.
+By comparing results across these data sources, this large-scale study provides a more robust understanding of transformative agreements' impact on the transition to full open access. 
+It also reveals the comparative strengths and weaknesses of different bibliometric databases in measuring scholarly publishing's shift toward openness
 
@@
 
-This study investigates the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access.
+<!-- This study investigates the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access. -->
 As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
@@
 These collections provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall [@Piwowar_2018], the same provider as OpenAlex.
-However, the Web of Science lacks information about journals and articles under transformative agreements.
+However, Web of Science lacks information about journals and articles under transformative agreements.
 
@@
 Elsevier's Scopus, launched in 2004, is another widely used proprietary bibliometric database for measuring research [@Baas_2020].
-Similar to the Web of Science, Scopus is selective with regard to the journals it indexes.
+Similar to Web of Science, Scopus is selective with regard to the journals it indexes.
 However, its journal coverage is much broader than that of the Web of Science collections considered in this study, as it also indexes a wider range of applied research journals [@Singh_2021; @Visser_2021; @Stahlschmidt_2022].
@@
 
-This study used the Scopus annual snapshot of April 2024 as provided by the KB (scp_b_202404).
-The same KB curation effort was applied to the  Scopus raw data as for the Web of Science [@schmidt_2024_13935407].
+This study used the Scopus annual snapshot of April 2024 as provided by the KB (scp_b_202404). The same KB curation effort as for the Web of Science was applied to the Scopus [@schmidt_2024_13935407].
 
@@
 The JCT journal data were enriched with ISSN variants linked to an ISSN-L. 
-To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists (DOAJ, OpenAlex, Bielefeld).
-These data were made available via hoaddata and used to determine the publication volume of hybrid journals for each database independently.
+To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists including the Directory of Open Access Journals (DOAJ).
+The resulting hybrid journal data were made available via hoaddata and used to determine the publication volume for each database independently.
 
@@
 This additional parsing remains necessary despite transformative agreement workflows recommend to deposit CC licenses during DOI registration [@Geschuhn_2017].
-Both Web of Science and Scopus defined hybrid open access consistently as content available under CC licenses on publisher platforms according to their documentations[^wosoa][^scopusoa], distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses [@Piwowar_2018].
+Both Web of Science[^wosoa] and Scopus[^scopusoa] defined hybrid open access consistently as content available under CC licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses [@Piwowar_2018].
 
@@
 Author affiliations were retrieved for both first and, if available, corresponding authors to prepare the linking between articles and institutions covered by transformative agreements.
-To improve the data retrieval, JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
+To improve data retrieval, JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
 To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifier.
 Additionally, ISO country codes were retrieved for each author's address to compile country-level statistics. 
-These country codes for Web of Science and Scopus were provided by the KB.
-
-Because neither Web of Science nor Scopus support ROR at the time of the data retrieval, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise affiliation data.
+<!-- These country codes for Web of Science and Scopus were provided by the KB.-->
+
+Because neither Web of Science nor Scopus supported ROR-ID at the time of the data retrieval, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise affiliation data.
 First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
@@
 
-This process linked `r length(unique(scp_ror$ror))` ROR IDs to  `r length(unique(scp_ror$vendor_org_id))` Scopus Affiliation IDs, and  `r length(unique(wos_ror$ror))` ROR IDs to `r length(unique(wos_ror$vendor_org_id))` enhanced affiliation strings in the Web of Science.
+This process linked `r length(unique(scp_ror$ror))` ROR-IDs to  `r length(unique(scp_ror$vendor_org_id))` Scopus Affiliation IDs, and  `r length(unique(wos_ror$ror))` ROR-IDs to `r length(unique(wos_ror$vendor_org_id))` enhanced affiliation strings in Web of Science.
 Quality evaluation through random sampling of 50 pairs revealed an error rate of 22% for Web of Science (11 mismatches) and 6% for Scopus (3 mismatches). 
 Upon inspection, these mismatches primarily occurred with less-represented institutions having only a few publications, introduced through multiple affiliations of single authors. 
-The difference between databases suggests that Scopus' affiliation control aligns more closely with ROR than that of the Web of Science.
+The difference between databases suggests that Scopus' affiliation control aligns more closely with ROR than that of Web of Science.
 
@@
 
-Based on these so-compiled matching tables, articles eligible under transformative agreements could also be obtained from Web of Science and Scopus, although they did not contain the ROR IDs used by the JCT.
+Based on these compiled matching tables, articles eligible under transformative agreements could also be obtained from Web of Science and Scopus, although they did not contain ROR-IDs used by the JCT.
 The estimation of eligible articles followed @Jahn_2025 and included a matching of both journals and participating institutions according to the JCT.
@@
 @tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023.
-It shows that the majority of hybrid journals published at least one original research article or review marked as original during the five-year period.
-These journals formed the basis for the subsequent calculation of article-level indicators.
+It shows that the majority of hybrid journals published at least one original research article or review, collectively marked as "original article" in the following, during the five-year period.
+These journals provided the basis for the subsequent calculation of article-level indicators.
 
@@
 This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
-Open access indicators were aggregated by DOI, as Unpaywall only collects information on open access status for articles with a DOI.
+Open access indicators were aggregated by DOI, because Unpaywall only collects information on open access status for articles with a DOI.
 A closer look at original articles with affiliation data, and in line with related research [@Zhang_2024], reveals a lack of affiliation data in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
@@
 For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(hoad) * 100`%.
-At the time of writing, OpenAlex disclosed limited coverage of corresponding authorhip data [^oacorresponding].
+At the time of writing, OpenAlex disclosed limited coverage of corresponding authorship data [^oacorresponding].
 Therefore, only first author data for hoaddata were considered in the following analysis.
@@
 
-This section first presents the indexing coverage of hybrid open access by comparing open data sources with the proprietary bibliometric databases Scopus and Web of Science. 
+This section first presents the indexing of hybrid open access by comparing open data sources with the proprietary bibliometric databases Scopus and Web of Science. 
 Then, using the same methods, indicators at the publisher and country level are calculated independently for each database and compared with each other to assess the suitability of the bibliometric databases for investigating transformative agreements.
@@
 
-Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid journals included in transformative agreements, visualising the intersections of journals and articles across the examined databases as an UpSet graph [@Lex_2014; @ComplexUpset].
+Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid journals included in transformative agreements, based on the intersections of the examined data sources. 
+The intersection sets of journals and articles are visualised as an UpSet graph [@Lex_2014; @ComplexUpset].
 The analysis included hybrid journals that published at least one open access article between 2019 and 2023, based on open access status information from each database.
-Only original research articles and reviews were considered in the analysis.
+Only original articles were considered in the analysis.
 
 ```{r upset_coverage_results, fig.asp=0.75}
-#| fig-cap: "Comparison of the hybrid open access indexing coverage of hoaddata (based on Crossref), Scopus and Web of Science, 2019-2023.
+#| fig-cap: "Comparison of hybrid journal indexing by data source, 2019-2023.
 Only hybrid journals present in the cOAlition S Transformative Agreement Data dump with at least one open access article are considered.
@@
 The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
-Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaddata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in the proprietory database Scopus.
-Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in the Web of Science.
+Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaddata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in Scopus.
+Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in Web of Science.
 Upon inspection, this group of hybrid journals exclusively covered in the proprietary data sources mainly represented hybrid journals for which no open access evidence could be retrieved from Crossref, the open access evidence source for hoaddata.
@@
 In terms of article coverage, Figure \ref{fig-upset_coverage_results}B shows the total publication volume per combination in terms of DOI availability.
-The largest set of hybrid journals, which includes all three data sources, also contains the largest number of articles. 
-In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` overlapping articles, represented by the blue bar.
-They represented `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(wos) * 100)`% of original articles with DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(scp) * 100)`% in Scopus.
-Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to both Scopus and hoaddata.
+The largest set of hybrid journals, covering all three data sources, also represents the largest number of articles. 
+In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` articles, represented by the blue bar.
+They represented `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(wos) * 100)`% of original articles with a DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(scp) * 100)`% in Scopus.
+Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to the intersection of Scopus and hoaddata.
 Exclusively in hoaddata were `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == FALSE) |> pull(n_cr) |> sum()` articles, and exclusively in the proprietary databases were `r upset_data |> filter(Crossref == FALSE, Scopus == TRUE) |> pull(overlap) |> sum(na.rm = TRUE)` articles.
@@
 Figure \ref{fig-upset_coverage_results}B also shows the surplus of articles with DOI that were only available via hoaddata (grey area).
-In case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
-After validation at the DOI level using the KB databases and manual inspection, the main reasons for missing DOI coverage in the proprietary database were insufficient classification of journal content as original research articles and reviews during the compilation of hoaddata.
+In the case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
+After validation at the DOI level using the KB databases and manual inspection, the main reason for missing DOI coverage in the proprietary database were insufficient classification of journal content as original articles during the compilation of hoaddata.
 Particularly, letters and editorials could not be fully detected.
-Paratext recognition failed in `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scholarly content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science.
+Moreover, paratext recognition failed for `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scholarly content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science.
 To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.
@@
 Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
-Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows a large spread across the journals jointly covered by all three data sources. 
-These journals published more on average than journals covered by only one or two data sources. 
-In particular, the journals covered exclusively by hoaddata were substantially smaller. 
+Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows a large spread oj journal size across the journals jointly covered by all three data sources. 
+These journals published more on average than journals covered by only one or two data sources, with the exception of journals exclusively covered by Scopus and Web of Science. 
+In particular, journals covered exclusively by hoaddata were substantially smaller. 
 Upon inspection, these were often newly launched hybrid journals, which explains the relatively low five-year publication volume.
@@
 In terms of article volume, Elsevier led with `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_art)*100`% of the total) published across `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_jn)*100`% of the total).
-Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)`, although recording the largest number of hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%).
+Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals, although recording the largest number of hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%).
 Wiley accounted for `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
@@
 ```{r upset_publisher_plot, fig.asp=0.75}
-#| fig-cap: "Comparison of the hybrid open access indexing coverage of hoaddata (based on Crossref), Scopus and Web of Science, 2019-2023 by publisher.
+#| fig-cap: "Comparison of hybrid journal indexing by data source and publisher, 2019-2023.
 Only hybrid journals present in the cOAlition S Transformative Agreement Data dump with at least one open access article are considered.
@@
 
-The three largest publishers, Elsevier, Springer Nature, and Wiley, were best represented in the exclusive intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
+The three largest publishers, Elsevier, Springer Nature, and Wiley, were best represented in the  intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
 Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage (n = `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()`; `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as determined through shared DOIs.
@@
 
-Comparing publisher portfolios across different indexing sets demonstrates that publisher were not represented uniformly. 
+Comparing publisher portfolios across different indexing sets demonstrates that publishers were not represented uniformly. 
 Notably, Springer Nature exhibited `r jn_upset_publisher |> filter(esac_publisher == "Springer Nature", WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow()` hybrid journals exclusively indexed in both hoaddata and Scopus.
 This set included journals from the Chinese Academy of Science, German-language medical journals, and Eastern European publications including the *Journal of Mathematical Sciences*, which also publishes English-language translations of Russian-language works.
-Additionally, this subset included titles with a more broader disciplinary focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
-The set also captured ceased journals, providing further insights into the dynamics of journal publishing.
+Additionally, this subset included titles with a broader disciplinary focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
+The set also captured discontinued journals, providing further insights into the dynamics of journal publishing.
 
@@
 
-Examining publisher portfolios not covered by hoaddata but present in Scopus or Web of Science identified several publishers with missing CC license information in Crossref.
+Examining publisher portfolios not covered by hoaddata but present in Scopus or Web of Science identified several publishers with missing Creative Commons (CC) license information in Crossref.
 In particular, `r miss_tas$esac_publisher[1]` represented `r miss_tas$journals[1]` journals with `r miss_tas$n[1]` articles, `r miss_tas$esac_publisher[2]` accounted for `r miss_tas$journals[2]` journals with `r miss_tas$n[2]` articles, and `r miss_tas$esac_publisher[3]` recorded `r miss_tas$journals[3]` journals and `r miss_tas$n[3]` articles.
@@
 
-An inspection of individual journals also uncovered discrepancies in Unpaywall's open access identification for certain publishers that typically deposit CC licenses with Crossref.
+An inspection of individual journals also uncovered discrepancies in Unpaywall's open access identification for certain publishers that typically share CC license metadata with Crossref.
 Notably, some subscription-only journals contained one or two articles erroneously tagged as hybrid open access by Unpaywall, which were subsequently reflected in Scopus and Web of Science.
@@
 
-## Open Access Indicator Comparision
+## Open Access Indicator Comparison
 
@@
 ```{r, plot_uptake_overview, fig.asp=.5}
-#| fig-cap: "Development of open access in hybrid journals included in transformative agreements, 2019 and 2023 by data source and author role.
-Figure shows the open access percentage (A), the number (B) and percentage (C) of open access articles enabled transformative agreements."
+#| fig-cap: "Development of open access in hybrid journals in transformative agreements by data source and author role, 2019-2023.
+Figure shows the open access percentage (A), the number (B) and the percentage (C) of open access articles enabled by transformative agreements."
 #| label: fig-uptake_overview
@@
 According to hoaddata, hybrid open access uptake increased from `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
-Similarly, Scopus recorded an growth from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
-However, Web of Science recorded higher open access uptake in early years, before converging in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023, suggesting an different approach towards labeling hybrid open access by the Web of Science.
-
-Similarly, hybrid open access by transformative agreements increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
+Similarly, Scopus recorded a growth from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
+However, Web of Science recorded higher open access uptake in early years, before converging to a similar level in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
+This suggest a different approach towards labelling hybrid open access by Web of Science.
+
+Similarly, hybrid open access by transformative agreements substantially increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
 Trends were consistent when measuring first (dashed line) and corresponding author (dotted line) affiliations. 
@@
 Web of Science recorded `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_first_author_articles) |> sum()` open access articles via transformative agreements using first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`)  and  `r uptake_by_year_short |> filter(src == "Web of Science") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
-hoaddata, lacking corresponding author data, linked `r uptake_by_year_short |> filter(src == "HOAD") |> pull(ta_oa_first_author_articles) |> sum()` articles to transformative agreements but showed slower growth than Scopus and Web of Science, from `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` in 2019 to `r uptake_by_year_short |> filter(src == "HOAD", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()` in 2023.
-
-From 2021 (hoadata, Scopus) resp. 2022 (Web of Science),  transformative agreements enabled the majority of hybrid open access.
+hoaddata, lacking corresponding author data, linked `r uptake_by_year_short |> filter(src == "hoaddata") |> pull(ta_oa_first_author_articles) |> sum()` articles to transformative agreements, increasing from `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` in 2019 to `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()` in 2023.
+
+From 2021 (hoadata, Scopus) resp. 2022 (Web of Science),  transformative agreements enabled the majority of open access articles in hybrid journals.
 For first authors, the share ranged between `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and   `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science) in 2023.
 For corresponding authors, the shares were slightly larger, with Scopus recording  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% and Web of Science `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% in 2023.
-However, substantial hybrid open access was still facilitated outside transformative agreements, likely through APCs paid from discretionary research funds [@Jahn_2021; @Suber_2012].
+However, substantial hybrid open access was still facilitated outside transformative agreements, likely through APCs paid from discretionary research funds [@Suber_2012].
 
@@
 ```{r plot_publisher_uptake_overview, fig.asp=1.1}
-#| fig-cap: "Development of open access in hybrid journals included in transformative agreements, 2019 and 2023, by data source, author role and publisher.
-Figure shows the open access percentage (A), the number (B) and percentage (C) of open access articles enabled transformative agreements."
+#| fig-cap: "Development of open access in hybrid journals included in transformative agreements by data source, author role and publisher, 2019-2023.
+Figure shows the open access percentage (A), the number (B) and the percentage (C) of open access articles enabled by transformative agreements."
 #| label: fig-uptake_publisher
@@
 When considering open access trends by publisher (see Figure \ref{fig-uptake_publisher}), the observed differences in early uptake rates between hoaddata and Scopus compared to Web of Science can be largely attributed to articles published in Elsevier hybrid journals, the largest publisher in our sample.
-Both hoadata and Scopus reported an steady increase in open access uptake between 2019 and 2023 (hoaddata from `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%; Scopus from `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%).
-In contrast, Elsevier's share remained relatively constant, increasing only slightly from `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`% according to the Web of Science.
-Upon inspection, this discrepancy primarily stemmed from articles in the publisher's open archive. 
+Both hoadata and Scopus reported a steady increase in open access uptake between 2019 and 2023 (hoaddata from `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "HOAD", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%; Scopus from `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Scopus", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`%).
+In contrast, Elsevier's share remained relatively constant, increasing only slightly from `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`% according to Web of Science.
+Upon inspection, this discrepancy is primarily due to articles in the publisher's open archive. 
 These articles, made freely available after an embargo period under Elsevier's user license, were tagged as hybrid open access in Web of Science, even though its documentation[^wosoa] specified that only articles under a CC license variant were considered.
@@
 For these two years, hoaddata reported `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Scopus", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more open access articles than Scopus and `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Web of Science", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more than Web of Science.
-This discrepancy is presumably because of challenges in fetching full-texts by Unpaywall, the open access evidence source for Scopus and Web of Science.
-According to Unpaywall's software version history, Wiley's publisher platform redirects prevented Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
+This discrepancy is presumably due to challenges in fetching full-texts by Unpaywall, the open access evidence source for Scopus and Web of Science.
+According to Unpaywall's software version history, HTTP redirects from Wiley's publisher platform prevented Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
 hoaddata, which relies solely on Crossref metadata for open access identification, was unaffected by these issues.
@@
 Despite these differences in open access evidence, the three data sources show consistent temporal trends in hybrid open access enabled by transformative agreements (see Figure \ref{fig-uptake_publisher}B and C).
-Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across the examined data sources, followed by Springer Nature and Elsevier, recording a comparable late uptake, which is consistent with the publisher's historical reluctance to engage in negotations with library consortia.
+Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across the examined data sources, followed by Springer Nature.
+Elsevier, recorded a later uptake, which is consistent with the publisher's historical reluctance to engage in negotiations with library consortia.
 However, by 2023, the share of open access enabled by transformative agreements appeared to stabilise for all three publishers (see Figure \ref{fig-uptake_publisher}C). 
 Interestingly, the differences between first and corresponding author affiliations were more pronounced at the publisher level. 
-In Scopus, for example, the share of open access via transformative agreements measured by corresponding authorship was greater for Springer Nature in 2023 than when using first authorship.
+In Scopus, for example, the share of open access via transformative agreements measured by corresponding authorship was higher for Springer Nature in 2023 than when using first authorship.
 
@@
 #| fig-cap: "Comparison of hybrid open access by country, 2019-2023.
-Scatterplots distinguish between proprietary database Scopus and Web of Science, and author role.
-x axis shows hoaddata indicators.
+Scatterplots distinguish between proprietary databases Scopus and Web of Science, and author role.
+The x-axis shows hoaddata indicators.
 A) Five-years article volume, B) open access percentage in hybrid journals, (A and B limited to countries with > 10.000 articles published), C) number and D) percentage of open access articles enabled by transformative agreements (limited to countries with > 1.000 open access articles enabled by transformative agreements).
 Line represents line of equality.
-An interactive version is accessible here: <https://najkoja-hoa-replication.hf.space/>"
+An interactive version is accessible via: <https://najkoja-hoa-replication.hf.space/>"
 #| label: fig-uptake_country
@@
 
-In terms of article output by country (see Figure \ref{fig-uptake_country}A), strong positive correlation was observed across the data sources and author roles (Spearman rank correlation $\rho > .9$, $p < 0.001$).
+In terms of article output by country (see Figure \ref{fig-uptake_country}A), strong positive correlations were observed across the data sources and author roles (Spearman rank correlation $\rho > .9$, $p < 0.001$).
 Between 2019 and 2023, China was the most productive country, followed by the United States and, by a certain margin, India, the United Kingdom, and Germany. 
@@
 OECD members accounted for the majority of open access articles enabled by transformative agreements.
-As a notable exception, South Africa also featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward.[^sa]
+As a notable exception, South Africa again featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward.[^sa]
 Results were consistent across data sources. 
@@
 In all cases, strong positive correlations were observed using Spearman's rank correlation: 
-$\rho > .9$, $p < 0.001$ between data sources and authorship roles, when considering countries with an minimum of 1,000 open access articles enabled by transformative agreements between 2019 and 2023.
-When this limitation was removed, the correlation remained strong but slightly lower ($\rho > .85$, $p < 0.001$).
+$\rho > .9$, $p < 0.001$ between data sources and authorship roles, when considering countries with a minimum of 1,000 open access articles enabled by transformative agreements between 2019 and 2023. 
+When this limitation was removed, the correlation remained strong ($\rho > .85$, $p < 0.001$).
 This difference may signal countries where only a few institutions had transformative agreements in place, as opposed to those participating in national consortia with broader participation.
@@
 
-This study of over 13,000 hybrid journals shows a rise in open access due to transformative agreements between 2019 and 2023, although most articles remained paywalled.
+This study of over 13,000 hybrid journals shows a substantial rise in open access due to transformative agreements between 2019 and 2023, although most articles remained paywalled.
 While transformative agreements accounted for the majority of open access, many articles continue to become open through the payment of individual publication fees.
@@
 with high-income European countries—alongside South Africa—showing high adoption rates.
-In contrast, the three most productive countries, China, the United States, and India, were substantially ranked lower in transformative agreement adoption.
+In contrast, the three most productive countries, China, the United States, and India, showcase a substantially lower adoption in transformative agreements.
 Open questions remain whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or a deliberate avoidance of such agreements.
@@
 The findings were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science. 
-Notably, and aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73], the results show strong correlations by country affiliation, which supports using open metadata for large-scale analyses of hybrid open access.
-However, the observed differences in journal coverage and metadata availability warrant discussion, affecting not only open data sources but also proprietary databases when used in isolation.
+Notably, and aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73].
+Overall, the results show strong correlations by country affiliation, which supports using open metadata for large-scale analyses of hybrid open access. 
+However, the observed differences in journal coverage and metadata availability warrant further discussion in the following, affecting not only open data sources but also proprietary databases when used in isolation.
 
@@
 Comparing the coverage of OpenAlex, Scopus and Web of Science, @Simard_2024 indicate that only half of the fully open access journals listed in the DOAJ are also indexed in Scopus and Web of Science.
-Notably, journals that charge no publication fees ("diamond journals") are absent from the selective Web of Science, which reinforces existing disparities in the indexing of under-represented research communities and regions in selective bibliometric databases [@Simard_2024]
+Notably, journals that charge no publication fees ("diamond journals") are absent from the selective Web of Science, which reinforces existing disparities in the indexing of under-represented research communities and regions in selective bibliometric databases [@Simard_2024].
 
@@
 <!-- Thus, first authorship can reliably be used in large-scale studies to assess the impact of transformative agreements. -->
-Despite this correlation, the study observed a recent slight increase in open access articles by corresponding authors over first authors across Springer Nature hybrid journals prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
-Previous research has highlighted how institutionalized bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], suggesting the need to monitor potential influences of the availability of open access funding on authorships [@Maddi_2024].
+Despite this correlation, the study observed a recent slight increase in open access articles by corresponding authors over first authors across Springer Nature hybrid journals.
+This prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
+Previous research has highlighted how institutionalised bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], suggesting the need to monitor potential influences of the availability of open access funding on authorship roles [@Maddi_2024].
 
@@
 Although OpenAlex's affiliation coverage is less comprehensive, which likely reduced the number of articles confidently attributed to transformative agreements in this study, it still shows high correlations with Scopus and Web of Science at the country level.
-However, OpenAlex's native ROR ID integration offers a distinct advantage, enabling more reliable identification of agreement-enabled articles compared to Scopus and Web of Science, which require reconciliation with proprietary organization identifiers.
+However, OpenAlex's native ROR-ID integration offers a distinct advantage, allowing a more reliable identification of agreement-enabled articles compared to Scopus and Web of Science, which require reconciliation with proprietary organisation identifiers.
 Future studies based on the Web of Science will benefit from the recent integration of ROR-IDs, announced by the end of 2024.
@@
 The database comparison revealed important discrepancies in open access evidence. 
-Crucially, not all publishers deposit CC licences with Crossref, a limitation that becomes apparent when contrasting Crossref with Unpaywall’s data in Scopus and Web of Science. 
+Crucially, not all publishers share CC licence metadata via Crossref, a limitation that becomes apparent when contrasting Crossref with Unpaywall’s data in Scopus and Web of Science. 
 While Unpaywall can detect such gaps by parsing publishers' websites for open access licences, it nevertheless missed a substantial number of CC-licensed open access articles from Wiley journals indexed in Crossref, likely due to parsing errors on the publisher's website.
@@
 It is important to note that the study's estimates of articles from institutions involved in transformative agreements are approximations due to a lack of access to invoice data, which is not usually shared by library consortia and publishers.
-Furthermore, the data sources about transformative agreements including the cOAlition S Journal Checker Tool and the ESAC registry are a voluntary effort, crowd-sourced from various consortia. 
+Furthermore, the data sources about transformative agreements including the cOAlition S Journal Checker Tool and the ESAC registry are based on a voluntary effort, crowd-sourced from various consortia.
 However, recent validation with Dutch research information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements [@de_Jonge_2025]. 
-With enhanced open metadata compliance, attributable to evolving standards and initiatives to support negotiations with publishers, particularly through the communities of the ESAC initiative and Barcelona Declaration supporters, the situation is likely to improve.
-In conclusion, this study has shown that open metadata are suitable for analysing the transition from subscription-based journal publishing to full open access.
+With enhanced open metadata compliance, attributable to evolving standards and initiatives in support of negotiations with publishers, particularly through the communities of the ESAC initiative and Barcelona Declaration supporters, the situation is likely to improve.
+In conclusion, this study has shown that open metadata are well-suited for analysing the transition from subscription-based journal publishing to full open access.
 

### 2025-04-15 analysis/manuscript.tex
Incorporate feedback, in particular from Mikael concerning the framing, and Birgit

@@
 \begin{large}
-\textbf{Replication: Hybrid Open Access in Transformative Agreements}
+\textbf{Estimating transformative agreement impact on hybrid open access: A comparative large-scale study using open metadata, Scopus and Web of Science}
 \end{large}
@@
 Transformative agreements are a much-discussed library licensing model
-to transition subscription-based journal publishing to full open access.
-While these agreements can vary substantially, they mainly target hybrid
-journal portfolios, enabling researchers from participating institutions
-to publish open access in them while also providing reading access to
-the entire portfolio. Responding to experiences with the subscription
-model, library consortia and funders have begun to coordinate the
-standardisation of negotiation, administration, and evaluation practices
-for these agreements. Simultaneously, efforts to improve transparency
-through open data about contracts and payments are ongoing. However,
-this information remains only partially available through bibliometric
-databases, limiting large-scale studies about the impact of
-transformative agreements on the transition of hybrid journals to full
-open access. This study compares findings from hoaddata, an openly
-available, continuously updated dataset on hybrid open access, built on
-open data from Crossref, OpenAlex and the cOAlition S Journal Checker
-Tool, with the established bibliometric databases Web of Science and
-Scopus. The purpose of this comparison is not only to evaluate the
-suitability of open metadata for measuring hybrid open access but also
-to inform bibliometric research and practice by highlighting the
-strengths and weaknesses of each data source.
-
-Since the initial proposal to further develop hybrid open access as a
-means to cost-effectively repurpose subscription spending toward open
-access (Schimmer et al., 2015), transformative agreements have led to
-considerable academic debate. While standardisation activities,
-particularly through the OA2020 and ESAC (Efficiency and Standards for
-Article Charges) initiatives, have led to improved workflows between
-library consortia and publishers, as well as increased open access
-publishing (Campbell et al., 2022; Geschuhn \& Stone, 2017), the initial
-focus on a few large publishers by national consortia from high-income
-countries has drawn criticism (Borrego et al., 2021). Critics have in
-particular raised concerns about market concentration (Butler et al.,
-2023; Shu \& Larivière, 2023), the failure of most hybrid journals to
-convert to full OA (Kiley, 2024; Momeni et al., 2021), reduced
-competition and incentives (McCabe \& Mueller-Langer, 2024; Schmal,
-2024), and the continuation of publication fees (APCs) that widen gaps
-between well-resourced and under-resourced institutions and researchers,
-raising equity concerns (Babini et al., 2022; Ross-Hellauer et al.,
-2022).
-
-In terms of bibliometric evidence, studying over 700 agreements, Jahn
-(2025) observed a strong growth in open access publishing to 15\% in
-2022 due to the varying implementation of transformative agreements
-across countries, predominantly driven by European countries with
-national consortia that successfully negotiated these agreements and the
-three commercial publishers Elsevier, Springer Nature, and Wiley. This
-aligns with the rapid proliferation of transformative agreements in
-recent years (McCabe \& Mueller-Langer, 2024). Before, only a small
-proportion of articles were made open access in hybrid journals Piwowar
-et al. (2018), with only few European countries with dedicated hybrid
-open access funding policies and the high-energy physics SCOAP\^{}3
-consortium standing out (Huang et al., 2020; Kohls \& Mele, 2018;
-Pinfield, 2015; Robinson-Garcia et al., 2020).
-
-While consortia evaluations confirm the growth in open access publishing
-through transformative agreements, they reach mixed policy conclusions.
-For instance, the British Joint Information Systems Committee's
-evaluation (Brayman et al., 2024) noted transformative agreements'
-significant impact on national open access growth but limited effect on
-global open access transition, suggesting a need to reevaluate their
-strategy. Similar policy reconsiderations emerged in Norway and Sweden
-(Holden et al., 2023; Widding, 2024). Meanwhile, cOAlition S recommended
+for transitioning subscription-based journal publishing to full open
+access. Although these agreements can vary considerably, they are mainly
+aimed at hybrid journal bundles, allowing authors from participating
+institutions to publish open access in these journals, while providing
+reading access to the entire portfolio. Library consortia and funders
+have begun coordinating how these agreements are negotiated, managed and
+evaluated. At the same time, efforts are underway to improve
+transparency through open data on contracts and payments. However, this
+information is still only partially available, which limits large-scale
+studies of the impact of transformative agreements on the transition of
+hybrid journals to full open access. By comparing open metadata from
+hoaddata, an openly available dataset on hybrid open access based on
+metadata from Crossref, OpenAlex and the cOAlition S Journal Checker
+Tool, with the established bibliometric databases Scopus and Web of
+Science, this study aims to provide a robust estimate of the impact of
+transformative agreements on hybrid open access. This comparative
+approach informs bibliometric research and practice by highlighting the
+strengths and weaknesses of each data source individually and
+collectively.
+
+Since their initial proposal by the Max Planck Digital Library (Schimmer
+et al., 2015), transformative agreements have led to a rapid growth in
+the number of negotiated agreements and open access articles (Jahn,
+2025; McCabe \& Mueller-Langer, 2024; Rothfritz et al., 2024). Before,
+only a small proportion of articles in hybrid journals were openly
+available (Björk, 2012; Laakso \& Björk, 2016; Piwowar et al., 2018),
+with publications from only a few European countries with dedicated
+funding policies and the high-energy physics SCOAP\^{}3 consortium
+standing out (Huang et al., 2020; Kohls \& Mele, 2018; Pinfield, 2015;
+Robinson-Garcia et al., 2020). Related coordination activities, in
+particular through the OA2020 and ESAC (Efficiency and Standards for
+Article Charges) initiatives, have also led to improved workflows
+between library consortia and publishers (Campbell et al., 2022;
+Geschuhn \& Stone, 2017). However, critics have raised concerns about
+these agreements, particularly regarding market concentration (Butler et
+al., 2023; Shu \& Larivière, 2023), the failure of most hybrid journals
+to convert to full open access (Kiley, 2024; Momeni et al., 2021),
+reduced competition and incentives (McCabe \& Mueller-Langer, 2024;
+Schmal, 2024), and continuation of publication fees that widen gaps
+between well-resourced and under-resourced institutions (Babini et al.,
+2022; Ross-Hellauer et al., 2022)
+
+While consortia evaluations confirm the growth of open access through
+transformative agreements and acknowledge increasing coordination, they
+reach mixed policy conclusions. For instance, the British Jisc
+evaluation (Brayman et al., 2024) found that the consortium's
+transformative agreements had a significant impact on national open
+access growth but a limited effect on the global open access transition,
+suggesting a need to reevaluate their strategy. Similar policy
+reconsiderations emerged in Norway and Sweden (Holden et al., 2023;
+Widding, 2024). Meanwhile, the funder initiative cOAlition S recommended
 ending financial support for these agreements in 2024, but still
-considers hybrid open access compliant with funder open access policies.
-Conversely, Germany's DEAL consortium extended agreements with major
-publishers through 2028, while Colombia established Latin America's
-first transformative agreements (Muñoz-Vélez et al., 2024). Presenting
-experiences at the Max Planck Digital Library, (Dér, 2025) outlines
-strategies to achieve full open access while controlling costs.
-
-Open research information can benefit from transformative agreements as
-library consortia also negotiate for more comprehensive open metadata to
-avoid commodification of data analytics capabilities needed to negotiate
-and evaluate agreements (Aspesi \& Brand, 2020; McCabe \&
+considers hybrid open access to be compliant with funders' open access
+policies. Conversely, the German DEAL consortium extended agreements
+with major publishers until 2028, while also outside of Europe more and
+more transformative agreements are being put in place (for instance in
+Latin America, as described in Muñoz-Vélez et al. (2024)). Responding to
+criticism, the Max Planck Digital Library recently outlined strategies
+to evolve the original model, focusing on achieving full open access
+while controlling costs and addressing identified weaknesses (Dér, 2025)
+
+Despite controversies, transformative agreements can benefit the
+landscape of open research information as library consortia negotiate
+for comprehensive open metadata, helping avoid data analytics becoming a
+commodity controlled by publishers (Aspesi \& Brand, 2020; McCabe \&
 Mueller-Langer, 2024). For instance, ESAC recommends using Crossref
@@
 
-Numerous open scholarly data services building upon Crossref benefit
-from these developments. A prominent example is OpenAlex, which has
-gained considerable attention by promising inclusivity, openness, and
-comprehensive coverage of publication metadata (Priem et al., 2022). In
-a recent evaluation of OpenAlex, (Céspedes et al., 2025) reviewed the
-current evidence base about its strengths and shortcomings. They
-highlight both its increasing use in bibliometric research and adoption
-across research organisations. Comparative studies generally perceive
-OpenAlex as a reliable alternative to commercial databases (Alperin et
-al., 2024; Culbert et al., 2024), although data quality issues were
-observed regarding important metadata fields needed to analyse open
-access, including evidence about open access status itself (Jahn et al.,
-2023; Simard et al., 2024), as well as authors (Culbert et al., 2024),
-affiliations (Zhang et al., 2024), author roles (Jahn, 2025), and
-document types used to distinguish between different article types
-(Haupka et al., 2024). However, these studies emphasise the new
-possibilities of combining open metadata and OpenAlex's curation efforts
-in conjunction with growing awareness about the importance of open
-metadata as demonstrated by the Barcelona Declaration on Open Research
+Many open scholarly data services that build on Crossref benefit from
+these developments. A prominent example is OpenAlex, which has gained
+attention for its inclusivity, openness and comprehensive coverage of
+publication metadata (Priem et al., 2022). In a recent evaluation,
+Céspedes et al. (2025) reviewed the current evidence base on the
+strengths and weaknesses of OpenAlex. They highlight both its increasing
+use in bibliometric research and its adoption by research organisations.
+Comparative studies have generally found OpenAlex to be a reliable
+alternative to commercial databases (Alperin et al., 2024; Culbert et
+al., 2024), although data quality issues occurred for key metadata
+fields needed for open access analysis, including evidence of open
+access status itself (Jahn et al., 2023; Simard et al., 2024), as well
+as authorships (Culbert et al., 2024), affiliations (Zhang et al.,
+2024), author roles (Jahn, 2025) and document types used to distinguish
+between different article types (Haupka et al., 2024). However, these
+studies highlight the rapid improvements and the new possibilities of
+combining open metadata and the curation efforts of OpenAlex, in
+conjunction with the growing awareness of the importance of open
+metadata, as demonstrated by the Barcelona Declaration on Open Research
 Information.
 
-Against this background, this study evaluates the suitability of open
-metadata to measure the impact of transformative agreements on hybrid
-open access. Few studies have attempted this analysis due to data
-limitations and the need to combine multiple sources. Jonge et al.
-(2025) provides a notable exception by comparing transformative
-agreement data from the cOAlition S Journal Checker tool and OpenAlex
-with institutional invoicing records for publications funded by Dutch
-research funders through the Dutch library consortium. Their analysis
-found 89\% accuracy in identifying open access articles enabled by
-transformative agreements, with discrepancies resulting from both open
-data limitations and the inherent complexity of agreement structures
-that require data from both library consortia and
-publishers---information that is rarely made publicly available. Recent
-large-scale studies (Bakker et al., 2024; Haustein et al., 2024; Jahn,
-2025; Kramer, 2024) confirm that these challenges extend beyond the
-Dutch context, highlighting monitoring difficulties related to varying
-journal portfolios, challenges in using author roles as a proxy for
-invoicing data, and article caps within agreements.
-
-The present study addresses these gaps by adapting the open approach
-used in Jahn (2025), which combines cOAlition S Journal Checker tool
-data, Crossref and OpenAlex to measure the impact of transformative
-agreements on the open access uptake of hybrid journals, and applying
-this same methodology to Web of Science and Scopus data. This
-comparative approach enables assessment of different bibliometric data
-sources for monitoring transformative agreements while evaluating their
-respective strengths and limitations. Specifically, this study compares
-the indexing coverage of more than 13,000 hybrid journals across these
-sources and examines whether consistent results can be obtained
-regarding open access uptake between 2019 and 2023, analysed by
-publisher and country. By comparing findings across these platforms,
-this large-scale study aims to provide a more comprehensive
-understanding of how transformative agreements impact the ongoing
-transition from subscription-based journal publishing to full open
-access.
+As a consequence, there is growing interest in using open bibliometric
+databases such as Crossref and OpenAlex to assess the impact of
+transformative agreements. Similar to established proprietary
+bibliometric databases such as Scopus and Web of Science, these data
+sources lack specific data points, particularly about open access
+payments, which are typically not shared (Haustein et al., 2024).
+However, bibliometric data can be enriched with open metadata about the
+journals and institutions involved in transformative agreements (Jahn,
+2025). Jonge et al. (2025) demonstrated this potential by combining data
+from the cOAlition S Journal Checker tool and OpenAlex, achieving 89\%
+accuracy when validated against Dutch research information. While open
+data approaches show promise, understanding how these results compare
+when the same methodology is applied to proprietary databases such as
+Web of Science and Scopus remains unexplored.
+
+This study compares open and proprietary data sources for evaluating
+transformative agreements. It adapts the approach from Jahn (2025),
+which combines data from the cOAlition S Journal Checker tool, Crossref
+and OpenAlex, and applies the same methods to Web of Science and Scopus
+data. This comparative approach allows the evaluation of different
+bibliometric data sources for monitoring transformative agreements,
+while assessing their respective strengths and limitations.
+Specifically, this study compares the indexing of more than 13,000
+hybrid journals across these sources, and examines whether consistent
+results can be obtained in terms of open access adoption between 2019
+and 2023, analysed by publisher and country. By comparing results across
+these data sources, this large-scale study provides a more robust
+understanding of transformative agreements' impact on the transition to
+full open access. It also reveals the comparative strengths and
+weaknesses of different bibliometric databases in measuring scholarly
+publishing's shift toward openness
 
@@
 
-This study investigates the suitability of open scholarly data sources
-for assessing the impact of transformative agreements on hybrid open
-access. As shown in Figure \ref{fig:workflow}, the methodology involved
+As shown in Figure \ref{fig:workflow}, the methodology involved
 comparing hoaddata, an openly available collection of open research
@@
 status information derived from OurResearch's Unpaywall (Piwowar et al.,
-2018), the same provider as OpenAlex. However, the Web of Science lacks
+2018), the same provider as OpenAlex. However, Web of Science lacks
 information about journals and articles under transformative agreements.
@@
 bibliometric database for measuring research (Baas et al., 2020).
-Similar to the Web of Science, Scopus is selective with regard to the
+Similar to Web of Science, Scopus is selective with regard to the
 journals it indexes. However, its journal coverage is much broader than
@@
 This study used the Scopus annual snapshot of April 2024 as provided by
-the KB (scp\_b\_202404). The same KB curation effort was applied to the
-Scopus raw data as for the Web of Science (Schmidt et al., 2024).
+the KB (scp\_b\_202404). The same KB curation effort as for the Web of
+Science was applied to the Scopus (Schmidt et al., 2024).
 
@@
 ISSN-L. To identify hybrid journals, a comprehensive exclusion of fully
-open access journals was performed using multiple journal lists (DOAJ,
-OpenAlex, Bielefeld). These data were made available via hoaddata and
-used to determine the publication volume of hybrid journals for each
-database independently.
+open access journals was performed using multiple journal lists
+including the Directory of Open Access Journals (DOAJ). The resulting
+hybrid journal data were made available via hoaddata and used to
+determine the publication volume for each database independently.
 
@@
 agreement workflows recommend to deposit CC licenses during DOI
-registration (Geschuhn \& Stone, 2017). Both Web of Science and Scopus
+registration (Geschuhn \& Stone, 2017). Both Web of Science\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
+and Scopus\footnote{\url{https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option}}
 defined hybrid open access consistently as content available under CC
-licenses on publisher platforms according to their
-documentations\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}\footnote{\url{https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option}},
-distinguishing it from bronze open access that lack such explicit
-license information, or use publisher-specific licenses (Piwowar et al.,
-2018).
+licenses on publisher platforms, distinguishing it from bronze open
+access that lack such explicit license information, or use
+publisher-specific licenses (Piwowar et al., 2018).
 
@@
 corresponding authors to prepare the linking between articles and
-institutions covered by transformative agreements. To improve the data
+institutions covered by transformative agreements. To improve data
 retrieval, JCT institution data was enriched with ROR-IDs from
@@
 retrieved for each author's address to compile country-level statistics.
-These country codes for Web of Science and Scopus were provided by the
-KB.
-
-Because neither Web of Science nor Scopus support ROR at the time of the
-data retrieval, the institution identifier used by the JCT, a two-step
-matching process was implemented to harmonise affiliation data. First,
-2,782,540 articles from 6,457 institutions with ROR-IDs in the JCT data
-since 2017 (according to hoaddata) were processed to map first authors'
-ROR-IDs to corresponding proprietary affiliation identifier in Web of
-Science and Scopus using DOI matching. Then, an algorithm selected the
-most frequent ROR ID and proprietary identifier pairs to handle multiple
-affiliations and organisational hierarchy differences.
-
-This process linked 6,375 ROR IDs to 4,894 Scopus Affiliation IDs, and
-6,034 ROR IDs to 2,422 enhanced affiliation strings in the Web of
-Science. Quality evaluation through random sampling of 50 pairs revealed
-an error rate of 22\% for Web of Science (11 mismatches) and 6\% for
-Scopus (3 mismatches). Upon inspection, these mismatches primarily
-occurred with less-represented institutions having only a few
-publications, introduced through multiple affiliations of single
-authors. The difference between databases suggests that Scopus'
-affiliation control aligns more closely with ROR than that of the Web of
-Science.
+
+Because neither Web of Science nor Scopus supported ROR-ID at the time
+of the data retrieval, the institution identifier used by the JCT, a
+two-step matching process was implemented to harmonise affiliation data.
+First, 2,782,540 articles from 6,457 institutions with ROR-IDs in the
+JCT data since 2017 (according to hoaddata) were processed to map first
+authors' ROR-IDs to corresponding proprietary affiliation identifier in
+Web of Science and Scopus using DOI matching. Then, an algorithm
+selected the most frequent ROR ID and proprietary identifier pairs to
+handle multiple affiliations and organisational hierarchy differences.
+
+This process linked 6,375 ROR-IDs to 4,894 Scopus Affiliation IDs, and
+6,034 ROR-IDs to 2,422 enhanced affiliation strings in Web of Science.
+Quality evaluation through random sampling of 50 pairs revealed an error
+rate of 22\% for Web of Science (11 mismatches) and 6\% for Scopus (3
+mismatches). Upon inspection, these mismatches primarily occurred with
+less-represented institutions having only a few publications, introduced
+through multiple affiliations of single authors. The difference between
+databases suggests that Scopus' affiliation control aligns more closely
+with ROR than that of Web of Science.
 
@@
 
-Based on these so-compiled matching tables, articles eligible under
+Based on these compiled matching tables, articles eligible under
 transformative agreements could also be obtained from Web of Science and
-Scopus, although they did not contain the ROR IDs used by the JCT. The
+Scopus, although they did not contain ROR-IDs used by the JCT. The
 estimation of eligible articles followed Jahn (2025) and included a
@@
 majority of hybrid journals published at least one original research
-article or review marked as original during the five-year period. These
-journals formed the basis for the subsequent calculation of
-article-level indicators.
+article or review, collectively marked as ``original article'' in the
+following, during the five-year period. These journals provided the
+basis for the subsequent calculation of article-level indicators.
 
@@
 indexed by Scopus (Donner, 2017). Open access indicators were aggregated
-by DOI, as Unpaywall only collects information on open access status for
-articles with a DOI. A closer look at original articles with affiliation
-data, and in line with related research (Zhang et al., 2024), reveals a
-lack of affiliation data in the case of OpenAlex, the affiliation data
-source used by hoaddata, compared to Web of Science and Scopus. In
-particular, only about two-thirds of the articles examined provided
-corresponding author affiliations. For first authors, the proportion was
-89\%. At the time of writing, OpenAlex disclosed limited coverage of
-corresponding authorhip data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
+by DOI, because Unpaywall only collects information on open access
+status for articles with a DOI. A closer look at original articles with
+affiliation data, and in line with related research (Zhang et al.,
+2024), reveals a lack of affiliation data in the case of OpenAlex, the
+affiliation data source used by hoaddata, compared to Web of Science and
+Scopus. In particular, only about two-thirds of the articles examined
+provided corresponding author affiliations. For first authors, the
+proportion was 89\%. At the time of writing, OpenAlex disclosed limited
+coverage of corresponding authorship data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
 Therefore, only first author data for hoaddata were considered in the
@@
 
-This section first presents the indexing coverage of hybrid open access
-by comparing open data sources with the proprietary bibliometric
-databases Scopus and Web of Science. Then, using the same methods,
-indicators at the publisher and country level are calculated
-independently for each database and compared with each other to assess
-the suitability of the bibliometric databases for investigating
-transformative agreements.
+This section first presents the indexing of hybrid open access by
+comparing open data sources with the proprietary bibliometric databases
+Scopus and Web of Science. Then, using the same methods, indicators at
+the publisher and country level are calculated independently for each
+database and compared with each other to assess the suitability of the
+bibliometric databases for investigating transformative agreements.
 
@@
 Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid
-journals included in transformative agreements, visualising the
-intersections of journals and articles across the examined databases as
-an UpSet graph (Krassowski, 2020; Lex et al., 2014). The analysis
-included hybrid journals that published at least one open access article
-between 2019 and 2023, based on open access status information from each
-database. Only original research articles and reviews were considered in
-the analysis.
+journals included in transformative agreements, based on the
+intersections of the examined data sources. The intersection sets of
+journals and articles are visualised as an UpSet graph (Krassowski,
+2020; Lex et al., 2014). The analysis included hybrid journals that
+published at least one open access article between 2019 and 2023, based
+on open access status information from each database. Only original
+articles were considered in the analysis.
 
@@
 
-\caption{\label{fig-upset_coverage_results}Comparison of the hybrid open
-access indexing coverage of hoaddata (based on Crossref), Scopus and Web
-of Science, 2019-2023. Only hybrid journals present in the cOAlition S
-Transformative Agreement Data dump with at least one open access article
-are considered. A) presents the number of journals, B) the number of
-articles (DOI), distinguishing between shared DOI corpus and surplus in
-hoaddata. Box plots (C) shows the five-years journal article volume
-(log-scale).}
+\caption{\label{fig-upset_coverage_results}Comparison of hybrid journal
+indexing by data source, 2019-2023. Only hybrid journals present in the
+cOAlition S Transformative Agreement Data dump with at least one open
+access article are considered. A) presents the number of journals, B)
+the number of articles (DOI), distinguishing between shared DOI corpus
+and surplus in hoaddata. Box plots (C) shows the five-years journal
+article volume (log-scale).}
 
@@
 of journals were exclusively contained in hoaddata, while another 6\% (n
-= 748) were only found in the proprietory database Scopus. Of these, 354
-were also available in the Web of Science. Upon inspection, this group
-of hybrid journals exclusively covered in the proprietary data sources
-mainly represented hybrid journals for which no open access evidence
-could be retrieved from Crossref, the open access evidence source for
-hoaddata.
+= 748) were only found in Scopus. Of these, 354 were also available in
+Web of Science. Upon inspection, this group of hybrid journals
+exclusively covered in the proprietary data sources mainly represented
+hybrid journals for which no open access evidence could be retrieved
+from Crossref, the open access evidence source for hoaddata.
 
@@
 shows the total publication volume per combination in terms of DOI
-availability. The largest set of hybrid journals, which includes all
-three data sources, also contains the largest number of articles. In
-total, these journals recorded 6,289,687 overlapping articles,
-represented by the blue bar. They represented 94\% of original articles
-with DOI indexed in the Web of Science sample, and 86\% in Scopus.
-Another 657,697 articles were exclusive to both Scopus and hoaddata.
-Exclusively in hoaddata were 177,110 articles, and exclusively in the
-proprietary databases were 325,194 articles.
+availability. The largest set of hybrid journals, covering all three
+data sources, also represents the largest number of articles. In total,
+these journals recorded 6,289,687 articles, represented by the blue bar.
+They represented 94\% of original articles with a DOI indexed in the Web
+of Science sample, and 86\% in Scopus. Another 657,697 articles were
+exclusive to the intersection of Scopus and hoaddata. Exclusively in
+hoaddata were 177,110 articles, and exclusively in the proprietary
+databases were 325,194 articles.
 
@@
 articles with DOI that were only available via hoaddata (grey area). In
-case of hybrid journals covered by all three data sources, 1,023,882
+the case of hybrid journals covered by all three data sources, 1,023,882
 DOIs were only present in hoaddata. After validation at the DOI level
-using the KB databases and manual inspection, the main reasons for
+using the KB databases and manual inspection, the main reason for
 missing DOI coverage in the proprietary database were insufficient
-classification of journal content as original research articles and
-reviews during the compilation of hoaddata. Particularly, letters and
-editorials could not be fully detected. Paratext recognition failed in
-37\% of DOIs to identify non-scholarly content such as front matters or
-reviewer lists, which are generally not indexed by Scopus and Web of
-Science. To a lesser extent, differences in publication and indexing
-dates were a reason for non-overlapping DOIs.
+classification of journal content as original articles during the
+compilation of hoaddata. Particularly, letters and editorials could not
+be fully detected. Moreover, paratext recognition failed for 37\% of
+DOIs to identify non-scholarly content such as front matters or reviewer
+lists, which are generally not indexed by Scopus and Web of Science. To
+a lesser extent, differences in publication and indexing dates were a
+reason for non-overlapping DOIs.
 
@@
 \ref{fig-upset_coverage_results}C illustrates the distribution for each
-combination. It shows a large spread across the journals jointly covered
-by all three data sources. These journals published more on average than
-journals covered by only one or two data sources. In particular, the
-journals covered exclusively by hoaddata were substantially smaller.
-Upon inspection, these were often newly launched hybrid journals, which
-explains the relatively low five-year publication volume. An example is
-\emph{Digital society} that published 86 articles. This hybrid journal
-was launched in 2022, being covered by various Springer Nature
-transformative agreements since then.
+combination. It shows a large spread oj journal size across the journals
+jointly covered by all three data sources. These journals published more
+on average than journals covered by only one or two data sources, with
+the exception of journals exclusively covered by Scopus and Web of
+Science. In particular, journals covered exclusively by hoaddata were
+substantially smaller. Upon inspection, these were often newly launched
+hybrid journals, which explains the relatively low five-year publication
+volume. An example is \emph{Digital society} that published 86 articles.
+This hybrid journal was launched in 2022, being covered by various
+Springer Nature transformative agreements since then.
 
@@
 highlights the dominance of the three largest publishers, Elsevier,
-Springer Nature, and Wiley, which collectively accounted for 47\% of
+Springer Nature, and Wiley, which collectively accounted for 46\% of
 hybrid journals and 62\% of articles published during this five-year
-period. In terms of article volume, Elsevier led with 2,441,358 articles
-(33\% of the total) published across 1,951 hybrid journals (16\% of the
-total). Springer Nature followed with 1,247,578 articles (17\%) in
-2,311, although recording the largest number of hybrid journals (19\%).
-Wiley accounted for 858,939 articles (12\%) in 1,382 hybrid journals
-(11\%). The remaining 54 publishers collectively accounted for 2,749,847
-articles (38\%) in 6,476 hybrid journals (53\%).
+period. In terms of article volume, Elsevier led with 2,435,478 articles
+(34\% of the total) published across 1,941 hybrid journals (16\% of the
+total). Springer Nature followed with 1,237,700 articles (17\%) in 2,294
+hybrid journals, although recording the largest number of hybrid
+journals (19\%). Wiley accounted for 814,349 articles (11\%) in 1,337
+hybrid journals (11\%). The remaining 54 publishers collectively
+accounted for 2,733,994 articles (38\%) in 6,442 hybrid journals (54\%).
 
@@
 
-\caption{\label{fig-upset_coverage_results_publisher}Comparison of the
-hybrid open access indexing coverage of hoaddata (based on Crossref),
-Scopus and Web of Science, 2019-2023 by publisher. Only hybrid journals
-present in the cOAlition S Transformative Agreement Data dump with at
-least one open access article are considered. A) presents the percentage
-of journals by publisher, B) the number of articles by publisher (shared
-DOI). Box plots (C) shows the five-years journal article volume
-(log-scale) by publisher.}
+\caption{\label{fig-upset_coverage_results_publisher}Comparison of
+hybrid journal indexing by data source and publisher, 2019-2023. Only
+hybrid journals present in the cOAlition S Transformative Agreement Data
+dump with at least one open access article are considered. A) presents
+the percentage of journals by publisher, B) the number of articles by
+publisher (shared DOI). Box plots (C) shows the five-years journal
+article volume (log-scale) by publisher.}
 
@@
 The three largest publishers, Elsevier, Springer Nature, and Wiley, were
-best represented in the exclusive intersection of all three data sources
-(hoaddata, Scopus, and Web of Science). Together, they comprised 4,384
+best represented in the intersection of all three data sources
+(hoaddata, Scopus, and Web of Science). Together, they comprised 4,347
 hybrid journals (55\% of the intersectional set) and dominated article
-coverage (n = 4,174,315; 66\%), as determined through shared DOIs. When
+coverage (n = 4,143,663; 66\%), as determined through shared DOIs. When
 examining publication volume per journal, (Figure
@@
 Comparing publisher portfolios across different indexing sets
-demonstrates that publisher were not represented uniformly. Notably,
-Springer Nature exhibited 519 hybrid journals exclusively indexed in
+demonstrates that publishers were not represented uniformly. Notably,
+Springer Nature exhibited 517 hybrid journals exclusively indexed in
 both hoaddata and Scopus. This set included journals from the Chinese
@@
 Russian-language works. Additionally, this subset included titles with a
-more broader disciplinary focus such as \emph{SN Computer Science} and
-newly launched hybrid journals like \emph{Nature Computational Science},
-which started in 2021 and was indexed in Scopus but not yet in Web of
-Science. The set also captured ceased journals, providing further
-insights into the dynamics of journal publishing.
+broader disciplinary focus such as \emph{SN Computer Science} and newly
+launched hybrid journals like \emph{Nature Computational Science}, which
+started in 2021 and was indexed in Scopus but not yet in Web of Science.
+The set also captured discontinued journals, providing further insights
+into the dynamics of journal publishing.
 
 Examining publisher portfolios not covered by hoaddata but present in
-Scopus or Web of Science identified several publishers with missing CC
-license information in Crossref. In particular, Emerald represented 322
-journals with 86,409 articles, AIP Publishing accounted for 24 journals
-with 64,898 articles, and World Scientific recorded 87 journals and
-42,531 articles. In total, 9 publishers did not share CC licenses with
-Crossref and were therefore not represented in hoaddata.
+Scopus or Web of Science identified several publishers with missing
+Creative Commons (CC) license information in Crossref. In particular,
+Emerald represented 321 journals with 86,269 articles, AIP Publishing
+accounted for 24 journals with 64,898 articles, and World Scientific
+recorded 87 journals and 42,531 articles. In total, 9 publishers did not
+share CC licenses with Crossref and were therefore not represented in
+hoaddata.
 
@@
 Unpaywall's open access identification for certain publishers that
-typically deposit CC licenses with Crossref. Notably, some
+typically share CC license metadata with Crossref. Notably, some
 subscription-only journals contained one or two articles erroneously
@@
 \subsection{Open Access Indicator
-Comparision}\label{open-access-indicator-comparision}
+Comparison}\label{open-access-indicator-comparison}
 
@@
 \caption{\label{fig-uptake_overview}Development of open access in hybrid
-journals included in transformative agreements, 2019 and 2023 by data
-source and author role. Figure shows the open access percentage (A), the
-number (B) and percentage (C) of open access articles enabled
-transformative agreements.}
+journals in transformative agreements by data source and author role,
+2019-2023. Figure shows the open access percentage (A), the number (B)
+and the percentage (C) of open access articles enabled by transformative
+agreements.}
 
@@
 hoaddata, hybrid open access uptake increased from 6.4\% (n = 85,071) in
-2019 to 19\% (n = 302,358) in 2023. Similarly, Scopus recorded an growth
+2019 to 19\% (n = 302,358) in 2023. Similarly, Scopus recorded a growth
 from 6.5\% (n = 84,648) in 2019 to 19\% (n = 322,850) in 2023. However,
 Web of Science recorded higher open access uptake in early years, before
-converging in 2023, from 11\% (n = 137,202) in 2019 to 20\% (n =
-255,481) in 2023, suggesting an different approach towards labeling
-hybrid open access by the Web of Science.
-
-Similarly, hybrid open access by transformative agreements increased
-between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C). Trends
-were consistent when measuring first (dashed line) and corresponding
-author (dotted line) affiliations. According to Scopus, 479,297 open
-access articles could be attributed to transformative agreements based
-on first author metadata (increasing from 15,341 to 206,084) and 489,262
-using corresponding author metadata (from 15,444 to 210,816). Web of
-Science recorded 493,028 open access articles via transformative
-agreements using first author metadata (increasing from 21,871 to
-174,126) and 500,076 using corresponding author metadata (from 22,092 to
-177,030). hoaddata, lacking corresponding author data, linked 0 articles
-to transformative agreements but showed slower growth than Scopus and
-Web of Science, from 0 in 2019 to 0 in 2023.
+converging to a similar level in 2023, from 11\% (n = 137,202) in 2019
+to 20\% (n = 255,481) in 2023. This suggest a different approach towards
+labelling hybrid open access by Web of Science.
+
+Similarly, hybrid open access by transformative agreements substantially
+increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and
+C). Trends were consistent when measuring first (dashed line) and
+corresponding author (dotted line) affiliations. According to Scopus,
+479,297 open access articles could be attributed to transformative
+agreements based on first author metadata (increasing from 15,341 to
+206,084) and 489,262 using corresponding author metadata (from 15,444 to
+210,816). Web of Science recorded 493,028 open access articles via
+transformative agreements using first author metadata (increasing from
+21,871 to 174,126) and 500,076 using corresponding author metadata (from
+22,092 to 177,030). hoaddata, lacking corresponding author data, linked
+501,649 articles to transformative agreements, increasing from 16,010 in
+2019 to 183,757 in 2023.
 
 From 2021 (hoadata, Scopus) resp. 2022 (Web of Science), transformative
-agreements enabled the majority of hybrid open access. For first
-authors, the share ranged between 61\% (hoaddata), 64\% (Scopus), and 68
-\% (Web of Science) in 2023. For corresponding authors, the shares were
-slightly larger, with Scopus recording 65\% and Web of Science 69\% in
-2023. However, substantial hybrid open access was still facilitated
-outside transformative agreements, likely through APCs paid from
-discretionary research funds (Jahn et al., 2022; Suber, 2012).
+agreements enabled the majority of open access articles in hybrid
+journals. For first authors, the share ranged between 61\% (hoaddata),
+64\% (Scopus), and 68 \% (Web of Science) in 2023. For corresponding
+authors, the shares were slightly larger, with Scopus recording 65\% and
+Web of Science 69\% in 2023. However, substantial hybrid open access was
+still facilitated outside transformative agreements, likely through APCs
+paid from discretionary research funds (Suber, 2012).
 
@@
 \caption{\label{fig-uptake_publisher}Development of open access in
-hybrid journals included in transformative agreements, 2019 and 2023, by
-data source, author role and publisher. Figure shows the open access
-percentage (A), the number (B) and percentage (C) of open access
-articles enabled transformative agreements.}
+hybrid journals included in transformative agreements by data source,
+author role and publisher, 2019-2023. Figure shows the open access
+percentage (A), the number (B) and the percentage (C) of open access
+articles enabled by transformative agreements.}
 
@@
 largely attributed to articles published in Elsevier hybrid journals,
-the largest publisher in our sample. Both hoadata and Scopus reported an
+the largest publisher in our sample. Both hoadata and Scopus reported a
 steady increase in open access uptake between 2019 and 2023 (hoaddata
@@
 share remained relatively constant, increasing only slightly from 14\%
-to 15\% according to the Web of Science. Upon inspection, this
-discrepancy primarily stemmed from articles in the publisher's open
-archive. These articles, made freely available after an embargo period
-under Elsevier's user license, were tagged as hybrid open access in Web
-of Science, even though its documentation\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
+to 15\% according to Web of Science. Upon inspection, this discrepancy
+is primarily due to articles in the publisher's open archive. These
+articles, made freely available after an embargo period under Elsevier's
+user license, were tagged as hybrid open access in Web of Science, even
+though its documentation\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
 specified that only articles under a CC license variant were considered.
@@
 Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022
-compared to hoaddata. For these two years, hoaddata reported 35,308 more
-open access articles than Scopus and 32,491 more than Web of Science.
-This discrepancy is presumably because of challenges in fetching
-full-texts by Unpaywall, the open access evidence source for Scopus and
-Web of Science. According to Unpaywall's software version history,
-Wiley's publisher platform redirects prevented Unpaywall from parsing
-license information from full-texts.\footnote{See Unpaywall version
-  history related to Wiley fixes:
+compared to hoaddata. For these two years, hoaddata reported 32,445 more
+open access articles than Scopus and 29,854 more than Web of Science.
+This discrepancy is presumably due to challenges in fetching full-texts
+by Unpaywall, the open access evidence source for Scopus and Web of
+Science. According to Unpaywall's software version history, HTTP
+redirects from Wiley's publisher platform prevented Unpaywall from
+parsing license information from full-texts.\footnote{See Unpaywall
+  version history related to Wiley fixes:
   \url{https://github.com/search?q=repo\%3Aourresearch\%2Foadoi+wiley&type=commits}}
@@
 reported as open access in 2023 across the examined data sources,
-followed by Springer Nature and Elsevier, recording a comparable late
-uptake, which is consistent with the publisher's historical reluctance
-to engage in negotations with library consortia. However, by 2023, the
-share of open access enabled by transformative agreements appeared to
-stabilise for all three publishers (see Figure
-\ref{fig-uptake_publisher}C). Interestingly, the differences between
-first and corresponding author affiliations were more pronounced at the
-publisher level. In Scopus, for example, the share of open access via
-transformative agreements measured by corresponding authorship was
-greater for Springer Nature in 2023 than when using first authorship.
+followed by Springer Nature. Elsevier, recorded a later uptake, which is
+consistent with the publisher's historical reluctance to engage in
+negotiations with library consortia. However, by 2023, the share of open
+access enabled by transformative agreements appeared to stabilise for
+all three publishers (see Figure \ref{fig-uptake_publisher}C).
+Interestingly, the differences between first and corresponding author
+affiliations were more pronounced at the publisher level. In Scopus, for
+example, the share of open access via transformative agreements measured
+by corresponding authorship was higher for Springer Nature in 2023 than
+when using first authorship.
 
@@
 country, 2019-2023. Scatterplots distinguish between proprietary
-database Scopus and Web of Science, and author role. x axis shows
+databases Scopus and Web of Science, and author role. The x-axis shows
 hoaddata indicators. A) Five-years article volume, B) open access
@@
 transformative agreements). Line represents line of equality. An
-interactive version is accessible here:
+interactive version is accessible via:
 \url{https://najkoja-hoa-replication.hf.space/}}
@@
 In terms of article output by country (see Figure
-\ref{fig-uptake_country}A), strong positive correlation was observed
+\ref{fig-uptake_country}A), strong positive correlations were observed
 across the data sources and author roles (Spearman rank correlation
@@
 accounted for the majority of open access articles enabled by
-transformative agreements. As a notable exception, South Africa also
+transformative agreements. As a notable exception, South Africa again
 featured prominently, as the South African National Library and
@@
 Spearman's rank correlation: \(\rho > .9\), \(p < 0.001\) between data
-sources and authorship roles, when considering countries with an minimum
+sources and authorship roles, when considering countries with a minimum
 of 1,000 open access articles enabled by transformative agreements
 between 2019 and 2023. When this limitation was removed, the correlation
-remained strong but slightly lower (\(\rho > .85\), \(p < 0.001\)). This
-difference may signal countries where only a few institutions had
-transformative agreements in place, as opposed to those participating in
-national consortia with broader participation.
+remained strong (\(\rho > .85\), \(p < 0.001\)). This difference may
+signal countries where only a few institutions had transformative
+agreements in place, as opposed to those participating in national
+consortia with broader participation.
 
@@
 
-This study of over 13,000 hybrid journals shows a rise in open access
-due to transformative agreements between 2019 and 2023, although most
-articles remained paywalled. While transformative agreements accounted
-for the majority of open access, many articles continue to become open
-through the payment of individual publication fees. Hybrid open access
-and transformative agreements remain concentrated among a small group of
-large commercial publishers, with high-income European
-countries---alongside South Africa---showing high adoption rates. In
-contrast, the three most productive countries, China, the United States,
-and India, were substantially ranked lower in transformative agreement
-adoption. Open questions remain whether this uneven distribution
-reflects temporary implementation gaps, inherent inequities in the
-transformative agreement model, or a deliberate avoidance of such
+This study of over 13,000 hybrid journals shows a substantial rise in
+open access due to transformative agreements between 2019 and 2023,
+although most articles remained paywalled. While transformative
+agreements accounted for the majority of open access, many articles
+continue to become open through the payment of individual publication
+fees. Hybrid open access and transformative agreements remain
+concentrated among a small group of large commercial publishers, with
+high-income European countries---alongside South Africa---showing high
+adoption rates. In contrast, the three most productive countries, China,
+the United States, and India, showcase a substantially lower adoption in
+transformative agreements. Open questions remain whether this uneven
+distribution reflects temporary implementation gaps, inherent inequities
+in the transformative agreement model, or a deliberate avoidance of such
 agreements.
@@
 and aligning with previous studies (Akbaritabar et al., 2024; Alperin et
-al., 2024) and rankings (N. J. van Eck et al., 2024), the results show
-strong correlations by country affiliation, which supports using open
-metadata for large-scale analyses of hybrid open access. However, the
-observed differences in journal coverage and metadata availability
-warrant discussion, affecting not only open data sources but also
-proprietary databases when used in isolation.
+al., 2024) and rankings (N. J. van Eck et al., 2024). Overall, the
+results show strong correlations by country affiliation, which supports
+using open metadata for large-scale analyses of hybrid open access.
+However, the observed differences in journal coverage and metadata
+availability warrant further discussion in the following, affecting not
+only open data sources but also proprietary databases when used in
+isolation.
 
@@
 under-represented research communities and regions in selective
-bibliometric databases (Simard et al., 2024)
+bibliometric databases (Simard et al., 2024).
 
@@
 articles by corresponding authors over first authors across Springer
-Nature hybrid journals prompts questions about how institutional open
-access sponsorship practices influence author roles and assignments
+Nature hybrid journals. This prompts questions about how institutional
+open access sponsorship practices influence author roles and assignments
 within co-author teams, especially as funding opportunities vary
 (Gumpenberger et al., 2018). Previous research has highlighted how
-institutionalized bibliometric practices can affect the valuation of
+institutionalised bibliometric practices can affect the valuation of
 authorship positions (Helgesson, 2020), suggesting the need to monitor
 potential influences of the availability of open access funding on
-authorships (Maddi \& Silva, 2024).
+authorship roles (Maddi \& Silva, 2024).
 
@@
 high correlations with Scopus and Web of Science at the country level.
-However, OpenAlex's native ROR ID integration offers a distinct
-advantage, enabling more reliable identification of agreement-enabled
+However, OpenAlex's native ROR-ID integration offers a distinct
+advantage, allowing a more reliable identification of agreement-enabled
 articles compared to Scopus and Web of Science, which require
-reconciliation with proprietary organization identifiers. Future studies
+reconciliation with proprietary organisation identifiers. Future studies
 based on the Web of Science will benefit from the recent integration of
@@
 The database comparison revealed important discrepancies in open access
-evidence. Crucially, not all publishers deposit CC licences with
+evidence. Crucially, not all publishers share CC licence metadata via
 Crossref, a limitation that becomes apparent when contrasting Crossref
@@
 transformative agreements including the cOAlition S Journal Checker Tool
-and the ESAC registry are a voluntary effort, crowd-sourced from various
-consortia. However, recent validation with Dutch research information
-demonstrates the reliability of such an open approach for the assessment
-of articles under transformative agreements (Jonge et al., 2025). With
-enhanced open metadata compliance, attributable to evolving standards
-and initiatives to support negotiations with publishers, particularly
-through the communities of the ESAC initiative and Barcelona Declaration
-supporters, the situation is likely to improve. In conclusion, this
-study has shown that open metadata are suitable for analysing the
-transition from subscription-based journal publishing to full open
-access.
+and the ESAC registry are based on a voluntary effort, crowd-sourced
+from various consortia. However, recent validation with Dutch research
+information demonstrates the reliability of such an open approach for
+the assessment of articles under transformative agreements (Jonge et
+al., 2025). With enhanced open metadata compliance, attributable to
+evolving standards and initiatives in support of negotiations with
+publishers, particularly through the communities of the ESAC initiative
+and Barcelona Declaration supporters, the situation is likely to
+improve. In conclusion, this study has shown that open metadata are
+well-suited for analysing the transition from subscription-based journal
+publishing to full open access.
 
@@
 
-\bibitem[\citeproctext]{ref-Bakker_2024}
-Bakker, C., Langham-Putrow, A., \& Riegelman, A. (2024). Impact of
-transformative agreements on publication patterns: An analysis based on
-agreements from the ESAC registry. \emph{International Journal of
-Librarianship}, \emph{8}(4), 67--96.
-\url{https://doi.org/10.23974/ijol.2024.vol8.4.341}
-
 \bibitem[\citeproctext]{ref-vanbellen2024oligopolyacademicpublisherspersists}
@@
 \url{https://doi.org/10.1162/qss_a_00018}
+
+\bibitem[\citeproctext]{ref-Bj_rk_2012}
+Björk, B.-C. (2012). The hybrid model for open access publication of
+scholarly articles: A failed experiment? \emph{Journal of the American
+Society for Information Science and Technology}, \emph{63}(8),
+1496--1504. \url{https://doi.org/10.1002/asi.22709}
 
@@
 
-\bibitem[\citeproctext]{ref-Kramer_2024}
-Kramer, B. (2024). \emph{Study on scientific publishing in {Europe} --
-{Development}, diversity, and transparency of costs}. Publications
-Office of the European Union. \url{https://doi.org/doi/10.2777/89349}
-
 \bibitem[\citeproctext]{ref-ComplexUpset}
@@
 
+\bibitem[\citeproctext]{ref-Rothfritz_2024}
+Rothfritz, L., Schmal, W. B., \& Herb, U. (2024). \emph{Trapped in
+transformative agreements? A multifaceted analysis of \textgreater1,000
+contracts}. \url{https://arxiv.org/abs/2409.20224}
+
 \bibitem[\citeproctext]{ref-Schimmer_2015}

### 2025-04-16 analysis/manuscript.tex
Add basic usage notes

@@
 \begin{large}
-\textbf{Estimating transformative agreement impact on hybrid open access: A comparative large-scale study using open metadata, Scopus and Web of Science}
+\textbf{Estimating transformative agreement impact on hybrid open access: A comparative large-scale study using Scopus, Web of Science and open metadata}
 \end{large}

### 2025-04-16 data-raw/002_wos_ta_matching.R
clean up

@@ -1,147 +1,0 @@
-#' Web of Science matching with Transformative Agreement data
-#' 
-#' The aim is to link bibliographic records with transformative agreement data 
-#' to determine, which publications were from an eligible institution under a 
-#' specific TA
-#' 
-#' ## Data prep
-#' 
-#' First, I need to add ror ids to the wos dataset using our matching.
-#' 
-#' To Do: Keep this data point in KB.
-library(tidyverse)
-wos_df <- readr::read_csv("data-raw/wos_jct_affiliations.csv") 
-# Matching
-wos_ror <- readr::read_csv("data-raw/ror_wos_matching.csv") |>
-  distinct(ror, vendor_org_id)
-
-wos_matching <- wos_df |> 
-  left_join(wos_ror, by = "vendor_org_id")
-#' The results is downloaded as csv and uploaded to Google BiGQuery using gcloud 
-#' command line tools (because file is large)
-#' `gcloud storage cp ~/Documents/thesis/hoa_validation/data-raw/wos_matching_20240717.csv gs://bigschol`
-write_csv(wos_matching, "data-raw/wos_matching_20240717.csv")
-#'
-#' 
-#' ## Matching
-#' 
-#' First, we combine data about journals and participating institutions per
-#' transformative agreements according to the JCT. Note, ISSN-L and ROR were 
-#' enriched. The resulting temporary table is `esac_data`
-#' 
-#' Then, we create a subset of publications from eligible institutions by merging
-#' `esac_data` with `wos_matching_20240717`. To improve the performance of the query
-#' not all fields will be returned. Also, records from institutions without TA were 
-#' excluded before the matching as indicated by NA. 
-library(bigrquery)
-
-my_sql <- "WITH
-  esac_data AS (
-  SELECT
-    jct_jn.esac_id,
-    issn_l,
-    jct_inst.ror_main,
-    jct_inst.ror as ror_matching,
-    EXTRACT(YEAR
-    FROM
-      start_date) AS start_year,
-    -- Extracting the year from start_date
-    EXTRACT(YEAR
-    FROM
-      end_date) AS end_year,
-    -- Extracting the year from end_date
-  FROM
-    `subugoe-collaborative.hoaddata.jct_hybrid_jns` AS jct_jn
-  INNER JOIN
[…]

### 2025-04-16 analysis/manuscript.qmd
Some minor language fixes

@@
 Transformative agreements are a much-discussed library licensing model for transitioning subscription-based journal publishing to full open access.
-Although these agreements can vary considerably, they are mainly aimed at hybrid journal bundles, allowing authors from participating institutions to publish open access in these journals, while providing reading access to the entire portfolio.
-Library consortia and funders have begun coordinating how these agreements are negotiated, managed and evaluated.
+Although these agreements can vary considerably, they are mainly aimed at hybrid journal bundles, allowing authors from participating institutions to publish open access in these journals while providing reading access to the entire portfolio.
+Library consortia and funders have begun to coordinate how these agreements are negotiated, managed and evaluated.
 At the same time, efforts are underway to improve transparency through open data on contracts and payments. 
@@
 
-Since their initial proposal by the Max Planck Digital Library [@Schimmer_2015], transformative agreements have led to a rapid growth in the number of negotiated agreements and open access articles [@Jahn_2025; @Rothfritz_2024; @McCabe_2024].
+Since their initial proposal by the Max Planck Digital Library [@Schimmer_2015], transformative agreements have led to rapid growth in the number of negotiated agreements and open access articles [@Jahn_2025; @Rothfritz_2024; @McCabe_2024].
 Before, only a small proportion of articles in hybrid journals were openly available [@Bj_rk_2012; @Laakso_2016; @Piwowar_2018], with publications from only a few European countries with dedicated funding policies and the high-energy physics SCOAP^3 consortium standing out [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020; @Kohls_2018].
 Related coordination activities, in particular through the OA2020 and ESAC (Efficiency and Standards for Article Charges) initiatives, have also led to improved workflows between library consortia and publishers [@ifla; @Geschuhn_2017].
-However, critics have raised concerns about these agreements, particularly regarding market concentration [@Shu_2023; @Butler_2023], the failure of most hybrid journals to convert to full open access [@Momeni_2021; @Kiley_2024], reduced competition and incentives [@McCabe_2024; @Schmal_competition], and continuation of publication fees that widen gaps between well-resourced and under-resourced institutions [@budapest; @Ross_Hellauer_2022]
+However, critics have raised concerns about these agreements, particularly regarding market concentration [@Shu_2023; @Butler_2023], the failure of most hybrid journals to convert to full open access [@Momeni_2021; @Kiley_2024], reduced competition and incentives [@McCabe_2024; @Schmal_competition], and the continuation of publication fees that widen gaps between well-resourced and under-resourced institutions [@budapest; @Ross_Hellauer_2022]
 
@@
 Meanwhile, the funder initiative cOAlition S recommended ending financial support for these agreements in 2024, but still considers hybrid open access to be compliant with funders' open access policies.
-Conversely, the German DEAL consortium extended agreements with major publishers until 2028, while also outside of Europe more and more transformative agreements are being put in place (for instance in Latin America, as described in @Mu_oz_V_lez_2024).
-Responding to criticism, the Max Planck Digital Library recently outlined strategies to evolve the original model, focusing on achieving full open access while controlling costs and addressing identified weaknesses [@D_r_2025]
-
-Despite controversies, transformative agreements can benefit the landscape of open research information as library consortia negotiate for comprehensive open metadata, helping avoid data analytics becoming a commodity controlled by publishers [@Aspesi_2020; @McCabe_2024].
+Conversely, the German DEAL consortium extended agreements with major publishers until 2028, while also outside of Europe, more transformative agreements are being put in place (for instance in Latin America, as described in @Mu_oz_V_lez_2024).
+Responding to criticism, the Max Planck Digital Library recently outlined strategies to evolve the original model, focusing on achieving fully open access while controlling costs and addressing identified weaknesses [@D_r_2025]
+
+Despite controversies, transformative agreements can benefit the landscape of open research information as library consortia negotiate for comprehensive open metadata, helping avoid data analytics from becoming a commodity controlled by publishers [@Aspesi_2020; @McCabe_2024].
 For instance, ESAC recommends using Crossref [@Hendricks_2020] to share metadata supporting workflows including open access licenses and funding information [@Geschuhn_2017]. 
-Simultaneously, the bibliometric community pushes for more open metadata through this publisher-driven DOI registration platform and continuously monitors progress [@eck_crossref_2022].
-
-Many open scholarly data services that build on Crossref benefit from these developments.
+Simultaneously, the bibliometric community pushes for more open metadata through this publisher-driven DOI registration platform and continuously monitors the progress [@eck_crossref_2022].
+
+Many open scholarly data services that build on Crossref have benefited from these developments.
 A prominent example is OpenAlex, which has gained attention for its inclusivity, openness and comprehensive coverage of publication metadata [@priem2022openalex].
@@
 They highlight both its increasing use in bibliometric research and its adoption by research organisations.
-Comparative studies have generally found OpenAlex to be a reliable alternative to commercial databases [@alperin2024analysissuitabilityopenalexbibliometric; @culbert2024referencecoverageanalysisopenalex], although data quality issues occurred for key metadata fields needed for open access analysis, including evidence of open access status itself [@jahn2023analysing; @Simard_2024], as well as authorships [@culbert2024referencecoverageanalysisopenalex], affiliations [@Zhang_2024], author roles [@Jahn_2025] and document types used to distinguish between different article types [@Haupka_2024].
-However, these studies highlight the rapid improvements and the new possibilities of combining open metadata and the curation efforts of OpenAlex, in conjunction with the growing awareness of the importance of open metadata, as demonstrated by the Barcelona Declaration on Open Research Information.
-
-As a consequence, there is growing interest in using open bibliometric databases such as Crossref and OpenAlex to assess the impact of transformative agreements. 
+Comparative studies have generally found OpenAlex to be a reliable alternative to commercial databases [@alperin2024analysissuitabilityopenalexbibliometric; @culbert2024referencecoverageanalysisopenalex], although data quality issues occurred for key metadata fields required for open access analysis, including evidence of open access status itself [@jahn2023analysing; @Simard_2024], as well as authorships [@culbert2024referencecoverageanalysisopenalex], affiliations [@Zhang_2024], author roles [@Jahn_2025] and document types used to distinguish between different article types [@Haupka_2024].
+However, these studies highlight the rapid improvements and new possibilities of combining open metadata and the curation efforts of OpenAlex in conjunction with the growing awareness of the importance of open metadata, as demonstrated by the Barcelona Declaration on Open Research Information.
+
+Consequently, there is growing interest in using open bibliometric databases such as Crossref and OpenAlex to assess the impact of transformative agreements. 
 Similar to established proprietary bibliometric databases such as Scopus and Web of Science, these data sources lack specific data points, particularly about open access payments, which are typically not shared [@Haustein_APC_2024].
-However, bibliometric data can be enriched with open metadata about the journals and institutions involved in transformative agreements [@Jahn_2025]. 
+However, bibliometric data can be enriched with open metadata about journals and institutions involved in transformative agreements [@Jahn_2025]. 
 @de_Jonge_2025 demonstrated this potential by combining data from the cOAlition S Journal Checker tool and OpenAlex, achieving 89% accuracy when validated against Dutch research information. 
@@
 It adapts the approach from @Jahn_2025, which combines data from the cOAlition S Journal Checker tool, Crossref and OpenAlex, and applies the same methods to Web of Science and Scopus data.
-This comparative approach allows the evaluation of different bibliometric data sources for monitoring transformative agreements, while assessing their respective strengths and limitations.
-Specifically, this study compares the indexing of more than 13,000 hybrid journals across these sources, and examines whether consistent results can be obtained in terms of open access adoption between 2019 and 2023, analysed by publisher and country.
-By comparing results across these data sources, this large-scale study provides a more robust understanding of transformative agreements' impact on the transition to full open access. 
-It also reveals the comparative strengths and weaknesses of different bibliometric databases in measuring scholarly publishing's shift toward openness
+This comparative approach allows the evaluation of different bibliometric data sources to monitor transformative agreements, while assessing their respective strengths and limitations.
+Specifically, this study compares the indexing of more than 13,000 hybrid journals across these sources, and examines whether consistent results can be obtained in terms of open access adoption between 2019 and 2023, analysed by publishers and countries.
+By comparing results across these data sources, this large-scale study provides a more robust understanding of the impact of transformative agreements on the transition to full open access. 
+It also reveals the comparative strengths and weaknesses of different bibliometric databases in measuring scholarly publishing's shift toward openness.
 
@@
 
-hoaddata follows good practices for computational reproducibility using R [@Marwick_2018].
+hoaddata followed good practices for computational reproducibility using R [@Marwick_2018].
 The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
@@
 Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020].
-Web of Science is a selective, base research-focused database [@Visser_2021; @Stahlschmidt_2022].
-The collections considered in this study were the Science Citation Index Expanded (SCIE), the Social Sciences Citation Index (SSCI) and the Arts & Humanities Citation Index (AHCI).
-
-These collections provides important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall [@Piwowar_2018], the same provider as OpenAlex.
+Web of Science is a selective database with a focus on base research [@Visser_2021; @Stahlschmidt_2022].
+The collections considered in this study were the Science Citation Index Expanded (SCIE), Social Sciences Citation Index (SSCI) and Arts & Humanities Citation Index (AHCI).
+
+These collections provide important data points for analysing open access: author affiliations and roles, differentiation of journal articles into document types representing different types of journal contributions, such as original articles or reviews, and open access status information derived from OurResearch's Unpaywall [@Piwowar_2018], the same provider as OpenAlex.
 However, Web of Science lacks information about journals and articles under transformative agreements.
@@
 For programmatic access to article-level data, this study used the database of the Kompetenznetzwerk Bibliometrie (KB) in Germany.
-The KB processes raw XML data provided by Clarivate Analytics, which is ingested into an in-house PostgreSQL database under a uniform schema.
+The KB processes raw XML data provided by Clarivate Analytics, which are ingested into an in-house PostgreSQL database under a uniform schema.
 To support reproducibility, KB maintains annual snapshots of the database. 
@@
 Similar to Web of Science, Scopus is selective with regard to the journals it indexes.
-However, its journal coverage is much broader than that of the Web of Science collections considered in this study, as it also indexes a wider range of applied research journals [@Singh_2021; @Visser_2021; @Stahlschmidt_2022].
+However, its journal coverage is much broader than that of the Web of Science collections considered in this study, as it indexes a wider range of applied research journals [@Singh_2021; @Visser_2021; @Stahlschmidt_2022].
 With detailed metadata about article types, open access status information derived from Unpaywall, author roles, and disambiguated affiliations, Scopus also contains important data to assess open access uptake, although no direct information regarding transformative agreements was available at the time of the study.
 
-This study used the Scopus annual snapshot of April 2024 as provided by the KB (scp_b_202404). The same KB curation effort as for the Web of Science was applied to the Scopus [@schmidt_2024_13935407].
+This study used the Scopus annual snapshot from April 2024 as provided by KB (scp_b_202404). The same KB curation effort as that for Web of Science was applied to Scopus [@schmidt_2024_13935407].
 
@@
 
-### Determining hybrid journal publication volume
+### Determining the hybrid journal publication volume
 
 Following @Jahn_2025, the starting point was a unified dataset of several safeguarded JCT snapshots[^jct_data].
-The JCT journal data were enriched with ISSN variants linked to an ISSN-L. 
+JCT journal data were enriched with ISSN variants linked to ISSN-L. 
 To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists including the Directory of Open Access Journals (DOAJ).
-The resulting hybrid journal data were made available via hoaddata and used to determine the publication volume for each database independently.
-
-hoaddata relies on Crossref for obtaining journal publication volume and open access status through Creative Commons (CC) licence information relative to the published version ("version of record").
+The resulting hybrid journal data were made available via hoaddata and were used to independently determine the publication volume for each database.
+
+hoaddata relies on Crossref to obtain journal publication volume and open access status through Creative Commons (CC) licence information relative to the published version ("version of record").
 <!-- Because of limited affiliation metadata in Crossref [@van_Eck_2022], hoaddata sources  affiliation metadata from OpenAlex.-->
@@
 -->
-The article metadata included DOIs, publication dates, open access information as well as author roles and affiliations.
+The article metadata included DOIs, publication dates, open access information, and author roles and affiliations.
 Publication years were determined using the earliest known date of publication in a journal.
@@
 Because of limited information on these document types in open scholarly data [@Haupka_2024], hoaddata used an extended version of Unpaywall's paratext recognition approach to exclude non-scholarly content [@Piwowar_2018].
-To exclude conference supplements, which are also often not covered by transformative agreements, only articles published in regular issues, indicated by numerical pagination, were considered.
-For Web of Science and Scopus, their established, mainly accurate document type classifications [@Maisano_2025; @Donner_2017] were used to identify original research articles and reviews, referred to as original articles throughout this study.
+To exclude conference supplements, which are  often not covered by transformative agreements, only articles published in regular issues, indicated by numerical pagination, were considered.
+For Web of Science and Scopus, their established and mainly accurate document type classifications [@Maisano_2025; @Donner_2017] were used to identify original research articles and reviews, referred to as original articles throughout this study.
 
@@
 
-Articles in hybrid journals were considered open access when they were made freely available under a CC license on publisher platforms. 
+Articles in hybrid journals were considered open access when publishers made them freely available under a CC license on their platforms. 
 While hoaddata obtained this information from Crossref license metadata, Web of Science and Scopus relied on Unpaywall as evidence source. 
 Unpaywall also uses Crossref license metadata, but supplements them by parsing publisher websites directly, addressing cases where publishers do not provide machine-readable CC license information [@Piwowar_2018].
-This additional parsing remains necessary despite transformative agreement workflows recommend to deposit CC licenses during DOI registration [@Geschuhn_2017].
+This additional parsing remains necessary despite transformative agreement workflows recommending the deposition of CC licenses during DOI registration [@Geschuhn_2017].
 Both Web of Science[^wosoa] and Scopus[^scopusoa] defined hybrid open access consistently as content available under CC licenses on publisher platforms, distinguishing it from bronze open access that lack such explicit license information, or use publisher-specific licenses [@Piwowar_2018].
@@
 
-Author affiliations were retrieved for both first and, if available, corresponding authors to prepare the linking between articles and institutions covered by transformative agreements.
-To improve data retrieval, JCT institution data was enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex' institution entity.
-To handle different address variants, database-specific affiliation identifiers were used: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifier.
+Author affiliations were retrieved for both the first and, if available, the corresponding authors to prepare the linking between articles and institutions covered by transformative agreements.
+To improve data retrieval, JCT institution data were enriched with ROR-IDs from associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society, according to OpenAlex's institution entity.
+Database-specific affiliation identifiers were used to handle different address variants: ROR-IDs from OpenAlex for hoaddata, affiliation enhanced names for Web of Science, and Scopus Affiliation Identifier.
 Additionally, ISO country codes were retrieved for each author's address to compile country-level statistics. 
@@
 
-Because neither Web of Science nor Scopus supported ROR-ID at the time of the data retrieval, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise affiliation data.
-First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map first authors' ROR-IDs to corresponding proprietary affiliation identifier in Web of Science and Scopus using DOI matching.
+Because neither Web of Science nor Scopus supported ROR-ID at the time of data retrieval, the institution identifier used by the JCT, a two-step matching process was implemented to harmonise the affiliation data.
+First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map the first authors' ROR-IDs to corresponding proprietary affiliation identifiers in Web of Science and Scopus using DOI matching.
 Then, an algorithm selected the most frequent ROR ID and proprietary identifier pairs to handle multiple affiliations and organisational hierarchy differences.
@@
 
-This process linked `r length(unique(scp_ror$ror))` ROR-IDs to  `r length(unique(scp_ror$vendor_org_id))` Scopus Affiliation IDs, and  `r length(unique(wos_ror$ror))` ROR-IDs to `r length(unique(wos_ror$vendor_org_id))` enhanced affiliation strings in Web of Science.
-Quality evaluation through random sampling of 50 pairs revealed an error rate of 22% for Web of Science (11 mismatches) and 6% for Scopus (3 mismatches). 
+This process linked `r length(unique(scp_ror$ror))` ROR-IDs to  `r length(unique(scp_ror$vendor_org_id))` Scopus Affiliation IDs and  `r length(unique(wos_ror$ror))` ROR-IDs to `r length(unique(wos_ror$vendor_org_id))` enhanced affiliation strings in Web of Science.
+Quality evaluation through random sampling of 50 pairs revealed an error rate of 22% for Web of Science (11 mismatches) and 6% for Scopus (three mismatches). 
 Upon inspection, these mismatches primarily occurred with less-represented institutions having only a few publications, introduced through multiple affiliations of single authors. 
-The difference between databases suggests that Scopus' affiliation control aligns more closely with ROR than that of Web of Science.
+The difference between databases suggests that Scopus' affiliation control aligns more closely with ROR than with Web of Science.
 
@@
 Based on these compiled matching tables, articles eligible under transformative agreements could also be obtained from Web of Science and Scopus, although they did not contain ROR-IDs used by the JCT.
-The estimation of eligible articles followed @Jahn_2025 and included a matching of both journals and participating institutions according to the JCT.
-The matching also took into account the duration of agreements according to the ESAC registry, with only those matches where an agreement was actually in place being considered for subsequent analysis.
+The estimation of eligible articles followed @Jahn_2025 and included matching of both journals and participating institutions according to the JCT.
+The matching also considered the duration of agreements according to the ESAC registry, with only those matches where an agreement was actually in place being considered for subsequent analysis.
 A related study [@de_Jonge_2025], applied to publications funded by the Dutch Research Council (NWO) and validated against internal invoicing data, confirmed that such matching can accurately identify most articles under transformative agreements.
@@
 
-As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were aggregated for each database at country and journal level by year.
-@tbl-methods_overview_tab provides a general overview of the coverage between 2019 and 2023.
+As a result of the comprehensive data processing described above, datasets on open access in hybrid journals included in transformative agreements were aggregated for each database at the country and journal levels by year.
+@tbl-methods_overview_tab provides a general overview of coverage between 2019 and 2023.
 It shows that the majority of hybrid journals published at least one original research article or review, collectively marked as "original article" in the following, during the five-year period.
-These journals provided the basis for the subsequent calculation of article-level indicators.
+These journals have provided the basis for the subsequent calculation of article-level indicators.
 
@@
 
-While hoaddata only covers articles with a DOI, Scopus and Web of Science publication indicators were calculated using their database identifier.
+While hoaddata only covered articles with a DOI, Scopus and Web of Science publication indicators were calculated using their database identifier.
 A subsequent comparison of DOI coverage shows that non-original articles in Web of Science often lacked a DOI.
 This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
-Open access indicators were aggregated by DOI, because Unpaywall only collects information on open access status for articles with a DOI.
-A closer look at original articles with affiliation data, and in line with related research [@Zhang_2024], reveals a lack of affiliation data in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
-In particular, only about two-thirds of the articles examined provided corresponding author affiliations.
-For first authors, the proportion was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(hoad) * 100`%.
+Open access indicators were aggregated by DOI, because Unpaywall only collects information on open access status for articles with DOI.
+A closer look at original articles with affiliation data, in line with related research [@Zhang_2024], reveals a lack of affiliation data in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
+In particular, only approximately two-thirds of the articles examined provided corresponding author affiliations.
+The proportion for first authors was `r methods_overview_table_df |> filter(indicator == "First author articles") |> pull(hoad) / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(hoad) * 100`%.
 At the time of writing, OpenAlex disclosed limited coverage of corresponding authorship data [^oacorresponding].
@@
 
-Throughout this largely automated data collection and analysis process, Tidyverse tools [@tidyverse] for the R programming language [@r] were used. 
+TTidyverse tools [@tidyverse] for the R programming language [@r] were used throughout this largely automated data collection and analysis process. 
 Rank correlations were calculated using the Hmisc package [@Harrell_Jr_2003].
-The source code analysis, including all queries used to obtain the data, is available on GitHub (add link). 
-An interactive supplement exploring correlations between the data sources examined by country and publisher is available on HuggingFace Spaces: https://huggingface.co/spaces/najkoja/hoa_replication.
+Source code analysis, including all queries used to obtain the data, is available on GitHub (<https://github.com/njahn82/hoa_validation>). 
+An interactive supplement exploring correlations between the data sources examined by country and publisher is available via HuggingFace Spaces: <https://huggingface.co/spaces/najkoja/hoa_replication>.
 
@@
 This section first presents the indexing of hybrid open access by comparing open data sources with the proprietary bibliometric databases Scopus and Web of Science. 
-Then, using the same methods, indicators at the publisher and country level are calculated independently for each database and compared with each other to assess the suitability of the bibliometric databases for investigating transformative agreements.
+Then, using the same methods, indicators at the publisher and country levels are calculated independently for each database and compared to assess the suitability of the bibliometric databases for investigating transformative agreements.
 
@@
 The analysis included hybrid journals that published at least one open access article between 2019 and 2023, based on open access status information from each database.
-Only original articles were considered in the analysis.
+Only original articles were included in the analysis.
 
@@
 ```
-Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of hybrid journals included in transformative agreements were indexed in all three databases (Figure \ref{fig-upset_coverage_results}A).
-The second-largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
-Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaddata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were only found in Scopus.
-Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in Web of Science.
-Upon inspection, this group of hybrid journals exclusively covered in the proprietary data sources mainly represented hybrid journals for which no open access evidence could be retrieved from Crossref, the open access evidence source for hoaddata.
+Journal coverage analysis revealed that `r round(nrow(union_jns) / nrow(all_jns) * 100)`% (n = `r nrow(union_jns)`) of the hybrid journals included in transformative agreements were indexed in all three databases (Figure \ref{fig-upset_coverage_results}A).
+The second largest set consisted of journals indexed in both hoaddata and Scopus, comprising `r round( jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |>  nrow()`) of hybrid journals.
+Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaddata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were found only in Scopus.
+Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were available in Web of Science.
+Upon inspection, this group of hybrid journals exclusively covered in proprietary data sources mainly represented hybrid journals for which no open access evidence could be retrieved from Crossref, the open access evidence source for hoaddata.
 
 In terms of article coverage, Figure \ref{fig-upset_coverage_results}B shows the total publication volume per combination in terms of DOI availability.
-The largest set of hybrid journals, covering all three data sources, also represents the largest number of articles. 
-In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` articles, represented by the blue bar.
-They represented `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(wos) * 100)`% of original articles with a DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(scp) * 100)`% in Scopus.
-Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to the intersection of Scopus and hoaddata.
+The largest set of hybrid journals covering all three data sources also represented the largest number of articles. 
+In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` articles, represented by the blue bars.
+They recorded `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(wos) * 100)`% of the original articles with a DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(scp) * 100)`% in Scopus.
+Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to the intersection between Scopus and hoaddata.
 Exclusively in hoaddata were `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == FALSE) |> pull(n_cr) |> sum()` articles, and exclusively in the proprietary databases were `r upset_data |> filter(Crossref == FALSE, Scopus == TRUE) |> pull(overlap) |> sum(na.rm = TRUE)` articles.
@@
 Figure \ref{fig-upset_coverage_results}B also shows the surplus of articles with DOI that were only available via hoaddata (grey area).
-In the case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were only present in hoaddata.
-After validation at the DOI level using the KB databases and manual inspection, the main reason for missing DOI coverage in the proprietary database were insufficient classification of journal content as original articles during the compilation of hoaddata.
-Particularly, letters and editorials could not be fully detected.
-Moreover, paratext recognition failed for `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scholarly content such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science.
-To a lesser extent, differences in publication and indexing dates were a reason for non-overlapping DOIs.
-
-Using overlapping DOIs, the publication volume between 2019 and 2023 was also calculated for each journal. 
-Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows a large spread oj journal size across the journals jointly covered by all three data sources. 
+In the case of hybrid journals covered by all three data sources, `r miss_dois |> nrow()` DOIs were present only in hoaddata.
+After validation at the DOI level using KB databases and manual inspection, the main reason for missing DOI coverage in the proprietary database was insufficient classification of journal content as original articles during the compilation of hoaddata.
+Particularly, letters and editorials could not be detected fully.
+Moreover, paratext recognition failed for `r round(miss_dois_val |> filter(is.na(scp_item_type), is.na(wos_item_type)) |> pull(n_doi) / miss_dois |> nrow() * 100)`% of DOIs to identify non-scholarly content, such as front matters or reviewer lists, which are generally not indexed by Scopus and Web of Science.
+To a lesser extent, differences in publication and indexing dates were reasons for non-overlapping DOIs.
+
+Using overlapping DOIs, the publication volume between 2019 and 2023 was calculated for each journal. 
+Figure \ref{fig-upset_coverage_results}C illustrates the distribution for each combination. It shows a large spread of journal size across journals mutually covered by all three data sources. 
 These journals published more on average than journals covered by only one or two data sources, with the exception of journals exclusively covered by Scopus and Web of Science. 
@@
 An example is *Digital society* that published `r jn_upset_articles |> filter(issn_l == "2731-4650") |> pull(n_cr)` articles.
-This hybrid journal was launched in 2022, being covered by various Springer Nature transformative agreements since then.
+This hybrid journal was launched in 2022, and had since been covered by various Springer Nature transformative agreements.
 
@@
 
-Figure \ref{fig-upset_coverage_results_publisher} presents the coverage of hybrid journals in transformative agreements across data sources from 2019 to 2023 with a focus on publisher portfolios.
-The analysis highlights the dominance of the three largest publishers, Elsevier, Springer Nature, and Wiley, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of articles published during this five-year period.
+Figure \ref{fig-upset_coverage_results_publisher} presents the coverage of hybrid journals in transformative agreements across data sources from 2019 to 2023 focusing on publisher portfolios.
+The analysis highlights the dominance of the three largest publishers, Elsevier, Springer Nature, and Wiley, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of the hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of the articles published during this five-year period.
 In terms of article volume, Elsevier led with `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_art)*100`% of the total) published across `r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Elsevier") |> pull(prop_jn)*100`% of the total).
-Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals, although recording the largest number of hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%).
+Springer Nature followed with `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(n_journals)` hybrid journals, although it recorded the largest number of hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Springer Nature") |> pull(prop_jn)*100`%).
 Wiley accounted for `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(articles)` articles (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_art)*100`%) in `r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(n_journals)` hybrid journals (`r jn_upset_publisher_all |> filter(publisher == "Wiley") |> pull(prop_jn)*100`%).
@@
 
-The three largest publishers, Elsevier, Springer Nature, and Wiley, were best represented in the  intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
-Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage (n = `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()`; `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%), as determined through shared DOIs.
-When examining publication volume per journal, (Figure \ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals, followed by Springer Nature and Wiley.
-
-Comparing publisher portfolios across different indexing sets demonstrates that publishers were not represented uniformly. 
+The three largest publishers, Elsevier, Springer Nature, and Wiley, were best represented at the  intersection of all three data sources  (hoaddata, Scopus, and Web of Science).
+Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage (n = `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()`; `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%) as determined through shared DOIs.
+When examining the publication volume per journal (Figure \ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals, followed by Springer Nature and Wiley.
+
+A comparison of publisher portfolios across different indexing sets demonstrates that publishers were not represented uniformly. 
 Notably, Springer Nature exhibited `r jn_upset_publisher |> filter(esac_publisher == "Springer Nature", WoS == FALSE, Crossref == TRUE, Scopus == TRUE) |> distinct(issn_l) |> nrow()` hybrid journals exclusively indexed in both hoaddata and Scopus.
 This set included journals from the Chinese Academy of Science, German-language medical journals, and Eastern European publications including the *Journal of Mathematical Sciences*, which also publishes English-language translations of Russian-language works.
-Additionally, this subset included titles with a broader disciplinary focus such as *SN Computer Science* and newly launched hybrid journals like *Nature Computational Science*, which started in 2021 and was indexed in Scopus but not yet in Web of Science.
-The set also captured discontinued journals, providing further insights into the dynamics of journal publishing.
+Additionally, this subset included titles with a broader disciplinary focus, such as *SN Computer Science*, and newly launched hybrid journals such as *Nature Computational Science*, which started in 2021 and were indexed in Scopus but not yet in Web of Science.
+The set also captures discontinued journals, providing further insights into the dynamics of journal publishing.
 
@@
 
-Examining publisher portfolios not covered by hoaddata but present in Scopus or Web of Science identified several publishers with missing Creative Commons (CC) license information in Crossref.
-In particular, `r miss_tas$esac_publisher[1]` represented `r miss_tas$journals[1]` journals with `r miss_tas$n[1]` articles, `r miss_tas$esac_publisher[2]` accounted for `r miss_tas$journals[2]` journals with `r miss_tas$n[2]` articles, and `r miss_tas$esac_publisher[3]` recorded `r miss_tas$journals[3]` journals and `r miss_tas$n[3]` articles.
+Examining publisher portfolios not covered by hoaddata, but present in Scopus or Web of Science, identified several publishers with missing Creative Commons (CC) license information in Crossref.
+In particular, `r miss_tas$esac_publisher[1]` represented `r miss_tas$journals[1]` journals with `r miss_tas$n[1]` articles, `r miss_tas$esac_publisher[2]` accounted for `r miss_tas$journals[2]` journals with `r miss_tas$n[2]` articles, and `r miss_tas$esac_publisher[3]` recorded `r miss_tas$journals[3]` journals with `r miss_tas$n[3]` articles.
 In total, `r nrow(miss_publishers)` publishers did not share CC licenses with Crossref and were therefore not represented in hoaddata.
@@
 An inspection of individual journals also uncovered discrepancies in Unpaywall's open access identification for certain publishers that typically share CC license metadata with Crossref.
-Notably, some subscription-only journals contained one or two articles erroneously tagged as hybrid open access by Unpaywall, which were subsequently reflected in Scopus and Web of Science.
-Examples of such misclassifications include Elsevier's *Journal of Bioscience and Bioengineering* and Springer Nature's *Journal of Mechanical Science and Technology*.
+Notably, some subscription-only journals contained one or two articles erroneously tagged as hybrid open access by Unpaywall, which was subsequently reflected in Scopus and Web of Science.
+Examples of such misclassifications include Elsevier's *Journal of Bioscience and Bioengineering*, and Springer Nature's *Journal of Mechanical Science and Technology*.
 
@@
 This section examines the uptake of open access in hybrid journals, focusing on the influence of transformative agreements across hoaddata, Scopus, and Web of Science.
-The aim was to assess whether consistent results can be derived from these data sources despite differences in coverage and methodologies. 
-Following @Jahn_2025, indicators were calculated for each data source and comprise the number and proportion of open access articles, including those enabled by transformative agreements, from 2019 to 2023.
-For Web of Science and Scopus, the impact of transformative agreements was estimated using both first and corresponding authorships, while hoaddata indicator calculation was limited to first author affiliations.
+The aim was to assess whether consistent results could be derived from these data sources despite differences in coverage and methodologies. 
+Following @Jahn_2025, indicators were calculated for each data source and comprised the number and proportion of open access articles, including those enabled by transformative agreements, from 2019 to 2023.
+For Web of Science and Scopus, the impact of transformative agreements was estimated using both first and corresponding authorships, while hoaddata indicator calculations were limited to first author affiliations.
 
@@
 Similarly, Scopus recorded a growth from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
-However, Web of Science recorded higher open access uptake in early years, before converging to a similar level in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
-This suggest a different approach towards labelling hybrid open access by Web of Science.
+However, Web of Science recorded higher open access uptake in the early years, before converging to a similar level in 2023, from `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2019) |> pull(n_oa_articles)`) in 2019 to `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(oa_prop)*100`% (n = `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(n_oa_articles)`) in 2023.
+This suggests a different approach towards labelling hybrid open access by Web of Science.
 
 Similarly, hybrid open access by transformative agreements substantially increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
-Trends were consistent when measuring first (dashed line) and corresponding author (dotted line) affiliations. 
+Trends were consistent when measuring the first (dashed line) and corresponding authorship (dotted line). 
 According to Scopus, `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_first_author_articles) |> sum()` open access articles could be attributed to transformative agreements based on first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`) and `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
@@
 
-From 2021 (hoadata, Scopus) resp. 2022 (Web of Science),  transformative agreements enabled the majority of open access articles in hybrid journals.
-For first authors, the share ranged between `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and   `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science) in 2023.
-For corresponding authors, the shares were slightly larger, with Scopus recording  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% and Web of Science `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% in 2023.
+Since 2021 (hoadata, Scopus) resp. 2022 (Web of Science),  transformative agreements have enabled the majority of open access articles in hybrid journals.
+For the first authors, the share ranged between `r uptake_by_year_short |> filter(src == "hoaddata", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (hoaddata),  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_prop) * 100`% (Scopus), and   `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_prop) * 100` % (Web of Science) by 2023.
+For the corresponding authors, the shares were slightly larger, with Scopus recording  `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% and Web of Science `r uptake_by_year_short |> filter(src == "Web of Science", earliest_year == 2023) |> pull(ta_oa_cor_prop) * 100`% by 2023.
 However, substantial hybrid open access was still facilitated outside transformative agreements, likely through APCs paid from discretionary research funds [@Suber_2012].
@@
 In contrast, Elsevier's share remained relatively constant, increasing only slightly from `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2019) |> pull(oa_prop)*100`% to `r uptake_by_publisher_short |> filter(src == "Web of Science", publisher == "Elsevier", earliest_year == 2023) |> pull(oa_prop)*100`% according to Web of Science.
-Upon inspection, this discrepancy is primarily due to articles in the publisher's open archive. 
+Upon inspection, this discrepancy was primarily due to articles under the publisher's open archive. 
 These articles, made freely available after an embargo period under Elsevier's user license, were tagged as hybrid open access in Web of Science, even though its documentation[^wosoa] specified that only articles under a CC license variant were considered.
@@
 Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022 compared to hoaddata. 
-For these two years, hoaddata reported `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Scopus", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more open access articles than Scopus and `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Web of Science", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more than Web of Science.
+For these two years, hoaddata reported `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Scopus", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more open access articles than Scopus and `r uptake_publisher_by_year |> filter(publisher == "Wiley", src == "HOAD", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum() - uptake_publisher_by_year |> filter(publisher == "Wiley", src == "Web of Science", earliest_year %in% 2021:2022) |> pull(n_oa_articles) |> sum()` more open access articles than Web of Science.
 This discrepancy is presumably due to challenges in fetching full-texts by Unpaywall, the open access evidence source for Scopus and Web of Science.
-According to Unpaywall's software version history, HTTP redirects from Wiley's publisher platform prevented Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
+According to Unpaywall's software version history, HTTP redirects from Wiley's publisher platform, preventing Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
 hoaddata, which relies solely on Crossref metadata for open access identification, was unaffected by these issues.
@@
 Wiley emerged as the fastest-growing publisher in terms of open access uptake, with more than 30% of articles in hybrid journals reported as open access in 2023 across the examined data sources, followed by Springer Nature.
-Elsevier, recorded a later uptake, which is consistent with the publisher's historical reluctance to engage in negotiations with library consortia.
+Elsevier recorded a later uptake, consistent with the publisher's historical reluctance to engage in negotiations with library consortia [@Fraser_2023].
 However, by 2023, the share of open access enabled by transformative agreements appeared to stabilise for all three publishers (see Figure \ref{fig-uptake_publisher}C). 
-Interestingly, the differences between first and corresponding author affiliations were more pronounced at the publisher level. 
-In Scopus, for example, the share of open access via transformative agreements measured by corresponding authorship was higher for Springer Nature in 2023 than when using first authorship.
+Interestingly, the differences between first and corresponding authorships were more pronounced at the publisher level. 
+In Scopus, for example, the share of open access via transformative agreements measured by the corresponding authorship was higher for Springer Nature in 2023 than when using the first authorship.
 
@@
 
-When comparing countries, consistent patterns were observed across data sources for the five-year period 2019 to 2023. 
+When comparing countries, consistent patterns were observed across data sources for the five-year period from 2019 to 2023. 
 Figure \ref{fig-uptake_country} presents hybrid open access indicators by country, comparing hoaddata (x-axis) with Web of Science and Scopus (y-axis).
-Indicators calculated from these proprietary databases are shown for both first and corresponding authors, with full counting used to account for multiple country affiliations [@Hottenrott_2021].
+Indicators calculated from these proprietary databases are shown for both the first and corresponding authors, with full counting used to account for multiple country affiliations [@Hottenrott_2021].
 
@@
 
-In terms of article output by country (see Figure \ref{fig-uptake_country}A), strong positive correlations were observed across the data sources and author roles (Spearman rank correlation $\rho > .9$, $p < 0.001$).
+In terms of article output by country (see Figure \ref{fig-uptake_country}A), strong positive correlations were observed across data sources and author roles (Spearman rank correlation $\rho > .9$, $p < 0.001$).
 Between 2019 and 2023, China was the most productive country, followed by the United States and, by a certain margin, India, the United Kingdom, and Germany. 
-Analysis of authorship roles revealed minimal variation, indicating that first and corresponding authors were typically from the same country.
-
-When examining the percentage of open access articles in hybrid journals (see Figure \ref{fig-uptake_country}B), a different pattern emerged.
-Authors affiliated with institutions from medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, provided a large proportion of their articles in open access.
+The analysis of authorship roles revealed minimal variation, indicating that the first and corresponding authors were typically from the same country.
+
+When examining the percentage of open access articles in hybrid journals (Figure \ref{fig-uptake_country}B), a different pattern emerged.
+Authors affiliated with institutions from medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, made a large proportion of their articles open access.
 Germany and the United Kingdom also had approximately 40% of their output available as open access.
@@
 These findings were consistent across all databases.
-However, France was better represented in Web of Science, likely due to its agreement with Elsevier starting in 2019, which allowed delayed open access under the publisher’s user license [@Rabesandratana_2019].
+However, France was better represented in Web of Science, likely because of its agreement with Elsevier starting in 2019, which allowed delayed open access under the publisher’s user license [@Rabesandratana_2019].
 This licence was not classified as hybrid open access in either Scopus or hoaddata.
-In all cases, Spearman rank correlations were $\rho > .9$, $p < 0.001$, showing a high level of correlations between the databases and authorship roles considered.
+In all cases, Spearman rank correlations were $\rho > .9$, $p < 0.001$, indicating a high level of correlation between the databases and authorship roles considered.
 
@@
 As a notable exception, South Africa again featured prominently, as the South African National Library and Information Consortium (SANLiC) successfully negotiated transformative agreements with major publishers from 2022 onward.[^sa]
-Results were consistent across data sources. 
-However, Wiley's open access surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with publisher that started in July 2019.
+The results were consistent across data sources. 
+However, Wiley's open access surplus in hoaddata led to better rankings for countries where Wiley played a substantial role, such as Germany, where the DEAL consortium negotiated its first transformative agreements with publishers that started in July 2019.
 
@@
 with high-income European countries—alongside South Africa—showing high adoption rates.
-In contrast, the three most productive countries, China, the United States, and India, showcase a substantially lower adoption in transformative agreements.
-Open questions remain whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or a deliberate avoidance of such agreements.
+In contrast, the three most productive countries, China, the United States, and India, show a substantially lower adoption in transformative agreements.
+Open questions remain as to whether this uneven distribution reflects temporary implementation gaps, inherent inequities in the transformative agreement model, or deliberate avoidance of such agreements.
 
 The findings were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science. 
-Notably, and aligning with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73].
-Overall, the results show strong correlations by country affiliation, which supports using open metadata for large-scale analyses of hybrid open access. 
-However, the observed differences in journal coverage and metadata availability warrant further discussion in the following, affecting not only open data sources but also proprietary databases when used in isolation.
-
-The coverage analysis reveals that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage, reflecting the dominance of major publishers whose established journal portfolios are comprehensively indexed in proprietary databases [@vanbellen2024oligopolyacademicpublisherspersists].
+This is consistent with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73].
+Overall, the results show strong correlations by country affiliation, which supports the use open metadata for large-scale analyses of hybrid open access. 
+However, the observed differences in journal coverage and metadata availability warrant further discussion, affecting not only open data sources but also proprietary databases when used in isolation.
+
+The coverage analysis revealed that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage, reflecting the dominance of major publishers whose established journal portfolios are comprehensively indexed in proprietary databases [@vanbellen2024oligopolyacademicpublisherspersists].
 Differences emerge for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
-Using Crossref as a bibliometric database in hoaddata demonstrated particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
+Using Crossref as a bibliometric database in hoaddata demonstrated a particular strength in identifying newly established hybrid journals, a notable finding given that transformative agreements primarily target existing subscription-based journals.
 The landscape of hybrid journal publishing thus differs markedly from that of fully open access journals.
@@
 However, this analysis demonstrates that indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, reflecting disciplinary norms in scholarly publishing with regard to contributorship and author roles and positions [@Larivi_re_2016].
-First authors typically conduct the main research underlying a paper, while the corresponding author often supervises the research [@Mattsson_2010; @Fox_2018]. 
+The first author typically conduct the main research underlying a paper, while the corresponding author often supervises the research [@Mattsson_2010; @Fox_2018]. 
 Unsurprisingly, measures based on first or corresponding authorships are strongly correlated, suggesting that these authors share the same country affiliation. 
-Moreover, in most cases, the first author is identical to the corresponding author [@Chinchilla_Rodr_guez_2024].
+In most cases, the first author is identical to the corresponding author [@Chinchilla_Rodr_guez_2024].
 <!-- Thus, first authorship can reliably be used in large-scale studies to assess the impact of transformative agreements. -->
@@
 This prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
-Previous research has highlighted how institutionalised bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], suggesting the need to monitor potential influences of the availability of open access funding on authorship roles [@Maddi_2024].
+Previous research has highlighted how institutionalised bibliometric practices can affect the valuation of authorship positions [@Helgesson_2020], suggesting the need to monitor the potential influences of the availability of open access funding on authorship roles [@Maddi_2024].
 
 Another critical data element in the study is affiliation data, which is essential for estimating open access enabled by transformative agreements. 
-Although OpenAlex's affiliation coverage is less comprehensive, which likely reduced the number of articles confidently attributed to transformative agreements in this study, it still shows high correlations with Scopus and Web of Science at the country level.
+Although OpenAlex's affiliation coverage is less comprehensive, which likely reduced the number of articles confidently attributed to transformative agreements in this study, it still showed high correlations with Scopus and Web of Science at the country level.
 However, OpenAlex's native ROR-ID integration offers a distinct advantage, allowing a more reliable identification of agreement-enabled articles compared to Scopus and Web of Science, which require reconciliation with proprietary organisation identifiers.
-Future studies based on the Web of Science will benefit from the recent integration of ROR-IDs, announced by the end of 2024.
+Future studies based on Web of Science will benefit from the recent integration of ROR-IDs, announced by the end of 2024.
 <!-- Dimension is another scholarly data source that uses open standards to identify journals, which enables analysing transformative agreements [@Fraser_2023].-->
@@
 Crucially, not all publishers share CC licence metadata via Crossref, a limitation that becomes apparent when contrasting Crossref with Unpaywall’s data in Scopus and Web of Science. 
-While Unpaywall can detect such gaps by parsing publishers' websites for open access licences, it nevertheless missed a substantial number of CC-licensed open access articles from Wiley journals indexed in Crossref, likely due to parsing errors on the publisher's website.
+While Unpaywall can detect such gaps by parsing publishers' websites for open access licences, it nevertheless missed a substantial number of CC-licensed open access articles from Wiley journals indexed in Crossref, likely because of  parsing errors on the publisher's website.
 This resulted in fewer open access articles being recorded in Scopus and Web of Science.
 Further inconsistencies emerged between Scopus and Web of Science, despite both relying on Unpaywall: Web of Science erroneously labels Elsevier’s delayed open access as hybrid, whereas Scopus correctly categorises it.
-Notably, Scopus and Crossref showed greater alignment than observed in a related comparison between Scopus and OpenAlex [@alperin2024analysissuitabilityopenalexbibliometric].
-This likely reflects the use of a curated list to identify hybrid open access journals, rather than relying solely on article-level tags.
-Given these variations, which may compromise comparability, research and monitoring exercises should avoid reliance on a single source.
+Notably, Scopus and Crossref showed greater alignment than that observed in a related comparison between Scopus and OpenAlex [@alperin2024analysissuitabilityopenalexbibliometric]. 
+This likely reflects the use of a curated list to identify hybrid open access journals, rather than depending solely on article-level tags.
+Given these variations, which may compromise comparability, research and monitoring exercises should avoid relying on a single source. 
 Instead, the selection of open access evidence should be cross-verified using multiple sources and snapshots that can be used to track changes in the data over time [@Huang_2020].
@@
 
-It is important to note that the study's estimates of articles from institutions involved in transformative agreements are approximations due to a lack of access to invoice data, which is not usually shared by library consortia and publishers.
-Furthermore, the data sources about transformative agreements including the cOAlition S Journal Checker Tool and the ESAC registry are based on a voluntary effort, crowd-sourced from various consortia.
-However, recent validation with Dutch research information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements [@de_Jonge_2025]. 
-With enhanced open metadata compliance, attributable to evolving standards and initiatives in support of negotiations with publishers, particularly through the communities of the ESAC initiative and Barcelona Declaration supporters, the situation is likely to improve.
+It is important to note that the study's estimates of articles from institutions involved in transformative agreements are approximations because of a lack of access to invoice data, which is not usually shared by library consortia and publishers.
+Furthermore, the open data sources for transformative agreements, including the cOAlition S Journal Checker Tool and the ESAC registry, are based on voluntary effort, crowd-sourced from various consortia.
+However, recent validation using Dutch research information demonstrates the reliability of such an open approach for the assessment of articles under transformative agreements [@de_Jonge_2025]. 
+With enhanced open metadata compliance attributable to evolving standards and initiatives in support of negotiations with publishers, particularly through the ESAC initiative and Barcelona Declaration on Open Research Information, the situation is likely to improve.
 In conclusion, this study has shown that open metadata are well-suited for analysing the transition from subscription-based journal publishing to full open access.
 
+# Author contributions {-}
+
+The author confirms the sole responsibility for the following: Conceptualization, Methodology, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Writing - Review & Editing, Visualization.
+
+# Competing interests {-}
+
+The author declares no competing interests.
+
+# Funding information {-}
+
+This work was supported by the Federal Ministry of Education and Research of Germany (BMBF) under grants 16WIK2301E / 16WIK2101F.
+
+The proprietary bibliometric data from Scopus and Web of Science were provided by the German Competence Network for Bibliometrics under BMBF grant 16WIK2101A.
+
+# Data and code availability {-}
+
+Data source code supplement is available via GitHub: <https://github.com/njahn82/hoa_validation>.
+An interactive dashboard, which allows the browsing of metrics by country and publisher, is available on HuggingFace Spaces: <https://huggingface.co/spaces/najkoja/hoa_replication>.
+
+Please note that according to the licence terms for Scopus and Web of Science, as provided by the  German Competence Network for Bibliometrics, only aggregated data accompanying the scientific publication can be made publicly available for replication purposes.
+
+Open metadata provided by Crossref, OpenAlex, and cOAlition S Journal Checker Transformative Agreement Data are freely available for re-use.
+
 # References
 
-<!-- Limitations: agreeemnts are complex, game auhtorship, invoicing not available, mapping, not feasible, requires more insitutional and publsihjer case studies involving financial data, but also improved and streamlined metadata compliance according to the evlolving stadatrd and iniatives (ESAC, Barcelona) in negotiation with publishers.
-
-Overall, -->
-
-<!--Journal indexing is also influenced by the open access evidence source used. 
-Using the same list of hybrid journals, the results show consoderable difference in indexing, because of different approach to identify open access.
-(coverage)
-
-(document types)
-( correpsonding auhtor)
-(open access)
-(pids)
-
-<!-- Hybird journals are very well indexed in both proprietory and open sources.
-confirms the  large concentrartion of publieher there is a much a less advantage of using openalex in terms of includindg missing ones as in comparision to full opena ccess jorunals with or without APCs (also called diamond oa).
-Previouly reported hybrid obsvered surplus is likely due to miscoding of full open access jorunals, as they cannot been found in list of transofrmative jorunals.
-However, uuing open schoalrly sources ournals with local audience  can be helpful to determine newly launched hybrid jorunals, as well as journals with local audience which is surprisingly given that transformative agreemnt are intented to shift the business modell of extisting ubscripton.based journal.
-
-Combining different open access evidence sources
-
-Authorship
-
-confirm metadata gaps, in particular document types.
-hoewever, it is often unclear, which are incldued in transformative agreements.
-but metadata gaps ar enot just to be found in openalex, but also in .
-Lack oif ror ids to incldue this.
-only a workaroudn made it possible to establish a link.
-a manual curation is not feaible given the amoutn of transformative agreements possible.
-
-the observed increae provide oppoirtunites for improved meatdata.
-negotiator have leverage, funder start using  open arrpaoches, standardisatuion activities in .
-These demosntrate the added value of transformative agrrements.
-
-The increasing trend with is consistent with previous research and case studie that describe an general increase with the introduction of transformative agreements. For example, Sweden reproted
-
-In terms of metadata coverage and quality, it shows the value of combining
-
-that,a t least at the country level, resuts between first and correpsodnign auhtor are storngyl correlated. 
-In line with empricial research, showing that morst first auhtros were also correposnding auhtor. 
-This can be explained by norm of auhtorship where fir
-
-The study deomnstrates the impact of transformative agreements on hybrid open access being consistent across biliometric data 
- [juan study] larger proportion of hybrid, suggest journals included which might not be covered by transformative agreement, oa journals not tagged as fully oa, 741268 withouht any publisher. -->
-
 [^unpaywallwiley]: See Unpaywall version history related to Wiley fixes: <https://github.com/search?q=repo%3Aourresearch%2Foadoi+wiley&type=commits>

### 2025-04-16 analysis/manuscript.tex
Some minor language fixes

@@
 \section*{Abstract}
-{}
-{\textbf{Keywords}: hybrid open access, transformative agreements, scholarly publishing, big deals, bibliometrics}
+{This study compares open metadata from hoaddata, an openly available dataset based on Crossref, OpenAlex and the cOAlition S Journal Checker Tool, with proprietary bibliometric databases Scopus and Web of Science to estimate the impact of transformative agreements on hybrid open access publishing.
+Analysing over 13,000 hybrid journals between 2019-2023, the research found substantial growth in open access due to these agreements, although most articles remain paywalled.
+The results were consistent across all three data sources, showing strong correlations in country-level metrics despite differences in journal coverage and metadata availability.
+By 2023, transformative agreements enabled the majority of hybrid open access articles, with particularly high adoption in European countries.
+The analysis revealed strong alignment between first and corresponding authorship when measuring agreement uptake by publisher and country.
+This comparative approach supports the use of open metadata for large-scale hybrid open access studies, while using multiple data sources together provides a more robust understanding of hybrid open access adoption than any single database can offer, overcoming individual limitations in coverage and metadata quality.}
+
+{\textbf{Keywords}: bibliometric data sources, open metadata, hybrid open access, transformative agreements, curative bibliometrics}
 
@@
 aimed at hybrid journal bundles, allowing authors from participating
-institutions to publish open access in these journals, while providing
+institutions to publish open access in these journals while providing
 reading access to the entire portfolio. Library consortia and funders
-have begun coordinating how these agreements are negotiated, managed and
-evaluated. At the same time, efforts are underway to improve
+have begun to coordinate how these agreements are negotiated, managed
+and evaluated. At the same time, efforts are underway to improve
 transparency through open data on contracts and payments. However, this
@@
 Since their initial proposal by the Max Planck Digital Library (Schimmer
-et al., 2015), transformative agreements have led to a rapid growth in
-the number of negotiated agreements and open access articles (Jahn,
-2025; McCabe \& Mueller-Langer, 2024; Rothfritz et al., 2024). Before,
-only a small proportion of articles in hybrid journals were openly
-available (Björk, 2012; Laakso \& Björk, 2016; Piwowar et al., 2018),
-with publications from only a few European countries with dedicated
-funding policies and the high-energy physics SCOAP\^{}3 consortium
-standing out (Huang et al., 2020; Kohls \& Mele, 2018; Pinfield, 2015;
+et al., 2015), transformative agreements have led to rapid growth in the
+number of negotiated agreements and open access articles (Jahn, 2025;
+McCabe \& Mueller-Langer, 2024; Rothfritz et al., 2024). Before, only a
+small proportion of articles in hybrid journals were openly available
+(Björk, 2012; Laakso \& Björk, 2016; Piwowar et al., 2018), with
+publications from only a few European countries with dedicated funding
+policies and the high-energy physics SCOAP\^{}3 consortium standing out
+(Huang et al., 2020; Kohls \& Mele, 2018; Pinfield, 2015;
 Robinson-Garcia et al., 2020). Related coordination activities, in
@@
 reduced competition and incentives (McCabe \& Mueller-Langer, 2024;
-Schmal, 2024), and continuation of publication fees that widen gaps
+Schmal, 2024), and the continuation of publication fees that widen gaps
 between well-resourced and under-resourced institutions (Babini et al.,
@@
 policies. Conversely, the German DEAL consortium extended agreements
-with major publishers until 2028, while also outside of Europe more and
-more transformative agreements are being put in place (for instance in
-Latin America, as described in Muñoz-Vélez et al. (2024)). Responding to
+with major publishers until 2028, while also outside of Europe, more
+transformative agreements are being put in place (for instance in Latin
+America, as described in Muñoz-Vélez et al. (2024)). Responding to
 criticism, the Max Planck Digital Library recently outlined strategies
-to evolve the original model, focusing on achieving full open access
+to evolve the original model, focusing on achieving fully open access
 while controlling costs and addressing identified weaknesses (Dér, 2025)
@@
 landscape of open research information as library consortia negotiate
-for comprehensive open metadata, helping avoid data analytics becoming a
-commodity controlled by publishers (Aspesi \& Brand, 2020; McCabe \&
-Mueller-Langer, 2024). For instance, ESAC recommends using Crossref
-(Hendricks et al., 2020) to share metadata supporting workflows
+for comprehensive open metadata, helping avoid data analytics from
+becoming a commodity controlled by publishers (Aspesi \& Brand, 2020;
+McCabe \& Mueller-Langer, 2024). For instance, ESAC recommends using
+Crossref (Hendricks et al., 2020) to share metadata supporting workflows
 including open access licenses and funding information (Geschuhn \&
@@
 open metadata through this publisher-driven DOI registration platform
-and continuously monitors progress (N. J. van Eck \& Waltman, 2022).
-
-Many open scholarly data services that build on Crossref benefit from
-these developments. A prominent example is OpenAlex, which has gained
-attention for its inclusivity, openness and comprehensive coverage of
-publication metadata (Priem et al., 2022). In a recent evaluation,
-Céspedes et al. (2025) reviewed the current evidence base on the
-strengths and weaknesses of OpenAlex. They highlight both its increasing
-use in bibliometric research and its adoption by research organisations.
-Comparative studies have generally found OpenAlex to be a reliable
-alternative to commercial databases (Alperin et al., 2024; Culbert et
-al., 2024), although data quality issues occurred for key metadata
-fields needed for open access analysis, including evidence of open
-access status itself (Jahn et al., 2023; Simard et al., 2024), as well
-as authorships (Culbert et al., 2024), affiliations (Zhang et al.,
+and continuously monitors the progress (N. J. van Eck \& Waltman, 2022).
+
+Many open scholarly data services that build on Crossref have benefited
+from these developments. A prominent example is OpenAlex, which has
+gained attention for its inclusivity, openness and comprehensive
+coverage of publication metadata (Priem et al., 2022). In a recent
+evaluation, Céspedes et al. (2025) reviewed the current evidence base on
+the strengths and weaknesses of OpenAlex. They highlight both its
+increasing use in bibliometric research and its adoption by research
+organisations. Comparative studies have generally found OpenAlex to be a
+reliable alternative to commercial databases (Alperin et al., 2024;
+Culbert et al., 2025), although data quality issues occurred for key
+metadata fields required for open access analysis, including evidence of
+open access status itself (Jahn et al., 2023; Simard et al., 2024), as
+well as authorships (Culbert et al., 2025), affiliations (Zhang et al.,
 2024), author roles (Jahn, 2025) and document types used to distinguish
 between different article types (Haupka et al., 2024). However, these
-studies highlight the rapid improvements and the new possibilities of
-combining open metadata and the curation efforts of OpenAlex, in
+studies highlight the rapid improvements and new possibilities of
+combining open metadata and the curation efforts of OpenAlex in
 conjunction with the growing awareness of the importance of open
@@
 
-As a consequence, there is growing interest in using open bibliometric
+Consequently, there is growing interest in using open bibliometric
 databases such as Crossref and OpenAlex to assess the impact of
@@
 payments, which are typically not shared (Haustein et al., 2024).
-However, bibliometric data can be enriched with open metadata about the
+However, bibliometric data can be enriched with open metadata about
 journals and institutions involved in transformative agreements (Jahn,
@@
 data. This comparative approach allows the evaluation of different
-bibliometric data sources for monitoring transformative agreements,
-while assessing their respective strengths and limitations.
-Specifically, this study compares the indexing of more than 13,000
-hybrid journals across these sources, and examines whether consistent
-results can be obtained in terms of open access adoption between 2019
-and 2023, analysed by publisher and country. By comparing results across
-these data sources, this large-scale study provides a more robust
-understanding of transformative agreements' impact on the transition to
-full open access. It also reveals the comparative strengths and
-weaknesses of different bibliometric databases in measuring scholarly
-publishing's shift toward openness
+bibliometric data sources to monitor transformative agreements, while
+assessing their respective strengths and limitations. Specifically, this
+study compares the indexing of more than 13,000 hybrid journals across
+these sources, and examines whether consistent results can be obtained
+in terms of open access adoption between 2019 and 2023, analysed by
+publishers and countries. By comparing results across these data
+sources, this large-scale study provides a more robust understanding of
+the impact of transformative agreements on the transition to full open
+access. It also reveals the comparative strengths and weaknesses of
+different bibliometric databases in measuring scholarly publishing's
+shift toward openness.
 
@@
 
-hoaddata follows good practices for computational reproducibility using
+hoaddata followed good practices for computational reproducibility using
 R (Marwick et al., 2018). The package, which includes data, code, a test
@@
 proprietary bibliometric database consisting of several collections
-(Birkle et al., 2020). Web of Science is a selective, base
-research-focused database (Stahlschmidt \& Stephen, 2022; Visser et al.,
+(Birkle et al., 2020). Web of Science is a selective database with a
+focus on base research (Stahlschmidt \& Stephen, 2022; Visser et al.,
 2021). The collections considered in this study were the Science
-Citation Index Expanded (SCIE), the Social Sciences Citation Index
-(SSCI) and the Arts \& Humanities Citation Index (AHCI).
-
-These collections provides important data points for analysing open
+Citation Index Expanded (SCIE), Social Sciences Citation Index (SSCI)
+and Arts \& Humanities Citation Index (AHCI).
+
+These collections provide important data points for analysing open
 access: author affiliations and roles, differentiation of journal
@@
 database of the Kompetenznetzwerk Bibliometrie (KB) in Germany. The KB
-processes raw XML data provided by Clarivate Analytics, which is
+processes raw XML data provided by Clarivate Analytics, which are
 ingested into an in-house PostgreSQL database under a uniform schema. To
@@
 that of the Web of Science collections considered in this study, as it
-also indexes a wider range of applied research journals (Singh et al.,
-2021; Stahlschmidt \& Stephen, 2022; Visser et al., 2021). With detailed
+indexes a wider range of applied research journals (Singh et al., 2021;
+Stahlschmidt \& Stephen, 2022; Visser et al., 2021). With detailed
 metadata about article types, open access status information derived
@@
 
-This study used the Scopus annual snapshot of April 2024 as provided by
-the KB (scp\_b\_202404). The same KB curation effort as for the Web of
-Science was applied to the Scopus (Schmidt et al., 2024).
+This study used the Scopus annual snapshot from April 2024 as provided
+by KB (scp\_b\_202404). The same KB curation effort as that for Web of
+Science was applied to Scopus (Schmidt et al., 2024).
 
@@
 
-\subsubsection{Determining hybrid journal publication
-volume}\label{determining-hybrid-journal-publication-volume}
+\subsubsection{Determining the hybrid journal publication
+volume}\label{determining-the-hybrid-journal-publication-volume}
 
@@
 several safeguarded JCT snapshots\footnote{\url{https://github.com/njahn82/jct_data}}.
-The JCT journal data were enriched with ISSN variants linked to an
-ISSN-L. To identify hybrid journals, a comprehensive exclusion of fully
-open access journals was performed using multiple journal lists
-including the Directory of Open Access Journals (DOAJ). The resulting
-hybrid journal data were made available via hoaddata and used to
-determine the publication volume for each database independently.
-
-hoaddata relies on Crossref for obtaining journal publication volume and
+JCT journal data were enriched with ISSN variants linked to ISSN-L. To
+identify hybrid journals, a comprehensive exclusion of fully open access
+journals was performed using multiple journal lists including the
+Directory of Open Access Journals (DOAJ). The resulting hybrid journal
+data were made available via hoaddata and were used to independently
+determine the publication volume for each database.
+
+hoaddata relies on Crossref to obtain journal publication volume and
 open access status through Creative Commons (CC) licence information
 relative to the published version (``version of record''). The article
-metadata included DOIs, publication dates, open access information as
-well as author roles and affiliations. Publication years were determined
-using the earliest known date of publication in a journal. In hoaddata,
-this corresponded to Crossref's issued date. For Web of Science and
-Scopus, the earliest publication date was used where available, with
-Scopus dates specifically determined by the KB through version tracking
-of the raw data.
+metadata included DOIs, publication dates, open access information, and
+author roles and affiliations. Publication years were determined using
+the earliest known date of publication in a journal. In hoaddata, this
+corresponded to Crossref's issued date. For Web of Science and Scopus,
+the earliest publication date was used where available, with Scopus
+dates specifically determined by the KB through version tracking of the
+raw data.
 
@@
 exclude non-scholarly content (Piwowar et al., 2018). To exclude
-conference supplements, which are also often not covered by
-transformative agreements, only articles published in regular issues,
-indicated by numerical pagination, were considered. For Web of Science
-and Scopus, their established, mainly accurate document type
-classifications (Donner, 2017; Maisano et al., 2025) were used to
-identify original research articles and reviews, referred to as original
-articles throughout this study.
+conference supplements, which are often not covered by transformative
+agreements, only articles published in regular issues, indicated by
+numerical pagination, were considered. For Web of Science and Scopus,
+their established and mainly accurate document type classifications
+(Donner, 2017; Maisano et al., 2025) were used to identify original
+research articles and reviews, referred to as original articles
+throughout this study.
 
@@
 
-Articles in hybrid journals were considered open access when they were
-made freely available under a CC license on publisher platforms. While
+Articles in hybrid journals were considered open access when publishers
+made them freely available under a CC license on their platforms. While
 hoaddata obtained this information from Crossref license metadata, Web
@@
 This additional parsing remains necessary despite transformative
-agreement workflows recommend to deposit CC licenses during DOI
-registration (Geschuhn \& Stone, 2017). Both Web of Science\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
+agreement workflows recommending the deposition of CC licenses during
+DOI registration (Geschuhn \& Stone, 2017). Both Web of
+Science\footnote{\url{https://webofscience.help.clarivate.com/en-us/Content/open-access.html}}
 and Scopus\footnote{\url{https://blog.scopus.com/posts/scopus-filters-for-open-access-type-and-green-oa-full-text-access-option}}
@@
 
-Author affiliations were retrieved for both first and, if available,
-corresponding authors to prepare the linking between articles and
+Author affiliations were retrieved for both the first and, if available,
+the corresponding authors to prepare the linking between articles and
 institutions covered by transformative agreements. To improve data
-retrieval, JCT institution data was enriched with ROR-IDs from
+retrieval, JCT institution data were enriched with ROR-IDs from
 associated institutions, such as university hospitals or institutes of
 large research organisations such as the Max Planck Society, according
-to OpenAlex' institution entity. To handle different address variants,
-database-specific affiliation identifiers were used: ROR-IDs from
+to OpenAlex's institution entity. Database-specific affiliation
+identifiers were used to handle different address variants: ROR-IDs from
 OpenAlex for hoaddata, affiliation enhanced names for Web of Science,
@@
 Because neither Web of Science nor Scopus supported ROR-ID at the time
-of the data retrieval, the institution identifier used by the JCT, a
-two-step matching process was implemented to harmonise affiliation data.
-First, 2,782,540 articles from 6,457 institutions with ROR-IDs in the
-JCT data since 2017 (according to hoaddata) were processed to map first
-authors' ROR-IDs to corresponding proprietary affiliation identifier in
-Web of Science and Scopus using DOI matching. Then, an algorithm
-selected the most frequent ROR ID and proprietary identifier pairs to
-handle multiple affiliations and organisational hierarchy differences.
-
-This process linked 6,375 ROR-IDs to 4,894 Scopus Affiliation IDs, and
+of data retrieval, the institution identifier used by the JCT, a
+two-step matching process was implemented to harmonise the affiliation
+data. First, 2,782,540 articles from 6,457 institutions with ROR-IDs in
+the JCT data since 2017 (according to hoaddata) were processed to map
+the first authors' ROR-IDs to corresponding proprietary affiliation
+identifiers in Web of Science and Scopus using DOI matching. Then, an
+algorithm selected the most frequent ROR ID and proprietary identifier
+pairs to handle multiple affiliations and organisational hierarchy
+differences.
+
+This process linked 6,375 ROR-IDs to 4,894 Scopus Affiliation IDs and
 6,034 ROR-IDs to 2,422 enhanced affiliation strings in Web of Science.
 Quality evaluation through random sampling of 50 pairs revealed an error
-rate of 22\% for Web of Science (11 mismatches) and 6\% for Scopus (3
-mismatches). Upon inspection, these mismatches primarily occurred with
-less-represented institutions having only a few publications, introduced
-through multiple affiliations of single authors. The difference between
-databases suggests that Scopus' affiliation control aligns more closely
-with ROR than that of Web of Science.
+rate of 22\% for Web of Science (11 mismatches) and 6\% for Scopus
+(three mismatches). Upon inspection, these mismatches primarily occurred
+with less-represented institutions having only a few publications,
+introduced through multiple affiliations of single authors. The
+difference between databases suggests that Scopus' affiliation control
+aligns more closely with ROR than with Web of Science.
 
@@
 Scopus, although they did not contain ROR-IDs used by the JCT. The
-estimation of eligible articles followed Jahn (2025) and included a
+estimation of eligible articles followed Jahn (2025) and included
 matching of both journals and participating institutions according to
-the JCT. The matching also took into account the duration of agreements
+the JCT. The matching also considered the duration of agreements
 according to the ESAC registry, with only those matches where an
@@
 datasets on open access in hybrid journals included in transformative
-agreements were aggregated for each database at country and journal
-level by year. Table~\ref{tbl-methods_overview_tab} provides a general
-overview of the coverage between 2019 and 2023. It shows that the
-majority of hybrid journals published at least one original research
-article or review, collectively marked as ``original article'' in the
-following, during the five-year period. These journals provided the
-basis for the subsequent calculation of article-level indicators.
+agreements were aggregated for each database at the country and journal
+levels by year. Table~\ref{tbl-methods_overview_tab} provides a general
+overview of coverage between 2019 and 2023. It shows that the majority
+of hybrid journals published at least one original research article or
+review, collectively marked as ``original article'' in the following,
+during the five-year period. These journals have provided the basis for
+the subsequent calculation of article-level indicators.
 
@@
 
-While hoaddata only covers articles with a DOI, Scopus and Web of
+While hoaddata only covered articles with a DOI, Scopus and Web of
 Science publication indicators were calculated using their database
@@
 by DOI, because Unpaywall only collects information on open access
-status for articles with a DOI. A closer look at original articles with
-affiliation data, and in line with related research (Zhang et al.,
-2024), reveals a lack of affiliation data in the case of OpenAlex, the
+status for articles with DOI. A closer look at original articles with
+affiliation data, in line with related research (Zhang et al., 2024),
+reveals a lack of affiliation data in the case of OpenAlex, the
 affiliation data source used by hoaddata, compared to Web of Science and
-Scopus. In particular, only about two-thirds of the articles examined
-provided corresponding author affiliations. For first authors, the
-proportion was 89\%. At the time of writing, OpenAlex disclosed limited
-coverage of corresponding authorship data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
+Scopus. In particular, only approximately two-thirds of the articles
+examined provided corresponding author affiliations. The proportion for
+first authors was 89\%. At the time of writing, OpenAlex disclosed
+limited coverage of corresponding authorship data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
 Therefore, only first author data for hoaddata were considered in the
@@
 
-Throughout this largely automated data collection and analysis process,
-Tidyverse tools (Wickham et al., 2019) for the R programming language (R
-Core Team, 2024) were used. Rank correlations were calculated using the
-Hmisc package (Harrell Jr, 2003). The source code analysis, including
-all queries used to obtain the data, is available on GitHub (add link).
-An interactive supplement exploring correlations between the data
-sources examined by country and publisher is available on HuggingFace
-Spaces: https://huggingface.co/spaces/najkoja/hoa\_replication.
+TTidyverse tools (Wickham et al., 2019) for the R programming language
+(R Core Team, 2024) were used throughout this largely automated data
+collection and analysis process. Rank correlations were calculated using
+the Hmisc package (Harrell Jr, 2003). Source code analysis, including
+all queries used to obtain the data, is available on GitHub
+(\url{https://github.com/njahn82/hoa_validation}). An interactive
+supplement exploring correlations between the data sources examined by
+country and publisher is available via HuggingFace Spaces:
+\url{https://huggingface.co/spaces/najkoja/hoa_replication}.
 
@@
 Scopus and Web of Science. Then, using the same methods, indicators at
-the publisher and country level are calculated independently for each
-database and compared with each other to assess the suitability of the
-bibliometric databases for investigating transformative agreements.
+the publisher and country levels are calculated independently for each
+database and compared to assess the suitability of the bibliometric
+databases for investigating transformative agreements.
 
@@
 on open access status information from each database. Only original
-articles were considered in the analysis.
+articles were included in the analysis.
 
@@
 
-Journal coverage analysis revealed that 66\% (n = 7,970) of hybrid
+Journal coverage analysis revealed that 66\% (n = 7,970) of the hybrid
 journals included in transformative agreements were indexed in all three
-databases (Figure \ref{fig-upset_coverage_results}A). The second-largest
+databases (Figure \ref{fig-upset_coverage_results}A). The second largest
 set consisted of journals indexed in both hoaddata and Scopus,
@@
 of journals were exclusively contained in hoaddata, while another 6\% (n
-= 748) were only found in Scopus. Of these, 354 were also available in
-Web of Science. Upon inspection, this group of hybrid journals
-exclusively covered in the proprietary data sources mainly represented
-hybrid journals for which no open access evidence could be retrieved
-from Crossref, the open access evidence source for hoaddata.
+= 748) were found only in Scopus. Of these, 354 were available in Web of
+Science. Upon inspection, this group of hybrid journals exclusively
+covered in proprietary data sources mainly represented hybrid journals
+for which no open access evidence could be retrieved from Crossref, the
+open access evidence source for hoaddata.
 
@@
 shows the total publication volume per combination in terms of DOI
-availability. The largest set of hybrid journals, covering all three
-data sources, also represents the largest number of articles. In total,
-these journals recorded 6,289,687 articles, represented by the blue bar.
-They represented 94\% of original articles with a DOI indexed in the Web
-of Science sample, and 86\% in Scopus. Another 657,697 articles were
-exclusive to the intersection of Scopus and hoaddata. Exclusively in
-hoaddata were 177,110 articles, and exclusively in the proprietary
+availability. The largest set of hybrid journals covering all three data
+sources also represented the largest number of articles. In total, these
+journals recorded 6,289,687 articles, represented by the blue bars. They
+recorded 94\% of the original articles with a DOI indexed in the Web of
+Science sample, and 86\% in Scopus. Another 657,697 articles were
+exclusive to the intersection between Scopus and hoaddata. Exclusively
+in hoaddata were 177,110 articles, and exclusively in the proprietary
 databases were 325,194 articles.
@@
 the case of hybrid journals covered by all three data sources, 1,023,882
-DOIs were only present in hoaddata. After validation at the DOI level
-using the KB databases and manual inspection, the main reason for
-missing DOI coverage in the proprietary database were insufficient
-classification of journal content as original articles during the
-compilation of hoaddata. Particularly, letters and editorials could not
-be fully detected. Moreover, paratext recognition failed for 37\% of
-DOIs to identify non-scholarly content such as front matters or reviewer
-lists, which are generally not indexed by Scopus and Web of Science. To
-a lesser extent, differences in publication and indexing dates were a
-reason for non-overlapping DOIs.
+DOIs were present only in hoaddata. After validation at the DOI level
+using KB databases and manual inspection, the main reason for missing
+DOI coverage in the proprietary database was insufficient classification
+of journal content as original articles during the compilation of
+hoaddata. Particularly, letters and editorials could not be detected
+fully. Moreover, paratext recognition failed for 37\% of DOIs to
+identify non-scholarly content, such as front matters or reviewer lists,
+which are generally not indexed by Scopus and Web of Science. To a
+lesser extent, differences in publication and indexing dates were
+reasons for non-overlapping DOIs.
 
 Using overlapping DOIs, the publication volume between 2019 and 2023 was
-also calculated for each journal. Figure
-\ref{fig-upset_coverage_results}C illustrates the distribution for each
-combination. It shows a large spread oj journal size across the journals
-jointly covered by all three data sources. These journals published more
-on average than journals covered by only one or two data sources, with
-the exception of journals exclusively covered by Scopus and Web of
-Science. In particular, journals covered exclusively by hoaddata were
-substantially smaller. Upon inspection, these were often newly launched
-hybrid journals, which explains the relatively low five-year publication
-volume. An example is \emph{Digital society} that published 86 articles.
-This hybrid journal was launched in 2022, being covered by various
-Springer Nature transformative agreements since then.
+calculated for each journal. Figure \ref{fig-upset_coverage_results}C
+illustrates the distribution for each combination. It shows a large
+spread of journal size across journals mutually covered by all three
+data sources. These journals published more on average than journals
+covered by only one or two data sources, with the exception of journals
+exclusively covered by Scopus and Web of Science. In particular,
+journals covered exclusively by hoaddata were substantially smaller.
+Upon inspection, these were often newly launched hybrid journals, which
+explains the relatively low five-year publication volume. An example is
+\emph{Digital society} that published 86 articles. This hybrid journal
+was launched in 2022, and had since been covered by various Springer
+Nature transformative agreements.
 
@@
 of hybrid journals in transformative agreements across data sources from
-2019 to 2023 with a focus on publisher portfolios. The analysis
-highlights the dominance of the three largest publishers, Elsevier,
-Springer Nature, and Wiley, which collectively accounted for 46\% of
-hybrid journals and 62\% of articles published during this five-year
+2019 to 2023 focusing on publisher portfolios. The analysis highlights
+the dominance of the three largest publishers, Elsevier, Springer
+Nature, and Wiley, which collectively accounted for 46\% of the hybrid
+journals and 62\% of the articles published during this five-year
 period. In terms of article volume, Elsevier led with 2,435,478 articles
@@
 total). Springer Nature followed with 1,237,700 articles (17\%) in 2,294
-hybrid journals, although recording the largest number of hybrid
+hybrid journals, although it recorded the largest number of hybrid
 journals (19\%). Wiley accounted for 814,349 articles (11\%) in 1,337
@@
 The three largest publishers, Elsevier, Springer Nature, and Wiley, were
-best represented in the intersection of all three data sources
+best represented at the intersection of all three data sources
 (hoaddata, Scopus, and Web of Science). Together, they comprised 4,347
 hybrid journals (55\% of the intersectional set) and dominated article
-coverage (n = 4,143,663; 66\%), as determined through shared DOIs. When
-examining publication volume per journal, (Figure
+coverage (n = 4,143,663; 66\%) as determined through shared DOIs. When
+examining the publication volume per journal (Figure
 \ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
@@
 
-Comparing publisher portfolios across different indexing sets
+A comparison of publisher portfolios across different indexing sets
 demonstrates that publishers were not represented uniformly. Notably,
@@
 Russian-language works. Additionally, this subset included titles with a
-broader disciplinary focus such as \emph{SN Computer Science} and newly
-launched hybrid journals like \emph{Nature Computational Science}, which
-started in 2021 and was indexed in Scopus but not yet in Web of Science.
-The set also captured discontinued journals, providing further insights
-into the dynamics of journal publishing.
-
-Examining publisher portfolios not covered by hoaddata but present in
-Scopus or Web of Science identified several publishers with missing
+broader disciplinary focus, such as \emph{SN Computer Science}, and
+newly launched hybrid journals such as \emph{Nature Computational
+Science}, which started in 2021 and were indexed in Scopus but not yet
+in Web of Science. The set also captures discontinued journals,
+providing further insights into the dynamics of journal publishing.
+
+Examining publisher portfolios not covered by hoaddata, but present in
+Scopus or Web of Science, identified several publishers with missing
 Creative Commons (CC) license information in Crossref. In particular,
@@
 accounted for 24 journals with 64,898 articles, and World Scientific
-recorded 87 journals and 42,531 articles. In total, 9 publishers did not
-share CC licenses with Crossref and were therefore not represented in
-hoaddata.
+recorded 87 journals with 42,531 articles. In total, 9 publishers did
+not share CC licenses with Crossref and were therefore not represented
+in hoaddata.
 
@@
 subscription-only journals contained one or two articles erroneously
-tagged as hybrid open access by Unpaywall, which were subsequently
+tagged as hybrid open access by Unpaywall, which was subsequently
 reflected in Scopus and Web of Science. Examples of such
 misclassifications include Elsevier's \emph{Journal of Bioscience and
-Bioengineering} and Springer Nature's \emph{Journal of Mechanical
+Bioengineering}, and Springer Nature's \emph{Journal of Mechanical
 Science and Technology}.
@@
 Scopus, and Web of Science. The aim was to assess whether consistent
-results can be derived from these data sources despite differences in
+results could be derived from these data sources despite differences in
 coverage and methodologies. Following Jahn (2025), indicators were
-calculated for each data source and comprise the number and proportion
+calculated for each data source and comprised the number and proportion
 of open access articles, including those enabled by transformative
@@
 of transformative agreements was estimated using both first and
-corresponding authorships, while hoaddata indicator calculation was
+corresponding authorships, while hoaddata indicator calculations were
 limited to first author affiliations.
@@
 from 6.5\% (n = 84,648) in 2019 to 19\% (n = 322,850) in 2023. However,
-Web of Science recorded higher open access uptake in early years, before
-converging to a similar level in 2023, from 11\% (n = 137,202) in 2019
-to 20\% (n = 255,481) in 2023. This suggest a different approach towards
-labelling hybrid open access by Web of Science.
+Web of Science recorded higher open access uptake in the early years,
+before converging to a similar level in 2023, from 11\% (n = 137,202) in
+2019 to 20\% (n = 255,481) in 2023. This suggests a different approach
+towards labelling hybrid open access by Web of Science.
 
@@
 increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and
-C). Trends were consistent when measuring first (dashed line) and
-corresponding author (dotted line) affiliations. According to Scopus,
-479,297 open access articles could be attributed to transformative
-agreements based on first author metadata (increasing from 15,341 to
-206,084) and 489,262 using corresponding author metadata (from 15,444 to
-210,816). Web of Science recorded 493,028 open access articles via
-transformative agreements using first author metadata (increasing from
-21,871 to 174,126) and 500,076 using corresponding author metadata (from
-22,092 to 177,030). hoaddata, lacking corresponding author data, linked
-501,649 articles to transformative agreements, increasing from 16,010 in
-2019 to 183,757 in 2023.
-
-From 2021 (hoadata, Scopus) resp. 2022 (Web of Science), transformative
-agreements enabled the majority of open access articles in hybrid
-journals. For first authors, the share ranged between 61\% (hoaddata),
-64\% (Scopus), and 68 \% (Web of Science) in 2023. For corresponding
-authors, the shares were slightly larger, with Scopus recording 65\% and
-Web of Science 69\% in 2023. However, substantial hybrid open access was
-still facilitated outside transformative agreements, likely through APCs
-paid from discretionary research funds (Suber, 2012).
+C). Trends were consistent when measuring the first (dashed line) and
+corresponding authorship (dotted line). According to Scopus, 479,297
+open access articles could be attributed to transformative agreements
+based on first author metadata (increasing from 15,341 to 206,084) and
+489,262 using corresponding author metadata (from 15,444 to 210,816).
+Web of Science recorded 493,028 open access articles via transformative
+agreements using first author metadata (increasing from 21,871 to
+174,126) and 500,076 using corresponding author metadata (from 22,092 to
+177,030). hoaddata, lacking corresponding author data, linked 501,649
+articles to transformative agreements, increasing from 16,010 in 2019 to
+183,757 in 2023.
+
+Since 2021 (hoadata, Scopus) resp. 2022 (Web of Science), transformative
+agreements have enabled the majority of open access articles in hybrid
+journals. For the first authors, the share ranged between 61\%
+(hoaddata), 64\% (Scopus), and 68 \% (Web of Science) by 2023. For the
+corresponding authors, the shares were slightly larger, with Scopus
+recording 65\% and Web of Science 69\% by 2023. However, substantial
+hybrid open access was still facilitated outside transformative
+agreements, likely through APCs paid from discretionary research funds
+(Suber, 2012).
 
@@
 to 15\% according to Web of Science. Upon inspection, this discrepancy
-is primarily due to articles in the publisher's open archive. These
+was primarily due to articles under the publisher's open archive. These
 articles, made freely available after an embargo period under Elsevier's
@@
 compared to hoaddata. For these two years, hoaddata reported 32,445 more
-open access articles than Scopus and 29,854 more than Web of Science.
-This discrepancy is presumably due to challenges in fetching full-texts
-by Unpaywall, the open access evidence source for Scopus and Web of
-Science. According to Unpaywall's software version history, HTTP
-redirects from Wiley's publisher platform prevented Unpaywall from
-parsing license information from full-texts.\footnote{See Unpaywall
-  version history related to Wiley fixes:
+open access articles than Scopus and 29,854 more open access articles
+than Web of Science. This discrepancy is presumably due to challenges in
+fetching full-texts by Unpaywall, the open access evidence source for
+Scopus and Web of Science. According to Unpaywall's software version
+history, HTTP redirects from Wiley's publisher platform, preventing
+Unpaywall from parsing license information from full-texts.\footnote{See
+  Unpaywall version history related to Wiley fixes:
   \url{https://github.com/search?q=repo\%3Aourresearch\%2Foadoi+wiley&type=commits}}
@@
 reported as open access in 2023 across the examined data sources,
-followed by Springer Nature. Elsevier, recorded a later uptake, which is
+followed by Springer Nature. Elsevier recorded a later uptake,
 consistent with the publisher's historical reluctance to engage in
-negotiations with library consortia. However, by 2023, the share of open
-access enabled by transformative agreements appeared to stabilise for
-all three publishers (see Figure \ref{fig-uptake_publisher}C).
-Interestingly, the differences between first and corresponding author
-affiliations were more pronounced at the publisher level. In Scopus, for
-example, the share of open access via transformative agreements measured
-by corresponding authorship was higher for Springer Nature in 2023 than
-when using first authorship.
+negotiations with library consortia (Fraser et al., 2023). However, by
+2023, the share of open access enabled by transformative agreements
+appeared to stabilise for all three publishers (see Figure
+\ref{fig-uptake_publisher}C). Interestingly, the differences between
+first and corresponding authorships were more pronounced at the
+publisher level. In Scopus, for example, the share of open access via
+transformative agreements measured by the corresponding authorship was
+higher for Springer Nature in 2023 than when using the first authorship.
 
@@
 When comparing countries, consistent patterns were observed across data
-sources for the five-year period 2019 to 2023. Figure
+sources for the five-year period from 2019 to 2023. Figure
 \ref{fig-uptake_country} presents hybrid open access indicators by
@@
 (y-axis). Indicators calculated from these proprietary databases are
-shown for both first and corresponding authors, with full counting used
-to account for multiple country affiliations (Hottenrott et al., 2021).
+shown for both the first and corresponding authors, with full counting
+used to account for multiple country affiliations (Hottenrott et al.,
+2021).
 
@@
 \ref{fig-uptake_country}A), strong positive correlations were observed
-across the data sources and author roles (Spearman rank correlation
+across data sources and author roles (Spearman rank correlation
 \(\rho > .9\), \(p < 0.001\)). Between 2019 and 2023, China was the most
 productive country, followed by the United States and, by a certain
-margin, India, the United Kingdom, and Germany. Analysis of authorship
-roles revealed minimal variation, indicating that first and
-corresponding authors were typically from the same country.
+margin, India, the United Kingdom, and Germany. The analysis of
+authorship roles revealed minimal variation, indicating that the first
+and corresponding authors were typically from the same country.
 
 When examining the percentage of open access articles in hybrid journals
-(see Figure \ref{fig-uptake_country}B), a different pattern emerged.
-Authors affiliated with institutions from medium-sized European
-countries, such as Sweden, the Netherlands, Finland, and Hungary,
-provided a large proportion of their articles in open access. Germany
-and the United Kingdom also had approximately 40\% of their output
-available as open access. In contrast, non-OECD countries showed notably
-lower adoption of hybrid open access, with South Africa being the only
-BRICS member well-represented in the data. The United States also
-demonstrated a relatively low proportion of open access articles. These
-findings were consistent across all databases. However, France was
-better represented in Web of Science, likely due to its agreement with
-Elsevier starting in 2019, which allowed delayed open access under the
+(Figure \ref{fig-uptake_country}B), a different pattern emerged. Authors
+affiliated with institutions from medium-sized European countries, such
+as Sweden, the Netherlands, Finland, and Hungary, made a large
+proportion of their articles open access. Germany and the United Kingdom
+also had approximately 40\% of their output available as open access. In
+contrast, non-OECD countries showed notably lower adoption of hybrid
+open access, with South Africa being the only BRICS member
+well-represented in the data. The United States also demonstrated a
+relatively low proportion of open access articles. These findings were
+consistent across all databases. However, France was better represented
+in Web of Science, likely because of its agreement with Elsevier
+starting in 2019, which allowed delayed open access under the
 publisher's user license (Rabesandratana, 2019). This licence was not
@@
 cases, Spearman rank correlations were \(\rho > .9\), \(p < 0.001\),
-showing a high level of correlations between the databases and
+indicating a high level of correlation between the databases and
 authorship roles considered.
@@
 agreements with major publishers from 2022 onward.\footnote{\url{https://sanlic.ac.za/read-and-publish-agreements/}}
-Results were consistent across data sources. However, Wiley's open
+The results were consistent across data sources. However, Wiley's open
 access surplus in hoaddata led to better rankings for countries where
 Wiley played a substantial role, such as Germany, where the DEAL
-consortium negotiated its first transformative agreements with publisher
-that started in July 2019.
+consortium negotiated its first transformative agreements with
+publishers that started in July 2019.
 
@@
 adoption rates. In contrast, the three most productive countries, China,
-the United States, and India, showcase a substantially lower adoption in
-transformative agreements. Open questions remain whether this uneven
-distribution reflects temporary implementation gaps, inherent inequities
-in the transformative agreement model, or a deliberate avoidance of such
-agreements.
+the United States, and India, show a substantially lower adoption in
+transformative agreements. Open questions remain as to whether this
+uneven distribution reflects temporary implementation gaps, inherent
+inequities in the transformative agreement model, or deliberate
+avoidance of such agreements.
 
@@
 hoaddata, derived from Crossref and OpenAlex, and the established
-proprietary bibliometric databases Scopus and Web of Science. Notably,
-and aligning with previous studies (Akbaritabar et al., 2024; Alperin et
+proprietary bibliometric databases Scopus and Web of Science. This is
+consistent with previous studies (Akbaritabar et al., 2024; Alperin et
 al., 2024) and rankings (N. J. van Eck et al., 2024). Overall, the
 results show strong correlations by country affiliation, which supports
-using open metadata for large-scale analyses of hybrid open access.
+the use open metadata for large-scale analyses of hybrid open access.
 However, the observed differences in journal coverage and metadata
-availability warrant further discussion in the following, affecting not
-only open data sources but also proprietary databases when used in
-isolation.
-
-The coverage analysis reveals that hybrid journals are well indexed in
+availability warrant further discussion, affecting not only open data
+sources but also proprietary databases when used in isolation.
+
+The coverage analysis revealed that hybrid journals are well indexed in
 all three data sources, particularly in terms of article coverage,
@@
 exclusively in Crossref and Scopus. Using Crossref as a bibliometric
-database in hoaddata demonstrated particular strength in identifying
+database in hoaddata demonstrated a particular strength in identifying
 newly established hybrid journals, a notable finding given that
@@
 regard to contributorship and author roles and positions (Larivière et
-al., 2016). First authors typically conduct the main research underlying
-a paper, while the corresponding author often supervises the research
-(Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly, measures
-based on first or corresponding authorships are strongly correlated,
-suggesting that these authors share the same country affiliation.
-Moreover, in most cases, the first author is identical to the
+al., 2016). The first author typically conduct the main research
+underlying a paper, while the corresponding author often supervises the
+research (Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly,
+measures based on first or corresponding authorships are strongly
+correlated, suggesting that these authors share the same country
+affiliation. In most cases, the first author is identical to the
 corresponding author (Chinchilla-Rodríguez et al., 2024). Despite this
@@
 authorship positions (Helgesson, 2020), suggesting the need to monitor
-potential influences of the availability of open access funding on
+the potential influences of the availability of open access funding on
 authorship roles (Maddi \& Silva, 2024).
@@
 comprehensive, which likely reduced the number of articles confidently
-attributed to transformative agreements in this study, it still shows
+attributed to transformative agreements in this study, it still showed
 high correlations with Scopus and Web of Science at the country level.
@@
 reconciliation with proprietary organisation identifiers. Future studies
-based on the Web of Science will benefit from the recent integration of
+based on Web of Science will benefit from the recent integration of
 ROR-IDs, announced by the end of 2024.
@@
 licences, it nevertheless missed a substantial number of CC-licensed
-open access articles from Wiley journals indexed in Crossref, likely due
-to parsing errors on the publisher's website. This resulted in fewer
-open access articles being recorded in Scopus and Web of Science.
+open access articles from Wiley journals indexed in Crossref, likely
+because of parsing errors on the publisher's website. This resulted in
+fewer open access articles being recorded in Scopus and Web of Science.
 Further inconsistencies emerged between Scopus and Web of Science,
@@
 categorises it. Notably, Scopus and Crossref showed greater alignment
-than observed in a related comparison between Scopus and OpenAlex
+than that observed in a related comparison between Scopus and OpenAlex
 (Alperin et al., 2024). This likely reflects the use of a curated list
-to identify hybrid open access journals, rather than relying solely on
+to identify hybrid open access journals, rather than depending solely on
 article-level tags. Given these variations, which may compromise
-comparability, research and monitoring exercises should avoid reliance
-on a single source. Instead, the selection of open access evidence
-should be cross-verified using multiple sources and snapshots that can
-be used to track changes in the data over time (Huang et al., 2020).
+comparability, research and monitoring exercises should avoid relying on
+a single source. Instead, the selection of open access evidence should
+be cross-verified using multiple sources and snapshots that can be used
+to track changes in the data over time (Huang et al., 2020).
 Incorporating expert-curated journal lists can also help prevent
@@
 institutions involved in transformative agreements are approximations
-due to a lack of access to invoice data, which is not usually shared by
-library consortia and publishers. Furthermore, the data sources about
-transformative agreements including the cOAlition S Journal Checker Tool
-and the ESAC registry are based on a voluntary effort, crowd-sourced
-from various consortia. However, recent validation with Dutch research
+because of a lack of access to invoice data, which is not usually shared
+by library consortia and publishers. Furthermore, the open data sources
+for transformative agreements, including the cOAlition S Journal Checker
+Tool and the ESAC registry, are based on voluntary effort, crowd-sourced
+from various consortia. However, recent validation using Dutch research
 information demonstrates the reliability of such an open approach for
 the assessment of articles under transformative agreements (Jonge et
-al., 2025). With enhanced open metadata compliance, attributable to
+al., 2025). With enhanced open metadata compliance attributable to
 evolving standards and initiatives in support of negotiations with
-publishers, particularly through the communities of the ESAC initiative
-and Barcelona Declaration supporters, the situation is likely to
+publishers, particularly through the ESAC initiative and Barcelona
+Declaration on Open Research Information, the situation is likely to
 improve. In conclusion, this study has shown that open metadata are
@@
 
-\section{References}\label{references}
+\section*{Author contributions}\label{author-contributions}
+\addcontentsline{toc}{section}{Author contributions}
+
+The author confirms the sole responsibility for the following:
+Conceptualization, Methodology, Formal analysis, Investigation, Data
+Curation, Writing - Original Draft, Writing - Review \& Editing,
+Visualization.
+
+\section*{Competing interests}\label{competing-interests}
+\addcontentsline{toc}{section}{Competing interests}
+
+The author declares no competing interests.
+
+\section*{Funding information}\label{funding-information}
+\addcontentsline{toc}{section}{Funding information}
+
+This work was supported by the Federal Ministry of Education and
+Research of Germany (BMBF) under grants 16WIK2301E / 16WIK2101F.
+
+The proprietary bibliometric data from Scopus and Web of Science were
+provided by the German Competence Network for Bibliometrics under BMBF
+grant 16WIK2101A.
+
+\section*{Data and code availability}\label{data-and-code-availability}
+\addcontentsline{toc}{section}{Data and code availability}
+
+Data source code supplement is available via GitHub:
+\url{https://github.com/njahn82/hoa_validation}. An interactive
+dashboard, which allows the browsing of metrics by country and
+publisher, is available on HuggingFace Spaces:
+\url{https://huggingface.co/spaces/najkoja/hoa_replication}.
+
+Please note that according to the licence terms for Scopus and Web of
+Science, as provided by the German Competence Network for Bibliometrics,
+only aggregated data accompanying the scientific publication can be made
+publicly available for replication purposes.
+
+Open metadata provided by Crossref, OpenAlex, and cOAlition S Journal
+Checker Transformative Agreement Data are freely available for re-use.
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
@@
 Alperin, J. P., Portenoy, J., Demes, K., Larivière, V., \& Haustein, S.
-(2024). An analysis of the suitability of OpenAlex for bibliometric
-analyses. \emph{arXiv}. \url{https://arxiv.org/abs/2404.17663}
+(2024). \emph{An analysis of the suitability of OpenAlex for
+bibliometric analyses}. arXiv. \url{https://arxiv.org/abs/2404.17663}
 
@@
 Bellen, S. van, Alperin, J. P., \& Larivière, V. (2024). \emph{The
-oligopoly of academic publishers persists in exclusive database}.
+oligopoly of academic publishers persists in exclusive database}. arXiv.
 \url{https://arxiv.org/abs/2406.17893}
@@
 \bibitem[\citeproctext]{ref-culbert2024referencecoverageanalysisopenalex}
-Culbert, J., Hobert, A., Jahn, N., Haupka, N., Schmidt, M., Donner, P.,
-\& Mayr, P. (2024). \emph{Reference coverage analysis of OpenAlex
-compared to web of science and scopus}.
-\url{https://arxiv.org/abs/2401.16359}
+Culbert, J. H., Hobert, A., Jahn, N., Haupka, N., Schmidt, M., Donner,
+P., \& Mayr, P. (2025). Reference coverage analysis of {OpenAlex}
+compared to {Web of Science} and {Scopus}. \emph{Scientometrics}.
+\url{https://doi.org/10.1007/s11192-025-05293-3}
 
@@
 \bibitem[\citeproctext]{ref-hoaddata}
-Jahn, N. (2024). \emph{Hoaddata: Data about hybrid open access journal
-publishing}. \url{https://github.com/subugoe/hoaddata}
+Jahn, N. (2024). \emph{{hoaddata}: Data about hybrid open access journal
+publishing (v.0.3)}.
+\url{https://github.com/subugoe/hoaddata/releases/tag/v.0.3}
 
@@
 \bibitem[\citeproctext]{ref-ComplexUpset}
-Krassowski, M. (2020). \emph{ComplexUpset}.
+Krassowski, M. (2020). \emph{ComplexUpset}. Zenodo.
 \url{https://doi.org/10.5281/zenodo.3700590}
@@
 transformative agreements? A multifaceted analysis of \textgreater1,000
-contracts}. \url{https://arxiv.org/abs/2409.20224}
+contracts}. arXiv. \url{https://arxiv.org/abs/2409.20224}
 
@@
 (2024). \emph{The open access coverage of {OpenAlex}, {Scopus} and {Web
-of Science}}. \url{https://doi.org/10.48550/ARXIV.2404.01985}
+of Science}}. arXiv. \url{https://doi.org/10.48550/ARXIV.2404.01985}
 
@@
 Singh, V. K., Singh, P., Karmakar, M., Leta, J., \& Mayr, P. (2021). The
-journal coverage of web of science, scopus and dimensions: A comparative
-analysis. \emph{Scientometrics}, \emph{126}(6), 5113--5142.
+journal coverage of {Web of Science}, {Scopus} and {Dimensions}: A
+comparative analysis. \emph{Scientometrics}, \emph{126}(6), 5113--5142.
 \url{https://doi.org/10.1007/s11192-021-03948-5}
@@
 Missing institutions in OpenAlex: Possible reasons, implications, and
-solutions. \emph{Scientometrics}.
+solutions. \emph{Scientometrics}, \emph{129}(10), 5869--5891.
 \url{https://doi.org/10.1007/s11192-023-04923-y}

### 2025-04-21 analysis/manuscript.qmd
Improve intro

@@
 
-Transformative agreements are a much-discussed library licensing model for transitioning subscription-based journal publishing to full open access.
+<!--Transformative agreements are a much-discussed library licensing model for transitioning subscription-based journal publishing to full open access.
 Although these agreements can vary considerably, they are mainly aimed at hybrid journal bundles, allowing authors from participating institutions to publish open access in these journals while providing reading access to the entire portfolio.
@@
 By comparing open metadata from hoaddata, an openly available dataset on hybrid open access based on metadata from Crossref, OpenAlex and the cOAlition S Journal Checker Tool, with the established bibliometric databases Scopus and Web of Science, this study aims to provide a robust estimate of the impact of transformative agreements on hybrid open access.
-This comparative approach informs bibliometric research and practice by highlighting the strengths and weaknesses of each data source individually and collectively.
-
-Since their initial proposal by the Max Planck Digital Library [@Schimmer_2015], transformative agreements have led to rapid growth in the number of negotiated agreements and open access articles [@Jahn_2025; @Rothfritz_2024; @McCabe_2024].
-Before, only a small proportion of articles in hybrid journals were openly available [@Bj_rk_2012; @Laakso_2016; @Piwowar_2018], with publications from only a few European countries with dedicated funding policies and the high-energy physics SCOAP^3 consortium standing out [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020; @Kohls_2018].
-Related coordination activities, in particular through the OA2020 and ESAC (Efficiency and Standards for Article Charges) initiatives, have also led to improved workflows between library consortia and publishers [@ifla; @Geschuhn_2017].
-However, critics have raised concerns about these agreements, particularly regarding market concentration [@Shu_2023; @Butler_2023], the failure of most hybrid journals to convert to full open access [@Momeni_2021; @Kiley_2024], reduced competition and incentives [@McCabe_2024; @Schmal_competition], and the continuation of publication fees that widen gaps between well-resourced and under-resourced institutions [@budapest; @Ross_Hellauer_2022]
-
-While consortia evaluations confirm the growth of open access through transformative agreements and acknowledge increasing coordination, they reach mixed policy conclusions.
-For instance, the British Jisc evaluation [@Jisc_2024] found that the consortium's transformative agreements had a significant impact on national open access growth but a limited effect on the global open access transition, suggesting a need to reevaluate their strategy.
+This comparative approach informs bibliometric research and practice by highlighting the strengths and weaknesses of each data source individually and collectively.-->
+
+Transformative agreements are a much-discussed library licensing model for transitioning subscription-based journal publishing to full open access.
+Although these agreements can vary considerably, they are mainly aimed at hybrid journal bundles, enabling authors from participating institutions to publish open access in these journals while providing reading access to the entire portfolio [@Borrego_2020; @Hinchliffe_2019].
+Following their initial proposal in 2015 [@Schimmer_2015], the number of transformative agreements has grown substantially, resulting in increased open access publishing in hybrid journals [@Jahn_2025; @Rothfritz_2024; @McCabe_2024].
+Prior to these agreements, only a few articles in hybrid journals were openly accessible [@Bj_rk_2012; @Laakso_2016; @Piwowar_2018], except for some European countries that had dedicated funding policies and the SCOAP$^3$ consortium for high-energy physics journals [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020; @Kohls_2018].
+Coordination efforts through initiatives such as OA2020 and Efficiency and Standards for Article Charges (ESAC) have further streamlined workflows between library consortia and publishers [@ifla; @Geschuhn_2017].
+By April 2025, the ESAC Transformative Agreement Registry, the primary community-driven evidence source, listed approximately 1,300 agreements.[^esac_registry]
+
+However, critics have raised concerns about these agreements, particularly regarding perpetuating issues related to hybrid open access such as charging twice for reading and publishing ("double dipping") [@Asai_2023], market concentration [@Shu_2023; @Butler_2023], and the failure of most hybrid journals to convert to full open access [@Momeni_2021; @Kiley_2024].
+Additional concerns include reduced competition and incentives [@McCabe_2024; @Schmal_competition], and the continuation of publication fees that widen gaps between well-resourced and under-resourced institutions [@budapest; @Ross_Hellauer_2022].
+
+Consortia evaluations generally confirm increased open access through transformative agreements and better coordination. 
+However, they reached mixed conclusions about the effectiveness.
+For instance, the British Jisc evaluation [@Jisc_2024] found that the consortium's transformative agreements had a significant impact on national open access growth but a limited effect on the global open access transition, suggesting a need to reevaluate its strategy.
 Similar policy reconsiderations emerged in Norway and Sweden [@norway; @Widding_2024].
 Meanwhile, the funder initiative cOAlition S recommended ending financial support for these agreements in 2024, but still considers hybrid open access to be compliant with funders' open access policies.
-Conversely, the German DEAL consortium extended agreements with major publishers until 2028, while also outside of Europe, more transformative agreements are being put in place (for instance in Latin America, as described in @Mu_oz_V_lez_2024).
-Responding to criticism, the Max Planck Digital Library recently outlined strategies to evolve the original model, focusing on achieving fully open access while controlling costs and addressing identified weaknesses [@D_r_2025]
-
-Despite controversies, transformative agreements can benefit the landscape of open research information as library consortia negotiate for comprehensive open metadata, helping avoid data analytics from becoming a commodity controlled by publishers [@Aspesi_2020; @McCabe_2024].
-For instance, ESAC recommends using Crossref [@Hendricks_2020] to share metadata supporting workflows including open access licenses and funding information [@Geschuhn_2017]. 
-Simultaneously, the bibliometric community pushes for more open metadata through this publisher-driven DOI registration platform and continuously monitors the progress [@eck_crossref_2022].
-
-Many open scholarly data services that build on Crossref have benefited from these developments.
-A prominent example is OpenAlex, which has gained attention for its inclusivity, openness and comprehensive coverage of publication metadata [@priem2022openalex].
-In a recent evaluation, @C_spedes_2025 reviewed the current evidence base on the strengths and weaknesses of OpenAlex.
-They highlight both its increasing use in bibliometric research and its adoption by research organisations.
-Comparative studies have generally found OpenAlex to be a reliable alternative to commercial databases [@alperin2024analysissuitabilityopenalexbibliometric; @culbert2024referencecoverageanalysisopenalex], although data quality issues occurred for key metadata fields required for open access analysis, including evidence of open access status itself [@jahn2023analysing; @Simard_2024], as well as authorships [@culbert2024referencecoverageanalysisopenalex], affiliations [@Zhang_2024], author roles [@Jahn_2025] and document types used to distinguish between different article types [@Haupka_2024].
-However, these studies highlight the rapid improvements and new possibilities of combining open metadata and the curation efforts of OpenAlex in conjunction with the growing awareness of the importance of open metadata, as demonstrated by the Barcelona Declaration on Open Research Information.
-
-Consequently, there is growing interest in using open bibliometric databases such as Crossref and OpenAlex to assess the impact of transformative agreements. 
-Similar to established proprietary bibliometric databases such as Scopus and Web of Science, these data sources lack specific data points, particularly about open access payments, which are typically not shared [@Haustein_APC_2024].
-However, bibliometric data can be enriched with open metadata about journals and institutions involved in transformative agreements [@Jahn_2025]. 
-@de_Jonge_2025 demonstrated this potential by combining data from the cOAlition S Journal Checker tool and OpenAlex, achieving 89% accuracy when validated against Dutch research information. 
-While open data approaches show promise, understanding how these results compare when the same methodology is applied to proprietary databases such as Web of Science and Scopus remains unexplored.
-
-This study compares open and proprietary data sources for evaluating transformative agreements.
-It adapts the approach from @Jahn_2025, which combines data from the cOAlition S Journal Checker tool, Crossref and OpenAlex, and applies the same methods to Web of Science and Scopus data.
-This comparative approach allows the evaluation of different bibliometric data sources to monitor transformative agreements, while assessing their respective strengths and limitations.
-Specifically, this study compares the indexing of more than 13,000 hybrid journals across these sources, and examines whether consistent results can be obtained in terms of open access adoption between 2019 and 2023, analysed by publishers and countries.
-By comparing results across these data sources, this large-scale study provides a more robust understanding of the impact of transformative agreements on the transition to full open access. 
-It also reveals the comparative strengths and weaknesses of different bibliometric databases in measuring scholarly publishing's shift toward openness.
+By contrast, the German DEAL consortium extended agreements with major publishers until 2028, and countries in the Global South have adopted transformative agreements [@Mu_oz_V_lez_2024]. 
+Responding to criticism, the Max Planck Digital Library recently outlined strategies to evolve the original model, which continue to focus on achieving fully open access while controlling costs and addressing identified shortcomings [@D_r_2025].
+
+These discussions highlight the need for robust data to evaluate transformative agreements. 
+Importantly, these agreements themselves can help improve this evidence base, as researchers and libraries advocate for including open metadata in negotiations to help avoid data analytics from becoming a commodity controlled by publishers [@Aspesi_2020; @McCabe_2024].
+For instance, ESAC recommends using Crossref to share metadata supporting workflows including open access licenses and funding information to support discovery and monitoring [@Geschuhn_2017]. 
+Simultaneously, the bibliometric community pushes for more open metadata through this publisher-driven DOI registration platform and continuously monitors progress [@van_Eck_2022].
+
+This push for comprehensive metadata has yielded results. 
+A growing number of open scholarly data services that build on publisher-provided metadata through Crossref have benefited from increasing metadata coverage, although differences between publishers can be observed [@van_Eck_2022].
+OpenAlex stands out as a prominent example that primarily retrieves updated records from Crossref and has gained attention for its inclusivity, openness and comprehensive coverage [@priem2022openalex].
+Comparative studies confirm it as a suitable alternative to proprietary bibliometric databases [@alperin2024analysissuitabilityopenalexbibliometric; @culbert2024referencecoverageanalysisopenalex].
+Notably, @C_spedes_2025 synthesised evidence on both OpenAlex's strengths and its shortcomings for bibliometric analysis.
+These include challenges with open access status identification [@jahn2023analysing; @Simard_2024], author attribution [@culbert2024referencecoverageanalysisopenalex], institutional affiliations [@Zhang_2024], and document type classification [@Haupka_2024].
+
+While promising for general bibliometric research, applying open metadata to transformative agreements presents distinct challenges due to their targeted journal coverage and complex funding attribution.
+With invoicing data largely unavailable, previous studies have relied on publication fee pricing lists or expenditures, and first and/or corresponding author affiliations as proxies to attribute funding to institutions [@Butler_2023; @Zhang_2022; @Haustein_APC_2024].
+Recent work has extended this approach and attributed open access articles to transformative agreements by using data from the cOAlition S Journal Checker Tool, which links journals and participating institutions to transformative agreements included in the ESAC registry [@Jahn_2025; @de_Jonge_2025].
+Yet, no systematic comparison exists of how open versus proprietary data sources perform when analysing the same hybrid journals covered by transformative agreements to identify comparative strengths and shortcomings of data sources.
+
+The present study addresses this gap by comparing open and proprietary data sources for evaluating transformative agreements. 
+It adapts the approach from @Jahn_2025, which combines data from the cOAlition S Journal Checker tool, Crossref as publication metadata source including open access and OpenAlex as author affiliation data source, and applies the same methods to the established databases Scopus and Web of Science.
+
+This comparative approach examines different bibliometric data sources through two primary research questions:
+
+1. How does the coverage of hybrid journals included in transformative agreements compare across Crossref, Scopus and Web of Science in terms of journal indexing and article representation?
+
+2.
+To what extent do estimates of articles enabled by transformative agreements between 2019 and 2023 differ when using different affiliation data sources (OpenAlex, Scopus and Web of Science) and different authorship attribution methods (first versus corresponding authorship)?
+
+By analysing more than 13,000 hybrid journals across these sources, this study provides insights into whether consistent results can be obtained when measuring open access adoption across publishers and countries.
+The findings enhance understanding of transformative agreements' impact on the transition to full open access while revealing the comparative strengths and limitations of different bibliometric databases in measuring scholarly publishing's shift toward openness.
 
@@
 <!-- This study investigates the suitability of open scholarly data sources for assessing the impact of transformative agreements on hybrid open access. -->
-As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Web of Science and Scopus.
+As shown in Figure \ref{fig:workflow}, the methodology involved comparing hoaddata, an openly available collection of open research information on hybrid open access, with the bibliometric databases Scopus and Web of Science.
 This section introduces the initial data sources, followed by a presentation of the necessary data processing steps to obtain eligible articles enabled by transformative agreements using author roles (first and corresponding) and harmonised affiliation data.
@@
 
-hoaddata, developed and maintained by the author to support open access monitoring and research (e.g.
-@Jahn_2025), is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
-It combines article-level metadata from Crossref [@Hendricks_2020] and affilation metadata from OpenAlex [@priem2022openalex] with transformative agreement information from the Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data about participating research organisations to agreements in the ESAC registry.
+hoaddata, developed and maintained by the author to support open access monitoring and research [@Achterberg_2023; @Jahn_2025], is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
+It combines article-level metadata from Crossref [@Hendricks_2020] and affilation metadata from OpenAlex [@priem2022openalex] with transformative agreement information from the Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data about participating research organisations to agreements in the ESAC Transformative Agreement Registry[^esac_registry].
 
@@
 
-hoaddata followed good practices for computational reproducibility using R [@Marwick_2018].
+hoaddata follows good practices for computational reproducibility using R [@Marwick_2018].
 The package, which includes data, code, a test suite and documentation, is openly available on GitHub.
@@
 Clarivate Analytics' Web of Science (WoS) is a well-established proprietary bibliometric database consisting of several collections [@Birkle_2020].
-Web of Science is a selective database with a focus on base research [@Visser_2021; @Stahlschmidt_2022].
+Web of Science is selective regarding journal indexing with a focus on base research [@Visser_2021; @Stahlschmidt_2022].
 The collections considered in this study were the Science Citation Index Expanded (SCIE), Social Sciences Citation Index (SSCI) and Arts & Humanities Citation Index (AHCI).
@@
 Elsevier's Scopus, launched in 2004, is another widely used proprietary bibliometric database for measuring research [@Baas_2020].
-Similar to Web of Science, Scopus is selective with regard to the journals it indexes.
-However, its journal coverage is much broader than that of the Web of Science collections considered in this study, as it indexes a wider range of applied research journals [@Singh_2021; @Visser_2021; @Stahlschmidt_2022].
+Similar to Web of Science, Scopus is selective, but its journal coverage is much broader than that of the Web of Science collections considered in this study, as it indexes a wider range of applied research journals [@Singh_2021; @Visser_2021; @Stahlschmidt_2022].
 With detailed metadata about article types, open access status information derived from Unpaywall, author roles, and disambiguated affiliations, Scopus also contains important data to assess open access uptake, although no direct information regarding transformative agreements was available at the time of the study.
@@
 Following @Jahn_2025, the starting point was a unified dataset of several safeguarded JCT snapshots[^jct_data].
-JCT journal data were enriched with ISSN variants linked to ISSN-L. 
+JCT journal data were enriched with ISSN variants linked to ISSN-L according to the ISSN agency. 
 To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists including the Directory of Open Access Journals (DOAJ).
@@
 First, `r length(unique(ta_oa_inst$doi))` articles from `r length(unique(ta_oa_inst$ror))` institutions with ROR-IDs in the JCT data since 2017 (according to hoaddata) were processed to map the first authors' ROR-IDs to corresponding proprietary affiliation identifiers in Web of Science and Scopus using DOI matching.
-Then, an algorithm selected the most frequent ROR ID and proprietary identifier pairs to handle multiple affiliations and organisational hierarchy differences.
+Then, an algorithm selected the most frequent ROR-ID and proprietary organisation identifier pairs to handle multiple affiliations and organisational hierarchy differences.
 
@@
 Upon inspection, these mismatches primarily occurred with less-represented institutions having only a few publications, introduced through multiple affiliations of single authors. 
-The difference between databases suggests that Scopus' affiliation control aligns more closely with ROR than with Web of Science.
+The difference between databases suggests that Scopus's affiliation control is more closely aligned with ROR than the Web of Science is with ROR.
 
@@
       indicator == "ind_active_jns" ~ "Active journals",
-      indicator == "ind_active_jns_core" ~ "Active journals (core)",
-      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (core) with OA",
+      indicator == "ind_active_jns_core" ~ "Active journals (original)",
+      indicator == "ind_active_jns_core_with_oa" ~ "Active journals (original) with OA",
       indicator == "ind_articles_total" ~ "Total published articles ",
@@
 This was particularly the case for meeting abstracts, which are notably prevalent in Health Sciences journals [@Melero_Fuentes_2025] and are not indexed by Scopus [@Donner_2017].
-Open access indicators were aggregated by DOI, because Unpaywall only collects information on open access status for articles with DOI.
-A closer look at original articles with affiliation data, in line with related research [@Zhang_2024], reveals a lack of affiliation data in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
+Open access indicators were aggregated by DOI, because Unpaywall only collects information on open access status for articles in Crossref.
+A closer look at original articles with affiliation data reveals a lack of data on affiliations and author roles in the case of OpenAlex, the affiliation data source used by hoaddata, compared to Web of Science and Scopus.
 In particular, only approximately two-thirds of the articles examined provided corresponding author affiliations.
@@
 
-TTidyverse tools [@tidyverse] for the R programming language [@r] were used throughout this largely automated data collection and analysis process. 
+Tidyverse tools [@tidyverse] for the R programming language [@r] were used throughout this largely automated data collection and analysis process. 
 Rank correlations were calculated using the Hmisc package [@Harrell_Jr_2003].
@@
 
-This section first presents the indexing of hybrid open access by comparing open data sources with the proprietary bibliometric databases Scopus and Web of Science. 
-Then, using the same methods, indicators at the publisher and country levels are calculated independently for each database and compared to assess the suitability of the bibliometric databases for investigating transformative agreements.
+This section first presents the indexing of hybrid open access by comparing open metadata with the proprietary bibliometric databases Scopus and Web of Science. 
+Then, using the same methods, indicators at the publisher and country levels are calculated independently for each database and compared with each other.
 
@@
 Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid journals included in transformative agreements, based on the intersections of the examined data sources. 
-The intersection sets of journals and articles are visualised as an UpSet graph [@Lex_2014; @ComplexUpset].
+The intersection sets of journals (\ref{fig-upset_coverage_results}A) and articles (\ref{fig-upset_coverage_results}B= are visualised as an UpSet graph [@Lex_2014; @ComplexUpset].
+The x-axis represents the set intersections in a matrix layout.
 The analysis included hybrid journals that published at least one open access article between 2019 and 2023, based on open access status information from each database.
@@
 Notably, `r round(jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(WoS == FALSE, Crossref == TRUE, Scopus == FALSE) |>  nrow()`) of journals were exclusively contained in hoaddata, while another `r round( jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |> distinct(issn_l) |> nrow() / nrow(all_jns) * 100)`% (n = `r jn_upset_articles |> filter(Crossref == FALSE, Scopus == TRUE) |>  nrow()`) were found only in Scopus.
-Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were available in Web of Science.
-Upon inspection, this group of hybrid journals exclusively covered in proprietary data sources mainly represented hybrid journals for which no open access evidence could be retrieved from Crossref, the open access evidence source for hoaddata.
+Of these, `r jn_upset_articles |> filter(WoS == TRUE, Crossref == FALSE, Scopus == TRUE) |>  nrow()` were also available in Web of Science.
+Upon inspection, the group of hybrid journals exclusively covered in proprietary data sources mainly represented hybrid journals for which no open access evidence could be retrieved from Crossref, the open access evidence source for hoaddata.
 
@@
 In total, these journals recorded `r upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()` articles, represented by the blue bars.
-They recorded `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(wos) * 100)`% of the original articles with a DOI indexed in the Web of Science sample, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(scp) * 100)`% in Scopus.
+They recorded `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(wos) * 100)`% of the original articles with a DOI indexed in Web of Science, and `r round(upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum() / methods_overview_table_df |> filter(indicator == "Original articles with DOI") |> pull(scp) * 100)`% in Scopus.
 Another `r upset_data |> filter(Crossref == TRUE, WoS == FALSE, Scopus == TRUE) |> pull(n_cr) |> sum()` articles were exclusive to the intersection between Scopus and hoaddata.
@@
 Together, they comprised `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum()` hybrid journals (`r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(n_journals) |> sum() / nrow(union_jns)*100`% of the intersectional set) and dominated article coverage (n = `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum()`; `r jn_upset_publisher_ind |> filter(publisher != "Other") |> pull(articles) |> sum() / upset_data |> filter(Crossref == TRUE, WoS == TRUE, Scopus == TRUE) |> pull(overlap) |> sum()*100`%) as determined through shared DOIs.
-When examining the publication volume per journal (Figure \ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals, followed by Springer Nature and Wiley.
+When examining the publication volume per journal (Figure \ref{fig-upset_coverage_results_publisher}C), Elsevier published, on average, the largest journals.
 
@@
 
-Differences in open access evidence are also evident for Wiley. 
+Differences in open access evidence are also apparent for Wiley. 
 Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022 compared to hoaddata. 
@@
 This discrepancy is presumably due to challenges in fetching full-texts by Unpaywall, the open access evidence source for Scopus and Web of Science.
-According to Unpaywall's software version history, HTTP redirects from Wiley's publisher platform, preventing Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
+According to Unpaywall's software version history, HTTP redirects from Wiley's publisher platform prevented Unpaywall from parsing license information from full-texts.[^unpaywallwiley]
 hoaddata, which relies solely on Crossref metadata for open access identification, was unaffected by these issues.
@@
 Between 2019 and 2023, China was the most productive country, followed by the United States and, by a certain margin, India, the United Kingdom, and Germany. 
-The analysis of authorship roles revealed minimal variation, indicating that the first and corresponding authors were typically from the same country.
+The analysis of authorship roles revealed minimal variation, indicating that the first and corresponding author of an article were typically from the same country.
 
@@
 Germany and the United Kingdom also had approximately 40% of their output available as open access.
-In contrast, non-OECD countries showed notably lower adoption of hybrid open access, with South Africa being the only BRICS member well-represented in the data.
+In contrast, non-OECD countries showed notably lower adoption of hybrid open access, with South Africa being the only BRICS member (as of 2022) well-represented in the data.
 The United States also demonstrated a relatively low proportion of open access articles.
@@
 Hybrid open access and transformative agreements remain concentrated among a small group of large commercial publishers, 
-with high-income European countries—alongside South Africa—showing high adoption rates.
+with European countries—alongside South Africa—showing high adoption rates.
 In contrast, the three most productive countries, China, the United States, and India, show a substantially lower adoption in transformative agreements.
@@
 The findings were consistent across the investigated open data source hoaddata, derived from Crossref and OpenAlex, and the established proprietary bibliometric databases Scopus and Web of Science. 
-This is consistent with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73].
+This aligns with previous studies [@Akbaritabar_2024; @alperin2024analysissuitabilityopenalexbibliometric] and rankings [@eck_2024_6bx73].
 Overall, the results show strong correlations by country affiliation, which supports the use open metadata for large-scale analyses of hybrid open access. 
@@
 
-The coverage analysis revealed that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage, reflecting the dominance of major publishers whose established journal portfolios are comprehensively indexed in proprietary databases [@vanbellen2024oligopolyacademicpublisherspersists].
+The coverage analysis revealed that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage, which reflects the dominance of major publishers whose established journal portfolios are comprehensively indexed in proprietary databases [@vanbellen2024oligopolyacademicpublisherspersists].
 Differences emerge for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
@@
 A frequently reported limitation of studying open access with less selective databases is the lack of corresponding authorship information [@Fraser_2023; @Haucap_2021; @Shu_2023]. 
-However, this analysis demonstrates that indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, reflecting disciplinary norms in scholarly publishing with regard to contributorship and author roles and positions [@Larivi_re_2016].
+However, this analysis demonstrates that indicators based on first authors, which have often been used as a proxy for determining open access funding, and corresponding authors show a high level of correlation, reflecting disciplinary norms in scholarly publishing with regard to contributions, author roles and positions [@Larivi_re_2016; @Zhang_2022].
 The first author typically conduct the main research underlying a paper, while the corresponding author often supervises the research [@Mattsson_2010; @Fox_2018]. 
@@
 <!-- Thus, first authorship can reliably be used in large-scale studies to assess the impact of transformative agreements. -->
-Despite this correlation, the study observed a recent slight increase in open access articles by corresponding authors over first authors across Springer Nature hybrid journals.
+Despite this correlation, the study observed a recent slight increase in open access articles attributed to transformative agreements by corresponding authors over first authors across Springer Nature hybrid journals.
 This prompts questions about how institutional open access sponsorship practices influence author roles and assignments within co-author teams, especially as funding opportunities vary [@Gumpenberger_2018].
@@
 In conclusion, this study has shown that open metadata are well-suited for analysing the transition from subscription-based journal publishing to full open access.
+But using them in conjunction with proprietary databases provides a more robust understanding that overcomes individual shortcomings in coverage and metadata quality.
 
@@
 [^oacorresponding]: <https://docs.openalex.org/api-entities/works/work-object/authorship-object#is_corresponding>
+
+[^esac_registry]: <https://esac-initiative.org/about/transformative-agreements/agreement-registry/>

### 2025-04-21 analysis/manuscript.tex
Improve intro

@@
 access. Although these agreements can vary considerably, they are mainly
-aimed at hybrid journal bundles, allowing authors from participating
+aimed at hybrid journal bundles, enabling authors from participating
 institutions to publish open access in these journals while providing
-reading access to the entire portfolio. Library consortia and funders
-have begun to coordinate how these agreements are negotiated, managed
-and evaluated. At the same time, efforts are underway to improve
-transparency through open data on contracts and payments. However, this
-information is still only partially available, which limits large-scale
-studies of the impact of transformative agreements on the transition of
-hybrid journals to full open access. By comparing open metadata from
-hoaddata, an openly available dataset on hybrid open access based on
-metadata from Crossref, OpenAlex and the cOAlition S Journal Checker
-Tool, with the established bibliometric databases Scopus and Web of
-Science, this study aims to provide a robust estimate of the impact of
-transformative agreements on hybrid open access. This comparative
-approach informs bibliometric research and practice by highlighting the
-strengths and weaknesses of each data source individually and
-collectively.
-
-Since their initial proposal by the Max Planck Digital Library (Schimmer
-et al., 2015), transformative agreements have led to rapid growth in the
-number of negotiated agreements and open access articles (Jahn, 2025;
-McCabe \& Mueller-Langer, 2024; Rothfritz et al., 2024). Before, only a
-small proportion of articles in hybrid journals were openly available
-(Björk, 2012; Laakso \& Björk, 2016; Piwowar et al., 2018), with
-publications from only a few European countries with dedicated funding
-policies and the high-energy physics SCOAP\^{}3 consortium standing out
-(Huang et al., 2020; Kohls \& Mele, 2018; Pinfield, 2015;
-Robinson-Garcia et al., 2020). Related coordination activities, in
-particular through the OA2020 and ESAC (Efficiency and Standards for
-Article Charges) initiatives, have also led to improved workflows
-between library consortia and publishers (Campbell et al., 2022;
-Geschuhn \& Stone, 2017). However, critics have raised concerns about
-these agreements, particularly regarding market concentration (Butler et
-al., 2023; Shu \& Larivière, 2023), the failure of most hybrid journals
-to convert to full open access (Kiley, 2024; Momeni et al., 2021),
-reduced competition and incentives (McCabe \& Mueller-Langer, 2024;
-Schmal, 2024), and the continuation of publication fees that widen gaps
-between well-resourced and under-resourced institutions (Babini et al.,
-2022; Ross-Hellauer et al., 2022)
-
-While consortia evaluations confirm the growth of open access through
-transformative agreements and acknowledge increasing coordination, they
-reach mixed policy conclusions. For instance, the British Jisc
-evaluation (Brayman et al., 2024) found that the consortium's
+reading access to the entire portfolio (Borrego et al., 2021;
+Hinchliffe, 2019). Following their initial proposal in 2015 (Schimmer et
+al., 2015), the number of transformative agreements has grown
+substantially, resulting in increased open access publishing in hybrid
+journals (Jahn, 2025; McCabe \& Mueller-Langer, 2024; Rothfritz et al.,
+2024). Prior to these agreements, only a few articles in hybrid journals
+were openly accessible (Björk, 2012; Laakso \& Björk, 2016; Piwowar et
+al., 2018), except for some European countries that had dedicated
+funding policies and the SCOAP\(^3\) consortium for high-energy physics
+journals (Huang et al., 2020; Kohls \& Mele, 2018; Pinfield, 2015;
+Robinson-Garcia et al., 2020). Coordination efforts through initiatives
+such as OA2020 and Efficiency and Standards for Article Charges (ESAC)
+have further streamlined workflows between library consortia and
+publishers (Campbell et al., 2022; Geschuhn \& Stone, 2017). By April
+2025, the ESAC Transformative Agreement Registry, the primary
+community-driven evidence source, listed approximately 1,300
+agreements.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}
+
+However, critics have raised concerns about these agreements,
+particularly regarding perpetuating issues related to hybrid open access
+such as charging twice for reading and publishing (``double dipping'')
+(Asai, 2023), market concentration (Butler et al., 2023; Shu \&
+Larivière, 2023), and the failure of most hybrid journals to convert to
+full open access (Kiley, 2024; Momeni et al., 2021). Additional concerns
+include reduced competition and incentives (McCabe \& Mueller-Langer,
+2024; Schmal, 2024), and the continuation of publication fees that widen
+gaps between well-resourced and under-resourced institutions (Babini et
+al., 2022; Ross-Hellauer et al., 2022).
+
+Consortia evaluations generally confirm increased open access through
+transformative agreements and better coordination. However, they reached
+mixed conclusions about the effectiveness. For instance, the British
+Jisc evaluation (Brayman et al., 2024) found that the consortium's
 transformative agreements had a significant impact on national open
 access growth but a limited effect on the global open access transition,
-suggesting a need to reevaluate their strategy. Similar policy
+suggesting a need to reevaluate its strategy. Similar policy
 reconsiderations emerged in Norway and Sweden (Holden et al., 2023;
@@
 considers hybrid open access to be compliant with funders' open access
-policies. Conversely, the German DEAL consortium extended agreements
-with major publishers until 2028, while also outside of Europe, more
-transformative agreements are being put in place (for instance in Latin
-America, as described in Muñoz-Vélez et al. (2024)). Responding to
-criticism, the Max Planck Digital Library recently outlined strategies
-to evolve the original model, focusing on achieving fully open access
-while controlling costs and addressing identified weaknesses (Dér, 2025)
-
-Despite controversies, transformative agreements can benefit the
-landscape of open research information as library consortia negotiate
-for comprehensive open metadata, helping avoid data analytics from
-becoming a commodity controlled by publishers (Aspesi \& Brand, 2020;
-McCabe \& Mueller-Langer, 2024). For instance, ESAC recommends using
-Crossref (Hendricks et al., 2020) to share metadata supporting workflows
-including open access licenses and funding information (Geschuhn \&
-Stone, 2017). Simultaneously, the bibliometric community pushes for more
-open metadata through this publisher-driven DOI registration platform
-and continuously monitors the progress (N. J. van Eck \& Waltman, 2022).
-
-Many open scholarly data services that build on Crossref have benefited
-from these developments. A prominent example is OpenAlex, which has
-gained attention for its inclusivity, openness and comprehensive
-coverage of publication metadata (Priem et al., 2022). In a recent
-evaluation, Céspedes et al. (2025) reviewed the current evidence base on
-the strengths and weaknesses of OpenAlex. They highlight both its
-increasing use in bibliometric research and its adoption by research
-organisations. Comparative studies have generally found OpenAlex to be a
-reliable alternative to commercial databases (Alperin et al., 2024;
-Culbert et al., 2025), although data quality issues occurred for key
-metadata fields required for open access analysis, including evidence of
-open access status itself (Jahn et al., 2023; Simard et al., 2024), as
-well as authorships (Culbert et al., 2025), affiliations (Zhang et al.,
-2024), author roles (Jahn, 2025) and document types used to distinguish
-between different article types (Haupka et al., 2024). However, these
-studies highlight the rapid improvements and new possibilities of
-combining open metadata and the curation efforts of OpenAlex in
-conjunction with the growing awareness of the importance of open
-metadata, as demonstrated by the Barcelona Declaration on Open Research
-Information.
-
-Consequently, there is growing interest in using open bibliometric
-databases such as Crossref and OpenAlex to assess the impact of
-transformative agreements. Similar to established proprietary
-bibliometric databases such as Scopus and Web of Science, these data
-sources lack specific data points, particularly about open access
-payments, which are typically not shared (Haustein et al., 2024).
-However, bibliometric data can be enriched with open metadata about
-journals and institutions involved in transformative agreements (Jahn,
-2025). Jonge et al. (2025) demonstrated this potential by combining data
-from the cOAlition S Journal Checker tool and OpenAlex, achieving 89\%
-accuracy when validated against Dutch research information. While open
-data approaches show promise, understanding how these results compare
-when the same methodology is applied to proprietary databases such as
-Web of Science and Scopus remains unexplored.
-
-This study compares open and proprietary data sources for evaluating
-transformative agreements. It adapts the approach from Jahn (2025),
-which combines data from the cOAlition S Journal Checker tool, Crossref
-and OpenAlex, and applies the same methods to Web of Science and Scopus
-data. This comparative approach allows the evaluation of different
-bibliometric data sources to monitor transformative agreements, while
-assessing their respective strengths and limitations. Specifically, this
-study compares the indexing of more than 13,000 hybrid journals across
-these sources, and examines whether consistent results can be obtained
-in terms of open access adoption between 2019 and 2023, analysed by
-publishers and countries. By comparing results across these data
-sources, this large-scale study provides a more robust understanding of
-the impact of transformative agreements on the transition to full open
-access. It also reveals the comparative strengths and weaknesses of
-different bibliometric databases in measuring scholarly publishing's
-shift toward openness.
+policies. By contrast, the German DEAL consortium extended agreements
+with major publishers until 2028, and countries in the Global South have
+adopted transformative agreements (Muñoz-Vélez et al., 2024). Responding
+to criticism, the Max Planck Digital Library recently outlined
+strategies to evolve the original model, which continue to focus on
+achieving fully open access while controlling costs and addressing
+identified shortcomings (Dér, 2025).
+
+These discussions highlight the need for robust data to evaluate
+transformative agreements. Importantly, these agreements themselves can
+help improve this evidence base, as researchers and libraries advocate
+for including open metadata in negotiations to help avoid data analytics
+from becoming a commodity controlled by publishers (Aspesi \& Brand,
+2020; McCabe \& Mueller-Langer, 2024). For instance, ESAC recommends
+using Crossref to share metadata supporting workflows including open
+access licenses and funding information to support discovery and
+monitoring (Geschuhn \& Stone, 2017). Simultaneously, the bibliometric
+community pushes for more open metadata through this publisher-driven
+DOI registration platform and continuously monitors progress (van Eck \&
+Waltman, 2024).
+
+This push for comprehensive metadata has yielded results. A growing
+number of open scholarly data services that build on publisher-provided
+metadata through Crossref have benefited from increasing metadata
+coverage, although differences between publishers can be observed (van
+Eck \& Waltman, 2024). OpenAlex stands out as a prominent example that
+primarily retrieves updated records from Crossref and has gained
+attention for its inclusivity, openness and comprehensive coverage
+(Priem et al., 2022). Comparative studies confirm it as a suitable
+alternative to proprietary bibliometric databases (Alperin et al., 2024;
+Culbert et al., 2025). Notably, Céspedes et al. (2025) synthesised
+evidence on both OpenAlex's strengths and its shortcomings for
+bibliometric analysis. These include challenges with open access status
+identification (Jahn et al., 2023; Simard et al., 2024), author
+attribution (Culbert et al., 2025), institutional affiliations (Zhang et
+al., 2024), and document type classification (Haupka et al., 2024).
+
+While promising for general bibliometric research, applying open
+metadata to transformative agreements presents distinct challenges due
+to their targeted journal coverage and complex funding attribution. With
+invoicing data largely unavailable, previous studies have relied on
+publication fee pricing lists or expenditures, and first and/or
+corresponding author affiliations as proxies to attribute funding to
+institutions (Butler et al., 2023; Haustein et al., 2024; Zhang et al.,
+2022). Recent work has extended this approach and attributed open access
+articles to transformative agreements by using data from the cOAlition S
+Journal Checker Tool, which links journals and participating
+institutions to transformative agreements included in the ESAC registry
+(Jahn, 2025; Jonge et al., 2025). Yet, no systematic comparison exists
+of how open versus proprietary data sources perform when analysing the
+same hybrid journals covered by transformative agreements to identify
+comparative strengths and shortcomings of data sources.
+
+The present study addresses this gap by comparing open and proprietary
+data sources for evaluating transformative agreements. It adapts the
+approach from Jahn (2025), which combines data from the cOAlition S
+Journal Checker tool, Crossref as publication metadata source including
+open access and OpenAlex as author affiliation data source, and applies
+the same methods to the established databases Scopus and Web of Science.
+
+This comparative approach examines different bibliometric data sources
+through two primary research questions:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  How does the coverage of hybrid journals included in transformative
+  agreements compare across Crossref, Scopus and Web of Science in terms
+  of journal indexing and article representation?
+\item
+  To what extent do estimates of articles enabled by transformative
+  agreements between 2019 and 2023 differ when using different
+  affiliation data sources (OpenAlex, Scopus and Web of Science) and
+  different authorship attribution methods (first versus corresponding
+  authorship)?
+\end{enumerate}
+
+By analysing more than 13,000 hybrid journals across these sources, this
+study provides insights into whether consistent results can be obtained
+when measuring open access adoption across publishers and countries. The
+findings enhance understanding of transformative agreements' impact on
+the transition to full open access while revealing the comparative
+strengths and limitations of different bibliometric databases in
+measuring scholarly publishing's shift toward openness.
 
@@
 comparing hoaddata, an openly available collection of open research
-information on hybrid open access, with the bibliometric databases Web
-of Science and Scopus. This section introduces the initial data sources,
-followed by a presentation of the necessary data processing steps to
-obtain eligible articles enabled by transformative agreements using
-author roles (first and corresponding) and harmonised affiliation data.
+information on hybrid open access, with the bibliometric databases
+Scopus and Web of Science. This section introduces the initial data
+sources, followed by a presentation of the necessary data processing
+steps to obtain eligible articles enabled by transformative agreements
+using author roles (first and corresponding) and harmonised affiliation
+data.
 
@@
 hoaddata, developed and maintained by the author to support open access
-monitoring and research (e.g. Jahn (2025)), is an R data package that
-regularly collects information on hybrid open access uptake from
-multiple openly available data sources (Jahn, 2024). It combines
-article-level metadata from Crossref (Hendricks et al., 2020) and
-affilation metadata from OpenAlex (Priem et al., 2022) with
+monitoring and research (Achterberg \& Jahn, 2023; Jahn, 2025), is an R
+data package that regularly collects information on hybrid open access
+uptake from multiple openly available data sources (Jahn, 2024). It
+combines article-level metadata from Crossref (Hendricks et al., 2020)
+and affilation metadata from OpenAlex (Priem et al., 2022) with
 transformative agreement information from the Transformative Agreement
@@
 which links journal and institutional data about participating research
-organisations to agreements in the ESAC registry.
-
-hoaddata followed good practices for computational reproducibility using
+organisations to agreements in the ESAC Transformative Agreement
+Registry\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}.
+
+hoaddata follows good practices for computational reproducibility using
 R (Marwick et al., 2018). The package, which includes data, code, a test
@@
 proprietary bibliometric database consisting of several collections
-(Birkle et al., 2020). Web of Science is a selective database with a
-focus on base research (Stahlschmidt \& Stephen, 2022; Visser et al.,
-2021). The collections considered in this study were the Science
-Citation Index Expanded (SCIE), Social Sciences Citation Index (SSCI)
-and Arts \& Humanities Citation Index (AHCI).
+(Birkle et al., 2020). Web of Science is selective regarding journal
+indexing with a focus on base research (Stahlschmidt \& Stephen, 2022;
+Visser et al., 2021). The collections considered in this study were the
+Science Citation Index Expanded (SCIE), Social Sciences Citation Index
+(SSCI) and Arts \& Humanities Citation Index (AHCI).
 
@@
 bibliometric database for measuring research (Baas et al., 2020).
-Similar to Web of Science, Scopus is selective with regard to the
-journals it indexes. However, its journal coverage is much broader than
-that of the Web of Science collections considered in this study, as it
-indexes a wider range of applied research journals (Singh et al., 2021;
-Stahlschmidt \& Stephen, 2022; Visser et al., 2021). With detailed
-metadata about article types, open access status information derived
-from Unpaywall, author roles, and disambiguated affiliations, Scopus
-also contains important data to assess open access uptake, although no
-direct information regarding transformative agreements was available at
-the time of the study.
+Similar to Web of Science, Scopus is selective, but its journal coverage
+is much broader than that of the Web of Science collections considered
+in this study, as it indexes a wider range of applied research journals
+(Singh et al., 2021; Stahlschmidt \& Stephen, 2022; Visser et al.,
+2021). With detailed metadata about article types, open access status
+information derived from Unpaywall, author roles, and disambiguated
+affiliations, Scopus also contains important data to assess open access
+uptake, although no direct information regarding transformative
+agreements was available at the time of the study.
 
@@
 several safeguarded JCT snapshots\footnote{\url{https://github.com/njahn82/jct_data}}.
-JCT journal data were enriched with ISSN variants linked to ISSN-L. To
-identify hybrid journals, a comprehensive exclusion of fully open access
-journals was performed using multiple journal lists including the
-Directory of Open Access Journals (DOAJ). The resulting hybrid journal
-data were made available via hoaddata and were used to independently
-determine the publication volume for each database.
+JCT journal data were enriched with ISSN variants linked to ISSN-L
+according to the ISSN agency. To identify hybrid journals, a
+comprehensive exclusion of fully open access journals was performed
+using multiple journal lists including the Directory of Open Access
+Journals (DOAJ). The resulting hybrid journal data were made available
+via hoaddata and were used to independently determine the publication
+volume for each database.
 
@@
 identifiers in Web of Science and Scopus using DOI matching. Then, an
-algorithm selected the most frequent ROR ID and proprietary identifier
-pairs to handle multiple affiliations and organisational hierarchy
-differences.
+algorithm selected the most frequent ROR-ID and proprietary organisation
+identifier pairs to handle multiple affiliations and organisational
+hierarchy differences.
 
@@
 introduced through multiple affiliations of single authors. The
-difference between databases suggests that Scopus' affiliation control
-aligns more closely with ROR than with Web of Science.
+difference between databases suggests that Scopus's affiliation control
+is more closely aligned with ROR than the Web of Science is with ROR.
 
@@
 \hspace{1em}Active journals & 12,890 & 8,655 & 11,888\\
-\hspace{1em}Active journals (core) & 12,888 & 8,655 & 11,878\\
-\hspace{1em}Active journals (core) with OA & 11,348 & 8,392 & 11,313\\
+\hspace{1em}Active journals (original) & 12,888 & 8,655 & 11,878\\
+\hspace{1em}Active journals (original) with OA & 11,348 & 8,392 & 11,313\\
 \addlinespace[0.3em]
@@
 by DOI, because Unpaywall only collects information on open access
-status for articles with DOI. A closer look at original articles with
-affiliation data, in line with related research (Zhang et al., 2024),
-reveals a lack of affiliation data in the case of OpenAlex, the
-affiliation data source used by hoaddata, compared to Web of Science and
-Scopus. In particular, only approximately two-thirds of the articles
-examined provided corresponding author affiliations. The proportion for
-first authors was 89\%. At the time of writing, OpenAlex disclosed
-limited coverage of corresponding authorship data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
+status for articles in Crossref. A closer look at original articles with
+affiliation data reveals a lack of data on affiliations and author roles
+in the case of OpenAlex, the affiliation data source used by hoaddata,
+compared to Web of Science and Scopus. In particular, only approximately
+two-thirds of the articles examined provided corresponding author
+affiliations. The proportion for first authors was 89\%. At the time of
+writing, OpenAlex disclosed limited coverage of corresponding authorship
+data \footnote{\url{https://docs.openalex.org/api-entities/works/work-object/authorship-object\#is_corresponding}}.
 Therefore, only first author data for hoaddata were considered in the
@@
 
-TTidyverse tools (Wickham et al., 2019) for the R programming language
-(R Core Team, 2024) were used throughout this largely automated data
+Tidyverse tools (Wickham et al., 2019) for the R programming language (R
+Core Team, 2024) were used throughout this largely automated data
 collection and analysis process. Rank correlations were calculated using
@@
 This section first presents the indexing of hybrid open access by
-comparing open data sources with the proprietary bibliometric databases
+comparing open metadata with the proprietary bibliometric databases
 Scopus and Web of Science. Then, using the same methods, indicators at
 the publisher and country levels are calculated independently for each
-database and compared to assess the suitability of the bibliometric
-databases for investigating transformative agreements.
+database and compared with each other.
 
@@
 intersections of the examined data sources. The intersection sets of
-journals and articles are visualised as an UpSet graph (Krassowski,
-2020; Lex et al., 2014). The analysis included hybrid journals that
-published at least one open access article between 2019 and 2023, based
-on open access status information from each database. Only original
-articles were included in the analysis.
+journals (\ref{fig-upset_coverage_results}A) and articles
+(\ref{fig-upset_coverage_results}B= are visualised as an UpSet graph
+(Krassowski, 2020; Lex et al., 2014). The x-axis represents the set
+intersections in a matrix layout. The analysis included hybrid journals
+that published at least one open access article between 2019 and 2023,
+based on open access status information from each database. Only
+original articles were included in the analysis.
 
@@
 of journals were exclusively contained in hoaddata, while another 6\% (n
-= 748) were found only in Scopus. Of these, 354 were available in Web of
-Science. Upon inspection, this group of hybrid journals exclusively
-covered in proprietary data sources mainly represented hybrid journals
-for which no open access evidence could be retrieved from Crossref, the
-open access evidence source for hoaddata.
+= 748) were found only in Scopus. Of these, 354 were also available in
+Web of Science. Upon inspection, the group of hybrid journals
+exclusively covered in proprietary data sources mainly represented
+hybrid journals for which no open access evidence could be retrieved
+from Crossref, the open access evidence source for hoaddata.
 
@@
 journals recorded 6,289,687 articles, represented by the blue bars. They
-recorded 94\% of the original articles with a DOI indexed in the Web of
-Science sample, and 86\% in Scopus. Another 657,697 articles were
-exclusive to the intersection between Scopus and hoaddata. Exclusively
-in hoaddata were 177,110 articles, and exclusively in the proprietary
-databases were 325,194 articles.
+recorded 94\% of the original articles with a DOI indexed in Web of
+Science, and 86\% in Scopus. Another 657,697 articles were exclusive to
+the intersection between Scopus and hoaddata. Exclusively in hoaddata
+were 177,110 articles, and exclusively in the proprietary databases were
+325,194 articles.
 
@@
 the dominance of the three largest publishers, Elsevier, Springer
-Nature, and Wiley, which collectively accounted for 46\% of the hybrid
+Nature, and Wiley, which collectively accounted for 47\% of the hybrid
 journals and 62\% of the articles published during this five-year
-period. In terms of article volume, Elsevier led with 2,435,478 articles
-(34\% of the total) published across 1,941 hybrid journals (16\% of the
-total). Springer Nature followed with 1,237,700 articles (17\%) in 2,294
+period. In terms of article volume, Elsevier led with 2,441,358 articles
+(33\% of the total) published across 1,951 hybrid journals (16\% of the
+total). Springer Nature followed with 1,247,578 articles (17\%) in 2,311
 hybrid journals, although it recorded the largest number of hybrid
-journals (19\%). Wiley accounted for 814,349 articles (11\%) in 1,337
+journals (19\%). Wiley accounted for 858,939 articles (12\%) in 1,382
 hybrid journals (11\%). The remaining 54 publishers collectively
-accounted for 2,733,994 articles (38\%) in 6,442 hybrid journals (54\%).
+accounted for 2,749,847 articles (38\%) in 6,476 hybrid journals (53\%).
 
@@
 best represented at the intersection of all three data sources
-(hoaddata, Scopus, and Web of Science). Together, they comprised 4,347
+(hoaddata, Scopus, and Web of Science). Together, they comprised 4,384
 hybrid journals (55\% of the intersectional set) and dominated article
-coverage (n = 4,143,663; 66\%) as determined through shared DOIs. When
+coverage (n = 4,174,315; 66\%) as determined through shared DOIs. When
 examining the publication volume per journal (Figure
 \ref{fig-upset_coverage_results_publisher}C), Elsevier published, on
-average, the largest journals, followed by Springer Nature and Wiley.
+average, the largest journals.
 
@@
 demonstrates that publishers were not represented uniformly. Notably,
-Springer Nature exhibited 517 hybrid journals exclusively indexed in
+Springer Nature exhibited 519 hybrid journals exclusively indexed in
 both hoaddata and Scopus. This set included journals from the Chinese
@@
 Creative Commons (CC) license information in Crossref. In particular,
-Emerald represented 321 journals with 86,269 articles, AIP Publishing
+Emerald represented 322 journals with 86,409 articles, AIP Publishing
 accounted for 24 journals with 64,898 articles, and World Scientific
@@
 
-Differences in open access evidence are also evident for Wiley.
+Differences in open access evidence are also apparent for Wiley.
 Specifically, Web of Science and Scopus recorded a drop in 2021 and 2022
-compared to hoaddata. For these two years, hoaddata reported 32,445 more
-open access articles than Scopus and 29,854 more open access articles
+compared to hoaddata. For these two years, hoaddata reported 35,308 more
+open access articles than Scopus and 32,491 more open access articles
 than Web of Science. This discrepancy is presumably due to challenges in
@@
 Scopus and Web of Science. According to Unpaywall's software version
-history, HTTP redirects from Wiley's publisher platform, preventing
+history, HTTP redirects from Wiley's publisher platform prevented
 Unpaywall from parsing license information from full-texts.\footnote{See
@@
 authorship roles revealed minimal variation, indicating that the first
-and corresponding authors were typically from the same country.
+and corresponding author of an article were typically from the same
+country.
 
@@
 contrast, non-OECD countries showed notably lower adoption of hybrid
-open access, with South Africa being the only BRICS member
+open access, with South Africa being the only BRICS member (as of 2022)
 well-represented in the data. The United States also demonstrated a
@@
 concentrated among a small group of large commercial publishers, with
-high-income European countries---alongside South Africa---showing high
-adoption rates. In contrast, the three most productive countries, China,
-the United States, and India, show a substantially lower adoption in
+European countries---alongside South Africa---showing high adoption
+rates. In contrast, the three most productive countries, China, the
+United States, and India, show a substantially lower adoption in
 transformative agreements. Open questions remain as to whether this
@@
 hoaddata, derived from Crossref and OpenAlex, and the established
-proprietary bibliometric databases Scopus and Web of Science. This is
-consistent with previous studies (Akbaritabar et al., 2024; Alperin et
-al., 2024) and rankings (N. J. van Eck et al., 2024). Overall, the
-results show strong correlations by country affiliation, which supports
-the use open metadata for large-scale analyses of hybrid open access.
-However, the observed differences in journal coverage and metadata
-availability warrant further discussion, affecting not only open data
-sources but also proprietary databases when used in isolation.
+proprietary bibliometric databases Scopus and Web of Science. This
+aligns with previous studies (Akbaritabar et al., 2024; Alperin et al.,
+2024) and rankings (Eck et al., 2024). Overall, the results show strong
+correlations by country affiliation, which supports the use open
+metadata for large-scale analyses of hybrid open access. However, the
+observed differences in journal coverage and metadata availability
+warrant further discussion, affecting not only open data sources but
+also proprietary databases when used in isolation.
 
 The coverage analysis revealed that hybrid journals are well indexed in
-all three data sources, particularly in terms of article coverage,
-reflecting the dominance of major publishers whose established journal
+all three data sources, particularly in terms of article coverage, which
+reflects the dominance of major publishers whose established journal
 portfolios are comprehensively indexed in proprietary databases (Bellen
@@
 correlation, reflecting disciplinary norms in scholarly publishing with
-regard to contributorship and author roles and positions (Larivière et
-al., 2016). The first author typically conduct the main research
-underlying a paper, while the corresponding author often supervises the
-research (Fox et al., 2018; Mattsson et al., 2010). Unsurprisingly,
-measures based on first or corresponding authorships are strongly
-correlated, suggesting that these authors share the same country
-affiliation. In most cases, the first author is identical to the
+regard to contributions, author roles and positions (Larivière et al.,
+2016; Zhang et al., 2022). The first author typically conduct the main
+research underlying a paper, while the corresponding author often
+supervises the research (Fox et al., 2018; Mattsson et al., 2010).
+Unsurprisingly, measures based on first or corresponding authorships are
+strongly correlated, suggesting that these authors share the same
+country affiliation. In most cases, the first author is identical to the
 corresponding author (Chinchilla-Rodríguez et al., 2024). Despite this
 correlation, the study observed a recent slight increase in open access
-articles by corresponding authors over first authors across Springer
-Nature hybrid journals. This prompts questions about how institutional
-open access sponsorship practices influence author roles and assignments
-within co-author teams, especially as funding opportunities vary
-(Gumpenberger et al., 2018). Previous research has highlighted how
-institutionalised bibliometric practices can affect the valuation of
-authorship positions (Helgesson, 2020), suggesting the need to monitor
-the potential influences of the availability of open access funding on
-authorship roles (Maddi \& Silva, 2024).
+articles attributed to transformative agreements by corresponding
+authors over first authors across Springer Nature hybrid journals. This
+prompts questions about how institutional open access sponsorship
+practices influence author roles and assignments within co-author teams,
+especially as funding opportunities vary (Gumpenberger et al., 2018).
+Previous research has highlighted how institutionalised bibliometric
+practices can affect the valuation of authorship positions (Helgesson,
+2020), suggesting the need to monitor the potential influences of the
+availability of open access funding on authorship roles (Maddi \& Silva,
+2024).
 
@@
 well-suited for analysing the transition from subscription-based journal
-publishing to full open access.
+publishing to full open access. But using them in conjunction with
+proprietary databases provides a more robust understanding that
+overcomes individual shortcomings in coverage and metadata quality.
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-Achterberg_2023}
+Achterberg, I., \& Jahn, N. (2023). Introducing the hybrid open access
+dashboard (HOAD). \emph{cOAlition S Blog}.
+\url{https://www.coalition-s.org/blog/introducing-the-hybrid-open-access-dashboard-hoad/}
+
 \bibitem[\citeproctext]{ref-Akbaritabar_2024}
@@
 bibliometric analyses}. arXiv. \url{https://arxiv.org/abs/2404.17663}
+
+\bibitem[\citeproctext]{ref-Asai_2023}
+Asai, S. (2023). Does double dipping occur? The case of wiley's hybrid
+journals. \emph{Scientometrics}, \emph{128}(9), 5159--5168.
+\url{https://doi.org/10.1007/s11192-023-04800-8}
 
@@
 
-\bibitem[\citeproctext]{ref-eck_crossref_2022}
-Eck, N. J. van, \& Waltman, L. (2022). \emph{Crossref as a source of
-open bibliographic metadata}. MetaArXiv.
-\url{https://doi.org/10.31222/osf.io/smxe5}
-
 \bibitem[\citeproctext]{ref-eck_2024_6bx73}
@@
 \url{https://doi.org/10.1162/qss_a_00022}
+
+\bibitem[\citeproctext]{ref-Hinchliffe_2019}
+Hinchliffe, L. J. (2019). \emph{Transformative agreements: A primer}.
+The Scholarly Kitchen.
+\url{https://web.archive.org/web/20210128170342/https://scholarlykitchen.sspnet.org/2019/04/23/transformative-agreements/}
 
@@
 
+\bibitem[\citeproctext]{ref-van_Eck_2022}
+van Eck, N. J., \& Waltman, L. (2024). \emph{Crossref as a source of
+open bibliographic metadata}.
+https://doi.org/\url{https://doi.org/10.31222/osf.io/smxe5}
+
 \bibitem[\citeproctext]{ref-Visser_2021}
@@
 
+\bibitem[\citeproctext]{ref-Zhang_2022}
+Zhang, L., Wei, Y., Huang, Y., \& Sivertsen, G. (2022). Should open
+access lead to closed research? The trends towards paying to perform
+research. \emph{Scientometrics}, \emph{127}(12), 7653--7679.
+\url{https://doi.org/10.1007/s11192-022-04407-5}
+
 \end{CSLReferences}

### 2025-04-21 analysis/manuscript.qmd
Arxiv submission

@@
 Following their initial proposal in 2015 [@Schimmer_2015], the number of transformative agreements has grown substantially, resulting in increased open access publishing in hybrid journals [@Jahn_2025; @Rothfritz_2024; @McCabe_2024].
-Prior to these agreements, only a few articles in hybrid journals were openly accessible [@Bj_rk_2012; @Laakso_2016; @Piwowar_2018], except for some European countries that had dedicated funding policies and the SCOAP$^3$ consortium for high-energy physics journals [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020; @Kohls_2018].
+Prior to these agreements, only a few articles in hybrid journals were openly accessible [@Bj_rk_2012; @Laakso_2016; @Piwowar_2018], except for some European countries that had dedicated funding policies and the SCOAP$^3$ consortium for high-energy physics journals [@Pinfield_2015; @Robinson_Garcia_2020; @Huang_2020_elife; @Kohls_2018].
 Coordination efforts through initiatives such as OA2020 and Efficiency and Standards for Article Charges (ESAC) have further streamlined workflows between library consortia and publishers [@ifla; @Geschuhn_2017].
@@
 To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists including the Directory of Open Access Journals (DOAJ).
-The resulting hybrid journal data were made available via hoaddata and were used to independently determine the publication volume for each database.
+The resulting hybrid journal data comprising more tahn 13,000 hybrid journals were made available via hoaddata and were used to independently determine the publication volume for each database.
 
@@
 
-The author confirms the sole responsibility for the following: Conceptualization, Methodology, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Writing - Review & Editing, Visualization.
+The author confirms the sole responsibility for the following: Conceptualisation, Methodology, Formal analysis, Investigation, Data Curation, Writing - Original Draft, Writing - Review & Editing, Visualization.
 
@@
 
-The author declares no competing interests.
+The author has participated in working groups of OA2020, ESAC and the Barcelona Declaration as part of his professional responsibilities.
+No funding was received from these initiatives for the present study, and they had no influence on its design, data collection, analysis, or interpretation of results.
 
@@
 This work was supported by the Federal Ministry of Education and Research of Germany (BMBF) under grants 16WIK2301E / 16WIK2101F.
-
 The proprietary bibliometric data from Scopus and Web of Science were provided by the German Competence Network for Bibliometrics under BMBF grant 16WIK2101A.

### 2025-04-21 analysis/manuscript.tex
Arxiv submission

@@
 The results were consistent across all three data sources, showing strong correlations in country-level metrics despite differences in journal coverage and metadata availability.
-By 2023, transformative agreements enabled the majority of hybrid open access articles, with particularly high adoption in European countries.
+By 2023, transformative agreements enabled the majority of open access in hybrid journals, with particularly high adoption in European countries.
 The analysis revealed strong alignment between first and corresponding authorship when measuring agreement uptake by publisher and country.
@@
 funding policies and the SCOAP\(^3\) consortium for high-energy physics
-journals (Huang et al., 2020; Kohls \& Mele, 2018; Pinfield, 2015;
-Robinson-Garcia et al., 2020). Coordination efforts through initiatives
-such as OA2020 and Efficiency and Standards for Article Charges (ESAC)
-have further streamlined workflows between library consortia and
-publishers (Campbell et al., 2022; Geschuhn \& Stone, 2017). By April
-2025, the ESAC Transformative Agreement Registry, the primary
-community-driven evidence source, listed approximately 1,300
+journals (C.-K. (Karl). Huang et al., 2020; Kohls \& Mele, 2018;
+Pinfield, 2015; Robinson-Garcia et al., 2020). Coordination efforts
+through initiatives such as OA2020 and Efficiency and Standards for
+Article Charges (ESAC) have further streamlined workflows between
+library consortia and publishers (Campbell et al., 2022; Geschuhn \&
+Stone, 2017). By April 2025, the ESAC Transformative Agreement Registry,
+the primary community-driven evidence source, listed approximately 1,300
 agreements.\footnote{\url{https://esac-initiative.org/about/transformative-agreements/agreement-registry/}}
@@
 institutions to transformative agreements included in the ESAC registry
-(Jahn, 2025; Jonge et al., 2025). Yet, no systematic comparison exists
-of how open versus proprietary data sources perform when analysing the
-same hybrid journals covered by transformative agreements to identify
-comparative strengths and shortcomings of data sources.
+(de Jonge et al., 2025; Jahn, 2025). Yet, no systematic comparison
+exists of how open versus proprietary data sources perform when
+analysing the same hybrid journals covered by transformative agreements
+to identify comparative strengths and shortcomings of data sources.
 
@@
 using multiple journal lists including the Directory of Open Access
-Journals (DOAJ). The resulting hybrid journal data were made available
-via hoaddata and were used to independently determine the publication
-volume for each database.
+Journals (DOAJ). The resulting hybrid journal data comprising more tahn
+13,000 hybrid journals were made available via hoaddata and were used to
+independently determine the publication volume for each database.
 
@@
 agreement was actually in place being considered for subsequent
-analysis. A related study (Jonge et al., 2025), applied to publications
-funded by the Dutch Research Council (NWO) and validated against
-internal invoicing data, confirmed that such matching can accurately
-identify most articles under transformative agreements.
+analysis. A related study (de Jonge et al., 2025), applied to
+publications funded by the Dutch Research Council (NWO) and validated
+against internal invoicing data, confirmed that such matching can
+accurately identify most articles under transformative agreements.
 
@@
 aligns with previous studies (Akbaritabar et al., 2024; Alperin et al.,
-2024) and rankings (Eck et al., 2024). Overall, the results show strong
-correlations by country affiliation, which supports the use open
+2024) and rankings (van Eck et al., 2024). Overall, the results show
+strong correlations by country affiliation, which supports the use open
 metadata for large-scale analyses of hybrid open access. However, the
@@
 be cross-verified using multiple sources and snapshots that can be used
-to track changes in the data over time (Huang et al., 2020).
+to track changes in the data over time (C.-K. Huang et al., 2020).
 Incorporating expert-curated journal lists can also help prevent
@@
 information demonstrates the reliability of such an open approach for
-the assessment of articles under transformative agreements (Jonge et
+the assessment of articles under transformative agreements (de Jonge et
 al., 2025). With enhanced open metadata compliance attributable to
@@
 The author confirms the sole responsibility for the following:
-Conceptualization, Methodology, Formal analysis, Investigation, Data
+Conceptualisation, Methodology, Formal analysis, Investigation, Data
 Curation, Writing - Original Draft, Writing - Review \& Editing,
@@
 
-The author declares no competing interests.
+The author has participated in working groups of OA2020, ESAC and the
+Barcelona Declaration as part of his professional responsibilities. No
+funding was received from these initiatives for the present study, and
+they had no influence on its design, data collection, analysis, or
+interpretation of results.
 
@@
 This work was supported by the Federal Ministry of Education and
-Research of Germany (BMBF) under grants 16WIK2301E / 16WIK2101F.
-
-The proprietary bibliometric data from Scopus and Web of Science were
+Research of Germany (BMBF) under grants 16WIK2301E / 16WIK2101F. The
+proprietary bibliometric data from Scopus and Web of Science were
 provided by the German Competence Network for Bibliometrics under BMBF
@@
 \bibitem[\citeproctext]{ref-Achterberg_2023}
-Achterberg, I., \& Jahn, N. (2023). Introducing the hybrid open access
-dashboard (HOAD). \emph{cOAlition S Blog}.
+Achterberg, I., \& Jahn, N. (2023). \emph{Introducing the hybrid open
+access dashboard (HOAD)}. {cOAlition S}.
 \url{https://www.coalition-s.org/blog/introducing-the-hybrid-open-access-dashboard-hoad/}
@@
 
+\bibitem[\citeproctext]{ref-de_Jonge_2025}
+de Jonge, H., Kramer, B., \& Sondervan, J. (2025). \emph{Tracking
+transformative agreements through open metadata: Method and validation
+using {Dutch Research Council NWO} funded papers}. MetaArXiv.
+\url{https://doi.org/10.31222/osf.io/tz6be_v1}
+
 \bibitem[\citeproctext]{ref-D_r_2025}
@@
 \emph{113}(1), 219--236. \url{https://doi.org/10.1007/s11192-017-2483-y}
-
-\bibitem[\citeproctext]{ref-eck_2024_6bx73}
-Eck, N. J. van, Waltman, L., \& Neijssel, M. (2024). \emph{{Launch of
-the CWTS Leiden Ranking Open Edition 2024}}. Leiden Madtrics.
-\url{https://doi.org/10.59350/r512t-r8h93}
 
@@
 
+\bibitem[\citeproctext]{ref-Huang_2020_elife}
+Huang, C.-K. (Karl)., Neylon, C., Hosking, R., Montgomery, L., Wilson,
+K. S., Ozaygen, A., \& Brookes-Kenworthy, C. (2020). Evaluating the
+impact of open access policies on research institutions. \emph{eLife},
+\emph{9}. \url{https://doi.org/10.7554/elife.57067}
+
 \bibitem[\citeproctext]{ref-Huang_2020}
@@
 \url{https://doi.org/10.1002/asi.24549}
-
-\bibitem[\citeproctext]{ref-de_Jonge_2025}
-Jonge, H. de, Kramer, B., \& Sondervan, J. (2025). \emph{Tracking
-transformative agreements through open metadata: Method and validation
-using {Dutch Research Council NWO} funded papers}. MetaArXiv.
-\url{https://doi.org/10.31222/osf.io/tz6be_v1}
 
@@
 \bibitem[\citeproctext]{ref-Robinson_Garcia_2020}
-Robinson-Garcia, N., Costas, R., \& Leeuwen, T. N. van. (2020). Open
+Robinson-Garcia, N., Costas, R., \& van Leeuwen, T. N. (2020). Open
 access uptake by universities worldwide. \emph{{PeerJ}}, \emph{8},
@@
 
+\bibitem[\citeproctext]{ref-eck_2024_6bx73}
+van Eck, N. J., Waltman, L., \& Neijssel, M. (2024). \emph{{Launch of
+the CWTS Leiden Ranking Open Edition 2024}}. Leiden Madtrics.
+\url{https://doi.org/10.59350/r512t-r8h93}
+
 \bibitem[\citeproctext]{ref-Visser_2021}
-Visser, M., Eck, N. J. van, \& Waltman, L. (2021). Large-scale
+Visser, M., van Eck, N. J., \& Waltman, L. (2021). Large-scale
 comparison of bibliographic data sources: {Scopus, Web of Science,

### 2025-05-07 analysis/manuscript.qmd
Added two important paper and submitted to journals

@@
     extra_dependencies: ["flafter"]
+  docx: default
 csl: apa.csl
@@
 Importantly, these agreements themselves can help improve this evidence base, as researchers and libraries advocate for including open metadata in negotiations to help avoid data analytics from becoming a commodity controlled by publishers [@Aspesi_2020; @McCabe_2024].
-For instance, ESAC recommends using Crossref to share metadata supporting workflows including open access licenses and funding information to support discovery and monitoring [@Geschuhn_2017]. 
+For instance, the ESAC Workflow Recommendations for Transformative and Open Access Agreements promote using Crossref to share metadata that supports workflows including open access licenses and funding information to support discovery and monitoring [@Geschuhn_2017].
 Simultaneously, the bibliometric community pushes for more open metadata through this publisher-driven DOI registration platform and continuously monitors progress [@van_Eck_2022].
@@
 
-While promising for general bibliometric research, applying open metadata to transformative agreements presents distinct challenges due to their targeted journal coverage and complex funding attribution.
-With invoicing data largely unavailable, previous studies have relied on publication fee pricing lists or expenditures, and first and/or corresponding author affiliations as proxies to attribute funding to institutions [@Butler_2023; @Zhang_2022; @Haustein_APC_2024].
-Recent work has extended this approach and attributed open access articles to transformative agreements by using data from the cOAlition S Journal Checker Tool, which links journals and participating institutions to transformative agreements included in the ESAC registry [@Jahn_2025; @de_Jonge_2025].
-Yet, no systematic comparison exists of how open versus proprietary data sources perform when analysing the same hybrid journals covered by transformative agreements to identify comparative strengths and shortcomings of data sources.
+While promising for general bibliometric research, measuring hybrid open accesss and transformative agreements presents particular challenges. 
+With invoicing data largely unavailable [@Kramer_2024], previous studies have relied on publication fee pricing lists or expenditures, and first and/or corresponding author affiliations as proxies to attribute funding to institutions [@Butler_2023; @Zhang_2022; @Haustein_APC_2024].
+Others manually obtained journals and institutions from transformative agreement documents before retrieving eligible publications from the Web of Science [@Bakker_2024].
+Recent work has extended this approach and programmatically attributed open access articles to transformative agreements by using data from the cOAlition S Journal Checker Tool, which links journals and participating institutions to transformative agreements included in the ESAC registry, and the open metadata sources Crossref and OpenAlex [@Jahn_2025; @de_Jonge_2025].
+Yet, no systematic comparison exists of how open versus proprietary bibliometric data sources perform when analysing the same hybrid journals covered by transformative agreements to identify comparative strengths and shortcomings of data sources.
 
@@
 
-hoaddata, developed and maintained by the author to support open access monitoring and research [@Achterberg_2023; @Jahn_2025], is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources [@hoaddata].
+hoaddata [@hoaddata], developed and maintained by the author to support open access monitoring and research [@Achterberg_2023; @Jahn_2025], is an R data package that regularly collects information on hybrid open access uptake from multiple openly available data sources.
 It combines article-level metadata from Crossref [@Hendricks_2020] and affilation metadata from OpenAlex [@priem2022openalex] with transformative agreement information from the Transformative Agreement Data dump used by the cOAlition S Journal Checker Tool (JCT)[^tadump], which links journal and institutional data about participating research organisations to agreements in the ESAC Transformative Agreement Registry[^esac_registry].
@@
 To identify hybrid journals, a comprehensive exclusion of fully open access journals was performed using multiple journal lists including the Directory of Open Access Journals (DOAJ).
-The resulting hybrid journal data comprising more tahn 13,000 hybrid journals were made available via hoaddata and were used to independently determine the publication volume for each database.
+The resulting hybrid journal data comprising more than 13,000 hybrid journals were made available via hoaddata and were used to independently determine the publication volume for each database.
 
@@
 Figure \ref{fig-upset_coverage_results} presents the coverage of hybrid journals included in transformative agreements, based on the intersections of the examined data sources. 
-The intersection sets of journals (\ref{fig-upset_coverage_results}A) and articles (\ref{fig-upset_coverage_results}B= are visualised as an UpSet graph [@Lex_2014; @ComplexUpset].
+The intersection sets of journals (\ref{fig-upset_coverage_results}A) and articles (\ref{fig-upset_coverage_results}B) are visualised as an UpSet graph [@Lex_2014; @ComplexUpset].
 The x-axis represents the set intersections in a matrix layout.
@@
 
-Figure \ref{fig-upset_coverage_results_publisher} presents the coverage of hybrid journals in transformative agreements across data sources from 2019 to 2023 focusing on publisher portfolios.
+Figure \ref{fig-upset_coverage_results_publisher} presents the coverage of hybrid journals in transformative agreements across data sources from 2019 to 2023 with a focus on publisher portfolios.
 The analysis highlights the dominance of the three largest publishers, Elsevier, Springer Nature, and Wiley, which collectively accounted for `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_jn)*100`% of the hybrid journals and `r 100 - jn_upset_publisher_all |> filter(publisher == "Other") |> pull(prop_art)*100`% of the articles published during this five-year period.
@@
     scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
-    guides(fill = guide_legend(reverse = TRUE)) +
+  #  guides(fill = guide_legend(reverse = TRUE)) +
     ylab("Journals") 
@@
 Similarly, hybrid open access by transformative agreements substantially increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and C).
-Trends were consistent when measuring the first (dashed line) and corresponding authorship (dotted line). 
+Trends were consistent when measuring first (dashed line) and corresponding authorship (dotted line). 
 According to Scopus, `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_first_author_articles) |> sum()` open access articles could be attributed to transformative agreements based on first author metadata (increasing from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_first_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_first_author_articles) |> sum()`) and `r uptake_by_year_short |> filter(src == "Scopus") |> pull(ta_oa_corresponding_author_articles) |> sum()` using corresponding author metadata (from `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2019) |> pull(ta_oa_corresponding_author_articles) |> sum()` to `r uptake_by_year_short |> filter(src == "Scopus", earliest_year == 2023) |> pull(ta_oa_corresponding_author_articles) |> sum()`).
@@
 
-When examining the percentage of open access articles in hybrid journals (Figure \ref{fig-uptake_country}B), a different pattern emerged.
+When examining the percentage of open access articles in hybrid journals (see Figure \ref{fig-uptake_country}B), a different pattern emerged.
 Authors affiliated with institutions from medium-sized European countries, such as Sweden, the Netherlands, Finland, and Hungary, made a large proportion of their articles open access.
@@
 
-The coverage analysis revealed that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage, which reflects the dominance of major publishers whose established journal portfolios are comprehensively indexed in proprietary databases [@vanbellen2024oligopolyacademicpublisherspersists].
+The coverage analysis revealed that hybrid journals are well indexed in all three data sources, particularly in terms of article coverage, which reflects the dominance of major publishers whose established journal portfolios are comprehensively indexed in proprietary databases [@vanbellen2024oligopolyacademicpublisherspersists; @Bakker_2024].
 Differences emerge for journals targeting practitioners or local non-English language communities, with many such titles indexed exclusively in Crossref and Scopus.
@@
 Further inconsistencies emerged between Scopus and Web of Science, despite both relying on Unpaywall: Web of Science erroneously labels Elsevier’s delayed open access as hybrid, whereas Scopus correctly categorises it.
-Notably, Scopus and Crossref showed greater alignment than that observed in a related comparison between Scopus and OpenAlex [@alperin2024analysissuitabilityopenalexbibliometric]. 
-This likely reflects the use of a curated list to identify hybrid open access journals, rather than depending solely on article-level tags.
+Notably, Scopus and hoaddata showed greater alignment than that observed in a related comparison between Scopus and OpenAlex [@alperin2024analysissuitabilityopenalexbibliometric]. 
+This likely reflects the advantage of using of a curated list to identify hybrid open access journals, rather than depending solely on article-level tags.
 Given these variations, which may compromise comparability, research and monitoring exercises should avoid relying on a single source. 
@@
 
-It is important to note that the study's estimates of articles from institutions involved in transformative agreements are approximations because of a lack of access to invoice data, which is not usually shared by library consortia and publishers.
+It is important to note that the study's estimates of articles from institutions involved in transformative agreements are approximations because of a lack of access to invoice data, which is not usually shared by library consortia and publishers [@Kramer_2024].
 Furthermore, the open data sources for transformative agreements, including the cOAlition S Journal Checker Tool and the ESAC registry, are based on voluntary effort, crowd-sourced from various consortia.

### 2025-05-07 analysis/manuscript.tex
Added two important paper and submitted to journals

@@
 from becoming a commodity controlled by publishers (Aspesi \& Brand,
-2020; McCabe \& Mueller-Langer, 2024). For instance, ESAC recommends
-using Crossref to share metadata supporting workflows including open
+2020; McCabe \& Mueller-Langer, 2024). For instance, the ESAC Workflow
+Recommendations for Transformative and Open Access Agreements promote
+using Crossref to share metadata that supports workflows including open
 access licenses and funding information to support discovery and
@@
 
-While promising for general bibliometric research, applying open
-metadata to transformative agreements presents distinct challenges due
-to their targeted journal coverage and complex funding attribution. With
-invoicing data largely unavailable, previous studies have relied on
-publication fee pricing lists or expenditures, and first and/or
-corresponding author affiliations as proxies to attribute funding to
-institutions (Butler et al., 2023; Haustein et al., 2024; Zhang et al.,
-2022). Recent work has extended this approach and attributed open access
+While promising for general bibliometric research, measuring hybrid open
+accesss and transformative agreements presents particular challenges.
+With invoicing data largely unavailable (Kramer, 2024), previous studies
+have relied on publication fee pricing lists or expenditures, and first
+and/or corresponding author affiliations as proxies to attribute funding
+to institutions (Butler et al., 2023; Haustein et al., 2024; Zhang et
+al., 2022). Others manually obtained journals and institutions from
+transformative agreement documents before retrieving eligible
+publications from the Web of Science (Bakker et al., 2024). Recent work
+has extended this approach and programmatically attributed open access
 articles to transformative agreements by using data from the cOAlition S
 Journal Checker Tool, which links journals and participating
-institutions to transformative agreements included in the ESAC registry
-(de Jonge et al., 2025; Jahn, 2025). Yet, no systematic comparison
-exists of how open versus proprietary data sources perform when
-analysing the same hybrid journals covered by transformative agreements
-to identify comparative strengths and shortcomings of data sources.
+institutions to transformative agreements included in the ESAC registry,
+and the open metadata sources Crossref and OpenAlex (de Jonge et al.,
+2025; Jahn, 2025). Yet, no systematic comparison exists of how open
+versus proprietary bibliometric data sources perform when analysing the
+same hybrid journals covered by transformative agreements to identify
+comparative strengths and shortcomings of data sources.
 
@@
 
-hoaddata, developed and maintained by the author to support open access
-monitoring and research (Achterberg \& Jahn, 2023; Jahn, 2025), is an R
-data package that regularly collects information on hybrid open access
-uptake from multiple openly available data sources (Jahn, 2024). It
-combines article-level metadata from Crossref (Hendricks et al., 2020)
-and affilation metadata from OpenAlex (Priem et al., 2022) with
+hoaddata (Jahn, 2024), developed and maintained by the author to support
+open access monitoring and research (Achterberg \& Jahn, 2023; Jahn,
+2025), is an R data package that regularly collects information on
+hybrid open access uptake from multiple openly available data sources.
+It combines article-level metadata from Crossref (Hendricks et al.,
+2020) and affilation metadata from OpenAlex (Priem et al., 2022) with
 transformative agreement information from the Transformative Agreement
@@
 using multiple journal lists including the Directory of Open Access
-Journals (DOAJ). The resulting hybrid journal data comprising more tahn
+Journals (DOAJ). The resulting hybrid journal data comprising more than
 13,000 hybrid journals were made available via hoaddata and were used to
@@
 journals (\ref{fig-upset_coverage_results}A) and articles
-(\ref{fig-upset_coverage_results}B= are visualised as an UpSet graph
+(\ref{fig-upset_coverage_results}B) are visualised as an UpSet graph
 (Krassowski, 2020; Lex et al., 2014). The x-axis represents the set
@@
 of hybrid journals in transformative agreements across data sources from
-2019 to 2023 focusing on publisher portfolios. The analysis highlights
-the dominance of the three largest publishers, Elsevier, Springer
-Nature, and Wiley, which collectively accounted for 47\% of the hybrid
-journals and 62\% of the articles published during this five-year
+2019 to 2023 with a focus on publisher portfolios. The analysis
+highlights the dominance of the three largest publishers, Elsevier,
+Springer Nature, and Wiley, which collectively accounted for 47\% of the
+hybrid journals and 62\% of the articles published during this five-year
 period. In terms of article volume, Elsevier led with 2,441,358 articles
@@
 increased between 2019 and 2023 (Figures \ref{fig-uptake_overview}B and
-C). Trends were consistent when measuring the first (dashed line) and
+C). Trends were consistent when measuring first (dashed line) and
 corresponding authorship (dotted line). According to Scopus, 479,297
@@
 When examining the percentage of open access articles in hybrid journals
-(Figure \ref{fig-uptake_country}B), a different pattern emerged. Authors
-affiliated with institutions from medium-sized European countries, such
-as Sweden, the Netherlands, Finland, and Hungary, made a large
-proportion of their articles open access. Germany and the United Kingdom
-also had approximately 40\% of their output available as open access. In
-contrast, non-OECD countries showed notably lower adoption of hybrid
-open access, with South Africa being the only BRICS member (as of 2022)
-well-represented in the data. The United States also demonstrated a
-relatively low proportion of open access articles. These findings were
+(see Figure \ref{fig-uptake_country}B), a different pattern emerged.
+Authors affiliated with institutions from medium-sized European
+countries, such as Sweden, the Netherlands, Finland, and Hungary, made a
+large proportion of their articles open access. Germany and the United
+Kingdom also had approximately 40\% of their output available as open
+access. In contrast, non-OECD countries showed notably lower adoption of
+hybrid open access, with South Africa being the only BRICS member (as of
+2022) well-represented in the data. The United States also demonstrated
+a relatively low proportion of open access articles. These findings were
 consistent across all databases. However, France was better represented
@@
 reflects the dominance of major publishers whose established journal
-portfolios are comprehensively indexed in proprietary databases (Bellen
-et al., 2024). Differences emerge for journals targeting practitioners
-or local non-English language communities, with many such titles indexed
-exclusively in Crossref and Scopus. Using Crossref as a bibliometric
-database in hoaddata demonstrated a particular strength in identifying
-newly established hybrid journals, a notable finding given that
-transformative agreements primarily target existing subscription-based
-journals. The landscape of hybrid journal publishing thus differs
-markedly from that of fully open access journals. Comparing the coverage
-of OpenAlex, Scopus and Web of Science, Simard et al. (2024) indicate
-that only half of the fully open access journals listed in the DOAJ are
-also indexed in Scopus and Web of Science. Notably, journals that charge
-no publication fees (``diamond journals'') are absent from the selective
-Web of Science, which reinforces existing disparities in the indexing of
-under-represented research communities and regions in selective
-bibliometric databases (Simard et al., 2024).
+portfolios are comprehensively indexed in proprietary databases (Bakker
+et al., 2024; Bellen et al., 2024). Differences emerge for journals
+targeting practitioners or local non-English language communities, with
+many such titles indexed exclusively in Crossref and Scopus. Using
+Crossref as a bibliometric database in hoaddata demonstrated a
+particular strength in identifying newly established hybrid journals, a
+notable finding given that transformative agreements primarily target
+existing subscription-based journals. The landscape of hybrid journal
+publishing thus differs markedly from that of fully open access
+journals. Comparing the coverage of OpenAlex, Scopus and Web of Science,
+Simard et al. (2024) indicate that only half of the fully open access
+journals listed in the DOAJ are also indexed in Scopus and Web of
+Science. Notably, journals that charge no publication fees (``diamond
+journals'') are absent from the selective Web of Science, which
+reinforces existing disparities in the indexing of under-represented
+research communities and regions in selective bibliometric databases
+(Simard et al., 2024).
 
@@
 Elsevier's delayed open access as hybrid, whereas Scopus correctly
-categorises it. Notably, Scopus and Crossref showed greater alignment
+categorises it. Notably, Scopus and hoaddata showed greater alignment
 than that observed in a related comparison between Scopus and OpenAlex
-(Alperin et al., 2024). This likely reflects the use of a curated list
-to identify hybrid open access journals, rather than depending solely on
-article-level tags. Given these variations, which may compromise
-comparability, research and monitoring exercises should avoid relying on
-a single source. Instead, the selection of open access evidence should
-be cross-verified using multiple sources and snapshots that can be used
-to track changes in the data over time (C.-K. Huang et al., 2020).
-Incorporating expert-curated journal lists can also help prevent
-misclassification based on business models (Visser et al., 2021).
+(Alperin et al., 2024). This likely reflects the advantage of using of a
+curated list to identify hybrid open access journals, rather than
+depending solely on article-level tags. Given these variations, which
+may compromise comparability, research and monitoring exercises should
+avoid relying on a single source. Instead, the selection of open access
+evidence should be cross-verified using multiple sources and snapshots
+that can be used to track changes in the data over time (C.-K. Huang et
+al., 2020). Incorporating expert-curated journal lists can also help
+prevent misclassification based on business models (Visser et al.,
+2021).
 
@@
 because of a lack of access to invoice data, which is not usually shared
-by library consortia and publishers. Furthermore, the open data sources
-for transformative agreements, including the cOAlition S Journal Checker
-Tool and the ESAC registry, are based on voluntary effort, crowd-sourced
-from various consortia. However, recent validation using Dutch research
-information demonstrates the reliability of such an open approach for
-the assessment of articles under transformative agreements (de Jonge et
-al., 2025). With enhanced open metadata compliance attributable to
-evolving standards and initiatives in support of negotiations with
-publishers, particularly through the ESAC initiative and Barcelona
-Declaration on Open Research Information, the situation is likely to
-improve. In conclusion, this study has shown that open metadata are
-well-suited for analysing the transition from subscription-based journal
-publishing to full open access. But using them in conjunction with
-proprietary databases provides a more robust understanding that
-overcomes individual shortcomings in coverage and metadata quality.
+by library consortia and publishers (Kramer, 2024). Furthermore, the
+open data sources for transformative agreements, including the cOAlition
+S Journal Checker Tool and the ESAC registry, are based on voluntary
+effort, crowd-sourced from various consortia. However, recent validation
+using Dutch research information demonstrates the reliability of such an
+open approach for the assessment of articles under transformative
+agreements (de Jonge et al., 2025). With enhanced open metadata
+compliance attributable to evolving standards and initiatives in support
+of negotiations with publishers, particularly through the ESAC
+initiative and Barcelona Declaration on Open Research Information, the
+situation is likely to improve. In conclusion, this study has shown that
+open metadata are well-suited for analysing the transition from
+subscription-based journal publishing to full open access. But using
+them in conjunction with proprietary databases provides a more robust
+understanding that overcomes individual shortcomings in coverage and
+metadata quality.
 
@@
 \url{https://www.budapestopenaccessinitiative.org/boai20/}
+
+\bibitem[\citeproctext]{ref-Bakker_2024}
+Bakker, C., Langham-Putrow, A., \& Riegelman, A. (2024). Impact of
+transformative agreements on publication patterns: An analysis based on
+agreements from the ESAC registry. \emph{International Journal of
+Librarianship}, \emph{8}(4), 67--96.
+\url{https://doi.org/10.23974/ijol.2024.vol8.4.341}
 
@@
 
+\bibitem[\citeproctext]{ref-Kramer_2024}
+Kramer, B. (2024). \emph{Study on scientific publishing in {Europe} --
+{Development}, diversity, and transparency of costs}. Publications
+Office of the European Union. \url{https://doi.org/doi/10.2777/89349}
+
 \bibitem[\citeproctext]{ref-ComplexUpset}

### 2025-05-07 data/README.Rmd
Add hoaddata indicators
Improve data documentation

@@
 
-This folder contains analytical datasets providing aggregated counts on the uptake of open access in hybrid journals included in transformative agreements by journal and country information between 2019 and 2023.
+This folder contains analytical datasets providing aggregated counts from proprietary bibliometric databases on the uptake of open access in hybrid journals included in transformative agreements by journal and country information between 2019 and 2023.
 A particular focus was on whether publications were from eligible authors who could use such agreements.
@@
 - Web of Science (`wos`): In-house Web of Science database provided by the Kompetenznetzwerk Bibliometrie (April 2024 snapshot)
+
+The folder also include respective datasets derived from the openly available hoaddata (`hoad`), v.0.3.
+It combines data from several open metadata sources, including cOAlition S Public Transformative Agreement Data powering the Journal Checker Tool, Crossref and OpenAlex.
+<https://github.com/subugoe/hoaddata/releases/tag/v.0.3>.
+Note that the OA indicators were calculated for first authors only.
 
